### 2019-12-29 background.tex
start intro

\section{Background and Related Work}
\label{sxn:background}

Here we will cite and discuss related work, including
\cite{MM17_TR},
\cite{MM18_TR},
\cite{MM19_HTSR_ICML},
\cite{weightwatcher_package}
\cite{MM19_KDD},
\cite{MM20_SDM},
\cite{MM20_unpub_work},

### 2019-12-29 conclusion.tex
start intro

\section{Conclusion}
\label{sxn:conc}
XXX.  CONCLUSION.

### 2019-12-29 introduction.tex
start intro

\section{Introduction}
\label{sxn:intro}

A common problem in machine learning (ML) 
%and artificial intelligence (AI) 
is to evaluate the quality of a given model.
A popular way to accomplish this
%, in particular in academic environments, 
is to train a model and then evaluate its training and/or testing error.
There are many problems with this approach.
Well-known problems with just examining training/testing curves include that 
they give very limited insight into the overall properties of the model, 
they do not take into account the (often extremely large human and CPU/GPU) time for hyperparameter fiddling,
they typically do not correlate with other properties of interest such as robustness or fairness or interpretability, 
%XXX SOMETHING ELSE, 
and so on.
A somewhat less well-known problem, but one that is increasingly important (in particular in industrial-scale ML---where the \emph{users} of models are not the \emph{developers} of the models) is that one may access to neither the training data nor the testing data.
Instead, one may simply be given a model that has already been trained---we will call such an already-trained model a \emph{pre-trained model}---and be told to use it.

Na\"{\i}vely---but in our experience commonly, among both ML practitioners and ML theorists---if one does not have access to training or testing data, then one can say absolutely nothing about the quality of a ML model.
This may be true in worst-case theory, but ML models are used in practice, and there is a need for theory to guide that practice.
Moreover, if ML is to become an industrial process, then the process will become siloed: some groups will gather data, other groups will develop models, and still other groups will use those models.
The users of models can not be expected to know the precise details of how the models were built, the specifics of the data that were used to train the model, what the loss of values of the hyperparameters were, how precisely the model was regularized, etc.
%
Having metrics to evaluate the quality of a ML model in the absence of training and testing data and without any detailed knowledge of the training and testing process---indeed, having theory for pre-trained models, to predict how, when, and why such models can be expected to perform well or poorly---is clearly of interest.

In this paper, we present and apply techniques to evaluate the generalization properties of large-scale state-of-the-art pre-trained neural network (NN) models.%
\footnote{We reiterate: One could use these techniques to improve training, and we have been asked about that, but we are not interested in that here. Our main goal here is to use these techniques to evaluate properties of state-of-the-art pre-trained NN models.}
To do so, we consider a large suite of publicly-available models from computer vision (CV) and natural language processing (NLP).
%
By now, there are many such state-of-the-art models that are publicly-available, e.g., 
there are now hundreds of pre-trained models in CV ($\ge 500$) and NLP ($\approx 100$).%
\footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
These provide a large corpus of models that by some community standard are state-of-the-art.%
\footnote{Clearly, there is a selection bias or survivorship bias here---people tend not to make publicly-available their poorly-performing models---but these models are things in the world that (like social networks or the internet) can be analyzed for their properties.}
XXX.  MORE DETAILS.
Importantly, for all of these models, we have no access to training data or testing data.

XXX.  LIST PLACES WHERE THEY ARE AVAIALBLE, HERE OR IN NEXT SECTION.

In more detail, our main contributions are the following.
\begin{itemize}
\item XXX TECHNICAL THING 1
\item XXX TECHNICAL THING 2
\item XXX TECHNICAL THING 3
\end{itemize}
Perhaps our most improtant contribution, however, is just doing things different, etc. ...

### 2019-12-31 all_models.tex
added outline

\section{Comparison of Metrics on PreTrained CV Models}
\label{sxn:all_models}

### 2019-12-31 cv_models.tex
added outline

\section{Comparison of 3 Computer Vision Models}
\label{sxn:cv}

### 2019-12-31 nlp_models.tex
added outline

\section{Comparison of NLP Transformer Models}
\label{sxn:cv}

### 2020-01-01 methods.tex
added methods section

\section{Methods}
\label{sxn:methods}

We assume we are given several pretrained Deep Neural Networks (DNNs), as part of a similar architecture.
We would like to estimate the trends in the reported test / generalization accuracy accross a series of similar archtectures.  
For example, below we compare the 8 pretrained models in the VGG series: (VGG11, VGG11\_BN$\cdots$ VGG19), with
and without Batch Normalization, trained on ImageNet, and widely available in the pyTorch distribution.

To do this, we will compute a variety of \emph{Complexity Metrics} based on the Product Norm of the layer weight matrics.
Note that unlike traditional ML approaches, however, we do not seek a bound on the complexity (i.e. test error), 
nor are we trying to evaluating a single model with differing hyperparmeters.  We wish to examine different models a 
common architecture series. And, also, compare different architectures themselves.

Let us write the Energy Landscape (or optimization function) for a typical DNN with $L$ layers, with activation functions $h_{l}(\cdot)$, and with $N\times M$)  weight matrices $\mathbf{W}_{l}$, and the biases $\mathbf{b}_{l}$, as follows:
\begin{equation}
E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})  .
\label{eqn:dnn_energy}
\end{equation}

We assume the model has been (or will be) trained on (unspecified) labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, 
using Backprop, by minimizing the loss $\mathcal{L}$ 
For simplicity, we do not indicate the structural details of the layers (e.g., Dense or not, Convolutions or not, Residual/Skip Connections, etc.).

Each layer contains by one or more layer 2D weight matrices $\mathbf{W}_{L}$, and/or the 2D feature maps $\mathbf{W}_{i,L}$ 
extracted from 2D Convolutional layers.

### 2020-01-01 methods.tex
added methods section

@@
 
-We assume the model has been (or will be) trained on (unspecified) labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, 
-using Backprop, by minimizing the loss $\mathcal{L}$ 
-For simplicity, we do not indicate the structural details of the layers (e.g., Dense or not, Convolutions or not, Residual/Skip Connections, etc.).
+The model has been (or will be) trained on (unspecified) labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, 
+using Backprop, by minimizing the loss $\mathcal{L}()$.
+(comment on regularization, dropout, etc).
+Each layer contains by one or more layer 2D weight matrices $\mathbf{W}_{L}$, and/or the 2D feature maps $\mathbf{W}_{i,L}$ extracted from 2D Convolutional layers.
+(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below.)
 
-Each layer contains by one or more layer 2D weight matrices $\mathbf{W}_{L}$, and/or the 2D feature maps $\mathbf{W}_{i,L}$ 
-extracted from 2D Convolutional layers.
+We assume the layer weight matrices are statistically independent, allowing us to estimate the Complexity, or test accuracy, with a standard Product Norm
+
+\begin{equation}
+\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert\cdots\Vert\mathbf{W}_{L}\Vert ,
+\end{equation}
+where $\Vert\mathbf{W}\Vert$ may be the Frobenius norm, the Spectral Norm, or even p-Norm (Shatten Norm).
+
+DEFINE EACH HERE
+
+One can view $\mathcal{C}$ as something akin to a data dependent VC complexity.
+
+CLARIFY HOW CONV2D lsyers are treated
+
+\begin{eqnarray*}
+\log\mathcal{C} &\sim& \log\bigg[\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert\cdots\Vert\mathbf{W}_{L}\Vert\bigg]  \\
+                &\sim& \bigg[\log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert\cdots\log\Vert\mathbf{W}_{L}\Vert\bigg]  ,
+\end{eqnarray*}
+and we define the average log norm of the weight matrices as

### 2020-01-01 methods.tex
added methods section

@@
 The model has been (or will be) trained on (unspecified) labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, 
-using Backprop, by minimizing the loss $\mathcal{L}()$.
-(comment on regularization, dropout, etc).
+using Backprop, by minimizing some (also unspecified) loss function $\mathcal{L}()$.
+Moreover, we expect that most well trained,.
+production quality models will employ 1 or more forms of on regularization, such as Batch Normalization, Dropout, etc, and will also contain additional structure such as Skip Connections etc.
+Here, we ignore these details, and focus only on the weight matrices.
 Each layer contains by one or more layer 2D weight matrices $\mathbf{W}_{L}$, and/or the 2D feature maps $\mathbf{W}_{i,L}$ extracted from 2D Convolutional layers.
@@
 
-We assume the layer weight matrices are statistically independent, allowing us to estimate the Complexity, or test accuracy, with a standard Product Norm
+We assume the layer weight matrices are statistically independent, allowing us to estimate the Complexity, or test accuracy, with a standard Product Norm, which resembles a data dependent VC complexity
 
@@
 \end{equation}
-where $\Vert\mathbf{W}\Vert$ may be the Frobenius norm, the Spectral Norm, or even p-Norm (Shatten Norm).
+where $mathbf{W}$ is an $(N\times M)$ weight matrix, with $N\ge M$, and 
+ $\Vert\mathbf{W}\Vert$ is some matrix Norm.
 
-DEFINE EACH HERE
+\begin{itemize}
+ \item Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{W}\Vert^{2}_{2}=\sum_{i,j}W^{2}_{i,j}$,
+ \item Spectral Norm:  $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$, or even the
+ \item $\alpha$-Norm (or Shatten Norm) $\Vert\mathbf{W}\Vert^{alpha}_{\alpha}=\sum_{i=1}^{M}\lambda^{\alpha}$,
+\end{itemize}
 
-One can view $\mathcal{C}$ as something akin to a data dependent VC complexity.
+where $\lamba_{i}$ is the i-th eigenvalue of the correlation matrix $\mathbf{X}$
 
+\begin{equation}
+\mathbf{X}=\dfrac{1}{N}\mathbf{W}^{T}\mathbf{W},
+\end{equation}
+
+$lambda_{max}$ is the maximum eigenvalue, and $\alpha$ is the power law exponent that
+
+MAKE THIS MORE CLEAR
 CLARIFY HOW CONV2D lsyers are treated

### 2020-01-01 methods.tex
added methods section

@@
 
-Let us write the Energy Landscape (or optimization function) for a typical DNN with $L$ layers, with activation functions $h_{l}(\cdot)$, and with $N\times M$)  weight matrices $\mathbf{W}_{l}$, and the biases $\mathbf{b}_{l}$, as follows:
+Let us write the Energy Landscape (or optimization function) for a typical DNN with $L$ layers as
 \begin{equation}
@@
 \end{equation}
+
+with activation functions $h_{l}(\cdot)$,  weight matrices $\mathbf{W}_{l}$, and the biases $\mathbf{b}_{l}$.
 
@@
 Here, we ignore these details, and focus only on the weight matrices.
+
 Each layer contains by one or more layer 2D weight matrices $\mathbf{W}_{L}$, and/or the 2D feature maps $\mathbf{W}_{i,L}$ extracted from 2D Convolutional layers.
-(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below.)
-
-We assume the layer weight matrices are statistically independent, allowing us to estimate the Complexity, or test accuracy, with a standard Product Norm, which resembles a data dependent VC complexity
+(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below.) We assume the layer weight matrices are statistically independent, allowing us to estimate the Complexity, or test accuracy, with a standard Product Norm, which resembles a data dependent VC complexity
 
@@
 \end{equation}
-where $mathbf{W}$ is an $(N\times M)$ weight matrix, with $N\ge M$, and 
- $\Vert\mathbf{W}\Vert$ is some matrix Norm.
+where $\mathbf{W}$ is an $(N\times M)$ weight matrix, with $N\ge M$, and 
+ $\Vert\mathbf{W}\Vert$ is some matrix Norm, such as the:
 
 \begin{itemize}
- \item Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{W}\Vert^{2}_{2}=\sum_{i,j}W^{2}_{i,j}$,
+ \item Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{W}\Vert^{2}_{2}=\sum_{i,j}w^{2}_{i,j}$,
  \item Spectral Norm:  $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$, or even the
- \item $\alpha$-Norm (or Shatten Norm) $\Vert\mathbf{W}\Vert^{alpha}_{\alpha}=\sum_{i=1}^{M}\lambda^{\alpha}$,
+ \item $\alpha$-Norm (or Shatten Norm) $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda^{\alpha}$,
 \end{itemize}
 
-where $\lamba_{i}$ is the i-th eigenvalue of the correlation matrix $\mathbf{X}$
+where $\lambda_{i}$ is the $i-th$ eigenvalue of the correlation matrix $\mathbf{X}=\dfrac{1}{N}\mathbf{W}^{T}\mathbf{W}$, and $\lambda_{max}$ is the maximum eigenvalue.
+
+The exponent $\alpha$ is the power law exponent that arise in our \emph{Theory of Heavy Tailed Self Regularization}, and is the determined by fitting the Empirical Spectral Density (ESD) of $\mathbf{X}$--i.e.
+a histogram of the eigenvalues--$\rho(\lambda)$ to a truncated power law
 
 \begin{equation}
-\mathbf{X}=\dfrac{1}{N}\mathbf{W}^{T}\mathbf{W},
+\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}
 \end{equation}
 
-$lambda_{max}$ is the maximum eigenvalue, and $\alpha$ is the power law exponent that
+---
 
-MAKE THIS MORE CLEAR
 CLARIFY HOW CONV2D lsyers are treated
+
+----
 

### 2020-01-01 methods.tex
added methods section

@@
 Each layer contains by one or more layer 2D weight matrices $\mathbf{W}_{L}$, and/or the 2D feature maps $\mathbf{W}_{i,L}$ extracted from 2D Convolutional layers.
-(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below.) We assume the layer weight matrices are statistically independent, allowing us to estimate the Complexity, or test accuracy, with a standard Product Norm, which resembles a data dependent VC complexity
+(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below.) We assume the layer weight matrices are statistically independent, allowing us to estimate the Complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm, which resembles a data dependent VC complexity
 
@@
 where $\mathbf{W}$ is an $(N\times M)$ weight matrix, with $N\ge M$, and 
- $\Vert\mathbf{W}\Vert$ is some matrix Norm, such as the:
+ $\Vert\mathbf{W}\Vert$ is some matrix Norm.  Here, we will consider:
 
@@
  \item Spectral Norm:  $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$, or even the
- \item $\alpha$-Norm (or Shatten Norm) $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda^{\alpha}$,
+ \item $\alpha-$Norm (or Shatten Norm) $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda^{\alpha}$,
 \end{itemize}
@@
 
+We will also consider an approximate capacity metric, $\hat{\alpha}$, shown previously to correlate well the reported test accuracy of pretrained DNNs
+
+\begin{itemize}
+ \item $\hat{\alpha}=\alpha\log\lambda_{max}\approx\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$
+\end{itemize}
+
+which approxmiates the log $\alpha-$Norm for both Very Heavy Tailed weight matrices as well as finite size, Moderately Heavy Tailed $\mathbf{W}$.
 ---

### 2020-01-01 methods.tex
added methods section

@@
 where $\mathbf{W}$ is an $(N\times M)$ weight matrix, with $N\ge M$, and 
- $\Vert\mathbf{W}\Vert$ is some matrix Norm.  Here, we will consider:
+ $\Vert\mathbf{W}\Vert$ is a matrix norm.   We will actually compute the log Complexity, which takes the form 
+of an Average Log Norm:
+\begin{eqnarray*}
+\log\mathcal{C} &\sim& \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert\cdots\log\Vert\mathbf{W}_{L}\Vert
+\end{eqnarray*}
+
+Here, we will consider the following Norms:
 
 \begin{itemize}
- \item Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{W}\Vert^{2}_{2}=\sum_{i,j}w^{2}_{i,j}$,
- \item Spectral Norm:  $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$, or even the
+ \item Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{W}\Vert^{2}_{2}=\sum_{i,j}w^{2}_{i,j}$
+ \item Spectral Norm:  $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
  \item $\alpha-$Norm (or Shatten Norm) $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda^{\alpha}$,
@@
 where $\lambda_{i}$ is the $i-th$ eigenvalue of the correlation matrix $\mathbf{X}=\dfrac{1}{N}\mathbf{W}^{T}\mathbf{W}$, and $\lambda_{max}$ is the maximum eigenvalue.
+(Note that eigenvalues are the square of the singular values $\sigma_{i}$ of $\mathbf{W}$ :  $\lambda_{i}=\sigma^{2}_{i}$.)
 
@@
 
-We will also consider an approximate capacity metric, $\hat{\alpha}$, shown previously to correlate well the reported test accuracy of pretrained DNNs
+We will also consider an approximate capacity metric, $\hat{\alpha}$, shown previously to correlate well with the trends in reported test accuracies of pretrained DNNs \cite{MM}
 
@@
 
-which approxmiates the log $\alpha-$Norm for both Very Heavy Tailed weight matrices as well as finite size, Moderately Heavy Tailed $\mathbf{W}$.
----
+which approximates the log $\alpha-$Norm for both Very Heavy Tailed weight matrices ($alpha < 2$  and reasoably well for finite size, Moderately Heavy Tailed ones $\alpha\in[2,5]$.
 
-CLARIFY HOW CONV2D lsyers are treated
+SOME MORE DISCUSSION HERE
 
+\paragraph{Spectral Analysis of Convolutional 2D Layers}
+While we can easily analyze Linear layers, there is some ambiguity in performing spectral analysis on Convolutional 2D (Conv2D) layers.
+A Conv2D layer is a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$.
+Typically, the $w=h=k$, giving $(k\times k)$ Feature Maps of dimension $(in\_ch\times out\_ch)$ each. 
+Usually $in\_ch\ll out\_ch$.
+There are at least 3 different approaches to computing the Singular Values Decomposition(s) (SVD) of an Conv2D layer
+\begin{enumerate}
+\item run SVD on each of the feature maps, yielding $(k\times k)$ sets of $M$ singular values. 
+\item stack the feature maps into a single rectangular matrix of, say, dimension $((k\times k\times out\_ch)\times in\_ch)$, yielding $in\_ch$ singular values
+\item Compute the 2D Fourier Transform (FFT) for each of the $(in\_ch, out\_ch)$ pairs, and run SVD on the resulting Fourier coeffients\cite{Long2019}.
+This leads to $(k\times in\_ch\times out\_ch)$ non-zero singular values.\footnote{assuming say no padding and a stride of 1}
+\end{enumerate}
+
+Each method has tradeoffs. For this study, because we are computing tens of thousands of calculations,  we select $1.$, which is numerically the fastest. 
 ----
-
-\begin{eqnarray*}
-\log\mathcal{C} &\sim& \log\bigg[\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert\cdots\Vert\mathbf{W}_{L}\Vert\bigg]  \\
-                &\sim& \bigg[\log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert\cdots\log\Vert\mathbf{W}_{L}\Vert\bigg]  ,
-\end{eqnarray*}
-and we define the average log norm of the weight matrices as

### 2020-01-01 methods.tex
added methods section

@@
 
-The exponent $\alpha$ is the power law exponent that arise in our \emph{Theory of Heavy Tailed Self Regularization}, and is the determined by fitting the Empirical Spectral Density (ESD) of $\mathbf{X}$--i.e.
+The exponent $\alpha$ is the power law exponent that arises in our \emph{Theory of Heavy Tailed Self Regularization}, and is the determined by fitting the Empirical Spectral Density (ESD) of $\mathbf{X}$--i.e.
 a histogram of the eigenvalues--$\rho(\lambda)$ to a truncated power law

### 2020-01-01 methods.tex
added methods section

@@
-\section{Methods}
+section{Methods}
 \label{sxn:methods}
@@
 While we can easily analyze Linear layers, there is some ambiguity in performing spectral analysis on Convolutional 2D (Conv2D) layers.
-A Conv2D layer is a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$.
-Typically, the $w=h=k$, giving $(k\times k)$ Feature Maps of dimension $(in\_ch\times out\_ch)$ each. 
-Usually $in\_ch\ll out\_ch$.
+A Conv2D layer is a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ kernel, and $in\_ch,out\_ch$ input and output channels, respectively.
+Typically, the $w=h=k$, giving $(k\times k)$ Feature Maps $\mathbf{W}_{i,L}$
+of dimension $(in\_ch\times out\_ch)$ each.  Usually $in\_ch\ll out\_ch$.
 There are at least 3 different approaches to computing the Singular Values Decomposition(s) (SVD) of an Conv2D layer
 \begin{enumerate}
-\item run SVD on each of the feature maps, yielding $(k\times k)$ sets of $M$ singular values. 
+\item run SVD on each of the feature maps $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values. 
 \item stack the feature maps into a single rectangular matrix of, say, dimension $((k\times k\times out\_ch)\times in\_ch)$, yielding $in\_ch$ singular values
@@
 
-Each method has tradeoffs. For this study, because we are computing tens of thousands of calculations,  we select $1.$, which is numerically the fastest. 
+Each method has tradeoffs.  While, in principle, 3. is mathematically sound, the actual data available
+in a pretrained model is usually the ourput of the Conv2D transformation, followed by Pooling and/or Batch Nprmalization,
+and this complicates the interpretation.  For this study, because we are computing tens of thousands of calculations,  
+we select $1.$, which is numerically the fastest and easiest to reproduce.\footnote{We will provide a Google Colab notebook where all results can be reproduced, with the option to redo the calculations with option 3 for the SVD of the Conv2D.}
 ----

### 2020-01-01 methods.tex
added methods section

@@
 
+\charles{not actually feature Maps, need to find the correct term: Tensor Slice}
+
 \paragraph{Spectral Analysis of Convolutional 2D Layers}
@@
 A Conv2D layer is a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ kernel, and $in\_ch,out\_ch$ input and output channels, respectively.
-Typically, the $w=h=k$, giving $(k\times k)$ Feature Maps $\mathbf{W}_{i,L}$
+Typically, the $w=h=k$, giving $(k\times k)$ Tensor Slices $\mathbf{W}_{i,L}$
 of dimension $(in\_ch\times out\_ch)$ each.  Usually $in\_ch\ll out\_ch$.
@@
 \begin{enumerate}
-\item run SVD on each of the feature maps $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values. 
+\item run SVD on each of the Tensor Slices $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values. 
 \item stack the feature maps into a single rectangular matrix of, say, dimension $((k\times k\times out\_ch)\times in\_ch)$, yielding $in\_ch$ singular values
 \item Compute the 2D Fourier Transform (FFT) for each of the $(in\_ch, out\_ch)$ pairs, and run SVD on the resulting Fourier coeffients\cite{Long2019}.
-This leads to $(k\times in\_ch\times out\_ch)$ non-zero singular values.\footnote{assuming say no padding and a stride of 1}
+This leads to $\sim(k\times in\_ch\times out\_ch)$ non-zero singular values.
 \end{enumerate}

### 2020-01-02 methods.tex
minor update

@@
 While we can easily analyze Linear layers, there is some ambiguity in performing spectral analysis on Convolutional 2D (Conv2D) layers.
-A Conv2D layer is a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ kernel, and $in\_ch,out\_ch$ input and output channels, respectively.
+A Conv2D layer is a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ filter, and $in\_ch,out\_ch$ input and output channels, respectively.
 Typically, the $w=h=k$,  giving $(k\times k)$ Tensor Slices $\mathbf{W}_{i,L}$

### 2020-01-04 methods.tex
more explanation in methods section

@@
 
-Each method has tradeoffs.  While, in principle, 3. is mathematically sound, the actual data available
-in a pretrained model is usually the ourput of the Conv2D transformation, followed by Pooling and/or Batch Nprmalization,
-and this complicates the interpretation.  For this study, because we are computing tens of thousands of calculations,  
-we select $1.$, which is numerically the fastest and easiest to reproduce.\footnote{We will provide a Google Colab notebook where all results can be reproduced, with the option to redo the calculations with option 3 for the SVD of the Conv2D.}
+Each method has tradeoffs.  While, in principle, 3. is mathematically sound, it is computationally expensive. 
+For this study, because we are computing tens of thousands of calculations, we select $1.$, which is 
+numerically the fastest and easiest to reproduce.\footnote{We will provide a Google Colab notebook where all results can be reproduced, with the option to redo the calculations with option 3 for the SVD of the Conv2D.}
+
+To verify that our approach is meaningful, we need to see the ESD is neither due to a random matrix, nor due to unsually large matrix elements.
+We examine typical Conv2D layer for the pretained AlexNet model (distributed with pyTorch).
+Figure \ref{fig:alexnet1} displays the ESD for the first slice (or matrix $\mathbf{W}$) of the third Conv2D layer, extracted from a 4-index Tensor of shape $(384, 192, 3, 3)$.
+The red line displays the best fit to a random matrix, using the Marchenko pastur theory\cite{MM}.
+We can see the random matrix model does not describe the ESD very well.
+For comparison, Figure \ref{fig:alexnet2} shows the ESD of the same matrix, randomly shuffled; here looks similar to the red line plot of the orginal ESD.
+In fact, the empircal ESD is better modeled with a truncated power law distribtion.
+
+Although the ESD is \emph{Heavy Tailed}, this does not imply that the orginal matrix $\mathbf{W}$ is itself heavy tailed--
+only the correlation matrix $\mathbf{X}$ is. (i.e. maximum eigenvalue) than the original. If $\mathbf{W}$ was, 
+then it would contain 1 or more unusually large matrix elements, and they would dominate the ESD.  
+Of course the randomized $\mathbf{W}$ would also be heavy tailed, but its ESD neither resembles the original
+nor is it heavy tailed.  So we can rule out $\mathbf{W}$ being heavy tailed.
+
+\begin{figure}[H]
+   \centering
+   \subfigure[ESD of AlexNet Conv2D Layer 3 Slice 1]{
+     \includegraphics[scale=0.5]{img/alexnet1.png} 
+     \label{fig:alexnet1}
+   }
+   \subfigure[ESD of the same matrix, randomly shuffled  ]{
+      \includegraphics[scale=0.5]{img/alexnet2.png}
+      \label{fig:alexnet2}
+   }
+   \caption{%
+    Caption
+   }
+   \label{fig:alexnet}
+\end{figure}
+
+These plots tell us that the Tensor Slice of the Conv2D contains significant correlations learned from the data.  
+By modeling the ESD with a power law distribution $\lambda^{\alpha}$, we can characterize the amount of correlation learned;
+the smaller the exponent $\alpha$, the more correlation in the weight matric.
+
 ----

### 2020-01-04 cv_models.tex
place holders for CV VGG

@@
 \label{sxn:cv}
+
+\begin{figure}[t]
+    \centering
+    \subfigure[ Frobenius Norm ]{
+        \includegraphics[width=5cm]{img/vgg-fnorm.png}
+        \label{fig:vgg-fnorm}
+    }
+    \qquad
+    \subfigure[ Spectral Norm ]{
+        \includegraphics[width=4.9cm]{img/vgg-snorm.png}
+        \label{fig:vgg-snorm}
+    }
+    \qquad
+    \subfigure[ Weighted Alpha ]{
+        \includegraphics[width=4.9cm]{img/vgg-walpha.png}
+        \label{fig:vgg-walpha}
+    }
+    \qquad
+    \subfigure[ Alpha-Norm ]{
+        \includegraphics[width=4.9cm]{img/vgg-pnorm.png}
+        \label{fig:vgg-pnorm}
+    }
+    \caption{Comparison of norm metrics vs reported test accuracy for pretrained VGG models, trained on ImageNet, available in pyTorch.  Plots will be updated and replace }
+
+    \label{fig:vgg-metrics}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.30]{img/vgg-alpha-layers.png}
+  \label{fig:vgg-alpha-layers}
+    \caption{Power law exponent $\alpha$ vs layer for VGG11 and VGG19. }
+\end{figure}

### 2020-01-04 methods.tex
place holders for CV VGG

@@
-section{Methods}
+\section{Methods}
 \label{sxn:methods}
@@
  \item Spectral Norm:  $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
- \item $\alpha-$Norm (or Shatten Norm) $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda^{\alpha}$,
+ \item $\alpha-$Norm (or Shatten Norm) $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda^{\alpha}$,
 \end{itemize}
@@
 \begin{itemize}
- \item $\hat{\alpha}=\alpha\log\lambda_{max}\approx\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$
+ \item $\hat{\alpha}=\alpha\log\lambda_{max}\approx\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$
 \end{itemize}
@@
 
-To verify that our approach is meaningful, we need to see the ESD is neither due to a random matrix, nor due to unsually large matrix elements.
+To verify that our approach is meaningful, we need to see the ESD is neither due to a random matrix, nor due to unsually large matrix elements, but, in fact, captures correlations learned from the data.
 We examine typical Conv2D layer for the pretained AlexNet model (distributed with pyTorch).
@@
 Although the ESD is \emph{Heavy Tailed}, this does not imply that the orginal matrix $\mathbf{W}$ is itself heavy tailed--
-only the correlation matrix $\mathbf{X}$ is. (i.e. maximum eigenvalue) than the original. If $\mathbf{W}$ was, 
+only the correlation matrix $\mathbf{X}$ is. If $\mathbf{W}$ was, 
 then it would contain 1 or more unusually large matrix elements, and they would dominate the ESD.  
@@
 By modeling the ESD with a power law distribution $\lambda^{\alpha}$, we can characterize the amount of correlation learned;
-the smaller the exponent $\alpha$, the more correlation in the weight matric.
+the smaller the exponent $\alpha$, the more correlation in the weight matrix.
 

### 2020-01-04 cv_models.tex
latest PDF
still working on CV
placeholders for CV images

@@
 \label{sxn:cv}
+
+\paragraph{Empirical Metrics vs Test Accuracies for VGG}
 
@@
 
-\begin{figure}
+ADD PLOTS FOR
+
+pNorm and ResNet
+pNorm and ResNet (ImageNet1K all models)
+pNorm and DenseNet
+
+\paragraph{Correlations and Information Flow in CV Models}
+
+\begin{figure}[t]
     \centering
-  \includegraphics[scale=0.30]{img/vgg-alpha-layers.png}
+
+    \subfigure[ VGG ]{
+        \includegraphics[width=4.1cm]{img/vgg-alpha-layers.png}
         \label{fig:vgg-alpha-layers}
-    \caption{Power law exponent $\alpha$ vs layer for VGG11 and VGG19. }
+    }
+    \qquad
+    \subfigure[ ResNet ]{
+        \includegraphics[width=3.8cm]{img/resnet-alpha-layers.png}
+        \label{fig:resnet-alpha-layer}
+    }
+    \qquad
+    \subfigure[ DenseNet ]{
+        \includegraphics[width=4.1cm]{img/densenet-alpha-layers.png}
+        \label{fig:densenet-alpha-layer}
+    }
+    \caption{Power law exponent $\alpha$ vs layer for VGG, ResNet, and DenseNet models}
+    \label{fig:vgg-alpha-layers}
 \end{figure}

### 2020-01-04 cv_models.tex
moved pdf

@@
 
-\paragraph{Empirical Metrics vs Test Accuracies for VGG}
+\paragraph{Empirical Metrics vs Test Accuracies for CV models}
 
@@
 
+VGG works remarkably well
+
+ResNet is also correlated, but the RMSE is much larger, and 2/5 of the models are almost outliers
+But note that the PyTorch distribution only has a few models.
+The ResNet results look better when considering all ResNet models, trained on ImageNet1K.
+
 ADD PLOTS FOR
@@
 \paragraph{Correlations and Information Flow in CV Models}
+
+Compare VGG, ResNet, and DenseNet in the context of how many connections that have
 

### 2020-01-04 methods.tex
latest PDF
pre-Activation Maps

@@
 We would like to estimate the trends in the reported test / generalization accuracy accross a series of similar archtectures.  
-For example, below we compare the 8 pretrained models in the VGG series: (VGG11, VGG11\_BN$\cdots$ VGG19), with
 and without Batch Normalization, trained on ImageNet, and widely available in the pyTorch distribution.
 
+\paragraph{Empirical Complexity Metrics}
 To do this, we will compute a variety of \emph{Complexity Metrics} based on the Product Norm of the layer weight matrics.
@@
 
-\charles{not actually feature Maps, need to find the correct term: Tensor Slice}
-
 \paragraph{Spectral Analysis of Convolutional 2D Layers}
 While we can easily analyze Linear layers, there is some ambiguity in performing spectral analysis on Convolutional 2D (Conv2D) layers.
-A Conv2D layer is a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ filter, and $in\_ch,out\_ch$ input and output channels, respectively.
-Typically, the $w=h=k$,  giving $(k\times k)$ Tensor Slices $\mathbf{W}_{i,L}$
-of dimension $(in\_ch\times out\_ch)$ each.  Usually $in\_ch\ll out\_ch$.
+A Conv2D layer is a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ filter (or kernel), and $in\_ch,out\_ch$ input and output channels, respectively.
+Typically, the $w=h=k$,  giving $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\_ch\times out\_ch)$ each.
+Usually $in\_ch\le out\_ch$.
+
 There are at least 3 different approaches to computing the Singular Values Decomposition(s) (SVD) of an Conv2D layer
 \begin{enumerate}
-\item run SVD on each of the Tensor Slices $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values. 
+\item run SVD on each of the pre-Activation Maps $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values. 
 \item stack the feature maps into a single rectangular matrix of, say, dimension $((k\times k\times out\_ch)\times in\_ch)$, yielding $in\_ch$ singular values

### 2020-01-04 methods.tex
pre-Activation Maps

@@
 
-which approximates the log $\alpha-$Norm for both Very Heavy Tailed weight matrices ($alpha < 2$  and reasoably well for finite size, Moderately Heavy Tailed ones $\alpha\in[2,5]$.
+which approximates the log $\alpha-$Norm for both Very Heavy Tailed weight matrices ($\alpha < 2)$  and reasoably well for finite size, Moderately Heavy Tailed ones $\alpha\in[2,5]$.
 
@@
    \centering
-   \subfigure[ESD of AlexNet Conv2D Layer 3 Slice 1]{
+   \subfigure[Actual ESD]{
      \includegraphics[scale=0.5]{img/alexnet1.png} 
@@
    }
-   \subfigure[ESD of the same matrix, randomly shuffled  ]{
+   \subfigure[ESD of randomly shuffled matrix]{
       \includegraphics[scale=0.5]{img/alexnet2.png}
@@
    }
-   \caption{%
-    Caption
-   }
+   \caption{ESD of AlexNet Conv2D pre-Activation map for Layer 3 Slice 1, actual and randomized}
    \label{fig:alexnet}
@@
 
-These plots tell us that the Tensor Slice of the Conv2D contains significant correlations learned from the data.  
+These plots tell us that the pre-activation maps of the Conv2D contains significant correlations learned from the data.  
 By modeling the ESD with a power law distribution $\lambda^{\alpha}$, we can characterize the amount of correlation learned;

### 2020-01-05 cv_models.tex
updated CV models

@@
 
-pNorm and ResNet
-pNorm and ResNet (ImageNet1K all models)
-pNorm and DenseNet
+\begin{figure}[t]
+    \centering
+
+    \subfigure[ ResNet ]{
+        \includegraphics[width=4.2cm]{img/resnet-accuracy.png}
+        \label{fig:resnet-accuracy}
+    }
+    \qquad
+    \subfigure[ ResNet-1K ]{
+        \includegraphics[width=4.5cm]{img/resnet1k-accuracy.png}
+        \label{fig:resnet1k-accuracy}
+    }
+    \qquad
+    \subfigure[ DenseNet ]{
+        \includegraphics[width=4.4cm]{img/densenet-accuracy.png}
+        \label{fig:densenet-accuracy}
+    }
+    \caption{$\alpha$-Norm vs reported Top1 Error for  ResNet, ResNet-1K, and DenseNet models}
+    \label{fig:cv2-accuracy}
+\end{figure}
 

### 2020-01-05 cv_models.tex
cv models updated

@@
 
+TODO:  add table of metrics for these 3 models, show that alpha-Norm is best again!
+
+\begin{table}[t]
+\small
+\begin{center}
+\begin{tabular}{|p{1in}|c|c|c|c|}
+\hline
+      & Frobenus Norm & Spectal Norm & Weighted Alpha & Alpha-Norm \\
+ Series   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+\hline
+ VGG & & & & \\
+ ResNet & & & & \\
+ ResNet-1K & & & & \\
+ DenseNet & & & & \\
+\hline
+\end{tabular}
+\end{center}
+\caption{RMSE for Linear Fits of Metric to Reported Top1 Test Error for all pre-trained models in the architecture series.
+All models trained on ImageNet except ResNet-1K, which was trained on ImageNet-1K.
+}
+\label{table:models}
+\end{table}
+
 \paragraph{Correlations and Information Flow in CV Models}

### 2020-01-05 cv_models.tex
cv models updated

@@
 \begin{center}
-\begin{tabular}{|p{1in}|c|c|c|c|}
+\begin{tabular}{|p{1in}|c|c|c|c|c|}
 \hline
-      & Frobenus Norm & Spectal Norm & Weighted Alpha & Alpha-Norm \\
- Series   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+   &    & Frobenus Norm & Spectal Norm & Weighted Alpha & Alpha-Norm \\
+ Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
 \hline

### 2020-01-05 cv_models.tex
cv models updated

@@
 \hline
-   &    & Frobenus Norm & Spectal Norm & Weighted Alpha & Alpha-Norm \\
- Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+   &    & Frobenus Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
+ Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
 \hline

### 2020-01-05 cv_models.tex
cv models updated

--- 2020-01-05 cv_models.tex
@@
 \hline
-   &    & Frobenus Norm & Spectal Norm & Weighted Alpha & Alpha-Norm \\
- Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
+ Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
 \hline

### 2020-01-09 nlp_models.tex
gpt2 figures added

@@
 \section{Comparison of NLP Transformer Models}
-\label{sxn:cv}
+\label{sxn:nlp}
+
+\paragraph{GPT vs GPT2}
+
+\paragraph{Power Law Exponents}
+
+\begin{figure}
+   \includegraphics[scale=0.25]{img/gpt-alpha-hist.png}
+   \caption{Comparison of heavy tailed power law exponent $\alpha$ for OpenAI GPT and GPT pretrained models}
+
+   \label{fig:gpt-alphs-hist}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+
+    \subfigure[ Power Law Exponent $\alpha$  ]{
+        \includegraphics[width=4.5cm]{img/gpt2-alpha-hist.png}
+        \label{fig:gpt-alpha-layer}
+    }
+    \qquad
+    \subfigure[ Spectral Norm $\Vert\mathbf{W}\Vert_{\infty}$]{
+        \includegraphics[width=4.5cm]{img/gpt2-snorm-hist.png}
+        \label{fig:resnet-snorm-layer}
+    }
+    \qquad
+    \subfigure[ Alpha-Norm $\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$]{
+        \includegraphics[width=4.5cm]{img/gpt2-pnorm-hist.png}
+        \label{fig:gpt-pnorm-layer}
+    }
+    \caption{Comparison of Power Law Exponnents, Spectral Norm, and Alpha-Norm for different size models in the GPT2 architecture series.    }
+    \label{fig:gpt-alpha-layers}
+\end{figure}
+
+\paragraph{Correlation Flow}
+
+\begin{figure}[t]
+    \centering
+
+    \subfigure[ Power Law Exponent $\alpha$  ]{
+        \includegraphics[width=4.0cm]{img/gpt-alpha-layer.png}
+        \label{fig:gpt-alpha-layer}
+    }
+    \qquad
+    \subfigure[ Spectral Norm $\Vert\mathbf{W}\Vert_{\infty}$]{
+        \includegraphics[width=4.5cm]{img/gpt-snorm-layer.png}
+        \label{fig:resnet-snorm-layer}
+    }
+    \qquad
+    \subfigure[ Alpha-Norm $\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$]{
+        \includegraphics[width=4.5cm]{img/gpt-pnorm-layer.png}
+        \label{fig:gpt-pnorm-layer}
+    }
+    \caption{Comparison of Correlation Flow and Spectral Norm for OpenAI GPT and GPT2   }
+    \label{fig:gpt-alpha-layers}
+\end{figure}
+
+\paragraph{GPT2: small, medium, large, extra-large}

### 2020-01-09 nlp_models.tex
more to nlp added

@@
 \paragraph{GPT vs GPT2}
+
+\begin{table}[t]
+\small
+\begin{center}
+\begin{tabular}{|p{1in}|c|c|c|c|c|}
+\hline
+   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
+ Series & \#Layers   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+\hline
+ GPT & & & & \\
+ GPT2 (small) & & & & \\
+ GPT2-medium & & & & \\
+ GPT2-large & & & & \\
+ GPT2-xls & & & & \\
+
+\hline
+\end{tabular}
+\end{center}
+\caption{Average metrics for pretrainnd OpenAI GPT and GPT2 models.}
+\label{table:models}
+\end{table}
 

### 2020-01-09 nlp_models.tex
minor reorg
discussion of GPT2 added

@@
 
+In the past two years, nearly 100 open source, pre-trained DNNs for Natural Language Processing (NLP) have emerged,
+based on the revolutionary Transformer architecture, including variants of BERT, Transformer-XML, GPT, etc.
+The Transformer architectures consist of blocks of Attention layers, containing of 2 large, Feed Forward (Linear)
+weight matrics \cite{Attn2017}. In contrast to the smaller pre-Activation maps arising in Cond2D layers,
+the Attention matrices ar
+
 \paragraph{GPT vs GPT2}
+Here, we use the \emph{WeightWatcher} tool to analyze the OpenAI GPT and GPT2 models, which gives us
+the opportunity to analyze the effect of both training the same model with different size data sets,
+These GPT models have generated significant media attention because of its remarkable ability to
+ generate fake text, and it's potential misuse. 
+The original GPT and later GPT model have the same architecture and number of layers, but
+the original GPT was released having been trained on a deficient data set.  Recently,
+ however, a much improved GPT2 model has been released, which is remarkably good at creating 
+auto-genetated text.
+
+\charles{more here ?}
+We analyze GPT models deployed with the popular HuggingFace PyTorch library.
+GPT has 12 layers, with 4 Multi-head Attention Blocksm, giving $48$ Layer Weight Matrices $\mathbf{W}$.
+Each Block has 2 components, the Self Attention (attn) and the Projection (proj) matrics.  
+The self-attention  matrices are larger, of dimension ($2304\times 768$) or ($3072\times 768$).
+The projection layer concatenates the self-attention results into a vector (of dimension $768$).
+This gives many large, typically sparse weight matrices, and they can be
+ \emph{poorly correlated} when trained with insufficient data--as we shall see below.
+
+Because GPT and GPT are trained on different data sets, the initial Embedding matrices differ in shape.
+GPT  has an initial Token and Positional Embedding layers, of dimension
+$(40478\times 768)$ and $(512\times 768)$, resp, whereas GPT2 has input Embeddings of shape
+$(50257\times 768)$ and $(1024\times 768)$, resp.  Interstingly, they also have very spectral properties,
+also shown below.
+
+\charles{Since we show this table, maybe we should discuss GPT2 scaling models}
 
@@
 
-\paragraph{Power Law Exponents}
+\paragraph{Heavy Tailed Power Law Exponents in GPT and GPT2}
 
@@
 
-\begin{figure}[t]
-    \centering
-
-    \subfigure[ Power Law Exponent $\alpha$  ]{
-        \includegraphics[width=4.5cm]{img/gpt2-alpha-hist.png}
-        \label{fig:gpt-alpha-layer}
-    }
-    \qquad
-    \subfigure[ Spectral Norm $\Vert\mathbf{W}\Vert_{\infty}$]{
-        \includegraphics[width=4.5cm]{img/gpt2-snorm-hist.png}
-        \label{fig:resnet-snorm-layer}
-    }
-    \qquad
-    \subfigure[ Alpha-Norm $\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$]{
-        \includegraphics[width=4.5cm]{img/gpt2-pnorm-hist.png}
-        \label{fig:gpt-pnorm-layer}
-    }
-    \caption{Comparison of Power Law Exponnents, Spectral Norm, and Alpha-Norm for different size models in the GPT2 architecture series.    }
-    \label{fig:gpt-alpha-layers}
-\end{figure}
-
-\paragraph{Correlation Flow}
+\paragraph{Correlation Flow in GPT and GPT2}
 
@@
 \paragraph{GPT2: small, medium, large, extra-large}
+
+\begin{figure}[t]
+    \centering
+
+    \subfigure[ Power Law Exponent $\alpha$  ]{
+        \includegraphics[width=4.5cm]{img/gpt2-alpha-hist.png}
+        \label{fig:gpt-alpha-layer}
+    }
+    \qquad
+    \subfigure[ Spectral Norm $\Vert\mathbf{W}\Vert_{\infty}$]{
+        \includegraphics[width=4.5cm]{img/gpt2-snorm-hist.png}
+        \label{fig:resnet-snorm-layer}
+    }
+    \qquad
+    \subfigure[ Alpha-Norm $\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$]{
+        \includegraphics[width=4.5cm]{img/gpt2-pnorm-hist.png}
+        \label{fig:gpt-pnorm-layer}
+    }
+    \caption{Comparison of Power Law Exponnents, Spectral Norm, and Alpha-Norm for different size models in the GPT2 architecture series.    }
+    \label{fig:gpt-alpha-layers}
+\end{figure}

### 2020-01-09 nlp_models.tex
some ebasic text added

@@
 We analyze GPT models deployed with the popular HuggingFace PyTorch library.
-GPT has 12 layers, with 4 Multi-head Attention Blocksm, giving $48$ Layer Weight Matrices $\mathbf{W}$.
+GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ Layer Weight Matrices $\mathbf{W}$.
 Each Block has 2 components, the Self Attention (attn) and the Projection (proj) matrics.  
@@
 The projection layer concatenates the self-attention results into a vector (of dimension $768$).
-This gives many large, typically sparse weight matrices, and they can be
+This gives $50$ large, typically sparse weight matrices, and they can be
  \emph{poorly correlated} when trained with insufficient data--as we shall see below.
@@
 
-\charles{Since we show this table, maybe we should discuss GPT2 scaling models}
+The additional OpenAI GPT2 (English) models are: \emph{gpt-medium, got-large, and gpt-xl}, 
+having include $12, 24, 36, and 48$ layers, resp., with increasingly larger weight matrices.
+The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
+Table \ref{table:nlp} reports results for the average log norm metrics, using \emph{weightwatcher (0.2.3)},
+and with fully reproducible Jupyter notebooks.\footnote{\url{https://github.com/CalculatedContent/kdd2020}}
 
@@
  GPT2-large & & & & \\
- GPT2-xls & & & & \\
+ GPT2-xl & & & & \\
 
@@
 \end{center}
-\caption{Average metrics for pretrainnd OpenAI GPT and GPT2 models.}
-\label{table:models}
+\caption{Average Log Norm Metrics for pretrainnd OpenAI GPT and GPT2 models.}
+\label{table:nlp}
 \end{table}
 
-\paragraph{Heavy Tailed Power Law Exponents in GPT and GPT2}
+\paragraph{The Heavy Tailed Power Law Exponents in GPT and GPT2}
+
+are very differenmt, with GPT2 having both a notably smaller mean $\alpha$, and far fewer, unusually large outliers.
+Figure \ref{fig:gpt-alphs-hist} shows the empirical density (histogram) of $\alpha$
+for all layers in GPT (blue) and GPT2 (red).  \charles{discuss more}
 
@@
 
-\paragraph{Correlation Flow in GPT and GPT2}
+\paragraph{The Correlation Flow in GPT and GPT2} also differs significantly between GPT and GPT2.
+Figure \ref{fig:gpt-alpha-layer} plots $\alpha$ vs the layer id for each model.
+
+\charles{Discuss Spectral Norm, alpha-Norm}
 
@@
 
+Figure \ref{fig:gpt-alpha-layers} ...
+
+For this series of GPT2 models, the average $\alpha$ still decreases with increasing model size,
+althogh, the differences are less noticible than between the GPT and GPT2 models.
+Unlike GPT, however, the Layer Spectral Norms $\vert\mathbf{W}\Vert_{\infty}$ 
+and Alpha-Norms $\vert\mathbf{W}\Vert_{\alpha}^{\alpha}$
+ behave  as expected for GPT2 layers, with the larger models 
+consistently having  smaller norms.
+
 \begin{figure}[t]

### 2020-01-15 cv_models.tex
minor reorg
updated images, can autogen now

@@
     \subfigure[ Frobenius Norm ]{
-        \includegraphics[width=5cm]{img/vgg-fnorm.png}
+        \includegraphics[width=5cm]{img/VGG_lognorm_accs.png}
         \label{fig:vgg-fnorm}
@@
     \subfigure[ Spectral Norm ]{
-        \includegraphics[width=4.9cm]{img/vgg-snorm.png}
+        \includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
         \label{fig:vgg-snorm}
@@
     \subfigure[ Weighted Alpha ]{
-        \includegraphics[width=4.9cm]{img/vgg-walpha.png}
+        \includegraphics[width=4.9cm]{img/VGG_alpha_weighted_accs.png}
         \label{fig:vgg-walpha}
@@
     \subfigure[ Alpha-Norm ]{
-        \includegraphics[width=4.9cm]{img/vgg-pnorm.png}
+        \includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
         \label{fig:vgg-pnorm}
@@
     \subfigure[ ResNet ]{
-        \includegraphics[width=4.2cm]{img/resnet-accuracy.png}
+        \includegraphics[width=4.2cm]{img/ResNet_logpnorm_accs.png}
         \label{fig:resnet-accuracy}
@@
     \subfigure[ ResNet-1K ]{
-        \includegraphics[width=4.5cm]{img/resnet1k-accuracy.png}
+        \includegraphics[width=4.5cm]{img/ResNet-1K_logpnorm_accs.png}
         \label{fig:resnet1k-accuracy}
@@
     \subfigure[ DenseNet ]{
-        \includegraphics[width=4.4cm]{img/densenet-accuracy.png}
+        \includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
         \label{fig:densenet-accuracy}
@@
 \hline
- VGG & & & & \\
- ResNet & & & & \\
- ResNet-1K & & & & \\
- DenseNet & & & & \\
+ VGG & & & & & \\
+ ResNet & & & & & \\
+ ResNet-1K & & & & & \\
+ DenseNet & & & &  &\\
 \hline
@@
 
-\paragraph{Correlations and Information Flow in CV Models}
+\paragraph{Correlation Flow in CV Models}
+
+Explain WHAT IS \emph{Correlation Flow}
 
@@
     \subfigure[ VGG ]{
-        \includegraphics[width=4.1cm]{img/vgg-alpha-layers.png}
+        \includegraphics[width=4.5cm]{img/VGG_fnl_alpha_depth.png} 
                 \label{fig:vgg-alpha-layers}
@@
     \subfigure[ ResNet ]{
-        \includegraphics[width=3.8cm]{img/resnet-alpha-layers.png}
+        \includegraphics[width=4.5cm]{img/ResNet_fnl_alpha_depth.png} 
         \label{fig:resnet-alpha-layer}
@@
     \subfigure[ DenseNet ]{
-        \includegraphics[width=4.1cm]{img/densenet-alpha-layers.png}
+        \includegraphics[width=4.5cm]{img/DenseNet_all_depth.png} 
         \label{fig:densenet-alpha-layer}

### 2020-01-15 nlp_models.tex
updated images, can autogen now

@@
 \hline
- GPT & & & & \\
- GPT2 (small) & & & & \\
- GPT2-medium & & & & \\
- GPT2-large & & & & \\
- GPT2-xl & & & & \\
+ GPT & & & & &\\
+ GPT2 (small) & & & & & \\
+ GPT2-medium & & & & & \\
+ GPT2-large & & & & & \\
+ GPT2-xl & & & & &\\
 
@@
    \caption{Comparison of heavy tailed power law exponent $\alpha$ for OpenAI GPT and GPT pretrained models}
-
-   \label{fig:gpt-alphs-hist}
+   \label{fig:gpt-alpha-hist.png}
 \end{figure}
@@
     \subfigure[ Power Law Exponent $\alpha$  ]{
-        \includegraphics[width=4.0cm]{img/gpt-alpha-layer.png}
+        \includegraphics[width=4.5cm]{img/GPT-alpha-depth.png}
         \label{fig:gpt-alpha-layer}
@@
     \subfigure[ Spectral Norm $\Vert\mathbf{W}\Vert_{\infty}$]{
-        \includegraphics[width=4.5cm]{img/gpt-snorm-layer.png}
+        \includegraphics[width=4.5cm]{img/GPT-snorm-depth.png}
         \label{fig:resnet-snorm-layer}
@@
     \subfigure[ Alpha-Norm $\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$]{
-        \includegraphics[width=4.5cm]{img/gpt-pnorm-layer.png}
+        \includegraphics[width=4.5cm]{img/GPT-pnorm-depth.png}
         \label{fig:gpt-pnorm-layer}
@@
     \subfigure[ Power Law Exponent $\alpha$  ]{
-        \includegraphics[width=4.5cm]{img/gpt2-alpha-hist.png}
+        \includegraphics[width=4.5cm]{img/GPT2_fnl_alpha_hist.png}
         \label{fig:gpt-alpha-layer}
@@
     \subfigure[ Spectral Norm $\Vert\mathbf{W}\Vert_{\infty}$]{
-        \includegraphics[width=4.5cm]{img/gpt2-snorm-hist.png}
+        \includegraphics[width=4.5cm]{img/GPT2_fnl_spectralnorm_hist.png}
         \label{fig:resnet-snorm-layer}
@@
     \subfigure[ Alpha-Norm $\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$]{
-        \includegraphics[width=4.5cm]{img/gpt2-pnorm-hist.png}
+        \includegraphics[width=4.5cm]{img/GPT2_fnl_alpha_weighted_hist.png}
         \label{fig:gpt-pnorm-layer}

### 2020-01-16 methods.tex
discussion of GPT2 added
normalization discussion added

@@
 
-where $\lambda_{i}$ is the $i-th$ eigenvalue of the correlation matrix $\mathbf{X}=\dfrac{1}{N}\mathbf{W}^{T}\mathbf{W}$, and $\lambda_{max}$ is the maximum eigenvalue.
-(Note that eigenvalues are the square of the singular values $\sigma_{i}$ of $\mathbf{W}$ :  $\lambda_{i}=\sigma^{2}_{i}$.)
+where $\lambda_{i}$ is the $i-th$ eigenvalue of the \emph{Empirical Correlation Matrix}
+
+\begin{equation}
+\mathbf{X}=\mathbf{W}^{T}\mathbf{W}
+\end{equation}
+
+and $\lambda_{max}$ is the maximum eigenvalue. These eigenvalues are the square of the singular values $\sigma_{i}$ of $\mathbf{W}$ :  $\lambda_{i}=\sigma^{2}_{i}$.
 
@@
 which approximates the log $\alpha-$Norm for both Very Heavy Tailed weight matrices ($\alpha < 2)$  and reasoably well for finite size, Moderately Heavy Tailed ones $\alpha\in[2,5]$.
-
-SOME MORE DISCUSSION HERE
 
@@
 
+\paragraph{Normalization of Empirical Matrices}.  Formally, to apply RMT, we would usually define Correlation Matrix with $\dfrac{1}{N}$ normalization,
+and assume that either the variance of $\mathbf{X}$ is unity, $\sigma^{2}=1$, or is known.
+\footnote{And for formal proofs of the Heavy Tailed results, we need a different normalization,$\dfrac{1}{N^{1-\alpha}}$}
+But for these empirical studies, the pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$ ,
+with the variance already normalized to $\sqrt{\frac{1}{N}}$, or some variant of this such as Glorot (or Xavier) Normalization\cite{GloRot},
+or as $\sqrt{\frac{2}{N*k^2}}$ for Convolutional 2D Layers. 
+We do not normalize (or renormalize) the Empirical Correlation Matrices, and use them as-is,
+except that we rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $\frac{k}{\sqrt{2}}$ so that they
+are on the same scale as the Linear layers.
+Note that in some pretrinaed models, like the pyTorch VGG models, the Linear weight matrices appear to have been 
+initialized with variance $\sigma\sim0.01$.  We do not attempt to correct for this, and, instead, simply try to treat all
+models the same.  
+Also, The normalization does not affect the Power Law fits because the Heavy Tailed exponent $\alpha$ is scale-invariant.
+As are metrics like the Stable Rank and our MP Soft Rank. In this study, however, all of the metrics strongly depend 
+on the scale of the weight matrix.
+
+\paragraph{The WeightWatcher Tool}
+
 ----

### 2020-01-16 nlp_models.tex
NLP tables added...debugging needed

@@
  GPT & & & & &\\
- GPT2 (small) & & & & & \\
- GPT2-medium & & & & & \\
- GPT2-large & & & & & \\
- GPT2-xl & & & & &\\
+GT2 small & 50 & 2.05 & 2.59& 9.78 & 10.03 \\
+GPT2 medium & 98 & 2.08 & 2.58& 9.74 & 10.01 \\
+GPT2 large & 146 & 1.85 & 1.99& 7.67 & 7.94 \\
+GPT2 xl & 194 & 1.86 &1.9 2& 7.17 & 7.51 \\
+
+SOMETHING IS WRONG HERE...SOME OUTLIERS OR WE NEED TO FIX THE NORMALIZATION
 

### 2020-01-16 methods.tex
updated methods, etc

@@
 
-\paragraph{Normalization of Empirical Matrices}.  Formally, to apply RMT, we would usually define Correlation Matrix with $\dfrac{1}{N}$ normalization,
+\paragraph{Normalization of Empirical Matrices}.  
+The normalization does not affect the Power Law fits because the Heavy Tailed exponent $\alpha$ is scale-invariant.
+As are metrics like the Stable Rank and our MP Soft Rank. In this study, however, all of the metrics strongly depend 
+on the scale of the weight matrix.
+Formally, to apply RMT, we would usually define Correlation Matrix with $\dfrac{1}{N}$ normalization,
 and assume that either the variance of $\mathbf{X}$ is unity, $\sigma^{2}=1$, or is known.
@@
 are on the same scale as the Linear layers.
-Note that in some pretrinaed models, like the pyTorch VGG models, the Linear weight matrices appear to have been 
+
+Note that in some pretrained models, like the pyTorch VGG models, the Linear weight matrices appear to have been 
 initialized with variance $\sigma\sim0.01$.  We do not attempt to correct for this, and, instead, simply try to treat all
 models the same.
-Also, The normalization does not affect the Power Law fits because the Heavy Tailed exponent $\alpha$ is scale-invariant.
-As are metrics like the Stable Rank and our MP Soft Rank. In this study, however, all of the metrics strongly depend 
-on the scale of the weight matrix.
+
+However, in the GPT2 and other NLP models, we find that the first embedding layer(s) have unusually large eigenvalues,
+reflecting a lack of normalization.  For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$, 
+but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
+ $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we layer all layers as-is (although one may to normalize
+the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier); including them does
+not significantly alter the results here.
+
+COMMENT ON HOW LOG NORM first and last layers behave, maybe somewhere else
+
+COMMENT ON HOW LOG PORM  for GPT includes unusually high alpha, not meaningful other than to show the trend.
 

### 2020-01-16 nlp_models.tex
updated discussion on first layer in GPT2
updated methods, etc

--- 2020-01-15 nlp_models.tex
@@
 \label{sxn:nlp}
+
+[THESE RESULTS HAVE TO BE REDONE BY REMOVING THE FIRST LAYER...]
 
@@
  GPT & & & & &\\
- GPT2 (small) & & & & & \\
- GPT2-medium & & & & & \\
- GPT2-large & & & & & \\
- GPT2-xl & & & & &\\
+GPT2 & 50 & 2.05 &2.59& 9.78 & 10.03 \\
+GPT2 medium & 98 & 2.08 &2.58& 9.74 & 10.01 \\
+GPT2 large & 146 & 1.85 &1.99& 7.67 & 7.94 \\
+GPT2 xl & 194 & 1.86 &1.92& 7.17 & 7.51 \\
+
+Explain why Log Norm is sooo large...alpha is off may need to prune the alpha
 

### 2020-01-17 all_models.tex
more to nlp added
added serenas text

@@
-\section{Comparison of Metrics on PreTrained CV Models}
-\label{sxn:all_models}
+\section{Comparison of all pretrained CV Models}
+\label{sxn:all_cv_models}
+
+Here, we use the Weightwatcher tool to analyze 466 pretrained computer vision models Pytorch.
+These image classification and segmentation models are pretrained on nine datasets, ImageNet-1K, and CIFAR-10, CIFAR-100, Street View House Numbers (SVHN), Caltech-UCSD Birds-200-2011 (CUB-200-2011), Pascal VOC2012, ADE20K, Cityscapes, and Common Objects in Context (COCO).
+The pretrained models and their accuracy metrics are summarized in the osmr github
+
+\serena{insert link in the footnote}
+
+and a full summary of all the models analyzed is included in the Appendix. For our analysis, we then group models by architecture and datasets for further analysis.
+
+In this paper, we propose that the Weightwatcher tool could be used to predict the trends in the generalization accuracy of deep neutral network without a test set.
+To test our proposition, we choose simple linear regression to analyze the relationship between the Weightwatcher metrics and the traditional accuracy metric obtained with a test set (we avoid polynomial regressions as they are more prone to overfitting and does not make economic sense).
+On the left-hand-side of regression, we have the Top1 errors, Top5 errors as reported for ImageNet-1K models, Error % as reported in CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy (Pix.Acc.,%) and Intersection-Over-Union (mIOU,%) as dependent variables.
+We then regress them individually on each of the Weightwatcher metrics, including  alpha, alpha_weighted, spectralnormlog, softranklog, lognorm, and logpnorm as defined in earlier section.
+
+\serena{Question: only the alpha-related metrics is the core metrics of the WW tool, right? Metrics such as stable rank or spectral norm have been suggested in other paper before???}
+
+\serena{insert a simple formula for linear regression}
+
+To further refine our analysis, we run three batches of linear regressions.
+First at the global level, we divide models by datasets and run regression separately on all models of a certain dataset, regardless of the architecture.
+At this level, the plots are quite noisy and clustered as each architecture has its own accuracy trend but, you could still see that most plots show positive relationship with positive coefficients
+
+\serena{see examples in Figure X}
+
+(Here we omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets as there are less than 15 models for those datasets and thus the regression is less statistically significant)
+
+\serena{insert plots for Figure X}
+
+For the second batch, we plot the regression for models of each architecture-datasets combination, which shows the relationship between the progression of the model accuracy and Weightwatcher metrics more clearly and precisely.
+For example, as you could see in the Figure X2,
+
+\serena{Add an example, UPDATE when we have the results from the new codes}
+
+Insert plots for Figure X2
+
+While running each regression, we record the R-squared and mean squared errors (MSE) for each regression.
+We then filter out regressions with less than five datapoints and models with structural outliers.
+
+\serena{Define and give an example of the structural outliers}
+
+Insert a plot for outliers, OPTIONAL

### 2020-01-18 all_models.tex
added results table

@@
 Insert a plot for outliers, OPTIONAL
+
+\charles{Still preliminary, only 309 data points here}
+\paragraph{Results}
+\begin{table}[t]
+\small
+\begin{center}
+\begin{tabular}{|p{1in}|c|c|c|c|}
+\hline
+    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
+    & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\langle\hat{\alpha}=\alpha\log\lambda_{max}\rangle$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
+\hline
+$R^{2}$ (mean)  & 0.61 & 0.59 & 0.66 & 0.66 \\
+$R^{2}$ (std)   & 0.35 & 0.37 & 0.30 & 0.30 \\
+\hline
+$MSE$ (mean)    & 3.60 &  7.19 & 2.44 & 2.28 \\
+$MSE$ (std)     & 7.64 & 20.13 & 4.57 & 4.43 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Comparison of linear regression fits for different average log norm metrics across XX datasets, covering 228 different pretrained DNNs.
+We only cinclude regressions for architecturs with 5 or more data points \charles{and which are postively correlated with the test error}.}
+\label{table:nlp}
+\end{table}

### 2020-01-18 all_models.tex
updated Table 3

@@
 \hline
-$R^{2}$ (mean)  & 0.61 & 0.59 & 0.66 & 0.66 \\
-$R^{2}$ (std)   & 0.35 & 0.37 & 0.30 & 0.30 \\
+$R^{2}$ (mean) & 0.63 &0.55 &0.64 &0.64 \\
+$R^{2}$ (std)  & 0.34 &0.36 &0.29 &0.30 \\
 \hline
-$MSE$ (mean)    & 3.60 &  7.19 & 2.44 & 2.28 \\
-$MSE$ (std)     & 7.64 & 20.13 & 4.57 & 4.43 \\
+$MSE$ (mean)   & 4.54 &9.62 &3.14 &2.92 \\
+$MSE$ (std)    & 8.69 &23.06 &5.14 &5.00 \\
 \hline
@@
 \end{center}
-\caption{Comparison of linear regression fits for different average log norm metrics across XX datasets, covering 228 different pretrained DNNs.
-We only cinclude regressions for architecturs with 5 or more data points \charles{and which are postively correlated with the test error}.}
+\caption{Comparison of linear regression fits for different average log norm metrics across XX datasets, covering 168 (out of 309) different pretrained DNNs.
+We only cinclude regressions for architecturs with 4 or more data points \charles{and which are postively correlated with the test error?}.
+These results can be readily reproduced using the Google Colab notebooks accompanying thie paper\cite{notebooks}}
 \label{table:nlp}

### 2020-01-18 all_models.tex
more tables added

@@
 
+\charles{What are the structrual outliers we chose ?}
+
 Insert a plot for outliers, OPTIONAL
+
+\charles{These tables could go to appendix. We need references}
+
+\begin{table}[t]
+\small
+\begin{center}
+\begin{tabular}{|p{1in}|c|}
+\hline
+Dataset & $\#$ of Models \\
+\hline
+imagenet-1k   &  78 \\
+svhn          &  30 \\
+cifar-100     &  30 \\
+cifar-10      &  18 \\
+cub-200-2011  &  12 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Datasets used}
+\label{table:datasets}
+\end{table}
+
+\begin{table}[t]
+\small
+\begin{center}
+\begin{tabular}{|p{3in}|c|}
+\hline
+Architecture & $\#$ of Models \\
+\hline
+ResNet                                     & 30 \\
+SENet/SE-ResNet/SE-PreResNet/SE-ResNeXt    & 24 \\
+DIA-ResNet/DIA-PreResNet                   & 18 \\
+ResNeXt                                    & 12 \\
+WRN                                        & 12 \\
+DLA                                        & 6 \\
+PreResNet                                  & 6 \\
+ProxylessNAS                               & 6 \\
+VGG/BN-VGG                                 & 6 \\
+IGCV3                                      & 6 \\
+EfficientNet                               & 6 \\
+SqueezeNext/SqNxt                          & 6 \\
+ShuffleNet                                 & 6 \\
+DRN-C/DRN-D                                & 6 \\
+ESPNetv2                                   & 6 \\
+HRNet                                      & 6 \\
+SqueezeNet/SqueezeResNet                   & 6 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Architectures used}
+\label{table:architectures}
+\end{table}
 
@@
 \end{center}
-\caption{Comparison of linear regression fits for different average log norm metrics across XX datasets, covering 168 (out of 309) different pretrained DNNs.
-We only cinclude regressions for architecturs with 4 or more data points \charles{and which are postively correlated with the test error?}.
+\caption{Comparison of linear regression fits for different average log norm metrics across 5 computer vision datasets, 17 Architectures, covering 168 (out of 309) different pretrained DNNs.
+We only conclude regressions for architecturs with 4 or more data points \charles{and which are postively correlated with the test error?}.
 These results can be readily reproduced using the Google Colab notebooks accompanying thie paper\cite{notebooks}}
-\label{table:nlp}
+\label{table:results}
 \end{table}

### 2020-01-18 all_models.tex
added prelim all models tex

@@
 The pretrained models and their accuracy metrics are summarized in the osmr github
+
+\charlesx{We actually don't run regressions on all these datasetsm, BUT we could present them in the Figure below to show that alpha is a good metric for these kinds of models, in contrast to the NLP, where alpha is frequently too large to run a regression}
+
+\begin{figure}[t]
+    \centering
+    \subfigure[ImageNet 1K]{
+        \includegraphics[width=4cm]{img/imagenet1k_alpha.png}
+        \label{fig:imagenet1k-alpha}
+    }
+    \qquad
+    \subfigure[ CIFAR 10 ]{
+        \includegraphics[width=4cm]{img/cifar10_alpha.png}
+        \label{fig:cifar10.alpha}
+    }
+    \qquad
+    \subfigure[ CIFAR 100 ]{
+        \includegraphics[width=4cm]{img/cifar100_alpha.png}
+        \label{fig:cifar100.alpha}
+    }
+    \qquad
+    \subfigure[ SVHN ]{
+        \includegraphics[width=4cm]{img/svhn_alpha.png}
+        \label{fig:svhn.alpha}
+    }
+    \qquad
+    \subfigure[ CUB 200 ]{
+        \includegraphics[width=4cm]{img/cub200_alpha.png}
+        \label{fig:cub200.alpha}
+    }
+    \caption{\charles{Preliminary charts:} Heavy Tailed Power Law exponent $\alpha$ vs. reported Top1 Test Accuracies for pretrained DNNs available\charles{ref} for 5 different data sets.}
+
+    \label{fig:dataset_alphas}
+\end{figure}
 

### 2020-01-18 all_models.tex
fixed GPT2 plots

@@
 
-\charlesx{We actually don't run regressions on all these datasetsm, BUT we could present them in the Figure below to show that alpha is a good metric for these kinds of models, in contrast to the NLP, where alpha is frequently too large to run a regression}
+\charles{We actually don't run regressions on all these datasetsm, BUT we could present them in the Figure below to show that alpha is a good metric for these kinds of models, in contrast to the NLP, where alpha is frequently too large to run a regression}
 

### 2020-01-18 nlp_models.tex
fixed something that couldn't compile
added serenas text
fixed GPT2 plots

@@
 GPT2 xl & 194 & 1.86 &1.92& 7.17 & 7.51 \\
-
-Explain why Log Norm is sooo large...alpha is off may need to prune the alpha
-
 \hline
@@
 \end{table}
+
+Explain why Log Norm is sooo large...alpha is off may need to prune the alpha
 
@@
 \end{figure}
+
+\charles{We have redo the spectral norm and remove the outliers}
 
@@
     \subfigure[ Power Law Exponent $\alpha$  ]{
-        \includegraphics[width=4.5cm]{img/GPT2_fnl_alpha_hist.png}
+        \includegraphics[width=4.5cm]{img/GPT2_all_alpha_hist.png}
         \label{fig:gpt-alpha-layer}
@@
     \subfigure[ Spectral Norm $\Vert\mathbf{W}\Vert_{\infty}$]{
-        \includegraphics[width=4.5cm]{img/GPT2_fnl_spectralnorm_hist.png}
+        \includegraphics[width=4.5cm]{img/GPT2_all_spectralnorm_hist.png}
         \label{fig:resnet-snorm-layer}
@@
     \subfigure[ Alpha-Norm $\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$]{
-        \includegraphics[width=4.5cm]{img/GPT2_fnl_alpha_weighted_hist.png}
+        \includegraphics[width=4.5cm]{img/GPT2_all_alpha_weighted_hist.png}
         \label{fig:gpt-pnorm-layer}

### 2020-01-18 nlp_models.tex
still need to fix snorm plots

@@
 The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
-Table \ref{table:nlp} reports results for the average log norm metrics, using \emph{weightwatcher (0.2.3)},
+Table \ref{table:nlp} reports results for the average log norm metrics, using \emph{weightwatcher (0.2.7)},
 and with fully reproducible Jupyter notebooks.\footnote{\url{https://github.com/CalculatedContent/kdd2020}}
@@
 \begin{figure}
-   \includegraphics[scale=0.25]{img/gpt-alpha-hist.png}
+   \includegraphics[scale=0.25]{img/GPT-alpha-hist.png}
    \caption{Comparison of heavy tailed power law exponent $\alpha$ for OpenAI GPT and GPT pretrained models}
@@
     }
-    \caption{Comparison of Power Law Exponnents, Spectral Norm, and Alpha-Norm for different size models in the GPT2 architecture series.    }
+    \caption{Comparison of Power Law Exponnents, Spectral Norm, and Alpha-Norm for different size models in the GPT2 architecture series.
+Spectral Norm dhistogram does not contain the first 2 layers because the values are unusually large.
+}
     \label{fig:gpt-alpha-layers}

### 2020-01-18 nlp_models.tex
still need to fix snorm plots
updated NLP plots

@@
 weight matrics \cite{Attn2017}. In contrast to the smaller pre-Activation maps arising in Cond2D layers,
-the Attention matrices ar
+the Attention matrices are significantly larger, and frequently under-correlated, with generally larger
+Heavy Tailed Power Law exponents $\alpha$ (when fitting the ESD).
 
-\paragraph{GPT vs GPT2}
+\paragraph{OpenAI GPT Models}
 Here, we use the \emph{WeightWatcher} tool to analyze the OpenAI GPT and GPT2 models, which gives us
 the opportunity to analyze the effect of both training the same model with different size data sets,
-These GPT models have generated significant media attention because of its remarkable ability to
- generate fake text, and it's potential misuse. 
+and increasing sizes of both the data set and architectures.
+These models have generated significant media attention because of their remarkable ability to
+ generate fake text that appears to be real and the potential misuse of this.
+For this reason, the original GPT model was trained on on a deficient data set, rendering
+the model interesting but not fully functional.  Later, the same GPT model 
 The original GPT and later GPT model have the same architecture and number of layers, but
-the original GPT was released having been trained on a deficient data set.  Recently,
+
  however, a much improved GPT2 model has been released, which is remarkably good at creating 
@@
 \begin{figure}
-   \includegraphics[scale=0.25]{img/GPT-alpha-hist.png}
-   \caption{Comparison of heavy tailed power law exponent $\alpha$ for OpenAI GPT and GPT pretrained models}
-   \label{fig:gpt-alpha-hist.png}
+    \centering
+    \subfigure[Power Law Exponent $\alpha$]{
+        \includegraphics[width=5cm]{img/GPT-alpha-hist.png}
+        \label{fig:GPT-alpha.hist.png}
+       }
+    \qquad
+    \subfigure[Log Spectral Norm $\log\Vert\mathbf{W}\Vert_{\infty}$]{
+        \includegraphics[width=5cm]{img/GPT-snorm-hist.png}
+        \label{fig:GPT-snorm.hist.png}
+       }
+
+   \caption{Comparison of Heavy Tailed Power Law exponents $\alpha$, and log Spectral Norms $\log\Vert\mathbf{W}\Vert_{\infty}$
+for the OpenAI GPT and GPT2 (small) pretrained models.}
+
 \end{figure}
@@
     \caption{Comparison of Power Law Exponnents, Spectral Norm, and Alpha-Norm for different size models in the GPT2 architecture series.
-Spectral Norm dhistogram does not contain the first 2 layers because the values are unusually large.
-}
+Spectral Norm dhistogram does not contain the first 2 layers because the values are unusually large.}
     \label{fig:gpt-alpha-layers}

### 2020-01-18 nlp_models.tex
updating text

@@
 the Attention matrices are significantly larger, and frequently under-correlated, with generally larger
-Heavy Tailed Power Law exponents $\alpha$ (when fitting the ESD).
+Heavy Tailed Power Law exponents $\alpha$ (when fitting the ESD).  Here, we breifly look at a few
+popular pretrained NLP DNNs to demonstrate that the Theory of Heavy Tailed Self-Regularization also
+applies to NLP models and not just CV models, and to highlight how to use the theory to identify
+poorly trained models where the empirical norm metrics will perform poorly.
 
 \paragraph{OpenAI GPT Models}
-Here, we use the \emph{WeightWatcher} tool to analyze the OpenAI GPT and GPT2 models, which gives us
+We use the \emph{WeightWatcher} tool to analyze the OpenAI GPT and GPT2 models, which gives us
 the opportunity to analyze the effect of both training the same model with different size data sets,
@@
 For this reason, the original GPT model was trained on on a deficient data set, rendering
-the model interesting but not fully functional.  Later, the same GPT model 
-The original GPT and later GPT model have the same architecture and number of layers, but
-
- however, a much improved GPT2 model has been released, which is remarkably good at creating 
-auto-genetated text.
+the model interesting but not fully functional.  Later, OpenAI released a much improved model--
+GPT2 (small)--which has the same architecture and number of layers as GPT, but has
+been trained on a larger and betterd data set (and with other changes), making it
+remarkably good at generating (near) human-quality fake text.  
+By comparing GPT and GPT2, we can indentify empirical markers indicating that
+a model has been poorly trained and may perform poorly when deployed.
 

### 2020-01-18 nlp_models.tex
updating NLP section'

@@
 GPT2 (small)--which has the same architecture and number of layers as GPT, but has
-been trained on a larger and betterd data set (and with other changes), making it
+been trained on a larger and better data set (and with other changes), making it
 remarkably good at generating (near) human-quality fake text.  
-By comparing GPT and GPT2, we can indentify empirical markers indicating that
-a model has been poorly trained and may perform poorly when deployed.
+By comparing the poorly trained GPT to the well trained GPT2, we
+indentify empirical indidcators for when  model has in fact
+been poorly trained and may perform poorly when deployed.
 

### 2020-01-18 methods.tex
moved methods nlp to nlp

@@
 
-However, in the GPT2 and other NLP models, we find that the first embedding layer(s) have unusually large eigenvalues,
-reflecting a lack of normalization.  For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$, 
-but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
- $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we layer all layers as-is (although one may to normalize
-the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier); including them does
-not significantly alter the results here.
-
 COMMENT ON HOW LOG NORM first and last layers behave, maybe somewhere else

### 2020-01-18 nlp_models.tex
moved methods nlp to nlp

@@
 
-Explain why Log Norm is sooo large...alpha is off may need to prune the alpha
+We find that in the pretrained GPT models, the first embedding layer(s) have unusually large eigenvalues,
+reflecting a lack of normalization.  For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$, 
+but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
+ $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we layer all layers as-is (although one may to normalize
+the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier); including them does
+not significantly alter the results here.
 

### 2020-01-18 nlp_models.tex
fixed GPT2 spectral norm
udpating nlp plots and tables

@@
 The projection layer concatenates the self-attention results into a vector (of dimension $768$).
-This gives $50$ large, typically sparse weight matrices, and they can be
- \emph{poorly correlated} when trained with insufficient data--as we shall see below.
+This gives $50$ large matrices.
 
@@
 
-The additional OpenAI GPT2 (English) models are: \emph{gpt-medium, got-large, and gpt-xl}, 
-having include $12, 24, 36, and 48$ layers, resp., with increasingly larger weight matrices.
+The OpenAI GPT2 (English) models are: \nred{gpt-small, gpt-medium, gpt-large, and gpt-xl}, \
+having include $12, 24, 36, \text{and }48$ layers, resp., with increasingly larger weight matrices.
 The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
@@
 \hline
- GPT & & & & &\\
-GPT2 & 50 & 2.05 &2.59& 9.78 & 10.03 \\
-GPT2 medium & 98 & 2.08 &2.58& 9.74 & 10.01 \\
-GPT2 large & 146 & 1.85 &1.99& 7.67 & 7.94 \\
-GPT2 xl & 194 & 1.86 &1.92& 7.17 & 7.51 \\
+GPT & 49 & 1.64  & 1.72 & 7.01 & 7.28 \\
+GPT (small) & 49 & 2.04  & 2.54& 9.62 & 9.87 \\
+GPT2 medium & 98 & 2.08 & 2.58& 9.74 & 10.01 \\
+GPT2 large & 146 & 1.85 & 1.99& 7.67 & 7.94 \\
+GPT2 xl & 194 & 1.86 & 1.92 & 7.17 & 7.51 \\
 \hline
@@
 \end{table}
-
-We find that in the pretrained GPT models, the first embedding layer(s) have unusually large eigenvalues,
-reflecting a lack of normalization.  For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$, 
-but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
- $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we layer all layers as-is (although one may to normalize
-the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier); including them does
-not significantly alter the results here.
 
@@
     \qquad
-    \subfigure[Log Spectral Norm $\log\Vert\mathbf{W}\Vert_{\infty}$]{
+    \subfigure[log Spectral Norm $\log\Vert\mathbf{W}\Vert_{\infty}$]{
         \includegraphics[width=5cm]{img/GPT-snorm-hist.png}
@@
 
-    \subfigure[ Power Law Exponent $\alpha$  ]{
+    \subfigure[Power Law Exponent $\alpha$]{
         \includegraphics[width=4.5cm]{img/GPT-alpha-depth.png}
@@
     \qquad
-    \subfigure[ Spectral Norm $\Vert\mathbf{W}\Vert_{\infty}$]{
+    \subfigure[log Spectral Norm $\log\Vert\mathbf{W}\Vert_{\infty}$]{
         \includegraphics[width=4.5cm]{img/GPT-snorm-depth.png}
@@
     \qquad
-    \subfigure[ Alpha-Norm $\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$]{
+    \subfigure[log Alpha-Norm $\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$]{
         \includegraphics[width=4.5cm]{img/GPT-pnorm-depth.png}
@@
 
+Notice the first embedding layer(s) have unusually large Spectral Norms,
+reflecting a lack of normalization.  For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,
+but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
+ $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we layer all layers as-is (although one may to normalize
+the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier); including them does
+not significantly alter the results here.
+
 \paragraph{GPT2: small, medium, large, extra-large}
@@
 althogh, the differences are less noticible than between the GPT and GPT2 models.
-Unlike GPT, however, the Layer Spectral Norms $\vert\mathbf{W}\Vert_{\infty}$ 
-and Alpha-Norms $\vert\mathbf{W}\Vert_{\alpha}^{\alpha}$
- behave  as expected for GPT2 layers, with the larger models 
-consistently having  smaller norms.
+Unlike GPT, however, the Layer (log) Spectral Norms $\log\vert\mathbf{W}\Vert_{\infty}$ 
+and (log) Alpha-Norms $\log\vert\mathbf{W}\Vert_{\alpha}^{\alpha}$
+ behave  as expected for GPT2 layers, with the larger models  consistently having  smaller norms.
 
@@
 
-    \subfigure[ Power Law Exponent $\alpha$  ]{
+    \subfigure[Power Law Exponent $\alpha$  ]{
         \includegraphics[width=4.5cm]{img/GPT2_all_alpha_hist.png}
@@
     \qquad
-    \subfigure[ Spectral Norm $\Vert\mathbf{W}\Vert_{\infty}$]{
-        \includegraphics[width=4.5cm]{img/GPT2_all_spectralnorm_hist.png}
+    \subfigure[log Spectral Norm $\log\Vert\mathbf{W}\Vert_{\infty}$]{
+        \includegraphics[width=4.5cm]{img/GPT2*_all_spectralnorm_hist.png}
         \label{fig:resnet-snorm-layer}
@@
     \qquad
-    \subfigure[ Alpha-Norm $\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$]{
-        \includegraphics[width=4.5cm]{img/GPT2_all_alpha_weighted_hist.png}
+    \subfigure[log Alpha-Norm $\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$]{
+        \includegraphics[width=4.5cm]{img/GPT2_all_logpnorm_hist.png}
         \label{fig:gpt-pnorm-layer}
@@
     \caption{Comparison of Power Law Exponnents, Spectral Norm, and Alpha-Norm for different size models in the GPT2 architecture series.
-Spectral Norm dhistogram does not contain the first 2 layers because the values are unusually large.}
+*Note: the log spectral norms (b) histogram omits the first 2 layers, corresponding to the embedding layer, which are normalized differently and have anamolously large values.}
     \label{fig:gpt-alpha-layers}

### 2020-01-18 nlp_models.tex
updated NLP section

@@
 
-\paragraph{The Heavy Tailed Power Law Exponents in GPT and GPT2}
-
-are very differenmt, with GPT2 having both a notably smaller mean $\alpha$, and far fewer, unusually large outliers.
+\paragraph{Empirical Quality Metrics for GPT and GPT2}
+We first compare the distribution of  Heavy Tailed Power Law exponents $\alpha$ in GPT and GPT2.
+They are very differenmt, with GPT2 having both a notably smaller mean $\alpha$, and far fewer, unusually large outliers.
 Figure \ref{fig:gpt-alphs-hist} shows the empirical density (histogram) of $\alpha$
@@
         \includegraphics[width=5cm]{img/GPT-alpha-hist.png}
-        \label{fig:GPT-alpha.hist.png}
+        \label{fig:GPT-alpha-hist}
        }
@@
         \includegraphics[width=5cm]{img/GPT-snorm-hist.png}
-        \label{fig:GPT-snorm.hist.png}
+        \label{fig:GPT-snorm-hist}
        }
@@
 
-\charles{We have redo the spectral norm and remove the outliers}
+Indeed, the $\alpha$ metric makes a good quality metric for comparing these models.
+Figure \ref{fig:GPT-alpha-hist} compares the PL exponent $\alpha$ for the GPT and GPT2 models for all layers,
+As expected, the improved GPT2 model has, on average, smaller $\alpha$ than the older GPT, with all
+$\alpha\le6$.  Indeed, the deficient GPT model has numerous weight matrices with unsually large fitted exponents,
+indicating that they are not Heavy Tailed at all.  Indeed, we may expect that a poorly trained model
+will not exhibit consistent Heavy Tailed behavior in all layers.
+
+However, as seen in Figure \ref{fig:GPT-snorm-hist},
+ the poorly trained GPT model has many smaller (log) Spectral Norms $\log\Vert\mathbf{W}\Vert_{\infty}$
+than the GPT2, which would be inconsistent with a belief that smaller Spectral Norm is always better.
+Indeed, because there are so many anamously small $\Vert\mathbf{W}\Vert_{\infty}$,
+it appears that the GPT model may be exhibiting a kind of rank collapse.
+This is an extremely important observation because it demonstrates that while the Spectral Norm
+may correlate well with predicted test error, it is not a good indicator of the overall quality of a model,
+and using it as an empirical metric may give spurious results when applied to poorly trained
+or otherwise deficient models.
+
+Note that Figure \ref{fig:GPT-snorm-hist} also shows a couple anomalously  large Spectral Norms.
+From Figure \ref{fig:gpt-snorm-layer} (below), we see that these correspond to the first embedding layer(s).
+These layers appear to have a normalization, and therefore a differenc scale.
+For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,
+but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
+ $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we layer all layers as-is (although one may to normalize
+the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier).
+Here, we do not include them in our computed average metrics in Table \ref{table:nlp},
+and do not include them in the histogram plot in Figure \ref{fig:GPT-snorm-hist}.
 
 \paragraph{The Correlation Flow in GPT and GPT2} also differs significantly between GPT and GPT2.
-Figure \ref{fig:gpt-alpha-layer} plots $\alpha$ vs the layer id for each model.
+Figure \ref{fig:gpt-alpha-layer} plots $\alpha$ vs the depth (i.e. a layer id) for each model.
 
@@
         \includegraphics[width=4.5cm]{img/GPT-snorm-depth.png}
-        \label{fig:resnet-snorm-layer}
+        \label{fig:gpt-snorm-layer}
     }
@@
 \end{figure}
-
-Notice the first embedding layer(s) have unusually large Spectral Norms,
-reflecting a lack of normalization.  For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,
-but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
- $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we layer all layers as-is (although one may to normalize
-the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier); including them does
-not significantly alter the results here.
 

### 2020-01-18 nlp_models.tex
more NLP text

@@
     \centering
-    \subfigure[Power Law Exponent $\alpha$]{
+    \subfigure[Power Law Exponent $(\alpha)$]{
         \includegraphics[width=5cm]{img/GPT-alpha-hist.png}
@@
     \qquad
-    \subfigure[log Spectral Norm $\log\Vert\mathbf{W}\Vert_{\infty}$]{
+    \subfigure[log Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
         \includegraphics[width=5cm]{img/GPT-snorm-hist.png}
@@
 
-    \subfigure[Power Law Exponent $\alpha$]{
+    \subfigure[Power Law Exponent $(\alpha)$]{
         \includegraphics[width=4.5cm]{img/GPT-alpha-depth.png}
@@
     \qquad
-    \subfigure[log Spectral Norm $\log\Vert\mathbf{W}\Vert_{\infty}$]{
+    \subfigure[log Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
         \includegraphics[width=4.5cm]{img/GPT-snorm-depth.png}
@@
     \qquad
-    \subfigure[log Alpha-Norm $\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$]{
+    \subfigure[log Alpha-Norm $(\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha})$]{
         \includegraphics[width=4.5cm]{img/GPT-pnorm-depth.png}
@@
 
-\paragraph{GPT2: small, medium, large, extra-large}
+\paragraph{GPT2: small, medium, large, xl}
 
-Figure \ref{fig:gpt-alpha-layers} ...
+We now look across the series of increasingly improving GPT2 models, examining both our
+PL exponent $\alpha$, as well as the various Norm metrics.
+As we move from small to xl, both the PL exponents $\alpha$, and the peak of distribtion of
+log Norm metrics shifts from larger to smaller values. 
+Figure \ref{fig:gpt2-histograms} shows the histograms over the layer weight matrics
+for fitted $\alpha$, and the log Spectral Norm
+ $\log\Vert\mathbf{W}\Vert_{\infty}$  
+and log Alpha-Norm
+ $\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$ 
+metrics.
 
-For this series of GPT2 models, the average $\alpha$ still decreases with increasing model size,
-althogh, the differences are less noticible than between the GPT and GPT2 models.
+We see that the average $\alpha$ decreases with increasing model size,
+althogh, the differences are less noticible between GTP2 mkodels than between the GPT and GPT2 models.
 Unlike GPT, however, the Layer (log) Spectral Norms $\log\vert\mathbf{W}\Vert_{\infty}$ 
 and (log) Alpha-Norms $\log\vert\mathbf{W}\Vert_{\alpha}^{\alpha}$
- behave  as expected for GPT2 layers, with the larger models  consistently having  smaller norms.
+behave more as expected for GPT2 layers, with the larger models  consistently having  smaller norms. 
+Likewise, Figure \ref{fig:gpt2-snorm-hist} shows decreasing average log Spectral Norms
+with the larger models.   As we have seen in the trends of other well trained models.
+
+We do notice, however, that while the peaks of the $\alpha$ is getting smaller, towards $2.0$,
+the tails of the distribution shifts right, with larger GPT2 models having more
+usually large $\alpha$.  We suspect this indicates that these larger GPT2 models are 
+overparameterized and/or not yet fully optimized and require datasets even larger
+than the recent XL $1.5B$ release\cite{gpt2-xl}.
 
@@
 
-    \subfigure[Power Law Exponent $\alpha$  ]{
+    \subfigure[Power Law Exponent $(\alpha)$]{
         \includegraphics[width=4.5cm]{img/GPT2_all_alpha_hist.png}
-        \label{fig:gpt-alpha-layer}
+        \label{fig:gpt2-alpha-hist}
     }
     \qquad
-    \subfigure[log Spectral Norm $\log\Vert\mathbf{W}\Vert_{\infty}$]{
+    \subfigure[log Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
         \includegraphics[width=4.5cm]{img/GPT2*_all_spectralnorm_hist.png}
-        \label{fig:resnet-snorm-layer}
+        \label{fig:gpt2-snorm-hist}
     }
     \qquad
-    \subfigure[log Alpha-Norm $\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$]{
+    \subfigure[log Alpha-Norm $(\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha})$]{
         \includegraphics[width=4.5cm]{img/GPT2_all_logpnorm_hist.png}
-        \label{fig:gpt-pnorm-layer}
+        \label{fig:gpt2-pnorm-hist}
     }
@@
 *Note: the log spectral norms (b) histogram omits the first 2 layers, corresponding to the embedding layer, which are normalized differently and have anamolously large values.}
-    \label{fig:gpt-alpha-layers}
+    \label{fig:gpt2-histograms}
 \end{figure}

### 2020-01-18 cv_models.tex
updated DenseNet
cv table added
comments add for mike

@@
 \label{sxn:cv}
+
+\michael{We set up the methdology we will use later, and give examples on the CV models
+and explain how to interpret things a bit. We study pretrained CV models.
+We look at correlation with the reportest test accuarcies.
+ We will run linear regressions. We will compute the RMSE.  
+We study alpha and the 4 norms.  we could add softrank with some work.
+
+We introduce the notion of Correlation Flow and 
+tell a nice story about residual connections
+with VGG RseNet and DenseNet}
 
@@
 \hline
- VGG & & & & & \\
- ResNet & & & & & \\
- ResNet-1K & & & & & \\
- DenseNet & & & &  &\\
+ VGG & 6 & 0.56 & 0.53 & 0.48 & 0.42  \\
+ ResNet & 5 & 0.9 & 1.4 & 0.61 & 0.66  \\
+ ResNet-1K & 19 & 2.4 & 3.6 & 1.8 & 1.9  \\
+ DenseNet & 4 & 0.3 & 0.26 & 0.16 & 0.21 \\
 \hline
@@
 \end{table}
+
+DISCUSS TABLE
 
@@
     \subfigure[ DenseNet ]{
-        \includegraphics[width=4.5cm]{img/DenseNet_all_depth.png} 
+        \includegraphics[width=4.5cm]{img/DenseNet_fnl_alpha_depth.png} 
         \label{fig:densenet-alpha-layer}

### 2020-01-18 nlp_models.tex
comments add for mike

@@
 
-[THESE RESULTS HAVE TO BE REDONE BY REMOVING THE FIRST LAYER...]
+\michael{We show that this not only works for CV, so we look at NLP.
+We now look in detail at the alpha metric, the spectral norm, and the new norm.
+We compare good and bad models, and a series of increasingly big models
+trained on big data sets.  We show why alpha works but spectral norm is not enough
+
+}
 

### 2020-01-20 methods.tex
updated methods

@@
 
+\paragraph{Organization of this paper}
+
+We want to study the effective of Product Norm Metrics in describing trends in the
+generalization accuracyof real-word, production quality, pretrained Deep Neural Networks.
+In Section \ref{sxn:cv}, we study three well known, and widely available DNN architectures:
+the VGG series, the ResNet series, and the DenseNet series of models.  Each have been pretrained
+on the full ImageNet dataset\cite{imagenet}, and are distributed with the current of pyTorch (1.4)\cite{pyTorch}.
+[Explain more]
+
 ----

### 2020-01-20 methods.tex
added org sec

--- 2020-01-18 methods.tex
@@
 
+\paragraph{Organization of this paper}
+
+We want to study the effective of Product Norm Metrics in describing trends in the
+generalization accuracyof real-word, production quality, pretrained Deep Neural Networks.
+In Section \ref{sxn:cv}, we study three well known, and widely available DNN CV architectures:
+the VGG series, the ResNet series, and the DenseNet series of models.  We lay out the
+basic methodology we use to evaluate the different metrics against the reported test accuracies. 
+We also introduce the notion of Correlation Flow, compare the three pretrained architectures in detail.  
+In Section  \ref{sxn:nlp}, we then look at several variations of a popular NLP DNN architecture,
+the OpenAI GPT and GPT2 models.  We discuss when and how the different metrics apply more generally.
+Finally, in Section \ref{sxn:all}, we examine over 100 pretraind DNN CV models, and show
+how well each metric is correlated with the reported test accuracies across different models
+in an architeture series.  
 ----

### 2020-01-20 methods.tex
minor update

--- 2020-01-18 methods.tex
@@
 
+\paragraph{Organization of this paper}
+\charles{This could be moved to the end of the into also}
+We want to study the effectiveness of product norm metrics in predicting trends in the
+generalization accuracy of real-word, production quality, pretrained Deep Neural Networks.
+In Section \ref{sxn:cv}, we study three well known, and widely available DNN CV architectures:
+the VGG, the ResNet, and the DenseNet series of models.  We lay out the
+basic methodology we use to evaluate the different metrics against the reported test accuracies. 
+We also introduce the notion of Correlation Flow, and use this to 
+compare the three pretrained architectures in detail.  
+Then, in Section  \ref{sxn:nlp}, we look at several variations of a popular NLP DNN architecture,
+the OpenAI GPT and GPT2 models.  We discuss when and how the different metrics apply more generally,
+and where they differ.  In particular, we show where the layer Spectral Norm
+ $\Vert\mathbf{W}\Vert_{\infty}$ behaves unexpectedly, and, contrastly, how
+the $\alpha$ parameter can give insight into how well trained a model is.
+Finally, in Section \ref{sxn:all}, we examine over 100 pretraind DNN CV models, and show
+how well each metric is correlated with the reported test accuracies across different models
+in an architeture series, and that the Alpha-Norm metric performs better.
 ----

### 2020-01-20 methods.tex
minor update

@@
 how well each metric is correlated with the reported test accuracies across different models
-in an architeture series, and that the Alpha-Norm metric performs better.
+in an architeture series, and argue the Alpha-Norm metric(s) perform remarkably well.
 ----

### 2020-01-20 cv_models.tex
minor update

--- 2020-01-15 cv_models.tex
@@
 \label{sxn:cv}
+
+We start our empirical studies by compute the differernt log 
+product norm metrics for three computer vision (CV) architectures,
+the VGG, and ResNet, and the DenseNet series.  Each
+series consists of several pretrained DNN models, 
+trained on the full ImageNet\cite{imagenet} dataset,
+and distributed with the current opensource pyTorch framework (vesion 1.4)\cite{pyTorch}.
+We also provide results for a larger set of ResNet models,
+trained on the ImageNet-1K dataset\cite{imagenet1k},  provided 
+on the OSMR ``Sandbox for training convolutional networks for computer vision''\cite{osmr};
+we call this the ResNet-1K series of models.
+We compute the metrics using the WeightWatcher tool (version 0.2.7),
+and provide Jupyter notebooks in the github repo for this paper\cite{repo},
+making the results fully reproducible.
+
+Figure \ref{fig:vgg-metrics} plots the different log norm metrics
+versus the reported Test accuracies\cite{pyTorchVgg}, for the
+pretrained models:  VGG11, VGG13, VGG16, and VGG19, with and without
+BatchNormalization.  The figure also shows a basic linear regression
+line, with the Root Mean Squared Error (RMSE) shown.
+All four metrics correlate well with the reported Top1 Accuracies,
+with smaller norms implying better generalziation 
+(greater accuracy, lower error).  Similar results are obtained
+for the Top5 Accuracies (not shown).  Notice that the log $\alpha$-Norm
+metric $\log\Vert\mathbf{W}\Vert)_{\infty}^{\infty}$ performs best,
+with an RMSE of $0.34$.  Also, the next best metric is the
+Weighted Alpha, $\hat\alpha$, which is an approximate log $\alpha$-Norm metric.
 
@@
 
-VGG works remarkably well
+Table \ref{table:cv-models} summarizes results from linear regressions, applied to the VGG, ResNet, ResNet-1K, and DenseNet series,
+and reports the RMSE for all four product norm metrics.  For all series, all four metrics correlates remarkably well with
+the reported Top1 test accuracies.  Moreover, the $\alpha$-Norm and/or the approximate Weighted Alpha metrics perform best.
 
-ResNet is also correlated, but the RMSE is much larger, and 2/5 of the models are almost outliers
-But note that the PyTorch distribution only has a few models.
-The ResNet results look better when considering all ResNet models, trained on ImageNet1K.
-
-ADD PLOTS FOR
+For a more visual depiction, Figure \ref{fig:cv2-accuracy} displays the log $\alpha$-Norm metric  $\log\Vert\mathbf{W}\Vert)_{\infty}^{\infty}$ 
+versus test accuracies for the ResNet, ResNet-1K, and DenseNet series. Notice that ResNet series, which has been trained on the full
+ImageNet dataset, has an RMSE of $0.66$, whereas the ResNet-1K series, which has been trained on the much smaller ImageNet-1K dataset,
+also correlated wells, but has a much larger RMSE of $1.9$.
 
@@
 
-TODO:  add table of metrics for these 3 models, show that alpha-Norm is best again!
-
 \begin{table}[t]
@@
 }
-\label{table:models}
+\label{table:cv-models}
 \end{table}
@@
 
-Explain WHAT IS \emph{Correlation Flow}
+We can learn even more about how a pretrained model will perform by examining what we call the
+\emph{Correlation Flow} down the layers.  By this, we mean the values of individual Heavy Tailed
+Power Law exponent $\alpha$ for each layer weight matric $\mathbf{W}$, as a function of depth
+(or layer id) in the network. (The first layer corresponds to the data, and the last layer the labels).
+Figure \ref{fig:vgg-alpha-layers} plots $\alpha$ for each layer for the VGG (no batchnorm), RseNet-1K, and DenseNet models,
+including the least accurate and most accurate model in each series.
 
-Compare VGG, ResNet, and DenseNet in the context of how many connections that have
+First observe that in the VGG models, $\alpha$ systematically increases as we move down the network from data to labels,
+in the Conv2D layers, starting with $\alpha\le 2.0$ and reaching all the way to $\alpha\sim 5.0$.
+Then, $\alpha\in[2,2.5]$ stabilizes back down to last three, large,  fully connected (FC) layers. 
+In contrast, in the larger ResNet-1K (152) model, $\alpha$ remaims stable at $\alpha\sim 2.0$ all the way
+until we reach the final layers.  Finally, in the DenseNet models, as in the VGG models, $\alpha$ increases
+as the layer id increases, although it displays a large variance, and can range all the way to $alpha\le 8.0$.
+
+\charles{INTERPRET}
 
@@
     \qquad
-    \subfigure[ ResNet ]{
+    \subfigure[ ResNet-1K ]{
         \includegraphics[width=4.5cm]{img/ResNet_fnl_alpha_depth.png} 
@@
     \subfigure[ DenseNet ]{
-        \includegraphics[width=4.5cm]{img/DenseNet_all_depth.png} 
+        \includegraphics[width=4.5cm]{img/DenseNet_fnl_alpha_depth.png} 
         \label{fig:densenet-alpha-layer}
     }
-    \caption{Power law exponent $\alpha$ vs layer for VGG, ResNet, and DenseNet models}
+    \caption{Correlation Flow: Heavy Tailed Power Law exponent $\alpha$ vs layer for VGG, ResNet-1K, and DenseNet models.}
     \label{fig:vgg-alpha-layers}

### 2020-01-20 cv_models.tex
minor update

@@
 \end{figure}
+
+We can understand the Correlation Flow by considering the differences between the VGG, ResNet, and DenseNet architectures,
+and, in particular, the number of residual connections.  VGG, while a good model, had been limited by needing the massive
+FC layers at the end of the model,requiring significant memory and computational resources.  ResNet greatly improved
+on VGG by introducing residual connections, allowing for greater accurary with far fewer parameters.
+Indeed, ResNet models of upto 1000 layers have been trained. And we see this in the Corraltion Flow.
+We conjuecture that the effeciency and effectivness of ResNet is reflected in the smaller and stable
+$\alpha\sim 2.0$,  indicating that the inner layers are very well correlated, really optimized.
+In contrast, the DenseNet model contains a massive number of connections and between every layer.
+Here, we suppose that DenseNet has too many connections, diluting the Correlation Flow, leaving
+many layers very poorly correlated, with large $\alpha$, meaning they are probably not even Heavy Tailed.

### 2020-01-24 abstract_file.tex
split text of abstract into separate file

In many applications,
% of machine learning and artificial intelligence, 
one must work with models that have been trained by someone else.
For such \emph{pretrained models}, one typically does not have access to training data or test data, and one does not know the details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
Here, we develop and evaluate quality metrics for pretrained neural network models at scale.
%%Our metrics are drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as Heavy-Tailed Random Matrix Theory (HT-RMT), as used in the recently-developed Theory of Heavy-Tailed Self Regularization (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models).
Most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision and natural language processing.
We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best,'' but that HT-SR metrics do much better.
We also find that, for models that are not well-trained (which, arguably is the point of needing metrics to evaluate the quality of pretrained models), norm-based metrics can qualitatively fail, while HT-SR metrics still go a very good job at predicting trends in model quality.
HT-SR metrics can also be used to characterize fine-scale properties of models, such as layer-wise \emph{correlation flow} and XXX SOMETHING ELSE.

### 2020-01-30 abstract_file.tex
mwm initial pass

@@
 For such \emph{pretrained models}, one typically does not have access to training data or test data, and one does not know the details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
-Here, we develop and evaluate quality metrics for pretrained neural network models at scale.
+Here, we present and evaluate quality metrics for pretrained neural network models at scale.
 %%Our metrics are drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as Heavy-Tailed Random Matrix Theory (HT-RMT), as used in the recently-developed Theory of Heavy-Tailed Self Regularization (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models).
-Most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
+The most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
 Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision and natural language processing.
-We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best,'' but that HT-SR metrics do much better.
-We also find that, for models that are not well-trained (which, arguably is the point of needing metrics to evaluate the quality of pretrained models), norm-based metrics can qualitatively fail, while HT-SR metrics still go a very good job at predicting trends in model quality.
-HT-SR metrics can also be used to characterize fine-scale properties of models, such as layer-wise \emph{correlation flow} and XXX SOMETHING ELSE.
+We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best.'' 
+On the other hand, for models that may not be well-trained (which, arguably is the point of needing metrics to evaluate the quality of pretrained models), i.e., when we want to distinguish ``good-bad,'' norm-based metrics can qualitatively fail.
+We also find that HT-SR metrics do much better, quantitatively better for at discriminating good-better-best and qualitatively better at discriminating good-bad.
+HT-SR metrics can also be used to characterize fine-scale properties of models, e.g., understanding layer-wise \emph{correlation flow}, and evaluate post-training modifications such as model distillation and model improvement.

### 2020-01-30 all_models.tex
comments add for mike
mwm initial pass

@@
 \label{sxn:all_cv_models}
+
+\michael{Now we go back to the CV models and look at over 100 of them, and draw'
+more broad conclusions about why alpha and the new norm is better, by looking at the MSE}
 
@@
 We only conclude regressions for architecturs with 4 or more data points \charles{and which are postively correlated with the test error?}.
-These results can be readily reproduced using the Google Colab notebooks accompanying thie paper\cite{notebooks}}
+These results can be readily reproduced using the Google Colab notebooks accompanying thie paper~\cite{notebooks}}
 \label{table:results}

### 2020-01-30 background.tex
cleaning up files
mwm initial pass

@@
 
-Here we will cite and discuss related work, including
+To our knowledge, there is very little work on the particular question we are addressing: namely, how to predict, in a theoretically-principled manner, the quality of large-scale state-of-the-art NNs, and to do so without access to training data or testing data or details of the training protocol, etc.
+Our work is, however, loosely related to several other lines of work, and we briefly discuss them here.
+
+\paragraph{Statistical mechanical theory of NNs.}
+
+XXX.
+Cite our stuff:
 \cite{MM17_TR},
@@
 \cite{MM20_SDM},
-\cite{MM20_unpub_work},
+\cite{MM20_unpub_work}.
+Cite also Ganguli review and maybe other stuff.
+
+XXX.
+Distinguish between what we will call a
+\emph{phenomenological theory}
+(that describes empirical relationship of phenomena to each other, in a way which is consistent with fundamental theory, but is not directly derived from that theory)
+and what can be called a 
+\emph{first principles theory} 
+(that is applicable to tiny things but has no hope of scaling up).
+\footnote{In most areas where there are complex highly-engineered systems (beyond complex AI/ML systems), one used phenomenological theory rather than first principles theory.
+For example, one does not try to solve the Schr\"odinger equation if one is interested in building a bridge or an airplane.}
+
+\paragraph{Norm-based capacity control theory.}
+XXX.  MOST IRRELEVANT, BUT LIAO AND OUR SDM ARE RELATED.  
+XXX.  MAYBE BEAT ON INFINITELY WIDE OR SOMETHING ELSE.
+
+\paragraph{Practical problems poorly addressed by theory.}
+There are many very practical problems in ML that are poorly addressed by existing theory and that either motivated our work or should be addressable by our techniques.
+Here are several.
+\begin{itemize}
+\item
+\textbf{Lack of accuracy metrics.}
+\item
+\textbf{Lack of data.}
+\item
+XXX. LACK OF SOMETHING ELSE.
+\end{itemize}
+Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text, when developing self driving car systems, and when distilling a reliable model in some way to obtain comparable training/test quality but that damages the model in some other subtle way.
+\michael{That last example is awkward.  It would be good to have a better example and plant seeds for model distillation elsewhere.}

### 2020-01-30 conclusion.tex
mwm initial pass

@@
 \label{sxn:conc}
-XXX.  CONCLUSION.
+
+XXX.  PUT CONCLUSION HERE AND WEAVE IN COMMENTS FROM BELOW.
+
+Some other comments that we need to weave into a narrative eventually after later sections are written:
+\begin{itemize}
+\item
+GPT versus GPT2.
+What happens when we don't have enough data?
+This is the main question, and we can use out metrics to evaluate that, but we also get very different results for GPT versus GPT2.
+\item
+The spectral norm is a regularizer, used to distinguish good-better-best, not a quality metric.
+For example, it can ``collapse,'' and for bad models we can have small spectral norm.
+So, it isn't really a quality metric.
+\item
+One question that isn't obvious is whether regularization metrics can be used as quality metrics.
+One might think so, but the answer isn't obviously yes.
+We show that the answer is No.
+A regularizer is designed to select a unique solution from a non-unique good-better-best.
+Quality metrics can also distinguish good versus bad.
+\item
+(We should at least mention this is like the statistical thing where we evaluate which model is better, as oposed to asking if a given model is good, I forget the name of that.)
+\item
+There are cases where the model is bad but regularization metric doesn't tell you that.
+Quality should be correlated in an empirical way.
+Correlated with good-better-best; but also tell good-bad.
+\item
+Question: why not use regularier for quality?
+Answer: A regularizer selects from a given set of degenerate models one which is nice or unique.
+It doesn't tell good versus bad, i.e., whether that model class is any good.
+\item
+Thus, it isn't obvious that norm-based metrics should do well, and they don't in general.
+\item
+We give examples of all of these: bad data; defective data; and distill models in a bad way.
+(Of course, bad data means bad model, at least indirectly, since the quality of the data affects the properties of the model.)
+\item
+We can select a model and change it, i.e., we don't just do hyperparameter fiddling.
+\end{itemize}
+
+%Perhaps our most improtant contribution, however, is just doing things different, etc. ...

### 2020-01-30 cv_models.tex
mwm initial pass

--- 2020-01-18 cv_models.tex
@@
-\section{Comparison of 3 Computer Vision Models}
+\section{Comparison across series of CV models}
 \label{sxn:cv}
 
-\michael{We set up the methdology we will use later, and give examples on the CV models
-and explain how to interpret things a bit. We study pretrained CV models.
-We look at correlation with the reportest test accuarcies.
- We will run linear regressions. We will compute the RMSE.  
-We study alpha and the 4 norms.  we could add softrank with some work.
+In this section, we examine the quality metrics described in Section~\ref{sxn:methods} for several CV model architecture series.
+This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current opensource pyTorch framework (version 1.4)~\cite{pyTorch}.
+This also includes results for a larger set of ResNet models, trained on the ImageNet-1K dataset~\cite{imagenet1k}, provided on the OSMR ``Sandbox for training convolutional networks for computer vision''~\cite{osmr}, which we call the ResNet-1K series.
+We use our quality metrics to understand the gross properties of these models, comparing and contrasting the three (of four, if ResNet-1K is a distinct series) model series.
+We also use our metrics to understand more fine-scale properties of these models as a function of model depth.
+\michael{And mention distillation, if we include that in this section.}
 
-We introduce the notion of Correlation Flow and 
-tell a nice story about residual connections
-with VGG RseNet and DenseNet}
+\paragraph{Quality Metrics versus Reported Test Accuracies.}
 
-\paragraph{Empirical Metrics vs Test Accuracies for CV models}
+We have examined the performance of the four quality (Frobenius norm, Spectral norm, Weighted Alpha, and $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
+To start, Figure~\ref{fig:vgg-metrics} plots the four quality metrics versus the reported Test accuracies~\cite{pyTorchVgg}, as well as a basic linear regression line, 
+%with the Root Mean Squared Error (RMSE) shown, 
+for the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BatchNormalization.
+\michael{These are all log norms, right?}
+\michael{XXX.  WHERE IS RMSE.  IN THAT TABLE?}
+All four metrics correlate quite well with the reported Top1 Accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalziation (i.e., greater accuracy, lower error). 
+Similar results (not shown) are obtained for the Top5 Accuracies.  
+Notice that the log $\alpha$-Norm metric $\log\Vert\mathbf{W}\Vert)_{\infty}^{\infty}$ 
+\michael{XXX IS THAT A TYPO} 
+performs best, with an RMSE of $0.34$; and the second best metric is $\hat\alpha$, which is an approximate log $\alpha$-Norm metric.
+\michael{Are those numbers stale, I don't see them, and some things seem in the other order.}
+\michael{We should probably point to the table if we cite quantitative results.}
+
+Figure~\ref{fig:cv2-accuracy} plots the best performing metric ($\alpha$-Norm) for the full ResNet, ResNet-1K, and DenseNet series.
+(A much more detailed set of plots---including the behavior of all four metrods on each of the series---are availale at \michael{give web link}.)
+For a more visual depiction, Figure~\ref{fig:cv2-accuracy} displays the log $\alpha$-Norm metric  $\log\Vert\mathbf{W}\Vert)_{\infty}^{\infty}$ versus test accuracies for the ResNet, ResNet-1K, and DenseNet series.
+Notice that ResNet series, which has been trained on the full
+ImageNet dataset, has an RMSE of $0.66$, whereas the ResNet-1K series, which has been trained on the much smaller ImageNet-1K dataset,
+also correlated wells, but has a much larger RMSE of $1.9$.
+
+Table \ref{table:cv-models} summarizes results for all four metrics (Frobenius norm, Spectral norm, Weighted Alpha, and $\alpha$-Norm) applied to the VGG, ResNet, ResNet-1K, and DenseNet series.
+\michael{XXX.  AGAIN, WHERE IS RMSE, IN THAT TABLE?}
+
+For all series, all four metrics correlate quite well with the reported Top1 test accuracies.  
+Moreover, the $\alpha$-Norm and/or the approximate Weighted Alpha metrics perform best.
 
@@
     \subfigure[ Frobenius Norm ]{
-        \includegraphics[width=5cm]{img/VGG_lognorm_accs.png}
+        %\includegraphics[width=5cm]{img/VGG_lognorm_accs.png}
+        \includegraphics[width=3.7cm]{img/VGG_lognorm_accs.png}
         \label{fig:vgg-fnorm}
     }
-    \qquad
+    %\qquad
     \subfigure[ Spectral Norm ]{
-        \includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
+        %\includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
+        \includegraphics[width=3.7cm]{img/VGG_spectralnorm_accs.png}
         \label{fig:vgg-snorm}
     }
-    \qquad
+    %\qquad
     \subfigure[ Weighted Alpha ]{
-        \includegraphics[width=4.9cm]{img/VGG_alpha_weighted_accs.png}
+        %\includegraphics[width=4.9cm]{img/VGG_alpha_weighted_accs.png}
+        \includegraphics[width=3.7cm]{img/VGG_alpha_weighted_accs.png}
         \label{fig:vgg-walpha}
     }
-    \qquad
+    %\qquad
     \subfigure[ Alpha-Norm ]{
-        \includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
+        %\includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
+        \includegraphics[width=3.7cm]{img/VGG_logpnorm_accs.png}
         \label{fig:vgg-pnorm}
     }
-    \caption{Comparison of norm metrics vs reported test accuracy for pretrained VGG models, trained on ImageNet, available in pyTorch.  Plots will be updated and replace }
-
+    \caption{Comparison of quality metrics versus reported test accuracy for pretrained VGG models, trained on ImageNet, available in pyTorch.  
+             XXX Plots will be updated and replaced. 
+             \michael{Figures are unreadable, can we make fonts bigger, etc.}
+            }
     \label{fig:vgg-metrics}
 \end{figure}
-
-VGG works remarkably well
-
-ResNet is also correlated, but the RMSE is much larger, and 2/5 of the models are almost outliers
-But note that the PyTorch distribution only has a few models.
-The ResNet results look better when considering all ResNet models, trained on ImageNet1K.
-
-ADD PLOTS FOR
 
@@
     \subfigure[ ResNet ]{
-        \includegraphics[width=4.2cm]{img/ResNet_logpnorm_accs.png}
+        %\includegraphics[width=4.2cm]{img/ResNet_logpnorm_accs.png}
+        \includegraphics[width=3.7cm]{img/ResNet_logpnorm_accs.png}
         \label{fig:resnet-accuracy}
@@
     \subfigure[ ResNet-1K ]{
-        \includegraphics[width=4.5cm]{img/ResNet-1K_logpnorm_accs.png}
+        %\includegraphics[width=4.5cm]{img/ResNet-1K_logpnorm_accs.png}
+        \includegraphics[width=3.7cm]{img/ResNet-1K_logpnorm_accs.png}
         \label{fig:resnet1k-accuracy}
@@
     \subfigure[ DenseNet ]{
-        \includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
+        %\includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
+        \includegraphics[width=3.7cm]{img/DenseNet_logpnorm_accs.png}
         \label{fig:densenet-accuracy}
     }
-    \caption{$\alpha$-Norm vs reported Top1 Error for  ResNet, ResNet-1K, and DenseNet models}
+    \caption{$\alpha$-Norm versus reported Top1 Error for  ResNet, ResNet-1K, and DenseNet models.}
     \label{fig:cv2-accuracy}
 \end{figure}
-
-TODO:  add table of metrics for these 3 models, show that alpha-Norm is best again!
 
@@
    &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
- Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+ Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
 \hline
@@
 \end{center}
-\caption{RMSE for Linear Fits of Metric to Reported Top1 Test Error for all pre-trained models in the architecture series.
-All models trained on ImageNet except ResNet-1K, which was trained on ImageNet-1K.
-}
-\label{table:models}
+\caption{RMSE for linear fits (smaller is better) of quality metrics to Reported Top1 Test Error (\cite{XXX-GIVE-REFS-AGAIN}\michael{Give refs again.}) for all pretrained models in the architecture series.
+VGG, ResNet, and DenseNet were pretrained on ImageNet, and ResNet-1K was pretrained on ImageNet-1K.
+\michael{Do we want Frob norm squared.} }
+\label{table:cv-models}
 \end{table}
 
-DISCUSS TABLE
+\paragraph{Correlation Flow in CV Models.}
 
-\paragraph{Correlation Flow in CV Models}
-
-Explain WHAT IS \emph{Correlation Flow}
-
-Compare VGG, ResNet, and DenseNet in the context of how many connections that have
+We can learn much more about a pretrained model by going beyond average values to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).
+% in the network.
+The most interesting results are seen when we plot the PL exponent, $\alpha$, as a function of depth.
+%
+See Figure~\ref{fig:vgg-alpha-layers}, which plots $\alpha$ for each layer (the first corresponds to data, the last to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (without BatchNormalization), ResNet-1K, and DenseNet~series.
 
@@
     }
-    \caption{Power law exponent $\alpha$ vs layer for VGG, ResNet, and DenseNet models}
+    \caption{PL exponent $\alpha$ versus layer for VGG, ResNet-1K, and DenseNet models.
+Y axes on each plot are different.
+ResNet-1K exhibits very different and much more stable behavior, which we interpret in the text as a \emph{correlation flow} across the layers.
+}
     \label{fig:vgg-alpha-layers}
 \end{figure}
+
+In the VGG models, $\alpha$ systematically increases as we move down the network, from data to labels, in the Conv2D layers, starting with $\alpha\lesssim 2.0$ and reaching all the way to $\alpha\sim 5.0$; and then, in the last three, large, fully-connected (FC) layers, $\alpha$ stabilizes back down to $\alpha\in[2,2.5]$.
+This is seen for all the VGG models (again, only the shallowest and deepest are shown in this figure), indicating that the main effect of increasing depth is to increase the range over which $\alpha$ can increase, thus leasing to larger $\alpha$ values in later Conv2D layers of the VGG models.
+This is quite different than the behavior of either the ResNet-1K models or the DenseNet models.
+
+For the ResNet-1K models, $\alpha$ also increases in the last few layers, but as the models get deeper, there is a wide range over which $\alpha$ values tend to remain small.
+\michael{Can we say something about exceptions which are larger than for VGG or DenseNet.}
+This is seen for other models in the ResNet-1K series, but it is most pronounced for the larger ResNet-1K (152) model, where $\alpha$ remains relatively stable at $\alpha\sim 2.0$, from the earliest layers all the way until we reach close to the final layers.
+
+For the DenseNet models, $\alpha$ tends to increase as the layer id increases, in particular for layers toward the end.
+This is similar to what is seen in the VGG models, but with the DenseNet models, the variance is much larger (in particular for the earlier and middle layers, where it can range all the way to $\alpha\lesssim 8.0$) and much less systematic.
+
+We can understand Figure~\ref{fig:vgg-alpha-layers} in terms of the \emph{flow of correlation}, as follows.
+Informally, one would expect a DNN model to perform well when it fascilitates the propagation of information/features across layers.
+In the absence of training/test data, we can to try to quantify this by measuring the PL properties of weight matrices.
+Smaller $\alpha$ values correspond to layers in which correlations across multiple scales are better captured~\cite{MM18_TR,SornetteBook}, and thus we expect that values of $\alpha$ that are stable across multiple layers enable better \emph{correlation flow} through the network.
+
+Consider the differences---in particular, the number of residual connections---between the VGG, ResNet, and DenseNet architectures.
+VGG, while a good model in many ways, was limited by needing a massive FC layers at the end of the model, requiring significant memory and computational resources, and it's poor conditioning led to problems with vanishing gradients.
+The empirical manifestations of this (on weight matrices) are that fitted $\alpha$ values get much worse for deeper layers.
+ResNet greatly improved on VGG by introducing residual connections, allowing for greater accurary with far fewer parameters.
+(Indeed, ResNet models of up to 1000 layers have been trained.) 
+We conjuecture that the effeciency and effectivness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and strongly optimized.
+This should also be contrasted with the DenseNet model, which contains many connections between every layer.
+Our results (large $\alpha$, meaning they even a HT model is probably a poor fit) suggest that DenseNet has too many connections, diluting the correlation flow across layers, and leaving many layers very poorly optimized.
+
+\paragraph{Behavior on Less Thoroughy Studied CV Models.}
+
+Figure~\ref{fig:vgg-metrics} and Table~\ref{table:cv-models} show that---for well-trained and widely-studied models---the weighted $\hat{\alpha}$ metric performs better than but similar to norm-based metrics.
+This suggests an obvious question: is the $\hat{\alpha}$ metric just a variation of these more familiar norm-based metrics, or does it capture something different?  
+To show that it is not just a variation and that it does capture something different, consider the following less thoroughy studied model, a compressed DNN model~\cite{CWZZ17_TR}, where we observed that the average (Frobenius or Spectral) norm increases with decreasing test error, whereas the average $\alpha$ decreases, as expected.
+
+\begin{figure}[t]
+   \centering
+   \subfigure[$\lambda_{max}$ for ResNet20 layers]{
+      %\includegraphics[scale=0.14]{img/resnet4d_maxev.png}
+      %\includegraphics[width=3.7cm]{img/resnet4d_maxev.png}
+      \includegraphics[width=4.5cm]{img/resnet4d_maxev.png}
+      \label{fig:resnet204Dalpha}
+   }
+   \qquad
+   \subfigure[$\alpha$ for ResNet20 layers]{
+      %\includegraphics[scale=0.14]{img/resnet4d_alphas.png}
+      %\includegraphics[width=3.7cm]{img/resnet4d_alphas.png}
+      \includegraphics[width=4.5cm]{img/resnet4d_alphas.png}
+      \label{fig:resnet204Dmaxev}
+   }
+   \caption{
+     Analysis of ResNet20, distilled with Group Regularization, as implemented in the \texttt{distiller} (4D\_regularized\_5Lremoved) pre-trained models.  
+     Comparison of individual layer $\mathbf{W}_{l}$ maximum eigenvalues ($\lambda_{max}$, or Spectral Norms) and  
+     PL exponent $\alpha$, between baseline (green) and fine-tuned (red)  pre-trained models.   
+           }
+   \label{fig:resnet204D5L}
+\end{figure}
+
+Consider average metrics measured on ResNet20, trained on CIFAR10, before and after applying the Group Regularization technique, as implemented in the \texttt{distiller} package.%
+\footnote{For details, see \url{https://nervanasystems.github.io/distiller/\#distiller-documentation} and also \url{https://github.com/NervanaSystems/distiller}.}
+We analyze the available pre-trained 4D\_regularized\_5Lremoved baseline and fine-tuned models.  %\cite{distiller_repo}.
+Figure~\ref{fig:resnet204D5L} presents the Spectral Norm ($\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}_{l}$.%
+\footnote{We only include layer matrices or feature maps with $M\ge50$.}
+The reported baseline test accuracies ($Top1=91.450$ and $Top5=99.750$) are better than the reported fine-tuned test accuracies ($Top1=91.020$ and $Top5=99.670$).
+Thus, the results of Figure~\ref{fig:vgg-metrics} and Figure~\ref{fig:cv2-accuracy} might suggest that the baseline Spectral (and Frobenius) Norm should be \emph{smaller} than those of the layers in the fine-tuned model.
+In both cases (Frobenius norm results not shown), we observe the opposite.
+On the other hand, the $\alpha$ values do not systematically differ between the baseline and fine-tuned models.
+Also (not shown), the average (unweighted) baseline $\alpha$ is smaller than the fine-tuned average (as would be predicted by HT-SR Theory, on which $\hat{\alpha}$ is based).
+
+The reason for this is that the \texttt{distiller} Group Regularization technique has the unusual effect of increasing the norms of the $\mathbf{W}$ feature maps for at least two of the Conv2D layers.
+\michael{This impedance shuffling or whatever it is called messes up norms, but the strong correlations are still preserved, as seen in the $\hat{\alpha}$ metric, and the model quality is good.
+CLARIFY.}

### 2020-01-30 introduction.tex
added methods section
mwm initial pass

@@
 they typically do not correlate with other properties of interest such as robustness or fairness or interpretability, 
-%XXX SOMETHING ELSE, 
 and so on.
 A somewhat less well-known problem, but one that is increasingly important (in particular in industrial-scale ML---where the \emph{users} of models are not the \emph{developers} of the models) is that one may access to neither the training data nor the testing data.
-Instead, one may simply be given a model that has already been trained---we will call such an already-trained model a \emph{pre-trained model}---and be told to use it.
+Instead, one may simply be given a model that has already been trained---we will call such an already-trained model a \emph{pretrained model}---and be told to use it.
 
 Na\"{\i}vely---but in our experience commonly, among both ML practitioners and ML theorists---if one does not have access to training or testing data, then one can say absolutely nothing about the quality of a ML model.
-This may be true in worst-case theory, but ML models are used in practice, and there is a need for theory to guide that practice.
-Moreover, if ML is to become an industrial process, then the process will become siloed: some groups will gather data, other groups will develop models, and still other groups will use those models.
-The users of models can not be expected to know the precise details of how the models were built, the specifics of the data that were used to train the model, what the loss of values of the hyperparameters were, how precisely the model was regularized, etc.
+This may be true in worst-case theory, but ML models are used in practice, and there is a need for a \emph{practical theory} to guide that practice.
+Moreover, if ML is to become an industrial process, then that process will become siloed: some groups will gather data, other groups will develop models, and still other groups will use those models.
+The users of models can not be expected to know the precise details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
 %
-Having metrics to evaluate the quality of a ML model in the absence of training and testing data and without any detailed knowledge of the training and testing process---indeed, having theory for pre-trained models, to predict how, when, and why such models can be expected to perform well or poorly---is clearly of interest.
+Having metrics to evaluate the quality of a ML model in the absence of training and testing data and without any detailed knowledge of the training and testing process---indeed, having theory for pretrained models, to predict how, when, and why such models can be expected to perform well or poorly---is clearly of interest.
 
-In this paper, we present and apply techniques to evaluate the generalization properties of large-scale state-of-the-art pre-trained neural network (NN) models.%
+In this paper, we present and evaluate quality metrics for pretrained neural network (NN) and deep neural network (DNN) models at scale.%
 \footnote{We reiterate: One could use these techniques to improve training, and we have been asked about that, but we are not interested in that here.
-Our main goal here is to use these techniques to evaluate properties of state-of-the-art pre-trained NN models.}
-To do so, we consider a large suite of publicly-available models from computer vision (CV) and natural language processing (NLP).
+Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained NN models.}
+To do so, we consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
 %
 By now, there are many such state-of-the-art models that are publicly-available, e.g., 
-there are now hundreds of pre-trained models in CV ($\ge 500$) and NLP ($\approx 100$).%
+there are now hundreds of pretrained models in CV ($\ge 500$) and NLP ($\approx 100$).%
 \footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
@@
 \footnote{Clearly, there is a selection bias or survivorship bias here---people tend not to make publicly-available their poorly-performing models---but these models are things in the world that (like social networks or the internet) can be analyzed for their properties.}
-XXX.  MORE DETAILS.
+\michael{XXX.  MORE DETAILS.}
 Importantly, for all of these models, we have no access to training data or testing data.
 
-XXX.  LIST PLACES WHERE THEY ARE AVAIALBLE, HERE OR IN NEXT SECTION.
+\michael{XXX.  LIST PLACES WHERE THEY ARE AVAIALBLE, HERE OR IN NEXT SECTION.}
+
+XXX.  
+To do this, we will compute a variety of \emph{quality metrics} based on the spectral properties of the layer weight matrics.
+Note that unlike traditional ML approaches, however, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), and \emph{we do not aim to evaluate a single model} (e.g., by training with differing hyperparmeters).
+Instead, we want to examine different models a common architecture series, and we want to compare models between different architectures themselves, and in both cases \emph{we aim to predict trends in the quality of pre-trained models without access to training or testing data}.
 
@@
 \item XXX TECHNICAL THING 3
+\item XXX TECHNICAL THING 4
 \end{itemize}
-Perhaps our most improtant contribution, however, is just doing things different, etc. ...
+
+\paragraph{Organization of this paper.}
+
+We start in Section~\ref{sxn:background} and Section~\ref{sxn:methods} with background and an overview of our general approach.
+In Section \ref{sxn:cv}, we study three well-known widely-available DNN CV architectures (the VGG, ResNet, and DenseNet series of models); and we provide an illustration of our basic methodology, both to evaluate the different metrics against reported test accuracies and to use quality metrics to understand model properties.
+Then, in Section~\ref{sxn:nlp}, we look at several variations of a popular NLP DNN architecture (the OpenAI GPT and GPT2 models); and we show how model quality and properties vary between several variants of GPT and GPT2, including how metrics behave similarly and differently.
+Then, in Section \ref{sxn:all_cv_models}, we present results based on an analysis of hundreds of pretraind DNN CV models, showing how well each metric is correlated with the reported test accuracies, and how the Alpha-Norm metric(s) perform remarkably well.
+Finally, in Section~\ref{sxn:conc}, we provide a brief discussion and conclusion.

### 2020-01-30 methods.tex
mwm initial pass

--- 2020-01-20 methods.tex
@@
 
-We assume we are given several pretrained Deep Neural Networks (DNNs), as part of a similar architecture.
-We would like to estimate the trends in the reported test / generalization accuracy accross a series of similar archtectures.  
-and without Batch Normalization, trained on ImageNet, and widely available in the pyTorch distribution.
-
-\paragraph{Empirical Complexity Metrics}
-To do this, we will compute a variety of \emph{Complexity Metrics} based on the Product Norm of the layer weight matrics.
-Note that unlike traditional ML approaches, however, we do not seek a bound on the complexity (i.e. test error), 
-nor are we trying to evaluating a single model with differing hyperparmeters.  We wish to examine different models a 
-common architecture series. And, also, compare different architectures themselves.
-
-Let us write the Energy Landscape (or optimization function) for a typical DNN with $L$ layers as
+Let us write the Energy Landscape (or optimization function, parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a typical DNN with $L$ layers, with activation functions $h_{l}(\cdot)$, and with $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$, as:
 \begin{equation}
@@
 \end{equation}
+We assume we are given several pretrained Deep Neural Networks (DNNs), e.g., as part of an architecture series.
+The models have been trained on (unspecified and unavailable) labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using Backprop, by minimizing some (also unspecified and unavailable) loss function $\mathcal{L}()$.
+We expect that most well-trained, production-quality models will employ one or more forms of on regularization, such as Batch Normalization, Dropout, etc., and many will also contain additional structure such as Skip Connections, etc.
+Here, we will ignore these details, and will focus only on the weight matrices.
 
-with activation functions $h_{l}(\cdot)$,  weight matrices $\mathbf{W}_{l}$, and the biases $\mathbf{b}_{l}$.
+\paragraph{DNN Quality Metrics.}
 
-The model has been (or will be) trained on (unspecified) labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, 
-using Backprop, by minimizing some (also unspecified) loss function $\mathcal{L}()$.
-Moreover, we expect that most well trained,.
-production quality models will employ 1 or more forms of on regularization, such as Batch Normalization, Dropout, etc, and will also contain additional structure such as Skip Connections etc.
-Here, we ignore these details, and focus only on the weight matrices.
-
-Each layer contains by one or more layer 2D weight matrices $\mathbf{W}_{L}$, and/or the 2D feature maps $\mathbf{W}_{i,L}$ extracted from 2D Convolutional layers.
-(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below.) We assume the layer weight matrices are statistically independent, allowing us to estimate the Complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm, which resembles a data dependent VC complexity
-
+Each DNN layer contains one or more layer 2D weight matrices $\mathbf{W}_{L}$, and/or 2D feature maps $\mathbf{W}_{i,L}$ extracted from 2D Convolutional layers. 
+(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below.) 
+%
+\michael{
+XXX.
+PROBABLY WE DO NOT WANT THE FOLLOWING.
+INSTEAD, JUST PRESENT METRICS AND POINT TO PREVIOUS PAPERS FOR JUSTIFICATION.
+E.G., WE DONT WANT TO JUSTIFY THE IMPLICIT STATISTICAL INDEPENDENCE ASSUMPTION HERE.
+IF OKAY, JUST DELETE THIS RED BLOCK.
+We assume the layer weight matrices are statistically independent, allowing us to estimate the Complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm, which resembles a data dependent VC complexity
 \begin{equation}
@@
 \end{eqnarray*}
+}
+%
+We have examined a large number of possible quality metrics.
+The best performing metrics (recall that we can only consider metrics that do not use training/test data) depend on the norm and/or spectral properties of weight matrices, $\mathbf{W}$.%
+\footnote{We do not use intra-layer information from the models in our quality metrics, but as we will describe our metrics can be used to learn about }
+We consider the following metrics.
+\begin{itemize}
+\item 
+Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{W}\Vert^{2}_{2}=\sum_{i,j}w^{2}_{i,j} = \sum_{i=1}^{M} \lambda_{i}^{2}$
+\michael{Do we use norm or log norm, here and with other norms.  Worth being very explicit and very consistent.}
+\item 
+Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
+\michael{We should probably call the spectral norm $\Vert\mathbf{W}\Vert_{2}$, or change Frob norm notation and have an explicit remork.  Let's decide.}
+\item 
+Weighted Alpha Metric: $\hat{\alpha}=\alpha\log\lambda_{max}$
+\item 
+$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda_{i}^{\alpha}$,
+\michael{Be clear about how we average this across layers.}
+\end{itemize}
+The first two metrics are well-known in ML.
+The last two deserve special mention.
+For all these metrics, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the \emph{Empirical Correlation Matrix},
+$ %% $$
+\mathbf{X}=\mathbf{W}^{T}\mathbf{W} ,
+$ %% $$
+and so $\lambda_{max}$ is the maximum eigenvalue of $\mathbf{X}$. 
+(These eigenvalues are the squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, i.e., $\lambda_{i}=\sigma^{2}_{i}$.)
+For the last two metrics, the exponent $\alpha$ is the power law exponent that arises in the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+Operationally, $\alpha$ is determined by using the publicly-available \emph{WeightWatcher} tool (\cite{weightwatcher_package}) to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated power law
+\begin{equation}
+\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}  .
+\end{equation}
+\michael{We need to be clear that this is a truncated power law fit and that $\lambda_{max}$ comes from that and not the largest empirical eigenvalue.}
+The Weighted Alpha Metric was introduced previously~\cite{MM20_SDM}, where (on a much smaller set of data than we consider here) it was shown to correlate well with the trends in reported test accuracies of pretrained DNNs.
+Based on this, here we introduce and evaluate the $\alpha$-Norm metric.
+One would expect $\hat{\alpha}$ to approximate the log $\alpha$-Norm very well for $\alpha < 2$ and reasoably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.
 
-Here, we will consider the following Norms:
+\michael{Need to say that such large $\alpha$ values don't mean much.}
 
-\begin{itemize}
- \item Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{W}\Vert^{2}_{2}=\sum_{i,j}w^{2}_{i,j}$
- \item Spectral Norm:  $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
- \item $\alpha-$Norm (or Shatten Norm) $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda^{\alpha}$,
-\end{itemize}
+\michael{Need to highlight difference between $\alpha$ and $\hat{\alpha}$.}
 
-where $\lambda_{i}$ is the $i-th$ eigenvalue of the \emph{Empirical Correlation Matrix}
+\paragraph{Spectral Analysis of Convolutional 2D Layers.}
 
-\begin{equation}
-\mathbf{X}=\mathbf{W}^{T}\mathbf{W}
-\end{equation}
+There is some ambiguity in performing spectral analysis on Convolutional 2D (Conv2D) layers.  
+A Conv2D layer can be represented as a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ filter (or kernel) and $in\_ch$ / $out\_ch$ input / output channels, respectively (usually $in\_ch\le out\_ch$).
+Typically, $w=h=k$,  giving $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\_ch\times out\_ch)$ each. 
+%
+There are at least three different approaches that have been advocated for applying the Singular Values Decomposition (SVD) to an Conv2D layer:
+run an SVD on each of the pre-Activation Maps $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values; 
+stack the feature maps into a single rectangular matrix of, say, dimension $((k\times k\times out\_ch)\times in\_ch)$, yielding $in\_ch$ singular values;
+compute the 2D Fourier Transform (FFT) for each of the $(in\_ch, out\_ch)$ pairs, and run SVD on the resulting Fourier coeffients~\cite{Long2019}, leading to $\sim(k\times in\_ch\times out\_ch)$ non-zero singular values.
+Each method has tradeoffs.  
+In principle, the third method is mathematically sound, but it is computationally expensive. 
+For our analysis, because we are performing tens of thousands of calculations, we select the first method, which is numerically the fastest and is easiest to reproduce.%
+\footnote{We provide a Google Colab notebook where all results can be reproduced, with the option to redo the calculations with the third option for the SVD of the Conv2D.}
 
-and $\lambda_{max}$ is the maximum eigenvalue. These eigenvalues are the square of the singular values $\sigma_{i}$ of $\mathbf{W}$ :  $\lambda_{i}=\sigma^{2}_{i}$.
-
-The exponent $\alpha$ is the power law exponent that arises in our \emph{Theory of Heavy Tailed Self Regularization}, and is the determined by fitting the Empirical Spectral Density (ESD) of $\mathbf{X}$--i.e.
-a histogram of the eigenvalues--$\rho(\lambda)$ to a truncated power law
-
-\begin{equation}
-\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}
-\end{equation}
-
-We will also consider an approximate capacity metric, $\hat{\alpha}$, shown previously to correlate well with the trends in reported test accuracies of pretrained DNNs \cite{MM}
-
-\begin{itemize}
- \item $\hat{\alpha}=\alpha\log\lambda_{max}\approx\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$
-\end{itemize}
-
-which approximates the log $\alpha-$Norm for both Very Heavy Tailed weight matrices ($\alpha < 2)$  and reasoably well for finite size, Moderately Heavy Tailed ones $\alpha\in[2,5]$.
-
-\paragraph{Spectral Analysis of Convolutional 2D Layers}
-While we can easily analyze Linear layers, there is some ambiguity in performing spectral analysis on Convolutional 2D (Conv2D) layers.
-A Conv2D layer is a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ filter (or kernel), and $in\_ch,out\_ch$ input and output channels, respectively.
-Typically, the $w=h=k$,  giving $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\_ch\times out\_ch)$ each.
-Usually $in\_ch\le out\_ch$.
-
-There are at least 3 different approaches to computing the Singular Values Decomposition(s) (SVD) of an Conv2D layer
-\begin{enumerate}
-\item run SVD on each of the pre-Activation Maps $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values. 
-\item stack the feature maps into a single rectangular matrix of, say, dimension $((k\times k\times out\_ch)\times in\_ch)$, yielding $in\_ch$ singular values
-\item Compute the 2D Fourier Transform (FFT) for each of the $(in\_ch, out\_ch)$ pairs, and run SVD on the resulting Fourier coeffients\cite{Long2019}.
-This leads to $\sim(k\times in\_ch\times out\_ch)$ non-zero singular values.
-\end{enumerate}
-
-Each method has tradeoffs.  While, in principle, 3. is mathematically sound, it is computationally expensive. 
-For this study, because we are computing tens of thousands of calculations, we select $1.$, which is 
-numerically the fastest and easiest to reproduce.\footnote{We will provide a Google Colab notebook where all results can be reproduced, with the option to redo the calculations with option 3 for the SVD of the Conv2D.}
-
-To verify that our approach is meaningful, we need to see the ESD is neither due to a random matrix, nor due to unsually large matrix elements, but, in fact, captures correlations learned from the data.
-We examine typical Conv2D layer for the pretained AlexNet model (distributed with pyTorch).
-Figure \ref{fig:alexnet1} displays the ESD for the first slice (or matrix $\mathbf{W}$) of the third Conv2D layer, extracted from a 4-index Tensor of shape $(384, 192, 3, 3)$.
-The red line displays the best fit to a random matrix, using the Marchenko pastur theory\cite{MM}.
+\paragraph{XXX.}
+XXX.  THIS VERIFICATION IS NOT ABOUT CONV LAYERS, IT IS ABOUT PL MORE GENERALLY, CORRECT?  WE SHOULD CLARIFY AND SQUISH.
+To verify that our approach is meaningful, we need to confirm that the ESD is neither due to a random matrix, nor due to unsually large matrix elements, but, in fact, captures correlations learned from the data.
+We examine typical layer for the pretained AlexNet model (distributed with pyTorch). 
+Figure~\ref{fig:alexnet1} displays the ESD for the first slice (or matrix $\mathbf{W}$) of the third Conv2D layer, extracted from a 4-index Tensor of shape $(384, 192, 3, 3)$.
+The red line displays the best fit to a random matrix, using the Marchenko pastur theory~\cite{MM}.
 We can see the random matrix model does not describe the ESD very well.
@@
 In fact, the empircal ESD is better modeled with a truncated power law distribtion.
-
-Although the ESD is \emph{Heavy Tailed}, this does not imply that the orginal matrix $\mathbf{W}$ is itself heavy tailed--
-only the correlation matrix $\mathbf{X}$ is. If $\mathbf{W}$ was, 
-then it would contain 1 or more unusually large matrix elements, and they would dominate the ESD.  
-Of course the randomized $\mathbf{W}$ would also be heavy tailed, but its ESD neither resembles the original
-nor is it heavy tailed.  So we can rule out $\mathbf{W}$ being heavy tailed.
+\michael{We may want to give a one-sentence summary of this par and fig at the end of the previous par.}
 
@@
    \subfigure[Actual ESD]{
-     \includegraphics[scale=0.5]{img/alexnet1.png} 
+     %\includegraphics[scale=0.5]{img/alexnet1.png} 
+     \includegraphics[scale=0.25]{img/alexnet1.png} 
      \label{fig:alexnet1}
@@
    \subfigure[ESD of randomly shuffled matrix]{
-      \includegraphics[scale=0.5]{img/alexnet2.png}
+      %\includegraphics[scale=0.5]{img/alexnet2.png}
+      \includegraphics[scale=0.25]{img/alexnet2.png}
       \label{fig:alexnet2}
    }
-   \caption{ESD of AlexNet Conv2D pre-Activation map for Layer 3 Slice 1, actual and randomized}
+   \caption{ESD of AlexNet Conv2D pre-Activation map for Layer 3 Slice 1, actual and randomized.
+            \michael{Need better figs here.}
+           }
    \label{fig:alexnet}
 \end{figure}
+
+Although the ESD is \emph{Heavy Tailed}, this does not imply that the orginal matrix $\mathbf{W}$ is itself heavy tailed--only the correlation matrix $\mathbf{X}$ is. 
+If $\mathbf{W}$ was, then it would contain 1 or more unusually large matrix elements, and they would dominate the ESD.  
+Of course the randomized $\mathbf{W}$ would also be heavy tailed, but its ESD neither resembles the original nor is it heavy tailed. 
+So we can rule out $\mathbf{W}$ being heavy tailed.
+\michael{These comments seem out of place, since they hold more generally than for the Conv2D layers.}
 
@@
 the smaller the exponent $\alpha$, the more correlation in the weight matrix. 
+\michael{These comments seem out of place, since they hold more generally than for the Conv2D layers.}
 
-\paragraph{Normalization of Empirical Matrices}.  
-The normalization does not affect the Power Law fits because the Heavy Tailed exponent $\alpha$ is scale-invariant.
-As are metrics like the Stable Rank and our MP Soft Rank. In this study, however, all of the metrics strongly depend 
-on the scale of the weight matrix.
-Formally, to apply RMT, we would usually define Correlation Matrix with $\dfrac{1}{N}$ normalization,
-and assume that either the variance of $\mathbf{X}$ is unity, $\sigma^{2}=1$, or is known.
-\footnote{And for formal proofs of the Heavy Tailed results, we need a different normalization,$\dfrac{1}{N^{1-\alpha}}$}
-But for these empirical studies, the pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$ ,
-with the variance already normalized to $\sqrt{\frac{1}{N}}$, or some variant of this such as Glorot (or Xavier) Normalization\cite{GloRot},
-or as $\sqrt{\frac{2}{N*k^2}}$ for Convolutional 2D Layers. 
-We do not normalize (or renormalize) the Empirical Correlation Matrices, and use them as-is,
-except that we rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $\frac{k}{\sqrt{2}}$ so that they
-are on the same scale as the Linear layers.
+\paragraph{Normalization of Empirical Matrices.}  
+Normalization is an important, if underappreciated, practical issue.
+Importantly, the normalization of weight matrices does \emph{not} affect the Power Law fits because the Heavy Tailed exponent $\alpha$ (as well as other metrics such as the Stable Rank and MP Soft Rank~\cite{MM18_TR,MM19_HTSR_ICML}) is scale-invariant.
+Norm-based metrics, however, do depend strongly on the scale of the weight matrix.
+Typically, to apply RMT, we would usually define Correlation Matrix with $1/N$ normalization and assume that the variance of $\mathbf{X}$ is either unity or a known constant.% 
+\footnote{For formal proofs of Heavy Tailed results, one typically needs a different normalization such as $1/N^{1-\alpha}$.}
+Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with the variance already normalized to $\sqrt{1/N}$, or some variant of this, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.
+We do not have conrol over the final empirical normalization of these models; and we do \emph{not} normalize (or renormalize) the Empirical Correlation Matrices, i.e., we use them as-is.
+The only exception to this is that we do rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $k/\sqrt{2}$ so that they are on the same scale as the Linear layers.
 
-Note that in some pretrained models, like the pyTorch VGG models, the Linear weight matrices appear to have been 
-initialized with variance $\sigma\sim0.01$.  We do not attempt to correct for this, and, instead, simply try to treat all
-models the same.
+\michael{MOVE TO LATER: COMMENT ON HOW LOG NORM first and last layers behave, maybe somewhere else.}
 
-COMMENT ON HOW LOG NORM first and last layers behave, maybe somewhere else
+\michael{CMOVE TO LATER: OMMENT ON HOW LOG PORM  for GPT includes unusually high alpha, not meaningful other than to show the trend.}
 
-COMMENT ON HOW LOG PORM  for GPT includes unusually high alpha, not meaningful other than to show the trend.
+\paragraph{The WeightWatcher Tool.}
 
-\paragraph{The WeightWatcher Tool}
+We compute the metrics using the WeightWatcher tool (version 0.2.7), and we provide Jupyter notebooks in the github repo for this paper~\cite{repo}, making the results fully reproducible.
 
-\paragraph{Organization of this paper}
-
-We want to study the effective of Product Norm Metrics in describing trends in the
-generalization accuracyof real-word, production quality, pretrained Deep Neural Networks.
-In Section \ref{sxn:cv}, we study three well known, and widely available DNN CV architectures:
-the VGG series, the ResNet series, and the DenseNet series of models.  We lay out the
-basic methodology we use to evaluate the different metrics against the reported test accuracies. 
-We also introduce the notion of Correlation Flow, compare the three pretrained architectures in detail.  
-In Section  \ref{sxn:nlp}, we then look at several variations of a popular NLP DNN architecture,
-the OpenAI GPT and GPT2 models.  We discuss when and how the different metrics apply more generally.
-Finally, in Section \ref{sxn:all}, we examine over 100 pretraind DNN CV models, and show
-how well each metric is correlated with the reported test accuracies across different models
-in an architeture series.  
-----
+\michael{XXX.  FEW SENTENCES ABOUT REPRODUCIBILITY MORE GENERALLY HERE.}

### 2020-01-30 nlp_models.tex
mwm initial pass

@@
 \label{sxn:nlp}
+
+In this section, we examine the quality metrics described in Section~\ref{sxn:methods} for several NLP model architectures.
+
+apply more generally, where they differ, e.g., how the layer Spectral Norm behaves unexpectedly, while the $\alpha$-based metrics can give insight into how well trained a model is.
 
@@
 
-In the past two years, nearly 100 open source, pre-trained DNNs for Natural Language Processing (NLP) have emerged,
-based on the revolutionary Transformer architecture, including variants of BERT, Transformer-XML, GPT, etc.
-The Transformer architectures consist of blocks of Attention layers, containing of 2 large, Feed Forward (Linear)
-weight matrics \cite{Attn2017}. In contrast to the smaller pre-Activation maps arising in Cond2D layers,
-the Attention matrices are significantly larger, and frequently under-correlated, with generally larger
-Heavy Tailed Power Law exponents $\alpha$ (when fitting the ESD).  Here, we breifly look at a few
-popular pretrained NLP DNNs to demonstrate that the Theory of Heavy Tailed Self-Regularization also
+Within the past two years, nearly 100 open source, pre-trained DNNs for NLP have emerged, based on the revolutionary Transformer architecture.
+This includes variants of BERT, Transformer-XML, GPT, etc.
+The Transformer architectures consist of blocks of Attention layers, containing two large, Feed Forward (Linear) weight matrics~\cite{Attn2017}. 
+In contrast to the smaller pre-Activation maps arising in Cond2D layers, Attention matrices are significantly larger, 
+and frequently under-correlated, with generally larger Heavy Tailed Power Law exponents $\alpha$ (when fitting the ESD).  
+Here, we breifly look at a few popular pretrained NLP DNNs to demonstrate that the Theory of Heavy Tailed Self-Regularization also
 applies to NLP models and not just CV models, and to highlight how to use the theory to identify
@@
 
-\paragraph{OpenAI GPT Models}
+\paragraph{OpenAI GPT Models.}
+
 We use the \emph{WeightWatcher} tool to analyze the OpenAI GPT and GPT2 models, which gives us
@@
 
-\paragraph{Empirical Quality Metrics for GPT and GPT2}
+\paragraph{Empirical Quality Metrics for GPT and GPT2.}
+
 We first compare the distribution of  Heavy Tailed Power Law exponents $\alpha$ in GPT and GPT2.
@@
 
-\paragraph{The Correlation Flow in GPT and GPT2} also differs significantly between GPT and GPT2.
+\paragraph{Correlation Flow in GPT and GPT2.}
+
+also differs significantly between GPT and GPT2.
 Figure \ref{fig:gpt-alpha-layer} plots $\alpha$ vs the depth (i.e. a layer id) for each model.
@@
 overparameterized and/or not yet fully optimized and require datasets even larger
-than the recent XL $1.5B$ release\cite{gpt2-xl}.
+than the recent XL $1.5B$ release~\cite{gpt2-xl}.
 

### 2020-01-30 introduction.tex
small update

@@
 A somewhat less well-known problem, but one that is increasingly important (in particular in industrial-scale ML---where the \emph{users} of models are not the \emph{developers} of the models) is that one may access to neither the training data nor the testing data.
-Instead, one may simply be given a model that has already been trained---we will call such an already-trained model a \emph{pretrained model}---and be told to use it.
+Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and be told to use it.
 
@@
 %
-Having metrics to evaluate the quality of a ML model in the absence of training and testing data and without any detailed knowledge of the training and testing process---indeed, having theory for pretrained models, to predict how, when, and why such models can be expected to perform well or poorly---is clearly of interest.
+\charles{Moreover, industry faces unique practical problems such as 'do we have enough data for this model?' It is great practical interest to have metrics to evaluate the quality of a ML model in the absence of training and testing data and without any detailed knowledge of the training and testing process.
+We seek a theory for pretrained models which can predict how, when, and why such models can be expected to perform well or poorly.}
 
@@
 To do this, we will compute a variety of \emph{quality metrics} based on the spectral properties of the layer weight matrics.
-Note that unlike traditional ML approaches, however, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), and \emph{we do not aim to evaluate a single model} (e.g., by training with differing hyperparmeters).
-Instead, we want to examine different models a common architecture series, and we want to compare models between different architectures themselves, and in both cases \emph{we aim to predict trends in the quality of pre-trained models without access to training or testing data}.
+Note that unlike traditional ML approaches, however, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., \ncyan{hyperparmeter optimization}).
+Instead, we want to examine different models \ncyan{across} common architecture series, and we want to compare models between different architectures themselves, and in both cases \emph{we aim to predict trends in the quality of pre-trained models without access to training or testing data}.
 

### 2020-01-31 background.tex
edit text

@@
 \cite{MM20_SDM},
-\cite{MM20_unpub_work}.
+\cite{MM20_unpub_work}.C
 Cite also Ganguli review and maybe other stuff.
+\charles{Gangulis work is not really close to what we are doing, although I get the politics.
+The closest useful work would be the older work by Bouchaud and Potters, which discusses the statisical mechanics of heavy tailed and strongly correlated systems form the 90s}
 
@@
 (that is applicable to tiny things but has no hope of scaling up).
+\charles{it is the oppposite...ab initio theory scales quite well...it is the spherical cow models of physics and ML that do not scale.
+What we have is a semi-empirical theory.
+We use real theory, but require empirical input, at least in the new stat mech work.
+What we have introduced is a phenomenology, which to me is different from a semi-empirical or phenomenological theory  }
 \footnote{In most areas where there are complex highly-engineered systems (beyond complex AI/ML systems), one used phenomenological theory rather than first principles theory.
@@
 \end{itemize}
-Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text, when developing self driving car systems, and when distilling a reliable model in some way to obtain comparable training/test quality but that damages the model in some other subtle way.
-\michael{That last example is awkward.  It would be good to have a better example and plant seeds for model distillation elsewhere.}
+Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text, when developing self driving cars, or creating realistic chatbots.\footnote{i.e.
+current chatbots use perplexity as a proxy for passing a Turing test}
+%, and when distilling a reliable model in some way to obtain comparable training/test quality but that damages the model in some other subtle way. 
+%\michael{That last example is awkward.  It would be good to have a better example and plant seeds for model distillation elsewhere.}

### 2020-01-31 cv_models.tex
more edits

@@
 \michael{Are those numbers stale, I don't see them, and some things seem in the other order.}
+\charles{See Figure 2}
 \michael{We should probably point to the table if we cite quantitative results.}
@@
 \michael{XXX.  AGAIN, WHERE IS RMSE, IN THAT TABLE?}
+\charles{The Table may be stale, see the figures to confirm}
 
@@
 \michael{Can we say something about exceptions which are larger than for VGG or DenseNet.}
+\charles{You have the same plots I do..what do you want to say >}
 This is seen for other models in the ResNet-1K series, but it is most pronounced for the larger ResNet-1K (152) model, where $\alpha$ remains relatively stable at $\alpha\sim 2.0$, from the earliest layers all the way until we reach close to the final layers.

### 2020-01-31 methods.tex
more edits

@@
 %
+\charles{hard to confer on this.
+I tried to state this as simply as possible.
+We have never stated this before, and I think its going to be confusing.
+We at least need to explain why we are looking at the average log norm.
+Otherwise we are forcing the reader to look back, and the paper is hard to read as a standalone entity.
+}
 \michael{
@@
 \michael{We may want to give a one-sentence summary of this par and fig at the end of the previous par.}
+\charles{Here, on the RMT MP stuff, I think it makes sense to point back}
 
@@
 \michael{These comments seem out of place, since they hold more generally than for the Conv2D layers.}
+\charles{Agreed. We could move this up.  We have never really talked about this, but it is essential to explain the difference between assuming W is heavy tailed , which confuses everyone}
 

### 2020-01-31 abstract_file.tex
mostly section 3, but touched every section

@@
 %%Our metrics are drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as Heavy-Tailed Random Matrix Theory (HT-RMT), as used in the recently-developed Theory of Heavy-Tailed Self Regularization (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models).
-The most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
+The most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in a system) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
 Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision and natural language processing.
@@
 On the other hand, for models that may not be well-trained (which, arguably is the point of needing metrics to evaluate the quality of pretrained models), i.e., when we want to distinguish ``good-bad,'' norm-based metrics can qualitatively fail.
-We also find that HT-SR metrics do much better, quantitatively better for at discriminating good-better-best and qualitatively better at discriminating good-bad.
+We also find that HT-SR metrics do much better, quantitatively better at discriminating good-better-best and qualitatively better at discriminating good-bad.
 HT-SR metrics can also be used to characterize fine-scale properties of models, e.g., understanding layer-wise \emph{correlation flow}, and evaluate post-training modifications such as model distillation and model improvement.

### 2020-01-31 appendix.tex
mostly section 3, but touched every section

--- 2020-01-30 methods.tex
@@
-\section{Methods}
-\label{sxn:methods}
+\section{Appendix}
 
-Let us write the Energy Landscape (or optimization function, parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a typical DNN with $L$ layers, with activation functions $h_{l}(\cdot)$, and with $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$, as:
-\begin{equation}
-E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})  .
-\label{eqn:dnn_energy}
-\end{equation}
-We assume we are given several pretrained Deep Neural Networks (DNNs), e.g., as part of an architecture series.
-The models have been trained on (unspecified and unavailable) labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using Backprop, by minimizing some (also unspecified and unavailable) loss function $\mathcal{L}()$.
-We expect that most well-trained, production-quality models will employ one or more forms of on regularization, such as Batch Normalization, Dropout, etc., and many will also contain additional structure such as Skip Connections, etc.
-Here, we will ignore these details, and will focus only on the weight matrices.
+XXX.  APPENDIX.
 
-\paragraph{DNN Quality Metrics.}
-
-Each DNN layer contains one or more layer 2D weight matrices $\mathbf{W}_{L}$, and/or 2D feature maps $\mathbf{W}_{i,L}$ extracted from 2D Convolutional layers. 
-(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below.) 
-%
-\michael{
-XXX.
-PROBABLY WE DO NOT WANT THE FOLLOWING.
-INSTEAD, JUST PRESENT METRICS AND POINT TO PREVIOUS PAPERS FOR JUSTIFICATION.
-E.G., WE DONT WANT TO JUSTIFY THE IMPLICIT STATISTICAL INDEPENDENCE ASSUMPTION HERE.
-IF OKAY, JUST DELETE THIS RED BLOCK.
-We assume the layer weight matrices are statistically independent, allowing us to estimate the Complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm, which resembles a data dependent VC complexity
-\begin{equation}
-\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert\cdots\Vert\mathbf{W}_{L}\Vert ,
-\end{equation}
-where $\mathbf{W}$ is an $(N\times M)$ weight matrix, with $N\ge M$, and 
- $\Vert\mathbf{W}\Vert$ is a matrix norm.   We will actually compute the log Complexity, which takes the form 
-of an Average Log Norm:
-\begin{eqnarray*}
-\log\mathcal{C} &\sim& \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert\cdots\log\Vert\mathbf{W}_{L}\Vert
-\end{eqnarray*}
-}
-%
-We have examined a large number of possible quality metrics.
-The best performing metrics (recall that we can only consider metrics that do not use training/test data) depend on the norm and/or spectral properties of weight matrices, $\mathbf{W}$.%
-\footnote{We do not use intra-layer information from the models in our quality metrics, but as we will describe our metrics can be used to learn about }
-We consider the following metrics.
-\begin{itemize}
-\item 
-Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{W}\Vert^{2}_{2}=\sum_{i,j}w^{2}_{i,j} = \sum_{i=1}^{M} \lambda_{i}^{2}$
-\michael{Do we use norm or log norm, here and with other norms.  Worth being very explicit and very consistent.}
-\item 
-Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
-\michael{We should probably call the spectral norm $\Vert\mathbf{W}\Vert_{2}$, or change Frob norm notation and have an explicit remork.  Let's decide.}
-\item 
-Weighted Alpha Metric: $\hat{\alpha}=\alpha\log\lambda_{max}$
-\item 
-$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda_{i}^{\alpha}$,
-\michael{Be clear about how we average this across layers.}
-\end{itemize}
-The first two metrics are well-known in ML.
-The last two deserve special mention.
-For all these metrics, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the \emph{Empirical Correlation Matrix},
-$ %% $$
-\mathbf{X}=\mathbf{W}^{T}\mathbf{W} ,
-$ %% $$
-and so $\lambda_{max}$ is the maximum eigenvalue of $\mathbf{X}$. 
-(These eigenvalues are the squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, i.e., $\lambda_{i}=\sigma^{2}_{i}$.)
-For the last two metrics, the exponent $\alpha$ is the power law exponent that arises in the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
-Operationally, $\alpha$ is determined by using the publicly-available \emph{WeightWatcher} tool (\cite{weightwatcher_package}) to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated power law
-\begin{equation}
-\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}  .
-\end{equation}
-\michael{We need to be clear that this is a truncated power law fit and that $\lambda_{max}$ comes from that and not the largest empirical eigenvalue.}
-The Weighted Alpha Metric was introduced previously~\cite{MM20_SDM}, where (on a much smaller set of data than we consider here) it was shown to correlate well with the trends in reported test accuracies of pretrained DNNs.
-Based on this, here we introduce and evaluate the $\alpha$-Norm metric.
-One would expect $\hat{\alpha}$ to approximate the log $\alpha$-Norm very well for $\alpha < 2$ and reasoably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.
-
-\michael{Need to say that such large $\alpha$ values don't mean much.}
-
-\michael{Need to highlight difference between $\alpha$ and $\hat{\alpha}$.}
-
-\paragraph{Spectral Analysis of Convolutional 2D Layers.}
-
-There is some ambiguity in performing spectral analysis on Convolutional 2D (Conv2D) layers.  
-A Conv2D layer can be represented as a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ filter (or kernel) and $in\_ch$ / $out\_ch$ input / output channels, respectively (usually $in\_ch\le out\_ch$).
-Typically, $w=h=k$,  giving $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\_ch\times out\_ch)$ each. 
-%
-There are at least three different approaches that have been advocated for applying the Singular Values Decomposition (SVD) to an Conv2D layer:
-run an SVD on each of the pre-Activation Maps $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values; 
-stack the feature maps into a single rectangular matrix of, say, dimension $((k\times k\times out\_ch)\times in\_ch)$, yielding $in\_ch$ singular values;
-compute the 2D Fourier Transform (FFT) for each of the $(in\_ch, out\_ch)$ pairs, and run SVD on the resulting Fourier coeffients~\cite{Long2019}, leading to $\sim(k\times in\_ch\times out\_ch)$ non-zero singular values.
-Each method has tradeoffs.  
-In principle, the third method is mathematically sound, but it is computationally expensive. 
-For our analysis, because we are performing tens of thousands of calculations, we select the first method, which is numerically the fastest and is easiest to reproduce.%
-\footnote{We provide a Google Colab notebook where all results can be reproduced, with the option to redo the calculations with the third option for the SVD of the Conv2D.}
+\michael{I put this here for now as a placeholder.  Where to put it.  Maybe in a discussion/conclusion if there is space.}
 
@@
 \michael{These comments seem out of place, since they hold more generally than for the Conv2D layers.}
-
-\paragraph{Normalization of Empirical Matrices.}  
-Normalization is an important, if underappreciated, practical issue.
-Importantly, the normalization of weight matrices does \emph{not} affect the Power Law fits because the Heavy Tailed exponent $\alpha$ (as well as other metrics such as the Stable Rank and MP Soft Rank~\cite{MM18_TR,MM19_HTSR_ICML}) is scale-invariant.
-Norm-based metrics, however, do depend strongly on the scale of the weight matrix.
-Typically, to apply RMT, we would usually define Correlation Matrix with $1/N$ normalization and assume that the variance of $\mathbf{X}$ is either unity or a known constant.% 
-\footnote{For formal proofs of Heavy Tailed results, one typically needs a different normalization such as $1/N^{1-\alpha}$.}
-Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with the variance already normalized to $\sqrt{1/N}$, or some variant of this, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.
-We do not have conrol over the final empirical normalization of these models; and we do \emph{not} normalize (or renormalize) the Empirical Correlation Matrices, i.e., we use them as-is.
-The only exception to this is that we do rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $k/\sqrt{2}$ so that they are on the same scale as the Linear layers.
-
-\michael{MOVE TO LATER: COMMENT ON HOW LOG NORM first and last layers behave, maybe somewhere else.}
-
-\michael{CMOVE TO LATER: OMMENT ON HOW LOG PORM  for GPT includes unusually high alpha, not meaningful other than to show the trend.}
-
-\paragraph{The WeightWatcher Tool.}
-
-We compute the metrics using the WeightWatcher tool (version 0.2.7), and we provide Jupyter notebooks in the github repo for this paper~\cite{repo}, making the results fully reproducible.
-
-\michael{XXX.  FEW SENTENCES ABOUT REPRODUCIBILITY MORE GENERALLY HERE.}

### 2020-01-31 background.tex
edit text
mostly section 3, but touched every section

@@
 \cite{MM20_SDM},
-\cite{MM20_unpub_work}.C
+\cite{MM20_unpub_work}.
 Cite also Ganguli review and maybe other stuff.
@@
 The closest useful work would be the older work by Bouchaud and Potters, which discusses the statisical mechanics of heavy tailed and strongly correlated systems form the 90s}
+\michael{Yes on both.}
 
@@
 What we have introduced is a phenomenology, which to me is different from a semi-empirical or phenomenological theory  }
+\michael{Let's touch base to get this right.}
 \footnote{In most areas where there are complex highly-engineered systems (beyond complex AI/ML systems), one used phenomenological theory rather than first principles theory.
@@
 \item
-\textbf{Lack of accuracy metrics.}
+\textbf{Simplicity, or lack of,  accuracy metrics.}
+XXX.
 \item
-\textbf{Lack of data.}
+\textbf{Information leakage in the production pipeline.}
+XXX.
 \item
-XXX. LACK OF SOMETHING ELSE.
+\textbf{Cost of acquring labeled data.}
+XXX.
 \end{itemize}
-Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text, when developing self driving cars, or creating realistic chatbots.\footnote{i.e.
-current chatbots use perplexity as a proxy for passing a Turing test}
+Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text, when developing self driving cars, or when creating realistic chatbots.%
+\footnote{For example, current chatbots use perplexity as a proxy for passing a Turing test.}
 %, and when distilling a reliable model in some way to obtain comparable training/test quality but that damages the model in some other subtle way. 

### 2020-01-31 cv_models.tex
mostly section 3, but touched every section

@@
 \label{sxn:cv}
+
+\michael{Need to say that such large $\alpha$ values don't mean much.}
+
+\michael{Need to highlight difference between $\alpha$ and $\hat{\alpha}$.}
 
@@
 This also includes results for a larger set of ResNet models, trained on the ImageNet-1K dataset~\cite{imagenet1k}, provided on the OSMR ``Sandbox for training convolutional networks for computer vision''~\cite{osmr}, which we call the ResNet-1K series.
-We use our quality metrics to understand the gross properties of these models, comparing and contrasting the three (of four, if ResNet-1K is a distinct series) model series.
-We also use our metrics to understand more fine-scale properties of these models as a function of model depth.
-\michael{And mention distillation, if we include that in this section.}
+
+We use our quality metrics to perform \emph{coarse model analysis}, comparing and contrasting the three (or four, if ResNet-1K is a distinct series) model series, permitting us to understand the gross properties of these models.
+We also use our metrics to perform \emph{fine layer analysis}, permitting us to understand more fine-scale properties of these models as a function of model depth.
 
@@
     \centering
-    \subfigure[ Frobenius Norm ]{
+    \subfigure[ Frobenius Norm, VGG ]{
         %\includegraphics[width=5cm]{img/VGG_lognorm_accs.png}
@@
     %\qquad
-    \subfigure[ Spectral Norm ]{
+    \subfigure[ Spectral Norm, VGG ]{
         %\includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
@@
     %\qquad
-    \subfigure[ Weighted Alpha ]{
+    \subfigure[ Weighted Alpha, VGG ]{
         %\includegraphics[width=4.9cm]{img/VGG_alpha_weighted_accs.png}
@@
     %\qquad
-    \subfigure[ Alpha-Norm ]{
+    \subfigure[ $\alpha$-Norm, VGG ]{
         %\includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
@@
 
-    \subfigure[ ResNet ]{
+    \subfigure[ ResNet, $\alpha$-Norm ]{
         %\includegraphics[width=4.2cm]{img/ResNet_logpnorm_accs.png}
@@
     \qquad
-    \subfigure[ ResNet-1K ]{
+    \subfigure[ ResNet-1K, $\alpha$-Norm ]{
         %\includegraphics[width=4.5cm]{img/ResNet-1K_logpnorm_accs.png}
@@
     \qquad
-    \subfigure[ DenseNet ]{
+    \subfigure[ DenseNet, $\alpha$-Norm ]{
         %\includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
@@
     }
-    \caption{$\alpha$-Norm versus reported Top1 Error for  ResNet, ResNet-1K, and DenseNet models.}
+    \caption{$\alpha$-Norm versus reported Top1 Error for  ResNet, ResNet-1K, and DenseNet models. The corresponding results for VGG are shown in Figure~\ref{fig:vgg-pnorm}.  }
     \label{fig:cv2-accuracy}
@@
 \begin{center}
-\begin{tabular}{|p{1in}|c|c|c|c|c|}
+%\begin{tabular}{|p{1in}|c|c|c|c|c|}
+\begin{tabular}{|p{0.75in}|c|c|c|c|c|}
+%\begin{tabular}{|l|c|c|c|c|c|}
 \hline
-   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
- Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+%   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
+% Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+        & Num.   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+ Series & Models & Metric                     & Metric                     & Metric         & Metric                                   \\
 \hline
@@
 \end{center}
-\caption{RMSE for linear fits (smaller is better) of quality metrics to Reported Top1 Test Error (\cite{XXX-GIVE-REFS-AGAIN}\michael{Give refs again.}) for all pretrained models in the architecture series.
+\caption{RMSE (smaller is better) for linear fits of quality metrics to Reported Top1 Test Error for pretrained models in each architecture series.
 VGG, ResNet, and DenseNet were pretrained on ImageNet, and ResNet-1K was pretrained on ImageNet-1K.
-\michael{Do we want Frob norm squared.} }
+}
 \label{table:cv-models}
@@
 
-\paragraph{Correlation Flow in CV Models.}
+\paragraph{Layer Analytics: Correlation Flow in CV Models.}
 
@@
     \subfigure[ VGG ]{
-        \includegraphics[width=4.5cm]{img/VGG_fnl_alpha_depth.png} 
+        %\includegraphics[width=4.5cm]{img/VGG_fnl_alpha_depth.png} 
+        \includegraphics[width=3.7cm]{img/VGG_fnl_alpha_depth.png} 
                 \label{fig:vgg-alpha-layers}
@@
     \subfigure[ ResNet-1K ]{
-        \includegraphics[width=4.5cm]{img/ResNet_fnl_alpha_depth.png} 
+        %\includegraphics[width=4.5cm]{img/ResNet_fnl_alpha_depth.png} 
+        \includegraphics[width=3.7cm]{img/ResNet_fnl_alpha_depth.png} 
         \label{fig:resnet-alpha-layer}
@@
     \subfigure[ DenseNet ]{
-        \includegraphics[width=4.5cm]{img/DenseNet_fnl_alpha_depth.png} 
+        %\includegraphics[width=4.5cm]{img/DenseNet_fnl_alpha_depth.png} 
+        \includegraphics[width=3.7cm]{img/DenseNet_fnl_alpha_depth.png} 
         \label{fig:densenet-alpha-layer}
@@
 Y axes on each plot are different.
-ResNet-1K exhibits very different and much more stable behavior, which we interpret in the text as a \emph{correlation flow} across the layers.
+ResNet-1K exhibits very different and much more stable behavior across layers, which we interpret in the text as a \emph{correlation flow} across the layers.
 }
@@
 
-\paragraph{Behavior on Less Thoroughy Studied CV Models.}
+\paragraph{Layer Analytics: Behavior on Less Thoroughy Studied CV Models.}
 
@@
       %\includegraphics[scale=0.14]{img/resnet4d_maxev.png}
-      %\includegraphics[width=3.7cm]{img/resnet4d_maxev.png}
-      \includegraphics[width=4.5cm]{img/resnet4d_maxev.png}
+      \includegraphics[width=3.7cm]{img/resnet4d_maxev.png}
+      %\includegraphics[width=4.5cm]{img/resnet4d_maxev.png}
       \label{fig:resnet204Dalpha}
@@
       %\includegraphics[scale=0.14]{img/resnet4d_alphas.png}
-      %\includegraphics[width=3.7cm]{img/resnet4d_alphas.png}
-      \includegraphics[width=4.5cm]{img/resnet4d_alphas.png}
+      \includegraphics[width=3.7cm]{img/resnet4d_alphas.png}
+      %\includegraphics[width=4.5cm]{img/resnet4d_alphas.png}
       \label{fig:resnet204Dmaxev}

### 2020-01-31 introduction.tex
mostly section 3, but touched every section

@@
 and so on.
-A somewhat less well-known problem, but one that is increasingly important (in particular in industrial-scale ML---where the \emph{users} of models are not the \emph{developers} of the models) is that one may access to neither the training data nor the testing data.
+A somewhat less well-known problem, but one that is increasingly important (in particular in industrial-scale ML---where the \emph{users} of models are not the \emph{developers} of the models) is that one may have access to neither the training data nor the testing data.
 Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and be told to use it.
@@
 %
-\charles{Moreover, industry faces unique practical problems such as 'do we have enough data for this model?' It is great practical interest to have metrics to evaluate the quality of a ML model in the absence of training and testing data and without any detailed knowledge of the training and testing process.
-We seek a theory for pretrained models which can predict how, when, and why such models can be expected to perform well or poorly.}
+Moreover, in industry, one faces unique practical problems such as: do we have enough data for this model? 
+Methods that are developed and evaluated on any well-defined publicly-available coprus of data, no matter how large or diverse or interesting, are clearly not going to be well-suited to address problems such as this.
+It is thus of great practical interest to have metrics to evaluate the quality of a ML model in the absence of training and testing data and without any detailed knowledge of the training and testing process.
+We seek a practical theory for pretrained models which can predict how, when, and why such models can be expected to perform well or poorly.
 
-In this paper, we present and evaluate quality metrics for pretrained neural network (NN) and deep neural network (DNN) models at scale.%
-\footnote{We reiterate: One could use these techniques to improve training, and we have been asked about that, but we are not interested in that here.
-Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained NN models.}
+In this paper, we present and evaluate quality metrics for pretrained neural network (NN) and deep neural network (DNN) models at scale.
 To do so, we consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
@@
 \footnote{Clearly, there is a selection bias or survivorship bias here---people tend not to make publicly-available their poorly-performing models---but these models are things in the world that (like social networks or the internet) can be analyzed for their properties.}
-\michael{XXX.  MORE DETAILS.}
-Importantly, for all of these models, we have no access to training data or testing data.
+They include 
+\michael{XXX, XXX and XXX.  MORE DETAILS.}
+\michael{XXX.  LIST PLACES WHERE THEY ARE AVAIALBLE, HERE OR IN NEXT SECTION.}
+Importantly, all of these models have been trained by someone else and have been viewed to be of sufficient interest/quality to be made publicly-available, and for all of these models, we have no access to training data or testing data, and we have no knowledge of the training/testing protocols.
 
-\michael{XXX.  LIST PLACES WHERE THEY ARE AVAIALBLE, HERE OR IN NEXT SECTION.}
-
-XXX.  
-To do this, we will compute a variety of \emph{quality metrics} based on the spectral properties of the layer weight matrics.
-Note that unlike traditional ML approaches, however, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., \ncyan{hyperparmeter optimization}).
-Instead, we want to examine different models \ncyan{across} common architecture series, and we want to compare models between different architectures themselves, and in both cases \emph{we aim to predict trends in the quality of pre-trained models without access to training or testing data}.
+The \emph{quality metrics} we consider are based on the spectral properties of the layer weight matrics.
+Note that unlike traditional ML approaches, however, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., hyperparmeter optimization).%
+\footnote{We reiterate: One could use these techniques to improve training, and we have been asked about that, but we are not interested in that here.
+Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained NN models.}
+Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves, and in both cases \emph{we aim to predict trends in the quality of pre-trained models without access to training or testing data}.
 
@@
 \begin{itemize}
-\item XXX TECHNICAL THING 1
-\item XXX TECHNICAL THING 2
-\item XXX TECHNICAL THING 3
-\item XXX TECHNICAL THING 4
+\item \michael{XXX TECHNICAL THING 1}
+\item \michael{XXX TECHNICAL THING 2}
+\item \michael{XXX TECHNICAL THING 3}
+\item \michael{XXX TECHNICAL THING 4}
 \end{itemize}

### 2020-01-31 methods.tex
mostly section 3, but touched every section

--- 2020-01-30 methods.tex
@@
 We assume we are given several pretrained Deep Neural Networks (DNNs), e.g., as part of an architecture series.
-The models have been trained on (unspecified and unavailable) labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using Backprop, by minimizing some (also unspecified and unavailable) loss function $\mathcal{L}()$.
+The models have been trained on (unspecified and unavailable) labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using Backprop, by minimizing some (also unspecified and unavailable) loss function $\mathcal{L}(\cdot)$.
 We expect that most well-trained, production-quality models will employ one or more forms of on regularization, such as Batch Normalization, Dropout, etc., and many will also contain additional structure such as Skip Connections, etc.
@@
 Each DNN layer contains one or more layer 2D weight matrices $\mathbf{W}_{L}$, and/or 2D feature maps $\mathbf{W}_{i,L}$ extracted from 2D Convolutional layers. 
-(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below.) 
-%
-\michael{
-XXX.
-PROBABLY WE DO NOT WANT THE FOLLOWING.
-INSTEAD, JUST PRESENT METRICS AND POINT TO PREVIOUS PAPERS FOR JUSTIFICATION.
-E.G., WE DONT WANT TO JUSTIFY THE IMPLICIT STATISTICAL INDEPENDENCE ASSUMPTION HERE.
-IF OKAY, JUST DELETE THIS RED BLOCK.
-We assume the layer weight matrices are statistically independent, allowing us to estimate the Complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm, which resembles a data dependent VC complexity
-\begin{equation}
-\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert\cdots\Vert\mathbf{W}_{L}\Vert ,
-\end{equation}
-where $\mathbf{W}$ is an $(N\times M)$ weight matrix, with $N\ge M$, and 
- $\Vert\mathbf{W}\Vert$ is a matrix norm.   We will actually compute the log Complexity, which takes the form 
-of an Average Log Norm:
-\begin{eqnarray*}
-\log\mathcal{C} &\sim& \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert\cdots\log\Vert\mathbf{W}_{L}\Vert
-\end{eqnarray*}
-}
+(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below; and we let $\mathbf{W}$ be a generic $N\times M$ weight matrix, with $N\ge M$.) 
 %
@@
 The best performing metrics (recall that we can only consider metrics that do not use training/test data) depend on the norm and/or spectral properties of weight matrices, $\mathbf{W}$.%
-\footnote{We do not use intra-layer information from the models in our quality metrics, but as we will describe our metrics can be used to learn about }
+\footnote{We do not use intra-layer information from the models in our quality metrics, but as we will describe our metrics can be used to learn about that.}
 We consider the following metrics.
@@
 \item 
-Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{W}\Vert^{2}_{2}=\sum_{i,j}w^{2}_{i,j} = \sum_{i=1}^{M} \lambda_{i}^{2}$
-\michael{Do we use norm or log norm, here and with other norms.  Worth being very explicit and very consistent.}
+Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\sum_{i,j}w^{2}_{i,j} = \sum_{i=1}^{M} \lambda_{i}^{2}$
 \item 
 Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
-\michael{We should probably call the spectral norm $\Vert\mathbf{W}\Vert_{2}$, or change Frob norm notation and have an explicit remork.  Let's decide.}
 \item 
@@
 \item 
-$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda_{i}^{\alpha}$,
-\michael{Be clear about how we average this across layers.}
+$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda_{i}^{\alpha}$
 \end{itemize}
@@
 
-\michael{Need to say that such large $\alpha$ values don't mean much.}
-
-\michael{Need to highlight difference between $\alpha$ and $\hat{\alpha}$.}
+Each of these metrics is defined for a given layer $\mathbf{W}$ matrix.
+For the Weighted Alpha Metric, we consider the average of $\hat{\alpha}$ over all layers, i.e., $\sum_l \hat{\alpha}_l = \sum_l \alpha_l\log\lambda_{max,l} $ .  
+\michael{Be careful to get that right.}
+For the norm-based metrics, we consider the average of the log of the norm to the appropriate power.
+For example, for the $\alpha$-Norm, we compute 
+\begin{equation}
+\label{eqn:sum_log_alpha_norm_alpha}
+\sum\nolimits_l \log \Vert\mathbf{W}_l\Vert_{\alpha_l}^{\alpha_l} 
+=
+\sum\nolimits_l \alpha_l \log \Vert\mathbf{W}_l\Vert_{\alpha_l} .
+\end{equation}
+Informallly, this amounts to assuming that the layer weight matrices are statistically independent, in which case we can estimate the model complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm (which resembles a data dependent VC complexity),
+\begin{equation}
+\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert \times \cdots \times \Vert\mathbf{W}_{L}\Vert ,
+\end{equation}
+where $\Vert\cdot\Vert$ is a matrix norm.   
+In this case, the log Complexity takes the form of an Average Log Norm,
+\begin{equation}
+\label{eqn:eqn:sum_log_norm}
+\log\mathcal{C} \sim \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert + \cdots + \log\Vert\mathbf{W}_{L}\Vert  .
+\end{equation}
+For the Frobenius and Spectral norms, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly. 
+For example, when taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.
+For the Weighted Alpha Norm, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it cannot be taken out of the sum.
 
@@
 
-\paragraph{XXX.}
-XXX.  THIS VERIFICATION IS NOT ABOUT CONV LAYERS, IT IS ABOUT PL MORE GENERALLY, CORRECT?  WE SHOULD CLARIFY AND SQUISH.
-To verify that our approach is meaningful, we need to confirm that the ESD is neither due to a random matrix, nor due to unsually large matrix elements, but, in fact, captures correlations learned from the data.
-We examine typical layer for the pretained AlexNet model (distributed with pyTorch). 
-Figure~\ref{fig:alexnet1} displays the ESD for the first slice (or matrix $\mathbf{W}$) of the third Conv2D layer, extracted from a 4-index Tensor of shape $(384, 192, 3, 3)$.
-The red line displays the best fit to a random matrix, using the Marchenko pastur theory~\cite{MM}.
-We can see the random matrix model does not describe the ESD very well.
-For comparison, Figure \ref{fig:alexnet2} shows the ESD of the same matrix, randomly shuffled; here looks similar to the red line plot of the orginal ESD.
-In fact, the empircal ESD is better modeled with a truncated power law distribtion.
-\michael{We may want to give a one-sentence summary of this par and fig at the end of the previous par.}
-
-\begin{figure}[H]
-   \centering
-   \subfigure[Actual ESD]{
-     %\includegraphics[scale=0.5]{img/alexnet1.png} 
-     \includegraphics[scale=0.25]{img/alexnet1.png} 
-     \label{fig:alexnet1}
-   }
-   \subfigure[ESD of randomly shuffled matrix]{
-      %\includegraphics[scale=0.5]{img/alexnet2.png}
-      \includegraphics[scale=0.25]{img/alexnet2.png}
-      \label{fig:alexnet2}
-   }
-   \caption{ESD of AlexNet Conv2D pre-Activation map for Layer 3 Slice 1, actual and randomized.
-            \michael{Need better figs here.}
-           }
-   \label{fig:alexnet}
-\end{figure}
-
-Although the ESD is \emph{Heavy Tailed}, this does not imply that the orginal matrix $\mathbf{W}$ is itself heavy tailed--only the correlation matrix $\mathbf{X}$ is. 
-If $\mathbf{W}$ was, then it would contain 1 or more unusually large matrix elements, and they would dominate the ESD.  
-Of course the randomized $\mathbf{W}$ would also be heavy tailed, but its ESD neither resembles the original nor is it heavy tailed. 
-So we can rule out $\mathbf{W}$ being heavy tailed.
-\michael{These comments seem out of place, since they hold more generally than for the Conv2D layers.}
-
-These plots tell us that the pre-activation maps of the Conv2D contains significant correlations learned from the data.  
-By modeling the ESD with a power law distribution $\lambda^{\alpha}$, we can characterize the amount of correlation learned;
-the smaller the exponent $\alpha$, the more correlation in the weight matrix. 
-\michael{These comments seem out of place, since they hold more generally than for the Conv2D layers.}
-
 \paragraph{Normalization of Empirical Matrices.}  
@@
 
-\michael{MOVE TO LATER: COMMENT ON HOW LOG NORM first and last layers behave, maybe somewhere else.}
-
-\michael{CMOVE TO LATER: OMMENT ON HOW LOG PORM  for GPT includes unusually high alpha, not meaningful other than to show the trend.}
-
 \paragraph{The WeightWatcher Tool.}
@@
 We compute the metrics using the WeightWatcher tool (version 0.2.7), and we provide Jupyter notebooks in the github repo for this paper~\cite{repo}, making the results fully reproducible.
-
-\michael{XXX.  FEW SENTENCES ABOUT REPRODUCIBILITY MORE GENERALLY HERE.}
+\michael{XXX.  FEW SENTENCES ABOUT REPRODUCIBILITY MORE GENERALLY HERE, HERE OR IN DISCUSSION/CONCLUSION.}

### 2020-01-31 nlp_models.tex
mostly section 3, but touched every section

@@
 
+%        &  Num.  & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+% Series & Models & Metric                     & Metric                     & Metric         & Metric                                   \\
+
 \begin{table}[t]
@@
 \begin{center}
-\begin{tabular}{|p{1in}|c|c|c|c|c|}
+%\begin{tabular}{|p{1in}|c|c|c|c|c|}
+\begin{tabular}{|p{0.75in}|c|c|c|c|c|}
 \hline
-   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
- Series & \#Layers   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+%   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
+% Series & \#Layers   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+        & Num.   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+ Series & Layers & Metric                     & Metric                          & Metric         & Metric                                   \\
 \hline
@@
 \end{center}
-\caption{Average Log Norm Metrics for pretrainnd OpenAI GPT and GPT2 models.}
+\caption{Average value for each of the four metrics for pretrainnd OpenAI GPT and GPT2 models.}
 \label{table:nlp}

### 2020-01-31 cv_models.tex
typo

@@
 
-\paragraph{Quality Metrics versus Reported Test Accuracies.}
+\paragraph{Coarse Analysis: Quality Metrics versus Reported Test Accuracies.}
 
@@
 
-\paragraph{Layer Analytics: Correlation Flow in CV Models.}
+\paragraph{Layer Analysis: Correlation Flow in CV Models.}
 
@@
 
-\paragraph{Layer Analytics: Behavior on Less Thoroughy Studied CV Models.}
+\paragraph{Layer Analysis: Behavior on Less Thoroughy Studied CV Models.}
 

### 2020-01-31 cv_models.tex
local opt on sec 4

@@
 
-\michael{Need to say that such large $\alpha$ values don't mean much.}
+In this section, we examine quality metrics described in Section~\ref{sxn:methods} for several CV model architecture series.
+This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current opensource pyTorch framework (version 1.4)~\cite{pyTorch}.
+This also includes a larger set of ResNet models, trained on the ImageNet-1K dataset~\cite{imagenet1k}, provided on the OSMR ``Sandbox for training convolutional networks for computer vision''~\cite{osmr}, which we call the ResNet-1K series.
 
-\michael{Need to highlight difference between $\alpha$ and $\hat{\alpha}$.}
-
-In this section, we examine the quality metrics described in Section~\ref{sxn:methods} for several CV model architecture series.
-This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current opensource pyTorch framework (version 1.4)~\cite{pyTorch}.
-This also includes results for a larger set of ResNet models, trained on the ImageNet-1K dataset~\cite{imagenet1k}, provided on the OSMR ``Sandbox for training convolutional networks for computer vision''~\cite{osmr}, which we call the ResNet-1K series.
-
-We use our quality metrics to perform \emph{coarse model analysis}, comparing and contrasting the three (or four, if ResNet-1K is a distinct series) model series, permitting us to understand the gross properties of these models.
+We use our quality metrics to perform \emph{coarse model analysis}, comparing and contrasting the three (or four, if ResNet-1K is a distinct series) model series, permitting us to predict trends in model quality for these series.
 We also use our metrics to perform \emph{fine layer analysis}, permitting us to understand more fine-scale properties of these models as a function of model depth.
@@
 
-We have examined the performance of the four quality (Frobenius norm, Spectral norm, Weighted Alpha, and $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
-To start, Figure~\ref{fig:vgg-metrics} plots the four quality metrics versus the reported Test accuracies~\cite{pyTorchVgg}, as well as a basic linear regression line, 
-%with the Root Mean Squared Error (RMSE) shown, 
-for the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BatchNormalization.
-\michael{These are all log norms, right?}
-\michael{XXX.  WHERE IS RMSE.  IN THAT TABLE?}
+We have examined the performance of the four quality metrics (Frobenius norm, Spectral norm, Weighted Alpha, and $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
+To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BatchNormalization), and it plots the four quality metrics versus the reported Test accuracies~\cite{pyTorchVgg}, as well as a basic linear regression line.
+%with the Root Mean Squared Error (RMSE) shown, \michael{XXX.  WHERE IS RMSE.  IN THAT TABLE?}
 All four metrics correlate quite well with the reported Top1 Accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalziation (i.e., greater accuracy, lower error). 
-Similar results (not shown) are obtained for the Top5 Accuracies.  
-Notice that the log $\alpha$-Norm metric $\log\Vert\mathbf{W}\Vert)_{\infty}^{\infty}$ 
-\michael{XXX IS THAT A TYPO} 
-performs best, with an RMSE of $0.34$; and the second best metric is $\hat\alpha$, which is an approximate log $\alpha$-Norm metric.
-\michael{Are those numbers stale, I don't see them, and some things seem in the other order.}
-\charles{See Figure 2}
-\michael{We should probably point to the table if we cite quantitative results.}
-
-Figure~\ref{fig:cv2-accuracy} plots the best performing metric ($\alpha$-Norm) for the full ResNet, ResNet-1K, and DenseNet series.
-(A much more detailed set of plots---including the behavior of all four metrods on each of the series---are availale at \michael{give web link}.)
-For a more visual depiction, Figure~\ref{fig:cv2-accuracy} displays the log $\alpha$-Norm metric  $\log\Vert\mathbf{W}\Vert)_{\infty}^{\infty}$ versus test accuracies for the ResNet, ResNet-1K, and DenseNet series.
-Notice that ResNet series, which has been trained on the full
-ImageNet dataset, has an RMSE of $0.66$, whereas the ResNet-1K series, which has been trained on the much smaller ImageNet-1K dataset,
-also correlated wells, but has a much larger RMSE of $1.9$.
-
-Table \ref{table:cv-models} summarizes results for all four metrics (Frobenius norm, Spectral norm, Weighted Alpha, and $\alpha$-Norm) applied to the VGG, ResNet, ResNet-1K, and DenseNet series.
-\michael{XXX.  AGAIN, WHERE IS RMSE, IN THAT TABLE?}
-\charles{The Table may be stale, see the figures to confirm}
-
-For all series, all four metrics correlate quite well with the reported Top1 test accuracies.  
-Moreover, the $\alpha$-Norm and/or the approximate Weighted Alpha metrics perform best.
+While all four metrics perform well, notice that the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) performs best (with an RMSE of $0.42$, see Table~\ref{table:cv-models}); and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $), which is an approximation to the Log $\alpha$-Norm metric~\cite{MM20_unpub_work}, performs second best (with an RMSE of $0.48$, see Table~\ref{table:cv-models}).
 
@@
 \hline
- VGG & 6 & 0.56 & 0.53 & 0.48 & 0.42  \\
- ResNet & 5 & 0.9 & 1.4 & 0.61 & 0.66  \\
- ResNet-1K & 19 & 2.4 & 3.6 & 1.8 & 1.9  \\
- DenseNet & 4 & 0.3 & 0.26 & 0.16 & 0.21 \\
+ VGG       &  6 & 0.56 & 0.53 & 0.48          & \textbf{0.42}  \\
+ ResNet    &  5 & 0.9  & 1.4  & \textbf{0.61} & 0.66           \\
+ ResNet-1K & 19 & 2.4  & 3.6  & \textbf{1.8}  & 1.9            \\
+ DenseNet  &  4 & 0.3  & 0.26 & \textbf{0.16} & 0.21           \\
 \hline
@@
 VGG, ResNet, and DenseNet were pretrained on ImageNet, and ResNet-1K was pretrained on ImageNet-1K.
+\michael{Are the numbers in this table stale; if they get updated, then we need to modify the test at at least two places, on actual numbers and on what is best.}
 }
@@
 \end{table}
+
+Figure~\ref{fig:cv2-accuracy} plots the 
+%overall best performing (i.e., 
+Log $\alpha$-Norm
+%) 
+metric for the full ResNet, ResNet-1K, and DenseNet series.
+(A much more detailed set of plots, including the behavior of all four methods on each of the series, is available at \cite{XXX-WEB-LINK}.)
+Here, the Log $\alpha$-Norm is much better thatn the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is actually slightly worse than the Weighted Alpha metric).
+Notice also that ResNet series, which has been trained on the full ImageNet dataset, has an RMSE of $0.66$, whereas the ResNet-1K series, which has been trained on the much smaller ImageNet-1K dataset, also correlated wells, but has a much larger RMSE of $1.9$.
+
+See Table~\ref{table:cv-models} for a summary of results for Top1 Accuracies for all four metrics on all four model series.
+Similar results (not shown) are obtained for the Top5 Accuracies.
+Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best. 
+These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish ``good-better-best,'' with PL-based metrics performing somewhat better.
 
@@
 % in the network.
-The most interesting results are seen when we plot the PL exponent, $\alpha$, as a function of depth.
+%The most interesting results are seen when we 
+For example, we can 
+plot the PL exponent, $\alpha$,%
+\footnote{That is, here we consider just $\alpha$ for each layer, not $\hat{\alpha}$ or $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$.} 
+as a function of depth.
 %
-See Figure~\ref{fig:vgg-alpha-layers}, which plots $\alpha$ for each layer (the first corresponds to data, the last to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (without BatchNormalization), ResNet-1K, and DenseNet~series.
+See Figure~\ref{fig:vgg-alpha-layers}, which plots $\alpha$ for each layer (the first layer corresponds to data, the last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (without BatchNormalization), ResNet-1K, and DenseNet series.
+(Again, much more detailed set of plots is available at \cite{XXX-WEB-LINK}; but note that the corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than the results we present here.)
 
@@
     }
-    \caption{PL exponent $\alpha$ versus layer for VGG, ResNet-1K, and DenseNet models.
-Y axes on each plot are different.
-ResNet-1K exhibits very different and much more stable behavior across layers, which we interpret in the text as a \emph{correlation flow} across the layers.
+    \caption{Layer Analysis. 
+             PL exponent $\alpha$ versus layer id, for VGG, ResNet-1K, and DenseNet models.  
+             (Y axes on each plot are different.)  
+             ResNet-1K exhibits different and much more stable behavior across layers.
+             This contrasts with how VGG gets gradually worse in deeper layers and how DenseNet is worse and much more erratic.  
+             We interpret this in the text as quantifying a \emph{correlation flow} across the layers.  
             }
@@
 
-In the VGG models, $\alpha$ systematically increases as we move down the network, from data to labels, in the Conv2D layers, starting with $\alpha\lesssim 2.0$ and reaching all the way to $\alpha\sim 5.0$; and then, in the last three, large, fully-connected (FC) layers, $\alpha$ stabilizes back down to $\alpha\in[2,2.5]$.
-This is seen for all the VGG models (again, only the shallowest and deepest are shown in this figure), indicating that the main effect of increasing depth is to increase the range over which $\alpha$ can increase, thus leasing to larger $\alpha$ values in later Conv2D layers of the VGG models.
+In the VGG models, Figure~\ref{fig:vgg-alpha-layers} shows that the PL exponent $\alpha$ systematically increases as we move down the network, from data to labels, in the Conv2D layers, starting with $\alpha\lesssim 2.0$ and reaching all the way to $\alpha\sim 5.0$; and then, in the last three, large, fully-connected (FC) layers, $\alpha$ stabilizes back down to $\alpha\in[2,2.5]$.
+This is seen for all the VGG models (again, only the shallowest and deepest are shown in this figure), indicating that the main effect of increasing depth is to increase the range over which $\alpha$ increases, thus leading to larger $\alpha$ values in later Conv2D layers of the VGG models.
 This is quite different than the behavior of either the ResNet-1K models or the DenseNet models.
 
-For the ResNet-1K models, $\alpha$ also increases in the last few layers, but as the models get deeper, there is a wide range over which $\alpha$ values tend to remain small.
+For the ResNet-1K models, $\alpha$ also increases in the last few layers (more dramatically, in fact, observe the differing scales on the Y axes).
+However, as the ResNet-1K models get deeper, there is a wide range over which $\alpha$ values tend to remain small.
 \michael{Can we say something about exceptions which are larger than for VGG or DenseNet.}
 \charles{You have the same plots I do..what do you want to say >}
+\michael{Are they small or different types of layers or something?}
 This is seen for other models in the ResNet-1K series, but it is most pronounced for the larger ResNet-1K (152) model, where $\alpha$ remains relatively stable at $\alpha\sim 2.0$, from the earliest layers all the way until we reach close to the final layers.
@@
 For the DenseNet models, $\alpha$ tends to increase as the layer id increases, in particular for layers toward the end.
-This is similar to what is seen in the VGG models, but with the DenseNet models, the variance is much larger (in particular for the earlier and middle layers, where it can range all the way to $\alpha\lesssim 8.0$) and much less systematic.
+While this is similar to what is seen in the VGG models, with the DenseNet models, $\alpha$ values increase almost immediately after the first few layers, and the variance is much larger (in particular for the earlier and middle layers, where it can range all the way to $\alpha\lesssim 8.0$) and much less systematic throughout the network.
 
-We can understand Figure~\ref{fig:vgg-alpha-layers} in terms of the \emph{flow of correlation}, as follows.
-Informally, one would expect a DNN model to perform well when it fascilitates the propagation of information/features across layers.
-In the absence of training/test data, we can to try to quantify this by measuring the PL properties of weight matrices.
-Smaller $\alpha$ values correspond to layers in which correlations across multiple scales are better captured~\cite{MM18_TR,SornetteBook}, and thus we expect that values of $\alpha$ that are stable across multiple layers enable better \emph{correlation flow} through the network.
-
-Consider the differences---in particular, the number of residual connections---between the VGG, ResNet, and DenseNet architectures.
-VGG, while a good model in many ways, was limited by needing a massive FC layers at the end of the model, requiring significant memory and computational resources, and it's poor conditioning led to problems with vanishing gradients.
+Recall the differences---in particular, the number of residual connections---between the VGG, ResNet, and DenseNet architectures.
+VGG, while a fine model in many ways, was limited by needing massive FC layers at the end of the model, requiring significant memory and computational resources, and it's poor conditioning led to problems with vanishing gradients.
 The empirical manifestations of this (on weight matrices) are that fitted $\alpha$ values get much worse for deeper layers.
@@
 This should also be contrasted with the DenseNet model, which contains many connections between every layer.
-Our results (large $\alpha$, meaning they even a HT model is probably a poor fit) suggest that DenseNet has too many connections, diluting the correlation flow across layers, and leaving many layers very poorly optimized.
+Our results (large $\alpha$, meaning they even a HT model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly optimized.
 
-\paragraph{Layer Analysis: Behavior on Less Thoroughy Studied CV Models.}
+More generally, we can understand Figure~\ref{fig:vgg-alpha-layers} in terms of the \emph{flow of correlation}, as follows.
+Recall that the Log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on ideas from the statisical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
+There, one expects well-trained systems which exhibit correlations over many size scales to have ESDs that can be well-fit by PLs, with PL exponents $\alpha\approx 2$.
+Much larger values of $\alpha$ indicate very poor fits, but the presence of many adjacent layers with relatively good PL fits \emph{suggests} that those adjacent layers ``impedance match'' well and that the correlations that are captured by one layer propagate well to subsequent layers.
+Informally, one would expect a DNN model to perform well when it fascilitates the propagation of information/features across layers.
+In the absence of training/test data, we can to try to quantify this by measuring the PL properties of empirical weight matrices.
+Smaller $\alpha$ values correspond to layers in which correlations across multiple scales are better captured~\cite{MM18_TR,SornetteBook}, and thus we expect that small values of $\alpha$ that are stable across multiple layers enable better \emph{correlation flow} through the network.
 
-Figure~\ref{fig:vgg-metrics} and Table~\ref{table:cv-models} show that---for well-trained and widely-studied models---the weighted $\hat{\alpha}$ metric performs better than but similar to norm-based metrics.
-This suggests an obvious question: is the $\hat{\alpha}$ metric just a variation of these more familiar norm-based metrics, or does it capture something different?  
-To show that it is not just a variation and that it does capture something different, consider the following less thoroughy studied model, a compressed DNN model~\cite{CWZZ17_TR}, where we observed that the average (Frobenius or Spectral) norm increases with decreasing test error, whereas the average $\alpha$ decreases, as expected.
+%\paragraph{Layer Analysis: Behavior on Less Thoroughy Studied CV Models.}
+\paragraph{Layer Analysis: Differences Between PL Metrics and Norm Metrics.}
+
+The similarity between norm-based metrics and $\alpha$-based metrics suggests a question: is the Weighted Alpha metric just a variation of the more familiar norm-based metrics, or does it capture something different?
+More generally, do fitted $\alpha$ values contain information not captured by norms? 
+To show that it is not just a variation and that it does capture something different, consider the following example, which involves DN model compressed/distillation~\cite{CWZZ17_TR}.
 
@@
    }
-   \caption{
-     Analysis of ResNet20, distilled with Group Regularization, as implemented in the \texttt{distiller} (4D\_regularized\_5Lremoved) pre-trained models.  
-     Comparison of individual layer $\mathbf{W}_{l}$ maximum eigenvalues ($\lambda_{max}$, or Spectral Norms) and  
-     PL exponent $\alpha$, between baseline (green) and fine-tuned (red)  pre-trained models.   
+   \caption{Layer Analysis.
+            ResNet20, distilled with Group Regularization, as implemented in the \texttt{distiller} (4D\_regularized\_5Lremoved) pre-trained models.  
+            Spectral Norm, $\lambda_{max}$, and, PL exponent, $\alpha$, for individual layer, versus layer id, for both between baseline (before distillation, shown in green) and fine-tuned (after distillation, shown in red) pre-trained models.
            }
@@
 
-Consider average metrics measured on ResNet20, trained on CIFAR10, before and after applying the Group Regularization technique, as implemented in the \texttt{distiller} package.%
+We consider ResNet20, trained on CIFAR10, before and after applying the Group Regularization technique, as implemented in the \texttt{distiller} package.%
 \footnote{For details, see \url{https://nervanasystems.github.io/distiller/\#distiller-documentation} and also \url{https://github.com/NervanaSystems/distiller}.}
 We analyze the available pre-trained 4D\_regularized\_5Lremoved baseline and fine-tuned models.  %\cite{distiller_repo}.
+The reported baseline test accuracies ($Top1=91.45$ and $Top5=99.75$) are better than the reported fine-tuned test accuracies ($Top1=91.02$ and $Top5=99.67$)~\cite{XXX-XXX}.
+The results of Figure~\ref{fig:vgg-metrics} and Figure~\ref{fig:cv2-accuracy} might suggest that the baseline Spectral Norm should be \emph{smaller} than those of the layers in the fine-tuned model.
+\michael{Do we have something backwards here: where we observed that the average norm to increase with decreasing test error (as it ``should'' not), whereas the average PL exponent $\alpha$ decreases (as ``expected'' from HT-SR Theory).
+}
+In both cases (Frobenius norm results not shown), we observe the opposite.
 Figure~\ref{fig:resnet204D5L} presents the Spectral Norm ($\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}_{l}$.%
-\footnote{We only include layer matrices or feature maps with $M\ge50$.}
-The reported baseline test accuracies ($Top1=91.450$ and $Top5=99.750$) are better than the reported fine-tuned test accuracies ($Top1=91.020$ and $Top5=99.670$).
-Thus, the results of Figure~\ref{fig:vgg-metrics} and Figure~\ref{fig:cv2-accuracy} might suggest that the baseline Spectral (and Frobenius) Norm should be \emph{smaller} than those of the layers in the fine-tuned model.
-In both cases (Frobenius norm results not shown), we observe the opposite.
-On the other hand, the $\alpha$ values do not systematically differ between the baseline and fine-tuned models.
-Also (not shown), the average (unweighted) baseline $\alpha$ is smaller than the fine-tuned average (as would be predicted by HT-SR Theory, on which $\hat{\alpha}$ is based).
+\footnote{Here, we only include layer matrices or feature maps with $M\ge50$.}
+On the other hand, the $\alpha$ values do not differ systematically between the baseline and fine-tuned models.
+\michael{Some comment about big jump in norm; but fix previous backwards question first.}
+Also (not shown), the average (unweighted) baseline $\alpha$ is \emph{smaller} than the fine-tuned average (as would be predicted by HT-SR Theory, on which $\hat{\alpha}$ is based).
 
+\michael{XXX.  WORK ON THIS PAR AFTER DO NLP EXAMPLES.}
+Our interpretation of this is the following.
+The pre-trained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the VGG/ResNet/DenseNet models we condisered previously.
+Norms are often used as regularizers, and thys they are a natural metric to consider, e.g., during training or distillation.
+The PL $\alpha$ captures more subtle correlations.
+%
 The reason for this is that the \texttt{distiller} Group Regularization technique has the unusual effect of increasing the norms of the $\mathbf{W}$ feature maps for at least two of the Conv2D layers.
@@
 CLARIFY.}
+\michael{XXX.  INTERPRET IN TERMS OF XXX GOOD VERSUS BAD, AS OPPOSED TO GOOD BETTER BEST.}

### 2020-01-31 nlp_models.tex
local opt on sec 4

@@
 \label{sxn:nlp}
+
+\michael{Need to say that such large $\alpha$ values don't mean much.}
+
+\michael{Need to highlight difference between $\alpha$ and $\hat{\alpha}$.}
 
@@
     \subfigure[Power Law Exponent $(\alpha)$]{
-        \includegraphics[width=5cm]{img/GPT-alpha-hist.png}
+        %\includegraphics[width=5cm]{img/GPT-alpha-hist.png}
+        \includegraphics[width=3.7cm]{img/GPT-alpha-hist.png}
         \label{fig:GPT-alpha-hist}
@@
     \subfigure[log Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
-        \includegraphics[width=5cm]{img/GPT-snorm-hist.png}
+        %\includegraphics[width=5cm]{img/GPT-snorm-hist.png}
+        \includegraphics[width=3.7cm]{img/GPT-snorm-hist.png}
         \label{fig:GPT-snorm-hist}
@@
     \subfigure[Power Law Exponent $(\alpha)$]{
-        \includegraphics[width=4.5cm]{img/GPT-alpha-depth.png}
+        %\includegraphics[width=4.5cm]{img/GPT-alpha-depth.png}
+        \includegraphics[width=3.7cm]{img/GPT-alpha-depth.png}
         \label{fig:gpt-alpha-layer}
@@
     \subfigure[log Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
-        \includegraphics[width=4.5cm]{img/GPT-snorm-depth.png}
+        %\includegraphics[width=4.5cm]{img/GPT-snorm-depth.png}
+        \includegraphics[width=3.7cm]{img/GPT-snorm-depth.png}
         \label{fig:gpt-snorm-layer}
@@
     \subfigure[log Alpha-Norm $(\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha})$]{
-        \includegraphics[width=4.5cm]{img/GPT-pnorm-depth.png}
+        %\includegraphics[width=4.5cm]{img/GPT-pnorm-depth.png}
+        \includegraphics[width=3.7cm]{img/GPT-pnorm-depth.png}
         \label{fig:gpt-pnorm-layer}
@@
     \subfigure[Power Law Exponent $(\alpha)$]{
-        \includegraphics[width=4.5cm]{img/GPT2_all_alpha_hist.png}
+        %\includegraphics[width=4.5cm]{img/GPT2_all_alpha_hist.png}
+        \includegraphics[width=3.7cm]{img/GPT2_all_alpha_hist.png}
         \label{fig:gpt2-alpha-hist}
@@
     \subfigure[log Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
-        \includegraphics[width=4.5cm]{img/GPT2*_all_spectralnorm_hist.png}
+        %\includegraphics[width=4.5cm]{img/GPT2*_all_spectralnorm_hist.png}
+        \includegraphics[width=3.7cm]{img/GPT2*_all_spectralnorm_hist.png}
         \label{fig:gpt2-snorm-hist}
@@
     \subfigure[log Alpha-Norm $(\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha})$]{
-        \includegraphics[width=4.5cm]{img/GPT2_all_logpnorm_hist.png}
+        %\includegraphics[width=4.5cm]{img/GPT2_all_logpnorm_hist.png}
+        \includegraphics[width=3.7cm]{img/GPT2_all_logpnorm_hist.png}
         \label{fig:gpt2-pnorm-hist}

### 2020-02-01 background.tex
local opt on sec 4
put in some refs

@@
 
-XXX.
-Cite our stuff:
-\cite{MM17_TR},
-\cite{MM18_TR},
-\cite{MM19_HTSR_ICML},
-\cite{weightwatcher_package}
-\cite{MM19_KDD},
-\cite{MM20_SDM},
-\cite{MM20_unpub_work}.
-Cite also Ganguli review and maybe other stuff.
-\charles{Gangulis work is not really close to what we are doing, although I get the politics.
-The closest useful work would be the older work by Bouchaud and Potters, which discusses the statisical mechanics of heavy tailed and strongly correlated systems form the 90s}
-\michael{Yes on both.}
+Statistical mechanical ideas have long had influence on NN theory and practice~\cite{EB01_BOOK, MM17_TR, BKPx20}; and 
+our best-performing metrics (i.e., those using fitted PL exponents and HT-SR Theory) are based on statistical mechanics ideas~\cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM19_KDD, MM20_SDM}.
+However, that the way in which we \emph{use} statistical mechanics theory is quite different than the way it is more commonly formulated.
+Several very good overviews of the more common approach are available~\cite{EB01_BOOK, BKPx20}.
+We will \emph{use} statistical mechanics theory more as it is used in quantitative finance.
+Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statisical mechanics of heavy tailed and strongly correlated systems.
 
@@
 \paragraph{Norm-based capacity control theory.}
-XXX.  MOST IRRELEVANT, BUT LIAO AND OUR SDM ARE RELATED.  
-XXX.  MAYBE BEAT ON INFINITELY WIDE OR SOMETHING ELSE.
+
+There is also a large body of work on using norm-based metrics to bound generalization error~\cite{NTS15, BFT17_TR, LMBx18_TR}.
+In this area, theoretical work aims to prove generalization bounds, and applied work uses these norms as a regularization function to improve training.
+While we do find that norms provide relatively good quality metrics, at least for distinguishing good-better-best among well-trained models, we are not interested in proving generalization bounds or improving training.
 

### 2020-02-01 introduction.tex
local opt on sec 4
put in some refs

@@
 \end{itemize}
+MENTION: The most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in a system) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
+AND DEFINE HT-SR ACRONYM.
+\michael{Sentence or two about how our metrics come from norm-based SLT and alpha-based HT-SR theory, since that is also highlighted in the abstract.}
+
+SOMEWHERE CITE WEIGHTWATCHER: \cite{weightwatcher_package}
 

### 2020-02-01 background.tex
minor update

@@
 \end{itemize}
-Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text, when developing self driving cars, or when creating realistic chatbots.%
+Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text or realiztic chatbots, developing self driving cars, or just clusterting user profiles.
 \footnote{For example, current chatbots use perplexity as a proxy for passing a Turing test.}

### 2020-02-01 methods.tex
minor update to methods

@@
 We expect that most well-trained, production-quality models will employ one or more forms of on regularization, such as Batch Normalization, Dropout, etc., and many will also contain additional structure such as Skip Connections, etc.
-Here, we will ignore these details, and will focus only on the weight matrices.
+Here, we will ignore these details, and will focus only on the layer weight matrices $\mathbf{W}_{l}$.
 
@@
 
-Each DNN layer contains one or more layer 2D weight matrices $\mathbf{W}_{L}$, and/or 2D feature maps $\mathbf{W}_{i,L}$ extracted from 2D Convolutional layers. 
+Each DNN layer contains one or more layer 2D weight matrices $\mathbf{W}_{l}$, and/or 2D feature maps $\mathbf{W}_{i,l}$ extracted from 2D Convolutional layers. 
 (For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below; and we let $\mathbf{W}$ be a generic $N\times M$ weight matrix, with $N\ge M$.) 
@@
 \michael{We need to be clear that this is a truncated power law fit and that $\lambda_{max}$ comes from that and not the largest empirical eigenvalue.}
+\charles{Actually $\lambda_{max}$ \emph{is} the largest empirical eigenvalue}
 The Weighted Alpha Metric was introduced previously~\cite{MM20_SDM}, where (on a much smaller set of data than we consider here) it was shown to correlate well with the trends in reported test accuracies of pretrained DNNs.
@@
 \michael{Be careful to get that right.}
-For the norm-based metrics, we consider the average of the log of the norm to the appropriate power.
+For the norm-based metrics, we consider the \emph{weighted} average of the log of the norm to the appropriate power.
 For example, for the $\alpha$-Norm, we compute 

### 2020-02-01 cv_models.tex
minor change

@@
 
-In this section, we examine quality metrics described in Section~\ref{sxn:methods} for several CV model architecture series.
+In this section, we examine quality metrics described in Section~\ref{sxn:methods} for several Computer Vision (CV) model architecture series.
 This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current opensource pyTorch framework (version 1.4)~\cite{pyTorch}.
@@
 
-We use our quality metrics to perform \emph{coarse model analysis}, comparing and contrasting the three (or four, if ResNet-1K is a distinct series) model series, permitting us to predict trends in model quality for these series.
-We also use our metrics to perform \emph{fine layer analysis}, permitting us to understand more fine-scale properties of these models as a function of model depth.
+We use our quality metrics to perform \emph{coarse model analysis}, comparing and contrasting the three model series (or four, if ResNet-1K is included), permitting us to predict trends in model quality.
+We also use perform a more \emph{fine layer analysis} as a function of depth for these models, providing novel insights among the VGG, ResNet, and DenseNet architectures.
 
@@
 %   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
-% Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
-        & Num.   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+% Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+        & Num.   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
  Series & Models & Metric                     & Metric                     & Metric         & Metric                                   \\

### 2020-02-01 cv_models.tex
cv model edits

@@
 (A much more detailed set of plots, including the behavior of all four methods on each of the series, is available at \cite{XXX-WEB-LINK}.)
-Here, the Log $\alpha$-Norm is much better thatn the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is actually slightly worse than the Weighted Alpha metric).
+Here, the Log $\alpha$-Norm is much better that the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is actually slightly worse than the Weighted Alpha metric).
 Notice also that ResNet series, which has been trained on the full ImageNet dataset, has an RMSE of $0.66$, whereas the ResNet-1K series, which has been trained on the much smaller ImageNet-1K dataset, also correlated wells, but has a much larger RMSE of $1.9$.
@@
 Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best. 
-These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish ``good-better-best,'' with PL-based metrics performing somewhat better.
+These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among ``good-better-best'' models, with PL-based metrics performing somewhat better.
 
@@
 For example, we can 
-plot the PL exponent, $\alpha$,%
+plot \nred{just} the PL exponent, $\alpha$,%
 \footnote{That is, here we consider just $\alpha$ for each layer, not $\hat{\alpha}$ or $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$.} 
@@
 \michael{Are they small or different types of layers or something?}
+\charles{Not that I am aware of.  Ill think on it}
 This is seen for other models in the ResNet-1K series, but it is most pronounced for the larger ResNet-1K (152) model, where $\alpha$ remains relatively stable at $\alpha\sim 2.0$, from the earliest layers all the way until we reach close to the final layers.
@@
 More generally, do fitted $\alpha$ values contain information not captured by norms? 
-To show that it is not just a variation and that it does capture something different, consider the following example, which involves DN model compressed/distillation~\cite{CWZZ17_TR}.
+To show that it is not just a variation and that $\alpha$ does capture something different, consider the following example, which looks at a compressed / distilled DNN model~\cite{CWZZ17_TR}.
 
@@
 The reported baseline test accuracies ($Top1=91.45$ and $Top5=99.75$) are better than the reported fine-tuned test accuracies ($Top1=91.02$ and $Top5=99.67$)~\cite{XXX-XXX}.
-The results of Figure~\ref{fig:vgg-metrics} and Figure~\ref{fig:cv2-accuracy} might suggest that the baseline Spectral Norm should be \emph{smaller} than those of the layers in the fine-tuned model.
+The previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy} might suggest that the baseline Spectral Norm should be \emph{smaller} than those of the layers in the fine-tuned model.
 \michael{Do we have something backwards here: where we observed that the average norm to increase with decreasing test error (as it ``should'' not), whereas the average PL exponent $\alpha$ decreases (as ``expected'' from HT-SR Theory).

### 2020-02-01 methods.tex
minor changes, starting nlp

@@
 Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
+\michael{Should this be $\sqrt{\lambda_{max}}$?}
 \item 
@@
 \begin{equation}
-\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}  .
+\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}  ,
 \end{equation}
-\michael{We need to be clear that this is a truncated power law fit and that $\lambda_{max}$ comes from that and not the largest empirical eigenvalue.}
-\charles{Actually $\lambda_{max}$ \emph{is} the largest empirical eigenvalue}
+where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
 The Weighted Alpha Metric was introduced previously~\cite{MM20_SDM}, where (on a much smaller set of data than we consider here) it was shown to correlate well with the trends in reported test accuracies of pretrained DNNs.

### 2020-02-01 nlp_models.tex
minor changes, starting nlp

@@
 
-\michael{Need to say that such large $\alpha$ values don't mean much.}
+In this section, we examine the quality metrics described in Section~\ref{sxn:methods} for several Natural Language Processing (NLP) model architectures.
+NLP models are different in many ways than the CV models we considered in Section~\ref{sxn:cv}.
+In particular, within the past two years, nearly 100 open source, pre-trained DNNs for NLP have emerged, based on the revolutionary Transformer architecture.
+This includes variants of BERT, Transformer-XML, GPT, etc.
+The Transformer architectures consist of blocks of so-called Attention layers, containing two large, Feed Forward (Linear) weight matrics~\cite{Attn2017}. 
+In contrast to the smaller pre-Activation maps arising in Cond2D layers, Attention matrices are significantly larger, 
+and frequently under-correlated, with generally larger Heavy Tailed Power Law exponents $\alpha$ (when fitting the ESD).
 
-\michael{Need to highlight difference between $\alpha$ and $\hat{\alpha}$.}
+Nevertheless, norm-base metrics perform reasonably well on good NLP models, but they have problems with less well-trained models.
 
-In this section, we examine the quality metrics described in Section~\ref{sxn:methods} for several NLP model architectures.
+Here, PL-based metrics 
+(the Log $\alpha$-Norm metric, $\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$, and the Weighted Alpha metric, $\hat\alpha =\alpha\log\lambda_{max} $)
 
-apply more generally, where they differ, e.g., how the layer Spectral Norm behaves unexpectedly, while the $\alpha$-based metrics can give insight into how well trained a model is.
+we breifly look at a few popular pretrained NLP DNNs to demonstrate that HT-SR Theory 
+also
+applies to NLP models and not just CV models, and to highlight how to use the theory to identify
+poorly trained models where the empirical norm metrics will perform poorly.
 
@@
 
-Within the past two years, nearly 100 open source, pre-trained DNNs for NLP have emerged, based on the revolutionary Transformer architecture.
-This includes variants of BERT, Transformer-XML, GPT, etc.
-The Transformer architectures consist of blocks of Attention layers, containing two large, Feed Forward (Linear) weight matrics~\cite{Attn2017}. 
-In contrast to the smaller pre-Activation maps arising in Cond2D layers, Attention matrices are significantly larger, 
-and frequently under-correlated, with generally larger Heavy Tailed Power Law exponents $\alpha$ (when fitting the ESD).  
-Here, we breifly look at a few popular pretrained NLP DNNs to demonstrate that the Theory of Heavy Tailed Self-Regularization also
-applies to NLP models and not just CV models, and to highlight how to use the theory to identify
-poorly trained models where the empirical norm metrics will perform poorly.
+\michael{Need to say that such large $\alpha$ values don't mean much.}
+
+\michael{Need to highlight difference between $\alpha$ and $\hat{\alpha}$.}
+
+apply more generally, where they differ, e.g., how the layer Spectral Norm behaves unexpectedly, while the $\alpha$-based metrics can give insight into how well trained a model is.
 

### 2020-02-01 nlp_models.tex
some progress on nlp section

@@
 In this section, we examine the quality metrics described in Section~\ref{sxn:methods} for several Natural Language Processing (NLP) model architectures.
-NLP models are different in many ways than the CV models we considered in Section~\ref{sxn:cv}.
+NLP models are different in many ways than the CV models we considered in Section~\ref{sxn:cv}, and thus they should provide a very different test of norm based metrics and HT-SR metrics than do CV models.
+
 In particular, within the past two years, nearly 100 open source, pre-trained DNNs for NLP have emerged, based on the revolutionary Transformer architecture.
@@
 
-We use the \emph{WeightWatcher} tool to analyze the OpenAI GPT and GPT2 models, which gives us
-the opportunity to analyze the effect of both training the same model with different size data sets,
-and increasing sizes of both the data set and architectures.
-These models have generated significant media attention because of their remarkable ability to
- generate fake text that appears to be real and the potential misuse of this.
-For this reason, the original GPT model was trained on on a deficient data set, rendering
-the model interesting but not fully functional.  Later, OpenAI released a much improved model--
-GPT2 (small)--which has the same architecture and number of layers as GPT, but has
-been trained on a larger and better data set (and with other changes), making it
-remarkably good at generating (near) human-quality fake text.  
-By comparing the poorly trained GPT to the well trained GPT2, we
-indentify empirical indidcators for when  model has in fact
-been poorly trained and may perform poorly when deployed.
+The OpenAI GPT and GPT2 models provide us with the opportunity to analyze two effects: training the same model with different data set sizes; and increasing sizes of both the data set and architectures.
+These models have the remarkable ability to generate fake text that appears to the human to be real, and they have generated significant media attention because of the potential for their misuse.
+For this reason, the original GPT model released by OpenAI was trained on on a deficient data set, rendering the model interesting but not fully functional.  
+Later, OpenAI released a much improved model, GPT2 (small), which has the same architecture and number of layers as GPT, but which has been trained on a larger and better data set (and with other changes), making it remarkably good at generating (near) human-quality fake text.
+%
+By comparing the poorly-trained GPT to the well-trained GPT2, we can indentify empirical indidcators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
 
-\charles{more here ?}
-We analyze GPT models deployed with the popular HuggingFace PyTorch library.
+\charles{More details here.}
+The GPT models we analyze are deployed with the popular HuggingFace PyTorch library.
 GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ Layer Weight Matrices $\mathbf{W}$.
@@
 This gives $50$ large matrices.
+\michael{Fix that.}
 
 Because GPT and GPT are trained on different data sets, the initial Embedding matrices differ in shape.
-GPT  has an initial Token and Positional Embedding layers, of dimension
-$(40478\times 768)$ and $(512\times 768)$, resp, whereas GPT2 has input Embeddings of shape
-$(50257\times 768)$ and $(1024\times 768)$, resp.  Interstingly, they also have very spectral properties,
-also shown below.
+GPT  has an initial Token and Positional Embedding layers, of dimension $(40478\times 768)$ and $(512\times 768)$, respectively, whereas GPT2 has input Embeddings of shape $(50257\times 768)$ and $(1024\times 768)$, respectively.
+Interstingly, they also have very spectral properties, also shown below. XXX.  WHICH.  MOVE.
+\michael{Fix that.}
 
@@
 The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
-Table \ref{table:nlp} reports results for the average log norm metrics, using \emph{weightwatcher (0.2.7)},
+Table \ref{table:nlp} reports results for the average log norm metrics and the Weighted Alpha metric, using \emph{weightwatcher (0.2.7)},
 and with fully reproducible Jupyter notebooks.\footnote{\url{https://github.com/CalculatedContent/kdd2020}}
+\michael{Fix that.}
 
@@
 
-\paragraph{Empirical Quality Metrics for GPT and GPT2.}
+\paragraph{Coarse Analysis: Empirical Quality Metrics for GPT and GPT2.}
 
-We first compare the distribution of  Heavy Tailed Power Law exponents $\alpha$ in GPT and GPT2.
-They are very differenmt, with GPT2 having both a notably smaller mean $\alpha$, and far fewer, unusually large outliers.
-Figure \ref{fig:gpt-alphs-hist} shows the empirical density (histogram) of $\alpha$
-for all layers in GPT (blue) and GPT2 (red).  \charles{discuss more}
+We have analyzed the four quality metrics described in Section~\ref{sxn:methods} for the OpenAI GPT and GPT2 pretrained models.
+We start by examining the PL exponents $\alpha$ for GPT and GPT2 (small).
+Figure~\ref{fig:GPT-alpha-hist} shows the histogram (empirical density), for all layers, of $\alpha$ for GPT (blue) and this GPT2 (red) model.  
+These two histograms are very different, with GPT2 having both a notably smaller mean $\alpha$, and far fewer unusually-large outlying $\alpha$ values.
+
+Observe that these results indicate that $\alpha$ provides a good quality metric for comparing these two models, one of which which was designed to be of low quality.
+The deficient GPT model has numerous weight matrices with unsually large fitted exponents (indicating that they are not HT at all).  
+Indeed, we may expect that a poorly trained model will not exhibit consistent Heavy Tailed behavior in all layers.
+On the other hand, as expected, the improved GPT2 model has, on average, smaller $\alpha$ than the older GPT, with all $\alpha\le6$.  
+\michael{XXX.  FLESH OUT.}
 
@@
         %\includegraphics[width=5cm]{img/GPT-alpha-hist.png}
-        \includegraphics[width=3.7cm]{img/GPT-alpha-hist.png}
+        \includegraphics[width=4.0cm]{img/GPT-alpha-hist.png}
         \label{fig:GPT-alpha-hist}
        }
-    \qquad
+    %\qquad
     \subfigure[log Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
         %\includegraphics[width=5cm]{img/GPT-snorm-hist.png}
-        \includegraphics[width=3.7cm]{img/GPT-snorm-hist.png}
+        \includegraphics[width=4.0cm]{img/GPT-snorm-hist.png}
         \label{fig:GPT-snorm-hist}
        }
-
-   \caption{Comparison of Heavy Tailed Power Law exponents $\alpha$, and log Spectral Norms $\log\Vert\mathbf{W}\Vert_{\infty}$
-for the OpenAI GPT and GPT2 (small) pretrained models.}
-
+   \caption{Comparison of PL exponents, $\alpha$, and log Spectral Norms, $\log\Vert\mathbf{W}\Vert_{\infty}$, for the OpenAI GPT and GPT2 (small) pretrained models.}
+\label{fig:GPT-hist}
 \end{figure}
-
-Indeed, the $\alpha$ metric makes a good quality metric for comparing these models.
-Figure \ref{fig:GPT-alpha-hist} compares the PL exponent $\alpha$ for the GPT and GPT2 models for all layers,
-As expected, the improved GPT2 model has, on average, smaller $\alpha$ than the older GPT, with all
-$\alpha\le6$.  Indeed, the deficient GPT model has numerous weight matrices with unsually large fitted exponents,
-indicating that they are not Heavy Tailed at all.  Indeed, we may expect that a poorly trained model
-will not exhibit consistent Heavy Tailed behavior in all layers.
 
@@
 
-\paragraph{Correlation Flow in GPT and GPT2.}
+\paragraph{Layer Analysis: Correlation Flow in GPT and GPT2.}
 

### 2020-02-01 abstract_file.tex
early sections and move stuff to conclusion as a placeholder for now

@@
-In many applications,
-% of machine learning and artificial intelligence, 
-one must work with models that have been trained by someone else.
+In many applications, one must work with models that have been trained by someone else.
 For such \emph{pretrained models}, one typically does not have access to training data or test data, and one does not know the details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
@@
 %%Our metrics are drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as Heavy-Tailed Random Matrix Theory (HT-RMT), as used in the recently-developed Theory of Heavy-Tailed Self Regularization (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models).
-The most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in a system) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
+%%The most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in a system) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
+The most promising are norm-based metrics (such as those used to provide capacity control in traditional statistical learning theory) and metrics based on fitting eigenvalue distributions to truncated power law (PL) distributions (which derive from statistical mechanics, in particular the recently-developed Theory of Heavy-Tailed Self Regularization, and which can be used to characterize the degree of strong correlations in a system).
+%
 Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision and natural language processing.
 We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best.'' 
-On the other hand, for models that may not be well-trained (which, arguably is the point of needing metrics to evaluate the quality of pretrained models), i.e., when we want to distinguish ``good-bad,'' norm-based metrics can qualitatively fail.
-We also find that HT-SR metrics do much better, quantitatively better at discriminating good-better-best and qualitatively better at discriminating good-bad.
-HT-SR metrics can also be used to characterize fine-scale properties of models, e.g., understanding layer-wise \emph{correlation flow}, and evaluate post-training modifications such as model distillation and model improvement.
+On the other hand, for models that may not be well-trained (which, arguably is the point of needing metrics to evaluate the quality of pretrained models), i.e., when we want to distinguish ``good-vs-bad,'' norm-based metrics can qualitatively fail.
+We also find that PL-based metrics do much better, quantitatively better at discriminating good-better-best and qualitatively better at discriminating good-vs-bad.
+PL-based metrics can also be used to characterize fine-scale properties of models, e.g., understanding layer-wise \emph{correlation flow}, and evaluate post-training modifications such as model distillation and model improvement.

### 2020-02-01 background.tex
early sections and move stuff to conclusion as a placeholder for now

@@
 
-To our knowledge, there is very little work on the particular question we are addressing: namely, how to predict, in a theoretically-principled manner, the quality of large-scale state-of-the-art NNs, and to do so without access to training data or testing data or details of the training protocol, etc.
-Our work is, however, loosely related to several other lines of work, and we briefly discuss them here.
+Most theory for DNNs is applied to small toy models and assumes access to data.
+There is very little work asking how to predict, in a theoretically-principled manner, the quality of large-scale state-of-the-art NNs, and how to do so without access to training data or testing data or details of the training protocol, etc.
+Our work is, however, loosely related to several other lines of work.
 
@@
 Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statisical mechanics of heavy tailed and strongly correlated systems.
-
-XXX.
-Distinguish between what we will call a
-\emph{phenomenological theory}
-(that describes empirical relationship of phenomena to each other, in a way which is consistent with fundamental theory, but is not directly derived from that theory)
-and what can be called a 
-\emph{first principles theory} 
-(that is applicable to tiny things but has no hope of scaling up).
-\charles{it is the oppposite...ab initio theory scales quite well...it is the spherical cow models of physics and ML that do not scale.
-What we have is a semi-empirical theory.
-We use real theory, but require empirical input, at least in the new stat mech work.
-What we have introduced is a phenomenology, which to me is different from a semi-empirical or phenomenological theory  }
-\michael{Let's touch base to get this right.}
-\footnote{In most areas where there are complex highly-engineered systems (beyond complex AI/ML systems), one used phenomenological theory rather than first principles theory.
-For example, one does not try to solve the Schr\"odinger equation if one is interested in building a bridge or an airplane.}
 
@@
 \textbf{Simplicity, or lack of,  accuracy metrics.}
-XXX.
+\michael{XXX.  Put in two or at most three sentences.}
 \item
 \textbf{Information leakage in the production pipeline.}
-XXX.
+\michael{XXX.  Put in two or at most three sentences.}
 \item
 \textbf{Cost of acquring labeled data.}
-XXX.
+\michael{XXX.  Put in two or at most three sentences.}
 \end{itemize}

### 2020-02-01 conclusion.tex
mostly section 3, but touched every section
early sections and move stuff to conclusion as a placeholder for now

@@
 \label{sxn:conc}
+
+We should point out that it is not obvious that norm-based metrics will perform well \emph{across} models, as they are usually used \emph{within} a given model or parameterized model class.
+That PL-based metrics perform better should not be surprising---while norms are very coarse metrics of the properties of a matrix,
+
+That being said, our goal is not to make a statement about every publicly-available model.
+Ther are exceptions to the general trends we state.
+Several of these are described below, and others are are available in what we release.
+
+\michael{MENTION:} 
+Have some sort of caveat about how we don't solve every problem in the world, in particular there are exceptions to ghe general trends we note, but that all of our work is publicly available in Weightwatcher and Google Colab, so people can reproduce.
+\michael{XXX.  FEW SENTENCES ABOUT REPRODUCIBILITY MORE GENERALLY HERE, HERE OR IN DISCUSSION/CONCLUSION.}
+
+\michael{MENTION:} 
+PL-based metrics have been used to characterize the degree of strong correlations in a system, and they are derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
+AND DEFINE HT-SR ACRONYM.
+
+\michael{
+Where, if anywhere to put the following:
+\begin{itemize}
+\item MOVE TO LATER: COMMENT ON HOW LOG NORM first and last layers behave, maybe somewhere else.
+\item MOVE TO LATER: COMMENT ON HOW LOG PORM  for GPT includes unusually high alpha, not meaningful other than to show the trend.
+\end{itemize}
+}
+
+XXX.  THE FOLLOWING IS PROBABLY BETTER HERE.
+Distinguish between what we will call a
+\emph{phenomenological theory}
+(that describes empirical relationship of phenomena to each other, in a way which is consistent with fundamental theory, but is not directly derived from that theory)
+and what can be called a 
+\emph{first principles theory} 
+(that is applicable to tiny things but has no hope of scaling up).
+\charles{it is the oppposite...ab initio theory scales quite well...it is the spherical cow models of physics and ML that do not scale.
+What we have is a semi-empirical theory.
+We use real theory, but require empirical input, at least in the new stat mech work.
+What we have introduced is a phenomenology, which to me is different from a semi-empirical or phenomenological theory  }
+\michael{Let's touch base to get this right.}
+\footnote{In most areas where there are complex highly-engineered systems (beyond complex AI/ML systems), one used phenomenological theory rather than first principles theory.
+For example, one does not try to solve the Schr\"odinger equation if one is interested in building a bridge or an airplane.}
 

### 2020-02-01 introduction.tex
early sections and move stuff to conclusion as a placeholder for now

--- 2020-01-31 introduction.tex
@@
 \michael{XXX.  LIST PLACES WHERE THEY ARE AVAIALBLE, HERE OR IN NEXT SECTION.}
-Importantly, all of these models have been trained by someone else and have been viewed to be of sufficient interest/quality to be made publicly-available, and for all of these models, we have no access to training data or testing data, and we have no knowledge of the training/testing protocols.
+Importantly, all of these models have been trained by someone else and have been viewed to be of sufficient interest/quality to be made publicly-available; and, for all of these models, we have no access to training data or testing data, and we have no knowledge of the training/testing protocols.
 
 The \emph{quality metrics} we consider are based on the spectral properties of the layer weight matrics.
-Note that unlike traditional ML approaches, however, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., hyperparmeter optimization).%
-\footnote{We reiterate: One could use these techniques to improve training, and we have been asked about that, but we are not interested in that here.
+In particular, they are based on the norms of weight matrices (such norms have been used in traditional statistical learning theory to bound capacity and construct regularizers) and parameters of power law (PL) fits of the eigenvalues of weight matrices (such PL fits are based on statistical mechanics approaches to NNs).
+Note that, while we use traditional norm-based and PL-based metrics, our goals are different.
+In particular, unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., hyperparmeter optimization).%
+\footnote{One could of course use these techniques to improve training, and we have been asked about that, but we are not interested in that here.
 Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained NN models.}
@@
 In more detail, our main contributions are the following.
-\begin{itemize}
-\item \michael{XXX TECHNICAL THING 1}
-\item \michael{XXX TECHNICAL THING 2}
-\item \michael{XXX TECHNICAL THING 3}
-\item \michael{XXX TECHNICAL THING 4}
-\end{itemize}
+%\begin{itemize}
+%\item 
+First, 
+motivated by practical problems, we formulate the question ``Can one predict trends in the quality of state-of-the-art neural networks without access to training or testing data?''
+%\item
+Second, 
+to answer this question, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in CV and NLP.
+%\item 
+Third, 
+we find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best'' models, but they can qualitatively fail for models that may not be well-trained, i.e., to distinguish ``good-vs-bad'' models.
+%\item 
+Fourth, 
+we find that PL-based metrics do much better, quantitatively better at discriminating good-better-best and qualitatively better at discriminating good-vs-bad.
+%\item 
+Finally, 
+we find that PL-based metrics can also be used to characterize fine-scale properties of models, including understanding layer-wise \emph{correlation flow}, and evaluating post-training modifications such as model distillation and model improvement.
+%\end{itemize}
 
@@
 Finally, in Section~\ref{sxn:conc}, we provide a brief discussion and conclusion.
+
+\paragraph{The WeightWatcher Tool.}
+
+All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package};
+and, in order to make our results fully reproducible, we provide Jupyter notebooks in the github repo for this paper~\cite{repo}.
+\michael{LINK TO: Google Colab, IS THIS THAT REPO.}

### 2020-02-01 methods.tex
early sections and move stuff to conclusion as a placeholder for now

@@
 
-Let us write the Energy Landscape (or optimization function, parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a typical DNN with $L$ layers, with activation functions $h_{l}(\cdot)$, and with $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$, as:
+Let us 
+write the Energy Landscape (or optimization function, parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$,~as:
 \begin{equation}
@@
 \end{equation}
-We assume we are given several pretrained Deep Neural Networks (DNNs), e.g., as part of an architecture series.
+Assume we are given several pretrained DNNs, e.g., as part of an architecture series.
 The models have been trained on (unspecified and unavailable) labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using Backprop, by minimizing some (also unspecified and unavailable) loss function $\mathcal{L}(\cdot)$.
@@
 (These eigenvalues are the squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, i.e., $\lambda_{i}=\sigma^{2}_{i}$.)
-For the last two metrics, the exponent $\alpha$ is the power law exponent that arises in the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
-Operationally, $\alpha$ is determined by using the publicly-available \emph{WeightWatcher} tool (\cite{weightwatcher_package}) to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated power law
+For the last two metrics, $\alpha$ is the PL exponent that arises in the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+Operationally, $\alpha$ is determined by using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated PL,
 \begin{equation}
@@
 The only exception to this is that we do rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $k/\sqrt{2}$ so that they are on the same scale as the Linear layers.
-
-\paragraph{The WeightWatcher Tool.}
-
-We compute the metrics using the WeightWatcher tool (version 0.2.7), and we provide Jupyter notebooks in the github repo for this paper~\cite{repo}, making the results fully reproducible.
-\michael{XXX.  FEW SENTENCES ABOUT REPRODUCIBILITY MORE GENERALLY HERE, HERE OR IN DISCUSSION/CONCLUSION.}

### 2020-02-01 background.tex
some more misc changes, still to do cv and nlp sections

@@
 our best-performing metrics (i.e., those using fitted PL exponents and HT-SR Theory) are based on statistical mechanics ideas~\cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM19_KDD, MM20_SDM}.
+\michael{Where first define HT-SR.}
 However, that the way in which we \emph{use} statistical mechanics theory is quite different than the way it is more commonly formulated.

### 2020-02-01 cv_models.tex
some more misc changes, still to do cv and nlp sections

@@
 
-In this section, we examine quality metrics described in Section~\ref{sxn:methods} for several Computer Vision (CV) model architecture series.
+In this section, we examine quality metrics described in Section~\ref{sxn:methods} for several CV model architecture series.
 This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current opensource pyTorch framework (version 1.4)~\cite{pyTorch}.
@@
 
-We use our quality metrics to perform \emph{coarse model analysis}, comparing and contrasting the three model series (or four, if ResNet-1K is included), permitting us to predict trends in model quality.
-We also use perform a more \emph{fine layer analysis} as a function of depth for these models, providing novel insights among the VGG, ResNet, and DenseNet architectures.
+We perform \emph{coarse model analysis}, comparing and contrasting the four model series, and predicting trends in model quality. 
+We also perform \emph{fine layer analysis}, as a function of depth for these models, illustrating that PL-based metrics can provide novel insights among the VGG, ResNet/ResNet-1K, and DenseNet architectures.
 
@@
 
-We have examined the performance of the four quality metrics (Frobenius norm, Spectral norm, Weighted Alpha, and $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
+We have examined the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
 To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BatchNormalization), and it plots the four quality metrics versus the reported Test accuracies~\cite{pyTorchVgg}, as well as a basic linear regression line.

### 2020-02-01 methods.tex
some more misc changes, still to do cv and nlp sections

@@
 \footnote{We do not use intra-layer information from the models in our quality metrics, but as we will describe our metrics can be used to learn about that.}
-We consider the following metrics.
+Consider the following.
 \begin{itemize}
@@
 \end{itemize}
-The first two metrics are well-known in ML.
+The first two quantities are well-known in ML.
 The last two deserve special mention.
-For all these metrics, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the \emph{Empirical Correlation Matrix},
+For all these quantities, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the \emph{Empirical Correlation Matrix},
 $ %% $$
@@
 (These eigenvalues are the squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, i.e., $\lambda_{i}=\sigma^{2}_{i}$.)
-For the last two metrics, $\alpha$ is the PL exponent that arises in the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+For the last two quantities, $\alpha$ is the PL exponent that arises in the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
 Operationally, $\alpha$ is determined by using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated PL,
@@
 where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
-The Weighted Alpha Metric was introduced previously~\cite{MM20_SDM}, where (on a much smaller set of data than we consider here) it was shown to correlate well with the trends in reported test accuracies of pretrained DNNs.
-Based on this, here we introduce and evaluate the $\alpha$-Norm metric.
-One would expect $\hat{\alpha}$ to approximate the log $\alpha$-Norm very well for $\alpha < 2$ and reasoably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.
+Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.
 
-Each of these metrics is defined for a given layer $\mathbf{W}$ matrix.
-For the Weighted Alpha Metric, we consider the average of $\hat{\alpha}$ over all layers, i.e., $\sum_l \hat{\alpha}_l = \sum_l \alpha_l\log\lambda_{max,l} $ .  
-\michael{Be careful to get that right.}
-For the norm-based metrics, we consider the \emph{weighted} average of the log of the norm to the appropriate power.
-For example, for the $\alpha$-Norm, we compute 
+For norm-based metrics, we consider the average log norm to the appropriate power.
+For example, for the \emph{$\alpha$-Norm metric}, 
 \begin{equation}
@@
 where $\Vert\cdot\Vert$ is a matrix norm.   
-In this case, the log Complexity takes the form of an Average Log Norm,
+The log Complexity,
 \begin{equation}
 \label{eqn:eqn:sum_log_norm}
-\log\mathcal{C} \sim \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert + \cdots + \log\Vert\mathbf{W}_{L}\Vert  .
+\log\mathcal{C} \sim \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert + \cdots + \log\Vert\mathbf{W}_{L}\Vert  ,
 \end{equation}
-For the Frobenius and Spectral norms, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly. 
-For example, when taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.
+ takes the form of an Average Log Norm.
+For the \emph{Frobenius norm metric} and \emph{Spectral norm metric}, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly. 
+(When taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.)
 For the Weighted Alpha Norm, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it cannot be taken out of the sum.
+
+For the Weighted Alpha Metric, we consider the average of $\hat{\alpha}$ over all layers, i.e., 
+$ %$$
+\sum_l \hat{\alpha}_l = \sum_l \alpha_l\log\lambda_{max,l}  .  
+$ %$$
+\michael{Be careful to get that right.}
+The Weighted Alpha Metric was introduced previously~\cite{MM20_SDM}, where (on a much more limited set of data than we consider here) it was shown to correlate well with the trends in reported test accuracies of pretrained DNNs.
+Based on this, in this paper, we introduce and evaluate the $\alpha$-Norm metric.
+One would expect $\hat{\alpha}$ to approximate the log $\alpha$-Norm very well for $\alpha < 2$ and reasoably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.
 

### 2020-02-01 background.tex
more small changes, still just starting those two sections

--- 2020-02-01 background.tex
@@
 
-\paragraph{Statistical mechanical theory of NNs.}
+\paragraph{Statistical mechanics theory for NNs.}
 
-Statistical mechanical ideas have long had influence on NN theory and practice~\cite{EB01_BOOK, MM17_TR, BKPx20}; and 
-our best-performing metrics (i.e., those using fitted PL exponents and HT-SR Theory) are based on statistical mechanics ideas~\cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM19_KDD, MM20_SDM}.
-However, that the way in which we \emph{use} statistical mechanics theory is quite different than the way it is more commonly formulated.
+Statistical mechanics ideas have long had influence on NN theory and practice~\cite{EB01_BOOK, MM17_TR, BKPx20}; and 
+our best-performing metrics (those using fitted PL exponents) are based on statistical mechanics~\cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM19_KDD, MM20_SDM}, in particular the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+However, we emphasize that the way in which we \emph{use} statistical mechanics theory is quite different than the way it is more commonly formulated.
 Several very good overviews of the more common approach are available~\cite{EB01_BOOK, BKPx20}.

### 2020-02-01 cv_models.tex
more small changes, still just starting those two sections

@@
 % Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
-        & Num.   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+        &        & Log                   & Log                        & Weighted       & Log                                 \\
+        & Num.   & $\Vert\cdot\Vert_{F}$ & $\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
  Series & Models & Metric                & Metric                     & Metric         & Metric                              \\

### 2020-02-01 introduction.tex
more small changes, still just starting those two sections

@@
 All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package};
-and, in order to make our results fully reproducible, we provide Jupyter notebooks in the github repo for this paper~\cite{repo}.
+and, in order to make our results fully reproducible, we provide Jupyter notebooks in the github repo for this paper~\cite{repo},
+with fully reproducible Jupyter notebooks.%
+\footnote{\url{https://github.com/CalculatedContent/kdd2020} \michael{DIFFERENT NAME?}}
 \michael{LINK TO: Google Colab, IS THIS THAT REPO.}
+\michael{FIX THIS PAR.}

### 2020-02-01 methods.tex
more small changes, still just starting those two sections

@@
 (These eigenvalues are the squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, i.e., $\lambda_{i}=\sigma^{2}_{i}$.)
-For the last two quantities, $\alpha$ is the PL exponent that arises in the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+For the last two quantities, $\alpha$ is the PL exponent that arises in the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
 Operationally, $\alpha$ is determined by using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated PL,
@@
 
-For norm-based metrics, we consider the average log norm to the appropriate power.
-For example, for the \emph{$\alpha$-Norm metric}, 
+For norm-based metrics, we use the average of the log norm to the appropriate power.
+For example, consider the \emph{$\alpha$-Norm metric}, 
 \begin{equation}
@@
 Typically, to apply RMT, we would usually define Correlation Matrix with $1/N$ normalization and assume that the variance of $\mathbf{X}$ is either unity or a known constant.% 
-\footnote{For formal proofs of Heavy Tailed results, one typically needs a different normalization such as $1/N^{1-\alpha}$.}
+\footnote{For formal proofs of Heavy Tailed results, one typically needs a normalization such as $1/N^{1-\alpha}$.}
 Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with the variance already normalized to $\sqrt{1/N}$, or some variant of this, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.

### 2020-02-01 nlp_models.tex
more small changes, still just starting those two sections

@@
 
-In this section, we examine the quality metrics described in Section~\ref{sxn:methods} for several Natural Language Processing (NLP) model architectures.
-NLP models are different in many ways than the CV models we considered in Section~\ref{sxn:cv}, and thus they should provide a very different test of norm based metrics and HT-SR metrics than do CV models.
+In this section, we examine quality metrics described in Section~\ref{sxn:methods} for several NLP model architectures.
+%
+%In particular, 
+Within the past two years, nearly 100 open source, pre-trained NLP DNNs based on the revolutionary Transformer architecture have emerged.
+These include variants of BERT, Transformer-XML, GPT, etc.
+%
+The Transformer architectures consist of blocks of so-called Attention layers, containing two large, Feed Forward (Linear) weight matrics~\cite{Attn2017}. 
+In contrast to smaller pre-Activation maps arising in Cond2D layers, Attention matrices are significantly larger.
+In general, we have found that they have larger $\alpha$ PL exponents.
+Based on HT-SR Theory (in particular, the interpretation of values of $\alpha \sim 2$ as modeling systems with good correlations over many size scales~\cite{BouchaudPotters03, SornetteBook}), this suggests that these models fail to capture successfully many of the correlations in the data (relative to their size) and thus that they are substantially \emph{under}-trained.
+%
+More generally, these NLP models are quite different than the CV models we considered in Section~\ref{sxn:cv}, both in terms of their architecture (e.g., wheter there are convolutions) as well as how they are used (e.g., whether one uses prediction error or perplexity).
+Thus, they provide a very different test for our metrics.
 
-In particular, within the past two years, nearly 100 open source, pre-trained DNNs for NLP have emerged, based on the revolutionary Transformer architecture.
-This includes variants of BERT, Transformer-XML, GPT, etc.
-The Transformer architectures consist of blocks of so-called Attention layers, containing two large, Feed Forward (Linear) weight matrics~\cite{Attn2017}. 
-In contrast to the smaller pre-Activation maps arising in Cond2D layers, Attention matrices are significantly larger, 
-and frequently under-correlated, with generally larger Heavy Tailed Power Law exponents $\alpha$ (when fitting the ESD).
+\michael{WHERE: Norm-base metrics perform reasonably well on good NLP models, but they have problems with less well-trained models.
+PL-based metrics (the Log $\alpha$-Norm metric, $\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$, and the Weighted Alpha metric, $\hat\alpha =\alpha\log\lambda_{max} $) do even better.
+We highlight how to use the theory to identify poorly trained models where the empirical norm metrics will perform poorly.}
 
-Nevertheless, norm-base metrics perform reasonably well on good NLP models, but they have problems with less well-trained models.
-
-Here, PL-based metrics 
-(the Log $\alpha$-Norm metric, $\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$, and the Weighted Alpha metric, $\hat\alpha =\alpha\log\lambda_{max} $)
-
-we breifly look at a few popular pretrained NLP DNNs to demonstrate that HT-SR Theory 
-also
-applies to NLP models and not just CV models, and to highlight how to use the theory to identify
-poorly trained models where the empirical norm metrics will perform poorly.
-
-\michael{We show that this not only works for CV, so we look at NLP.
+\michael{WHERE: We show that this not only works for CV, so we look at NLP.
 We now look in detail at the alpha metric, the spectral norm, and the new norm.
 We compare good and bad models, and a series of increasingly big models
-trained on big data sets.  We show why alpha works but spectral norm is not enough
+trained on big data sets.  We show why alpha works but spectral norm is not enough.
+The layer Spectral Norm behaves unexpectedly, while $\alpha$-based metrics can give insight into how well trained a model is.}
 
-}
+\michael{WHERE: Say large $\alpha$ values don't mean much.}
 
-\michael{Need to say that such large $\alpha$ values don't mean much.}
-
-\michael{Need to highlight difference between $\alpha$ and $\hat{\alpha}$.}
-
-apply more generally, where they differ, e.g., how the layer Spectral Norm behaves unexpectedly, while the $\alpha$-based metrics can give insight into how well trained a model is.
+\michael{WHERE: Highlight difference between $\alpha$ and $\hat{\alpha}$.}
 
@@
 \charles{More details here.}
-The GPT models we analyze are deployed with the popular HuggingFace PyTorch library.
+The GPT models we analyze are deployed with the popular HuggingFace PyTorch library~\cite{XXX-XXX}.
 GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ Layer Weight Matrices $\mathbf{W}$.
@@
 This gives $50$ large matrices.
-\michael{Fix that.}
-
+%
 Because GPT and GPT are trained on different data sets, the initial Embedding matrices differ in shape.
 GPT  has an initial Token and Positional Embedding layers, of dimension $(40478\times 768)$ and $(512\times 768)$, respectively, whereas GPT2 has input Embeddings of shape $(50257\times 768)$ and $(1024\times 768)$, respectively.
-Interstingly, they also have very spectral properties, also shown below. XXX.  WHICH.  MOVE.
-\michael{Fix that.}
-
+%
 The OpenAI GPT2 (English) models are: \nred{gpt-small, gpt-medium, gpt-large, and gpt-xl}, \
-having include $12, 24, 36, \text{and }48$ layers, resp., with increasingly larger weight matrices.
+having include $12, 24, 36, \text{and }48$ layers, respectively, with increasingly larger weight matrices.
 The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
-Table \ref{table:nlp} reports results for the average log norm metrics and the Weighted Alpha metric, using \emph{weightwatcher (0.2.7)},
-and with fully reproducible Jupyter notebooks.\footnote{\url{https://github.com/CalculatedContent/kdd2020}}
+Table \ref{table:nlp} reports results for the average log norm metrics and the Weighted Alpha metric, 
 \michael{Fix that.}
@@
 % Series & \#Layers   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
-        & Num.   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
+        &        & Log                   & Log                        & Weighted       & Log                                 \\
+        & Num.   & $\Vert\cdot\Vert_{F}$ & $\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
  Series & Layers & Metric                & Metric                     & Metric         & Metric                              \\

### 2020-02-02 all_models.tex
more edits
pass over most sections, more to do on sec 5 and 6

@@
 
+\charles{This section needs a lot of work}
+
 \serena{insert link in the footnote}
@@
 We then regress them individually on each of the Weightwatcher metrics, including  alpha, alpha_weighted, spectralnormlog, softranklog, lognorm, and logpnorm as defined in earlier section.
-
-\serena{Question: only the alpha-related metrics is the core metrics of the WW tool, right? Metrics such as stable rank or spectral norm have been suggested in other paper before???}
-
-\serena{insert a simple formula for linear regression}
 
@@
 \begin{center}
-\begin{tabular}{|p{3in}|c|}
+\begin{tabular}{|p{2in}|c|}
 \hline
@@
 \begin{center}
-\begin{tabular}{|p{1in}|c|c|c|c|}
+%\begin{tabular}{|p{1in}|c|c|c|c|}
+\begin{tabular}{|p{0.75in}|c|c|c|c|}
 \hline
-    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
-    & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\langle\hat{\alpha}=\alpha\log\lambda_{max}\rangle$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
+%    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
+%    & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\langle\hat{\alpha}=\alpha\log\lambda_{max}\rangle$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
+        & Log                   & Log                        & Weighted       & Log                                 \\
+        & $\Vert\cdot\Vert_{F}$ & $\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
+ Series & Metric                & Metric                     & Metric         & Metric                              \\
 \hline

### 2020-02-02 appendix.tex
pass over most sections, more to do on sec 5 and 6

@@
 
-XXX.  APPENDIX.
+\subsection{XXX}
 
@@
 Figure~\ref{fig:alexnet1} displays the ESD for the first slice (or matrix $\mathbf{W}$) of the third Conv2D layer, extracted from a 4-index Tensor of shape $(384, 192, 3, 3)$.
-The red line displays the best fit to a random matrix, using the Marchenko pastur theory~\cite{MM}.
+The red line displays the best fit to a random matrix, using the Marchenko pastur theory~\cite{MM18_TR}.
 We can see the random matrix model does not describe the ESD very well.
@@
 \michael{These comments seem out of place, since they hold more generally than for the Conv2D layers.}
+
+\subsection{XXX}
+
+Some other comments that we need to weave into a narrative eventually after later sections are written:
+\begin{itemize}
+\item
+GPT versus GPT2.
+What happens when we don't have enough data?
+This is the main question, and we can use out metrics to evaluate that, but we also get very different results for GPT versus GPT2.
+\item
+The spectral norm is a regularizer, used to distinguish good-better-best, not a quality metric.
+For example, it can ``collapse,'' and for bad models we can have small spectral norm.
+So, it isn't really a quality metric.
+\item
+One question that isn't obvious is whether regularization metrics can be used as quality metrics.
+One might think so, but the answer isn't obviously yes.
+We show that the answer is No.
+A regularizer is designed to select a unique solution from a non-unique good-better-best.
+Quality metrics can also distinguish good versus bad.
+\item
+(We should at least mention this is like the statistical thing where we evaluate which model is better, as oposed to asking if a given model is good, I forget the name of that.)
+\item
+There are cases where the model is bad but regularization metric doesn't tell you that.
+Quality should be correlated in an empirical way.
+Correlated with good-better-best; but also tell good-bad.
+\item
+Question: why not use regularier for quality?
+Answer: A regularizer selects from a given set of degenerate models one which is nice or unique.
+It doesn't tell good versus bad, i.e., whether that model class is any good.
+\item
+Thus, it isn't obvious that norm-based metrics should do well, and they don't in general.
+\item
+We give examples of all of these: bad data; defective data; and distill models in a bad way.
+(Of course, bad data means bad model, at least indirectly, since the quality of the data affects the properties of the model.)
+\item
+We can select a model and change it, i.e., we don't just do hyperparameter fiddling.
+\end{itemize}

### 2020-02-02 conclusion.tex
pass over most sections, more to do on sec 5 and 6

--- 2019-12-29 conclusion.tex
@@
 \label{sxn:conc}
-XXX.  CONCLUSION.
+
+We have developed (based on strong theory) and evaluated (on a large corpus of publicly-available pretrained models from CV and NLP) methods to predict trends in the quality of state-of-the-art neural networks---without access to training or testing data.
+Prior to our work, it was not obvious that norm-based metrics would perform well to predict trends \emph{across} models (as they are usually used \emph{within} a given model or parameterized model class, e.g., to bound generalization error or to construct regularizers), and our results are the first to demonstrate that they can be used for this important practical problem.
+That PL-based metrics perform better (than norm-based metrics) should not be surprising---at least to those familiar with the statisical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}, since our use of PL exponents is designed to capture the idea that well-trained models capture correlations over many size scales in the data), but again,  our results are the first to demonstrate this.
+It is gratifying, also, that PL exponents can be used to provide fine-scale insight such as rationalizing the flow of correlations through a network.
+
+We conclude with a few thoughts on what a \emph{practical theory} of NNs should look like.
+%
+We distinguish between what we will call a
+\emph{phenomenological theory}
+(that describes empirical relationship of phenomena to each other, in a way which is consistent with fundamental theory, but is not directly derived from that theory)
+and what can be called a 
+\emph{first principles theory} 
+(that is applicable to toy models, but that does not scale up to realistic systems if one includes realistic apsects of realistic systems).
+%
+For most complex highly-engineered systems (aside from complex AI/ML systems), one \emph{uses} phenomenological theory rather than first principles theory. 
+(One does not try to solve the Schr\"odinger equation if one is interested in building a bridge or flying an airplane.)
+%
+Our results, which are based on our \emph{use} of sophisticated statistical mechanics theory to solve an important practical NN problems, suggests that this approach should be of interest more generally for those interested in developing a practical NN theory.
+
+\charles{it is the oppposite...ab initio theory scales quite well...it is the spherical cow models of physics and ML that do not scale.
+What we have is a semi-empirical theory.
+We use real theory, but require empirical input, at least in the new stat mech work.
+What we have introduced is a phenomenology, which to me is different from a semi-empirical or phenomenological theory  }
+\michael{Let's touch base to get this right.}

### 2020-02-02 cv_models.tex
pass over most sections, more to do on sec 5 and 6

@@
         %\includegraphics[width=5cm]{img/VGG_lognorm_accs.png}
-        \includegraphics[width=3.7cm]{img/VGG_lognorm_accs.png}
+        \includegraphics[width=4.0cm]{img/VGG_lognorm_accs.png}
         \label{fig:vgg-fnorm}
@@
         %\includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
-        \includegraphics[width=3.7cm]{img/VGG_spectralnorm_accs.png}
+        \includegraphics[width=4.0cm]{img/VGG_spectralnorm_accs.png}
         \label{fig:vgg-snorm}
@@
         %\includegraphics[width=4.9cm]{img/VGG_alpha_weighted_accs.png}
-        \includegraphics[width=3.7cm]{img/VGG_alpha_weighted_accs.png}
+        \includegraphics[width=4.0cm]{img/VGG_alpha_weighted_accs.png}
         \label{fig:vgg-walpha}
@@
         %\includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
-        \includegraphics[width=3.7cm]{img/VGG_logpnorm_accs.png}
+        \includegraphics[width=4.0cm]{img/VGG_logpnorm_accs.png}
         \label{fig:vgg-pnorm}
@@
         %\includegraphics[width=4.2cm]{img/ResNet_logpnorm_accs.png}
-        \includegraphics[width=3.7cm]{img/ResNet_logpnorm_accs.png}
+        \includegraphics[width=4.0cm]{img/ResNet_logpnorm_accs.png}
         \label{fig:resnet-accuracy}
     }
-    \qquad
+    %\qquad
     \subfigure[ ResNet-1K, $\alpha$-Norm ]{
         %\includegraphics[width=4.5cm]{img/ResNet-1K_logpnorm_accs.png}
-        \includegraphics[width=3.7cm]{img/ResNet-1K_logpnorm_accs.png}
+        \includegraphics[width=4.0cm]{img/ResNet-1K_logpnorm_accs.png}
         \label{fig:resnet1k-accuracy}
     }
-    \qquad
+    %\qquad
     \subfigure[ DenseNet, $\alpha$-Norm ]{
         %\includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
-        \includegraphics[width=3.7cm]{img/DenseNet_logpnorm_accs.png}
+        \includegraphics[width=4.0cm]{img/DenseNet_logpnorm_accs.png}
         \label{fig:densenet-accuracy}
@@
 For example, we can 
-plot \nred{just} the PL exponent, $\alpha$,%
-\footnote{That is, here we consider just $\alpha$ for each layer, not $\hat{\alpha}$ or $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$.} 
+plot (just) the PL exponent, $\alpha$,%
+\footnote{That is, here we consider just $\alpha$ for each layer, i.e., not $\hat{\alpha}$ or $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$.} 
 as a function of depth.
 %
-See Figure~\ref{fig:vgg-alpha-layers}, which plots $\alpha$ for each layer (the first layer corresponds to data, the last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (without BatchNormalization), ResNet-1K, and DenseNet series.
+See Figure~\ref{fig:3models-alpha-layers}, which plots $\alpha$ for each layer (the first layer corresponds to data, the last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (without BatchNormalization), ResNet-1K, and DenseNet series.
 (Again, much more detailed set of plots is available at \cite{XXX-WEB-LINK}; but note that the corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than the results we present here.)
@@
         %\includegraphics[width=4.5cm]{img/VGG_fnl_alpha_depth.png} 
-        \includegraphics[width=3.7cm]{img/VGG_fnl_alpha_depth.png} 
+        \includegraphics[width=4.0cm]{img/VGG_fnl_alpha_depth.png} 
                 \label{fig:vgg-alpha-layers}
     }
-    \qquad
+    %\qquad
     \subfigure[ ResNet-1K ]{
         %\includegraphics[width=4.5cm]{img/ResNet_fnl_alpha_depth.png} 
-        \includegraphics[width=3.7cm]{img/ResNet_fnl_alpha_depth.png} 
+        \includegraphics[width=4.0cm]{img/ResNet_fnl_alpha_depth.png} 
         \label{fig:resnet-alpha-layer}
     }
-    \qquad
+    %\qquad
     \subfigure[ DenseNet ]{
         %\includegraphics[width=4.5cm]{img/DenseNet_fnl_alpha_depth.png} 
-        \includegraphics[width=3.7cm]{img/DenseNet_fnl_alpha_depth.png} 
+        \includegraphics[width=4.0cm]{img/DenseNet_fnl_alpha_depth.png} 
         \label{fig:densenet-alpha-layer}
@@
             }
-    \label{fig:vgg-alpha-layers}
+    \label{fig:3models-alpha-layers}
 \end{figure}
@@
 
-For the ResNet-1K models, $\alpha$ also increases in the last few layers (more dramatically, in fact, observe the differing scales on the Y axes).
+For the ResNet-1K models, Figure~\ref{fig:resnet-alpha-layer} shows that $\alpha$ also increases in the last few layers (more dramatically, in fact, observe the differing scales on the Y axes).
 However, as the ResNet-1K models get deeper, there is a wide range over which $\alpha$ values tend to remain small.
-\michael{Can we say something about exceptions which are larger than for VGG or DenseNet.}
-\charles{You have the same plots I do..what do you want to say >}
-\michael{Are they small or different types of layers or something?}
-\charles{Not that I am aware of.  Ill think on it}
+%\michael{Can we say something about exceptions which are larger than for VGG or DenseNet.}
+%\charles{You have the same plots I do..what do you want to say >}
+%\michael{Are they small or different types of layers or something?}
+%\charles{Not that I am aware of.  Ill think on it}
 This is seen for other models in the ResNet-1K series, but it is most pronounced for the larger ResNet-1K (152) model, where $\alpha$ remains relatively stable at $\alpha\sim 2.0$, from the earliest layers all the way until we reach close to the final layers.
 
-For the DenseNet models, $\alpha$ tends to increase as the layer id increases, in particular for layers toward the end.
+For the DenseNet models, Figure~\ref{fig:densenet-alpha-layer} shows that $\alpha$ tends to increase as the layer id increases, in particular for layers toward the end.
 While this is similar to what is seen in the VGG models, with the DenseNet models, $\alpha$ values increase almost immediately after the first few layers, and the variance is much larger (in particular for the earlier and middle layers, where it can range all the way to $\alpha\lesssim 8.0$) and much less systematic throughout the network.
@@
       %\includegraphics[scale=0.14]{img/resnet4d_maxev.png}
-      \includegraphics[width=3.7cm]{img/resnet4d_maxev.png}
+      \includegraphics[width=4.0cm]{img/resnet4d_maxev.png}
       %\includegraphics[width=4.5cm]{img/resnet4d_maxev.png}
@@
    }
-   \qquad
+   %\qquad
    \subfigure[$\alpha$ for ResNet20 layers]{
       %\includegraphics[scale=0.14]{img/resnet4d_alphas.png}
-      \includegraphics[width=3.7cm]{img/resnet4d_alphas.png}
+      \includegraphics[width=4.0cm]{img/resnet4d_alphas.png}
       %\includegraphics[width=4.5cm]{img/resnet4d_alphas.png}

### 2020-02-02 introduction.tex
pass over most sections, more to do on sec 5 and 6

@@
 %, in particular in academic environments, 
-is to train a model and then evaluate its training and/or testing error.
+is to train a model and then evaluate its training/testing error.
 There are many problems with this approach.
@@
 they give very limited insight into the overall properties of the model, 
-they do not take into account the (often extremely large human and CPU/GPU) time for hyperparameter fiddling,
+they do not take into account the (often large human and CPU/GPU) time for hyperparameter fiddling,
 they typically do not correlate with other properties of interest such as robustness or fairness or interpretability, 
 and so on.
-A somewhat less well-known problem, but one that is increasingly important (in particular in industrial-scale ML---where the \emph{users} of models are not the \emph{developers} of the models) is that one may have access to neither the training data nor the testing data.
+A less well-known problem, but one that is increasingly important (in particular in industrial-scale ML---where the \emph{users} of models are not the \emph{developers} of the models) is that one may have access to neither the training data nor the testing data.
 Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and be told to use it.
 
-Na\"{\i}vely---but in our experience commonly, among both ML practitioners and ML theorists---if one does not have access to training or testing data, then one can say absolutely nothing about the quality of a ML model.
-This may be true in worst-case theory, but ML models are used in practice, and there is a need for a \emph{practical theory} to guide that practice.
-Moreover, if ML is to become an industrial process, then that process will become siloed: some groups will gather data, other groups will develop models, and still other groups will use those models.
-The users of models can not be expected to know the precise details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
-%
+Na\"{\i}vely---but in our experience commonly, among ML practitioners and ML theorists---if one does not have access to training or testing data, then one can say absolutely nothing about the quality of a ML model.
+This may be true in worst-case theory, but models are used in practice, and there is a need for a \emph{practical theory} to guide that practice.
+Moreover, if ML is to become an industrial process, then that process will become siloed: some groups will gather data, other groups will develop models, and other groups will use those models.
+Users of models cannot be expected to know the precise details of how models were built, the specifics of data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized,~etc.
+
 Moreover, in industry, one faces unique practical problems such as: do we have enough data for this model? 
 Methods that are developed and evaluated on any well-defined publicly-available coprus of data, no matter how large or diverse or interesting, are clearly not going to be well-suited to address problems such as this.
-It is thus of great practical interest to have metrics to evaluate the quality of a ML model in the absence of training and testing data and without any detailed knowledge of the training and testing process.
-We seek a practical theory for pretrained models which can predict how, when, and why such models can be expected to perform well or poorly.
+It is of great practical interest to have metrics to evaluate the quality of a ML model---in the absence of training/testing data and without any detailed knowledge of the training/testing process.  
+We seek a practical theory for pretrained models which can predict how, when, and why such models can be expected to perform well or~poorly.
 
@@
 \footnote{Clearly, there is a selection bias or survivorship bias here---people tend not to make publicly-available their poorly-performing models---but these models are things in the world that (like social networks or the internet) can be analyzed for their properties.}
-They include 
-\michael{XXX, XXX and XXX.  MORE DETAILS.}
-\michael{XXX.  LIST PLACES WHERE THEY ARE AVAIALBLE, HERE OR IN NEXT SECTION.}
 Importantly, all of these models have been trained by someone else and have been viewed to be of sufficient interest/quality to be made publicly-available; and, for all of these models, we have no access to training data or testing data, and we have no knowledge of the training/testing protocols.
@@
 In more detail, our main contributions are the following.
-%\begin{itemize}
-%\item 
-First, 
-motivated by practical problems, we formulate the question ``Can one predict trends in the quality of state-of-the-art neural networks without access to training or testing data?''
-%\item
-Second, 
-to answer this question, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in CV and NLP.
-%\item 
-Third, 
-we find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best'' models, but they can qualitatively fail for models that may not be well-trained, i.e., to distinguish ``good-vs-bad'' models.
-%\item 
-Fourth, 
-we find that PL-based metrics do much better, quantitatively better at discriminating good-better-best and qualitatively better at discriminating good-vs-bad.
-%\item 
-Finally, 
-we find that PL-based metrics can also be used to characterize fine-scale properties of models, including understanding layer-wise \emph{correlation flow}, and evaluating post-training modifications such as model distillation and model improvement.
-%\end{itemize}
+\begin{itemize}[leftmargin=*]
+\item 
+%First, 
+Motivated by practical problems, we formulate the question ``Can one predict trends in the quality of state-of-the-art neural networks without access to training or testing data?''
+\item
+%Second, 
+%To answer this question, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in CV and NLP.
+To answer this question, we analyze hundreds of publicly-available pretrained state-of-the-art models from CV and NLP.
+\item 
+%Third, 
+We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best'' models, but they can qualitatively fail for models that may not be well-trained, i.e., to distinguish ``good-vs-bad'' models.
+\item 
+%Fourth, 
+We find that PL-based metrics do much better, quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-vs-bad models.
+\item 
+%Finally, 
+We find that PL-based metrics can also be used to characterize fine-scale properties of models, including understanding layer-wise \emph{correlation flow}, and evaluating post-training modifications such as model distillation and model improvement.
+\end{itemize}
 
-\paragraph{Organization of this paper.}
-
-We start in Section~\ref{sxn:background} and Section~\ref{sxn:methods} with background and an overview of our general approach.
-In Section \ref{sxn:cv}, we study three well-known widely-available DNN CV architectures (the VGG, ResNet, and DenseNet series of models); and we provide an illustration of our basic methodology, both to evaluate the different metrics against reported test accuracies and to use quality metrics to understand model properties.
-Then, in Section~\ref{sxn:nlp}, we look at several variations of a popular NLP DNN architecture (the OpenAI GPT and GPT2 models); and we show how model quality and properties vary between several variants of GPT and GPT2, including how metrics behave similarly and differently.
-Then, in Section \ref{sxn:all_cv_models}, we present results based on an analysis of hundreds of pretraind DNN CV models, showing how well each metric is correlated with the reported test accuracies, and how the Alpha-Norm metric(s) perform remarkably well.
-Finally, in Section~\ref{sxn:conc}, we provide a brief discussion and conclusion.
+\noindent
+We emphasize that our goal is a practical theory to predict trends in the quality of state-of-the-art NN models, i.e., not to make a statement about every publicly-available model.
+We have examined hundreds of models, and are exceptions to the general trends we state.
+Several of the most interesting are described below.
 
@@
 \michael{FIX THIS PAR.}
+
+\paragraph{Organization of this paper.}
+
+We start in Section~\ref{sxn:background} and Section~\ref{sxn:methods} with background and an overview of our general approach.
+In Section \ref{sxn:cv}, we study three well-known widely-available DNN CV architectures (the VGG, ResNet, and DenseNet series of models); and we provide an illustration of our basic methodology, both to evaluate the different metrics against reported test accuracies and to use quality metrics to understand model properties.
+Then, in Section~\ref{sxn:nlp}, we look at several variations of a popular NLP DNN architecture (the OpenAI GPT and GPT2 models); and we show how model quality and properties vary between several variants of GPT and GPT2, including how metrics behave similarly and differently.
+Then, in Section \ref{sxn:all_cv_models}, we present results based on an analysis of hundreds of pretraind DNN CV models, showing how well each metric is correlated with the reported test accuracies, and how the Alpha-Norm metric(s) perform remarkably well.
+Finally, in Section~\ref{sxn:conc}, we provide a brief discussion and conclusion.

### 2020-02-02 methods.tex
pass over most sections, more to do on sec 5 and 6

@@
 Typically, to apply RMT, we would usually define Correlation Matrix with $1/N$ normalization and assume that the variance of $\mathbf{X}$ is either unity or a known constant.% 
-\footnote{For formal proofs of Heavy Tailed results, one typically needs a normalization such as $1/N^{1-\alpha}$.}
+\footnote{For Heavy Tailed theorems, one typically needs a normalization such as $1/N^{1-\alpha}$.}
 Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with the variance already normalized to $\sqrt{1/N}$, or some variant of this, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.

### 2020-02-02 nlp_models.tex
pass over most sections, more to do on sec 5 and 6

@@
 Because GPT and GPT are trained on different data sets, the initial Embedding matrices differ in shape.
+\michael{Clarify that, which is second GPT.}
 GPT  has an initial Token and Positional Embedding layers, of dimension $(40478\times 768)$ and $(512\times 768)$, respectively, whereas GPT2 has input Embeddings of shape $(50257\times 768)$ and $(1024\times 768)$, respectively.
@@
 The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
-Table \ref{table:nlp} reports results for the average log norm metrics and the Weighted Alpha metric, 
-\michael{Fix that.}
 
@@
 \hline
-%   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
+%   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha Norm \\
 % Series & \#Layers   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
@@
 \end{center}
-\caption{Average value for each of the four metrics for pretrainnd OpenAI GPT and GPT2 models.}
+\caption{Average value for the Average Log Norm metrics and the Weighted Alpha metric for pretrainnd OpenAI GPT and GPT2 models.  }
 \label{table:nlp}
@@
 We have analyzed the four quality metrics described in Section~\ref{sxn:methods} for the OpenAI GPT and GPT2 pretrained models.
+See Table \ref{table:nlp} for a summary of results.
 We start by examining the PL exponents $\alpha$ for GPT and GPT2 (small).
+Observe that all four metrics increase when going from GPT to GPT (small), i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
+Observe also that (with one minor exception involving the Log Frobenius norm metric) all four metrics decrease as one goes from GPT2 medium to GPT2 large to GPT2 xl, indicating that the larger models are indeed better than the smaller models.
+%(The one minor exception is the Log Frobenius norm increasing from $1.85$ to $1.86$, for GPT2 large to GPT2 xl.)
+\michael{We seem to have typos(s), so I'm not going to flesh out all conclusions I see yet.  The comments I just mad may be wrong, if GPT (small) is really GPT2 (small).  FIX TYPOS AND CLARIFY.}
+These results are broadly consistent with the results in Section~\ref{sxn:cv} on series of CV models.
+
 Figure~\ref{fig:GPT-alpha-hist} shows the histogram (empirical density), for all layers, of $\alpha$ for GPT (blue) and this GPT2 (red) model.  
@@
     \centering
-    \subfigure[Power Law Exponent $(\alpha)$]{
+    \subfigure[PL exponent ($\alpha$)]{
         %\includegraphics[width=5cm]{img/GPT-alpha-hist.png}
@@
     %\qquad
-    \subfigure[log Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
+    \subfigure[Log Spectral Norm]{
         %\includegraphics[width=5cm]{img/GPT-snorm-hist.png}
@@
 
-    \subfigure[Power Law Exponent $(\alpha)$]{
+    \subfigure[PL exponent ($\alpha$)]{
         %\includegraphics[width=4.5cm]{img/GPT-alpha-depth.png}
-        \includegraphics[width=3.7cm]{img/GPT-alpha-depth.png}
+        \includegraphics[width=4.0cm]{img/GPT-alpha-depth.png}
         \label{fig:gpt-alpha-layer}
     }
-    \qquad
-    \subfigure[log Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
+    %\qquad
+    \subfigure[Log Spectral Norm]{
         %\includegraphics[width=4.5cm]{img/GPT-snorm-depth.png}
-        \includegraphics[width=3.7cm]{img/GPT-snorm-depth.png}
+        \includegraphics[width=4.0cm]{img/GPT-snorm-depth.png}
         \label{fig:gpt-snorm-layer}
     }
-    \qquad
-    \subfigure[log Alpha-Norm $(\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha})$]{
+    %\qquad
+    \subfigure[Log Alpha Norm]{
         %\includegraphics[width=4.5cm]{img/GPT-pnorm-depth.png}
-        \includegraphics[width=3.7cm]{img/GPT-pnorm-depth.png}
+        \includegraphics[width=4.0cm]{img/GPT-pnorm-depth.png}
         \label{fig:gpt-pnorm-layer}
@@
 
-We now look across the series of increasingly improving GPT2 models, examining both our
-PL exponent $\alpha$, as well as the various Norm metrics.
-As we move from small to xl, both the PL exponents $\alpha$, and the peak of distribtion of
-log Norm metrics shifts from larger to smaller values. 
-Figure \ref{fig:gpt2-histograms} shows the histograms over the layer weight matrics
+We now look across the series of increasingly improving GPT2 models, examining both the PL exponent $\alpha$ as well as the various Log Norm metrics.
+In general, as we move from small to xl, the histograms for both $\alpha$ exponents and the log Norm metrics shift from larger to smaller values. 
+See Figure \ref{fig:gpt2-histograms}, which shows the histograms over the layer weight matrics
 for fitted $\alpha$, and the log Spectral Norm
  $\log\Vert\mathbf{W}\Vert_{\infty}$  
-and log Alpha-Norm
+and log Alpha Norm
  $\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$ 
@@
 
-We see that the average $\alpha$ decreases with increasing model size,
-althogh, the differences are less noticible between GTP2 mkodels than between the GPT and GPT2 models.
+We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticible between different GTP2 models than between the GPT and GPT2 models.
 Unlike GPT, however, the Layer (log) Spectral Norms $\log\vert\mathbf{W}\Vert_{\infty}$ 
-and (log) Alpha-Norms $\log\vert\mathbf{W}\Vert_{\alpha}^{\alpha}$
+and (log) Alpha Norms $\log\vert\mathbf{W}\Vert_{\alpha}^{\alpha}$
 behave more as expected for GPT2 layers, with the larger models  consistently having  smaller norms. 
-Likewise, Figure \ref{fig:gpt2-snorm-hist} shows decreasing average log Spectral Norms
-with the larger models.   As we have seen in the trends of other well trained models.
+\michael{Clarify.}
+Likewise, Figure \ref{fig:gpt2-snorm-hist} shows decreasing average log Spectral Norms with the larger models.  
+As we have seen in the trends of other well trained models.
 
-We do notice, however, that while the peaks of the $\alpha$ is getting smaller, towards $2.0$,
+We do notice, however, that while the peaks of the $\alpha$ is getting smaller, 
+towards $2.0$,
 the tails of the distribution shifts right, with larger GPT2 models having more
-usually large $\alpha$.  We suspect this indicates that these larger GPT2 models are 
-overparameterized and/or not yet fully optimized and require datasets even larger
-than the recent XL $1.5B$ release~\cite{gpt2-xl}.
+usually large $\alpha$.  
+We suspect this indicates that these larger GPT2 models are over-parameterized/under-optimized and 
+that they have capacity to support datasets even larger than the recent XL $1.5B$ release~\cite{gpt2-xl}.
 
@@
 
-    \subfigure[Power Law Exponent $(\alpha)$]{
+    \subfigure[PL exponent ($\alpha$)]{
         %\includegraphics[width=4.5cm]{img/GPT2_all_alpha_hist.png}
-        \includegraphics[width=3.7cm]{img/GPT2_all_alpha_hist.png}
+        \includegraphics[width=4.0cm]{img/GPT2_all_alpha_hist.png}
         \label{fig:gpt2-alpha-hist}
     }
-    \qquad
-    \subfigure[log Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
+    %\qquad
+    \subfigure[Log Spectral Norm]{
         %\includegraphics[width=4.5cm]{img/GPT2*_all_spectralnorm_hist.png}
-        \includegraphics[width=3.7cm]{img/GPT2*_all_spectralnorm_hist.png}
+        \includegraphics[width=4.0cm]{img/GPT2*_all_spectralnorm_hist.png}
         \label{fig:gpt2-snorm-hist}
     }
-    \qquad
-    \subfigure[log Alpha-Norm $(\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha})$]{
+    %\qquad
+    \subfigure[Log Alpha Norm]{
         %\includegraphics[width=4.5cm]{img/GPT2_all_logpnorm_hist.png}
-        \includegraphics[width=3.7cm]{img/GPT2_all_logpnorm_hist.png}
+        \includegraphics[width=4.0cm]{img/GPT2_all_logpnorm_hist.png}
         \label{fig:gpt2-pnorm-hist}
     }
-    \caption{Comparison of Power Law Exponnents, Spectral Norm, and Alpha-Norm for different size models in the GPT2 architecture series.
-*Note: the log spectral norms (b) histogram omits the first 2 layers, corresponding to the embedding layer, which are normalized differently and have anamolously large values.}
+    \caption{PL exponents ($\alpha$), Log Spectral norm, and Log Alpha norm for different size models in the GPT2 architecture series.
+(The Log Spectral norm histogram omits the first 2 layers, corresponding to the embedding layer, as these are normalized differently and have anamolously large values.)}
     \label{fig:gpt2-histograms}

### 2020-02-02 abstract_file.tex
minor changes

@@
-In many applications, one must work with models that have been trained by someone else.
-For such \emph{pretrained models}, one typically does not have access to training data or test data, and one does not know the details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
+In many practical AI applications, one works with deep neural network (DNN) models trained by someone else.
+For such \emph{pretrained models}, one typically does not have access to training data or test data, and, moreover,  does not know the details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
 Here, we present and evaluate quality metrics for pretrained neural network models at scale.
@@
 %%The most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in a system) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
-The most promising are norm-based metrics (such as those used to provide capacity control in traditional statistical learning theory) and metrics based on fitting eigenvalue distributions to truncated power law (PL) distributions (which derive from statistical mechanics, in particular the recently-developed Theory of Heavy-Tailed Self Regularization, and which can be used to characterize the degree of strong correlations in a system).
+The most promising are norm-based metrics (such as those used to provide capacity control in traditional statistical learning theory) and metrics based on fitting eigenvalue distributions to truncated power law (PL) distributions (which derive from statistical mechanics, in particular the recently-developed Theory of Heavy-Tailed Self Regularization, and characterize the strength of correlations in a DNN).
 %
-Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision and natural language processing.
-We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best.'' 
-On the other hand, for models that may not be well-trained (which, arguably is the point of needing metrics to evaluate the quality of pretrained models), i.e., when we want to distinguish ``good-vs-bad,'' norm-based metrics can qualitatively fail.
-We also find that PL-based metrics do much better, quantitatively better at discriminating good-better-best and qualitatively better at discriminating good-vs-bad.
-PL-based metrics can also be used to characterize fine-scale properties of models, e.g., understanding layer-wise \emph{correlation flow}, and evaluate post-training modifications such as model distillation and model improvement.
+Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision (ZCV) and natural language processing (NLP)
+We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can discriminate amoing ``good-better-best''  models.
+On the other hand, for poorly trained models--which, arguably why we need quality metrics-- we want to distinguish which and when norm-based metrics can distiungush ''good-vs-bad'' models.
+We also find that PL-based metrics do much better, quantitatively better, at discriminating aming ''good-better-best'' models, and qualitatively better at discriminating ''good-vs-bad'' models.
+PL-based metrics can also be used to characterize fine-scale properties of models, e.g., understanding layer-wise \emph{correlation flow}, and evaluate post-training modifications such as model distillation, increasing the data set size, and other model improvements.

### 2020-02-02 background.tex
pass over most sections, more to do on sec 5 and 6
minor changes

@@
 Most theory for DNNs is applied to small toy models and assumes access to data.
-There is very little work asking how to predict, in a theoretically-principled manner, the quality of large-scale state-of-the-art NNs, and how to do so without access to training data or testing data or details of the training protocol, etc.
+There is very little work asking how to predict, in a theoretically-principled manner, the quality of large-scale state-of-the-art DNNs, and how to do so without access to training data or testing data or details of the training protocol, etc.
 Our work is, however, loosely related to several other lines of work.
 
-\paragraph{Statistical mechanics theory for NNs.}
+\paragraph{Statistical mechanics theory for DNNs.}
 
-Statistical mechanics ideas have long had influence on NN theory and practice~\cite{EB01_BOOK, MM17_TR, BKPx20}; and 
+Statistical mechanics ideas have long had influence on DNN theory and practice~\cite{EB01_BOOK, MM17_TR, BKPx20}; and 
 our best-performing metrics (those using fitted PL exponents) are based on statistical mechanics~\cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM19_KDD, MM20_SDM}, in particular the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
@@
 Several very good overviews of the more common approach are available~\cite{EB01_BOOK, BKPx20}.
-We will \emph{use} statistical mechanics theory more as it is used in quantitative finance.
-Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statisical mechanics of heavy tailed and strongly correlated systems.
+We will \emph{use} statistical mechanics in a broader sense, drawing upon results from quantitative finance and random matrix theory (RMT).
+Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statisical mechanics of heavy tailed and strongly correlated systems\charles{, and current work (in progress)\cite{blog}}
 
@@
 In this area, theoretical work aims to prove generalization bounds, and applied work uses these norms as a regularization function to improve training.
-While we do find that norms provide relatively good quality metrics, at least for distinguishing good-better-best among well-trained models, we are not interested in proving generalization bounds or improving training.
+While we do find that norms provide relatively good quality metrics, at least for distinguishing good-better-best among well-trained models, we are not interested in proving generalization bounds \nred{or developing new regularizers.}
 
-\paragraph{Practical problems poorly addressed by theory.}
-There are many very practical problems in ML that are poorly addressed by existing theory and that either motivated our work or should be addressable by our techniques.
+\paragraph{Practical problems need a practical theory.}
+\nred{Exiting ML theory does not address many practical AI problems that our techiques directly target.}
+%There are many very practical problems in ML that are poorly addressed by existing theory and that either motivated our work or should be addressable by our techniques.
 Here are several.
-\begin{itemize}
+\begin{itemize}[leftmargin=*]
 \item
@@
 \end{itemize}
-Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text or realiztic chatbots, developing self driving cars, or just clusterting user profiles.
-\footnote{For example, current chatbots use perplexity as a proxy for passing a Turing test.}
+
+\charles{see comments in tex file}
+% we adress this above
+%Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text or realiztic chatbots, developing self driving cars, or just clusterting user profiles.
+%\footnote{For example, current chatbots use perplexity as a proxy for passing a Turing test.}
 %, and when distilling a reliable model in some way to obtain comparable training/test quality but that damages the model in some other subtle way. 

### 2020-02-02 conclusion.tex
minor changes

@@
 
-We conclude with a few thoughts on what a \emph{practical theory} of NNs should look like.
+We conclude with a few thoughts on what a \emph{practical theory} of DNNs should look like.
 %
@@
 %
-Our results, which are based on our \emph{use} of sophisticated statistical mechanics theory to solve an important practical NN problems, suggests that this approach should be of interest more generally for those interested in developing a practical NN theory.
+Our results, which are based on our \emph{use} of sophisticated statistical mechanics theory to solve an important practical DNN problems, suggests that this approach should be of interest more generally for those interested in developing a practical DNN theory.
 

### 2020-02-02 cv_models.tex
some misc stuff
minor changes

@@
 \end{figure}
+
+%MM% \begin{figure}
+%MM% \centering
+%MM% \begin{minipage}[b]{.22\textwidth}
+%MM% %<Code for the first figure>
+%MM% XXX
+%MM% \caption{Caption}\label{label-a}
+%MM% \end{minipage}\qquad
+%MM% \begin{minipage}[b]{.22\textwidth}
+%MM% %<Code for the second figure>
+%MM% XXX
+%MM% \caption{Caption}\label{label-b}
+%MM% \end{minipage}
+%MM% \end{figure}
 
@@
 The pre-trained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the VGG/ResNet/DenseNet models we condisered previously.
-Norms are often used as regularizers, and thys they are a natural metric to consider, e.g., during training or distillation.
-The PL $\alpha$ captures more subtle correlations.
+\charles{MOvE THIS SOMEWHERE ELSE AND CLEAN UP THIS SECTION Notice while norms make good regularizers for a single model, there is no reason \emph{a priori} 
+to expect them correlate so well with the test accuracies across different models.
+However, we do expect the PL $\alpha$ to do so because it effectively measures the amount of correlation in the model}
 %
-The reason for this is that the \texttt{distiller} Group Regularization technique has the unusual effect of increasing the norms of the $\mathbf{W}$ feature maps for at least two of the Conv2D layers.
+The reason for this is that the \texttt{distiller} Group Regularization technique 
+%%has the unusual effect of increasing
+\nred{spuriously increases}
+ the norms of the $\mathbf{W}$ pre-activation maps for at least two of the Conv2D layers.
 \michael{This impedance shuffling or whatever it is called messes up norms, but the strong correlations are still preserved, as seen in the $\hat{\alpha}$ metric, and the model quality is good.

### 2020-02-02 introduction.tex
minor changes

@@
 %, in particular in academic environments, 
-is to train a model and then evaluate its training/testing error.
-There are many problems with this approach.
-Well-known problems with just examining training/testing curves include that 
-they give very limited insight into the overall properties of the model, 
+is to train a model and then evaluate its training/testing error--but
+there are many problems with this approach.
+The training/testing curves give very limited insight into the overall properties of the model, 
 they do not take into account the (often large human and CPU/GPU) time for hyperparameter fiddling,
@@
 and so on.
-A less well-known problem, but one that is increasingly important (in particular in industrial-scale ML---where the \emph{users} of models are not the \emph{developers} of the models) is that one may have access to neither the training data nor the testing data.
-Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and be told to use it.
+A less well-known problem, but one that is increasingly important (in particular in industrial-scale 
+artificial intelligence (AI), arises when the model \emph{user} is not the model \emph{developer}.
+here, one may not have access to either the training data nor the testing data.
+\nred{Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it "as is'', or fine tune it and/or compress it.}
 
@@
 
-Moreover, in industry, one faces unique practical problems such as: do we have enough data for this model? 
+Moreover, for many large scale, practical applications, one can not define a true test metric.
+For example, NLP models that generate fake text or conversational chatbots may use a proxy, like perplexity, as a test metric, but, in the end, really require human evaluation.
+Or, say, models that cluster user profiles, which are widely used in marketing, are unsupervised and have no labels for comparison.
+\charles{more here, clean up?}
+
+Most importantly, in industry, one faces unique practical problems such as: do we have enough data for this model? 
+\nred{Indeed, high quality, labeled data can be very expensive to acquire, and this cost can make or break a project.}
 Methods that are developed and evaluated on any well-defined publicly-available coprus of data, no matter how large or diverse or interesting, are clearly not going to be well-suited to address problems such as this.
-It is of great practical interest to have metrics to evaluate the quality of a ML model---in the absence of training/testing data and without any detailed knowledge of the training/testing process.  
+It is of great practical interest to have metrics to evaluate the quality of a trained model---in the absence of training/testing data and without any detailed knowledge of the training/testing process.
 We seek a practical theory for pretrained models which can predict how, when, and why such models can be expected to perform well or~poorly.
 
-In this paper, we present and evaluate quality metrics for pretrained neural network (NN) and deep neural network (DNN) models at scale.
+In this paper, we present and evaluate quality metrics for pretrained deep neural network (DNN) models, and at scale.
 To do so, we consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
@@
 The \emph{quality metrics} we consider are based on the spectral properties of the layer weight matrics.
-In particular, they are based on the norms of weight matrices (such norms have been used in traditional statistical learning theory to bound capacity and construct regularizers) and parameters of power law (PL) fits of the eigenvalues of weight matrices (such PL fits are based on statistical mechanics approaches to NNs).
+In particular, they are based on the norms of weight matrices (such norms have been used in traditional statistical learning theory to bound capacity and construct regularizers) and parameters of power law (PL) fits of the eigenvalues of weight matrices (such PL fits are based on statistical mechanics approaches to DNNs).
 Note that, while we use traditional norm-based and PL-based metrics, our goals are different.
@@
 \footnote{One could of course use these techniques to improve training, and we have been asked about that, but we are not interested in that here.
-Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained NN models.}
+Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained DNN models.}
 Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves, and in both cases \emph{we aim to predict trends in the quality of pre-trained models without access to training or testing data}.
@@
 \noindent
-We emphasize that our goal is a practical theory to predict trends in the quality of state-of-the-art NN models, i.e., not to make a statement about every publicly-available model.
-We have examined hundreds of models, and are exceptions to the general trends we state.
-Several of the most interesting are described below.
+We emphasize that our goal is a practical theory to predict trends in the quality of state-of-the-art DNN models, i.e., not to make a statement about every publicly-available model.
+\nred{We have examined hundreds of models, we indentify the general trends, and highlight interesting exceptions.}
+%%Several of the most interesting are described below.
 
@@
 and, in order to make our results fully reproducible, we provide Jupyter notebooks in the github repo for this paper~\cite{repo},
-with fully reproducible Jupyter notebooks.%
-\footnote{\url{https://github.com/CalculatedContent/kdd2020} \michael{DIFFERENT NAME?}}
+and with fully reproducible Google colab notebooks for the larger calculations.%
+\footnote{\url{https://github.com/CalculatedContent/kdd2020} \michael{DIFFERENT NAME?}\charles{suggestions?}}
 \michael{LINK TO: Google Colab, IS THIS THAT REPO.}

### 2020-02-02 methods.tex
minor changes

@@
 Consider the following.
+\charles{Note I write X for the $\alpha$-norm and not W.  This is because it would be $2\alpha$.  I know its a bit confusing.
+We have to be careful here with the Frobenius norm also and check that we are not off by a factor of 2}
 \begin{itemize}
@@
 Importantly, the normalization of weight matrices does \emph{not} affect the Power Law fits because the Heavy Tailed exponent $\alpha$ (as well as other metrics such as the Stable Rank and MP Soft Rank~\cite{MM18_TR,MM19_HTSR_ICML}) is scale-invariant.
-Norm-based metrics, however, do depend strongly on the scale of the weight matrix.
+Norm-based metrics, however, do depend strongly on the scale of the weight matrix--\nred{that's the point.}
+\nred{Indeed, early theoretical work by Bartlett suggests that the test accuracy depends strongly on the ``total size'' of the weight matrics.}
 Typically, to apply RMT, we would usually define Correlation Matrix with $1/N$ normalization and assume that the variance of $\mathbf{X}$ is either unity or a known constant.% 
-\footnote{For Heavy Tailed theorems, one typically needs a normalization such as $1/N^{1-\alpha}$.}
+\footnote{For Heavy Tailed theorems, one typically needs a normalization such as \nred{$1/N^{\alpha-1}$. check this}}
 Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with the variance already normalized to $\sqrt{1/N}$, or some variant of this, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.

### 2020-02-02 nlp_models.tex
updates

@@
 In general, we have found that they have larger $\alpha$ PL exponents.
-Based on HT-SR Theory (in particular, the interpretation of values of $\alpha \sim 2$ as modeling systems with good correlations over many size scales~\cite{BouchaudPotters03, SornetteBook}), this suggests that these models fail to capture successfully many of the correlations in the data (relative to their size) and thus that they are substantially \emph{under}-trained.
+Based on HT-SR Theory, 
+%(in particular, the interpretation of values of $\alpha \sim 2$ as modeling systems with good correlations over many size scales~\cite{BouchaudPotters03, SornetteBook}), 
+this suggests that these models fail to successfully capture many of the correlations in the data (relative to their size) and thus are substantially \emph{under}-trained.
 %
-More generally, these NLP models are quite different than the CV models we considered in Section~\ref{sxn:cv}, both in terms of their architecture (e.g., wheter there are convolutions) as well as how they are used (e.g., whether one uses prediction error or perplexity).
-Thus, they provide a very different test for our metrics.
+More generally, compared to the CV models of Section~\ref{sxn:cv},
+modern NLP models have larger weight matrices and display different spectral properties, and 
+%, both in terms of their architecture (e.g., wheter there are convolutions) as well as how they are used (e.g., whether one uses prediction error or perplexity).
+thus, they provide a very different test for our empirical quality metrics.
 
-\michael{WHERE: Norm-base metrics perform reasonably well on good NLP models, but they have problems with less well-trained models.
-PL-based metrics (the Log $\alpha$-Norm metric, $\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$, and the Weighted Alpha metric, $\hat\alpha =\alpha\log\lambda_{max} $) do even better.
-We highlight how to use the theory to identify poorly trained models where the empirical norm metrics will perform poorly.}
+Where Norm-base metrics perform reasonably well on well-trained NLP models, they behave anomalously poorly-trained models.
+Indeed, in ``bad models'', the weight matrics may display rank collapse, decreased Frobeinus mass, or unsually small Spectral norms, 
+which may be mis-interpreted as ``smaller is better.''
+In contast, PL-based metrics (the Log $\alpha$-Norm metric, $\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$, and the Weighted Alpha metric, $\hat\alpha =\alpha\log\lambda_{max} $) display consistant behavior even on poorly trained models, and we can use these metrics to help indentify
+where the archiectures need repair, and/or more and/or better data is needed.
 
-\michael{WHERE: We show that this not only works for CV, so we look at NLP.
-We now look in detail at the alpha metric, the spectral norm, and the new norm.
-We compare good and bad models, and a series of increasingly big models
-trained on big data sets.  We show why alpha works but spectral norm is not enough.
-The layer Spectral Norm behaves unexpectedly, while $\alpha$-based metrics can give insight into how well trained a model is.}
+%We now look in detail at the alpha metric, the spectral norm, and the new norm.
+%We compare good and bad models, and a series of increasingly big models
+%trained on big data sets.  We show why alpha works but spectral norm is not enough.
+%The layer Spectral Norm behaves unexpectedly, while $\alpha$-based metrics can give insight into how well trained a model is.
 
 \michael{WHERE: Say large $\alpha$ values don't mean much.}
+We do note, however, that many NLP models, such as the GPT and BERT, may have weight matrices with unusually large PL exponents
+($\alpha>6$), which indicates these matrices may be \emph{under}-correlated (i.e over-paramaterized), although the truncated PL fit
+itself may not be very reliable because the MLE estimator is unreliable in this range.  Phenomenologically, if we examine
+the ESD visually, we can usually describe these $\mathbf{W}$ as in the \emph{Bulk-Decay} or \emph{Bulk-plus-Spikes} phase\cite{MM}.
+We have conjectured previously that very well-trained DNNs would not have many \emph{outlier} $\alpha>6$, 
+and improved versions of GPT (shown below, and BERT (not shown) confirm this.
 
 \michael{WHERE: Highlight difference between $\alpha$ and $\hat{\alpha}$.}
+To avoid confusion, let us clarify the differences between $\alpha$ and $\hat{\alpha}$.  
+We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated power law (PL), parameterized by 2 values:
+the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.  (Technically, we also need the minimum
+eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.)
+The PL exponent $\alpha$ measures of the 
+amount of correlation in a DNN layer weight matrix $\mathbf{W}$.  It is valid for $\lambda<\lambda_{max}$, and
+it scale invariant (does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.)
+The Weighted-Alpha metric, $\hat{\alpha}=\alpha\log\lambda_{max}$, approximates the Log $\alpha$ (Shatten) Norm 
+$(\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha})$ (derived using statistical mechanics and RMT \nred{(in progress)}, 
+and lets us compute a balanced, weighted average log PL norm metrix for the entire DNN.  
+The $\hat{\alpha}$ metric also behaves like an imrpoved, weighted log average Spectral Norm, and may track this metric in some cases.
 

### 2020-02-02 nlp_models.tex
nlp edits

@@
 Observe that all four metrics increase when going from GPT to GPT (small), i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
-Observe also that (with one minor exception involving the Log Frobenius norm metric) all four metrics decrease as one goes from GPT2 medium to GPT2 large to GPT2 xl, indicating that the larger models are indeed better than the smaller models.
+Observe also that (with one minor exception involving the Log Frobenius norm metric) all four metrics decrease as one goes from GPT2 medium to large to xl, indicating that the larger models indeed look better than the smaller models.
 %(The one minor exception is the Log Frobenius norm increasing from $1.85$ to $1.86$, for GPT2 large to GPT2 xl.)
 \michael{We seem to have typos(s), so I'm not going to flesh out all conclusions I see yet.  The comments I just mad may be wrong, if GPT (small) is really GPT2 (small).  FIX TYPOS AND CLARIFY.}
-These results are broadly consistent with the results in Section~\ref{sxn:cv} on series of CV models.
+\charles{GPT2 (small) is GPT--they are the same model}
+These results coincide with the results in Section~\ref{sxn:cv} on CV models.
 

### 2020-02-02 nlp_models.tex
nlp update

@@
 %
-Because GPT and GPT are trained on different data sets, the initial Embedding matrices differ in shape.
-\michael{Clarify that, which is second GPT.}
+Because GPTand GPT2 are trained on different data sets, the initial Embedding matrices differ in shape.
 GPT has an initial Token and Positional Embedding layers, of dimension $(40478\times 768)$ and $(512\times 768)$, respectively, whereas GPT2 has input Embeddings of shape $(50257\times 768)$ and $(1024\times 768)$, respectively.
@@
 \michael{We seem to have typos(s), so I'm not going to flesh out all conclusions I see yet.  The comments I just mad may be wrong, if GPT (small) is really GPT2 (small).  FIX TYPOS AND CLARIFY.}
-\charles{GPT2 (small) is GPT--they are the same model}
+\charles{GPT2 (small) is GPT--I think they are the same model--but the embeddings may be different I will check again}
 These results coincide with the results in Section~\ref{sxn:cv} on CV models.

### 2020-02-03 nlp_models.tex
GPT small to GPT2 small

@@
 GPT & 49 & 1.64  & 1.72 & 7.01 & 7.28 \\
-GPT (small) & 49 & 2.04  & 2.54& 9.62 & 9.87 \\
+GPT2 small & 49 & 2.04  & 2.54& 9.62 & 9.87 \\
 GPT2 medium & 98 & 2.08 & 2.58& 9.74 & 10.01 \\
@@
 See Table \ref{table:nlp} for a summary of results.
-We start by examining the PL exponents $\alpha$ for GPT and GPT2 (small).
-Observe that all four metrics increase when going from GPT to GPT (small), i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
+We start by examining the PL exponents $\alpha$ for GPT and GPT2 small.
+Observe that all four metrics increase when going from GPT to GPT2 small, i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
 Observe also that (with one minor exception involving the Log Frobenius norm metric) all four metrics decrease as one goes from GPT2 medium to large to xl, indicating that the larger models indeed look better than the smaller models.
 %(The one minor exception is the Log Frobenius norm increasing from $1.85$ to $1.86$, for GPT2 large to GPT2 xl.)
-\michael{We seem to have typos(s), so I'm not going to flesh out all conclusions I see yet.  The comments I just mad may be wrong, if GPT (small) is really GPT2 (small).  FIX TYPOS AND CLARIFY.}
-\charles{GPT2 (small) is GPT--I think they are the same model--but the embeddings may be different I will check again}
-These results coincide with the results in Section~\ref{sxn:cv} on CV models.
 

### 2020-02-03 nlp_models.tex
updating nlp

@@
 
-Observe that these results indicate that $\alpha$ provides a good quality metric for comparing these two models, one of which which was designed to be of low quality.
-The deficient GPT model has numerous weight matrices with unsually large fitted exponents (indicating that they are not HT at all).  
-Indeed, we may expect that a poorly trained model will not exhibit consistent Heavy Tailed behavior in all layers.
+From this, we see that $\alpha$ provides a good quality metric for comparing these two models, the``bad'' GPT vs. the ``good'' GPT2.
+The deficient GPT has numerous unsually large $\alpha$ exponents--meaning they are not Heavy Tailed at all.
+Indeed, we expect that a poorly trained model lack Heavy Tailed behavior in all layers.
 On the other hand, as expected, the improved GPT2 model has, on average, smaller $\alpha$ than the older GPT, with all $\alpha\le6$.
-\michael{XXX.  FLESH OUT.}
 
@@
 However, as seen in Figure \ref{fig:GPT-snorm-hist},
- the poorly trained GPT model has many smaller (log) Spectral Norms $\log\Vert\mathbf{W}\Vert_{\infty}$
-than the GPT2, which would be inconsistent with a belief that smaller Spectral Norm is always better.
-Indeed, because there are so many anamously small $\Vert\mathbf{W}\Vert_{\infty}$,
-it appears that the GPT model may be exhibiting a kind of rank collapse.
+the ``bad'' GPT model has, spuriously, many smaller Spectral Norms $\log\Vert\mathbf{W}\Vert_{\infty}$,
+compared to the ``good'' GPT2--which violates the beleif that that smaller Spectral Norms are always better.
+Indeed, because there are so many anonymously small $\Vert\mathbf{W}\Vert_{\infty}$,
+it appears that the GPT model may be exhibiting a kind of soft rank collapse.
 This is an extremely important observation because it demonstrates that while the Spectral Norm
@@
 
-Note that Figure \ref{fig:GPT-snorm-hist} also shows a couple anomalously  large Spectral Norms.
+Note that Figure \ref{fig:GPT-snorm-hist} also shows a couple unusually large Spectral Norms, but,
+here, simply correspond to the change in normalization of the word embedding layer(s) and can be ignored.
 From Figure \ref{fig:gpt-snorm-layer} (below), we see that these correspond to the first embedding layer(s).
-These layers appear to have a normalization, and therefore a differenc scale.
+These layers appear to have a different effectvie normalization, and therefore a different scale.
 For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,

### 2020-02-03 nlp_models.tex
updating nlp

@@
 Indeed, because there are so many anonymously small $\Vert\mathbf{W}\Vert_{\infty}$,
-it appears that the GPT model may be exhibiting a kind of soft rank collapse.
+it appears that the GPT model may be exhibiting a kind of soft rank collapse and/or undesired
+loss of Frobeinus mass.
 This is an extremely important observation because it demonstrates that while the Spectral Norm

### 2020-02-03 nlp_models.tex
updating nlp

@@
 the ``bad'' GPT model has, spuriously, many smaller Spectral Norms $\log\Vert\mathbf{W}\Vert_{\infty}$,
-compared to the ``good'' GPT2--which violates the beleif that that smaller Spectral Norms are always better.
+compared to the ``good'' GPT2--which violates the belief that smaller Spectral Norms are always better.
 Indeed, because there are so many anonymously small $\Vert\mathbf{W}\Vert_{\infty}$,

### 2020-02-03 nlp_models.tex
updating nlp

@@
 
-Note that Figure \ref{fig:GPT-snorm-hist} also shows a couple unusually large Spectral Norms, but,
+Note that Figure \ref{fig:GPT-snorm-hist} also shows some unusually large Spectral Norms, but,
 here, simply correspond to the change in normalization of the word embedding layer(s) and can be ignored.

### 2020-02-03 nlp_models.tex
updating nlp

@@
 but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
- $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we layer all layers as-is (although one may to normalize
+ $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we treat all layers as-is (although one may to normalize
 the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier).

### 2020-02-03 nlp_models.tex
updating nlp

@@
 Indeed, because there are so many anonymously small $\Vert\mathbf{W}\Vert_{\infty}$,
-it appears that the GPT model may be exhibiting a kind of soft rank collapse and/or undesired
+it appears that the GPT model may be exhibiting a kind of \emph{scale collapse} and/or undesired
 loss of Frobeinus mass.
@@
 Figure \ref{fig:gpt-alpha-layer} plots $\alpha$ vs the depth (i.e. a layer id) for each model.
+The deficient GPT model displays two trends in $\alpha$ , one stable with $\alpha\sim 4$,
+and one increasing with layer id, with $\alpha$ reaching as high as $12$.
+In contrast, the well-trained GPT2 model shows consistant and stable patterns, again
+with one stable $\alpha\sim 3.5$ (and below the GPT trend), and the other only
+slightly trending up, with $alpha\le 6$. 
+Again, the scale-invariant $\alpha$ metric lets us identify potentially
+poorly-trained models.
+
+Notice also that the Spectral Norms
+(Figure \ref{fig:gpt-snorm-layer})
+and the $\alpha-$Norms
+(Figure \ref{fig:gpt-pnorm-layer})
+display an unusually some mall values for GPT,
+and some increasing trending values for GPT2.
+Generally speaking, we do not think scale-dependennt 
+norm metrics can be directly applied to distinguish ``good'' vs ``bad''  models 
+because of the anamalous scale collapse that is observed in bad models.
 

### 2020-02-03 nlp_models.tex
updating nlp

@@
 slightly trending up, with $alpha\le 6$. 
-Again, the scale-invariant $\alpha$ metric lets us identify potentially
+The scale-invariant $\alpha$ metric lets us identify potentially
 poorly-trained models.
@@
 norm metrics can be directly applied to distinguish ``good'' vs ``bad'' models 
-because of the anamalous scale collapse that is observed in bad models.
-
-\charles{Discuss Spectral Norm, alpha-Norm}
+because of the anamalous scale collapse that is frequently observed in bad models.
 

### 2020-02-03 nlp_models.tex
updating nlp

@@
 The GPT models we analyze are deployed with the popular HuggingFace PyTorch library~\cite{XXX-XXX}.
-GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ Layer Weight Matrices $\mathbf{W}$.
+GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ layer Weight Matrices $\mathbf{W}$.
 Each Block has 2 components, the Self Attention (attn) and the Projection (proj) matrics.  
@@
 %
-The OpenAI GPT2 (English) models are: \nred{gpt-small, gpt-medium, gpt-large, and gpt-xl}, \
+The OpenAI GPT2 (English) models are: \nred{GPT2-small, GPT2-medium, GPT2-large, and GPT2-xl}, \
 having include $12, 24, 36, \text{and }48$ layers, respectively, with increasingly larger weight matrices.
@@
 GPT & 49 & 1.64  & 1.72 & 7.01 & 7.28 \\
-GPT2 small & 49 & 2.04  & 2.54& 9.62 & 9.87 \\
+GPT2-small & 49 & 2.04  & 2.54& 9.62 & 9.87 \\
 GPT2 medium & 98 & 2.08 & 2.58& 9.74 & 10.01 \\
@@
 \end{center}
-\caption{Average value for the Average Log Norm metrics and the Weighted Alpha metric for pretrainnd OpenAI GPT and GPT2 models.  }
+\caption{Average value for the Average log Norm metrics and the Weighted Alpha metric for pretrainnd OpenAI GPT and GPT2 models.  }
 \label{table:nlp}
@@
 See Table \ref{table:nlp} for a summary of results.
-We start by examining the PL exponents $\alpha$ for GPT and GPT2 small.
-Observe that all four metrics increase when going from GPT to GPT2 small, i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
+We start by examining the PL exponents $\alpha$ for GPT and GPT2-small.
+Observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
 Observe also that (with one minor exception involving the Log Frobenius norm metric) all four metrics decrease as one goes from GPT2 medium to large to xl, indicating that the larger models indeed look better than the smaller models.
@@
 
-We now look across the series of increasingly improving GPT2 models, examining both the PL exponent $\alpha$ as well as the various Log Norm metrics.
-In general, as we move from small to xl, the histograms for both $\alpha$ exponents and the log Norm metrics shift from larger to smaller values. 
+We now look across the series of increasingly improving GPT2 models, (i.e. ``good, better, best''), 
+by examining both the PL exponent $\alpha$, as well as the various log Norm metrics.  
+In general, as we move from GPT2-small to GPT2-xl, 
+the histograms for both $\alpha$ exponents and the log Norm metrics downshift from larger to smaller values. 
 See Figure \ref{fig:gpt2-histograms}, which shows the histograms over the layer weight matrics
@@
 
-We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticible between different GTP2 models than between the GPT and GPT2 models.
-Unlike GPT, however, the Layer (log) Spectral Norms $\log\vert\mathbf{W}\Vert_{\infty}$ 
+We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticible 
+between the differing ``good, better, best'' GTP2 models than between the ``good vs bad'' GPT2 and GPT models.
+Unlike GPT, however, the layer (log) Spectral Norms $\log\vert\mathbf{W}\Vert_{\infty}$ 
 and (log) Alpha Norms $\log\vert\mathbf{W}\Vert_{\alpha}^{\alpha}$
-behave more as expected for GPT2 layers, with the larger models  consistently having  smaller norms. 
-\michael{Clarify.}
+behave more as expected for GPT2 layers, with the larger models, consistently having smaller norms. 
+\michael{Clarify.}\charles{???}
 Likewise, Figure \ref{fig:gpt2-snorm-hist} shows decreasing average log Spectral Norms with the larger models.  

### 2020-02-03 nlp_models.tex
updating nlp

@@
 which may be mis-interpreted as ``smaller is better.''
-In contast, PL-based metrics (the Log $\alpha$-Norm metric, $\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$, and the Weighted Alpha metric, $\hat\alpha =\alpha\log\lambda_{max} $) display consistant behavior even on poorly trained models, and we can use these metrics to help indentify
+In contast, PL-based metrics (the log $\alpha$-Norm metric, 
+$(\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha})$, and the Weighted Alpha metric, $\hat\alpha =\alpha\log\lambda_{max} $) display consistant behavior even on poorly trained models, and we can use these metrics to help indentify
 where the archiectures need repair, and/or more and/or better data is needed.
@@
 it scale invariant (does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.)
-The Weighted-Alpha metric, $\hat{\alpha}=\alpha\log\lambda_{max}$, approximates the Log $\alpha$ (Shatten) Norm 
+The Weighted-Alpha metric, $\hat{\alpha}=\alpha\log\lambda_{max}$, approximates the log $\alpha$ (Shatten) Norm 
 $(\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha})$ (derived using statistical mechanics and RMT \nred{(in progress)}, 
@@
 % Series & \#Layers   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
-        &        & Log                   & Log                        & Weighted       & Log                                 \\
+        &        & log                   & log                        & Weighted       & log                                 \\
         & Num.   & $\Vert\cdot\Vert_{F}$ & $\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
@@
 Observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
-Observe also that (with one minor exception involving the Log Frobenius norm metric) all four metrics decrease as one goes from GPT2 medium to large to xl, indicating that the larger models indeed look better than the smaller models.
-%(The one minor exception is the Log Frobenius norm increasing from $1.85$ to $1.86$, for GPT2 large to GPT2 xl.)
+Observe also that (with one minor exception involving the log Frobenius norm metric) all four metrics decrease as one goes from GPT2 medium to large to xl, indicating that the larger models indeed look better than the smaller models.
+%(The one minor exception is the log Frobenius norm increasing from $1.85$ to $1.86$, for GPT2 large to GPT2 xl.)
 
@@
     %\qquad
-    \subfigure[Log Spectral Norm]{
+    \subfigure[log Spectral Norm]{
         %\includegraphics[width=5cm]{img/GPT-snorm-hist.png}
@@
        }
-   \caption{Comparison of PL exponents, $\alpha$, and log Spectral Norms, $\log\Vert\mathbf{W}\Vert_{\infty}$, for the OpenAI GPT and GPT2 (small) pretrained models.}
+   \caption{Comparison of PL exponents, $\alpha$, and log Spectral Norms, 
+$(\log\Vert\mathbf{W}\Vert_{\infty})$, for the OpenAI GPT and GPT2 (small) pretrained models.}
 \label{fig:GPT-hist}
@@
 
+\paragraph{Scale Collapse.}
 However, as seen in Figure \ref{fig:GPT-snorm-hist},
-the ``bad'' GPT model has, spuriously, many smaller Spectral Norms $\log\Vert\mathbf{W}\Vert_{\infty}$,
+the ``bad'' GPT model has, spuriously, many smaller Spectral Norms $(\log\Vert\mathbf{W}\Vert_{\infty})$,
 compared to the ``good'' GPT2--which violates the belief that smaller Spectral Norms are always better.
@@
     %\qquad
-    \subfigure[Log Spectral Norm]{
+    \subfigure[log Spectral Norm]{
         %\includegraphics[width=4.5cm]{img/GPT-snorm-depth.png}
@@
     %\qquad
-    \subfigure[Log Alpha Norm]{
+    \subfigure[log Alpha Norm]{
         %\includegraphics[width=4.5cm]{img/GPT-pnorm-depth.png}
@@
 for fitted $\alpha$, and the log Spectral Norm
- $\log\Vert\mathbf{W}\Vert_{\infty}$  
+ $(\log\Vert\mathbf{W}\Vert_{\infty})$  
 and log Alpha Norm
- $\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$ 
+ $(\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha})$ 
 metrics.
@@
 between the differing ``good, better, best'' GTP2 models than between the ``good vs bad'' GPT2 and GPT models.
-Unlike GPT, however, the layer (log) Spectral Norms $\log\vert\mathbf{W}\Vert_{\infty}$ 
-and (log) Alpha Norms $\log\vert\mathbf{W}\Vert_{\alpha}^{\alpha}$
+Unlike GPT, however, the layer (log) Spectral Norms $(\log\vert\mathbf{W}\Vert_{\infty})$ 
+and (log) Alpha Norms $(\log\vert\mathbf{W}\Vert_{\alpha}^{\alpha})$
 behave more as expected for GPT2 layers, with the larger models, consistently having smaller norms. 
@@
 Likewise, Figure \ref{fig:gpt2-snorm-hist} shows decreasing average log Spectral Norms with the larger models.  
-As we have seen in the trends of other well trained models.
+As expected, the norm metrics can indeed distinguish among ``good, better, best'' models
+among a series well trained models.
 
@@
     %\qquad
-    \subfigure[Log Spectral Norm]{
+    \subfigure[log Spectral Norm]{
         %\includegraphics[width=4.5cm]{img/GPT2*_all_spectralnorm_hist.png}
@@
     %\qquad
-    \subfigure[Log Alpha Norm]{
+    \subfigure[log Alpha Norm]{
         %\includegraphics[width=4.5cm]{img/GPT2_all_logpnorm_hist.png}
@@
     }
-    \caption{PL exponents ($\alpha$), Log Spectral norm, and Log Alpha norm for different size models in the GPT2 architecture series.
-(The Log Spectral norm histogram omits the first 2 layers, corresponding to the embedding layer, as these are normalized differently and have anamolously large values.)}
+    \caption{PL exponents ($\alpha$), log Spectral norm, and log Alpha norm for different size models in the GPT2 architecture series.
+(The log Spectral norm histogram omits the first 2 layers, corresponding to the embedding layer, as these are normalized differently and have anamolously large values.)}
     \label{fig:gpt2-histograms}

### 2020-02-03 all_models.tex
updating sec 6

@@
 
-Here, we use the Weightwatcher tool to analyze 466 pretrained computer vision models Pytorch.
-These image classification and segmentation models are pretrained on nine datasets, ImageNet-1K, and CIFAR-10, CIFAR-100, Street View House Numbers (SVHN), Caltech-UCSD Birds-200-2011 (CUB-200-2011), Pascal VOC2012, ADE20K, Cityscapes, and Common Objects in Context (COCO).
-The pretrained models and their accuracy metrics are summarized in the osmr github
+Here, we use the Weightwatcher tool to analyze over 100 pretrained computer vision (CV) models Pytorch, including image classification and segmentation models.
+They have been pretrained on nine datasets, ImageNet-1K, and CIFAR-10, CIFAR-100, Street View House Numbers (SVHN), Caltech-UCSD Birds-200-2011 (CUB-200-2011), Pascal VOC2012, ADE20K, Cityscapes, and Common Objects in Context (COCO).
+The pretrained models and their accuracy metrics are summarized in the osmr github \charles{add link}
 
@@
 In this paper, we propose that the Weightwatcher tool could be used to predict the trends in the generalization accuracy of deep neutral network without a test set.
-To test our proposition, we choose simple linear regression to analyze the relationship between the Weightwatcher metrics and the traditional accuracy metric obtained with a test set (we avoid polynomial regressions as they are more prone to overfitting and does not make economic sense).
-On the left-hand-side of regression, we have the Top1 errors, Top5 errors as reported for ImageNet-1K models, Error % as reported in CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy (Pix.Acc.,%) and Intersection-Over-Union (mIOU,%) as dependent variables.
-We then regress them individually on each of the Weightwatcher metrics, including  alpha, alpha_weighted, spectralnormlog, softranklog, lognorm, and logpnorm as defined in earlier section.
+To test our proposition, we choose simple linear regression to analyze the relationship between the Weightwatcher metrics and the traditional accuracy metric obtained with a test set.
+% (we avoid polynomial regressions as they are more prone to overfitting and does not make economic sense).
+We regress the metrics on the Top1 (and Top5) reported errors (as dependent variables).
+These include the Top5 errors for the ImageNet-1K mode, the percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and the Pixel accuracy (Pix.Acc.) and Intersection-Over-Union (mIOU) for XXX.
+We regress them individually on each of the Weightwatcher log Norm metrics, as described above.
 
@@
 \serena{Define and give an example of the structural outliers}
-
-\charles{What are the structrual outliers we chose ?}
 

### 2020-02-03 all_models.tex
updating sec 6

@@
 
-To further refine our analysis, we run three batches of linear regressions.
+%To further refine our analysis, we run three batches of linear regressions.
 First at the global level, we divide models by datasets and run regression separately on all models of a certain dataset, regardless of the architecture.
@@
 
-\serena{see examples in Figure X}
+%\serena{see examples in Figure X}
 
-(Here we omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets as there are less than 15 models for those datasets and thus the regression is less statistically significant)
+%(Here we omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets as there are less than 15 models for those datasets and thus the regression is less statistically significant)
 
-\serena{insert plots for Figure X}
+%\serena{insert plots for Figure X}
 
-For the second batch, we plot the regression for models of each architecture-datasets combination, which shows the relationship between the progression of the model accuracy and Weightwatcher metrics more clearly and precisely.
+%e compite the regression R-squared and RMSE  for models of each architecture-datasets combination, which shows the relationship between the progression of the model accuracy and Weightwatcher metrics more clearly and precisely.
 For example, as you could see in the Figure X2,
 
-\serena{Add an example, UPDATE when we have the results from the new codes}
+%\serena{Add an example, UPDATE when we have the results from the new codes}
 
-Insert plots for Figure X2
-
-While running each regression, we record the R-squared and mean squared errors (MSE) for each regression.
-We then filter out regressions with less than five datapoints and models with structural outliers.
-
-\serena{Define and give an example of the structural outliers}
-
-Insert a plot for outliers, OPTIONAL
+For each regression, we record the R-squared and mean squared errors (MSE). We filter out regressions with less than five datapoints,
+ \nred{and models with structural outliers (where the trend is anti-correlated?) ?}
 
@@
 %    & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\langle\hat{\alpha}=\alpha\log\lambda_{max}\rangle$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
-        & Log                   & Log                        & Weighted       & Log                                 \\
-        & $\Vert\cdot\Vert_{F}$ & $\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
- Series & Metric                & Metric                     & Metric         & Metric                              \\
+Series        & $\log\Vert\cdot\Vert_{F}$ & $\log\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
 \hline

### 2020-02-05 abstract_file.tex
typo in ad

@@
 %
-Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision (ZCV) and natural language processing (NLP)
+Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision (CV) and natural language processing (NLP)
 We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can discriminate amoing ``good-better-best''  models.

### 2020-02-06 all_models.tex
move things to appendix for charles to do

@@
 
-\charles{We actually don't run regressions on all these datasetsm, BUT we could present them in the Figure below to show that alpha is a good metric for these kinds of models, in contrast to the NLP, where alpha is frequently too large to run a regression}
-
-\begin{figure}[t]
-    \centering
-    \subfigure[ImageNet 1K]{
-        \includegraphics[width=4cm]{img/imagenet1k_alpha.png}
-        \label{fig:imagenet1k-alpha}
-    }
-    \qquad
-    \subfigure[ CIFAR 10 ]{
-        \includegraphics[width=4cm]{img/cifar10_alpha.png}
-        \label{fig:cifar10.alpha}
-    }
-    \qquad
-    \subfigure[ CIFAR 100 ]{
-        \includegraphics[width=4cm]{img/cifar100_alpha.png}
-        \label{fig:cifar100.alpha}
-    }
-    \qquad
-    \subfigure[ SVHN ]{
-        \includegraphics[width=4cm]{img/svhn_alpha.png}
-        \label{fig:svhn.alpha}
-    }
-    \qquad
-    \subfigure[ CUB 200 ]{
-        \includegraphics[width=4cm]{img/cub200_alpha.png}
-        \label{fig:cub200.alpha}
-    }
-    \caption{\charles{Preliminary charts:} Heavy Tailed Power Law exponent $\alpha$ vs. reported Top1 Test Accuracies for pretrained DNNs available\charles{ref} for 5 different data sets.}
-
-    \label{fig:dataset_alphas}
-\end{figure}
-
-\charles{This section needs a lot of work}
-
-\serena{insert link in the footnote}
-
-and a full summary of all the models analyzed is included in the Appendix. For our analysis, we then group models by architecture and datasets for further analysis.
+We provide more details in Appendix~\ref{sxn:appendix}.
 
@@
 
-%To further refine our analysis, we run three batches of linear regressions.
-First at the global level, we divide models by datasets and run regression separately on all models of a certain dataset, regardless of the architecture.
-At this level, the plots are quite noisy and clustered as each architecture has its own accuracy trend but, you could still see that most plots show positive relationship with positive coefficients
-
-%\serena{see examples in Figure X}
-
-%(Here we omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets as there are less than 15 models for those datasets and thus the regression is less statistically significant)
-
-%\serena{insert plots for Figure X}
-
-%e compite the regression R-squared and RMSE  for models of each architecture-datasets combination, which shows the relationship between the progression of the model accuracy and Weightwatcher metrics more clearly and precisely.
-For example, as you could see in the Figure X2,
-
-%\serena{Add an example, UPDATE when we have the results from the new codes}
-
-For each regression, we record the R-squared and mean squared errors (MSE). We filter out regressions with less than five datapoints,
- \nred{and models with structural outliers (where the trend is anti-correlated?) ?}
-
-\charles{These tables could go to appendix. We need references}
-
-\begin{table}[t]
-\small
-\begin{center}
-\begin{tabular}{|p{1in}|c|}
-\hline
-Dataset & $\#$ of Models \\
-\hline
-imagenet-1k   &  78 \\
-svhn          &  30 \\
-cifar-100     &  30 \\
-cifar-10      &  18 \\
-cub-200-2011  &  12 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{Datasets used}
-\label{table:datasets}
-\end{table}
-
-\begin{table}[t]
-\small
-\begin{center}
-\begin{tabular}{|p{2in}|c|}
-\hline
-Architecture & $\#$ of Models \\
-\hline
-ResNet                                     & 30 \\
-SENet/SE-ResNet/SE-PreResNet/SE-ResNeXt    & 24 \\
-DIA-ResNet/DIA-PreResNet                   & 18 \\
-ResNeXt                                    & 12 \\
-WRN                                        & 12 \\
-DLA                                        & 6 \\
-PreResNet                                  & 6 \\
-ProxylessNAS                               & 6 \\
-VGG/BN-VGG                                 & 6 \\
-IGCV3                                      & 6 \\
-EfficientNet                               & 6 \\
-SqueezeNext/SqNxt                          & 6 \\
-ShuffleNet                                 & 6 \\
-DRN-C/DRN-D                                & 6 \\
-ESPNetv2                                   & 6 \\
-HRNet                                      & 6 \\
-SqueezeNet/SqueezeResNet                   & 6 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{Architectures used}
-\label{table:architectures}
-\end{table}
-
-\charles{Still preliminary, only 309 data points here}
-\paragraph{Results}
 \begin{table}[t]

### 2020-02-06 appendix.tex
move things to appendix for charles to do

@@
 \section{Appendix}
+\label{sxn:appendix}
 
-\subsection{XXX}
+In this appendix, we provide more details on several issues that are important for reproducibility of our results.
+
+\subsection{Convolutional layers and normalization issues}
+
+\paragraph{Spectral Analysis of Convolutional 2D Layers.}
+
+There is some ambiguity in performing spectral analysis on Convolutional 2D (Conv2D) layers.  
+A Conv2D layer can be represented as a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ filter (or kernel) and $in\_ch$ / $out\_ch$ input / output channels, respectively (usually $in\_ch\le out\_ch$).
+Typically, $w=h=k$,  giving $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\_ch\times out\_ch)$ each. 
+%
+There are at least three different approaches that have been advocated for applying the Singular Values Decomposition (SVD) to an Conv2D layer:
+run an SVD on each of the pre-Activation Maps $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values; 
+stack the feature maps into a single rectangular matrix of, say, dimension $((k\times k\times out\_ch)\times in\_ch)$, yielding $in\_ch$ singular values;
+compute the 2D Fourier Transform (FFT) for each of the $(in\_ch, out\_ch)$ pairs, and run SVD on the resulting Fourier coeffients~\cite{Long2019}, leading to $\sim(k\times in\_ch\times out\_ch)$ non-zero singular values.
+Each method has tradeoffs.  
+In principle, the third method is mathematically sound, but it is computationally expensive. 
+For our analysis, because we are performing tens of thousands of calculations, we select the first method, which is numerically the fastest and is easiest to reproduce.%
+\footnote{We provide a Google Colab notebook where all results can be reproduced, with the option to redo the calculations with the third option for the SVD of the Conv2D.}
+
+\paragraph{Normalization of Empirical Matrices.}  
+Normalization is an important, if underappreciated, practical issue.
+Importantly, the normalization of weight matrices does \emph{not} affect the Power Law fits because the Heavy Tailed exponent $\alpha$ (as well as other metrics such as the Stable Rank and MP Soft Rank~\cite{MM18_TR,MM19_HTSR_ICML}) is scale-invariant.
+Norm-based metrics, however, do depend strongly on the scale of the weight matrix--\nred{that's the point.}
+\nred{Indeed, early theoretical work by Bartlett suggests that the test accuracy depends strongly on the ``total size'' of the weight matrics.}
+Typically, to apply RMT, we would usually define Correlation Matrix with $1/N$ normalization and assume that the variance of $\mathbf{X}$ is either unity or a known constant.% 
+\footnote{For Heavy Tailed theorems, one typically needs a normalization such as \nred{$1/N^{\alpha-1}$. check this}}
+Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with the variance already normalized to $\sqrt{1/N}$, or some variant of this, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.
+We do not have conrol over the final empirical normalization of these models; and we do \emph{not} normalize (or renormalize) the Empirical Correlation Matrices, i.e., we use them as-is.
+The only exception to this is that we do rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $k/\sqrt{2}$ so that they are on the same scale as the Linear layers.
+
+\subsection{Reproducing results on hundreds of models}
+
+\charles{This section needs a lot of work}
+
+\serena{insert link in the footnote}
+
+and a full summary of all the models analyzed is included in the Appendix. For our analysis, we then group models by architecture and datasets for further analysis.
+
+%To further refine our analysis, we run three batches of linear regressions.
+First at the global level, we divide models by datasets and run regression separately on all models of a certain dataset, regardless of the architecture.
+At this level, the plots are quite noisy and clustered as each architecture has its own accuracy trend but, you could still see that most plots show positive relationship with positive coefficients
+
+%\serena{see examples in Figure X}
+
+%(Here we omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets as there are less than 15 models for those datasets and thus the regression is less statistically significant)
+
+%\serena{insert plots for Figure X}
+
+%e compite the regression R-squared and RMSE  for models of each architecture-datasets combination, which shows the relationship between the progression of the model accuracy and Weightwatcher metrics more clearly and precisely.
+For example, as you could see in the Figure X2,
+
+%\serena{Add an example, UPDATE when we have the results from the new codes}
+
+For each regression, we record the R-squared and mean squared errors (MSE). We filter out regressions with less than five datapoints,
+ \nred{and models with structural outliers (where the trend is anti-correlated?) ?}
+
+\charles{Still preliminary, only 309 data points here}
+\paragraph{Results}
+
+\charles{These tables could go to appendix. We need references}
+
+\begin{table}[t]
+\small
+\begin{center}
+\begin{tabular}{|p{1in}|c|}
+\hline
+Dataset & $\#$ of Models \\
+\hline
+imagenet-1k   &  78 \\
+svhn          &  30 \\
+cifar-100     &  30 \\
+cifar-10      &  18 \\
+cub-200-2011  &  12 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Datasets used}
+\label{table:datasets}
+\end{table}
+
+\begin{table}[t]
+\small
+\begin{center}
+\begin{tabular}{|p{2in}|c|}
+\hline
+Architecture & $\#$ of Models \\
+\hline
+ResNet                                     & 30 \\
+SENet/SE-ResNet/SE-PreResNet/SE-ResNeXt    & 24 \\
+DIA-ResNet/DIA-PreResNet                   & 18 \\
+ResNeXt                                    & 12 \\
+WRN                                        & 12 \\
+DLA                                        & 6 \\
+PreResNet                                  & 6 \\
+ProxylessNAS                               & 6 \\
+VGG/BN-VGG                                 & 6 \\
+IGCV3                                      & 6 \\
+EfficientNet                               & 6 \\
+SqueezeNext/SqNxt                          & 6 \\
+ShuffleNet                                 & 6 \\
+DRN-C/DRN-D                                & 6 \\
+ESPNetv2                                   & 6 \\
+HRNet                                      & 6 \\
+SqueezeNet/SqueezeResNet                   & 6 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Architectures used}
+\label{table:architectures}
+\end{table}
+
+\charles{We actually don't run regressions on all these datasetsm, BUT we could present them in the Figure below to show that alpha is a good metric for these kinds of models, in contrast to the NLP, where alpha is frequently too large to run a regression}
+
+\begin{figure}[t]
+    \centering
+    \subfigure[ImageNet 1K]{
+        \includegraphics[width=2.5cm]{img/imagenet1k_alpha.png}
+        \label{fig:imagenet1k-alpha}
+    }
+    \qquad
+    \subfigure[ CIFAR 10 ]{
+        \includegraphics[width=2.5cm]{img/cifar10_alpha.png}
+        \label{fig:cifar10.alpha}
+    }
+    \qquad
+    \subfigure[ CIFAR 100 ]{
+        \includegraphics[width=2.5cm]{img/cifar100_alpha.png}
+        \label{fig:cifar100.alpha}
+    }
+    \qquad
+    \subfigure[ SVHN ]{
+        \includegraphics[width=2.5cm]{img/svhn_alpha.png}
+        \label{fig:svhn.alpha}
+    }
+    \qquad
+    \subfigure[ CUB 200 ]{
+        \includegraphics[width=2.5cm]{img/cub200_alpha.png}
+        \label{fig:cub200.alpha}
+    }
+    \caption{\charles{Preliminary charts:} Heavy Tailed Power Law exponent $\alpha$ vs. reported Top1 Test Accuracies for pretrained DNNs available\charles{ref} for 5 different data sets.}
+
+    \label{fig:dataset_alphas}
+\end{figure}
+
+\section{XXX: PLACEHOLDER STUFF PROBABLY TO BE REMOVED}
+
+\subsection{XXX: PLACEHOLDER STUFF PROBABLY TO BE REMOVED}
 
@@
 
-\subsection{XXX}
+\subsection{XXX: PLACEHOLDER STUFF PROBABLY TO BE REMOVED}
 

### 2020-02-06 methods.tex
move things to appendix for charles to do

@@
 
-\paragraph{Spectral Analysis of Convolutional 2D Layers.}
-
-There is some ambiguity in performing spectral analysis on Convolutional 2D (Conv2D) layers.  
-A Conv2D layer can be represented as a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ filter (or kernel) and $in\_ch$ / $out\_ch$ input / output channels, respectively (usually $in\_ch\le out\_ch$).
-Typically, $w=h=k$,  giving $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\_ch\times out\_ch)$ each. 
-%
-There are at least three different approaches that have been advocated for applying the Singular Values Decomposition (SVD) to an Conv2D layer:
-run an SVD on each of the pre-Activation Maps $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values; 
-stack the feature maps into a single rectangular matrix of, say, dimension $((k\times k\times out\_ch)\times in\_ch)$, yielding $in\_ch$ singular values;
-compute the 2D Fourier Transform (FFT) for each of the $(in\_ch, out\_ch)$ pairs, and run SVD on the resulting Fourier coeffients~\cite{Long2019}, leading to $\sim(k\times in\_ch\times out\_ch)$ non-zero singular values.
-Each method has tradeoffs.  
-In principle, the third method is mathematically sound, but it is computationally expensive. 
-For our analysis, because we are performing tens of thousands of calculations, we select the first method, which is numerically the fastest and is easiest to reproduce.%
-\footnote{We provide a Google Colab notebook where all results can be reproduced, with the option to redo the calculations with the third option for the SVD of the Conv2D.}
-
-\paragraph{Normalization of Empirical Matrices.}  
-Normalization is an important, if underappreciated, practical issue.
-Importantly, the normalization of weight matrices does \emph{not} affect the Power Law fits because the Heavy Tailed exponent $\alpha$ (as well as other metrics such as the Stable Rank and MP Soft Rank~\cite{MM18_TR,MM19_HTSR_ICML}) is scale-invariant.
-Norm-based metrics, however, do depend strongly on the scale of the weight matrix--\nred{that's the point.}
-\nred{Indeed, early theoretical work by Bartlett suggests that the test accuracy depends strongly on the ``total size'' of the weight matrics.}
-Typically, to apply RMT, we would usually define Correlation Matrix with $1/N$ normalization and assume that the variance of $\mathbf{X}$ is either unity or a known constant.% 
-\footnote{For Heavy Tailed theorems, one typically needs a normalization such as \nred{$1/N^{\alpha-1}$. check this}}
-Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with the variance already normalized to $\sqrt{1/N}$, or some variant of this, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.
-We do not have conrol over the final empirical normalization of these models; and we do \emph{not} normalize (or renormalize) the Empirical Correlation Matrices, i.e., we use them as-is.
-The only exception to this is that we do rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $k/\sqrt{2}$ so that they are on the same scale as the Linear layers.
+\paragraph{Convolutional Layers and Normalization issues.}
+There are several technical issues (regarding spectral analysis of convolutional layers and normalization of empirical matrices) that are important for reproducibility of our results.
+See Appendix \ref{sxn:appendix} for a discussion.

### 2020-02-06 appendix.tex
few typos in appedix

@@
 Normalization is an important, if underappreciated, practical issue.
-Importantly, the normalization of weight matrices does \emph{not} affect the Power Law fits because the Heavy Tailed exponent $\alpha$ (as well as other metrics such as the Stable Rank and MP Soft Rank~\cite{MM18_TR,MM19_HTSR_ICML}) is scale-invariant.
+Importantly, the normalization of weight matrices does \emph{not} affect the PL fits because the PL exponent $\alpha$ (as well as other metrics such as the Stable Rank and MP Soft Rank~\cite{MM18_TR,MM19_HTSR_ICML}) is scale-invariant.
 Norm-based metrics, however, do depend strongly on the scale of the weight matrix--\nred{that's the point.}
@@
     }
-    \caption{\charles{Preliminary charts:} Heavy Tailed Power Law exponent $\alpha$ vs. reported Top1 Test Accuracies for pretrained DNNs available\charles{ref} for 5 different data sets.}
+    \caption{\charles{Preliminary charts:} PL exponent $\alpha$ vs. reported Top1 Test Accuracies for pretrained DNNs available\charles{ref} for 5 different data sets.}
 

### 2020-02-07 abstract_file.tex
pass after we talked, must reread sec 5, and I didnot do sec 6

@@
-In many practical AI applications, one works with deep neural network (DNN) models trained by someone else.
+In many practical applications, one works with deep neural networks (DNNs) models trained by someone else.
 For such \emph{pretrained models}, one typically does not have access to training data or test data, and, moreover,  does not know the details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
@@
 %%The most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in a system) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
-The most promising are norm-based metrics (such as those used to provide capacity control in traditional statistical learning theory) and metrics based on fitting eigenvalue distributions to truncated power law (PL) distributions (which derive from statistical mechanics, in particular the recently-developed Theory of Heavy-Tailed Self Regularization, and characterize the strength of correlations in a DNN).
+The most promising are norm-based metrics (such as those used to provide capacity control in traditional statistical learning theory) and metrics based on fitting eigenvalue distributions to truncated power law (PL) distributions (which derive from statistical mechanics, in particular the recently-developed Theory of Heavy-Tailed Self Regularization, and which characterize the strength of correlations in a DNN).
 %
-Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision (CV) and natural language processing (NLP)
+Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision (CV) and natural language processing (NLP).
 We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can discriminate amoing ``good-better-best''  models.
-On the other hand, for poorly trained models--which, arguably why we need quality metrics-- we want to distinguish which and when norm-based metrics can distiungush ''good-vs-bad'' models.
-We also find that PL-based metrics do much better, quantitatively better, at discriminating aming ''good-better-best'' models, and qualitatively better at discriminating ''good-vs-bad'' models.
+On the other hand, for poorly trained models, i.e., to distinguish ``good-versus-bad'' models---which, arguably is the point of needing quality metrics---norm-based metrics can qualitatively fail.  
+We also find that PL-based metrics do much better---quantitatively better at discriminating aming ``good-better-best'' models, and qualitatively better at discriminating ``good-versus-bad'' models.
 PL-based metrics can also be used to characterize fine-scale properties of models, e.g., understanding layer-wise \emph{correlation flow}, and evaluate post-training modifications such as model distillation, increasing the data set size, and other model improvements.

### 2020-02-07 cv_models.tex
pass after we talked, must reread sec 5, and I didnot do sec 6

--- 2020-02-02 cv_models.tex
@@
 We have examined the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
-To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BatchNormalization), and it plots the four quality metrics versus the reported Test accuracies~\cite{pyTorchVgg}, as well as a basic linear regression line.
+To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BatchNormalization), and it plots the four quality metrics versus the \emph{reported} Test accuracies, as well as a basic linear regression line.%
+\footnote{We emphasize that these \emph{reported} accuracies were previously computed and reported in a publicly-available forum by others~\cite{pyTorchVgg}, they were \emph{not} computed/recomputed by~us.}
 %with the Root Mean Squared Error (RMSE) shown, \michael{XXX.  WHERE IS RMSE.  IN THAT TABLE?}
@@
 While all four metrics perform well, notice that the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) performs best (with an RMSE of $0.42$, see Table~\ref{table:cv-models}); and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $), which is an approximation to the Log $\alpha$-Norm metric~\cite{MM20_unpub_work}, performs second best (with an RMSE of $0.48$, see Table~\ref{table:cv-models}).
-
-\begin{figure}[t]
-    \centering
-    \subfigure[ Frobenius Norm, VGG ]{
-        %\includegraphics[width=5cm]{img/VGG_lognorm_accs.png}
-        \includegraphics[width=4.0cm]{img/VGG_lognorm_accs.png}
-        \label{fig:vgg-fnorm}
-    }
-    %\qquad
-    \subfigure[ Spectral Norm, VGG ]{
-        %\includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
-        \includegraphics[width=4.0cm]{img/VGG_spectralnorm_accs.png}
-        \label{fig:vgg-snorm}
-    }
-    %\qquad
-    \subfigure[ Weighted Alpha, VGG ]{
-        %\includegraphics[width=4.9cm]{img/VGG_alpha_weighted_accs.png}
-        \includegraphics[width=4.0cm]{img/VGG_alpha_weighted_accs.png}
-        \label{fig:vgg-walpha}
-    }
-    %\qquad
-    \subfigure[ $\alpha$-Norm, VGG ]{
-        %\includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
-        \includegraphics[width=4.0cm]{img/VGG_logpnorm_accs.png}
-        \label{fig:vgg-pnorm}
-    }
-    \caption{Comparison of quality metrics versus reported test accuracy for pretrained VGG models, trained on ImageNet, available in pyTorch.  
-             XXX Plots will be updated and replaced. 
-             \michael{Figures are unreadable, can we make fonts bigger, etc.}
-            }
-    \label{fig:vgg-metrics}
-\end{figure}
-
-\begin{figure}[t]
-    \centering
-
-    \subfigure[ ResNet, $\alpha$-Norm ]{
-        %\includegraphics[width=4.2cm]{img/ResNet_logpnorm_accs.png}
-        \includegraphics[width=4.0cm]{img/ResNet_logpnorm_accs.png}
-        \label{fig:resnet-accuracy}
-    }
-    %\qquad
-    \subfigure[ ResNet-1K, $\alpha$-Norm ]{
-        %\includegraphics[width=4.5cm]{img/ResNet-1K_logpnorm_accs.png}
-        \includegraphics[width=4.0cm]{img/ResNet-1K_logpnorm_accs.png}
-        \label{fig:resnet1k-accuracy}
-    }
-    %\qquad
-    \subfigure[ DenseNet, $\alpha$-Norm ]{
-        %\includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
-        \includegraphics[width=4.0cm]{img/DenseNet_logpnorm_accs.png}
-        \label{fig:densenet-accuracy}
-    }
-    \caption{$\alpha$-Norm versus reported Top1 Error for  ResNet, ResNet-1K, and DenseNet models. The corresponding results for VGG are shown in Figure~\ref{fig:vgg-pnorm}.  }
-    \label{fig:cv2-accuracy}
-\end{figure}
 
@@
 
-Figure~\ref{fig:cv2-accuracy} plots the 
-%overall best performing (i.e., 
-Log $\alpha$-Norm
-%) 
-metric for the full ResNet, ResNet-1K, and DenseNet series.
-(A much more detailed set of plots, including the behavior of all four methods on each of the series, is available at \cite{XXX-WEB-LINK}.)
-Here, the Log $\alpha$-Norm is much better that the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is actually slightly worse than the Weighted Alpha metric).
-Notice also that ResNet series, which has been trained on the full ImageNet dataset, has an RMSE of $0.66$, whereas the ResNet-1K series, which has been trained on the much smaller ImageNet-1K dataset, also correlated wells, but has a much larger RMSE of $1.9$.
-
-See Table~\ref{table:cv-models} for a summary of results for Top1 Accuracies for all four metrics on all four model series.
-Similar results (not shown) are obtained for the Top5 Accuracies.
-Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best. 
-These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among ``good-better-best'' models, with PL-based metrics performing somewhat better.
-
-\paragraph{Layer Analysis: Correlation Flow in CV Models.}
-
-We can learn much more about a pretrained model by going beyond average values to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).
-% in the network.
-%The most interesting results are seen when we 
-For example, we can 
-plot (just) the PL exponent, $\alpha$,%
-\footnote{That is, here we consider just $\alpha$ for each layer, i.e., not $\hat{\alpha}$ or $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$.} 
-as a function of depth.
-%
-See Figure~\ref{fig:3models-alpha-layers}, which plots $\alpha$ for each layer (the first layer corresponds to data, the last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (without BatchNormalization), ResNet-1K, and DenseNet series.
-(Again, much more detailed set of plots is available at \cite{XXX-WEB-LINK}; but note that the corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than the results we present here.)
-
 \begin{figure}[t]
     \centering
+    \subfigure[ Frobenius Norm, VGG ]{
+        %\includegraphics[width=5cm]{img/VGG_lognorm_accs.png}
+        \includegraphics[width=4.0cm]{img/VGG_lognorm_accs.png}
+        \label{fig:vgg-fnorm}
+    }
+    %\qquad
+    \subfigure[ Spectral Norm, VGG ]{
+        %\includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
+        \includegraphics[width=4.0cm]{img/VGG_spectralnorm_accs.png}
+        \label{fig:vgg-snorm}
+    }
+    %\qquad
+    \subfigure[ Weighted Alpha, VGG ]{
+        %\includegraphics[width=4.9cm]{img/VGG_alpha_weighted_accs.png}
+        \includegraphics[width=4.0cm]{img/VGG_alpha_weighted_accs.png}
+        \label{fig:vgg-walpha}
+    }
+    %\qquad
+    \subfigure[ $\alpha$-Norm, VGG ]{
+        %\includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
+        \includegraphics[width=4.0cm]{img/VGG_logpnorm_accs.png}
+        \label{fig:vgg-pnorm}
+    }
+    \caption{Comparison of quality metrics versus reported test accuracy for pretrained VGG models, trained on ImageNet, available in pyTorch.  
+             \michael{Figures are unreadable, can we make fonts bigger, etc.}
+            }
+    \label{fig:vgg-metrics}
+\end{figure}
 
+%MM%    \begin{figure}[t]
+%MM%        \centering
+%MM%        \subfigure[ ResNet, $\alpha$-Norm ]{
+%MM%            %\includegraphics[width=4.2cm]{img/ResNet_logpnorm_accs.png}
+%MM%            %\includegraphics[width=4.0cm]{img/ResNet_logpnorm_accs.png}
+%MM%            \includegraphics[width=2.6cm]{img/ResNet_logpnorm_accs.png}
+%MM%            \label{fig:resnet-accuracy}
+%MM%        }
+%MM%        %\qquad
+%MM%        \subfigure[ ResNet-1K, $\alpha$-Norm ]{
+%MM%            %\includegraphics[width=4.5cm]{img/ResNet-1K_logpnorm_accs.png}
+%MM%            %\includegraphics[width=4.0cm]{img/ResNet-1K_logpnorm_accs.png}
+%MM%            \includegraphics[width=2.6cm]{img/ResNet-1K_logpnorm_accs.png}
+%MM%            \label{fig:resnet1k-accuracy}
+%MM%        }
+%MM%        %\qquad
+%MM%        \subfigure[ DenseNet, $\alpha$-Norm ]{
+%MM%            %\includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
+%MM%            %\includegraphics[width=4.0cm]{img/DenseNet_logpnorm_accs.png}
+%MM%            \includegraphics[width=2.6cm]{img/DenseNet_logpnorm_accs.png}
+%MM%            \label{fig:densenet-accuracy}
+%MM%        }
+%MM%        \caption{$\alpha$-Norm versus reported Top1 Error for  ResNet, ResNet-1K, and DenseNet. Corresponding VGG results in Figure~\ref{fig:vgg-pnorm}.  }
+%MM%        \label{fig:cv2-accuracy}
+%MM%    \end{figure}
+%MM%    
+%MM%    \begin{figure}[t]
+%MM%        \centering
+%MM%        \subfigure[ VGG ]{
+%MM%            %\includegraphics[width=4.5cm]{img/VGG_fnl_alpha_depth.png} 
+%MM%            %\includegraphics[width=4.0cm]{img/VGG_fnl_alpha_depth.png} 
+%MM%            \includegraphics[width=2.6cm]{img/VGG_fnl_alpha_depth.png} 
+%MM%                    \label{fig:vgg-alpha-layers}
+%MM%        }
+%MM%        %\qquad
+%MM%        \subfigure[ ResNet-1K ]{
+%MM%            %\includegraphics[width=4.5cm]{img/ResNet_fnl_alpha_depth.png} 
+%MM%            %\includegraphics[width=4.0cm]{img/ResNet_fnl_alpha_depth.png} 
+%MM%            \includegraphics[width=2.6cm]{img/ResNet_fnl_alpha_depth.png} 
+%MM%            \label{fig:resnet-alpha-layer}
+%MM%        }
+%MM%        %\qquad
+%MM%        \subfigure[ DenseNet ]{
+%MM%            %\includegraphics[width=4.5cm]{img/DenseNet_fnl_alpha_depth.png} 
+%MM%            %\includegraphics[width=4.0cm]{img/DenseNet_fnl_alpha_depth.png} 
+%MM%            \includegraphics[width=2.6cm]{img/DenseNet_fnl_alpha_depth.png} 
+%MM%            \label{fig:densenet-alpha-layer}
+%MM%        }
+%MM%        \caption{Layer Analysis. 
+%MM%                 PL exponent $\alpha$ versus layer id, for VGG, ResNet-1K, and DenseNet models.  
+%MM%                 (Y axes on each plot are different.)  
+%MM%                 ResNet-1K exhibits different and much more stable behavior across layers.
+%MM%                 This contrasts with how VGG gets gradually worse in deeper layers and how DenseNet is worse and much more erratic.  
+%MM%                 In the text, we interpret this as quantifying a \emph{correlation flow} across the layers.  
+%MM%                }
+%MM%        \label{fig:3models-alpha-layers}
+%MM%    \end{figure}
+
+\begin{figure}
+\centering
+\begin{minipage}[b]{.22\textwidth}
+%<Code for the first figure>
+%MM%   \begin{figure}[t]
+%MM%       \centering
+       \subfigure[ ResNet, $\alpha$-Norm ]{
+           %\includegraphics[width=4.2cm]{img/ResNet_logpnorm_accs.png}
+           \includegraphics[width=4.0cm]{img/ResNet_logpnorm_accs.png}
+           \label{fig:resnet-accuracy}
+       }
+       %\qquad
+       \subfigure[ ResNet-1K, $\alpha$-Norm ]{
+           %\includegraphics[width=4.5cm]{img/ResNet-1K_logpnorm_accs.png}
+           \includegraphics[width=4.0cm]{img/ResNet-1K_logpnorm_accs.png}
+           \label{fig:resnet1k-accuracy}
+       }
+       %\qquad
+       \subfigure[ DenseNet, $\alpha$-Norm ]{
+           %\includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
+           \includegraphics[width=4.0cm]{img/DenseNet_logpnorm_accs.png}
+           \label{fig:densenet-accuracy}
+       }
+       \caption{$\alpha$-Norm versus reported Top1 Error for  ResNet, ResNet-1K, and DenseNet. Corresponding VGG results in Figure~\ref{fig:vgg-pnorm}.  }
+       \label{fig:cv2-accuracy}
+%MM%   \end{figure}
+%MM% \caption{Caption}\label{label-a}
+\end{minipage}\qquad
+\begin{minipage}[b]{.22\textwidth}
+%<Code for the second figure>
+%MM%    \begin{figure}[t]
+%MM%        \centering
        \subfigure[ VGG ]{
@@
                 (Y axes on each plot are different.)  
-             ResNet-1K exhibits different and much more stable behavior across layers.
-             This contrasts with how VGG gets gradually worse in deeper layers and how DenseNet is worse and much more erratic.  
-             We interpret this in the text as quantifying a \emph{correlation flow} across the layers.  
+%                ResNet-1K exhibits different and much more stable behavior across layers.
+%                This contrasts with how VGG gets gradually worse in deeper layers and how DenseNet is worse and much more erratic.  
+%                In the text, we interpret this as quantifying a \emph{correlation flow} across the layers.  
                }
        \label{fig:3models-alpha-layers}
+%MM%    \end{figure}
+%MM% \caption{Caption}\label{label-b}
+\end{minipage}
 \end{figure}
 
+Figure~\ref{fig:cv2-accuracy} plots the 
+%overall best performing (i.e., 
+Log $\alpha$-Norm
+%) 
+metric for the full ResNet, ResNet-1K, and DenseNet series.
+(A much more detailed set of plots, including the behavior of all four methods on each of the series, is available at \cite{XXX-WEB-LINK}.)
+Here, the Log $\alpha$-Norm metric is much better that the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is actually slightly worse than the Weighted Alpha metric).
+Also, the ResNet series, which has been trained on the full ImageNet dataset, has an RMSE of $0.66$, whereas the ResNet-1K series, which has been trained on the much smaller ImageNet-1K dataset, also correlated wells, but has a much larger RMSE of $1.9$.
+
+See Table~\ref{table:cv-models} for a summary of results for Top1 Accuracies for all four metrics on all four model series.
+Similar results (not shown) are obtained for the Top5 Accuracies.
+Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best. 
+These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among ``good-better-best'' models, with PL-based metrics performing somewhat better.
+
+\paragraph{Layer Analysis: Correlation Flow in CV Models.}
+
+We can learn much more about a pretrained model by going beyond average values to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).
+% in the network.
+%The most interesting results are seen when we 
+For example, we can 
+plot the PL exponent, $\alpha$,%
+\footnote{That is, here we consider just $\alpha$ for each layer, i.e., not $\hat{\alpha}$ or $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$.} 
+as a function of depth.
+%
+See Figure~\ref{fig:3models-alpha-layers}, which plots $\alpha$ for each layer (the first corresponds to data, the last to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (without BatchNormalization), ResNet-1K, and DenseNet series.
+(A much more detailed set of plots is available at \cite{XXX-WEB-LINK}.  The corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than these results.)
+
 In the VGG models, Figure~\ref{fig:vgg-alpha-layers} shows that the PL exponent $\alpha$ systematically increases as we move down the network, from data to labels, in the Conv2D layers, starting with $\alpha\lesssim 2.0$ and reaching all the way to $\alpha\sim 5.0$; and then, in the last three, large, fully-connected (FC) layers, $\alpha$ stabilizes back down to $\alpha\in[2,2.5]$.
-This is seen for all the VGG models (again, only the shallowest and deepest are shown in this figure), indicating that the main effect of increasing depth is to increase the range over which $\alpha$ increases, thus leading to larger $\alpha$ values in later Conv2D layers of the VGG models.
-This is quite different than the behavior of either the ResNet-1K models or the DenseNet models.
+This is seen for all the VGG models (again, only the shallowest and deepest are shown), indicating that the main effect of increasing depth is to increase the range over which $\alpha$ increases, thus leading to larger $\alpha$ values in later Conv2D layers of VGG models.
+This is quite different than the behavior of either the ResNet-1K or the DenseNet models.
 
@@
 
-Recall the differences---in particular, the number of residual connections---between the VGG, ResNet, and DenseNet architectures.
-VGG, while a fine model in many ways, was limited by needing massive FC layers at the end of the model, requiring significant memory and computational resources, and it's poor conditioning led to problems with vanishing gradients.
+To understand these results, recall the differences---in particular, the number of residual connections---between the VGG, ResNet, and DenseNet architectures.
+VGG, while a fine model in many ways, was limited by needing massive FC layers at the end, requiring significant memory and computational resources, and it's poor conditioning led to problems with vanishing gradients.
 The empirical manifestations of this (on weight matrices) are that fitted $\alpha$ values get much worse for deeper layers.
@@
 This should also be contrasted with the DenseNet model, which contains many connections between every layer.
-Our results (large $\alpha$, meaning they even a HT model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly optimized.
+Our results (large $\alpha$, meaning that even a HT model is a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly optimized.
 
@@
 More generally, do fitted $\alpha$ values contain information not captured by norms? 
-To show that it is not just a variation and that $\alpha$ does capture something different, consider the following example, which looks at a compressed / distilled DNN model~\cite{CWZZ17_TR}.
+To show that $\alpha$ is not just a variation and that it does capture something different, consider the following example, which looks at a compressed/distilled DNN model~\cite{CWZZ17_TR}.
 
@@
 The previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy} might suggest that the baseline Spectral Norm should be \emph{smaller} than those of the layers in the fine-tuned model.
-\michael{Do we have something backwards here: where we observed that the average norm to increase with decreasing test error (as it ``should'' not), whereas the average PL exponent $\alpha$ decreases (as ``expected'' from HT-SR Theory).
+\michael{FIX: Do we have something backwards here: where we observed that the average norm to increase with decreasing test error (as it ``should'' not), whereas the average PL exponent $\alpha$ decreases (as ``expected'' from HT-SR Theory).
 }
@@
 On the other hand, the $\alpha$ values do not differ systematically between the baseline and fine-tuned models.
-\michael{Some comment about big jump in norm; but fix previous backwards question first.}
 Also (not shown), the average (unweighted) baseline $\alpha$ is \emph{smaller} than the fine-tuned average (as would be predicted by HT-SR Theory, on which $\hat{\alpha}$ is based).
 
-\michael{XXX.  WORK ON THIS PAR AFTER DO NLP EXAMPLES.}
 Our interpretation of this is the following.
-The pre-trained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the VGG/ResNet/DenseNet models we condisered previously.
-Norms are often used as regularizers, and thys they are a natural metric to consider, e.g., during training or distillation.
-The PL $\alpha$ captures more subtle correlations.
+The pre-trained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than 
+%any of 
+the VGG/ResNet/DenseNet models we condisered previously.
+While norms often make good regularizers for a single model, there is no reason \emph{a priori} to expect them correlate so well with the test accuracies across different models (and so it is noteworthy when they do, but not surprising when the do not).
+We do expect, however, the PL $\alpha$ to correlate so well with the test accuracies across different models because it effectively measures the amount of correlation in the model.
 %
-The reason for this is that the \texttt{distiller} Group Regularization technique has the unusual effect of increasing the norms of the $\mathbf{W}$ feature maps for at least two of the Conv2D layers.
-\michael{This impedance shuffling or whatever it is called messes up norms, but the strong correlations are still preserved, as seen in the $\hat{\alpha}$ metric, and the model quality is good.
-CLARIFY.}
-\michael{XXX.  INTERPRET IN TERMS OF XXX GOOD VERSUS BAD, AS OPPOSED TO GOOD BETTER BEST.}
+The reason for this is that the \texttt{distiller} Group Regularization technique has the unusual effect of spuriously increasing the norms of the $\mathbf{W}$ pre-activation maps for at least two of the Conv2D layers.
+\michael{FIX: Where is that increase?  Or is that a typo?}
+In these, cases, the Spectral Norms of certain layers exhibit a form of \emph{scale collapse} and become anomalously small.

### 2020-02-07 methods.tex
pass after we talked, must reread sec 5, and I didnot do sec 6

@@
 
-Each DNN layer contains one or more layer 2D weight matrices $\mathbf{W}_{l}$, and/or 2D feature maps $\mathbf{W}_{i,l}$ extracted from 2D Convolutional layers. 
+Each DNN layer contains one or more layer 2D weight matrices, $\mathbf{W}_{l}$, and/or 2D feature maps, $\mathbf{W}_{i,l}$, extracted from 2D Convolutional layers. 
 (For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below; and we let $\mathbf{W}$ be a generic $N\times M$ weight matrix, with $N\ge M$.) 
@@
 \item 
-Weighted Alpha Metric: $\hat{\alpha}=\alpha\log\lambda_{max}$
+Weighted Alpha: $\hat{\alpha}=\alpha\log\lambda_{max}$
 \item 
@@
 $ %% $$
-and so $\lambda_{max}$ is the maximum eigenvalue of $\mathbf{X}$. 
+and $\lambda_{max}$ is the maximum eigenvalue of $\mathbf{X}$. 
 (These eigenvalues are the squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, i.e., $\lambda_{i}=\sigma^{2}_{i}$.)
@@
 For norm-based metrics, we use the average of the log norm to the appropriate power.
-For example, consider the \emph{$\alpha$-Norm metric}, 
+Consider, e.g., the \emph{$\alpha$-Shatten Norm metric}, 
 \begin{equation}
@@
 (When taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.)
-For the Weighted Alpha Norm, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it cannot be taken out of the sum.
+For the $\alpha$-Shatten Norm metric, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it cannot be taken out of the sum.
 
-For the Weighted Alpha Metric, we consider the average of $\hat{\alpha}$ over all layers, i.e., 
+For the \emph{Weighted Alpha metric}, we average of $\hat{\alpha}$ over all layers, i.e., 
 $ %$$
@@
 $ %$$
-\michael{Be careful to get that right.}
-The Weighted Alpha Metric was introduced previously~\cite{MM20_SDM}, where (on a much more limited set of data than we consider here) it was shown to correlate well with the trends in reported test accuracies of pretrained DNNs.
+\michael{Be careful to get right.}
+The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where (on a much more limited dataset than we consider here) it was shown to correlate well with trends in reported test accuracies of pretrained DNNs.
 Based on this, in this paper, we introduce and evaluate the $\alpha$-Norm metric.
 One would expect $\hat{\alpha}$ to approximate the log $\alpha$-Norm very well for $\alpha < 2$ and reasoably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.
+
+To avoid confusion, let us clarify the relationship between $\alpha$ and $\hat{\alpha}$.  
+We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated PL, parameterized by 2 values: the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.  
+(Technically, we also need the minimum eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.)
+The PL exponent $\alpha$ measures of the amount of correlation in a DNN layer weight matrix $\mathbf{W}$. 
+It is valid for $\lambda<\lambda_{max}$, and it is scale-invariant, i.e., it does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.
+The $\lambda_{max}$ is a measure of the size of $\mathbf{W}$.
+%
+Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric for the entire~DNN.  
+%%%
+%%For small values of $\alpha$, this Weighted Alpha metric approximates the log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}.
+%%%
+%%The Weighted Alpha metric and $\alpha$-Shatten norm metric also often behave like an improved, weighted log average Spectral Norm, and may track this metric in some cases.
 

### 2020-02-07 nlp_models.tex
pass after we talked, must reread sec 5, and I didnot do sec 6

@@
 %(in particular, the interpretation of values of $\alpha \sim 2$ as modeling systems with good correlations over many size scales~\cite{BouchaudPotters03, SornetteBook}), 
-this suggests that these models fail to successfully capture many of the correlations in the data (relative to their size) and thus are substantially \emph{under}-trained.
+this suggests that these models fail to successfully capture (relative to their size) many of the correlations in the data, and thus are substantially \emph{under}-trained.
 %
 More generally, compared to the CV models of Section~\ref{sxn:cv},
-modern NLP models have larger weight matrices and display different spectral properties, and 
+modern NLP models have larger weight matrices and display different spectral properties.
 %, both in terms of their architecture (e.g., wheter there are convolutions) as well as how they are used (e.g., whether one uses prediction error or perplexity).
-thus, they provide a very different test for our empirical quality metrics.
+Thus, they provide a very different test for our empirical quality metrics.
 
-Where Norm-base metrics perform reasonably well on well-trained NLP models, they behave anomalously poorly-trained models.
-Indeed, in ``bad models'', the weight matrics may display rank collapse, decreased Frobeinus mass, or unsually small Spectral norms, 
-which may be mis-interpreted as ``smaller is better.''
-In contast, PL-based metrics (the log $\alpha$-Norm metric, 
-$(\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha})$, and the Weighted Alpha metric, $\hat\alpha =\alpha\log\lambda_{max} $) display consistant behavior even on poorly trained models, and we can use these metrics to help indentify
-where the archiectures need repair, and/or more and/or better data is needed.
+Overall, our results indicate that norm-based metrics perform reasonably-well on NLP models that are reasonably well-trained, but that they behave anomalously on poorly-trained models.
+Indeed, for these ``bad'' models, weight matrics may display rank collapse, unsually small Spectral norms, or decreased Frobeinus mass. 
+This may be mis-interpreted as ``smaller is better.''
+In contast, the PL-based metrics (the log $\alpha$-Norm metric, 
+$\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$, and the Weighted Alpha metric, $\hat\alpha =\alpha\log\lambda_{max} $) display consistant behavior, both when comparing more-or-less well-trained models as well as when looking at poorly-trained or ``bad'' models.
+Indeed, we can use these metrics to help indentify where the archiectures need repair, and/or more and/or better data is needed.
 
@@
 
-\michael{WHERE: Say large $\alpha$ values don't mean much.}
-We do note, however, that many NLP models, such as the GPT and BERT, may have weight matrices with unusually large PL exponents
-($\alpha>6$), which indicates these matrices may be \emph{under}-correlated (i.e over-paramaterized), although the truncated PL fit
-itself may not be very reliable because the MLE estimator is unreliable in this range.  Phenomenologically, if we examine
-the ESD visually, we can usually describe these $\mathbf{W}$ as in the \emph{Bulk-Decay} or \emph{Bulk-plus-Spikes} phase\cite{MM}.
-We have conjectured previously that very well-trained DNNs would not have many \emph{outlier} $\alpha>6$, 
-and improved versions of GPT (shown below, and BERT (not shown) confirm this.
-
-\michael{WHERE: Highlight difference between $\alpha$ and $\hat{\alpha}$.}
-To avoid confusion, let us clarify the differences between $\alpha$ and $\hat{\alpha}$.  
-We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated power law (PL), parameterized by 2 values:
-the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.  (Technically, we also need the minimum
-eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.)
-The PL exponent $\alpha$ measures of the 
-amount of correlation in a DNN layer weight matrix $\mathbf{W}$.  It is valid for $\lambda<\lambda_{max}$, and
-it scale invariant (does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.)
-The Weighted-Alpha metric, $\hat{\alpha}=\alpha\log\lambda_{max}$, approximates the log $\alpha$ (Shatten) Norm 
-$(\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha})$ (derived using statistical mechanics and RMT \nred{(in progress)}, 
-and lets us compute a balanced, weighted average log PL norm metrix for the entire DNN.  
-The $\hat{\alpha}$ metric also behaves like an imrpoved, weighted log average Spectral Norm, and may track this metric in some cases.
+Our results also indicate that many NLP models, such as GPT and BERT, also have a few weight matrices with unusually large PL exponents ($\alpha>6$).
+This indicates these matrices (and thus these models) may be \emph{under}-correlated (i.e, under-trained or over-paramaterized).
+Of course, in this regime, the specific values returned by the truncated PL fits are less reliable because the MLE estimator is unreliable in this range. 
+Phenomenologically, if we examine the ESD visually, we can usually describe these $\mathbf{W}$ as in the \emph{Bulk-Decay} or \emph{Bulk-plus-Spikes} phase (from HT-SR Theory)~\cite{MM18_TR,MM19_HTSR_ICML}.
+HT-SR Theory predicts that very well-trained DNNs sould not have many \emph{outlier} $\alpha>6$, and improved versions of GPT (shown below) and BERT (not shown) confirm this.
 
@@
 These models have the remarkable ability to generate fake text that appears to the human to be real, and they have generated significant media attention because of the potential for their misuse.
-For this reason, the original GPT model released by OpenAI was trained on on a deficient data set, rendering the model interesting but not fully functional.  
-Later, OpenAI released a much improved model, GPT2 (small), which has the same architecture and number of layers as GPT, but which has been trained on a larger and better data set (and with other changes), making it remarkably good at generating (near) human-quality fake text.
+For this reason, the original GPT model released by OpenAI was trained on a deficient data set, rendering the model interesting but not fully functional (i.e., ``bad'').  
+Later, OpenAI released a much improved model, GPT2-small, which has the same architecture and number of layers as GPT, but which has been trained on a larger and better data set (and with other changes), making it remarkably good.
 %
-By comparing the poorly-trained GPT to the well-trained GPT2, we can indentify empirical indidcators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
+By comparing poorly-trained GPT to the well-trained GPT2 models, we can indentify empirical indidcators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
 
-\charles{More details here.}
 The GPT models we analyze are deployed with the popular HuggingFace PyTorch library~\cite{XXX-XXX}.
@@
 %
-The OpenAI GPT2 (English) models are: \nred{GPT2-small, GPT2-medium, GPT2-large, and GPT2-xl}, \
-having include $12, 24, 36, \text{and }48$ layers, respectively, with increasingly larger weight matrices.
+The OpenAI GPT2 (English) models are: GPT2-small, GPT2-medium, GPT2-large, and GPT2-xl, \
+having $12, 24, 36, \text{and }48$ layers, respectively, with increasingly larger weight matrices.
 The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
@@
 GPT2-small & 49 & 2.04  & 2.54& 9.62 & 9.87 \\
+\hline
 GPT2 medium & 98 & 2.08 & 2.58& 9.74 & 10.01 \\
@@
 \end{center}
-\caption{Average value for the Average log Norm metrics and the Weighted Alpha metric for pretrainnd OpenAI GPT and GPT2 models.  }
+\caption{Average values of Average log Norm and Weighted Alpha metrics for pretrainnd OpenAI GPT/GPT2~models.}
 \label{table:nlp}
@@
 Observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
-Observe also that (with one minor exception involving the log Frobenius norm metric) all four metrics decrease as one goes from GPT2 medium to large to xl, indicating that the larger models indeed look better than the smaller models.
+\michael{FIX: typo in that last sentence?}
+Observe also that (with one minor exception involving the log Frobenius norm metric) all four metrics decrease as one goes from GPT2-medium to GPT2-large to GPT2-xl, indicating that the larger models indeed look better than the smaller~models.
 %(The one minor exception is the log Frobenius norm increasing from $1.85$ to $1.86$, for GPT2 large to GPT2 xl.)
 
-Figure~\ref{fig:GPT-alpha-hist} shows the histogram (empirical density), for all layers, of $\alpha$ for GPT (blue) and this GPT2 (red) model.  
-These two histograms are very different, with GPT2 having both a notably smaller mean $\alpha$, and far fewer unusually-large outlying $\alpha$ values.
-
-From this, we see that $\alpha$ provides a good quality metric for comparing these two models, the``bad'' GPT vs. the ``good'' GPT2.
-The deficient GPT has numerous unsually large $\alpha$ exponents--meaning they are not Heavy Tailed at all.
-Indeed, we expect that a poorly trained model lack Heavy Tailed behavior in all layers.
-On the other hand, as expected, the improved GPT2 model has, on average, smaller $\alpha$ than the older GPT, with all $\alpha\le6$.
+Figure~\ref{fig:GPT-alpha-hist} shows the histogram (empirical density), for all layers, of $\alpha$ for GPT (blue) and GPT2-small (red).  
+These two histograms are very different, with GPT2-small having both a notably smaller mean/median $\alpha$, and also far fewer unusually-large outlying $\alpha$ values than GPT.
+The deficient GPT has numerous unsually large $\alpha$ exponents---meaning they are not well-described by PL.
+Indeed, we expect that a poorly-trained model lack PL behavior in most layers.
+On the other hand, as expected, the improved GPT2 model has, on average, smaller $\alpha$ than the older GPT, with all $\alpha\le6$ and with smaller mean/median $\alpha$.  
+From this (and other results not shown), we see that $\alpha$ provides a good quality metric for comparing these two models, the``bad'' GPT versus the ``good'' GPT2-small.
 
@@
        }
-   \caption{Comparison of PL exponents, $\alpha$, and log Spectral Norms, 
-$(\log\Vert\mathbf{W}\Vert_{\infty})$, for the OpenAI GPT and GPT2 (small) pretrained models.}
+   \caption{Comparison of PL exponents, $\alpha$, and log Spectral Norm, $\log\Vert\mathbf{W}\Vert_{\infty}$, for OpenAI GPT and GPT2-small pretrained models.}
 \label{fig:GPT-hist}
@@
 
-\paragraph{Scale Collapse.}
-However, as seen in Figure \ref{fig:GPT-snorm-hist},
-the ``bad'' GPT model has, spuriously, many smaller Spectral Norms $(\log\Vert\mathbf{W}\Vert_{\infty})$,
-compared to the ``good'' GPT2--which violates the belief that smaller Spectral Norms are always better.
-Indeed, because there are so many anonymously small $\Vert\mathbf{W}\Vert_{\infty}$,
-it appears that the GPT model may be exhibiting a kind of \emph{scale collapse} and/or undesired
-loss of Frobeinus mass.
-This is an extremely important observation because it demonstrates that while the Spectral Norm
-may correlate well with predicted test error, it is not a good indicator of the overall quality of a model,
-and using it as an empirical metric may give spurious results when applied to poorly trained
-or otherwise deficient models.
+\paragraph{Coarse Analysis: Scale Collapse.}
 
-Note that Figure \ref{fig:GPT-snorm-hist} also shows some unusually large Spectral Norms, but,
-here, simply correspond to the change in normalization of the word embedding layer(s) and can be ignored.
-From Figure \ref{fig:gpt-snorm-layer} (below), we see that these correspond to the first embedding layer(s).
-These layers appear to have a different effectvie normalization, and therefore a different scale.
-For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,
-but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
- $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we treat all layers as-is (although one may to normalize
-the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier).
-Here, we do not include them in our computed average metrics in Table \ref{table:nlp},
-and do not include them in the histogram plot in Figure \ref{fig:GPT-snorm-hist}.
+This should be contrasted with the behavior displayed by the Frobenius norm (not shown) and the Spectral norm.
+See Figure~\ref{fig:GPT-snorm-hist}, which shows that the ``bad'' GPT model has, spuriously, a smaller mean/median Spectral norm as well as many much smaller Spectral Norms, compared to the ``good'' GPT2-small, violating the conventional wisdom that smaller Spectral Norms are better.
+Indeed, because there are so many anonymously small Spectral norms, it appears that the GPT model may be exhibiting a kind of \emph{scale collapse} and/or related undesired loss of Frobeinus mass.
+This is important because it demonstrates that, while the Spectral Norm may correlate well with predicted test error, it is \emph{not} a good indicator of the overall quality of a model---it can mis-predict good-versus-bad questions in ways not seen with PL-based metrics.
+Using it as an empirical quality metric may give spurious results when applied to poorly-trained or otherwise deficient models.
 
-\paragraph{Layer Analysis: Correlation Flow in GPT and GPT2.}
+We should note that these models have a few layers with unusually large Spectral Norms.
+From Figure~\ref{fig:gpt-snorm-layer} (below), we see that these correspond to the first embedding layer(s).
+These layers appear to have a different effective normalization, simply corresponding to the change in normalization of the word embedding layer(s), and therefore a different scale.
+For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$, but in the first embedding layer, the maximum is of order $N$ (the number of words in the embedding), or $\lambda_{max}\sim\mathcal{O}(10^{5})$.
+For GPT and GPT2, we treat all layers as-is (although one may to normalize the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier).
+Here, we do not include them in our computed average metrics in Table~\ref{table:nlp}, and we do not include them in the histogram plot in Figure~\ref{fig:GPT-snorm-hist}.
 
-also differs significantly between GPT and GPT2.
-Figure \ref{fig:gpt-alpha-layer} plots $\alpha$ vs the depth (i.e. a layer id) for each model.
-The deficient GPT model displays two trends in $\alpha$ , one stable with $\alpha\sim 4$,
-and one increasing with layer id, with $\alpha$ reaching as high as $12$.
-In contrast, the well-trained GPT2 model shows consistant and stable patterns, again
-with one stable $\alpha\sim 3.5$ (and below the GPT trend), and the other only
-slightly trending up, with $alpha\le 6$. 
-The scale-invariant $\alpha$ metric lets us identify potentially
-poorly-trained models.
+\paragraph{Layer Analysis: Correlation Flow and Scale Collapse in GPT and GPT2.}
 
-Notice also that the Spectral Norms
-(Figure \ref{fig:gpt-snorm-layer})
-and the $\alpha-$Norms
-(Figure \ref{fig:gpt-pnorm-layer})
-display an unusually some mall values for GPT,
-and some increasing trending values for GPT2.
-Generally speaking, we do not think scale-dependennt 
-norm metrics can be directly applied to distinguish ``good'' vs ``bad'' models 
-because of the anamalous scale collapse that is frequently observed in bad models.
+We also examine in Figure~\ref{fig:gpt-alpha-layers} the PL exponent $\alpha$, log Spectral Norm, and log Alpha Norm versus layer id, for GPT and GPT2-small.
+Let's start with Figure~\ref{fig:gpt-alpha-layer}, which plots $\alpha$ versus the depth (i.e., layer id) for each model.
+The deficient GPT model displays two trends in $\alpha$, one stable with $\alpha\sim 4$, and one increasing with layer id, with $\alpha$ reaching as high as $12$.
+In contrast, the well-trained GPT2-small model shows consistant and stable patterns, again with one stable $\alpha\sim 3.5$ (and below the GPT trend), and the other only slightly trending up, with $\alpha\le 6$.
+These results suggest that the correlation flow differs significantly between GPT and GPT2-small (with the better GPT2-small looking more like the better ResNet-1K from Figure~\ref{fig:resnet-alpha-layer}) and that the scale-invariant $\alpha$ metric lets us identify potentially poorly-trained models.
 
-\begin{figure}[t]
+These results should be contrasted with results for Spectral Norms (Figure~\ref{fig:gpt-snorm-layer}) and the $\alpha$-Norms (Figure~\ref{fig:gpt-pnorm-layer}).
+Attention models have two types of layers, one small and large; and the Spectral Norm, in particular, displays unusually small values for some of these layers for GPT.
+This scale collapse for the poorly-trained GPT is similar to what we observed in Figure\ref{fig:resnet204Dalpha} for the distilled ResNet20 model.
+Because of the anomalous scale collapse that is frequently observed in poorly-trained models, these results suggest that scale-dependennt norm metrics should not be directly applied to distinguish good-versus-bad models.
+
+%MM%    \begin{figure}[t]
+%MM%        \centering
+%MM%        \subfigure[PL exponent ($\alpha$)]{
+%MM%            %\includegraphics[width=4.5cm]{img/GPT-alpha-depth.png}
+%MM%            %\includegraphics[width=4.0cm]{img/GPT-alpha-depth.png}
+%MM%            \includegraphics[width=2.6cm]{img/GPT-alpha-depth.png}
+%MM%            \label{fig:gpt-alpha-layer}
+%MM%        }
+%MM%        %\qquad
+%MM%        \subfigure[log Spectral Norm]{
+%MM%            %\includegraphics[width=4.5cm]{img/GPT-snorm-depth.png}
+%MM%            %\includegraphics[width=4.0cm]{img/GPT-snorm-depth.png}
+%MM%            \includegraphics[width=2.6cm]{img/GPT-snorm-depth.png}
+%MM%            \label{fig:gpt-snorm-layer}
+%MM%        }
+%MM%        %\qquad
+%MM%        \subfigure[log Alpha Norm]{
+%MM%            %\includegraphics[width=4.5cm]{img/GPT-pnorm-depth.png}
+%MM%            %\includegraphics[width=4.0cm]{img/GPT-pnorm-depth.png}
+%MM%            \includegraphics[width=2.6cm]{img/GPT-pnorm-depth.png}
+%MM%            \label{fig:gpt-pnorm-layer}
+%MM%        }
+%MM%        \caption{
+%MM%                 PL exponent $\alpha$, log Spectral Norm, and log Alpha Norm versus layer id, for OpenAI GPT and GPT2-small models.
+%MM%                 (Y axes on each plot are different.)  
+%MM%                }
+%MM%        \label{fig:gpt-alpha-layers}
+%MM%    \end{figure}
+%MM%    
+%MM%    
+%MM%    \begin{figure}[t]
+%MM%        \centering
+%MM%        \subfigure[PL exponent ($\alpha$)]{
+%MM%            %\includegraphics[width=4.5cm]{img/GPT2_all_alpha_hist.png}
+%MM%            %\includegraphics[width=4.0cm]{img/GPT2_all_alpha_hist.png}
+%MM%            \includegraphics[width=2.6cm]{img/GPT2_all_alpha_hist.png}
+%MM%            \label{fig:gpt2-alpha-hist}
+%MM%        }
+%MM%        %\qquad
+%MM%        \subfigure[log Spectral Norm]{
+%MM%            %\includegraphics[width=4.5cm]{img/GPT2*_all_spectralnorm_hist.png}
+%MM%            %\includegraphics[width=4.0cm]{img/GPT2*_all_spectralnorm_hist.png}
+%MM%            \includegraphics[width=2.6cm]{img/GPT2*_all_spectralnorm_hist.png}
+%MM%            \label{fig:gpt2-snorm-hist}
+%MM%        }
+%MM%        %\qquad
+%MM%        \subfigure[log Alpha Norm]{
+%MM%            %\includegraphics[width=4.5cm]{img/GPT2_all_logpnorm_hist.png}
+%MM%            %\includegraphics[width=4.0cm]{img/GPT2_all_logpnorm_hist.png}
+%MM%            \includegraphics[width=2.6cm]{img/GPT2_all_logpnorm_hist.png}
+%MM%            \label{fig:gpt2-pnorm-hist}
+%MM%        }
+%MM%        \caption{PL exponents $\alpha$, log Spectral norm (omitting first two embedding layers), and log Alpha norm for different size models in GPT2 architecture series.  
+%MM%                 (The log Spectral norm histogram omits the first 2 layers, corresponding to the embedding layer, as these are normalized differently and have anamolously large values.)
+%MM%                }
+%MM%        \label{fig:gpt2-histograms}
+%MM%    \end{figure}
+
+\begin{figure}
 \centering
-
+\begin{minipage}[b]{.22\textwidth}
+%<Code for the first figure>
+%MM%   \begin{figure}[t]
+%MM%       \centering
        \subfigure[PL exponent ($\alpha$)]{
@@
        }
-    \caption{Comparison of Correlation Flow and Spectral Norm for OpenAI GPT and GPT2   }
+      \caption{
+               PL exponent $\alpha$, log Spectral Norm, and log Alpha Norm versus layer id, for OpenAI GPT and GPT2-small models.
+               (Y axes on each plot are different.)  
+%%               \michael{In the text, we interpret the behavior of $\alpha$ as quantifying a \emph{correlation flow} and the behavior of log Spectral Norm in terms of a \emph{scale collapse}.}
+              }
       \label{fig:gpt-alpha-layers}
-\end{figure}
-
-\paragraph{GPT2: small, medium, large, xl}
-
-We now look across the series of increasingly improving GPT2 models, (i.e. ``good, better, best''), 
-by examining both the PL exponent $\alpha$, as well as the various log Norm metrics.  
-In general, as we move from GPT2-small to GPT2-xl, 
-the histograms for both $\alpha$ exponents and the log Norm metrics downshift from larger to smaller values. 
-See Figure \ref{fig:gpt2-histograms}, which shows the histograms over the layer weight matrics
-for fitted $\alpha$, and the log Spectral Norm
- $(\log\Vert\mathbf{W}\Vert_{\infty})$  
-and log Alpha Norm
- $(\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha})$ 
-metrics.
-
-We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticible 
-between the differing ``good, better, best'' GTP2 models than between the ``good vs bad'' GPT2 and GPT models.
-Unlike GPT, however, the layer (log) Spectral Norms $(\log\vert\mathbf{W}\Vert_{\infty})$ 
-and (log) Alpha Norms $(\log\vert\mathbf{W}\Vert_{\alpha}^{\alpha})$
-behave more as expected for GPT2 layers, with the larger models, consistently having smaller norms. 
-\michael{Clarify.}\charles{???}
-Likewise, Figure \ref{fig:gpt2-snorm-hist} shows decreasing average log Spectral Norms with the larger models.  
-As expected, the norm metrics can indeed distinguish among ``good, better, best'' models
-among a series well trained models.
-
-We do notice, however, that while the peaks of the $\alpha$ is getting smaller, 
-towards $2.0$,
-the tails of the distribution shifts right, with larger GPT2 models having more
-usually large $\alpha$.  
-We suspect this indicates that these larger GPT2 models are over-parameterized/under-optimized and 
-that they have capacity to support datasets even larger than the recent XL $1.5B$ release~\cite{gpt2-xl}.
-
-\begin{figure}[t]
-    \centering
-
+%MM%   \end{figure}
+%MM% \caption{Caption}\label{label-a}
+\end{minipage}\qquad
+\begin{minipage}[b]{.22\textwidth}
+%<Code for the second figure>
+%MM%    \begin{figure}[t]
+%MM%        \centering
        \subfigure[PL exponent ($\alpha$)]{
@@
        }
-    \caption{PL exponents ($\alpha$), log Spectral norm, and log Alpha norm for different size models in the GPT2 architecture series.
-(The log Spectral norm histogram omits the first 2 layers, corresponding to the embedding layer, as these are normalized differently and have anamolously large values.)}
+       \caption{PL exponents $\alpha$, log Spectral norm (omitting first two embedding layers), and log Alpha norm for different size models in GPT2 architecture series.  
+%%                \michael{(The log Spectral norm histogram omits the first 2 layers, corresponding to the embedding layer, as these are normalized differently and have anamolously large values.)}
+               }
        \label{fig:gpt2-histograms}
+%MM%    \end{figure}
+%MM% \caption{Caption}\label{label-b}
+\end{minipage}
 \end{figure}
+
+\paragraph{GPT2: medium, large, xl}
+
+We now look across the series of increasingly improving GPT2 models, i.e., good-better-best, by examining both the PL exponent $\alpha$ as well as the various log Norm metrics.  
+In general, as we move from GPT2-medium to GPT2-xl, the histograms for both $\alpha$ exponents and the log Norm metrics downshift from larger to smaller values. 
+See Figure \ref{fig:gpt2-histograms}, which shows the histograms over the layer weight matrics
+for fitted $\alpha$, and the log Spectral Norm
+ $(\log\Vert\mathbf{W}\Vert_{\infty})$  
+and log Alpha Norm
+ $(\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha})$ 
+metrics.
+
+We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticible between the differing good-better-best GTP2 models than between the good-versus-bad GPT and GPT-small models.
+Unlike GPT, however, the layer (log) Spectral Norms $(\log\vert\mathbf{W}\Vert_{\infty})$ and (log) Alpha Norms $(\log\vert\mathbf{W}\Vert_{\alpha}^{\alpha})$ behave more as expected for GPT2 layers, with the larger models, consistently having smaller norms.
+Likewise, Figure \ref{fig:gpt2-snorm-hist} shows decreasing average log Spectral Norms with the larger models.  
+As expected, the norm metrics can indeed distinguish among good-better-best models, i.e., when the comparison is among a series well trained models.
+
+We do notice, however, that while the peaks of the $\alpha$ plots are getting smaller, towards $\alpha\approx 2.0$, the tails of the distribution shifts right, with larger GPT2 models having more usually large $\alpha$.
+We suspect this indicates that these larger GPT2 models are still under-optimized/over-parameterized and that they have capacity to support datasets even larger than the recent XL $1.5B$ release~\cite{gpt2-xl}.

### 2020-02-08 all_models.tex
updating tex

@@
 \caption{Comparison of linear regression fits for different average log norm metrics across 5 computer vision datasets, 17 Architectures, covering 168 (out of 309) different pretrained DNNs.
-We only conclude regressions for architecturs with 4 or more data points \charles{and which are postively correlated with the test error?}.
-These results can be readily reproduced using the Google Colab notebooks accompanying thie paper~\cite{notebooks}}
+We only conclude regressions for architecturs with 4 or more data points, and which are postively correlated with the test error.
+These results can be readily reproduced using the Google Colab notebooks (see Appendix)}
 \label{table:results}

### 2020-02-08 cv_models.tex
updating tex

--- 2020-02-02 cv_models.tex
@@
     }
-    \caption{Comparison of quality metrics versus reported test accuracy for pretrained VGG models, trained on ImageNet, available in pyTorch.  
-             XXX Plots will be updated and replaced. 
-             \michael{Figures are unreadable, can we make fonts bigger, etc.}
+    \caption{Comparison of average log Norm empirical quality metrics versus reported test accuracy for pretrained VGG models (with and without BatchNormalization), trained on ImageNet, available in pyTorch (v1.x).
+Metrics fit by linear regression, RMSE reported.
 }
@@
     %\qquad
-    \subfigure[ DenseNet, $\alpha$-Norm ]{
-        %\includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
-        \includegraphics[width=4.0cm]{img/DenseNet_logpnorm_accs.png}
-        \label{fig:densenet-accuracy}
-    }
-    \caption{$\alpha$-Norm versus reported Top1 Error for  ResNet, ResNet-1K, and DenseNet models. The corresponding results for VGG are shown in Figure~\ref{fig:vgg-pnorm}.  }
+    %\subfigure[ DenseNet, $\alpha$-Norm ]{
+    %    %\includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
+    %    \includegraphics[width=4.0cm]{img/DenseNet_logpnorm_accs.png}
+    %    \label{fig:densenet-accuracy}
+    %}
+    \caption{Comparison of the avergage $\alpha$-Norm empirical quality metric $\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle$ 
+versus reported Top 1 reported Test Accuracy for the ResNet and ResNet-1K pretrained (pyTorch) models. }
+%, and DenseNet models. }The corresponding results for VGG are shown in Figure~\ref{fig:vgg-pnorm}.  }
     \label{fig:cv2-accuracy}
@@
 \hline
-%   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
-% Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
-        &        & Log                   & Log                        & Weighted       & Log                                 \\
-        & Num.   & $\Vert\cdot\Vert_{F}$ & $\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
- Series & Models & Metric                & Metric                     & Metric         & Metric                              \\
+ Series    &\#   & $\log\langle\Vert\mathbf{W}\Vert_{F}\rangle$ & $\log\langle\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\hat{\alpha}$ & $\log\langle\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
 \hline
@@
 
-Figure~\ref{fig:cv2-accuracy} plots the 
-%overall best performing (i.e., 
+See Table~\ref{table:cv-models} for a summary of results for Top1 Accuracies for all four metrics on for the VGG, ResNet, and DenseNet series.
+Similar results (not shown) are obtained for the Top5 Accuracies.
+Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and th\
+e Weighted Alpha metric performs best.
+These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among ``good-better-bes\
+t'' models, with PL-based metrics performing somewhat better.
+
+Notice that the DenseNet series,  only has 4 data points.  In our larger analysis, in Section~\ref{sxn:all} we only include series with 5 or more models.
+For completemnss and reproducibility here, the DenseNet Log $\alpha$-Norm plot is given in the Appendix.
+
+\paragraph{Variations in Data Set Size}
+We want to point out, in particular, how the empirical log Norm metrics vary with data set size.
+Figure~\ref{fig:cv2-accuracy} plots and compares the 
 Log $\alpha$-Norm
-%) 
-metric for the full ResNet, ResNet-1K, and DenseNet series.
-(A much more detailed set of plots, including the behavior of all four methods on each of the series, is available at \cite{XXX-WEB-LINK}.)
+metric--the best performing metric--for the full ResNet, trained on the full ImageNet dataset, against the ResNet-1K, which has been trained on a much smaller ImageNet-1K data set.
+%(A much more detailed set of plots, including the behavior of all four methods on each of the series, is available at \cite{XXX-WEB-LINK}.)
 Here, the Log $\alpha$-Norm is much better that the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is actually slightly worse than the Weighted Alpha metric).
-Notice also that ResNet series, which has been trained on the full ImageNet dataset, has an RMSE of $0.66$, whereas the ResNet-1K series, which has been trained on the much smaller ImageNet-1K dataset, also correlated wells, but has a much larger RMSE of $1.9$.
-
-See Table~\ref{table:cv-models} for a summary of results for Top1 Accuracies for all four metrics on all four model series.
-Similar results (not shown) are obtained for the Top5 Accuracies.
-Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best. 
-These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among ``good-better-best'' models, with PL-based metrics performing somewhat better.
+The ResNet series, has an RMSE of $0.66$, whereas the ResNet-1K series also shows good correlation, but has a much larger RMSE of $1.9$.
+As expected, the higher quality data set shows a better fit, even with less data points.
 
 \paragraph{Layer Analysis: Correlation Flow in CV Models.}
-
 We can learn much more about a pretrained model by going beyond average values to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).
@@
     %\qquad
-    \subfigure[ ResNet-1K ]{
+    \subfigure[ ResNet ]{
         %\includegraphics[width=4.5cm]{img/ResNet_fnl_alpha_depth.png} 
@@
     }
-    \caption{Layer Analysis. 
-             PL exponent $\alpha$ versus layer id, for VGG, ResNet-1K, and DenseNet models.  
-             (Y axes on each plot are different.)  
-             ResNet-1K exhibits different and much more stable behavior across layers.
-             This contrasts with how VGG gets gradually worse in deeper layers and how DenseNet is worse and much more erratic.  
-             We interpret this in the text as quantifying a \emph{correlation flow} across the layers.  
+    \caption{Correlation Flow: Power Law exponent $\alpha$ versus layer id, for the least and the most accurate
+      models in VGG (a), ResNet (b), and DenseNet (c) series (VGG without BatchNorm, and the Y axes on each plot are different.)  
+             Figure (d) displays the ResNet models (b), zoomed in to $alpha\in[1,5]$, and with the layer ids overlaid on the X-axis, to
+             allow a more detailed analysis for the most strongly correlated layers.
+             Notuece that ResNet152 exhibits different and much more stable behavior across layers
+             This contrasts with how both VGG models gradually worsen in deeper layers and how the DenseNet models are much more erratic.  
             }

### 2020-02-08 nlp_models.tex
updating tex

--- 2020-02-03 nlp_models.tex
@@
 Based on HT-SR Theory, 
-%(in particular, the interpretation of values of $\alpha \sim 2$ as modeling systems with good correlations over many size scales~\cite{BouchaudPotters03, SornetteBook}), 
+%(in particular, the interpretation of values of $\alpha \sim 2$ as modeling systems with good correations over many size scales~\cite{BouchaudPotters03, SornetteBook}), 
 this suggests that these models fail to successfully capture many of the correlations in the data (relative to their size) and thus are substantially \emph{under}-trained.
@@
     \centering
-    \subfigure[PL exponent ($\alpha$)]{
-        %\includegraphics[width=5cm]{img/GPT-alpha-hist.png}
+    \subfigure[Power law exponent ($\alpha$)]{
         \includegraphics[width=4.0cm]{img/GPT-alpha-hist.png}
@@
     %\qquad
-    \subfigure[log Spectral Norm]{
-        %\includegraphics[width=5cm]{img/GPT-snorm-hist.png}
+    \subfigure[Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
         \includegraphics[width=4.0cm]{img/GPT-snorm-hist.png}
@@
 
-\begin{figure}[t]
+\begin{figure}[H]
     \centering
-
-    \subfigure[PL exponent ($\alpha$)]{
-        %\includegraphics[width=4.5cm]{img/GPT-alpha-depth.png}
+    \subfigure[Power Law exponent $(\alpha)$]{
         \includegraphics[width=4.0cm]{img/GPT-alpha-depth.png}
@@
     }
-    %\qquad
-    \subfigure[log Spectral Norm]{
-        %\includegraphics[width=4.5cm]{img/GPT-snorm-depth.png}
+    \subfigure[Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
         \includegraphics[width=4.0cm]{img/GPT-snorm-depth.png}
         \label{fig:gpt-snorm-layer}
-    }
-    %\qquad
-    \subfigure[log Alpha Norm]{
-        %\includegraphics[width=4.5cm]{img/GPT-pnorm-depth.png}
-        \includegraphics[width=4.0cm]{img/GPT-pnorm-depth.png}
-        \label{fig:gpt-pnorm-layer}
     }
@@
 
-\paragraph{GPT2: small, medium, large, xl}
-
+\paragraph{GPT2: medium, large, xl} 
 We now look across the series of increasingly improving GPT2 models, (i.e. ``good, better, best''), 
@@
 behave more as expected for GPT2 layers, with the larger models, consistently having smaller norms. 
-\michael{Clarify.}\charles{???}
 Likewise, Figure \ref{fig:gpt2-snorm-hist} shows decreasing average log Spectral Norms with the larger models.  
@@
     \subfigure[PL exponent ($\alpha$)]{
-        %\includegraphics[width=4.5cm]{img/GPT2_all_alpha_hist.png}
-        \includegraphics[width=4.0cm]{img/GPT2_all_alpha_hist.png}
+        \includegraphics[width=4.0cm]{img/GPT2*_fnl_alpha_hist.png}
         \label{fig:gpt2-alpha-hist}
@@
     %\qquad
-    \subfigure[log Spectral Norm]{
-        %\includegraphics[width=4.5cm]{img/GPT2*_all_spectralnorm_hist.png}
-        \includegraphics[width=4.0cm]{img/GPT2*_all_spectralnorm_hist.png}
-        \label{fig:gpt2-snorm-hist}
-    }
+    %%\subfigure[log Spectral Norm]{
+    %    \includegraphics[width=4.0cm]{img/GPT2*_fnl_spectralnorm_hist.png}
+    %    \label{fig:gpt2-snorm-hist}
+   % }
     %\qquad
     \subfigure[log Alpha Norm]{
-        %\includegraphics[width=4.5cm]{img/GPT2_all_logpnorm_hist.png}
-        \includegraphics[width=4.0cm]{img/GPT2_all_logpnorm_hist.png}
+        \includegraphics[width=4.0cm]{img/GPT2*_fnl_logpnorm_hist.png}
         \label{fig:gpt2-pnorm-hist}
     }
-    \caption{PL exponents ($\alpha$), log Spectral norm, and log Alpha norm for different size models in the GPT2 architecture series.
-(The log Spectral norm histogram omits the first 2 layers, corresponding to the embedding layer, as these are normalized differently and have anamolously large values.)}
+    \caption{PL exponents ($\alpha$), log Spectral norm $(\log\Vert\mathbf{W}\Vert_{\infty})$, and log Alpha norm $(\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha})$ for different size models in the GPT2 architecture series.
+(Plots omit the first 2 (embedding) layers, because they are normalized differently giving anamolously large values.)}
     \label{fig:gpt2-histograms}
 \end{figure}
+
+?

### 2020-02-08 cv_models.tex
updated densenet images

@@
 
-In this section, we examine quality metrics described in Section~\ref{sxn:methods} for several CV model architecture series.
+In this section, we examine empirical quality metrics described in Section~\ref{sxn:methods} for several CV model architecture series.
 This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current opensource pyTorch framework (version 1.4)~\cite{pyTorch}.
@@
         \label{fig:densenet-alpha-layer}
+
+}    \subfigure[ DenseNet ]{
+        \includegraphics[width=4.0cm]{img/resnet_alpha_overlaid_depth.png} 
+        \label{fig:resnet_alpha_overlaid_depth}
     }

### 2020-02-08 cv_models.tex
updated cv section

@@
 
-\paragraph{Layer Analysis: Correlation Flow in CV Models.}
+\paragraph{Correlation Flow}
 We can learn much more about a pretrained model by going beyond average values to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).
@@
 
-}    \subfigure[ DenseNet ]{
+}    \subfigure[ ResNet (overlaid) ]{
         \includegraphics[width=4.0cm]{img/resnet_alpha_overlaid_depth.png} 
@@
              allow a more detailed analysis for the most strongly correlated layers.
-             Notuece that ResNet152 exhibits different and much more stable behavior across layers
+             Notice that ResNet152 exhibits different and much more stable behavior across layers
              This contrasts with how both VGG models gradually worsen in deeper layers and how the DenseNet models are much more erratic.  
@@
 
-Recall the differences---in particular, the number of residual connections---between the VGG, ResNet, and DenseNet architectures.
-VGG, while a fine model in many ways, was limited by needing massive FC layers at the end of the model, requiring significant memory and computational resources, and it's poor conditioning led to problems with vanishing gradients.
-The empirical manifestations of this (on weight matrices) are that fitted $\alpha$ values get much worse for deeper layers.
-ResNet greatly improved on VGG by introducing residual connections, allowing for greater accurary with far fewer parameters.
-(Indeed, ResNet models of up to 1000 layers have been trained.) 
+%Recall the differences---in particular, the number of residual connections---between the VGG, ResNet, and DenseNet architectures.
+%VGG, while a fine model in many ways, was limited by needing massive FC layers at the end of the model, requiring significant memory and computational resources, and it's poor conditioning led to problems with vanishing gradients.
+%The empirical manifestations of this (on weight matrices) are that fitted $\alpha$ values get much worse for deeper layers.
+\ncyan{We can interpret these observations by recalling the architectural differences between the VGG, ResNet, and DenseNet architectures,
+and, in particular, the number of of residual connections. VGG resembles the  traditional convolutional architectures,
+such as LeNet5, and consists of several [Conv2D-Maxpool-ReLu] blocks, followed by 3 large Fully Connected (FC) layers.
+ResNet greatly improved on VGG by replacing the large FC layers, shrinking the Conv2D blocks, and introducing \emph{residual connections}.
+This optimized approach allows for greater accurary with far fewer parameters (and GPU memory requirements), and
+ResNet models of up to 1000 layers have been trained.\cite{resnet1000}.}
 We conjuecture that the effeciency and effectivness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and strongly optimized.
-This should also be contrasted with the DenseNet model, which contains many connections between every layer.
+Contrast this with the DenseNet models, which contains many connections between every layer.
 Our results (large $\alpha$, meaning they even a HT model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly optimized.
@@
 %\paragraph{Layer Analysis: Behavior on Less Thoroughy Studied CV Models.}
-\paragraph{Layer Analysis: Differences Between PL Metrics and Norm Metrics.}
+\paragraph{Corraltion Flow: PL vs. Norm Metrics.}
 
@@
    }
-   \caption{Layer Analysis.
+   \caption{Correlation Flow.
             ResNet20, distilled with Group Regularization, as implemented in the \texttt{distiller} (4D\_regularized\_5Lremoved) pre-trained models.  
-            Spectral Norm, $\lambda_{max}$, and, PL exponent, $\alpha$, for individual layer, versus layer id, for both between baseline (before distillation, shown in green) and fine-tuned (after distillation, shown in red) pre-trained models.
+            Spectral Norm, $\log\lambda_{max}$, and, PL exponent, $\alpha$, for individual layer, versus layer id, for both baseline (before distillation, green) and finetuned (after distillation, red) pre-trained models.
            }

### 2020-02-08 cv_models.tex
updated main sections

@@
-\section{Comparison across series of CV models}
+\section{Comparison of CV models}
 \label{sxn:cv}
@@
 
-\paragraph{Coarse Analysis: Quality Metrics versus Reported Test Accuracies.}
-
+\paragraph{Average Quality Metrics versus Reported Test Accuracies.}
 We have examined the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
@@
 VGG, ResNet, and DenseNet were pretrained on ImageNet, and ResNet-1K was pretrained on ImageNet-1K.
-\michael{Are the numbers in this table stale; if they get updated, then we need to modify the test at at least two places, on actual numbers and on what is best.}
+\michael{double check numbers}
 }
@@
 
-More generally, we can understand Figure~\ref{fig:vgg-alpha-layers} in terms of the \emph{flow of correlation}, as follows.
-Recall that the Log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on ideas from the statisical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
-There, one expects well-trained systems which exhibit correlations over many size scales to have ESDs that can be well-fit by PLs, with PL exponents $\alpha\approx 2$.
-Much larger values of $\alpha$ indicate very poor fits, but the presence of many adjacent layers with relatively good PL fits \emph{suggests} that those adjacent layers ``impedance match'' well and that the correlations that are captured by one layer propagate well to subsequent layers.
+More generally, we can understand Figure~\ref{fig:vgg-alpha-layers} in terms of the \emph{Correlation Flow}. 
+Recall that the log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on ideas from the statisical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
+There, one expects the weight matrices of well-trained DNNs will 
+exhibit correlations over many size scales. Their ESDs can be well-fit by a (truncated) Power law (PL), with exponents $\alpha\in[2,4]$.
+Much larger values $(\alpha\gg 5\;or\;6)$ may relfect poorer PL fits, whereas smaller values $(\alpha\rightarrow 2)$, are associated with models
+that generalize better.
+%, but the presence of many adjacent layers with relatively good PL fits \emph{suggests} that those adjacent layers ``impedance match'' well and that the correlations that are captured by one layer propagate well to subsequent layers.
 Informally, one would expect a DNN model to perform well when it fascilitates the propagation of information/features across layers.
+Previous studies argue this by computing the gradients over the input data.
 In the absence of training/test data, we can to try to quantify this by measuring the PL properties of empirical weight matrices.
@@
 %\paragraph{Layer Analysis: Behavior on Less Thoroughy Studied CV Models.}
-\paragraph{Corraltion Flow: PL vs. Norm Metrics.}
+\paragraph{Correlation Flow: PL vs. Norm Metrics.}
 
@@
       %\includegraphics[width=4.5cm]{img/resnet4d_maxev.png}
-      \label{fig:resnet204Dalpha}
+      \label{fig:resnet204Dmaxev}
    }
@@
       %\includegraphics[width=4.5cm]{img/resnet4d_alphas.png}
-      \label{fig:resnet204Dmaxev}
+      \label{fig:resnet204Dalpha}
    }
@@
 
-We consider ResNet20, trained on CIFAR10, before and after applying the Group Regularization technique, as implemented in the \texttt{distiller} package.%
+\paragraph{Scale Collapse; or, Distillation may break models}
+
+In examining hundreds of pretrained models, we have found several anamolies that demonsrate the power of our approach.
+Here, we demonstrate that some distillations methods may actually \emph{break} models unexpectedly by introducing what we call \emph{Scale Collapse},
+where several distilled layers have unexpectedly small Spectral Norms.
+We consider ResNet20, trained on CIFAR10, before and after applying the Group Regularization distillation technique, as implemented in the \texttt{distiller} package.%
 \footnote{For details, see \url{https://nervanasystems.github.io/distiller/\#distiller-documentation} and also \url{https://github.com/NervanaSystems/distiller}.}
-We analyze the available pre-trained 4D\_regularized\_5Lremoved baseline and fine-tuned models.  %\cite{distiller_repo}.
-The reported baseline test accuracies ($Top1=91.45$ and $Top5=99.75$) are better than the reported fine-tuned test accuracies ($Top1=91.02$ and $Top5=99.67$)~\cite{XXX-XXX}.
-The previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy} might suggest that the baseline Spectral Norm should be \emph{smaller} than those of the layers in the fine-tuned model.
-\michael{Do we have something backwards here: where we observed that the average norm to increase with decreasing test error (as it ``should'' not), whereas the average PL exponent $\alpha$ decreases (as ``expected'' from HT-SR Theory).
+We analyze the pre-trained 4D\_regularized\_5Lremoved baseline and finetuned models.  %\cite{distiller_repo}.
+The reported baseline test accuracies ($Top1=91.45$ and $Top5=99.75$) are better than the reported finetuned test accuracies ($Top1=91.02$ and $Top5=99.67$)~\cite{XXX-XXX}.
+Because the baseline accuracy is greater,  the previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy}) suggests that the baseline Spectral Norms should be \emph{smaller} on average than the finetuned ones should be smaller.
+\emph{The oppopsite is observed.}
+%\michael{Do we have something backwards here: where we observed that the average norm to increase with decreasing test error (as it ``should'' not), whereas the average PL exponent $\alpha$ decreases (as ``expected'' from HT-SR Theory).
 }
-In both cases (Frobenius norm results not shown), we observe the opposite.
-Figure~\ref{fig:resnet204D5L} presents the Spectral Norm ($\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}_{l}$.%
+%In both cases (Frobenius norm results not shown), we observe the opposite.
+Figure~\ref{fig:resnet204D5L} presents the Spectral Norm (here denoteds $\log\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}$.%
 \footnote{Here, we only include layer matrices or feature maps with $M\ge50$.}
-On the other hand, the $\alpha$ values do not differ systematically between the baseline and fine-tuned models.
-\michael{Some comment about big jump in norm; but fix previous backwards question first.}
-Also (not shown), the average (unweighted) baseline $\alpha$ is \emph{smaller} than the fine-tuned average (as would be predicted by HT-SR Theory, on which $\hat{\alpha}$ is based).
+On the other hand, the $\alpha$ values do not differ systematically between the baseline and finetuned models.
+%\michael{Some comment about big jump in norm; but fix previous backwards question first.}--SEE BELOW
+Also (not shown), the average (unweighted) baseline $\alpha$ is \emph{smaller} than the finetuned average (as predicted by HT-SR Theory, the basis of $\hat{\alpha}$).
+
+Note that Figure ~\ref{fig:resnet204Dalph} depcits 2 very large $\alpha\gg 6$ for the baseline model, but not for the finetuned model.
+This suggests that the baseline model has at least 2 over-parameterized layers, and the distillation method does, in fact, 
+improve the finetuned model by compressing these layers.s
 
@@
 Our interpretation of this is the following.
-The pre-trained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the VGG/ResNet/DenseNet models we condisered previously.
+The pre-trained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the 
+VGG, ResNet, or DenseNet series we cosidered above.
 \charles{MOvE THIS SOMEWHERE ELSE AND CLEAN UP THIS SECTION Notice while norms make good regularizers for a single model, there is no reason \emph{a priori} 

### 2020-02-08 methods.tex
updated main sections

@@
 $ %$$
-\michael{Be careful to get right.}
-The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where (on a much more limited dataset than we consider here) it was shown to correlate well with trends in reported test accuracies of pretrained DNNs.
+%\michael{Be careful to get right.}
+The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a limited set of models.
 Based on this, in this paper, we introduce and evaluate the $\alpha$-Norm metric.
@@
 We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated PL, parameterized by 2 values: the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.  
-(Technically, we also need the minimum eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.)
+%(Technically, we also need the minimum eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.)
 The PL exponent $\alpha$ measures of the amount of correlation in a DNN layer weight matrix $\mathbf{W}$. 
 It is valid for $\lambda<\lambda_{max}$, and it is scale-invariant, i.e., it does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.
-The $\lambda_{max}$ is a measure of the size of $\mathbf{W}$.
+The $\lambda_{max}$ is a measure of the size, or scale, of $\mathbf{W}$.
 %

### 2020-02-08 nlp_models.tex
updated main sections

@@
-\section{Comparison of NLP Transformer Models}
+\section{Comparison of NLP Models}
 \label{sxn:nlp}
@@
 having include $12, 24, 36, \text{and }48$ layers, respectively, with increasingly larger weight matrices.
-The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
-
-%        &  Num.  & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
-% Series & Models & Metric                     & Metric                     & Metric         & Metric                                   \\
+%The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
 
@@
 \hline
-%   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha Norm \\
-% Series & \#Layers   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
-        &        & log                   & log                        & Weighted       & log                                 \\
-        & Num.   & $\Vert\cdot\Vert_{F}$ & $\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
- Series & Layers & Metric                & Metric                     & Metric         & Metric                              \\
+ Series  & \#   & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\hat{\alpha}$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
 \hline
@@
 \end{center}
-\caption{Average value for the Average log Norm metrics and the Weighted Alpha metric for pretrainnd OpenAI GPT and GPT2 models.  }
+\caption{Average value for the Average log Norm metrics and the Weighted Alpha metric for pretrainnd OpenAI GPT and GPT2 models.
+Column \# refers to number of layers treated.
+Note averages do not include the first embedding layer(s) because they are not (implicitly) normalized.
+}
 \label{table:nlp}

### 2020-02-08 nlp_models.tex
updated main sections

@@
 it scale invariant (does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.)
-The Weighted-Alpha metric, $\hat{\alpha}=\alpha\log\lambda_{max}$, approximates the log $\alpha$ (Shatten) Norm 
+The Weighted-Alpha metric, $\hat{\alpha}=\alpha\log\lambda_{max}$, approximates the log $\alpha$ Norm.
 $(\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha})$ (derived using statistical mechanics and RMT \nred{(in progress)}, 

### 2020-02-08 nlp_models.tex
updated main sections

@@
 
-\michael{WHERE: Say large $\alpha$ values don't mean much.}
+%\michael{WHERE: Say large $\alpha$ values don't mean much.}
+\paragraph{What does large $\alpha$ mean?}
 We do note, however, that many NLP models, such as the GPT and BERT, may have weight matrices with unusually large PL exponents
-($\alpha>6$), which indicates these matrices may be \emph{under}-correlated (i.e over-paramaterized), although the truncated PL fit
+($\alpha\gg 6$), which indicates these matrices may be \emph{under}-correlated (i.e over-parameterized), although the truncated PL fit
 itself may not be very reliable because the MLE estimator is unreliable in this range.  Phenomenologically, if we examine
@@
 
-\michael{WHERE: Highlight difference between $\alpha$ and $\hat{\alpha}$.}
+%\michael{WHERE: Highlight difference between $\alpha$ and $\hat{\alpha}$.}
+\paragraph{The differences between $\alpha$ and $\hat{\alpha}$.}
 To avoid confusion, let us clarify the differences between $\alpha$ and $\hat{\alpha}$.  

### 2020-02-08 cv_models.tex
updated main sections

@@
 We have examined the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
-To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BatchNormalization), and it plots the four quality metrics versus the reported Test accuracies~\cite{pyTorchVgg}, as well as a basic linear regression line.
+To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without Batch Normalization), and it plots the four quality metrics versus the reported Test accuracies~\cite{pyTorchVgg}, as well as a basic linear regression line.
 %with the Root Mean Squared Error (RMSE) shown, \michael{XXX.  WHERE IS RMSE.  IN THAT TABLE?}
@@
     }
-    \caption{Comparison of average log Norm empirical quality metrics versus reported test accuracy for pretrained VGG models (with and without BatchNormalization), trained on ImageNet, available in pyTorch (v1.x).
+    \caption{Comparison of average log Norm empirical quality metrics versus reported test accuracy for pretrained VGG models (with and without Batch Normalization (BN)), trained on ImageNet, available in pyTorch (v1.x).
 Metrics fit by linear regression, RMSE reported.
@@
 VGG, ResNet, and DenseNet were pretrained on ImageNet, and ResNet-1K was pretrained on ImageNet-1K.
-\michael{double check numbers}
 }
@@
 %
-See Figure~\ref{fig:3models-alpha-layers}, which plots $\alpha$ for each layer (the first layer corresponds to data, the last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (without BatchNormalization), ResNet-1K, and DenseNet series.
+See Figure~\ref{fig:3models-alpha-layers}, which plots $\alpha$ for each layer (the first layer corresponds to data, the last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (no BN), ResNet, and DenseNet series.
 (Again, much more detailed set of plots is available at \cite{XXX-WEB-LINK}; but note that the corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than the results we present here.)

### 2020-02-08 nlp_models.tex
updated main sections

@@
-\section{Comparison of NLP Models}
+section{Comparison of NLP Models}
 \label{sxn:nlp}
@@
 
-\begin{figure}
+\paragraph{Scale Collapse in Poorly Trained Models}
+However, as seen in Figure \ref{fig:GPT-snorm-hist},
+the ``bad'' GPT model has, spuriously, many smaller Spectral Norms $(\log\Vert\mathbf{W}\Vert_{\infty})$,
+compared to the ``good'' GPT2--which violates the belief that smaller Spectral Norms are always better.
+Indeed, because there are so many anonymously small $\Vert\mathbf{W}\Vert_{\infty}$,
+it appears that the GPT model may be exhibiting the \emph{scale collapse} observed
+in the distilled CV models (above).
+This is an extremely important observation because it demonstrates that while the Spectral Norm
+may correlate well with predicted test error, it is not a good indicator of the overall quality of a model,
+and using it as an empirical metric may give spurious results when applied to poorly trained
+or otherwise deficient models.
+
+\begin{figure}[h]
     \centering
@@
 $(\log\Vert\mathbf{W}\Vert_{\infty})$, for the OpenAI GPT and GPT2 (small) pretrained models.}
+
 \label{fig:GPT-hist}
 \end{figure}
-
-\paragraph{Scale Collapse.}
-However, as seen in Figure \ref{fig:GPT-snorm-hist},
-the ``bad'' GPT model has, spuriously, many smaller Spectral Norms $(\log\Vert\mathbf{W}\Vert_{\infty})$,
-compared to the ``good'' GPT2--which violates the belief that smaller Spectral Norms are always better.
-Indeed, because there are so many anonymously small $\Vert\mathbf{W}\Vert_{\infty}$,
-it appears that the GPT model may be exhibiting a kind of \emph{scale collapse} and/or undesired
-loss of Frobeinus mass.
-This is an extremely important observation because it demonstrates that while the Spectral Norm
-may correlate well with predicted test error, it is not a good indicator of the overall quality of a model,
-and using it as an empirical metric may give spurious results when applied to poorly trained
-or otherwise deficient models.
 
@@
 From Figure \ref{fig:gpt-snorm-layer} (below), we see that these correspond to the first embedding layer(s).
-These layers appear to have a different effectvie normalization, and therefore a different scale.
-For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,
-but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
- $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we treat all layers as-is (although one may to normalize
-the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier).
+These layers appear to have a different effective normalization, and therefore a different scale.
+\charles{We discuss this further in the Appendix}
+%For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,
+%but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
+% $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we treat all layers as-is (although one may to normalize
+%the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier).
 Here, we do not include them in our computed average metrics in Table \ref{table:nlp},
@@
 
-\begin{figure}[H]
+\begin{figure}[htb]
     \centering
@@
 
-\begin{figure}[t]
+\begin{figure}[htb]
     \centering
@@
     }
-    %\qquad
-    %%\subfigure[log Spectral Norm]{
-    %    \includegraphics[width=4.0cm]{img/GPT2*_fnl_spectralnorm_hist.png}
-    %    \label{fig:gpt2-snorm-hist}
-   % }
-    %\qquad
     \subfigure[log Alpha Norm]{
@@
 \end{figure}
-
-?

### 2020-02-08 appendix.tex
updated main sections

@@
 The only exception to this is that we do rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $k/\sqrt{2}$ so that they are on the same scale as the Linear layers.
+
+Special consideration is necessary for NLP models.
+For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,
+but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
+ $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we treat all layers as-is (although one may to normalize
+the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier).
 

### 2020-02-08 appendix.tex
updated main sections

@@
 
-\subsection{Convolutional layers and normalization issues}
+\subsection{Reproducibility and Normalization issues}
 
@@
 
-Special consideration is necessary for NLP models.
+\paragraph{Special consideration for NLP models.}
+NLP models, and other models with large initial embeddings require special care because the
+embedding layers frequently lack the implicit $latex\frac{1}{\sqrt{N}}$ normalization present in other layers.
 For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,

### 2020-02-08 appendix.tex
appendix

@@
 
-\subsection{Reproducibility and Normalization issues}
+\subsection{Reproducibility Considerations}
 
-\paragraph{Spectral Analysis of Convolutional 2D Layers.}
+\paragraph{SVD of Convolutional 2D Layers.}
 
-There is some ambiguity in performing spectral analysis on Convolutional 2D (Conv2D) layers.  
-A Conv2D layer can be represented as a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ filter (or kernel) and $in\_ch$ / $out\_ch$ input / output channels, respectively (usually $in\_ch\le out\_ch$).
-Typically, $w=h=k$,  giving $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\_ch\times out\_ch)$ each. 
+There is some ambiguity in performing spectral analysis on Conv2D layers.  
+Each layer is a 4-index tensor of dimension $(w,h,in,out)$, with an $(w\times h)$ filter (or kernel) and $(in, out)$
+channels. When $w=h=k$,  giving $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\times out)$ each. 
 %
-There are at least three different approaches that have been advocated for applying the Singular Values Decomposition (SVD) to an Conv2D layer:
-run an SVD on each of the pre-Activation Maps $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values; 
-stack the feature maps into a single rectangular matrix of, say, dimension $((k\times k\times out\_ch)\times in\_ch)$, yielding $in\_ch$ singular values;
-compute the 2D Fourier Transform (FFT) for each of the $(in\_ch, out\_ch)$ pairs, and run SVD on the resulting Fourier coeffients~\cite{Long2019}, leading to $\sim(k\times in\_ch\times out\_ch)$ non-zero singular values.
+We identify 3 different approaches for running SVD on a Conv2D layer:
+\begin{enumerate}
+\item run SVD on each pre-Activation Map $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values
+\item stack the maps into a single matrix of, say, dimension $((k\times k\times out)\times in)$, run SVD to get $in$ singular values
+\item compute the 2D Fourier Transform (FFT) for each of the $(in, out)$ pairs, and run SVD on the Fourier coeffients~\cite{Long2019}, leading to $\sim(k\times in\times out)$ non-zero singular values.
+\end{enumerate}
 Each method has tradeoffs.  
-In principle, the third method is mathematically sound, but it is computationally expensive. 
-For our analysis, because we are performing tens of thousands of calculations, we select the first method, which is numerically the fastest and is easiest to reproduce.%
-\footnote{We provide a Google Colab notebook where all results can be reproduced, with the option to redo the calculations with the third option for the SVD of the Conv2D.}
+Method (3) is mathematically sound, but computationally expensive. Method (2) is ambiguous.
+For our analysis, because we need thousands of runs, we select method (1), which is the fastest (and is easiest to reproduce).
 
@@
 Normalization is an important, if underappreciated, practical issue.
-Importantly, the normalization of weight matrices does \emph{not} affect the PL fits because the PL exponent $\alpha$ (as well as other metrics such as the Stable Rank and MP Soft Rank~\cite{MM18_TR,MM19_HTSR_ICML}) is scale-invariant.
-Norm-based metrics, however, do depend strongly on the scale of the weight matrix--\nred{that's the point.}
-\nred{Indeed, early theoretical work by Bartlett suggests that the test accuracy depends strongly on the ``total size'' of the weight matrics.}
-Typically, to apply RMT, we would usually define Correlation Matrix with $1/N$ normalization and assume that the variance of $\mathbf{X}$ is either unity or a known constant.% 
-\footnote{For Heavy Tailed theorems, one typically needs a normalization such as \nred{$1/N^{\alpha-1}$. check this}}
-Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with the variance already normalized to $\sqrt{1/N}$, or some variant of this, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.
-We do not have conrol over the final empirical normalization of these models; and we do \emph{not} normalize (or renormalize) the Empirical Correlation Matrices, i.e., we use them as-is.
-The only exception to this is that we do rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $k/\sqrt{2}$ so that they are on the same scale as the Linear layers.
+Importantly, the normalization of weight matrices does \emph{not} affect the PL fits because $\alpha$ is scale-invariant.
+Norm-based metrics, however, do depend strongly on the scale of the weight matrix--\nred{that is the point.}
+%\nred{Indeed, early theoretical work by Bartlett suggests that the test accuracy depends strongly on the ``total size'' of the weight matrics.}
+To apply RMT, we usually define $\mathbf{X}$ with $1/N$ normalization, assuming variance of $\sigma^{2}=1.0$.
+%\footnote{For Heavy Tailed theorems, one typically needs a normalization such as \nred{$1/N^{\alpha-1}$. check this}}
+Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with
+ $\sigma^{2}\sim 1/\sqrt{N}$, or some variant, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers. With this implicit scale, 
+we do \emph{not} ``renormalize'' the empirical weight matrices, i.e., we use them as-is.
+The only exception is that we do rescale the Conv2D pre-activation maps $\mathbf{W}_{i,L}$ 
+by $k/\sqrt{2}$ so that they are on the same scale as the Linear / Fully Connected (FC) layers.
 

### 2020-02-08 appendix.tex
appendix

@@
 
-\subsection{Reproducing results on hundreds of models}
+\subsection{Reproducing the CV and NLP models }
+
+We provide a github repository for this paper that includes jupyter notsbooks that fully reproduce all results.
+All results have been produced using the weightwatcher tool (v0.2.7).
+The ImageNet and OpenAI GPT pretrained models are provided in the current pyTorch, torchvision, and huggingface distributions,
+as specified in the \texttt{requirements.txt} file.
+The ImageNet-1K models, and other models in section 6), are taken from the \textttr{omsr/imgclsmob}
+``Sandbox for training convolutional networks for computer vision'' github repository.
+
+\begin{table}[t]
+\small
+\begin{center}
+\begin{tabular}{|p{1in}|c|}
+\hline
+Figure & Notebook \\
+\hline
+1  &  WeightWatcher-VGG.ipynb \\
+2(a)  &  WeightWatcher-ResNet.ipynb \\
+2(b)  &  WeightWatcher-ResNet-1K.ipynb \\
+3(a)  &  WeightWatcher-VGG.ipynb \\
+3(b)  &  WeightWatcher-ResNet.ipynb \\
+3(c)  &  WeightWatcher-DenseNet.ipynb \\
+\hline
+5 & WeightWatcher-OpenAI-GPT2.ipynb \\
+6, 7 & WeightWatcher-OpenAI-GPT2.ipynb \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Jupyter notesbooks used to reproduce sections 4 and 5}
+\label{table:notebooks}
+\end{table}
+
+\subsection{Reproducing Section 6}
 

### 2020-02-08 appendix.tex
appendix

@@
 as specified in the \texttt{requirements.txt} file.
-The ImageNet-1K models, and other models in section 6), are taken from the \textttr{omsr/imgclsmob}
-``Sandbox for training convolutional networks for computer vision'' github repository.
 
@@
 \hline
-Figure & Notebook \\
+Figure & Jupyter Notebook \\
 \hline
@@
 \end{center}
-\caption{Jupyter notesbooks used to reproduce sections 4 and 5}
+\caption{Jupyter notebooks used to reproduce Tables 1 and 2, and Figures 1,2,3,5,6, and 7.}
 \label{table:notebooks}
@@
 
-\charles{This section needs a lot of work}
+We provide several Google Colab notebooks which can be used to reproduce the result in Table 3.
+The ImageNet-1K and other pretrained models are taken from the pytorch models in the \texttt{omsr/imgclsmob} 
+``Sandbox for training convolutional networks for computer vision'' github repository.
+\footnote{\url{https://github.com/osmr/imgclsmob}}
+The data can be generated in parallel by running each Colab notebook simultaneously on the same account,
+The final results are generated with \charles{move notebooks to repo and finish}
 
-\serena{insert link in the footnote}
+We attempt to run linear regressions for all pyTorch models for each architecture series for all datasets provided.  
+There are over $450$ models in all, and we note that the \texttt{osmr/imgclsmob} repository is constantly being updated with new models.
+We omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets as there are less than 15 models
+for those datasets. The final datasets used are shown in Table~\ref{table:datasets}
+The final architecture series used are shown in  Table~\ref{table:architectures}, with the number of models in each.
 
-and a full summary of all the models analyzed is included in the Appendix. For our analysis, we then group models by architecture and datasets for further analysis.
-
-%To further refine our analysis, we run three batches of linear regressions.
+To further explain how to reproduce our analysis, we run three batches of linear regressions.
 First at the global level, we divide models by datasets and run regression separately on all models of a certain dataset, regardless of the architecture.
 At this level, the plots are quite noisy and clustered as each architecture has its own accuracy trend but, you could still see that most plots show positive relationship with positive coefficients
+The regressions are shown in Figure~\ref{fig:DSalphas}.
 
-%\serena{see examples in Figure X}
-
-%(Here we omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets as there are less than 15 models for those datasets and thus the regression is less statistically significant)
-
-%\serena{insert plots for Figure X}
-
-%e compite the regression R-squared and RMSE  for models of each architecture-datasets combination, which shows the relationship between the progression of the model accuracy and Weightwatcher metrics more clearly and precisely.
-For example, as you could see in the Figure X2,
-
-%\serena{Add an example, UPDATE when we have the results from the new codes}
-
-For each regression, we record the R-squared and mean squared errors (MSE). We filter out regressions with less than five datapoints,
- \nred{and models with structural outliers (where the trend is anti-correlated?) ?}
-
-\charles{Still preliminary, only 309 data points here}
-\paragraph{Results}
-
-\charles{These tables could go to appendix. We need references}
+To generate the results in Table 3, we run linear for each architecture series in Table~\ref{table:architectures},
+regressing each empirical log norm metric against the reported Top 1 (and Top 5) errors (as listed on the \texttt{osmr/imgclsmob} github 
+repository README file. 
+We filter out regressions with less than five datapoints.
+We record the R-squared and mean squared errors (MSE). 
+The final results for all models is provided in the \texttt{XXXXXX}.
 
@@
 
-\charles{We actually don't run regressions on all these datasetsm, BUT we could present them in the Figure below to show that alpha is a good metric for these kinds of models, in contrast to the NLP, where alpha is frequently too large to run a regression}
-
 \begin{figure}[t]
@@
 
-    \label{fig:dataset_alphas}
+    \label{fig:DSalphas}
 \end{figure}

### 2020-02-08 appendix.tex
appendix

@@
 
-\subsection{Reproducing the CV and NLP models }
+\subsection{Reproducing Sections 4 and 5}
 
@@
 
-\subsection{Reproducing Section 6}
+\subsection{Reproducing Section 3, Distiller Model}
+
+\subsection{Reproducing Table 3, Section 6}
 
@@
 
-\section{XXX: PLACEHOLDER STUFF PROBABLY TO BE REMOVED}
-
-\subsection{XXX: PLACEHOLDER STUFF PROBABLY TO BE REMOVED}
-
-\michael{I put this here for now as a placeholder.  Where to put it.  Maybe in a discussion/conclusion if there is space.}
-
-\paragraph{XXX.}
-XXX.  THIS VERIFICATION IS NOT ABOUT CONV LAYERS, IT IS ABOUT PL MORE GENERALLY, CORRECT?  WE SHOULD CLARIFY AND SQUISH.
-To verify that our approach is meaningful, we need to confirm that the ESD is neither due to a random matrix, nor due to unsually large matrix elements, but, in fact, captures correlations learned from the data.
-We examine typical layer for the pretained AlexNet model (distributed with pyTorch). 
-Figure~\ref{fig:alexnet1} displays the ESD for the first slice (or matrix $\mathbf{W}$) of the third Conv2D layer, extracted from a 4-index Tensor of shape $(384, 192, 3, 3)$.
-The red line displays the best fit to a random matrix, using the Marchenko pastur theory~\cite{MM18_TR}.
-We can see the random matrix model does not describe the ESD very well.
-For comparison, Figure \ref{fig:alexnet2} shows the ESD of the same matrix, randomly shuffled; here looks similar to the red line plot of the orginal ESD.
-In fact, the empircal ESD is better modeled with a truncated power law distribtion.
-\michael{We may want to give a one-sentence summary of this par and fig at the end of the previous par.}
-\charles{Here, on the RMT MP stuff, I think it makes sense to point back}
-
-\begin{figure}[H]
-   \centering
-   \subfigure[Actual ESD]{
-     %\includegraphics[scale=0.5]{img/alexnet1.png} 
-     \includegraphics[scale=0.25]{img/alexnet1.png} 
-     \label{fig:alexnet1}
-   }
-   \subfigure[ESD of randomly shuffled matrix]{
-      %\includegraphics[scale=0.5]{img/alexnet2.png}
-      \includegraphics[scale=0.25]{img/alexnet2.png}
-      \label{fig:alexnet2}
-   }
-   \caption{ESD of AlexNet Conv2D pre-Activation map for Layer 3 Slice 1, actual and randomized.
-            \michael{Need better figs here.}
-           }
-   \label{fig:alexnet}
-\end{figure}
-
-Although the ESD is \emph{Heavy Tailed}, this does not imply that the orginal matrix $\mathbf{W}$ is itself heavy tailed--only the correlation matrix $\mathbf{X}$ is. 
-If $\mathbf{W}$ was, then it would contain 1 or more unusually large matrix elements, and they would dominate the ESD.  
-Of course the randomized $\mathbf{W}$ would also be heavy tailed, but its ESD neither resembles the original nor is it heavy tailed. 
-So we can rule out $\mathbf{W}$ being heavy tailed.
-\michael{These comments seem out of place, since they hold more generally than for the Conv2D layers.}
-\charles{Agreed. We could move this up.  We have never really talked about this, but it is essential to explain the difference between assuming W is heavy tailed , which confuses everyone}
-
-These plots tell us that the pre-activation maps of the Conv2D contains significant correlations learned from the data.  
-By modeling the ESD with a power law distribution $\lambda^{\alpha}$, we can characterize the amount of correlation learned;
-the smaller the exponent $\alpha$, the more correlation in the weight matrix. 
-\michael{These comments seem out of place, since they hold more generally than for the Conv2D layers.}
-
 \subsection{XXX: PLACEHOLDER STUFF PROBABLY TO BE REMOVED}

### 2020-02-09 submission/background.tex
clean up for submission

--- 2020-02-02 background.tex
@@
 our best-performing metrics (those using fitted PL exponents) are based on statistical mechanics~\cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM19_KDD, MM20_SDM}, in particular the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
-However, we emphasize that the way in which we \emph{use} statistical mechanics theory is quite different than the way it is more commonly formulated.
+We emphasize that the way in which we (and HT-SR Theory) \emph{use} statistical mechanics theory is quite different than the way it is more commonly formulated.
 Several very good overviews of the more common approach are available~\cite{EB01_BOOK, BKPx20}.
-We will \emph{use} statistical mechanics in a broader sense, drawing upon results from quantitative finance and random matrix theory (RMT).
-Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statisical mechanics of heavy tailed and strongly correlated systems\charles{, and current work (in progress)\cite{blog}}
+We \emph{use} statistical mechanics in a broader sense, drawing upon results from quantitative finance and random matrix theory.  % (RMT).
+Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statisical mechanics of heavy tailed and strongly correlated systems.
+% Probably best not to cite at this point %\charles{, and current work (in progress)\cite{blog}}
 
@@
 In this area, theoretical work aims to prove generalization bounds, and applied work uses these norms as a regularization function to improve training.
-While we do find that norms provide relatively good quality metrics, at least for distinguishing good-better-best among well-trained models, we are not interested in proving generalization bounds \nred{or developing new regularizers.}
+While we do find that norms provide relatively good quality metrics, at least for distinguishing ``good-better-best'' among well-trained models, we are not interested in proving generalization bounds or developing new regularizers.
 
@@
 
-\charles{see comments in tex file}
+\charles{see comments in tex file}\michael{Let's touch base, not sure what to do.}
 % we adress this above

### 2020-02-09 submission/conclusion.tex
clean up for submission

--- 2020-02-02 conclusion.tex
@@
 Our results, which are based on our \emph{use} of sophisticated statistical mechanics theory to solve an important practical DNN problems, suggests that this approach should be of interest more generally for those interested in developing a practical DNN theory.
-
-\charles{it is the oppposite...ab initio theory scales quite well...it is the spherical cow models of physics and ML that do not scale.
+\charles{see latex file}
+%\charles{it is the oppposite...ab initio theory scales quite well...it is the spherical cow models of physics and ML that do not scale.
 What we have is a semi-empirical theory.

### 2020-02-09 submission/introduction.tex
clean up for submission

--- 2020-02-02 introduction.tex
@@
 %, in particular in academic environments, 
-is to train a model and then evaluate its training/testing error--but
-there are many problems with this approach.
+is to train a model and then evaluate its training/testing error.
+There are many problems with this approach.
 The training/testing curves give very limited insight into the overall properties of the model, 
@@
 and so on.
-A less well-known problem, but one that is increasingly important (in particular in industrial-scale 
-artificial intelligence (AI), arises when the model \emph{user} is not the model \emph{developer}.
-here, one may not have access to either the training data nor the testing data.
-\nred{Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it "as is'', or fine tune it and/or compress it.}
+A less well-known problem, but one that is increasingly important, in particular in industrial-scale artificial intelligence (AI), arises when the model \emph{user} is not the model \emph{developer}.
+Here, one may not have access to either the training data nor the testing data.
+Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it "as is'', or to fine-tune and/or compress it and then use it.
 
@@
 
-Moreover, for many large scale, practical applications, one can not define a true test metric.
-For example, NLP models that generate fake text or conversational chatbots may use a proxy, like perplexity, as a test metric, but, in the end, really require human evaluation.
-Or, say, models that cluster user profiles, which are widely used in marketing, are unsupervised and have no labels for comparison.
-\charles{more here, clean up?}
+Moreover, for many large scale, practical applications, there is no obvious way to define an ideal test metric. 
+For example, models that generate fake text or conversational chatbots may use a proxy, like perplexity, as a test metric.
+In the end, however, they really require human evaluation. 
+Alternatively, models that cluster user profiles, which are widely used in marketing, are unsupervised and have no obvious labels for comparison and/or evaluation.
 
 Most importantly, in industry, one faces unique practical problems such as: do we have enough data for this model? 
-\nred{Indeed, high quality, labeled data can be very expensive to acquire, and this cost can make or break a project.}
+Indeed, high quality, labeled data can be very expensive to acquire, and this cost can make or break a project.
 Methods that are developed and evaluated on any well-defined publicly-available coprus of data, no matter how large or diverse or interesting, are clearly not going to be well-suited to address problems such as this.
@@
 In this paper, we present and evaluate quality metrics for pretrained deep neural network (DNN) models, and at scale.
-To do so, we consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
+We consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
 %
@@
 The \emph{quality metrics} we consider are based on the spectral properties of the layer weight matrics.
-In particular, they are based on the norms of weight matrices (such norms have been used in traditional statistical learning theory to bound capacity and construct regularizers) and parameters of power law (PL) fits of the eigenvalues of weight matrices (such PL fits are based on statistical mechanics approaches to DNNs).
+They are based on norms of weight matrices (such norms have been used in traditional statistical learning theory to bound capacity and construct regularizers) and/or parameters of power law (PL) fits of the eigenvalues of weight matrices (such PL fits are based on statistical mechanics approaches to DNNs).
 Note that, while we use traditional norm-based and PL-based metrics, our goals are different.
-In particular, unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., hyperparmeter optimization).%
+Unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., hyperparmeter optimization).%
 \footnote{One could of course use these techniques to improve training, and we have been asked about that, but we are not interested in that here.
@@
 %Third, 
-We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best'' models, but they can qualitatively fail for models that may not be well-trained, i.e., to distinguish ``good-vs-bad'' models.
+We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best'' models, but they can qualitatively fail for models that may not be well-trained, i.e., to distinguish ``good-versus-bad'' models.
 \item 
 %Fourth, 
-We find that PL-based metrics do much better, quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-vs-bad models.
+We find that PL-based metrics do much better---quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-versus-bad models.
 \item 
 %Finally, 
-We find that PL-based metrics can also be used to characterize fine-scale properties of models, including understanding layer-wise \emph{correlation flow}, and evaluating post-training modifications such as model distillation and model improvement.
+We find that PL-based metrics can also be used to characterize fine-scale model properties, including layer-wise \emph{correlation flow}, and evaluating post-training modifications such as model distillation, increasing the dataset size, and other model improvements.
 \end{itemize}
@@
 We emphasize that our goal is a practical theory to predict trends in the quality of state-of-the-art DNN models, i.e., not to make a statement about every publicly-available model.
-\nred{We have examined hundreds of models, we indentify the general trends, and highlight interesting exceptions.}
+We have examined hundreds of models, and we indentify general trends, but we also highlight interesting exceptions.
 %%Several of the most interesting are described below.
@@
 and with fully reproducible Google colab notebooks for the larger calculations.%
-\footnote{\url{https://github.com/CalculatedContent/kdd2020} \michael{DIFFERENT NAME?}\charles{suggestions?}}
-\michael{LINK TO: Google Colab, IS THIS THAT REPO.}
-\michael{FIX THIS PAR.}
+\footnote{\url{https://github.com/CalculatedContent/kdd2020} \michael{different name?}\charles{suggestions?}\michael{Anonymous, just for anonymized submission.}}
+\michael{FIX THIS LINK TO: Google Colab OR THE REPO.}
 
@@
 Then, in Section~\ref{sxn:nlp}, we look at several variations of a popular NLP DNN architecture (the OpenAI GPT and GPT2 models); and we show how model quality and properties vary between several variants of GPT and GPT2, including how metrics behave similarly and differently.
-Then, in Section \ref{sxn:all_cv_models}, we present results based on an analysis of hundreds of pretraind DNN CV models, showing how well each metric is correlated with the reported test accuracies, and how the Alpha-Norm metric(s) perform remarkably well.
+Then, in Section \ref{sxn:all_cv_models}, we present results based on an analysis of hundreds of pretraind DNN CV models, showing how well each metric predicts the reported test accuracies, and how the Alpha-Norm metric(s) perform remarkably well.
 Finally, in Section~\ref{sxn:conc}, we provide a brief discussion and conclusion.

### 2020-02-09 submission/nlp_models.tex
clean up for submission

--- 2020-02-08 nlp_models.tex
@@
 
-\paragraph{Coarse Analysis: Empirical Quality Metrics for GPT and GPT2.}
+\paragraph{Average Empirical Quality Metrics for GPT and GPT2.}
 
@@
 
-\paragraph{Layer Analysis: Correlation Flow in GPT and GPT2.}
-
+\paragraph{Correlation Flow in GPT and GPT2.} 
+The correlation flow ($\alpha$ as a function of depth / layer id)
 also differs significantly between GPT and GPT2.

### 2020-02-09 submission/appendix.tex
clean up for submission

--- 2020-02-08 appendix.tex
@@
 \hline
-5 & WeightWatcher-OpenAI-GPT2.ipynb \\
+4 & WeightWatcher-Intel-Distiller-ResNet20.ipynb \\
+\hline
+5 & WeightWatcher-OpenAI-GPT.ipynb \\
 6, 7 & WeightWatcher-OpenAI-GPT2.ipynb \\
@@
 \end{center}
-\caption{Jupyter notebooks used to reproduce Tables 1 and 2, and Figures 1,2,3,5,6, and 7.}
+\caption{Jupyter notebooks used to reproduce all results in sections ~\ref{sxn:cv} and ~\ref{sxn:nlp}}
 \label{table:notebooks}
@@
 
-\subsection{Reproducing Section 3, Distiller Model}
+\subsection{Reproducing Figure~\ref{fig:resnet204D5L}, Distiller Model}
 
-\subsection{Reproducing Table 3, Section 6}
+We provide the orginal Jupyter Notebooks, which uses the Intel \texttt{distiller} framework
+.\footnote{\url{https://nervanasystems.github.io/distiller}} in the \texttt{distiller} folder of our 
+github repo. Figure ~\ref{fig:resnet204D5L} is from the  \texttt{``...-Distiller-ResNet20.ipynb''} 
+notebook (see Table~\ref{table:notebooks}).  
+For completeness, we provide both the results described here, and additional results
+on other pretrained and distilled models using the \texttt{weightwatcher} tool.
 
-We provide several Google Colab notebooks which can be used to reproduce the result in Table 3.
+\subsection{Reproducing Table 3, Section~\ref{xsn:all}}
+
+We provide several Google Colab notebooks which can be used to reproduce the results in Table 3,
+in the \texttt{ww-colab} folder in our github repo.  
 The ImageNet-1K and other pretrained models are taken from the pytorch models in the \texttt{omsr/imgclsmob} 
-``Sandbox for training convolutional networks for computer vision'' github repository.
-\footnote{\url{https://github.com/osmr/imgclsmob}}
-The data can be generated in parallel by running each Colab notebook simultaneously on the same account,
-The final results are generated with \charles{move notebooks to repo and finish}
+``Sandbox for training convolutional networks for computer vision'' github repository
+.\footnote{\url{https://github.com/osmr/imgclsmob}}
+The data can be generated in parallel by running each Google Colab notebook (i.e. \texttt{ww\_kdd2020\_0\_100.ipynb}) 
+simultaneously on the same account.
+The final results are generated with the \texttt{ww\_kdd2020\_results.ipynb} notebook.
 
@@
 regressing each empirical log norm metric against the reported Top 1 (and Top 5) errors (as listed on the \texttt{osmr/imgclsmob} github 
-repository README file. 
+repository README file, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}
 We filter out regressions with less than five datapoints.
 We record the R-squared and mean squared errors (MSE). 
-The final results for all models is provided in the \texttt{XXXXXX}.
+The final results for all series regressions are provided in the \texttt{ww-colab/df\_all.xslx} python pandas dataframe file.
 

### 2020-02-09 submission/introduction.tex
clean up for submission

@@
 Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained DNN models.}
-Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves, and in both cases \emph{we aim to predict trends in the quality of pre-trained models without access to training or testing data}.
+Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves, and in both cases, we ask
 
-In more detail, our main contributions are the following.
-\begin{itemize}[leftmargin=*]
-\item 
+\emph{Can we predict trends in the quality of pre-trained models without access to training or testing data ?}.
+
+%\begin{itemize}[leftmargin=*]
+%\item 
 %First, 
-Motivated by practical problems, we formulate the question ``Can one predict trends in the quality of state-of-the-art neural networks without access to training or testing data?''
-\item
+%Motivated by practical problems, we formulate the question ``Can one predict trends in the quality of state-of-the-art neural networks without access to training or testing data''
+%\item
 %Second, 
 %To answer this question, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in CV and NLP.
-To answer this question, we analyze hundreds of publicly-available pretrained state-of-the-art models from CV and NLP.
+To answer this, we analyze hundreds of publicly-available pretrained state-of-the-art CV and NLP models. And we find that:
+%\end{itemize}
+%Third, 
+%\paragraph{Our main contributions are:}
+\begin{itemize}
 \item
-%Third, 
-We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best'' models, but they can qualitatively fail for models that may not be well-trained, i.e., to distinguish ``good-versus-bad'' models.
+Norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best'' models, but they can qualitatively fail for models that may not be well-trained, i.e., to distinguish ``good-versus-bad'' models.
 \item 
 %Fourth, 
-We find that PL-based metrics do much better---quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-versus-bad models.
+Power Law (PL)-based metrics do much better---quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-versus-bad models.
 \item 
 %Finally, 
-We find that PL-based metrics can also be used to characterize fine-scale model properties, including layer-wise \emph{correlation flow}, and evaluating post-training modifications such as model distillation, increasing the dataset size, and other model improvements.
+PL-based metrics can also be used to characterize fine-scale model properties, including layer-wise \emph{correlation flow}, and evaluating post-training modifications such as model distillation, increasing the dataset size, and other model improvements.
 \end{itemize}
@@
 All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package};
-and, in order to make our results fully reproducible, we provide Jupyter notebooks in the github repo for this paper~\cite{repo},
-and with fully reproducible Google colab notebooks for the larger calculations.%
-\footnote{\url{https://github.com/CalculatedContent/kdd2020} \michael{different name?}\charles{suggestions?}\michael{Anonymous, just for anonymized submission.}}
-\michael{FIX THIS LINK TO: Google Colab OR THE REPO.}
+and, to be fully reproducible, we only examine publically available, pretrained models, and
+also provide all Jupyter and Google Colab notebooks used accompanying github 
+repository\footnote{\url{https://github.com/CalculatedContent/kdd2020}
+\michael{To be anonymized}}
+See the Appendix for more details on how to reproduce all results.
 

### 2020-02-09 submission/methods.tex
clean up for submission

--- 2020-02-08 methods.tex
@@
 \end{equation}
+
+Each DNN layer contains one or more layer 2D  $N\times M$ weight matrices, $\mathbf{W}_{l}$ or pre-activation maps, $\mathbf{W}_{i,l}$, extracted from 2D Convolutional layers, and where $N > M$.
+\footnote{We do not use intra-layer information from the models in our quality metrics, but as we will describe our metrics can be used to learn about that.}
+Notice we may drop the $i$ and/or $i,l$ subscripts below.
+See the Appendix for how we define the Conv2D layer matrixes, and our choices of normalization.
+
 Assume we are given several pretrained DNNs, e.g., as part of an architecture series.
-The models have been trained on (unspecified and unavailable) labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using Backprop, by minimizing some (also unspecified and unavailable) loss function $\mathcal{L}(\cdot)$.
+The models have been trained and evaluated on labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$,
+using standard techniques.  The pretained pytroch model files are publically available,
+and the test accuracies have been reported online.  
+In this study, we do not have access to this data, and we have not trained any of the models ourselves,
+nor have we re-evaluated the test accuracies.
 We expect that most well-trained, production-quality models will employ one or more forms of on regularization, such as Batch Normalization, Dropout, etc., and many will also contain additional structure such as Skip Connections, etc.
-Here, we will ignore these details, and will focus only on the layer weight matrices $\mathbf{W}_{l}$.
+Here, we will ignore these details, and will focus only on the pretrained layer weight matrices $\mathbf{W}_{l}$.
 
-\paragraph{DNN Quality Metrics.}
+\paragraph{DNN Empirical Quality Metrics.}
 
-Each DNN layer contains one or more layer 2D weight matrices, $\mathbf{W}_{l}$, and/or 2D feature maps, $\mathbf{W}_{i,l}$, extracted from 2D Convolutional layers. 
-(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below; and we let $\mathbf{W}$ be a generic $N\times M$ weight matrix, with $N\ge M$.) 
-%
-We have examined a large number of possible quality metrics.
-The best performing metrics (recall that we can only consider metrics that do not use training/test data) depend on the norm and/or spectral properties of weight matrices, $\mathbf{W}$.%
-\footnote{We do not use intra-layer information from the models in our quality metrics, but as we will describe our metrics can be used to learn about that.}
-Consider the following.
-\charles{Note I write X for the $\alpha$-norm and not W.  This is because it would be $2\alpha$.  I know its a bit confusing.
-We have to be careful here with the Frobenius norm also and check that we are not off by a factor of 2}
+The best performing empirical quality metrics depend on the norms and/or spectral properties of each weight matrix,
+$\mathbf{W}$ and/or, equivalently, it's \emph{Empirical Correlation Matrix}: $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$ .%
+
+Here, we consider the following metrics:.
+
 \begin{itemize}
 \item 
-Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\sum_{i,j}w^{2}_{i,j} = \sum_{i=1}^{M} \lambda_{i}^{2}$
+Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{X}\Vert_{F}=\sum_{i=1}^{M} \lambda_{i}$
 \item 
-Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
-\michael{Should this be $\sqrt{\lambda_{max}}$?}
+Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}^{2}=\Vert\mathbf{X}\Vert_{\infty}=\lambda_{max}$
 \item 
@@
 \item 
-$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda_{i}^{\alpha}$
+$\alpha$-Norm (or $\alpha$-Shatten Norm):
+\footnote{Notice  $\Vert\mathbf{W}\Vert^{2\alpha}_{2\alpha}=\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$. We use $\mathbf{X}$ to emphasize that $\alpha$ depends on the ESD of $\mathbf{X}$}
+ $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda_{i}^{\alpha}$
 \end{itemize}
-The first two quantities are well-known in ML.
-The last two deserve special mention.
-For all these quantities, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the \emph{Empirical Correlation Matrix},
-$ %% $$
-\mathbf{X}=\mathbf{W}^{T}\mathbf{W} ,
-$ %% $$
-and $\lambda_{max}$ is the maximum eigenvalue of $\mathbf{X}$. 
-(These eigenvalues are the squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, i.e., $\lambda_{i}=\sigma^{2}_{i}$.)
-For the last two quantities, $\alpha$ is the PL exponent that arises in the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+
+The  $\lambda_{i}$ is the $i^{th}$ eigenvalue of the $\mathbf{X}$, 
+and $\lambda_{max}$ is the maximum eigenvalue.
+Note the squared singular values $\sigma_{i}$ of $\mathbf{W}$, $\lambda_{i}=\sigma^{2}_{i}$.)
+Also, note we do \emph{not} normalize $\mathbf{X}$ by $1/N$; see the Appendix for a discussion why.
+
+The first two norms are well-known in ML; the last two deserve special mention.
+The empirical parameter $\alpha$ is the Power Law (PL) exponent that arises in the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
 Operationally, $\alpha$ is determined by using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated PL,
@@
 For the \emph{Frobenius norm metric} and \emph{Spectral norm metric}, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly. 
-(When taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.)
+\footnote{When taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.}
 For the $\alpha$-Shatten Norm metric, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it cannot be taken out of the sum.
 
-For the \emph{Weighted Alpha metric}, we average of $\hat{\alpha}$ over all layers, i.e., 
-$ %$$
-\sum_l \hat{\alpha}_l = \sum_l \alpha_l\log\lambda_{max,l}  .  
-$ %$$
-%\michael{Be careful to get right.}
+\emph{The Weighted Alpha metric} is an average of $\hat{\alpha}$ over all layers, weighted by the size, or scale, or each matrix
+(and approximates the average log $\alpha$-Norm metric).
+\begin{equation}
+\hat{\alpha} = \dfrac{1}{L}\sum_l \alpha_l\log\lambda_{max,l}\approx\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle  
+\end{equation}
+where $L$ is the total nuumber of layer weight matrices.
 The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a limited set of models.
 Based on this, in this paper, we introduce and evaluate the $\alpha$-Norm metric.
-One would expect $\hat{\alpha}$ to approximate the log $\alpha$-Norm very well for $\alpha < 2$ and reasoably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.
+One expects$\hat{\alpha}$ approximates the log $\alpha$-Norm very well for $\alpha < 2$ and reasoably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.
 

### 2020-02-09 submission/abstract_file.tex
clean up for submission

--- 2020-02-07 abstract_file.tex
@@
+\charles{THis IS TOO LONG AND TO MANY (). IT DOES NOT FLOW.  TOO COMPLICATED FOR INDUSTRY PEOPLE--they are simpletons}
 In many practical applications, one works with deep neural networks (DNNs) models trained by someone else.

### 2020-02-09 submission/abstract_file.tex
clean up for submission

@@
-\charles{THis IS TOO LONG AND TO MANY (). IT DOES NOT FLOW.  TOO COMPLICATED FOR INDUSTRY PEOPLE--they are simpletons}
+\charles{THis IS TOO LONG AND TO MANY (). IT DOES NOT FLOW.  TOO TIME CONSUMING FOR INDUSTRY PEOPLE..THEY NEED TO GET THE POINT IN 30 seconds or they wil not read it}
 In many practical applications, one works with deep neural networks (DNNs) models trained by someone else.

### 2020-02-09 submission/methods.tex
clean up for submission

@@
 and $\lambda_{max}$ is the maximum eigenvalue.
-Note the squared singular values $\sigma_{i}$ of $\mathbf{W}$, $\lambda_{i}=\sigma^{2}_{i}$.)
-Also, note we do \emph{not} normalize $\mathbf{X}$ by $1/N$; see the Appendix for a discussion why.
+Note, the eigenvalues are square of the singular values $\sigma_{i}$ of $\mathbf{W}$: $\lambda_{i}=\sigma^{2}_{i}$.
+Also, note that we do \emph{not} normalize $\mathbf{X}$ by $1/N$; see the Appendix for a discussion why.
 

### 2020-02-09 submission/abstract_file.tex
final cleanup

--- 2020-02-07 abstract_file.tex
@@
 In many practical applications, one works with deep neural networks (DNNs) models trained by someone else.
-For such \emph{pretrained models}, one typically does not have access to training data or test data, and, moreover,  does not know the details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
-Here, we present and evaluate quality metrics for pretrained neural network models at scale.
+For such \emph{pretrained models}, one typically does not have access to training data or test data, 
+and, moreover,  does not know many detals aout the  model, such as
+the specifics of the training data, the loss function, the hyperparameter values, etc.
+Given just one or more pretrained models, can one say anything about the expected performance or quality of the models ?
+Here, we present and evaluate empirical quality metrics for pretrained neural network models at scale.
 %%Our metrics are drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as Heavy-Tailed Random Matrix Theory (HT-RMT), as used in the recently-developed Theory of Heavy-Tailed Self Regularization (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models).
 %%The most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in a system) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
-The most promising are norm-based metrics (such as those used to provide capacity control in traditional statistical learning theory) and metrics based on fitting eigenvalue distributions to truncated power law (PL) distributions (which derive from statistical mechanics, in particular the recently-developed Theory of Heavy-Tailed Self Regularization, and which characterize the strength of correlations in a DNN).
+%The most promising are norm-based metrics
+% (such as those used to provide capacity control in traditional statistical learning theory) 
+%and metrics based on fitting eigenvalue distributions to truncated power law (PL) distributions 
+%(which derive from statistical mechanics, in particular the recently-developed Theory of Heavy-Tailed Self Regularization, and which characterize the strength of correlations in a DNN). 
 %
 Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision (CV) and natural language processing (NLP).
+We examine both familiar norm-based capacity control metrics (the Frobenius and  Spectral norms). as well as
+new Power-Law (PL) based metrics,  the Weighted Alpha ($\alpha$) and the $\alpha$-(Shatten) Norm,
+from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
 We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can discriminate amoing ``good-better-best''  models.

### 2020-02-09 submission/abstract_file.tex
final cleanup

@@
 We examine both familiar norm-based capacity control metrics (the Frobenius and  Spectral norms). as well as
-new Power-Law (PL) based metrics,  the Weighted Alpha ($\alpha$) and the $\alpha$-(Shatten) Norm,
+new Power-Law (PL) based metrics, the PL exponent $\alpha$ and the Weighted Alpha ($\hat{\alpha}$) metrics,
 from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
-We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can discriminate amoing ``good-better-best''  models.
-On the other hand, for poorly trained models, i.e., to distinguish ``good-versus-bad'' models---which, arguably is the point of needing quality metrics---norm-based metrics can qualitatively fail.  
-We also find that PL-based metrics do much better---quantitatively better at discriminating aming ``good-better-best'' models, and qualitatively better at discriminating ``good-versus-bad'' models.
-PL-based metrics can also be used to characterize fine-scale properties of models, e.g., understanding layer-wise \emph{correlation flow}, and evaluate post-training modifications such as model distillation, increasing the data set size, and other model improvements.
+We also introduce the $\alpha$ (Shatten) Norm metric.
+We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models
+On the other hand, for poorly trained models, they can not  to distinguish ``good-versus-bad'' models--which, 
+arguably is the point of needing quality metrics.
+We also find that PL-based metrics do much better--quantitatively-- at discriminating among
+a series of ``good-better-best'' models, and qualitatively better at discriminating ``good-versus-bad'' models.
+PL-based metrics can also be used to characterize fine-scale properties of models, and
+we the layer-wise \emph{Correlation Flow} as new quality assesment.
+We also define and show how poorly-trained (and/or poorly-finetuned) models can exhbit both \emph{Scale Collapse}
+and unsually large PL exponents $\alpha\gg 6$.  In particular, we show that the \emph{WeightWatcher} tool
+can be used to identify when a pretrained DNN has problems that can not be detected simply
+by examining the test accuracy.

### 2020-02-09 submission/abstract_file.tex
final cleanup

@@
 We also introduce the $\alpha$ (Shatten) Norm metric.
-We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models
-On the other hand, for poorly trained models, they can not  to distinguish ``good-versus-bad'' models--which, 
-arguably is the point of needing quality metrics.
+We find that norm-based metrics correlate well with reported test accuracies of well-trained models
+across nearly all CV architecture series.
+PL-based metrics can also be used to characterize fine-scale properties of these models, and
+we define the layer-wise \emph{Correlation Flow} as new quality assesment.
 We also find that PL-based metrics do much better--quantitatively-- at discriminating among
 a series of ``good-better-best'' models, and qualitatively better at discriminating ``good-versus-bad'' models.
-PL-based metrics can also be used to characterize fine-scale properties of models, and
-we the layer-wise \emph{Correlation Flow} as new quality assesment.
-We also define and show how poorly-trained (and/or poorly-finetuned) models can exhbit both \emph{Scale Collapse}
+On the other hand, we find that norm-based metrics can not to distinguish ``good-versus-bad'' models--which, 
+arguably is the point of needing quality metrics.  Indeed, they may give spurious results.
+We show how poorly-trained (and/or poorly-finetuned) models may exhbit both \emph{Scale Collapse}
 and unsually large PL exponents $\alpha\gg 6$.  In particular, we show that the \emph{WeightWatcher} tool

### 2020-02-09 submission/abstract_file.tex
final cleanup

@@
 a series of ``good-better-best'' models, and qualitatively better at discriminating ``good-versus-bad'' models.
-On the other hand, we find that norm-based metrics can not to distinguish ``good-versus-bad'' models--which, 
+On the other hand, we find that norm-based metrics can not distinguish ``good-versus-bad'' models--which, 
 arguably is the point of needing quality metrics.  Indeed, they may give spurious results.

### 2020-02-09 submission/abstract_file.tex
final cleanup

@@
 %
-Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision (CV) and natural language processing (NLP).
+Using the open-source \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision (CV) and natural language processing (NLP).
 We examine both familiar norm-based capacity control metrics (the Frobenius and  Spectral norms). as well as

### 2020-02-09 submission/abstract_file.tex
final cleanup

@@
 For such \emph{pretrained models}, one typically does not have access to training data or test data, 
-and, moreover,  does not know many detals aout the  model, such as
+and, moreover,  does not know many details aout the  model, such as
 the specifics of the training data, the loss function, the hyperparameter values, etc.

### 2020-02-09 submission/methods.tex
Update methods.tex
Corrected a typo on line 92

@@
 Based on this, in this paper, we introduce and evaluate the $\alpha$-Norm metric.
-One expects$\hat{\alpha}$ approximates the log $\alpha$-Norm very well for $\alpha < 2$ and reasoably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.
+One expects$\hat{\alpha}$ approximates the log $\alpha$-Norm very well for $\alpha < 2$ and reasonably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.
 

### 2020-02-10 submission/abstract_file.tex
full pass

--- 2020-02-09 submission/abstract_file.tex
@@
-In many practical applications, one works with deep neural networks (DNNs) models trained by someone else.
-For such \emph{pretrained models}, one typically does not have access to training data or test data, 
-and, moreover,  does not know many detals aout the  model, such as
-the specifics of the training data, the loss function, the hyperparameter values, etc.
-Given just one or more pretrained models, can one say anything about the expected performance or quality of the models ?
-Here, we present and evaluate empirical quality metrics for pretrained neural network models at scale.
+In many practical applications, one works with deep neural network (DNN) models trained by someone else.
+For such \emph{pretrained models}, one typically does not have access to training data or test data.
+Moreover, one does not know many details about the model, such as the specifics of the training data, the loss function, the hyperparameter values, etc.
+Given one or many pretrained models, can one say anything about the expected performance or quality of the models?
+Here, we present and evaluate empirical quality metrics for pretrained DNN models at scale.
 %%Our metrics are drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as Heavy-Tailed Random Matrix Theory (HT-RMT), as used in the recently-developed Theory of Heavy-Tailed Self Regularization (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models).
@@
 Using the open-source \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision (CV) and natural language processing (NLP).
-We examine both familiar norm-based capacity control metrics (the Frobenius and  Spectral norms). as well as
-new Power-Law (PL) based metrics, the PL exponent $\alpha$ and the Weighted Alpha ($\hat{\alpha}$) metrics,
-from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
-We also introduce the $\alpha$ (Shatten) Norm metric.
-We find that norm-based metrics correlate well with reported test accuracies of well-trained models
-across nearly all CV architecture series.
-PL-based metrics can also be used to characterize fine-scale properties of these models, and
-we define the layer-wise \emph{Correlation Flow} as new quality assesment.
-We also find that PL-based metrics do much better--quantitatively-- at discriminating among
-a series of ``good-better-best'' models, and qualitatively better at discriminating ``good-versus-bad'' models.
-On the other hand, we find that norm-based metrics can not distinguish ``good-versus-bad'' models--which, 
-arguably is the point of needing quality metrics.  Indeed, they may give spurious results.
-We show how poorly-trained (and/or poorly-finetuned) models may exhbit both \emph{Scale Collapse}
-and unsually large PL exponents $\alpha\gg 6$.  In particular, we show that the \emph{WeightWatcher} tool
-can be used to identify when a pretrained DNN has problems that can not be detected simply
-by examining the test accuracy.
+We examine both familiar norm-based capacity control metrics (Frobenius and Spectral norms) as well as newer Power-Law (PL) based metrics (including fitted PL exponents, $\alpha$, and the Weighted Alpha metric, $\hat{\alpha}$), from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
+We also introduce the $\alpha$-Shatten Norm metric.
+We find that norm-based metrics correlate well with reported test accuracies for well-trained models across nearly all CV architecture series.
+On the other hand, we find that norm-based metrics can not distinguish ``good-versus-bad'' models---which, arguably is the point of needing quality metrics.  
+Indeed, they may give spurious results.
+We also find that PL-based metrics do much better---quantitatively better at discriminating among a series of ``good-better-best'' models, and qualitatively better at discriminating ``good-versus-bad'' models.
+PL-based metrics can also be used to characterize fine-scale properties of these models, and we introduce the layer-wise \emph{Correlation Flow} as new quality assesment.
+We show how poorly-trained (and/or poorly fine-tuned) models may exhbit both \emph{Scale Collapse} and unsually large PL exponents, $\alpha\gg 6$, in particular for recent NLP models.
+Our techniques, as implemented in the \emph{WeightWatcher} tool, can be used to identify when a pretrained DNN has problems that can not be detected simply by examining training/test accuracies.

### 2020-02-10 submission/all_models.tex
clean up for submission
full pass

--- 2020-02-08 all_models.tex
@@
-\section{Comparison of all pretrained CV Models}
+\section{Comparing Hundreds of CV Models}
 \label{sxn:all_cv_models}
 
-\michael{Now we go back to the CV models and look at over 100 of them, and draw'
-more broad conclusions about why alpha and the new norm is better, by looking at the MSE}
+In this section, we summarize results from a large-scale analysis of hundreds of CV models, including models developed for image classification, segmentation, and a range of related tasks.  
+Our aim is to complement the detailed results from Sections~\ref{sxn:cv} and~\ref{sxn:nlp} by providing broader conclusions. % about norm-based and PL-based metrics.
+The models we consider have been pretrained on nine datasets, including ImageNet-1K, and CIFAR-10, CIFAR-100, Street View House Numbers (SVHN), Caltech-UCSD Birds-200-2011 (CUB-200-2011), Pascal VOC2012, ADE20K, Cityscapes, and Common Objects in Context (COCO).
+%%The pretrained models and their accuracy metrics are summarized in the osmr github~\cite{charles-add-link}.
+We provide more details about our experimental setup in Appendix~\ref{sxn:appendix}.
 
-Here, we use the Weightwatcher tool to analyze over 100 pretrained computer vision (CV) models Pytorch, including image classification and segmentation models.
-They have been pretrained on nine datasets, ImageNet-1K, and CIFAR-10, CIFAR-100, Street View House Numbers (SVHN), Caltech-UCSD Birds-200-2011 (CUB-200-2011), Pascal VOC2012, ADE20K, Cityscapes, and Common Objects in Context (COCO).
-The pretrained models and their accuracy metrics are summarized in the osmr github \charles{add link}
-
-We provide more details in Appendix~\ref{sxn:appendix}.
-
-In this paper, we propose that the Weightwatcher tool could be used to predict the trends in the generalization accuracy of deep neutral network without a test set.
-To test our proposition, we choose simple linear regression to analyze the relationship between the Weightwatcher metrics and the traditional accuracy metric obtained with a test set.
+We choose simple linear regression to analyze the relationship between quality metrics (computed with the \emph{WeightWatcher} tool) and traditional accuracy metrics (publicly-reported from a test set).
 % (we avoid polynomial regressions as they are more prone to overfitting and does not make economic sense).
 We regress the metrics on the Top1 (and Top5) reported errors (as dependent variables).  
-These include the Top5 errors for the ImageNet-1K mode, the percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and the Pixel accuracy (Pix.Acc.) and Intersection-Over-Union (mIOU) for XXX.
-We regress them individually on each of the Weightwatcher log Norm metrics, as described above.
+These include Top5 errors for the ImageNet-1K model, percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy (Pix.Acc.) and Intersection-Over-Union (IOU) for other models.
+We regress them individually on each of the norm-based and PL-based metrics, as described above.
+
+Our results are summarized in Table~\ref{table:results}.
+For the mean, larger $R^{2}$ and smaller $MSE$ are desirable; and for the standard deviation, smaller values are desirable.
+Taken as a whole, over the entire corpus of data, PL-based metrics are somewhat better for both the $R^{2}$ mean and standard deviation; and PL-based metrics are much better for $MSE$ mean and standard deviation.
+These (and other) results suggest our conclusions from Sections~\ref{sxn:cv} and~\ref{sxn:nlp} hold much more generally, and they suggest ovbious questions for future work.
 
@@
 \hline
-$R^{2}$ (mean) & 0.63 &0.55 &0.64 &0.64 \\
-$R^{2}$ (std)  & 0.34 &0.36 &0.29 &0.30 \\
+$R^{2}$ (mean) & 0.63 &  0.55 & \textbf{0.64} & \textbf{0.64} \\
+$R^{2}$ (std)  & 0.34 &  0.36 & \textbf{0.29} &          0.30 \\
 \hline
-$MSE$ (mean)   & 4.54 &9.62 &3.14 &2.92 \\
-$MSE$ (std)    & 8.69 &23.06 &5.14 &5.00 \\
+$MSE$ (mean)   & 4.54 &  9.62 &          3.14 & \textbf{2.92} \\
+$MSE$ (std)    & 8.69 & 23.06 &          5.14 & \textbf{5.00} \\
 \hline
@@
 \end{center}
-\caption{Comparison of linear regression fits for different average log norm metrics across 5 computer vision datasets, 17 Architectures, covering 168 (out of 309) different pretrained DNNs.
-We only conclude regressions for architecturs with 4 or more data points, and which are postively correlated with the test error.
-These results can be readily reproduced using the Google Colab notebooks (see Appendix)}
+\caption{Comparison of linear regression fits for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering 168 (out of 309) different pretrained DNNs.  
+         We include regressions only for architectures with 4 or more data points, and which are postively correlated with test error.
+         These results can be readily reproduced using the Google Colab notebooks (see Appendix~\ref{sxn:appendix}). 
+        }
 \label{table:results}

### 2020-02-10 submission/background.tex
full pass

@@
 There is very little work asking how to predict, in a theoretically-principled manner, the quality of large-scale state-of-the-art DNNs, and how to do so without access to training data or testing data or details of the training protocol, etc.
-Our work is, however, loosely related to several other lines of work.
+Our 
+approach  %%  work 
+is, however, 
+%% loosely 
+related to 
+two  %%  several 
+other lines of work.
 
@@
 Several very good overviews of the more common approach are available~\cite{EB01_BOOK, BKPx20}.
-We \emph{use} statistical mechanics in a broader sense, drawing upon results from quantitative finance and random matrix theory.  % (RMT).
+We \emph{use} statistical mechanics in a broader sense, drawing upon techniques from quantitative finance and random matrix theory.  % (RMT).
 Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statisical mechanics of heavy tailed and strongly correlated systems.
@@
 There is also a large body of work on using norm-based metrics to bound generalization error~\cite{NTS15, BFT17_TR, LMBx18_TR}.
-In this area, theoretical work aims to prove generalization bounds, and applied work uses these norms as a regularization function to improve training.
-While we do find that norms provide relatively good quality metrics, at least for distinguishing ``good-better-best'' among well-trained models, we are not interested in proving generalization bounds or developing new regularizers.
+In this area, theoretical work aims to prove generalization bounds, and applied work uses these norms to construct regularizers to improve training.
+While we do find that norms provide relatively good quality metrics, at least for distinguishing good-better-best among well-trained models, we are not interested in proving generalization bounds or developing new regularizers.
 
-\paragraph{Practical problems need a practical theory.}
-\nred{Exiting ML theory does not address many practical AI problems that our techiques directly target.}
-%There are many very practical problems in ML that are poorly addressed by existing theory and that either motivated our work or should be addressable by our techniques.
-Here are several.
-\begin{itemize}[leftmargin=*]
-\item
-\textbf{Simplicity, or lack of,  accuracy metrics.}
-\michael{XXX.  Put in two or at most three sentences.}
-\item
-\textbf{Information leakage in the production pipeline.}
-\michael{XXX.  Put in two or at most three sentences.}
-\item
-\textbf{Cost of acquring labeled data.}
-\michael{XXX.  Put in two or at most three sentences.}
-\end{itemize}
-
-\charles{see comments in tex file}\michael{Let's touch base, not sure what to do.}
-% we adress this above
-%Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text or realiztic chatbots, developing self driving cars, or just clusterting user profiles.
-%\footnote{For example, current chatbots use perplexity as a proxy for passing a Turing test.}
-%, and when distilling a reliable model in some way to obtain comparable training/test quality but that damages the model in some other subtle way. 
-%\michael{That last example is awkward.  It would be good to have a better example and plant seeds for model distillation elsewhere.}
+%% \paragraph{Practical problems need a practical theory.}
+%% \nred{Exiting ML theory does not address many practical AI problems that our techiques directly target.}
+%% %There are many very practical problems in ML that are poorly addressed by existing theory and that either motivated our work or should be addressable by our techniques.
+%% Here are several.
+%% \begin{itemize}[leftmargin=*]
+%% \item
+%% \textbf{Simplicity, or lack of,  accuracy metrics.}
+%% \michael{XXX.  Put in two or at most three sentences.}
+%% \item
+%% \textbf{Information leakage in the production pipeline.}
+%% \michael{XXX.  Put in two or at most three sentences.}
+%% \item
+%% \textbf{Cost of acquring labeled data.}
+%% \michael{XXX.  Put in two or at most three sentences.}
+%% \end{itemize}
+%% 
+%% \charles{see comments in tex file}\michael{Let's touch base, not sure what to do.}
+%% % we adress this above
+%% %Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text or realiztic chatbots, developing self driving cars, or just clusterting user profiles.
+%% %\footnote{For example, current chatbots use perplexity as a proxy for passing a Turing test.}
+%% %, and when distilling a reliable model in some way to obtain comparable training/test quality but that damages the model in some other subtle way. 
+%% %\michael{That last example is awkward.  It would be good to have a better example and plant seeds for model distillation elsewhere.}

### 2020-02-10 submission/conclusion.tex
full pass

@@
 We have developed (based on strong theory) and evaluated (on a large corpus of publicly-available pretrained models from CV and NLP) methods to predict trends in the quality of state-of-the-art neural networks---without access to training or testing data.
-Prior to our work, it was not obvious that norm-based metrics would perform well to predict trends \emph{across} models (as they are usually used \emph{within} a given model or parameterized model class, e.g., to bound generalization error or to construct regularizers), and our results are the first to demonstrate that they can be used for this important practical problem.
-That PL-based metrics perform better (than norm-based metrics) should not be surprising---at least to those familiar with the statisical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}, since our use of PL exponents is designed to capture the idea that well-trained models capture correlations over many size scales in the data), but again,  our results are the first to demonstrate this.
-It is gratifying, also, that PL exponents can be used to provide fine-scale insight such as rationalizing the flow of correlations through a network.
+Prior to our work, it was not obvious that norm-based metrics would perform well to predict trends in quality \emph{across} models (as they are usually used \emph{within} a given model or parameterized model class, e.g., to bound generalization error or to construct regularizers).
+Our results are the first to demonstrate that they can be used for this important practical problem.
+That PL-based metrics perform better (than norm-based metrics) should not be surprising---at least to those familiar with the statisical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} (since our use of PL exponents is designed to capture the idea that well-trained models capture correlations over many size scales in the data).
+Again, though, our results are the first to demonstrate this.
+It is also gratifying that this approach can be used to provide fine-scale insight (such as rationalizing the flow of correlations or the collapse of size scale) throughout a network.
 
+\ncyan{
 We conclude with a few thoughts on what a \emph{practical theory} of DNNs should look like.
+\michael{MM TO DO.}
 %
@@
 Our results, which are based on our \emph{use} of sophisticated statistical mechanics theory to solve an important practical DNN problems, suggests that this approach should be of interest more generally for those interested in developing a practical DNN theory.
-\charles{see latex file}
 %\charles{it is the oppposite...ab initio theory scales quite well...it is the spherical cow models of physics and ML that do not scale.
@@
 What we have introduced is a phenomenology, which to me is different from a semi-empirical or phenomenological theory  }
-\michael{Let's touch base to get this right.}
+}

### 2020-02-10 submission/cv_models.tex
full pass

--- 2020-02-08 cv_models.tex
@@
 \paragraph{Average Quality Metrics versus Reported Test Accuracies.}
+
 We have examined the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
-To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without Batch Normalization), and it plots the four quality metrics versus the reported Test accuracies~\cite{pyTorchVgg}, as well as a basic linear regression line.
+To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BN), and it plots the four quality metrics versus the reported test accuracies~\cite{pyTorchVgg},%
+\footnote{That is, these test accuracies have been previously reported and made publicly-avaiable by others.
+We take them as given, and we do not attempt to reproduce/verify them, since we do not permit ourselves any access to training/test data.}
+as well as a basic linear regression line. 
 %with the Root Mean Squared Error (RMSE) shown, \michael{XXX.  WHERE IS RMSE.  IN THAT TABLE?}
-All four metrics correlate quite well with the reported Top1 Accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalziation (i.e., greater accuracy, lower error). 
+All four metrics correlate quite well with the reported Top1 accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalziation (i.e., greater accuracy, lower error). 
 While all four metrics perform well, notice that the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) performs best (with an RMSE of $0.42$, see Table~\ref{table:cv-models}); and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $), which is an approximation to the Log $\alpha$-Norm metric~\cite{MM20_unpub_work}, performs second best (with an RMSE of $0.48$, see Table~\ref{table:cv-models}).
@@
     \centering
-    \subfigure[ Frobenius Norm, VGG ]{
+    \subfigure[Log Frobenius Norm, VGG ]{
         %\includegraphics[width=5cm]{img/VGG_lognorm_accs.png}
@@
     %\qquad
-    \subfigure[ Spectral Norm, VGG ]{
+    \subfigure[Log Spectral Norm, VGG ]{
         %\includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
@@
     %\qquad
-    \subfigure[ $\alpha$-Norm, VGG ]{
+    \subfigure[Log $\alpha$-Norm, VGG ]{
         %\includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
@@
     }
-    \caption{Comparison of average log Norm empirical quality metrics versus reported test accuracy for pretrained VGG models (with and without Batch Normalization (BN)), trained on ImageNet, available in pyTorch (v1.x).
+    \caption{Comparison of Average Log Norm and Weighted Alpha quality metrics versus reported test accuracy for pretrained VGG models (with and without BN), trained on ImageNet, available in pyTorch (v1.x).
 Metrics fit by linear regression, RMSE reported.
@@
 
-    \subfigure[ ResNet, $\alpha$-Norm ]{
+    \subfigure[ ResNet, Log $\alpha$-Norm ]{
         %\includegraphics[width=4.2cm]{img/ResNet_logpnorm_accs.png}
@@
     %\qquad
-    \subfigure[ ResNet-1K, $\alpha$-Norm ]{
+    \subfigure[ ResNet-1K, Log $\alpha$-Norm ]{
         %\includegraphics[width=4.5cm]{img/ResNet-1K_logpnorm_accs.png}
@@
     %\qquad
-    %\subfigure[ DenseNet, $\alpha$-Norm ]{
+    %\subfigure[ DenseNet, Log $\alpha$-Norm ]{
     %    %\includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
@@
     %}
-    \caption{Comparison of the avergage $\alpha$-Norm empirical quality metric $\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle$ 
-versus reported Top 1 reported Test Accuracy for the ResNet and ResNet-1K pretrained (pyTorch) models. }
+    \caption{Comparison of Avergage $\alpha$-Norm quality metric ($\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle$) versus reported Top1 test accuracy for the ResNet and ResNet-1K pretrained (pyTorch) models.
+}
 %, and DenseNet models. }The corresponding results for VGG are shown in Figure~\ref{fig:vgg-pnorm}.  }
@@
 %\begin{tabular}{|p{1in}|c|c|c|c|c|}
-\begin{tabular}{|p{0.75in}|c|c|c|c|c|}
+\begin{tabular}{|p{0.65in}|c|c|c|c|c|}
 %\begin{tabular}{|l|c|c|c|c|c|}
 \hline
- Series    &\#   & $\log\langle\Vert\mathbf{W}\Vert_{F}\rangle$ & $\log\langle\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\hat{\alpha}$ & $\log\langle\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
+ Series    &\#   & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\hat{\alpha}$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
 \hline
@@
 \end{center}
-\caption{RMSE (smaller is better) for linear fits of quality metrics to Reported Top1 Test Error for pretrained models in each architecture series.
+\caption{RMSE (smaller is better) for linear fits of 
+         %Average Log Norm and Weighted Alpha 
+         quality metrics to reported Top1 test error for pretrained models in each architecture series.
+Column \# refers to number of models.
 VGG, ResNet, and DenseNet were pretrained on ImageNet, and ResNet-1K was pretrained on ImageNet-1K.
@@
 
-See Table~\ref{table:cv-models} for a summary of results for Top1 Accuracies for all four metrics on for the VGG, ResNet, and DenseNet series.
-Similar results (not shown) are obtained for the Top5 Accuracies.
-Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and th\
-e Weighted Alpha metric performs best.
-These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among ``good-better-bes\
-t'' models, with PL-based metrics performing somewhat better.
+See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics on for the VGG, ResNet, and DenseNet series.
+Similar results (not shown) are obtained for the Top5 accuracies.
+Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best.
+These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among ``good-better-best'' models, with PL-based metrics performing somewhat (i.e., quantitatively) better.
 
-Notice that the DenseNet series,  only has 4 data points.  In our larger analysis, in Section~\ref{sxn:all} we only include series with 5 or more models.
-For completemnss and reproducibility here, the DenseNet Log $\alpha$-Norm plot is given in the Appendix.
+The DenseNet series has similar behavior to what we see in Figures~\ref{fig:vgg-metrics} and~\ref{fig:cv2-accuracy} for the other models.
+However, as noted in Table~\ref{table:cv-models}, it has only 4 data points.  
+In our larger analysis, in Section~\ref{sxn:all_cv_models}, we will only include series with 5 or more models.
+%For completemnss and reproducibility here, the DenseNet Log $\alpha$-Norm plot is given in Appendix~\ref{sxn:appendix}.
+(Note that these and many other such plots can be seen on our publicly-available~repo.)
 
-\paragraph{Variations in Data Set Size}
-We want to point out, in particular, how the empirical log Norm metrics vary with data set size.
-Figure~\ref{fig:cv2-accuracy} plots and compares the 
-Log $\alpha$-Norm
-metric--the best performing metric--for the full ResNet, trained on the full ImageNet dataset, against the ResNet-1K, which has been trained on a much smaller ImageNet-1K data set.
-%(A much more detailed set of plots, including the behavior of all four methods on each of the series, is available at \cite{XXX-WEB-LINK}.)
-Here, the Log $\alpha$-Norm is much better that the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is actually slightly worse than the Weighted Alpha metric).
-The ResNet series, has an RMSE of $0.66$, whereas the ResNet-1K series also shows good correlation, but has a much larger RMSE of $1.9$.
-As expected, the higher quality data set shows a better fit, even with less data points.
+\paragraph{Variation in Data Set Size.}
 
-\paragraph{Correlation Flow}
-We can learn much more about a pretrained model by going beyond average values to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).
+We are interested in how our four quality metrics depend on data set size.
+To examine this, we look at results on ResNet versus ResNet-1K.
+See Figure~\ref{fig:cv2-accuracy}, which plots and compares the Log $\alpha$-Norm metric 
+%% MM: From Table 1 it is slightly worse than alpha-hat and we say that below %% --the best performing metric--
+for the full ResNet model, trained on the full ImageNet dataset, against the ResNet-1K model, which has been trained on a much smaller ImageNet-1K data set.
+The Log $\alpha$-Norm is much better than the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is actually slightly worse than the Weighted Alpha metric).
+The ResNet series has strong correlation, with an RMSE of $0.66$, whereas the ResNet-1K series also shows good correlation, but has a much larger RMSE of $1.9$.
+(Other metrics exhibit similar behavior.)
+As expected, the higher quality data set shows a better fit, even with fewer data points.
+
+\paragraph{Layer Analysis: Metrics as a Function of Depth.}
+
+We can learn much more about a pretrained model by going beyond average values of quality metrics to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).
 % in the network.
@@
 For example, we can 
-plot (just) the PL exponent, $\alpha$,%
-\footnote{That is, here we consider just $\alpha$ for each layer, i.e., not $\hat{\alpha}$ or $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$.} 
+plot (just) the PL exponent, $\alpha$, for each layer, %
+%\footnote{That is, here we consider just $\alpha$ for each layer, i.e., not $\hat{\alpha}$ or $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$.} 
 as a function of depth.
@@
     %\qquad
-    \subfigure[ ResNet ]{
+    \subfigure[ ResNet-1K ]{
         %\includegraphics[width=4.5cm]{img/ResNet_fnl_alpha_depth.png} 
@@
 
-}    \subfigure[ ResNet (overlaid) ]{
+}    \subfigure[ ResNet-1K (overlaid) ]{
         \includegraphics[width=4.0cm]{img/resnet_alpha_overlaid_depth.png} 
@@
     }
-    \caption{Correlation Flow: Power Law exponent $\alpha$ versus layer id, for the least and the most accurate
-      models in VGG (a), ResNet (b), and DenseNet (c) series (VGG without BatchNorm, and the Y axes on each plot are different.)  
-             Figure (d) displays the ResNet models (b), zoomed in to $alpha\in[1,5]$, and with the layer ids overlaid on the X-axis, to
+    \caption{PL exponent ($\alpha$) versus layer id, for the least and the most accurate models in VGG (a), ResNet (b), and DenseNet (c) series. 
+             (VGG is without BN; and note that the Y axes on each plot are different.)  
+             Subfigure (d) displays the ResNet models (b), zoomed in to $\alpha\in[1,5]$, and with the layer ids overlaid on the X-axis, from smallest to largest, to
              allow a more detailed analysis for the most strongly correlated layers.
-             Notice that ResNet152 exhibits different and much more stable behavior across layers
+             Notice that ResNet152 exhibits different and much more stable behavior of $\alpha$ across layers.
              This contrasts with how both VGG models gradually worsen in deeper layers and how the DenseNet models are much more erratic.  
+             In the text, this is interpreted in terms of \emph{Correlation Flow}.
             }
@@
 
-For the ResNet-1K models, Figure~\ref{fig:resnet-alpha-layer} shows that $\alpha$ also increases in the last few layers (more dramatically, in fact, observe the differing scales on the Y axes).
-However, as the ResNet-1K models get deeper, there is a wide range over which $\alpha$ values tend to remain small.
+For the ResNet-1K models, Figure~\ref{fig:resnet-alpha-layer} shows that $\alpha$ also increases in the last few layers (more dramatically, in fact, than for VGG, observe the differing scales on the Y axes).
+However, as the ResNet-1K models get deeper, there is a wide range over which $\alpha$ values tend to remain quite small.
 %\michael{Can we say something about exceptions which are larger than for VGG or DenseNet.}
@@
 For the DenseNet models, Figure~\ref{fig:densenet-alpha-layer} shows that $\alpha$ tends to increase as the layer id increases, in particular for layers toward the end.
-While this is similar to what is seen in the VGG models, with the DenseNet models, $\alpha$ values increase almost immediately after the first few layers, and the variance is much larger (in particular for the earlier and middle layers, where it can range all the way to $\alpha\lesssim 8.0$) and much less systematic throughout the network.
+While this is similar to what is seen in the VGG models, with the DenseNet models, $\alpha$ values increase almost immediately after the first few layers, and the variance is much larger (in particular for the earlier and middle layers, where it can range all the way to $\alpha\sim 8.0$) and much less systematic throughout the network.
 
@@
 %The empirical manifestations of this (on weight matrices) are that fitted $\alpha$ values get much worse for deeper layers.
-\ncyan{We can interpret these observations by recalling the architectural differences between the VGG, ResNet, and DenseNet architectures,
+%%\ncyan{
+We can interpret these observations by recalling the architectural differences between the VGG, ResNet, and DenseNet architectures,
 and, in particular, the number of of residual connections. VGG resembles the  traditional convolutional architectures,
@@
 This optimized approach allows for greater accurary with far fewer parameters (and GPU memory requirements), and
-ResNet models of up to 1000 layers have been trained.\cite{resnet1000}.}
+ResNet models of up to 1000 layers have been trained.\cite{resnet1000}.
+%%}
 We conjuecture that the effeciency and effectivness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and strongly optimized.
@@
 
-More generally, we can understand Figure~\ref{fig:vgg-alpha-layers} in terms of the \emph{Correlation Flow}. 
-Recall that the log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on ideas from the statisical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
-There, one expects the weight matrices of well-trained DNNs will 
-exhibit correlations over many size scales. Their ESDs can be well-fit by a (truncated) Power law (PL), with exponents $\alpha\in[2,4]$.
-Much larger values $(\alpha\gg 5\;or\;6)$ may relfect poorer PL fits, whereas smaller values $(\alpha\rightarrow 2)$, are associated with models
-that generalize better.
+\paragraph{Correlation Flow.}
+
+More generally, we can understand the results presented in Figure~\ref{fig:3models-alpha-layers} in terms of what we will call the \emph{Correlation Flow} of the model. 
+Recall that the average Log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on ideas from the statisical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
+There, one expects the weight matrices of well-trained DNNs will exhibit correlations over many size scales. 
+Their ESDs can be well-fit by a (truncated) PL, with exponents $\alpha\in[2,4]$.
+Much larger values $(\alpha\gg 5\;\mbox{or}\;6)$ may reflect poorer PL fits, whereas smaller values $(\alpha\sim 2)$, are associated with models that generalize better.
 %, but the presence of many adjacent layers with relatively good PL fits \emph{suggests} that those adjacent layers ``impedance match'' well and that the correlations that are captured by one layer propagate well to subsequent layers.
 Informally, one would expect a DNN model to perform well when it fascilitates the propagation of information/features across layers.
-Previous studies argue this by computing the gradients over the input data.
-In the absence of training/test data, we can to try to quantify this by measuring the PL properties of empirical weight matrices.
-Smaller $\alpha$ values correspond to layers in which correlations across multiple scales are better captured~\cite{MM18_TR,SornetteBook}, and thus we expect that small values of $\alpha$ that are stable across multiple layers enable better \emph{correlation flow} through the network.
-
-%\paragraph{Layer Analysis: Behavior on Less Thoroughy Studied CV Models.}
-\paragraph{Correlation Flow: PL vs. Norm Metrics.}
-
-The similarity between norm-based metrics and $\alpha$-based metrics suggests a question: is the Weighted Alpha metric just a variation of the more familiar norm-based metrics, or does it capture something different?
-More generally, do fitted $\alpha$ values contain information not captured by norms? 
-To show that it is not just a variation and that $\alpha$ does capture something different, consider the following example, which looks at a compressed / distilled DNN model~\cite{CWZZ17_TR}.
+Previous work argues this by computing the gradients over the input data.
+In the absence of training/test data, one might hope that this leaves empirical signatures on weight matrices, and thus we can to try to quantify this by measuring the PL properties of weight matrices.
+In this case, smaller $\alpha$ values correspond to layers in which correlations across multiple scales are better captured~\cite{MM18_TR,SornetteBook}, and we expect that small $\alpha$ values that are stable across multiple layers enable better \emph{correlation flow} through the network.
+We have seen this in many models, including those shown in Figure~\ref{fig:3models-alpha-layers}.
 
@@
    }
-   \caption{Correlation Flow.
-            ResNet20, distilled with Group Regularization, as implemented in the \texttt{distiller} (4D\_regularized\_5Lremoved) pre-trained models.  
-            Spectral Norm, $\log\lambda_{max}$, and, PL exponent, $\alpha$, for individual layer, versus layer id, for both baseline (before distillation, green) and finetuned (after distillation, red) pre-trained models.
+   \caption{%Correlation Flow.
+            ResNet20, distilled with Group Regularization, as implemented in the \texttt{distiller} (4D\_regularized\_5Lremoved) pretrained models.  
+            Log Spectral Norm ($\log\lambda_{max}$) and PL exponent ($\alpha$) for individual layers, versus layer id, for both baseline (before distillation, green) and fine-tuned (after distillation, red) pretrained models.
            }
@@
 
-\paragraph{Scale Collapse; or, Distillation may break models}
+%% %\paragraph{Layer Analysis: Behavior on Less Thoroughy Studied CV Models.}
+%% \paragraph{Correlation Flow: PL vs. Norm Metrics.}
+\paragraph{Scale Collapse; or How Distillation May Break Models.}
 
+The similarity between norm-based metrics and PL-based metrics suggests a question: is the Weighted Alpha metric just a variation of the more familiar norm-based metrics?
+%%, or does it capture something different?
+More generally, do fitted $\alpha$ values contain information not captured by norms? 
 In examining hundreds of pretrained models, we have found several anamolies that demonsrate the power of our approach.
-Here, we demonstrate that some distillations methods may actually \emph{break} models unexpectedly by introducing what we call \emph{Scale Collapse},
-where several distilled layers have unexpectedly small Spectral Norms.
+In particular, to show that $\alpha$ does capture something different, consider the following example, which looks at a compressed/distilled DNN model~\cite{CWZZ17_TR}.
+In this example, we show that some distillation methods may actually \emph{break} models unexpectedly by introducing what we call \emph{Scale Collapse}, where several distilled layers have unexpectedly small Spectral Norms.
+
 We consider ResNet20, trained on CIFAR10, before and after applying the Group Regularization distillation technique, as implemented in the \texttt{distiller} package.%
 \footnote{For details, see \url{https://nervanasystems.github.io/distiller/\#distiller-documentation} and also \url{https://github.com/NervanaSystems/distiller}.}
-We analyze the pre-trained 4D\_regularized\_5Lremoved baseline and finetuned models.  %\cite{distiller_repo}.
-The reported baseline test accuracies ($Top1=91.45$ and $Top5=99.75$) are better than the reported finetuned test accuracies ($Top1=91.02$ and $Top5=99.67$)~\cite{XXX-XXX}.
-Because the baseline accuracy is greater,  the previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy}) suggests that the baseline Spectral Norms should be \emph{smaller} on average than the finetuned ones should be smaller.
+We analyze the pretrained 4D\_regularized\_5Lremoved baseline and fine-tuned models.  %\cite{distiller_repo}.
+The reported baseline test accuracies (Top1$=91.45$ and Top5$=99.75$) are better than the reported fine-tuned test accuracies (Top1$=91.02$ and Top5$=99.67$)~\cite{XXX-XXX}.
+Because the baseline accuracy is greater,  the previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy}) suggests that the baseline Spectral Norms should be \emph{smaller} on average than the fine-tuned ones should be smaller.
+\michael{IS THERE A TYPO, THAT SOUNDS LIKE BOTH ARE SMALLER.}
 \emph{The oppopsite is observed.}
@@
 %In both cases (Frobenius norm results not shown), we observe the opposite.
-Figure~\ref{fig:resnet204D5L} presents the Spectral Norm (here denoteds $\log\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}$.%
+Figure~\ref{fig:resnet204D5L} presents the Spectral Norm (here denoted $\log\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}$.%
 \footnote{Here, we only include layer matrices or feature maps with $M\ge50$.}
-On the other hand, the $\alpha$ values do not differ systematically between the baseline and finetuned models.
+On the other hand, the $\alpha$ values (in Figure~\ref{fig:resnet204Dalpha}) do not differ systematically between the baseline and fine-tuned models.
 %\michael{Some comment about big jump in norm; but fix previous backwards question first.}--SEE BELOW
-Also (not shown), the average (unweighted) baseline $\alpha$ is \emph{smaller} than the finetuned average (as predicted by HT-SR Theory, the basis of $\hat{\alpha}$).
+Also (not shown), the average (unweighted) baseline $\alpha$ is \emph{smaller} than the fine-tuned average (as predicted by HT-SR Theory, the basis of $\hat{\alpha}$).
 
-Note that Figure ~\ref{fig:resnet204Dalph} depcits 2 very large $\alpha\gg 6$ for the baseline model, but not for the finetuned model.
-This suggests that the baseline model has at least 2 over-parameterized layers, and the distillation method does, in fact, 
-improve the finetuned model by compressing these layers.s
+(That being said, Figure~\ref{fig:resnet204Dalpha} also depcits two very large $\alpha\gg 6$ values for the baseline, but not for the fine-tuned, model.
+This suggests the baseline model has at least two over-parameterized/under-trained layers, and that the distillation method does, in fact, improve the fine-tuned model by compressing these layers.)
 
-\michael{XXX.  WORK ON THIS PAR AFTER DO NLP EXAMPLES.}
-Our interpretation of this is the following.
-The pre-trained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the 
-VGG, ResNet, or DenseNet series we cosidered above.
-\charles{MOvE THIS SOMEWHERE ELSE AND CLEAN UP THIS SECTION Notice while norms make good regularizers for a single model, there is no reason \emph{a priori} 
-to expect them correlate so well with the test accuracies across different models.
-However, we do expect the PL $\alpha$ to do so because it effectively measures the amount of correlation in the model}
-%
-The reason for this is that the \texttt{distiller} Group Regularization technique 
+The pretrained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the 
+VGG, ResNet, or DenseNet series.  %% we considered above.
+While norms make good regularizers for a single model, there is no reason \emph{a priori} to expect them correlate so well with test accuracies across different models.
+We do expect, however, the PL $\alpha$ to do so because it effectively measures the amount of correlation in the model~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+The reason for the anomalous behavior shown in 
+Figure~\ref{fig:resnet204D5L}
+is that the \texttt{distiller} Group Regularization technique 
 %%has the unusual effect of increasing
-\nred{spuriously increases}
- the norms of the $\mathbf{W}$ pre-activation maps for at least two of the Conv2D layers.
-\michael{This impedance shuffling or whatever it is called messes up norms, but the strong correlations are still preserved, as seen in the $\hat{\alpha}$ metric, and the model quality is good.
-CLARIFY.}
-\michael{XXX.  INTERPRET IN TERMS OF XXX GOOD VERSUS BAD, AS OPPOSED TO GOOD BETTER BEST.}
+spuriously increases the norms of the $\mathbf{W}$ pre-activation maps for at least two of the Conv2D layers.
+This is difficult to diagnose by analyzing training/test curves, but it is easy to diagnose with our approach.

### 2020-02-10 submission/introduction.tex
Update introduction.tex
Corrected a typo in line 80
full pass

@@
 There are many problems with this approach.
-The training/testing curves give very limited insight into the overall properties of the model, 
-they do not take into account the (often large human and CPU/GPU) time for hyperparameter fiddling,
-they typically do not correlate with other properties of interest such as robustness or fairness or interpretability, 
+The training/testing curves give very limited insight into the overall properties of the model; 
+they do not take into account the (often large human and CPU/GPU) time for hyperparameter fiddling;
+they typically do not correlate with other properties of interest such as robustness or fairness or interpretability; 
 and so on.
 A less well-known problem, but one that is increasingly important, in particular in industrial-scale artificial intelligence (AI), arises when the model \emph{user} is not the model \emph{developer}.
-Here, one may not have access to either the training data nor the testing data.
-Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it "as is'', or to fine-tune and/or compress it and then use it.
+Here, one may not have access to either the training data or the testing data.
+Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it "as is,'' or to fine-tune and/or compress it and then use it.
 
@@
 Moreover, if ML is to become an industrial process, then that process will become siloed: some groups will gather data, other groups will develop models, and other groups will use those models.
-Users of models cannot be expected to know the precise details of how models were built, the specifics of data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized,~etc.
+Users of models can not be expected to know the precise details of how models were built, the specifics of data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized,~etc.
 
@@
 In the end, however, they really require human evaluation. 
-Alternatively, models that cluster user profiles, which are widely used in marketing, are unsupervised and have no obvious labels for comparison and/or evaluation.
+Alternatively, models that cluster user profiles, which are widely used in areas such as marketing and advertising, are unsupervised and have no obvious labels for comparison and/or evaluation.
+In these and other areas, ML objectives can be poor proxies for downstream goals.
 
@@
 
-In this paper, we present and evaluate quality metrics for pretrained deep neural network (DNN) models, and at scale.
+In this paper, we present and evaluate quality metrics for pretrained deep neural network (DNN) models, and we do so at scale.
 We consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
@@
 They are based on norms of weight matrices (such norms have been used in traditional statistical learning theory to bound capacity and construct regularizers) and/or parameters of power law (PL) fits of the eigenvalues of weight matrices (such PL fits are based on statistical mechanics approaches to DNNs).
-Note that, while we use traditional norm-based and PL-based metrics, our goals are different.
-Unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., hyperparmeter optimization).%
+Note that, while we use traditional norm-based and PL-based metrics, our goals are not the traditional goals.
+Unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., as with hyperparmeter optimization).%
 \footnote{One could of course use these techniques to improve training, and we have been asked about that, but we are not interested in that here.
 Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained DNN models.}
-Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves, and in both cases, we ask
-
-\emph{Can we predict trends in the quality of pre-trained models without access to training or testing data ?}.
+Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves, and in both cases, we ask:
+\begin{quote}
+\emph{Can we predict trends in the quality of pretrained DNN models without access to training or testing data?}  
+\end{quote}
 
@@
 %To answer this question, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in CV and NLP.
-To answer this, we analyze hundreds of publicly-available pretrained state-of-the-art CV and NLP models. And we find that:
 %\end{itemize}
-%Third, 
-%\paragraph{Our main contributions are:}
+
+To answer this question, we analyze hundreds of publicly-available pretrained state-of-the-art CV and NLP models. 
+Here is a summary of our main results.
 \begin{itemize}
 \item
-Norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best'' models, but they can qualitatively fail for models that may not be well-trained, i.e., to distinguish ``good-versus-bad'' models.
+Norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models, i.e., they can be used to discriminate between ``good-better-best'' models.
+However, they can qualitatively fail for models that may not be well-trained, i.e., to distinguish ``good-versus-bad''~models.
 \item 
-%Fourth, 
-Power Law (PL)-based metrics do much better---quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-versus-bad models.
+PL-based metrics do much better at predicting quality trends in CV/NLP models---quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-versus-bad models.
 \item 
-%Finally, 
-PL-based metrics can also be used to characterize fine-scale model properties, including layer-wise \emph{correlation flow}, and evaluating post-training modifications such as model distillation, increasing the dataset size, and other model improvements.
+PL-based metrics can also be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow} and \emph{Scale Collapse} in poorly-trained models) and evaluate post-training modifications (such as model distillation, increasing the dataset size, and other model improvements).
 \end{itemize}
@@
 We emphasize that our goal is a practical theory to predict trends in the quality of state-of-the-art DNN models, i.e., not to make a statement about every publicly-available model.
-We have examined hundreds of models, and we indentify general trends, but we also highlight interesting exceptions.
+We have examined hundreds of models, and we identify general trends, but we also highlight interesting exceptions.
 %%Several of the most interesting are described below.
@@
 
-All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package};
-and, to be fully reproducible, we only examine publically available, pretrained models, and
-also provide all Jupyter and Google Colab notebooks used accompanying github 
-repository\footnote{\url{https://github.com/CalculatedContent/kdd2020}
-\michael{To be anonymized}}
-See the Appendix for more details on how to reproduce all results.
+All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
+To be fully reproducible, we only examine publically-available, pretrained models, and we also provide all Jupyter and Google Colab notebooks used in an accompanying github repository.%
+\footnote{\url{https://github.com/CalculatedContent/kdd2020} \michael{TO BE ANONYMIZED.}}
+See Appendix~\ref{sxn:appendix} for details on how to reproduce all results.
 
@@
 Then, in Section~\ref{sxn:nlp}, we look at several variations of a popular NLP DNN architecture (the OpenAI GPT and GPT2 models); and we show how model quality and properties vary between several variants of GPT and GPT2, including how metrics behave similarly and differently.
-Then, in Section \ref{sxn:all_cv_models}, we present results based on an analysis of hundreds of pretraind DNN CV models, showing how well each metric predicts the reported test accuracies, and how the Alpha-Norm metric(s) perform remarkably well.
+Then, in Section \ref{sxn:all_cv_models}, we present results based on an analysis of hundreds of pretraind DNN models, showing how well each metric predicts the reported test accuracies, and how the PL-based metrics perform remarkably well.
 Finally, in Section~\ref{sxn:conc}, we provide a brief discussion and conclusion.

### 2020-02-10 submission/methods.tex
full pass

--- 2020-02-09 submission/methods.tex
@@
 
-Let us 
-write the Energy Landscape (or optimization function, parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$,~as:
+Let us write the Energy Landscape (or optimization function, parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$,~as:
 \begin{equation}
@@
 \end{equation}
-
-Each DNN layer contains one or more layer 2D  $N\times M$ weight matrices, $\mathbf{W}_{l}$ or pre-activation maps, $\mathbf{W}_{i,l}$, extracted from 2D Convolutional layers, and where $N > M$.
-\footnote{We do not use intra-layer information from the models in our quality metrics, but as we will describe our metrics can be used to learn about that.}
-Notice we may drop the $i$ and/or $i,l$ subscripts below.
-See the Appendix for how we define the Conv2D layer matrixes, and our choices of normalization.
+Each DNN layer contains one or more layer 2D  $N\times M$ weight matrices, $\mathbf{W}_{l}$, or pre-activation maps, $\mathbf{W}_{i,l}$, extracted from 2D Convolutional layers, and where $N > M$.% 
+\footnote{We do not use intra-layer information from the models in our quality metrics, but (as we will describe) our metrics can be used to learn about intra-layer model properties.}
+(We may drop the $i$ and/or $i,l$ subscripts below.)
+See Appendix~\ref{sxn:appendix} for how we define the Conv2D layer matrixes and for our choices of normalization.
 
 Assume we are given several pretrained DNNs, e.g., as part of an architecture series.
-The models have been trained and evaluated on labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$,
-using standard techniques.  The pretained pytroch model files are publically available,
-and the test accuracies have been reported online.  
+The models have been trained and evaluated on labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using standard techniques.  
+The pretained pytroch model files are publicly-available, and the test accuracies have been reported online.  
 In this study, we do not have access to this data, and we have not trained any of the models ourselves,
 nor have we re-evaluated the test accuracies.
-We expect that most well-trained, production-quality models will employ one or more forms of on regularization, such as Batch Normalization, Dropout, etc., and many will also contain additional structure such as Skip Connections, etc.
+We expect that most well-trained, production-quality models will employ one or more forms of on regularization, such as Batch Normalization (BN), Dropout, etc., and many will also contain additional structure such as Skip Connections, etc.
 Here, we will ignore these details, and will focus only on the pretrained layer weight matrices $\mathbf{W}_{l}$.
@@
 
-Here, we consider the following metrics:.
+Here, we consider the following metrics.
 
@@
 \item 
-$\alpha$-Norm (or $\alpha$-Shatten Norm):
-\footnote{Notice  $\Vert\mathbf{W}\Vert^{2\alpha}_{2\alpha}=\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$. We use $\mathbf{X}$ to emphasize that $\alpha$ depends on the ESD of $\mathbf{X}$}
+$\alpha$-Norm (or $\alpha$-Shatten Norm):%
+\footnote{Notice $\Vert\mathbf{W}\Vert^{2\alpha}_{2\alpha}=\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$. We use $\mathbf{X}$ to emphasize that $\alpha$ depends on the ESD of $\mathbf{X}$.}
  $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda_{i}^{\alpha}$
 \end{itemize}
-
-The  $\lambda_{i}$ is the $i^{th}$ eigenvalue of the $\mathbf{X}$, 
-and $\lambda_{max}$ is the maximum eigenvalue.
-Note the squared singular values $\sigma_{i}$ of $\mathbf{W}$, $\lambda_{i}=\sigma^{2}_{i}$.)
-Also, note we do \emph{not} normalize $\mathbf{X}$ by $1/N$; see the Appendix for a discussion why.
+Here, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the $\mathbf{X}$, and $\lambda_{max}$ is the maximum eigenvalue.
+Recall that the eigenvalues are square of the singular values $\sigma_{i}$ of $\mathbf{W}$: $\lambda_{i}=\sigma^{2}_{i}$.
+Also, note that we do \emph{not} normalize $\mathbf{X}$ by $1/N$; see Appendix~\ref{sxn:appendix} for a discussion of this issue.
 
@@
 where $\Vert\cdot\Vert$ is a matrix norm.   
-The log Complexity,
+The log complexity,
 \begin{equation}
@@
 \end{equation}
- takes the form of an Average Log Norm.
-For the \emph{Frobenius norm metric} and \emph{Spectral norm metric}, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly. 
+ takes the form of an average Log Norm.
+For the \emph{Frobenius norm metric} and \emph{Spectral norm metric}, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly.% 
 \footnote{When taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.}
-For the $\alpha$-Shatten Norm metric, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it cannot be taken out of the sum.
+For the $\alpha$-Shatten Norm metric, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it can not be taken out of the sum.
 
-\emph{The Weighted Alpha metric} is an average of $\hat{\alpha}$ over all layers, weighted by the size, or scale, or each matrix
-(and approximates the average log $\alpha$-Norm metric).
+The \emph{Weighted Alpha metric} is an average of $\alpha_l$ over all layers $l \in \{1,\ldots,l\}$, weighted by the size, or scale, or each matrix
+(and it approximates the average log $\alpha$-Shatten Norm metric),
 \begin{equation}
-\hat{\alpha} = \dfrac{1}{L}\sum_l \alpha_l\log\lambda_{max,l}\approx\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle  
+\hat{\alpha} = \dfrac{1}{L}\sum_l \alpha_l\log\lambda_{max,l}\approx\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle    ,
 \end{equation}
@@
 The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a limited set of models.
-Based on this, in this paper, we introduce and evaluate the $\alpha$-Norm metric.
-One expects$\hat{\alpha}$ approximates the log $\alpha$-Norm very well for $\alpha < 2$ and reasoably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.
+Based on this, in this paper, we introduce and evaluate the $\alpha$-Shatten Norm metric.
+One expects $\hat{\alpha}$ approximates the average log $\alpha$-Shatten Norm very well for $\alpha < 2$ and reasonably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.
 
@@
 We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated PL, parameterized by 2 values: the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.  
-%(Technically, we also need the minimum eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.)
+(Technically, we also need the minimum eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.)
 The PL exponent $\alpha$ measures of the amount of correlation in a DNN layer weight matrix $\mathbf{W}$. 
@@
 %
-Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric for the entire~DNN.  
-%%%
-%%For small values of $\alpha$, this Weighted Alpha metric approximates the log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}.
-%%%
-%%The Weighted Alpha metric and $\alpha$-Shatten norm metric also often behave like an improved, weighted log average Spectral Norm, and may track this metric in some cases.
+Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric for the entire~DNN.%
+\footnote{For small $\alpha$, this Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}; and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm, and may track this metric in some cases.}
 
@@
 There are several technical issues (regarding spectral analysis of convolutional layers and normalization of empirical matrices) that are important for reproducibility of our results.
-See Appendix \ref{sxn:appendix} for a discussion.
+See Appendix~\ref{sxn:appendix} for a discussion.

### 2020-02-10 submission/nlp_models.tex
full pass

--- 2020-02-08 nlp_models.tex
@@
-section{Comparison of NLP Models}
+\section{Comparison of NLP Models}
 \label{sxn:nlp}
 
-In this section, we examine quality metrics described in Section~\ref{sxn:methods} for several NLP model architectures.
+In this section, we examine empirical quality metrics described in Section~\ref{sxn:methods} for several NLP model architectures.
 %
 %In particular, 
-Within the past two years, nearly 100 open source, pre-trained NLP DNNs based on the revolutionary Transformer architecture have emerged.
+Within the past two years, nearly 100 open source, pretrained NLP DNNs based on the revolutionary Transformer architecture have emerged.
 These include variants of BERT, Transformer-XML, GPT, etc.
@@
 In contrast to smaller pre-Activation maps arising in Cond2D layers, Attention matrices are significantly larger.
-In general, we have found that they have larger $\alpha$ PL exponents.
+In general, we have found that they have larger PL exponents $\alpha$.
 Based on HT-SR Theory, 
@@
 More generally, compared to the CV models of Section~\ref{sxn:cv},
-modern NLP models have larger weight matrices and display different spectral properties, and 
+modern NLP models have larger weight matrices and display different spectral properties.
 %, both in terms of their architecture (e.g., wheter there are convolutions) as well as how they are used (e.g., whether one uses prediction error or perplexity).
-thus, they provide a very different test for our empirical quality metrics.
+Thus, they provide a very different test for our empirical quality metrics.
 
-Where Norm-base metrics perform reasonably well on well-trained NLP models, they behave anomalously poorly-trained models.
-Indeed, in ``bad models'', the weight matrics may display rank collapse, decreased Frobeinus mass, or unsually small Spectral norms, 
-which may be mis-interpreted as ``smaller is better.''
-In contast, PL-based metrics (the log $\alpha$-Norm metric, 
-$(\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha})$, and the Weighted Alpha metric, $\hat\alpha =\alpha\log\lambda_{max} $) display consistant behavior even on poorly trained models, and we can use these metrics to help indentify
-where the archiectures need repair, and/or more and/or better data is needed.
+While norm-base metrics perform reasonably well on well-trained NLP models, they often behave anomalously on poorly-trained models.
+Indeed, for such ``bad'' models, weight matrics may display rank collapse, decreased Frobeinus mass, or unsually small Spectral norms.
+(This may be misinterpreted as ``smaller is better.'')
+In contrast, PL-based metrics, including the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $) display consistant behavior, even on poorly trained models.
+Indeed, we can use these metrics to help indentify when archiectures need repair and when more and/or better data are needed.
 
-%We now look in detail at the alpha metric, the spectral norm, and the new norm.
-%We compare good and bad models, and a series of increasingly big models
-%trained on big data sets.  We show why alpha works but spectral norm is not enough.
-%The layer Spectral Norm behaves unexpectedly, while $\alpha$-based metrics can give insight into how well trained a model is.
+%\paragraph{What does large $\alpha$ mean?}
+\paragraph{What do large values of $\alpha$ mean?}
 
-%\michael{WHERE: Say large $\alpha$ values don't mean much.}
-\paragraph{What does large $\alpha$ mean?}
-We do note, however, that many NLP models, such as the GPT and BERT, may have weight matrices with unusually large PL exponents
-($\alpha\gg 6$), which indicates these matrices may be \emph{under}-correlated (i.e over-parameterized), although the truncated PL fit
-itself may not be very reliable because the MLE estimator is unreliable in this range.  Phenomenologically, if we examine
-the ESD visually, we can usually describe these $\mathbf{W}$ as in the \emph{Bulk-Decay} or \emph{Bulk-plus-Spikes} phase\cite{MM}.
-We have conjectured previously that very well-trained DNNs would not have many \emph{outlier} $\alpha>6$, 
-and improved versions of GPT (shown below, and BERT (not shown) confirm this.
-
-%\michael{WHERE: Highlight difference between $\alpha$ and $\hat{\alpha}$.}
-\paragraph{The differences between $\alpha$ and $\hat{\alpha}$.}
-To avoid confusion, let us clarify the differences between $\alpha$ and $\hat{\alpha}$.  
-We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated power law (PL), parameterized by 2 values:
-the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.  (Technically, we also need the minimum
-eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.)
-The PL exponent $\alpha$ measures of the 
-amount of correlation in a DNN layer weight matrix $\mathbf{W}$.  It is valid for $\lambda<\lambda_{max}$, and
-it scale invariant (does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.)
-The Weighted-Alpha metric, $\hat{\alpha}=\alpha\log\lambda_{max}$, approximates the log $\alpha$ Norm.
-$(\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha})$ (derived using statistical mechanics and RMT \nred{(in progress)}, 
-and lets us compute a balanced, weighted average log PL norm metrix for the entire DNN.  
-The $\hat{\alpha}$ metric also behaves like an imrpoved, weighted log average Spectral Norm, and may track this metric in some cases.
+Many NLP models, such as GPT and BERT, have some weight matrices with unusually large PL exponents ($\alpha\gg 6$).
+This indicates these matrices may be \emph{under}-correlated (i.e., over-parameterized, relative to the amount of data).
+In this regime, the truncated PL fit itself may not be very reliable because the MLE estimator it uses is unreliable in this range (i.e., the specific $\alpha$ values returned by the truncated PL fits are less reliable, but having large versus small values of $\alpha$ is reliable).
+Phenomenologically, if we examine the ESD visually, we can usually describe these $\mathbf{W}$ as in the \emph{Bulk-Decay} or \emph{Bulk-plus-Spikes} phase~\cite{MM18_TR,MM19_HTSR_ICML}.
+Previous work~\cite{MM18_TR,MM19_HTSR_ICML} has conjectured that very well-trained DNNs would not have many \emph{outlier} $\alpha>6$; and improved versions of GPT (shown below) and BERT (not shown) confirm this.
 
@@
 
-The OpenAI GPT and GPT2 models provide us with the opportunity to analyze two effects: training the same model with different data set sizes; and increasing sizes of both the data set and architectures.
+The OpenAI GPT and GPT2 models provide us with the opportunity to analyze two effects: training the same model with different data set sizes; and increasing the sizes of both the data set and the architectures simultaneously.
 These models have the remarkable ability to generate fake text that appears to the human to be real, and they have generated significant media attention because of the potential for their misuse.
-For this reason, the original GPT model released by OpenAI was trained on on a deficient data set, rendering the model interesting but not fully functional.  
-Later, OpenAI released a much improved model, GPT2 (small), which has the same architecture and number of layers as GPT, but which has been trained on a larger and better data set (and with other changes), making it remarkably good at generating (near) human-quality fake text.
+For this reason, the original GPT model released by OpenAI was trained on a deficient data set, rendering the model interesting but not fully functional.  
+Later, OpenAI released a much improved model, GPT2-small, which has the same architecture and number of layers as GPT, but which has been trained on a larger and better data set (and with other changes), making it remarkably good at generating (near) human-quality fake text.
 %
-By comparing the poorly-trained GPT to the well-trained GPT2, we can indentify empirical indidcators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
+By comparing the poorly-trained (i.e., ``bad'') GPT to the well-trained (i.e., ``good'') GPT2, we can indentify empirical indidcators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
+By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing simultaneously data set and model size.
 
-\charles{More details here.}
 The GPT models we analyze are deployed with the popular HuggingFace PyTorch library~\cite{XXX-XXX}.
-GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ layer Weight Matrices $\mathbf{W}$.
+GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ layer Weight Matrices, $\mathbf{W}$.
 Each Block has 2 components, the Self Attention (attn) and the Projection (proj) matrics.  
@@
 %
-The OpenAI GPT2 (English) models are: \nred{GPT2-small, GPT2-medium, GPT2-large, and GPT2-xl}, \
-having include $12, 24, 36, \text{and }48$ layers, respectively, with increasingly larger weight matrices.
+The OpenAI GPT2 (English) models are: GPT2-small, GPT2-medium, GPT2-large, and GPT2-xl, having $12, 24, 36, \text{and } 48$ layers, respectively, with increasingly larger weight matrices.
 %The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
@@
 %\begin{tabular}{|p{1in}|c|c|c|c|c|}
-\begin{tabular}{|p{0.75in}|c|c|c|c|c|}
+\begin{tabular}{|p{0.67in}|c|c|c|c|c|}
 \hline
@@
 \hline
-GPT2 medium & 98 & 2.08 & 2.58& 9.74 & 10.01 \\
-GPT2 large & 146 & 1.85 & 1.99& 7.67 & 7.94 \\
-GPT2 xl & 194 & 1.86 & 1.92 & 7.17 & 7.51 \\
+GPT2-medium & 98 & 2.08 & 2.58& 9.74 & 10.01 \\
+GPT2-large & 146 & 1.85 & 1.99& 7.67 & 7.94 \\
+GPT2-xl & 194 & 1.86 & 1.92 & 7.17 & 7.51 \\
 \hline
@@
 \end{center}
-\caption{Average value for the Average log Norm metrics and the Weighted Alpha metric for pretrainnd OpenAI GPT and GPT2 models.
+\caption{Average value for the average Log Norm and Weighted Alpha metrics for pretrainnd OpenAI GPT and GPT2 models.
 Column \# refers to number of layers treated.
-Note averages do not include the first embedding layer(s) because they are not (implicitly) normalized.
+Note that the averages do not include the first embedding layer(s) because they are not (implicitly) normalized.
 }
@@
 
-\paragraph{Coarse Analysis: Empirical Quality Metrics for GPT and GPT2.}
+\paragraph{Average Quality Metrics for GPT and GPT2.}
 
@@
 See Table \ref{table:nlp} for a summary of results.
-We start by examining the PL exponents $\alpha$ for GPT and GPT2-small.
+We start by examining trends between GPT and GPT2-small.
 Observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
-Observe also that (with one minor exception involving the log Frobenius norm metric) all four metrics decrease as one goes from GPT2 medium to large to xl, indicating that the larger models indeed look better than the smaller models.
-%(The one minor exception is the log Frobenius norm increasing from $1.85$ to $1.86$, for GPT2 large to GPT2 xl.)
+\michael{IS THERE A TYPO THERE, OR AM I MISSING SOMETHING.}
+We next examine trends between GPT2-medium to GPT2-large to GPT2-xl.
+Observe that (with one minor exception involving the log Frobenius norm metric) all four metrics decrease as one goes from medium to large to xl, indicating that the larger models indeed look better than the smaller models.
+\michael{WE SHOULD PROBABLY REWORD, SINCE IT IS OPPOSITE OF GPT TO GPT2.}
 
-Figure~\ref{fig:GPT-alpha-hist} shows the histogram (empirical density), for all layers, of $\alpha$ for GPT (blue) and this GPT2 (red) model.  
-These two histograms are very different, with GPT2 having both a notably smaller mean $\alpha$, and far fewer unusually-large outlying $\alpha$ values.
+Going beyond average values, Figure~\ref{fig:GPT-alpha-hist} shows the histogram (empirical density), for all layers, of $\alpha$ for GPT and GPT2-small.  
+These two histograms are very different.
+The older deficient GPT has numerous unsually large $\alpha$ exponents---meaning they are not really well-described by a PL fit.
+Indeed, we expect that a poorly-trained model will lack good (i.e., small $\alpha$) PL behavior in many/most layers.
+On the other hand, as expected, the newer improved GPT2-small model has, on average, smaller $\alpha$ values than the older GPT, with all $\alpha\le6$ and with smaller mean/median $\alpha$, and it also has far fewer unusually-large outlying $\alpha$ values than GPT.
+From this (and other results not shown), we see that $\alpha$ provides a good quality metric for comparing these two models, the``bad'' GPT versus the ``good'' GPT2-small.
+%
+This should be contrasted with the behavior displayed by the Frobenius norm (not shown) and the Spectral norm.
 
-From this, we see that $\alpha$ provides a good quality metric for comparing these two models, the``bad'' GPT vs. the ``good'' GPT2.
-The deficient GPT has numerous unsually large $\alpha$ exponents--meaning they are not Heavy Tailed at all.
-Indeed, we expect that a poorly trained model lack Heavy Tailed behavior in all layers.
-On the other hand, as expected, the improved GPT2 model has, on average, smaller $\alpha$ than the older GPT, with all $\alpha\le6$.
+\paragraph{Scale Collapse in Poorly Trained Models.}
 
-\paragraph{Scale Collapse in Poorly Trained Models}
-However, as seen in Figure \ref{fig:GPT-snorm-hist},
-the ``bad'' GPT model has, spuriously, many smaller Spectral Norms $(\log\Vert\mathbf{W}\Vert_{\infty})$,
-compared to the ``good'' GPT2--which violates the belief that smaller Spectral Norms are always better.
-Indeed, because there are so many anonymously small $\Vert\mathbf{W}\Vert_{\infty}$,
-it appears that the GPT model may be exhibiting the \emph{scale collapse} observed
-in the distilled CV models (above).
-This is an extremely important observation because it demonstrates that while the Spectral Norm
-may correlate well with predicted test error, it is not a good indicator of the overall quality of a model,
-and using it as an empirical metric may give spurious results when applied to poorly trained
-or otherwise deficient models.
+We next describe the behavior of the Spectral norm in GPT versus GPT2-small.
+In Figure~\ref{fig:GPT-snorm-hist}, the ``bad'' GPT model has a smaller mean/median Spectral norm as well as, spuriously, many much smaller Spectral norms, compared to the ``good'' GPT2-small, violating the conventional wisdom that smaller Spectral norms are better.
+Indeed, because there are so many anonymously small Spectral norms, it appears that the GPT model may be exhibiting a kind of \emph{Scale Collapse}, like that observed in the in the distilled CV models (in Figure~\ref{fig:resnet204D5L}).
+This is important because it demonstrates that, while the Spectral norm may correlate well with predicted test error, it is \emph{not} a good indicator of the overall model quality.
+It can mispredict good-versus-bad questions in ways not seen with PL-based metrics.
+Using it as an empirical quality metric may give spurious results when applied to poorly-trained or otherwise deficient models.
 
@@
     \centering
-    \subfigure[Power law exponent ($\alpha$)]{
+    \subfigure[PL exponent ($\alpha$)]{
         \includegraphics[width=4.0cm]{img/GPT-alpha-hist.png}
@@
     %\qquad
-    \subfigure[Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
+    \subfigure[Log Spectral Norm ($\log\Vert\mathbf{W}\Vert_{\infty}$)]{
         \includegraphics[width=4.0cm]{img/GPT-snorm-hist.png}
@@
        }
-   \caption{Comparison of PL exponents, $\alpha$, and log Spectral Norms, 
-$(\log\Vert\mathbf{W}\Vert_{\infty})$, for the OpenAI GPT and GPT2 (small) pretrained models.}
+   \caption{Histogram of PL exponents ($\alpha$) and Log Spectral Norms ($\log\Vert\mathbf{W}\Vert_{\infty}$) for weight matrices from the OpenAI GPT and GPT2-small pretrained models.}
 
@@
 
-Note that Figure \ref{fig:GPT-snorm-hist} also shows some unusually large Spectral Norms, but,
-here, simply correspond to the change in normalization of the word embedding layer(s) and can be ignored.
-From Figure \ref{fig:gpt-snorm-layer} (below), we see that these correspond to the first embedding layer(s).
-These layers appear to have a different effective normalization, and therefore a different scale.
-\charles{We discuss this further in the Appendix}
-%For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,
-%but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
-% $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we treat all layers as-is (although one may to normalize
-%the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier).
-Here, we do not include them in our computed average metrics in Table \ref{table:nlp},
-and do not include them in the histogram plot in Figure \ref{fig:GPT-snorm-hist}.
+Note that Figure~\ref{fig:GPT-snorm-hist} also shows some unusually large Spectral Norms.
+Upon examination, e.g., from Figure \ref{fig:gpt-snorm-layer} (below), we see that these correspond to the first embedding layer(s).
+%That is, they correspond to the change in normalization of the word embedding layer(s), and thus they can be ignored.
+These layers have a different effective normalization, and therefore a different scale.
+% \charles{We discuss this further in Appendix~\ref{sxn:appendix}.}
+% For example, in GPT, for most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$, but in the first embedding layer, the maximum is of order $N$ (the number of words in the embedding), or $\lambda_{max}\sim\mathcal{O}(10^{5})$.
+% For GPT and GPT2, we treat all layers as-is (although one may want to normalize the first 2 layers by $\mathbf{X}$ by $1/N$ or treat them as an outlier).
+We discuss this further in Appendix~\ref{sxn:appendix}.
+Here, we do not include them in our computed average metrics in Table~\ref{table:nlp}, and we do not include them in the histogram plot in Figure~\ref{fig:GPT-snorm-hist}.
 
-\paragraph{Layer Analysis: Correlation Flow in GPT and GPT2.}
+\paragraph{Layer Analysis: Correlation Flow and Scale Collapse in GPT and GPT2.}
 
-also differs significantly between GPT and GPT2.
-Figure \ref{fig:gpt-alpha-layer} plots $\alpha$ vs the depth (i.e. a layer id) for each model.
-The deficient GPT model displays two trends in $\alpha$ , one stable with $\alpha\sim 4$,
-and one increasing with layer id, with $\alpha$ reaching as high as $12$.
-In contrast, the well-trained GPT2 model shows consistant and stable patterns, again
-with one stable $\alpha\sim 3.5$ (and below the GPT trend), and the other only
-slightly trending up, with $alpha\le 6$. 
-The scale-invariant $\alpha$ metric lets us identify potentially
-poorly-trained models.
+We also examine in Figure~\ref{fig:gpt-alpha-layers} the PL exponent $\alpha$ and Log Spectral Norm versus layer id, for GPT and GPT2-small.
+Let's start with Figure~\ref{fig:gpt-alpha-layer}, which plots $\alpha$ versus the depth (i.e., layer id) for each model.
+The deficient GPT model displays two trends in $\alpha$, one stable with $\alpha\sim 4$, and one increasing with layer id, with $\alpha$ reaching as high as $12$.
+In contrast, the well-trained GPT2-small model shows consistant and stable patterns, again with one stable $\alpha\sim 3.5$ (and below the GPT trend), and the other only slightly trending up, with $\alpha\le 6$.
+The scale-invariant $\alpha$ metric lets us identify potentially poorly-trained models; and these results show that the correlation flow differs significantly between GPT and GPT2-small (with the better GPT2-small looking more like the better ResNet-1K from Figure~\ref{fig:resnet-alpha-layer}).
 
-Notice also that the Spectral Norms
-(Figure \ref{fig:gpt-snorm-layer})
-and the $\alpha-$Norms
-(Figure \ref{fig:gpt-pnorm-layer})
-display an unusually some mall values for GPT,
-and some increasing trending values for GPT2.
-Generally speaking, we do not think scale-dependennt 
-norm metrics can be directly applied to distinguish ``good'' vs ``bad'' models 
-because of the anamalous scale collapse that is frequently observed in bad models.
+These results should be contrasted with the corresponding results for Spectral Norms, shown in Figure~\ref{fig:gpt-snorm-layer}.
+Attention models have two types of layers, one small and large; and the Spectral Norm, in particular, displays unusually small values for some of these layers for GPT.
+This scale collapse for the poorly-trained GPT is similar to what we observed for the distilled ResNet20 model in Figure\ref{fig:resnet204Dalpha}.
+Because of the anomalous scale collapse that is frequently observed in poorly-trained models, these results suggest that scale-dependennt norm metrics should not be directly applied to distinguish good-versus-bad models.
 
@@
     \centering
-    \subfigure[Power Law exponent $(\alpha)$]{
+    \subfigure[PL exponent ($\alpha$)]{
         \includegraphics[width=4.0cm]{img/GPT-alpha-depth.png}
@@
     }
-    \subfigure[Spectral Norm $(\log\Vert\mathbf{W}\Vert_{\infty})$]{
+    \subfigure[Log Spectral Norm ($\log\Vert\mathbf{W}\Vert_{\infty}$)]{
         \includegraphics[width=4.0cm]{img/GPT-snorm-depth.png}
@@
     }
-    \caption{Comparison of Correlation Flow and Spectral Norm for OpenAI GPT and GPT2   }
+    \caption{PL exponents ($\alpha$) (in (a)) and Log Spectral Norms ($\log\Vert\mathbf{W}\Vert_{\infty}$) (in (b)) for weight matrices from the OpenAI GPT and GPT2-small pretrained models.
+(Note that the Y axes on each plot are different.)
+             In the text, this is interpreted in terms of \emph{Correlation Flow} and \emph{Scale Collapse}.
+            }
     \label{fig:gpt-alpha-layers}
@@
 
-\paragraph{GPT2: medium, large, xl} 
-We now look across the series of increasingly improving GPT2 models, (i.e. ``good, better, best''), 
-by examining both the PL exponent $\alpha$, as well as the various log Norm metrics.  
-In general, as we move from GPT2-small to GPT2-xl, 
-the histograms for both $\alpha$ exponents and the log Norm metrics downshift from larger to smaller values. 
-See Figure \ref{fig:gpt2-histograms}, which shows the histograms over the layer weight matrics
-for fitted $\alpha$, and the log Spectral Norm
- $(\log\Vert\mathbf{W}\Vert_{\infty})$  
-and log Alpha Norm
- $(\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha})$ 
-metrics.
+\paragraph{GPT2: medium, large, xl.}
 
-We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticible 
-between the differing ``good, better, best'' GTP2 models than between the ``good vs bad'' GPT2 and GPT models.
-Unlike GPT, however, the layer (log) Spectral Norms $(\log\vert\mathbf{W}\Vert_{\infty})$ 
-and (log) Alpha Norms $(\log\vert\mathbf{W}\Vert_{\alpha}^{\alpha})$
-behave more as expected for GPT2 layers, with the larger models, consistently having smaller norms. 
-Likewise, Figure \ref{fig:gpt2-snorm-hist} shows decreasing average log Spectral Norms with the larger models.  
-As expected, the norm metrics can indeed distinguish among ``good, better, best'' models
-among a series well trained models.
+We now look across series of increasingly improving GPT2 models (i.e., we consider good-better-best questions), by examining both the PL exponent $\alpha$ as well as the Log Norm metrics.  
+In general, as we move from GPT2-medium to GPT2-xl, histograms for both $\alpha$ exponents and the Log Norm metrics downshift from larger to smaller values. 
+For example, see Figure \ref{fig:gpt2-histograms}, which shows the histograms over the layer weight matrics for fitted PL exponent ($\alpha$) and the Log Alpha Norm ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$)~metric.
 
-We do notice, however, that while the peaks of the $\alpha$ is getting smaller, 
-towards $2.0$,
-the tails of the distribution shifts right, with larger GPT2 models having more
-usually large $\alpha$.  
-We suspect this indicates that these larger GPT2 models are over-parameterized/under-optimized and 
-that they have capacity to support datasets even larger than the recent XL $1.5B$ release~\cite{gpt2-xl}.
+We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticible between the differing good-better-best GTP2 models than between the good-versus-bad GPT and GPT2-small models.
+Unlike GPT, however, the layer Log Alpha Norms behave more as expected for GPT2 layers, with the larger models consistently having smaller norms. 
+Similarly, the Log Spectral Norm also decreases on average with the larger models (not shown).  
+As expected, the norm metrics can indeed distinguish among good-better-best models among a series well-trained models.
+
+We do notice, however, that while the peaks of the $\alpha$ is getting smaller, towards $2.0$, the tails of the distribution shifts right, with larger GPT2 models having more usually large $\alpha$ (also not shown).
+We suspect this indicates that these larger GPT2 models are still under-optimized/over-parameterized (relative to the data on which they were trained) and that they have capacity to support datasets even larger than the recent XL $1.5B$ release~\cite{gpt2-xl}.
 
@@
     }
-    \subfigure[log Alpha Norm]{
+    \subfigure[Log Alpha Norm]{
         \includegraphics[width=4.0cm]{img/GPT2*_fnl_logpnorm_hist.png}
@@
     }
-    \caption{PL exponents ($\alpha$), log Spectral norm $(\log\Vert\mathbf{W}\Vert_{\infty})$, and log Alpha norm $(\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha})$ for different size models in the GPT2 architecture series.
-(Plots omit the first 2 (embedding) layers, because they are normalized differently giving anamolously large values.)}
+    \caption{Histogram of PL exponents ($\alpha$)
+             %Log Spectral norm ($\log\Vert\mathbf{W}\Vert_{\infty}$), 
+              and Log Alpha Norm ($\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$) for weight matrices from models of different sizes in the GPT2 architecture series.
+(Plots omit the first 2 (embedding) layers, because they are normalized differently giving anamolously large values.)
+              \michael{CAN WE HAVE OTHER FIGS, THEY WERE MORE VISUALLY COMPELLING.}
+             }
     \label{fig:gpt2-histograms}

### 2020-02-10 submission/cv_models.tex
figures updated, minor text changes

@@
 %%\ncyan{
+\paragraph{Comparison of VGG, ResNet, and DenseNet Architectures.}
 We can interpret these observations by recalling the architectural differences between the VGG, ResNet, and DenseNet architectures,

### 2020-02-10 submission/introduction.tex
figures updated, minor text changes

@@
 \item
-Norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models, i.e., they can be used to discriminate between ``good-better-best'' models.
-However, they can qualitatively fail for models that may not be well-trained, i.e., to distinguish ``good-versus-bad''~models.
+Norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models.
+%, i.e., they can be used to discriminate between ``good-better-best'' models.
 \item
-PL-based metrics do much better at predicting quality trends in CV/NLP models---quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-versus-bad models.
+However, norm-based meterics may fail for poorly-trained well-trained, i.e., to distinguish ``good-versus-bad''~models.
 \item 
-PL-based metrics can also be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow} and \emph{Scale Collapse} in poorly-trained models) and evaluate post-training modifications (such as model distillation, increasing the dataset size, and other model improvements).
+PL-based metrics do much bette--quanttiatviely--at predicting quality trends in pretrained models%---quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-versus-bad models.
+\item 
+PL-based metrics can be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow} and \emph{Scale Collapse} in poorly-trained models) and model enhancements (i.e.
+distillation, finetuning, etc.)
 \end{itemize}

### 2020-02-10 submission/introduction.tex
fix typo

@@
 \item 
-PL-based metrics do much bette--quanttiatviely--at predicting quality trends in pretrained models%---quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-versus-bad models.
+PL-based metrics do much better at predicting quality trends in CV/NLP models---quantitatively--at predicting quality trends in pretrained models%---quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-versus-bad models.
 \item 

### 2020-02-10 submission/introduction.tex
fixed more typos

--- 2020-02-10 submission/introduction.tex
@@
 \item 
-PL-based metrics do much bette--quanttiatviely--at predicting quality trends in pretrained models%---quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-versus-bad models.
+PL-based metrics do much better at predicting quality trends in CV/NLP models---quantitatively
+%--at predicting quality trends in pretrained models%---quantitatively 
+better at discriminating good-better-best, and qualitatively better at discriminating good-versus-bad models.
 \item 

### 2020-02-10 submission/introduction.tex
updated intro

--- 2020-02-10 submission/introduction.tex
@@
 \item
-However, norm-based meterics may fail for poorly-trained well-trained, i.e., to distinguish ``good-versus-bad''~models.
+Norm-based meterics, howevver, may give spurious results, such as \emph{Scale\  Collapse} for poorly-trained modeles (i.e. models trained without enough data, etc.).  
 \item 
-PL-based metrics do much bette--quanttiatviely--at predicting quality trends in pretrained models%---quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-versus-bad models.
+PL-based metrics do much better--quanttiatviely--at predicting quality trends in pretrained models.
+That is, they are  quantitatively better at discriminating among a series of ``good-better-best", and are qualitatively better at discriminating good-versus-bad models.
 \item 

### 2020-02-10 submission/introduction.tex

@@
 \item
-Norm-based meterics, howevver, may give spurious results, such as \emph{Scale\  Collapse} for poorly-trained modeles (i.e. models trained without enough data, etc.).  
+Norm-based meterics, however, may give spurious results, such as \emph{Scale\  Collapse} for poorly-trained modeles (i.e. models trained without enough data, etc.).  
 \item 
@@
 \item 
-PL-based metrics can be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow} and \emph{Scale Collapse} in poorly-trained models) and model enhancements (i.e.
+PL-based metrics can be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow}) in both well-trained and /or poorly-trained models), and can be used to evaluate model enhancements (i.e.
 distillation, finetuning, etc.)

### 2020-02-10 submission/introduction.tex
updated intro

@@
 \item 
-PL-based metrics can be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow}) in both well-trained and /or poorly-trained models), and can be used to evaluate model enhancements (i.e.
+PL-based metrics can be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow}) in both well- and /or poorly-trained models), and can be used to evaluate model enhancements (i.e.
 distillation, finetuning, etc.)

### 2020-02-10 submission/introduction.tex
updated intro

@@
 \item 
-PL-based metrics can be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow}) in both well- and /or poorly-trained models), and can be used to evaluate model enhancements (i.e.
+PL-based metrics can be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow}) in both well- and /or poorly-trained models), and can be used to evaluate enhancements (i.e.
 distillation, finetuning, etc.)

### 2020-02-10 submission/introduction.tex
revision of bullets in intro

@@
 \item
+\textbf{Norm-based metrics and well-trained models.}
 Norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models.
-%, i.e., they can be used to discriminate between ``good-better-best'' models.
 \item
-Norm-based meterics, however, may give spurious results, such as \emph{Scale\  Collapse} for poorly-trained modeles (i.e. models trained without enough data, etc.).  
+\textbf{Norm-based metrics and poorly-trained models.}
+Norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.), exhibiting \emph{Scale Collapse} for these models. 
 \item 
-PL-based metrics do much better--quanttiatviely--at predicting quality trends in pretrained models.
-That is, they are  quantitatively better at discriminating among a series of ``good-better-best", and are qualitatively better at discriminating good-versus-bad models.
+\textbf{PL-based metrics and model quality.}
+PL-based metrics do much better at predicting quality trends in pretrained CV/NLP models.  
+They are quantitatively better at discriminating ``good-better-best'' trends, and qualitatively better at distinguishing ``good-versus-bad'' models.
 \item 
-PL-based metrics can be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow}) in both well- and /or poorly-trained models), and can be used to evaluate enhancements (i.e.
-distillation, finetuning, etc.)
+\textbf{PL-based metrics and model diagnostics.}
+PL-based metrics can also be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow}) in well-trained and poorly-trained models, and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
 \end{itemize}

### 2020-02-11 submission/appendix.tex
all updates

@@
 \item stack the maps into a single matrix of, say, dimension $((k\times k\times out)\times in)$, run SVD to get $in$ singular values
-\item compute the 2D Fourier Transform (FFT) for each of the $(in, out)$ pairs, and run SVD on the Fourier coeffients~\cite{Long2019}, leading to $\sim(k\times in\times out)$ non-zero singular values.
+\item compute the 2D Fourier Transform (FFT) for each of the $(in, out)$ pairs, and run SVD on the Fourier coeffients~\cite{CNNSVD}, leading to $\sim(k\times in\times out)$ non-zero singular values.
 \end{enumerate}
@@
 Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with
- $\sigma^{2}\sim 1/\sqrt{N}$, or some variant, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers. With this implicit scale, 
+ $\sigma^{2}\sim 1/\sqrt{N}$, or some variant, e.g., the Glorot/Xavier normalization~\cite{GloBen10}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers. With this implicit scale, 
 we do \emph{not} ``renormalize'' the empirical weight matrices, i.e., we use them as-is.
@@
 
-\subsection{Reproducing Table 3, Section~\ref{xsn:all}}
+\subsection{Reproducing Table 3, Section~\ref{sxn:all_cv_models}}
 
@@
 The ImageNet-1K and other pretrained models are taken from the pytorch models in the \texttt{omsr/imgclsmob} 
-``Sandbox for training convolutional networks for computer vision'' github repository
-.\footnote{\url{https://github.com/osmr/imgclsmob}}
+``Sandbox for training convolutional networks for computer vision'' github repository~\cite{osmr}
 The data can be generated in parallel by running each Google Colab notebook (i.e. \texttt{ww\_kdd2020\_0\_100.ipynb}) 
@@
 regressing each empirical log norm metric against the reported Top 1 (and Top 5) errors (as listed on the \texttt{osmr/imgclsmob} github 
-repository README file, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}
+repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}
 We filter out regressions with less than five datapoints.

### 2020-02-11 submission/cv_models.tex
all updates

@@
 In this section, we examine empirical quality metrics described in Section~\ref{sxn:methods} for several CV model architecture series.
-This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current opensource pyTorch framework (version 1.4)~\cite{pyTorch}.
-This also includes a larger set of ResNet models, trained on the ImageNet-1K dataset~\cite{imagenet1k}, provided on the OSMR ``Sandbox for training convolutional networks for computer vision''~\cite{osmr}, which we call the ResNet-1K series.
+This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current opensource pyTorch framework (version 1.4)~\cite{pytorch}.
+This also includes a larger set of ResNet models, trained on the ImageNet-1K dataset~\cite{imagenet}, provided on the OSMR ``Sandbox for training convolutional networks for computer vision''~\cite{osmr}, which we call the ResNet-1K series.
 
@@
 We have examined the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
-To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BN), and it plots the four quality metrics versus the reported test accuracies~\cite{pyTorchVgg},%
+To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BN), and it plots the four quality metrics versus the reported test accuracies~\cite{pytorch},%
 \footnote{That is, these test accuracies have been previously reported and made publicly-avaiable by others.
@@
 
-We consider ResNet20, trained on CIFAR10, before and after applying the Group Regularization distillation technique, as implemented in the \texttt{distiller} package.%
-\footnote{For details, see \url{https://nervanasystems.github.io/distiller/\#distiller-documentation} and also \url{https://github.com/NervanaSystems/distiller}.}
-We analyze the pretrained 4D\_regularized\_5Lremoved baseline and fine-tuned models.  %\cite{distiller_repo}.
+We consider ResNet20, trained on CIFAR10, before and after applying the Group Regularization distillation technique, as implemented in the \texttt{distiller} package.\cite{distiller}
+We analyze the pretrained 4D\_regularized\_5Lremoved baseline and fine-tuned models. 
 The reported baseline test accuracies (Top1$=91.45$ and Top5$=99.75$) are better than the reported fine-tuned test accuracies (Top1$=91.02$ and Top5$=99.67$)~\cite{XXX-XXX}.
-Because the baseline accuracy is greater,  the previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy}) suggests that the baseline Spectral Norms should be \emph{smaller} on average than the fine-tuned ones should be smaller.
-\michael{IS THERE A TYPO, THAT SOUNDS LIKE BOTH ARE SMALLER.}
+Because the baseline accuracy is greater,  the previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy}) suggests that the baseline Spectral Norms should be \emph{smaller} on average than the fine-tuned ones.
 \emph{The oppopsite is observed.}
@@
 %In both cases (Frobenius norm results not shown), we observe the opposite.
+\charles{WHy is this confusing? Call me...}
 Figure~\ref{fig:resnet204D5L} presents the Spectral Norm (here denoted $\log\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}$.%

### 2020-02-11 submission/into.tex
all updates

--- 2020-02-10 submission/introduction.tex
@@
 There are many problems with this approach.
-The training/testing curves give very limited insight into the overall properties of the model; 
-they do not take into account the (often large human and CPU/GPU) time for hyperparameter fiddling;
-they typically do not correlate with other properties of interest such as robustness or fairness or interpretability; 
+The training/testing curves give very limited insight into the overall properties of the model, 
+they do not take into account the (often large human and CPU/GPU) time for hyperparameter fiddling,
+they typically do not correlate with other properties of interest such as robustness or fairness or interpretability, 
 and so on.
 A less well-known problem, but one that is increasingly important, in particular in industrial-scale artificial intelligence (AI), arises when the model \emph{user} is not the model \emph{developer}.
-Here, one may not have access to either the training data or the testing data.
-Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it "as is,'' or to fine-tune and/or compress it and then use it.
+Here, one may not have access to either the training data nor the testing data.
+Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it "as is'', or to fine-tune and/or compress it and then use it.
 
@@
 Moreover, if ML is to become an industrial process, then that process will become siloed: some groups will gather data, other groups will develop models, and other groups will use those models.
-Users of models can not be expected to know the precise details of how models were built, the specifics of data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized,~etc.
+Users of models cannot be expected to know the precise details of how models were built, the specifics of data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized,~etc.
 
@@
 In the end, however, they really require human evaluation. 
-Alternatively, models that cluster user profiles, which are widely used in areas such as marketing and advertising, are unsupervised and have no obvious labels for comparison and/or evaluation.
-In these and other areas, ML objectives can be poor proxies for downstream goals.
+Alternatively, models that cluster user profiles, which are widely used in marketing, are unsupervised and have no obvious labels for comparison and/or evaluation.
 
@@
 
-In this paper, we present and evaluate quality metrics for pretrained deep neural network (DNN) models, and we do so at scale.
+In this paper, we present and evaluate quality metrics for pretrained deep neural network (DNN) models, and at scale.
 We consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
@@
 They are based on norms of weight matrices (such norms have been used in traditional statistical learning theory to bound capacity and construct regularizers) and/or parameters of power law (PL) fits of the eigenvalues of weight matrices (such PL fits are based on statistical mechanics approaches to DNNs).
-Note that, while we use traditional norm-based and PL-based metrics, our goals are not the traditional goals.
-Unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., as with hyperparmeter optimization).%
+Note that, while we use traditional norm-based and PL-based metrics, our goals are different.
+Unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., hyperparmeter optimization).%
 \footnote{One could of course use these techniques to improve training, and we have been asked about that, but we are not interested in that here.
 Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained DNN models.}
-Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves, and in both cases, we ask:
-\begin{quote}
-\emph{Can we predict trends in the quality of pretrained DNN models without access to training or testing data?}  
-\end{quote}
+Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves, and in both cases, we ask
+
+\emph{Can we predict trends in the quality of pre-trained models without access to training or testing data ?}.
 
@@
 %To answer this question, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in CV and NLP.
+To answer this, we analyze hundreds of publicly-available pretrained state-of-the-art CV and NLP models. And we find that:
+\charles{Not visually easy to digest.  Simplify so the reader can quickly glance at this and decide, is this useful or not. }
 %\end{itemize}
-
-To answer this question, we analyze hundreds of publicly-available pretrained state-of-the-art CV and NLP models. 
-Here is a summary of our main results.
+%Third, 
+%\paragraph{Our main contributions are:}
 \begin{itemize}
 \item
-Norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models.
-%, i.e., they can be used to discriminate between ``good-better-best'' models.
+Norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best'' models, but they can qualitatively fail for models that may not be well-trained, i.e., to distinguish ``good-versus-bad'' models.
 \item 
-Norm-based meterics, however, may give spurious results, such as \emph{Scale\  Collapse} for poorly-trained modeles (i.e. models trained without enough data, etc.).  
+%Fourth, 
+Power Law (PL)-based metrics do much better---quantitatively better at discriminating good-better-best, and qualitatively better at discriminating good-versus-bad models.
 \item 
-PL-based metrics do much better--quanttiatviely--at predicting quality trends in pretrained models.
-That is, they are  quantitatively better at discriminating among a series of ``good-better-best", and are qualitatively better at discriminating good-versus-bad models.
-\item 
-PL-based metrics can be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow}) in both well- and /or poorly-trained models), and can be used to evaluate enhancements (i.e.
-distillation, finetuning, etc.)
+%Finally, 
+PL-based metrics can also be used to characterize fine-scale model properties, including layer-wise \emph{correlation flow}, and evaluating post-training modifications such as model distillation, increasing the dataset size, and other model improvements.
 \end{itemize}
@@
 We emphasize that our goal is a practical theory to predict trends in the quality of state-of-the-art DNN models, i.e., not to make a statement about every publicly-available model.
-We have examined hundreds of models, and we identify general trends, but we also highlight interesting exceptions.
+We have examined hundreds of models, and we indentify general trends, but we also highlight interesting exceptions.
 %%Several of the most interesting are described below.
@@
 
-All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
-To be fully reproducible, we only examine publically-available, pretrained models, and we also provide all Jupyter and Google Colab notebooks used in an accompanying github repository.%
-\footnote{\url{https://github.com/CalculatedContent/kdd2020} \michael{TO BE ANONYMIZED.}}
-See Appendix~\ref{sxn:appendix} for details on how to reproduce all results.
+All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package};
+and, to be fully reproducible, we only examine publically available, pretrained models, and
+also provide all Jupyter and Google Colab notebooks used accompanying github 
+repository\footnote{\url{https://github.com/CalculatedContent/kdd2020}
+\michael{To be anonymized}}
+See the Appendix for more details on how to reproduce all results.
 
@@
 Then, in Section~\ref{sxn:nlp}, we look at several variations of a popular NLP DNN architecture (the OpenAI GPT and GPT2 models); and we show how model quality and properties vary between several variants of GPT and GPT2, including how metrics behave similarly and differently.
-Then, in Section \ref{sxn:all_cv_models}, we present results based on an analysis of hundreds of pretraind DNN models, showing how well each metric predicts the reported test accuracies, and how the PL-based metrics perform remarkably well.
+Then, in Section \ref{sxn:all_cv_models}, we present results based on an analysis of hundreds of pretraind DNN CV models, showing how well each metric predicts the reported test accuracies, and how the Alpha-Norm metric(s) perform remarkably well.
 Finally, in Section~\ref{sxn:conc}, we provide a brief discussion and conclusion.

### 2020-02-11 submission/nlp_models.tex
all updates

@@
 
-The GPT models we analyze are deployed with the popular HuggingFace PyTorch library~\cite{XXX-XXX}.
+The GPT models we analyze are deployed with the popular HuggingFace PyTorch library~\cite{huggingface}.
 GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ layer Weight Matrices, $\mathbf{W}$.
@@
 Observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
-\michael{IS THERE A TYPO THERE, OR AM I MISSING SOMETHING.}
+Notice that in the GPT model, being poorly trained, the norm metrics all exhibit \emph{Scale Collapse} compared to GPT2-small.
+
 We next examine trends between GPT2-medium to GPT2-large to GPT2-xl.
 Observe that (with one minor exception involving the log Frobenius norm metric) all four metrics decrease as one goes from medium to large to xl, indicating that the larger models indeed look better than the smaller models.
-\michael{WE SHOULD PROBABLY REWORD, SINCE IT IS OPPOSITE OF GPT TO GPT2.}
+Notice
+Being well-trained models, the norm metrics now behave as expected, decreasing with increasing accuracy
 

### 2020-02-11 submission/nlp_models.tex
all updates

@@
 (Plots omit the first 2 (embedding) layers, because they are normalized differently giving anamolously large values.)
-              \michael{CAN WE HAVE OTHER FIGS, THEY WERE MORE VISUALLY COMPELLING.}
              }

### 2020-02-11 submission/all_models.tex
all updates

@@
 
-In this section, we summarize results from a large-scale analysis of hundreds of CV models, including models developed for image classification, segmentation, and a range of related tasks.  
+In this section, we summarize results from a large-scale analysis of hundreds of CV models, including models developed for image 
+classification, segmentation, and a range of related tasks.  
 Our aim is to complement the detailed results from Sections~\ref{sxn:cv} and~\ref{sxn:nlp} by providing broader conclusions. % about norm-based and PL-based metrics.
-The models we consider have been pretrained on nine datasets, including ImageNet-1K, and CIFAR-10, CIFAR-100, Street View House Numbers (SVHN), Caltech-UCSD Birds-200-2011 (CUB-200-2011), Pascal VOC2012, ADE20K, Cityscapes, and Common Objects in Context (COCO).
+The models we consider have been pretrained on nine datasets, shown in Table~\ref{table:datasets}.
+%including ImageNet-1K, and CIFAR-10, CIFAR-100, Street View House Numbers (SVHN), Caltech-UCSD Birds-200-2011 (CUB-200-2011), Pascal VOC2012, ADE20K, Cityscapes, and Common Objects in Context (COCO).
 %%The pretrained models and their accuracy metrics are summarized in the osmr github~\cite{charles-add-link}.
-We provide more details about our experimental setup in Appendix~\ref{sxn:appendix}.
-
-We choose simple linear regression to analyze the relationship between quality metrics (computed with the \emph{WeightWatcher} tool) and traditional accuracy metrics (publicly-reported from a test set).
-% (we avoid polynomial regressions as they are more prone to overfitting and does not make economic sense).
-We regress the metrics on the Top1 (and Top5) reported errors (as dependent variables).  
-These include Top5 errors for the ImageNet-1K model, percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy (Pix.Acc.) and Intersection-Over-Union (IOU) for other models.
-We regress them individually on each of the norm-based and PL-based metrics, as described above.
-
-Our results are summarized in Table~\ref{table:results}.
-For the mean, larger $R^{2}$ and smaller $MSE$ are desirable; and for the standard deviation, smaller values are desirable.
-Taken as a whole, over the entire corpus of data, PL-based metrics are somewhat better for both the $R^{2}$ mean and standard deviation; and PL-based metrics are much better for $MSE$ mean and standard deviation.
-These (and other) results suggest our conclusions from Sections~\ref{sxn:cv} and~\ref{sxn:nlp} hold much more generally, and they suggest ovbious questions for future work.
+We provide full details about how to reproduce these results in Appendix~\ref{sxn:appendix}.
 
@@
 \end{table}
+
+We choose simple linear regression to analyze the relationship between quality metrics (computed with the \emph{WeightWatcher} tool) 
+and traditional accuracy metrics (publicly-reported from a test set).
+% (we avoid polynomial regressions as they are more prone to overfitting and does not make economic sense).                            
+We regress the metrics on the Top1 (and Top5) reported errors (as dependent variables).
+These include Top5 errors for the ImageNet-1K model, percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy
+(Pix.Acc.) and Intersection-Over-Union (IOU) for other models.
+We regress them individually on each of the norm-based and PL-based metrics, as described above.
+
+Our results are summarized in Table~\ref{table:results}.
+For the mean, larger $R^{2}$ and smaller $MSE$ are desirable; and for the standard deviation, smaller values are desirable.
+Taken as a whole, over the entire corpus of data, PL-based metrics are somewhat better for both the $R^{2}$ mean and standard deviation;
+and PL-based metrics are much better for $MSE$ mean and standard deviation.
+These (and other) results suggest our conclusions from Sections~\ref{sxn:cv} and~\ref{sxn:nlp} hold much more generally, and they suggest
+obvious questions for future work.

### 2020-02-11 submission/methods.tex
all updates

@@
 
-For norm-based metrics, we use the average of the log norm to the appropriate power.
-Consider, e.g., the \emph{$\alpha$-Shatten Norm metric}, 
-\begin{equation}
-\label{eqn:sum_log_alpha_norm_alpha}
-\sum\nolimits_l \log \Vert\mathbf{W}_l\Vert_{\alpha_l}^{\alpha_l} 
-=
-\sum\nolimits_l \alpha_l \log \Vert\mathbf{W}_l\Vert_{\alpha_l} .
-\end{equation}
+For norm-based metrics, we use the average of the log norm, and to the appropriate power.
 Informallly, this amounts to assuming that the layer weight matrices are statistically independent, in which case we can estimate the model complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm (which resembles a data dependent VC complexity),
@@
 \footnote{When taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.}
-For the $\alpha$-Shatten Norm metric, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it can not be taken out of the sum.
+For the $\alpha$-Shatten Norm metric, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it can not be taken out of the sum:
+
+\begin{equation}
+\label{eqn:sum_log_alpha_norm_alpha}
+\sum\nolimits_l \log \Vert\mathbf{X}_l\Vert_{\alpha_l}^{\alpha_l} 
+=
+\sum\nolimits_l \alpha_l \log \Vert\mathbf{X}_l\Vert_{\alpha_l} .
+\end{equation}
 

### 2020-02-11 submission/methods.tex
all updates

@@
 \footnote{When taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.}
-For the $\alpha$-Shatten Norm metric, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it can not be taken out of the sum:
+
+The \emph{Weighted Alpha metric} is an average of $\alpha_l$ over all layers $l \in \{1,\ldots,l\}$, weighted by the size, or scale, or each matrix
+(and it approximates the average log $\alpha$-Shatten Norm metric),
+\begin{equation}
+\hat{\alpha} = \dfrac{1}{L}\sum_l \alpha_l\log\lambda_{max,l}\approx\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle    ,
+\end{equation}
+where $L$ is the total nuumber of layer weight matrices.
+The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a limited set of models.
+
+Based on this, in this paper, we introduce and evaluate the $\alpha$-Shatten Norm metric.
+Notice for the $\alpha$-Shatten Norm metric, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it can not be taken out of the sum:
 
@@
 
-The \emph{Weighted Alpha metric} is an average of $\alpha_l$ over all layers $l \in \{1,\ldots,l\}$, weighted by the size, or scale, or each matrix
-(and it approximates the average log $\alpha$-Shatten Norm metric),
-\begin{equation}
-\hat{\alpha} = \dfrac{1}{L}\sum_l \alpha_l\log\lambda_{max,l}\approx\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle    ,
-\end{equation}
-where $L$ is the total nuumber of layer weight matrices.
-The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a limited set of models.
-Based on this, in this paper, we introduce and evaluate the $\alpha$-Shatten Norm metric.
 One expects $\hat{\alpha}$ approximates the average log $\alpha$-Shatten Norm very well for $\alpha < 2$ and reasonably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.

### 2020-02-11 submission/cv_models.tex
fixed bib refs

@@
 See Figure~\ref{fig:3models-alpha-layers}, which plots $\alpha$ for each layer (the first layer corresponds to data, the last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (no BN), ResNet, and DenseNet series.
-(Again, much more detailed set of plots is available at \cite{XXX-WEB-LINK}; but note that the corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than the results we present here.)
+(Again, much more detailed set of plots is available at our repo; but note that the corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than the results we present here.)
 
@@
 We analyze the pretrained 4D\_regularized\_5Lremoved baseline and fine-tuned models. 
-The reported baseline test accuracies (Top1$=91.45$ and Top5$=99.75$) are better than the reported fine-tuned test accuracies (Top1$=91.02$ and Top5$=99.67$)~\cite{XXX-XXX}.
+The reported baseline test accuracies (Top1$=91.45$ and Top5$=99.75$) are better than the reported fine-tuned test accuracies (Top1$=91.02$ and Top5$=99.67$).
 Because the baseline accuracy is greater,  the previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy}) suggests that the baseline Spectral Norms should be \emph{smaller} on average than the fine-tuned ones.

### 2020-02-11 submission/empirical2_TRfmt.tex
fixed bib refs

%\documentclass[12pt]{article}
\documentclass[11pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\evensidemargin}{-0.5in}
%\addtolength{\topmargin}{-0.5in}
\addtolength{\topmargin}{-1.0in}
\addtolength{\textheight}{1.7in}
\newlength{\defbaselineskip}
\setlength{\defbaselineskip}{\baselineskip}

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{framed}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsmath}
%%% \usepackage{verbatim}
\usepackage{graphicx}
%\usepackage{caption}
%%\DeclareCaptionType{copyrightbox}
%\usepackage{subcaption}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{color}
\usepackage{xcolor}
%%\usepackage{enumitem}
\usepackage{paralist}

%\usepackage{subfig}
\usepackage{subfigure}

\usepackage{longtable} 
\usepackage{makecell}

\usepackage[multiple]{footmisc}
\usepackage[section]{placeins}

\usepackage{hyperref}
\hypersetup{
     colorlinks   = true,
     linkcolor    = blue,
     citecolor    = green
}

%-----------------------------------------------------------------------

\newcommand{\fix}[1]{\textcolor{red}{#1}}
\newcommand{\comment}[1]{\textcolor{blue}{#1}}
\newcommand{\awk}[1]{\textcolor{green}{#1}}

\newcommand{\argmin}{\text{argmin}}
\newcommand{\Probab}[1]{\mbox{}{\bf{Pr}}\left[#1\right]}
\newcommand{\Expect}[1]{\mbox{}{\bf{E}}\left[#1\right]}
\newcommand{\ExpectBracket}[1]{\mbox{}\langle#1\rangle}

% Here are two macros for comments.
\newcommand {\nred}[1]{{\color{red}\sf{[#1]}}}
\newcommand {\ngreen}[1]{{\color{green}\sf{[#1]}}}
\newcommand {\ncyan}[1]{{\color{cyan}\sf{[#1]}}}
\newcommand {\michael}[1]{{\color{red}\sf{[michael: #1]}}}
\newcommand {\charles}[1]{{\color{blue}\sf{[charles: #1]}}}
\newcommand {\serena}[1]{{\color{orange}\sf{[charles: #1]}}}

\usepackage[normalem]{ulem}

%-----------------------------------------------------------------------

\begin{document}

\title{%
Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data
}

\author{%
Charles H. Martin\thanks{Calculation Consulting, 8 Locksley Ave, 6B, San Francisco, CA 94122, \texttt{charles@CalculationConsulting.com}.} 
\and 
Tongsu (Serena) Peng\thanks{Calculation Consulting, 8 Locksley Ave, 6B, San Francisco, CA 94122, \texttt{serenapeng7@gmail.com}.}
\and
Michael W. Mahoney\thanks{ICSI and Department of Statistics, University of California at Berkeley, Berkeley, CA 94720, \texttt{mmahoney@stat.berkeley.edu}.}
}

\date{}
\maketitle

\begin{abstract}
\input{abstract_file}
\end{abstract}

%GLG | Powered by Box

\input{introduction}
\input{background}
\input{methods}
\input{cv_models}
\input{nlp_models}
\input{all_models}
\input{conclusion}

\noindent
\paragraph{Acknowledgements.}
MWM would like to acknowledge ARO, DARPA, NSF, and ONR for providing partial support of this work.
We would also like to thank Amir Khosrowshahi and colleagues at Intel for helpful discussion regarding the Group Regularization distillation technique.

\bibliographystyle{unsrt}
%\bibliographystyle{plain}
{\small
%\bibliography{gen_gap}
%\bibliography{dnns}
\bibliography{dnns,gen_gap}
}

\appendix
\input{appendix}

\end{document}

### 2020-02-11 submission/appendix.tex
appendix done

@@
 Importantly, the normalization of weight matrices does \emph{not} affect the PL fits because $\alpha$ is scale-invariant.
-Norm-based metrics, however, do depend strongly on the scale of the weight matrix--\nred{that is the point.}
+Norm-based metrics, however, do depend strongly on the scale of the weight matrix--that is the point.
 %\nred{Indeed, early theoretical work by Bartlett suggests that the test accuracy depends strongly on the ``total size'' of the weight matrics.}
@@
 we do \emph{not} ``renormalize'' the empirical weight matrices, i.e., we use them as-is.
-The only exception is that we do rescale the Conv2D pre-activation maps $\mathbf{W}_{i,L}$ 
+The only exception is that \emph{we do rescale} the Conv2D pre-activation maps $\mathbf{W}_{i,L}$ 
 by $k/\sqrt{2}$ so that they are on the same scale as the Linear / Fully Connected (FC) layers.
@@
 NLP models, and other models with large initial embeddings require special care because the
-embedding layers frequently lack the implicit $latex\frac{1}{\sqrt{N}}$ normalization present in other layers.
+embedding layers frequently lack the implicit $\frac{1}{\sqrt{N}}$ normalization present in other layers.
 For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,
@@
 
-We provide a github repository for this paper that includes jupyter notsbooks that fully reproduce all results.
+We provide a github repository for this paper that includes Jupyter notebooks that fully reproduce all results.
 All results have been produced using the weightwatcher tool (v0.2.7).
-The ImageNet and OpenAI GPT pretrained models are provided in the current pyTorch, torchvision, and huggingface distributions,
-as specified in the \texttt{requirements.txt} file.
+The ImageNet and OpenAI GPT pretrained models are provided in the current 
+pyTorch~\cite{pytorch} and huggingface~\cite{higgingface} distributions, as specified in the \texttt{requirements.txt} file.
 
@@
 
-\subsection{Reproducing Table 3, Section~\ref{sxn:all_cv_models}}
+\subsection{Reproducing Table~\ref{table:results}, Section~\ref{sxn:all_cv_models} }
 
-We provide several Google Colab notebooks which can be used to reproduce the results in Table 3,
+We provide several Google Colab notebooks which can be used to reproduce the of Section~\ref{sxn:all_cv_models}.
 in the \texttt{ww-colab} folder in our github repo.  
@@
 ``Sandbox for training convolutional networks for computer vision'' github repository~\cite{osmr}
-The data can be generated in parallel by running each Google Colab notebook (i.e. \texttt{ww\_kdd2020\_0\_100.ipynb}) 
+The data for each regression can be generated in parallel by running each Google Colab notebook (i.e. \texttt{ww\_kdd2020\_0\_100.ipynb}) 
 simultaneously on the same account.
-The final results are generated with the \texttt{ww\_kdd2020\_results.ipynb} notebook.
+The data generated is analyzed by the \texttt{ww\_kdd2020\_results.ipynb} notebook,
+which runs all regressions and tabulates the results presented in Table~\ref{table:results}.
 
@@
 We omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets as there are less than 15 models
-for those datasets. The final datasets used are shown in Table~\ref{table:datasets}
+for those datasets.  And we filter out regressions with less than five datapoints.
+The final datasets used are shown in Table~\ref{table:datasets}.
 The final architecture series used are shown in  Table~\ref{table:architectures}, with the number of models in each.
@@
 First at the global level, we divide models by datasets and run regression separately on all models of a certain dataset, regardless of the architecture.
-At this level, the plots are quite noisy and clustered as each architecture has its own accuracy trend but, you could still see that most plots show positive relationship with positive coefficients
-The regressions are shown in Figure~\ref{fig:DSalphas}.
+At this level, the plots are quite noisy and clustered as each architecture has its own accuracy trend but, you could still see that most plots show positive relationship with positive coefficients.
+The regressions are shown in Figure~\ref{fig:DSalphas}, as available in the results notebook.
 
@@
 repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}
-We filter out regressions with less than five datapoints.
-We record the R-squared and mean squared errors (MSE). 
-The final results for all series regressions are provided in the \texttt{ww-colab/df\_all.xslx} python pandas dataframe file.
+We record the R-squared and mean squared errors (MSE) for each norm metric, averaged over all regressions for all architectures and datasets.
+Plots are provided for every regression, and more fine grained results may be computed by the reader by 
+analyzing the data in the \texttt{df\_all.xlsx} file.
 
@@
 ESPNetv2                                   & 6 \\
-HRNet                                      & 6 \\
+HRNet                      z                & 6 \\
 SqueezeNet/SqueezeResNet                   & 6 \\
@@
 \end{figure}
-
-\subsection{XXX: PLACEHOLDER STUFF PROBABLY TO BE REMOVED}
-
-Some other comments that we need to weave into a narrative eventually after later sections are written:
-\begin{itemize}
-\item
-GPT versus GPT2.
-What happens when we don't have enough data?
-This is the main question, and we can use out metrics to evaluate that, but we also get very different results for GPT versus GPT2.
-\item
-The spectral norm is a regularizer, used to distinguish good-better-best, not a quality metric.
-For example, it can ``collapse,'' and for bad models we can have small spectral norm.
-So, it isn't really a quality metric.
-\item
-One question that isn't obvious is whether regularization metrics can be used as quality metrics.
-One might think so, but the answer isn't obviously yes.
-We show that the answer is No.
-A regularizer is designed to select a unique solution from a non-unique good-better-best.
-Quality metrics can also distinguish good versus bad.
-\item
-(We should at least mention this is like the statistical thing where we evaluate which model is better, as oposed to asking if a given model is good, I forget the name of that.)
-\item
-There are cases where the model is bad but regularization metric doesn't tell you that.
-Quality should be correlated in an empirical way.
-Correlated with good-better-best; but also tell good-bad.
-\item
-Question: why not use regularier for quality?
-Answer: A regularizer selects from a given set of degenerate models one which is nice or unique.
-It doesn't tell good versus bad, i.e., whether that model class is any good.
-\item
-Thus, it isn't obvious that norm-based metrics should do well, and they don't in general.
-\item
-We give examples of all of these: bad data; defective data; and distill models in a bad way.
-(Of course, bad data means bad model, at least indirectly, since the quality of the data affects the properties of the model.)
-\item
-We can select a model and change it, i.e., we don't just do hyperparameter fiddling.
-\end{itemize}

### 2020-02-12 submission/appendix.tex
typo fix

@@
 The ImageNet and OpenAI GPT pretrained models are provided in the current 
-pyTorch~\cite{pytorch} and huggingface~\cite{higgingface} distributions, as specified in the \texttt{requirements.txt} file.
+pyTorch~\cite{pytorch} and huggingface~\cite{huggingface} distributions, as specified in the \texttt{requirements.txt} file.
 
@@
     \caption{\charles{Preliminary charts:} PL exponent $\alpha$ vs. reported Top1 Test Accuracies for pretrained DNNs available\charles{ref} for 5 different data sets.}
-
     \label{fig:DSalphas}

### 2020-02-12 submission/cv_models.tex
typo fix

@@
 
-We consider ResNet20, trained on CIFAR10, before and after applying the Group Regularization distillation technique, as implemented in the \texttt{distiller} package.\cite{distiller}
+We consider ResNet20, trained on CIFAR10, before and after applying the Group Regularization distillation technique, as implemented in the \texttt{distiller} package~\cite{distiller}.
 We analyze the pretrained 4D\_regularized\_5Lremoved baseline and fine-tuned models. 

### 2020-02-12 submission/appendix.tex
quick pass on appendix, fixed problem but not prefectly, more later

@@
 \paragraph{Normalization of Empirical Matrices.}
+
 Normalization is an important, if underappreciated, practical issue.
@@
 \paragraph{Special consideration for NLP models.}
+
 NLP models, and other models with large initial embeddings require special care because the
@@
 
-\subsection{Reproducing Sections 4 and 5}
+\subsection{Reproducing Sections \ref{sxn:cv} and \ref{sxn:nlp} }   %% \subsection{Reproducing Sections 4 and 5}
 
-We provide a github repository for this paper that includes Jupyter notebooks that fully reproduce all results.
-All results have been produced using the weightwatcher tool (v0.2.7).
+We provide a github repository for this paper that includes Jupyter notebooks that fully reproduce all results (as well as many other results).
+All results have been produced using the \emph{WeightWatcher} tool (v0.2.7).
 The ImageNet and OpenAI GPT pretrained models are provided in the current 
-pyTorch~\cite{pytorch} and huggingface~\cite{huggingface} distributions, as specified in the \texttt{requirements.txt} file.
+pyTorch~\cite{pytorch} and Huggingface~\cite{huggingface} distributions, as specified in the \texttt{requirements.txt} file.
 
@@
 \hline
-1  &  WeightWatcher-VGG.ipynb \\
-2(a)  &  WeightWatcher-ResNet.ipynb \\
-2(b)  &  WeightWatcher-ResNet-1K.ipynb \\
-3(a)  &  WeightWatcher-VGG.ipynb \\
-3(b)  &  WeightWatcher-ResNet.ipynb \\
-3(c)  &  WeightWatcher-DenseNet.ipynb \\
+\ref{fig:vgg-metrics}                                 & WeightWatcher-VGG.ipynb \\
+\ref{fig:resnet-accuracy}                             & WeightWatcher-ResNet.ipynb \\
+\ref{fig:resnet1k-accuracy}                           & WeightWatcher-ResNet-1K.ipynb \\
+\ref{fig:vgg-alpha-layers}                            & WeightWatcher-VGG.ipynb \\
+\ref{fig:resnet-alpha-layer}                          & WeightWatcher-ResNet.ipynb \\
+\ref{fig:densenet-alpha-layer}                        & WeightWatcher-DenseNet.ipynb \\
 \hline
-4 & WeightWatcher-Intel-Distiller-ResNet20.ipynb \\
+\ref{fig:resnet204D5L}                                & WeightWatcher-Intel-Distiller-ResNet20.ipynb \\
 \hline
-5 & WeightWatcher-OpenAI-GPT.ipynb \\
-6, 7 & WeightWatcher-OpenAI-GPT2.ipynb \\
+\ref{fig:GPT-hist}                                    & WeightWatcher-OpenAI-GPT.ipynb \\
+\ref{fig:gpt-alpha-layers}, \ref{fig:gpt2-histograms} & WeightWatcher-OpenAI-GPT2.ipynb \\
 \hline
@@
 \end{center}
-\caption{Jupyter notebooks used to reproduce all results in sections ~\ref{sxn:cv} and ~\ref{sxn:nlp}}
+\caption{Jupyter notebooks used to reproduce all results in Sections~\ref{sxn:cv} and~\ref{sxn:nlp}}
 \label{table:notebooks}
@@
 
-We provide the orginal Jupyter Notebooks, which uses the Intel \texttt{distiller} framework
-.\footnote{\url{https://nervanasystems.github.io/distiller}} in the \texttt{distiller} folder of our 
-github repo. Figure ~\ref{fig:resnet204D5L} is from the  \texttt{``...-Distiller-ResNet20.ipynb''} 
-notebook (see Table~\ref{table:notebooks}).  
-For completeness, we provide both the results described here, and additional results
-on other pretrained and distilled models using the \texttt{weightwatcher} tool.
+In the \texttt{distiller} folder of our github repo, 
+we provide the orginal Jupyter Notebooks, which use the Intel \texttt{distiller} framework.%
+\footnote{\url{https://nervanasystems.github.io/distiller}} 
+Figure~\ref{fig:resnet204D5L} is from the  \texttt{``...-Distiller-ResNet20.ipynb''} notebook (see Table~\ref{table:notebooks}).  
+For completeness, we provide both the results described here, as well as additional results on other pretrained and distilled models using the \emph{WeightWatcher} tool.
 
-\subsection{Reproducing Table~\ref{table:results}, Section~\ref{sxn:all_cv_models} }
+\subsection{Reproducing Table~\ref{table:results} in Section~\ref{sxn:all_cv_models} }
 
-We provide several Google Colab notebooks which can be used to reproduce the of Section~\ref{sxn:all_cv_models}.
-in the \texttt{ww-colab} folder in our github repo.  
+In the \texttt{ww-colab} folder of our github repo, we provide several Google Colab notebooks which can be used to reproduce the results of Section~\ref{sxn:all_cv_models}.
 The ImageNet-1K and other pretrained models are taken from the pytorch models in the \texttt{omsr/imgclsmob} 
-``Sandbox for training convolutional networks for computer vision'' github repository~\cite{osmr}
-The data for each regression can be generated in parallel by running each Google Colab notebook (i.e. \texttt{ww\_kdd2020\_0\_100.ipynb}) 
-simultaneously on the same account.
-The data generated is analyzed by the \texttt{ww\_kdd2020\_results.ipynb} notebook,
-which runs all regressions and tabulates the results presented in Table~\ref{table:results}.
+``Sandbox for training convolutional networks for computer vision'' github repository~\cite{osmr}.
+The data for each regression can be generated in parallel by running each Google Colab notebook (i.e., \texttt{ww\_kdd2020\_0\_100.ipynb}) simultaneously on the same account.
+The data generated are analyzed with \texttt{ww\_kdd2020\_results.ipynb}, which runs all regressions and which tabulates the results presented in Table~\ref{table:results}.
 
@@
 There are over $450$ models in all, and we note that the \texttt{osmr/imgclsmob} repository is constantly being updated with new models.
-We omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets as there are less than 15 models
-for those datasets.  And we filter out regressions with less than five datapoints.
+We omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO, datasets as there are fewer than 15 models for those datasets.  
+Also, we filter out regressions with fewer than five datapoints.
 The final datasets used are shown in Table~\ref{table:datasets}.
 The final architecture series used are shown in  Table~\ref{table:architectures}, with the number of models in each.
-
-To further explain how to reproduce our analysis, we run three batches of linear regressions.
-First at the global level, we divide models by datasets and run regression separately on all models of a certain dataset, regardless of the architecture.
-At this level, the plots are quite noisy and clustered as each architecture has its own accuracy trend but, you could still see that most plots show positive relationship with positive coefficients.
-The regressions are shown in Figure~\ref{fig:DSalphas}, as available in the results notebook.
-
-To generate the results in Table 3, we run linear for each architecture series in Table~\ref{table:architectures},
-regressing each empirical log norm metric against the reported Top 1 (and Top 5) errors (as listed on the \texttt{osmr/imgclsmob} github 
-repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}
-We record the R-squared and mean squared errors (MSE) for each norm metric, averaged over all regressions for all architectures and datasets.
-Plots are provided for every regression, and more fine grained results may be computed by the reader by 
-analyzing the data in the \texttt{df\_all.xlsx} file.
 
@@
 \end{figure}
+
+To explain further how to reproduce our analysis, we run three batches of linear regressions. 
+First, at the global level, we divide models by datasets and run regressions separately on all models of a certain dataset, regardless of the architecture. 
+At this level, the plots are quite noisy and clustered, as each architecture has its own accuracy trend; but one can still see that most plots show positive relationship with positive coefficients. 
+Example regressions are shown in Figure~\ref{fig:DSalphas}, as available in the results notebook.
+
+To generate the results in Table~\ref{table:results}, we run linear regressions for each architecture series in Table~\ref{table:architectures}, regressing each empirical log norm metric against the reported Top 1 (and Top 5) errors (as listed on the \texttt{osmr/imgclsmob} github repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}).
+We record the R-squared and mean squared errors (MSE) for each metric, averaged over all regressions for all architectures and datasets.
+Plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.

### 2020-02-12 submission/appendix.tex
results added

@@
 \label{table:datasets}
-\end{table}
+\end{table}`
 
@@
 Plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.
+
+\begin{table}[t]
+\small
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|}
+Dataset & Model & Fnorm & Snorm & wAlpha & aNorm \\
+\hline
+ imagenet-1k & ResNet  & 5.96 &  11.03 & 3.51 & 4.01 \\
+ cifar-10 & ResNet  & 0.31 &  0.30 & 0.28 & 0.28 \\
+ cifar-100 & ResNet  & 4.13 &  4.50 & 3.06 & 3.06 \\
+ svhn & ResNet  & 0.04 &  0.04 & 0.02 & 0.02 \\
+ cub-200-2011 & ResNet  & 0.20 &  0.18 & 3.19 & 3.21 \\
+ imagenet-1k & EfficientNet  & 9.25 &  9.28 & 9.28 & 9.27 \\
+ imagenet-1k & PreResNet  & 6.59 &  15.44 & 3.59 & 3.71 \\
+ imagenet-1k & VGG  & 0.84 &  0.68 & 1.89 & 1.59 \\
+ imagenet-1k & ShuffleNet  & 35.38 &  89.58 & 19.54 & 18.48 \\
+ imagenet-1k & DLA  & 22.41 &  8.49 & 14.69 & 15.68 \\
+ imagenet-1k & HRNet  & 0.47 &  0.51 & 0.16 & 0.16 \\
+ imagenet-1k & DRN  & 0.60 &  0.66 & 0.40 & 0.48 \\
+ imagenet-1k & SqueezeNext  & 21.94 &  21.39 & 13.31 & 13.23 \\
+ imagenet-1k & ESPNetv2  & 13.77 &  14.74 & 1.87 & 2.53 \\
+ imagenet-1k & ProxylessNAS  & 0.19 &  0.26 & 0.28 & 0.26 \\
+ imagenet-1k & SqueezeNet  & 0.11 &  0.11 & 0.07 & 0.08 \\
+ imagenet-1k & IGCV3  & 1.94 &  87.76 & 8.48 & 1.09 \\
+ cifar-10 & DIA-ResNet  & 0.05 &  0.08 & 0.28 & 0.32 \\
+ cifar-100 & DIA-ResNet  & 0.36 &  1.38 & 0.93 & 1.02 \\
+ svhn & DIA-ResNet  & 0.00 &  0.00 & 0.02 & 0.02 \\
+ cifar-10 & SENet  & 0.09 &  0.09 & 0.04 & 0.04 \\
+ cifar-100 & SENet  & 0.36 &  0.43 & 0.26 & 0.26 \\
+ svhn & SENet  & 0.00 &  0.00 & 0.00 & 0.00 \\
+ cub-200-2011 & SENet  & 1.07 &  1.29 & 1.85 & 1.95 \\
+ cifar-100 & WRN  & 0.14 &  0.20 & 0.07 & 0.06 \\
+ svhn & WRN  & 0.01 &  0.01 & 0.01 & 0.01 \\
+ cifar-100 & ResNeXt  & 0.80 &  0.95 & 0.78 & 0.79 \\
+ svhn & ResNeXt  & 0.00 &  0.00 & 0.01 & 0.01 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{MSE Results}
+\label{table:MSEresults}
+\end{table}`
+
+\begin{table}[t]
+\small
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|}
+Dataset & Model & Fnorm & Snorm & wAlpha & aNorm \\
+\hline
+imagenet-1k & ResNet  & 0.82 &  0.67 & 0.90 & 0.88 \\
+ cifar-10 & ResNet  & 0.58 &  0.59 & 0.62 & 0.61 \\
+ cifar-100 & ResNet  & 0.61 &  0.58 & 0.71 & 0.71 \\
+ svhn & ResNet  & 0.69 &  0.70 & 0.82 & 0.81 \\
+ cub-200-2011 & ResNet  & 0.94 &  0.95 & 0.08 & 0.08 \\
+ imagenet-1k & EfficientNet  & 0.00 &  0.00 & 0.00 & 0.00 \\
+ imagenet-1k & PreResNet  & 0.73 &  0.36 & 0.85 & 0.85 \\
+ imagenet-1k & VGG  & 0.71 &  0.76 & 0.35 & 0.45 \\
+ imagenet-1k & ShuffleNet  & 0.63 &  0.06 & 0.80 & 0.81 \\
+ imagenet-1k & DLA  & 0.13 &  0.67 & 0.43 & 0.39 \\
+ imagenet-1k & HRNet  & 0.91 &  0.90 & 0.97 & 0.97 \\
+ imagenet-1k & DRN-C  & 0.81 &  0.79 & 0.87 & 0.85 \\
+ imagenet-1k & SqueezeNext  & 0.05 &  0.07 & 0.42 & 0.43 \\
+ imagenet-1k & ESPNetv2  & 0.42 &  0.38 & 0.92 & 0.89 \\
+ imagenet-1k & ProxylessNAS  & 0.68 &  0.56 & 0.53 & 0.58 \\
+ imagenet-1k & SqueezeNet  & 0.01 &  0.00 & 0.38 & 0.26 \\
+ imagenet-1k & IGCV3  & 0.98 &  0.12 & 0.92 & 0.99 \\
+ cifar-10 & DIA-ResNett  & 0.96 &  0.93 & 0.74 & 0.71 \\
+ cifar-100 & DIA-ResNet  & 0.96 &  0.85 & 0.90 & 0.89 \\
+ svhn & DIA-ResNet  & 0.94 &  0.95 & 0.78 & 0.77 \\
+ cifar-10 & SENet  & 0.91 &  0.91 & 0.96 & 0.96 \\
+ cifar-100 & SENet  & 0.97 &  0.96 & 0.98 & 0.98 \\
+ svhn & SENet & 0.99 &  0.96 & 0.98 & 0.98 \\
+ cub-200-2011 & SENet  & 0.66 &  0.59 & 0.41 & 0.38 \\
+ cifar-100 & WRN  & 0.32 &  0.04 & 0.66 & 0.69 \\
+ svhn & WRN  & 0.13 &  0.10 & 0.20 & 0.21 \\
+ cifar-100 & ResNeXt  & 0.17 &  0.02 & 0.20 & 0.18 \\
+ svhn & ResNeXt  & 0.87 &  0.90 & 0.64 & 0.75 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{$R^{2}$ Results}
+\label{table:R2results}
+\end{table}`

### 2020-02-12 submission/abstract_file.tex
fix typos after full read

@@
 Using the open-source \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision (CV) and natural language processing (NLP).
-We examine both familiar norm-based capacity control metrics (Frobenius and Spectral norms) as well as newer Power-Law (PL) based metrics (including fitted PL exponents, $\alpha$, and the Weighted Alpha metric, $\hat{\alpha}$), from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
+We examine both familiar norm-based capacity control metrics (Frobenius and Spectral norms) as well as newer Power Law (PL) based metrics (including fitted PL exponents, $\alpha$, and the Weighted Alpha metric, $\hat{\alpha}$), from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
 We also introduce the $\alpha$-Shatten Norm metric.

### 2020-02-12 submission/all_models.tex
fix typos after full read

@@
 Our aim is to complement the detailed results from Sections~\ref{sxn:cv} and~\ref{sxn:nlp} by providing broader conclusions. % about norm-based and PL-based metrics.
-The models we consider have been pretrained on nine datasets, shown in Table~\ref{table:datasets}.
+The models we consider have been pretrained on nine datasets.   %, shown in Table~\ref{table:datasets}.
 %including ImageNet-1K, and CIFAR-10, CIFAR-100, Street View House Numbers (SVHN), Caltech-UCSD Birds-200-2011 (CUB-200-2011), Pascal VOC2012, ADE20K, Cityscapes, and Common Objects in Context (COCO).
@@
 \caption{Comparison of linear regression fits for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering 168 (out of 309) different pretrained DNNs.  
-         We include regressions only for architectures with 4 or more data points, and which are postively correlated with test error.
+         We include regressions only for architectures with five or more data points, and which are postively correlated with test error.
          These results can be readily reproduced using the Google Colab notebooks (see Appendix~\ref{sxn:appendix}). 

### 2020-02-12 submission/appendix.tex
fix typos after full read

@@
 Importantly, the normalization of weight matrices does \emph{not} affect the PL fits because $\alpha$ is scale-invariant.
-Norm-based metrics, however, do depend strongly on the scale of the weight matrix--that is the point.
+Norm-based metrics, however, do depend strongly on the scale of the weight matrix---that is the point.
 %\nred{Indeed, early theoretical work by Bartlett suggests that the test accuracy depends strongly on the ``total size'' of the weight matrics.}
@@
 NLP models, and other models with large initial embeddings require special care because the
-embedding layers frequently lack the implicit $\frac{1}{\sqrt{N}}$ normalization present in other layers.
-For example, in GPT, most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,
+embedding layers frequently lack the implicit $1/\sqrt{N}$ normalization present in other layers.
+For example, in GPT, for most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,
 but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
  $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we treat all layers as-is (although one may to normalize
-the first 2 layers by  $\mathbf{X}$ by $\frac{1}{N}$, or to treat them as an outlier).
+the first 2 layers by  $\mathbf{X}$ by $1/N$, or to treat them as an outlier).
 
@@
 
-\subsection{Reproducing Figure~\ref{fig:resnet204D5L}, Distiller Model}
+\subsection{Reproducing Figure~\ref{fig:resnet204D5L}, for the Distiller Model}
 
@@
 There are over $450$ models in all, and we note that the \texttt{osmr/imgclsmob} repository is constantly being updated with new models.
-We omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO, datasets as there are fewer than 15 models for those datasets.  
+We omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets as there are fewer than 15 models for those datasets.  
 Also, we filter out regressions with fewer than five datapoints.
@@
 \label{table:datasets}
-\end{table}`
+\end{table}
 
@@
 
-To generate the results in Table~\ref{table:results}, we run linear regressions for each architecture series in Table~\ref{table:architectures}, regressing each empirical log norm metric against the reported Top 1 (and Top 5) errors (as listed on the \texttt{osmr/imgclsmob} github repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}).
-We record the R-squared and mean squared errors (MSE) for each metric, averaged over all regressions for all architectures and datasets.
+To generate the results in Table~\ref{table:results}, we run linear regressions for each architecture series in Table~\ref{table:architectures}, regressing each empirical Log Norm metric against the reported Top1 (and Top5) errors (as listed on the \texttt{osmr/imgclsmob} github repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}).
+We record the $R^{2}$ and $MSE$ for each metric, averaged over all regressions for all architectures and datasets.
 Plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.
@@
 \end{center}
-\caption{MSE Results}
+\caption{$MSE$ Results}
 \label{table:MSEresults}
-\end{table}`
+\end{table}
 
@@
 \label{table:R2results}
-\end{table}`
+\end{table}

### 2020-02-12 submission/conclusion.tex
fix typos after full read

@@
 We conclude with a few thoughts on what a \emph{practical theory} of DNNs should look like.
-\michael{MM TO DO.}
-%
-We distinguish between what we will call a
-\emph{phenomenological theory}
-(that describes empirical relationship of phenomena to each other, in a way which is consistent with fundamental theory, but is not directly derived from that theory)
-and what can be called a 
-\emph{first principles theory} 
-(that is applicable to toy models, but that does not scale up to realistic systems if one includes realistic apsects of realistic systems).
-%
-For most complex highly-engineered systems (aside from complex AI/ML systems), one \emph{uses} phenomenological theory rather than first principles theory. 
-(One does not try to solve the Schr\"odinger equation if one is interested in building a bridge or flying an airplane.)
-%
-Our results, which are based on our \emph{use} of sophisticated statistical mechanics theory to solve an important practical DNN problems, suggests that this approach should be of interest more generally for those interested in developing a practical DNN theory.
-%\charles{it is the oppposite...ab initio theory scales quite well...it is the spherical cow models of physics and ML that do not scale.
-What we have is a semi-empirical theory.
-We use real theory, but require empirical input, at least in the new stat mech work.
-What we have introduced is a phenomenology, which to me is different from a semi-empirical or phenomenological theory  }
+
+To do so, we distinguish between two types of theories:
+what may be called an \emph{non-empirical or analogical theory}, in which one creates a very simple model that can be analyzed rigorously, and one argues that the model has properties that are analogous the the system of interest; and
+what may be called a \emph{semi-empirical theory}, in which there exists a rigorous asymptotic theory that has asymptotic parameters, and one adjusts or fits the parameters to the finite data.
+
+often one simple enough to perform ``rigorous'' analysis, and one argues that although simple
+
+%\michael{MM TO DO.}
+%%
+%We distinguish between what we will call a
+%\emph{phenomenological theory}
+%(that describes empirical relationship of phenomena to each other, in a way which is consistent with fundamental theory, but is not directly derived from that theory)
+%and what can be called a 
+%\emph{first principles theory} 
+%(that is applicable to toy models, but that does not scale up to realistic systems if one includes realistic apsects of realistic systems).
+%%
+%For most complex highly-engineered systems (aside from complex AI/ML systems), one \emph{uses} phenomenological theory rather than first principles theory. 
+%(One does not try to solve the Schr\"odinger equation if one is interested in building a bridge or flying an airplane.)
+%%
+
+Much of the work on the theory of DNNs focuses on the former type of theory, but our approach is of the latter type.
+Our results, however, which are based on our \emph{use} of sophisticated statistical mechanics theory to solve an important practical DNN problems, suggests that our approach should be of interest more generally for those interested in developing a practical DNN theory.
 }

### 2020-02-12 submission/cv_models.tex
fix typos after full read

@@
 In this section, we examine empirical quality metrics described in Section~\ref{sxn:methods} for several CV model architecture series.
-This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current opensource pyTorch framework (version 1.4)~\cite{pytorch}.
+This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current open source pyTorch framework (version 1.4)~\cite{pytorch}.
 This also includes a larger set of ResNet models, trained on the ImageNet-1K dataset~\cite{imagenet}, provided on the OSMR ``Sandbox for training convolutional networks for computer vision''~\cite{osmr}, which we call the ResNet-1K series.
@@
 %with the Root Mean Squared Error (RMSE) shown, \michael{XXX.  WHERE IS RMSE.  IN THAT TABLE?}
-All four metrics correlate quite well with the reported Top1 accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalziation (i.e., greater accuracy, lower error). 
+All four metrics correlate quite well with the reported Top1 accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalization (i.e., greater accuracy, lower error). 
 While all four metrics perform well, notice that the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) performs best (with an RMSE of $0.42$, see Table~\ref{table:cv-models}); and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $), which is an approximation to the Log $\alpha$-Norm metric~\cite{MM20_unpub_work}, performs second best (with an RMSE of $0.48$, see Table~\ref{table:cv-models}).
@@
 
-See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics on for the VGG, ResNet, and DenseNet series.
+See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics for the VGG, ResNet, and DenseNet series.
 Similar results (not shown) are obtained for the Top5 accuracies.
@@
 (Other metrics exhibit similar behavior.)
-As expected, the higher quality data set shows a better fit, even with fewer data points.
+As expected, the higher quality data set shows a better fit, even with fewer data~points.
 
@@
              Subfigure (d) displays the ResNet models (b), zoomed in to $\alpha\in[1,5]$, and with the layer ids overlaid on the X-axis, from smallest to largest, to
-             allow a more detailed analysis for the most strongly correlated layers.
+             allow a more detailed analysis of the most strongly correlated layers.
              Notice that ResNet152 exhibits different and much more stable behavior of $\alpha$ across layers.
@@
 This optimized approach allows for greater accurary with far fewer parameters (and GPU memory requirements), and
-ResNet models of up to 1000 layers have been trained.\cite{resnet1000}.
+ResNet models of up to 1000 layers have been trained~\cite{resnet1000}.
 %%}
@@
 Contrast this with the DenseNet models, which contains many connections between every layer.
-Our results (large $\alpha$, meaning they even a HT model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly optimized.
+Our results (large $\alpha$, meaning they even a PL model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly optimized.
 
@@
 %In both cases (Frobenius norm results not shown), we observe the opposite.
-\charles{WHy is this confusing? Call me...}
+%%\charles{WHy is this confusing? Call me...}
 Figure~\ref{fig:resnet204D5L} presents the Spectral Norm (here denoted $\log\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}$.%

### 2020-02-12 submission/introduction.tex
fix typos after full read

@@
 Here, one may not have access to either the training data or the testing data.
-Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it "as is,'' or to fine-tune and/or compress it and then use it.
+Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it ``as is,'' or to fine-tune and/or compress it and then use it.
 
@@
 PL-based metrics do much better at predicting quality trends in pretrained CV/NLP models.  
-They are quantitatively better at discriminating ``good-better-best'' trends, and qualitatively better at distinguishing ``good-versus-bad'' models.
+They are quantitatively better at discriminating ``good-better-best'' trends, and qualitatively better at distinguishing ``good-versus-bad''~models.
 \item 
@@
 All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
-To be fully reproducible, we only examine publically-available, pretrained models, and we also provide all Jupyter and Google Colab notebooks used in an accompanying github repository.%
-\footnote{\url{https://github.com/CalculatedContent/kdd2020} \michael{TO BE ANONYMIZED.}}
+To be fully reproducible, we only examine publically-available, pretrained models, and we also provide all Jupyter and Google Colab notebooks used in an accompanying github repository~\cite{kdd20_sub_repo}.
 See Appendix~\ref{sxn:appendix} for details on how to reproduce all results.

### 2020-02-12 submission/methods.tex
fix typos after full read

@@
  takes the form of an average Log Norm.
-For the \emph{Frobenius norm metric} and \emph{Spectral norm metric}, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly.% 
+For the \emph{Frobenius Norm metric} and \emph{Spectral Norm metric}, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly.% 
 \footnote{When taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.}
 
-The \emph{Weighted Alpha metric} is an average of $\alpha_l$ over all layers $l \in \{1,\ldots,l\}$, weighted by the size, or scale, or each matrix
-(and it approximates the average log $\alpha$-Shatten Norm metric),
+The \emph{Weighted Alpha metric} is an average of $\alpha_l$ over all layers $l \in \{1,\ldots,l\}$, weighted by the size, or scale, or each matrix,
+%(and it approximates the average log $\alpha$-Shatten Norm metric),
 \begin{equation}
@@
 
-Based on this, in this paper, we introduce and evaluate the $\alpha$-Shatten Norm metric.
+Based on this, in this paper, we introduce and evaluate the \emph{$\alpha$-Shatten Norm metric}.
 Notice for the $\alpha$-Shatten Norm metric, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it can not be taken out of the sum:
@@
 
-One expects $\hat{\alpha}$ approximates the average log $\alpha$-Shatten Norm very well for $\alpha < 2$ and reasonably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.
+\noindent
+For small $\alpha$, the Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}; and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm, and may track this metric in some~cases.
 
@@
 %
-Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric for the entire~DNN.%
-\footnote{For small $\alpha$, this Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}; and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm, and may track this metric in some cases.}
+Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric for the entire~DNN.
 

### 2020-02-12 submission/nlp_models.tex
typo fix
fix typos after full read

@@
 %(in particular, the interpretation of values of $\alpha \sim 2$ as modeling systems with good correations over many size scales~\cite{BouchaudPotters03, SornetteBook}), 
-this suggests that these models fail to successfully capture many of the correlations in the data (relative to their size) and thus are substantially \emph{under}-trained.
+this suggests that these models fail to capture successfully many of the correlations in the data (relative to their size) and thus are substantially \emph{under}-trained.
 %
@@
 
-While norm-base metrics perform reasonably well on well-trained NLP models, they often behave anomalously on poorly-trained models.
+While norm-based metrics perform reasonably well on well-trained NLP models, they often behave anomalously on poorly-trained models.
 Indeed, for such ``bad'' models, weight matrics may display rank collapse, decreased Frobeinus mass, or unsually small Spectral norms.
@@
 
-Many NLP models, such as GPT and BERT, have some weight matrices with unusually large PL exponents ($\alpha\gg 6$).
+Many NLP models, such as GPT and BERT, have some weight matrices with unusually large PL exponents (e.g., $\alpha\gg 6$).
 This indicates these matrices may be \emph{under}-correlated (i.e., over-parameterized, relative to the amount of data).
@@
 By comparing the poorly-trained (i.e., ``bad'') GPT to the well-trained (i.e., ``good'') GPT2, we can indentify empirical indidcators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
-By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing simultaneously data set and model size.
+By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously.
 
@@
 %
-Because GPTand GPT2 are trained on different data sets, the initial Embedding matrices differ in shape.
+Because GPT and GPT2 are trained on different data sets, the initial Embedding matrices differ in shape.
 GPT has an initial Token and Positional Embedding layers, of dimension $(40478\times 768)$ and $(512\times 768)$, respectively, whereas GPT2 has input Embeddings of shape $(50257\times 768)$ and $(1024\times 768)$, respectively.
@@
 Observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
-Notice that in the GPT model, being poorly trained, the norm metrics all exhibit \emph{Scale Collapse} compared to GPT2-small.
+Notice that in the GPT model, being poorly trained, the norm metrics all exhibit \emph{Scale Collapse}, compared to GPT2-small.
 
@@
 Observe that (with one minor exception involving the log Frobenius norm metric) all four metrics decrease as one goes from medium to large to xl, indicating that the larger models indeed look better than the smaller models.
-Notice
-Being well-trained models, the norm metrics now behave as expected, decreasing with increasing accuracy
+Notice that, for these well-trained models, the norm metrics now behave as expected, decreasing with increasing accuracy.
 
@@
 On the other hand, as expected, the newer improved GPT2-small model has, on average, smaller $\alpha$ values than the older GPT, with all $\alpha\le6$ and with smaller mean/median $\alpha$, and it also has far fewer unusually-large outlying $\alpha$ values than GPT.
-From this (and other results not shown), we see that $\alpha$ provides a good quality metric for comparing these two models, the``bad'' GPT versus the ``good'' GPT2-small.
+From this (and other results not shown), we see that $\alpha$ provides a good quality metric for comparing these two models, the ``bad'' GPT versus the ``good'' GPT2-small.
 %
-This should be contrasted with the behavior displayed by the Frobenius norm (not shown) and the Spectral norm.
+This should be contrasted with the behavior displayed by the Frobenius norm (not shown) and the Spectral~norm.
 
@@
 In Figure~\ref{fig:GPT-snorm-hist}, the ``bad'' GPT model has a smaller mean/median Spectral norm as well as, spuriously, many much smaller Spectral norms, compared to the ``good'' GPT2-small, violating the conventional wisdom that smaller Spectral norms are better.
-Indeed, because there are so many anonymously small Spectral norms, it appears that the GPT model may be exhibiting a kind of \emph{Scale Collapse}, like that observed in the in the distilled CV models (in Figure~\ref{fig:resnet204D5L}).
+Indeed, because there are so many anonymously small Spectral norms, it appears that the GPT model may be exhibiting a kind of \emph{Scale Collapse}, like that observed in the distilled CV models (in Figure~\ref{fig:resnet204D5L}).
 This is important because it demonstrates that, while the Spectral norm may correlate well with predicted test error, it is \emph{not} a good indicator of the overall model quality.
@@
 In contrast, the well-trained GPT2-small model shows consistant and stable patterns, again with one stable $\alpha\sim 3.5$ (and below the GPT trend), and the other only slightly trending up, with $\alpha\le 6$.
-The scale-invariant $\alpha$ metric lets us identify potentially poorly-trained models; and these results show that the correlation flow differs significantly between GPT and GPT2-small (with the better GPT2-small looking more like the better ResNet-1K from Figure~\ref{fig:resnet-alpha-layer}).
+The scale-invariant $\alpha$ metric lets us identify potentially poorly-trained models; and these results show that the Correlation Flow differs significantly between GPT and GPT2-small (with the better GPT2-small looking more like the better ResNet-1K from Figure~\ref{fig:resnet-alpha-layer}).
 
@@
 Attention models have two types of layers, one small and large; and the Spectral Norm, in particular, displays unusually small values for some of these layers for GPT.
-This scale collapse for the poorly-trained GPT is similar to what we observed for the distilled ResNet20 model in Figure\ref{fig:resnet204Dalpha}.
-Because of the anomalous scale collapse that is frequently observed in poorly-trained models, these results suggest that scale-dependennt norm metrics should not be directly applied to distinguish good-versus-bad models.
+This Scale Collapse for the poorly-trained GPT is similar to what we observed for the distilled ResNet20 model in Figure~\ref{fig:resnet204Dalpha}.
+Because of the anomalous scale collapse that is frequently observed in poorly-trained models, these results suggest that scale-dependent norm metrics should not be directly applied to distinguish good-versus-bad models.
 
@@
 (Note that the Y axes on each plot are different.)
-             In the text, this is interpreted in terms of \emph{Correlation Flow} and \emph{Scale Collapse}.
+             In the text, this is interpreted in terms of \emph{Correlation Flow} and \emph{Scale~Collapse}.
             }
@@
 
-We do notice, however, that while the peaks of the $\alpha$ is getting smaller, towards $2.0$, the tails of the distribution shifts right, with larger GPT2 models having more usually large $\alpha$ (also not shown).
+We do notice, however, that while the peaks of the $\alpha$ are getting smaller, towards $2.0$, the tails of the distribution shifts right, with larger GPT2 models having more usually large $\alpha$ (also not shown).
 We suspect this indicates that these larger GPT2 models are still under-optimized/over-parameterized (relative to the data on which they were trained) and that they have capacity to support datasets even larger than the recent XL $1.5B$ release~\cite{gpt2-xl}.

### 2020-02-12 submission/conclusion.tex
last par, whitespace, misc stuff

--- 2020-02-10 submission/conclusion.tex
@@
+\vspace{-1mm}
 \section{Conclusion}
 \label{sxn:conc}
+%\vspace{-1mm}
 
@@
 
-\ncyan{
-We conclude with a few thoughts on what a \emph{practical theory} of DNNs should look like.
-\michael{MM TO DO.}
-%
-We distinguish between what we will call a
-\emph{phenomenological theory}
-(that describes empirical relationship of phenomena to each other, in a way which is consistent with fundamental theory, but is not directly derived from that theory)
-and what can be called a 
-\emph{first principles theory} 
-(that is applicable to toy models, but that does not scale up to realistic systems if one includes realistic apsects of realistic systems).
-%
-For most complex highly-engineered systems (aside from complex AI/ML systems), one \emph{uses} phenomenological theory rather than first principles theory. 
-(One does not try to solve the Schr\"odinger equation if one is interested in building a bridge or flying an airplane.)
-%
-Our results, which are based on our \emph{use} of sophisticated statistical mechanics theory to solve an important practical DNN problems, suggests that this approach should be of interest more generally for those interested in developing a practical DNN theory.
-%\charles{it is the oppposite...ab initio theory scales quite well...it is the spherical cow models of physics and ML that do not scale.
-What we have is a semi-empirical theory.
-We use real theory, but require empirical input, at least in the new stat mech work.
-What we have introduced is a phenomenology, which to me is different from a semi-empirical or phenomenological theory  }
-}
+We conclude with a few comments on what a \emph{practical theory} of DNNs should look like.
+To do so, we distinguish between two types of theories:
+\emph{non-empirical or analogical theories}, in which one creates, often from general principles, a very simple toy model that can be analyzed rigorously, and one then argues that the model is relevant to the system of interest; and
+\emph{semi-empirical theories}, in which there exists a rigorous asymptotic theory, which comes with parameters, for the system of interest, and one then adjusts or fits those parameters to the finite non-asymptotic data.
+A drawback of the former approach is that it typically makes very strong assumptions on the data, and the strength of those assumptions can limit the practical applicability of the theory.
+Much of the work on the theory of DNNs focuses on the former type of theory.
+Our approach focuses on the latter type of theory.
+Our results, which are based on our \emph{use} of sophisticated statistical mechanics theory to solve important practical DNN problems, suggests that our approach should be of interest more generally for those interested in developing a practical DNN theory.

### 2020-02-12 submission/all_models.tex
last par, whitespace, misc stuff
updated notebooks and results

@@
+\vspace{-1mm}
 \section{Comparing Hundreds of CV Models}
 \label{sxn:all_cv_models}
+%\vspace{-1mm}
 
@@
 \end{center}
-\caption{Comparison of linear regression fits for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering 168 (out of 309) different pretrained DNNs.  
+\caption{Comparison of linear regression fits for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering 108 (out of over 400) different pretrained DNNs.
          We include regressions only for architectures with five or more data points, and which are postively correlated with test error.

### 2020-02-12 submission/appendix.tex
updated notebooks and results

--- 2020-02-12 submission/appendix.tex
@@
 ``Sandbox for training convolutional networks for computer vision'' github repository~\cite{osmr}.
-The data for each regression can be generated in parallel by running each Google Colab notebook (i.e., \texttt{ww\_kdd2020\_0\_100.ipynb}) simultaneously on the same account.
-The data generated are analyzed with \texttt{ww\_kdd2020\_results.ipynb}, which runs all regressions and which tabulates the results presented in Table~\ref{table:results}.
+The data for each regression can be generated in parallel by running each Google Colab notebook (i.e., \texttt{ww\_colab\_0\_100.ipynb}) simultaneously on the same account.
+The data generated are analyzed with \texttt{ww\_colab\_results.ipynb}, which runs all regressions and which tabulates the results presented in Table~\ref{table:results}.
 
@@
 There are over $450$ models in all, and we note that the \texttt{osmr/imgclsmob} repository is constantly being updated with new models.
-We omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO, datasets as there are fewer than 15 models for those datasets.  
+We omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets as there are fewer than 15 models for those datasets.  
 Also, we filter out regressions with fewer than five datapoints.
+
+We also remove the following outliers, as identified by visual inspection: \texttt{efficient\_b0,\_b2}.
+And we remove the entire \texttt{cifar100} \texttt{ResNeXT} series, which is the only example show no trends with the norm metrics.
+
 The final datasets used are shown in Table~\ref{table:datasets}.
@@
 \hline
-imagenet-1k   &  78 \\
+imagenet-1k   &  76 \\
 svhn          &  30 \\
-cifar-100     &  30 \\
+cifar-100     &  26 \\
 cifar-10      &  18 \\
@@
 First, at the global level, we divide models by datasets and run regressions separately on all models of a certain dataset, regardless of the architecture. 
-At this level, the plots are quite noisy and clustered, as each architecture has its own accuracy trend; but one can still see that most plots show positive relationship with positive coefficients. 
+At this level, the plots are quite noisy and clustered, as each architecture has its own accuracy trend; but one can
+ still see that most plots show positive relationship with positive coefficients. 
 Example regressions are shown in Figure~\ref{fig:DSalphas}, as available in the results notebook.
 
-To generate the results in Table~\ref{table:results}, we run linear regressions for each architecture series in Table~\ref{table:architectures}, regressing each empirical log norm metric against the reported Top 1 (and Top 5) errors (as listed on the \texttt{osmr/imgclsmob} github repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}).
-We record the R-squared and mean squared errors (MSE) for each metric, averaged over all regressions for all architectures and datasets.
+To generate the results in Table~\ref{table:results}, we run linear regressions for each architecture series in Table~\ref{table:architectures}, regressing each empirical Log Norm metric against the reported Top1 (and Top5) errors (as listed on the \texttt{osmr/imgclsmob} github repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}).
+We record the $R^{2}$ and $MSE$ for each metric, averaged over all regressions for all architectures and datasets.
 Plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.
+The final analysis includes 108 regressions total with 4 or more models, with a positive $R^2$.
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+Dataset & Model & Fnorm & Snorm & wAlpha & aNorm \\
+\hline
+ imagenet-1k & ResNet  & 5.96 &  11.03 & \textbf{3.51} & 4.01 \\
+ imagenet-1k & EfficientNet  & 2.67 &  \textbf{1.23} & 2.56 & 2.50 \\
+ imagenet-1k & PreResNet  & 6.59 &  15.44 & \textbf{3.59} & 3.71 \\
+ imagenet-1k & ShuffleNet  & 35.38 &  89.58 & 19.54 & \textbf{18.48} \\
+ imagenet-1k & VGG  & 0.84 &  \textbf{0.68} & 1.89 & 1.59 \\
+ imagenet-1k & DLA  & 22.41 &  \textbf{8.49} & 14.69 & 15.68 \\
+ imagenet-1k & HRNet  & 0.47 &  0.51 & \textbf{0.16} & 0.16 \\
+ imagenet-1k & DRN-C  & 0.60 &  0.66 & \textbf{0.40} & 0.48 \\
+ imagenet-1k & SqueezeNext  & 21.94 &  21.39 & 13.31 & \textbf{13.23} \\
+ imagenet-1k & ESPNetv2  & 13.77 &  14.74 & \textbf{1.87} & 2.53 \\
+ imagenet-1k & IGCV3  & 1.94 &  87.76 & 8.48 & \textbf{1.09} \\
+ imagenet-1k & ProxylessNAS  & \textbf{0.19} &  0.26 & 0.28 & 0.26 \\
+ imagenet-1k & SqueezeNet  & 0.11 &  0.11 & \textbf{0.07} & 0.08 \\
+\hline
+ cifar-10 & ResNet  & 0.31 &  0.30 & \textbf{0.28} & 0.28 \\
+ cifar-10 & DIA-ResNet  & \textbf{0.05} &  0.08 & 0.28 & 0.32 \\
+ cifar-10 & SENet  & 0.09 &  0.09 & \textbf{0.04} & 0.04 \\
+\hline
+ cifar-100 & ResNet  & 4.13 &  4.50 & \textbf{3.06} & 3.06 \\
+ cifar-100 & DIA-ResNet  & \textbf{0.36} &  1.38 & 0.93 & 1.02 \\
+ cifar-100 & SENet  & 0.36 &  0.43 & \textbf{0.26} & 0.26 \\
+ cifar-100 & WRN  & 0.14 &  0.20 & 0.07 & \textbf{0.06} \\
+\hline
+ svhn & ResNet  & 0.04 &  0.04 & \textbf{0.02} & 0.02 \\
+ svhn & DIA-ResNet  & 0.00 &  \textbf{0.00} & 0.02 & 0.02 \\
+ svhn & SENet  & \textbf{0.00} &  0.00 & 0.00 & 0.00 \\
+ svhn & WRN  & 0.01 &  0.01 & 0.01 & \textbf{0.01} \\
+ svhn & ResNeXt  & 0.00 &  \textbf{0.00} & 0.01 & 0.01 \\
+\hline
+ cub-200-2011 & ResNet  & 0.20 &  \textbf{0.18} & 3.19 & 3.21 \\
+ cub-200-2011 & SENet  & \textbf{1.07} &  1.29 & 1.85 & 1.95 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{MSE Results for all CV model regressions.}
+\label{table:MSEresults}
+\end{table}
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+Dataset & Model & Fnorm & Snorm & wAlpha & aNorm \\
+\hline
+ imagenet-1k & ResNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
+ imagenet-1k & EfficientNet  & 1.07 &  \textbf{1.29} & 1.85 & 1.95 \\
+ imagenet-1k & PreResNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
+ imagenet-1k & ShuffleNet  & 1.07 &  1.29 & 1.85 & \textbf{1.95} \\
+ imagenet-1k & VGG  & 1.07 &  \textbf{1.29} & 1.85 & 1.95 \\
+ imagenet-1k & DLA  & 1.07 &  \textbf{1.29} & 1.85 & 1.95 \\
+ imagenet-1k & HRNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
+ imagenet-1k & DRN-C  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
+ imagenet-1k & SqueezeNext  & 1.07 &  1.29 & 1.85 & \textbf{1.95} \\
+ imagenet-1k & ESPNetv2  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
+ imagenet-1k & IGCV3  & 1.07 &  1.29 & 1.85 & \textbf{1.95} \\
+ imagenet-1k & ProxylessNAS  & \textbf{1.07} &  1.29 & 1.85 & 1.95 \\
+ imagenet-1k & SqueezeNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
+\hline
+ cifar-10 & ResNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
+ cifar-10 & DIA-ResNet  & \textbf{1.07} &  1.29 & 1.85 & 1.95 \\
+ cifar-10 & SENet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
+\hline
+ cifar-100 & ResNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
+ cifar-100 & DIA-ResNet  & \textbf{1.07} &  1.29 & 1.85 & 1.95 \\
+ cifar-100 & SENet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
+ cifar-100 & WRN  & 1.07 &  1.29 & 1.85 & \textbf{1.95} \\
+\hline
+ svhn & ResNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
+ svhn & DIA-ResNet  & 1.07 &  \textbf{1.29} & 1.85 & 1.95 \\
+ svhn & SENet  & \textbf{1.07} &  1.29 & 1.85 & 1.95 \\
+ svhn & WRN  & 1.07 &  1.29 & 1.85 & \textbf{1.95} \\
+ svhn & ResNeXt  & 1.07 &  \textbf{1.29} & 1.85 & 1.95 \\
+\hline
+ cub-200-2011 & ResNet  & 1.07 &  \textbf{1.29} & 1.85 & 1.95 \\
+ cub-200-2011 & SENet  & \textbf{1.07} &  1.29 & 1.85 & 1.95 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{$R^{2}$ Results for all CV model regressions.}
+\label{table:R2results}
+\end{table}

### 2020-02-12 submission/appendix.tex
removed R2 table

@@
 \end{table}
-
-\begin{table}[t]
-\begin{center}
-\begin{tabular}{|c|c|c|c|c|c|}
-\hline
-Dataset & Model & Fnorm & Snorm & wAlpha & aNorm \\
-\hline
- imagenet-1k & ResNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
- imagenet-1k & EfficientNet  & 1.07 &  \textbf{1.29} & 1.85 & 1.95 \\
- imagenet-1k & PreResNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
- imagenet-1k & ShuffleNet  & 1.07 &  1.29 & 1.85 & \textbf{1.95} \\
- imagenet-1k & VGG  & 1.07 &  \textbf{1.29} & 1.85 & 1.95 \\
- imagenet-1k & DLA  & 1.07 &  \textbf{1.29} & 1.85 & 1.95 \\
- imagenet-1k & HRNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
- imagenet-1k & DRN-C  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
- imagenet-1k & SqueezeNext  & 1.07 &  1.29 & 1.85 & \textbf{1.95} \\
- imagenet-1k & ESPNetv2  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
- imagenet-1k & IGCV3  & 1.07 &  1.29 & 1.85 & \textbf{1.95} \\
- imagenet-1k & ProxylessNAS  & \textbf{1.07} &  1.29 & 1.85 & 1.95 \\
- imagenet-1k & SqueezeNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
-\hline
- cifar-10 & ResNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
- cifar-10 & DIA-ResNet  & \textbf{1.07} &  1.29 & 1.85 & 1.95 \\
- cifar-10 & SENet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
-\hline
- cifar-100 & ResNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
- cifar-100 & DIA-ResNet  & \textbf{1.07} &  1.29 & 1.85 & 1.95 \\
- cifar-100 & SENet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
- cifar-100 & WRN  & 1.07 &  1.29 & 1.85 & \textbf{1.95} \\
-\hline
- svhn & ResNet  & 1.07 &  1.29 & \textbf{1.85} & 1.95 \\
- svhn & DIA-ResNet  & 1.07 &  \textbf{1.29} & 1.85 & 1.95 \\
- svhn & SENet  & \textbf{1.07} &  1.29 & 1.85 & 1.95 \\
- svhn & WRN  & 1.07 &  1.29 & 1.85 & \textbf{1.95} \\
- svhn & ResNeXt  & 1.07 &  \textbf{1.29} & 1.85 & 1.95 \\
-\hline
- cub-200-2011 & ResNet  & 1.07 &  \textbf{1.29} & 1.85 & 1.95 \\
- cub-200-2011 & SENet  & \textbf{1.07} &  1.29 & 1.85 & 1.95 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{$R^{2}$ Results for all CV model regressions.}
-\label{table:R2results}
-\end{table}

### 2020-02-12 submission/appendix.tex
updated r2

@@
 \hline
-Dataset & Model & Fnorm & Snorm & wAlpha & aNorm \\
+Dataset & Model  & $\log\Vert\cdot\Vert_{F}$ & $\log\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
 \hline
@@
 \end{center}
-\caption{MSE Results for all CV model regressions.}
+\caption{MSE Results for all CV model regressions. }
 \label{table:MSEresults}
 \end{table}
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+Dataset & Model  & $\log\Vert\cdot\Vert_{F}$ & $\log\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
+\hline
+imagenet-1k & ResNet  & 0.82 &  0.67 & \textbf[0.90] & 0.88 \\
+ imagenet-1k & EfficientNet  & 0.65 &  \textbf[0.84] & 0.67 & 0.67 \\
+ imagenet-1k & PreResNet  & 0.73 &  0.36 & \textbf[0.85] & 0.85 \\
+ imagenet-1k & ShuffleNet  & 0.63 &  0.06 & 0.80 & \textbf[0.81] \\
+ imagenet-1k & VGG  & 0.71 &  \textbf[0.76] & 0.35 & 0.45 \\
+ imagenet-1k & DLA  & 0.13 &  \textbf[0.67] & 0.43 & 0.39 \\
+ imagenet-1k & HRNet  & 0.91 &  0.90 & \textbf[0.97] & 0.97 \\
+ imagenet-1k & DRN-C  & 0.81 &  0.79 & \textbf[0.87] & 0.85 \\
+ imagenet-1k & SqueezeNext  & 0.05 &  0.07 & 0.42 & \textbf[0.43] \\
+ imagenet-1k & ESPNetv2  & 0.42 &  0.38 & \textbf[0.92] & 0.89 \\
+ imagenet-1k & IGCV3  & 0.98 &  0.12 & 0.92 & \textbf[0.99] \\
+ imagenet-1k & SqueezeNet  & 0.01 &  0.00 & \textbf[0.38] & 0.26 \\
+ imagenet-1k & ProxylessNAS  & \textbf[0.68] &  0.56 & 0.53 & 0.58 \\
+\hline
+ cifar-10 & ResNet  & 0.58 &  0.59 & \textbf[0.62] & 0.61 \\
+ cifar-10 & DIA-ResNet  & \textbf[0.96] &  0.93 & 0.74 & 0.71 \\
+ cifar-10 & SENet  & 0.91 &  0.91 & \textbf[0.96] & 0.96 \\
+\hline
+ cifar-100 & ResNet  & 0.61 &  0.58 & \textbf[0.71] & 0.71 \\
+ cifar-100 & DIA-ResNet  & \textbf[0.96] &  0.85 & 0.90 & 0.89 \\
+ cifar-100 & SENet  & 0.97 &  0.96 & \textbf[0.98] & 0.98 \\
+ cifar-100 & WRN  & 0.32 &  0.04 & 0.66 & \textbf[0.69] \\
+\hline
+ svhn & ResNet  & 0.69 &  0.70 & \textbf[0.82] & 0.81 \\
+ svhn & DIA-ResNet  & 0.94 &  \textbf[0.95] & 0.78 & 0.77 \\
+ svhn & SENet  & \textbf[0.99] &  0.96 & 0.98 & 0.98 \\
+ svhn & WRN  & 0.13 &  0.10 & 0.20 & \textbf[0.21] \\
+ svhn & ResNeXt  & 0.87 &  \textbf[0.90] & 0.64 & 0.75 \\
+\hline
+ cub-200-2011 & ResNet  & 0.94 &  \textbf[0.95] & 0.08 & 0.08 \\
+ cub-200-2011 & SENet  & \textbf[0.66] &  0.59 & 0.41 & 0.38 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{$R^{2}$ Results for all CV model regressions. }
+\label{table:R2results}
+\end{table}

### 2020-02-12 submission/appendix.tex
updated r2

--- 2020-02-12 submission/appendix.tex
@@
 \end{center}
-\caption{MSE Results for all CV model regressions.}
+\caption{MSE Results for all CV model regressions. }
 \label{table:MSEresults}
 \end{table}
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+Dataset & Model  & $\log\Vert\cdot\Vert_{F}$ & $\log\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
+\hline
+imagenet-1k & ResNet  & 0.82 &  0.67 & \textbf{0.90} & 0.88 \\
+ imagenet-1k & EfficientNet  & 0.65 &  \textbf{0.84} & 0.67 & 0.67 \\
+ imagenet-1k & PreResNet  & 0.73 &  0.36 & \textbf{0.85} & 0.85 \\
+ imagenet-1k & ShuffleNet  & 0.63 &  0.06 & 0.80 & \textbf{0.81} \\
+ imagenet-1k & VGG  & 0.71 &  \textbf{0.76} & 0.35 & 0.45 \\
+ imagenet-1k & DLA  & 0.13 &  \textbf{0.67} & 0.43 & 0.39 \\
+ imagenet-1k & HRNet  & 0.91 &  0.90 & \textbf{0.97} & 0.97 \\
+ imagenet-1k & DRN-C  & 0.81 &  0.79 & \textbf{0.87} & 0.85 \\
+ imagenet-1k & SqueezeNext  & 0.05 &  0.07 & 0.42 & \textbf{0.43} \\
+ imagenet-1k & ESPNetv2  & 0.42 &  0.38 & \textbf{0.92} & 0.89 \\
+ imagenet-1k & IGCV3  & 0.98 &  0.12 & 0.92 & \textbf{0.99} \\
+ imagenet-1k & SqueezeNet  & 0.01 &  0.00 & \textbf{0.38} & 0.26 \\
+ imagenet-1k & ProxylessNAS  & \textbf{0.68} &  0.56 & 0.53 & 0.58 \\
+\hline
+ cifar-10 & ResNet  & 0.58 &  0.59 & \textbf{0.62} & 0.61 \\
+ cifar-10 & DIA-ResNet  & \textbf{0.96} &  0.93 & 0.74 & 0.71 \\
+ cifar-10 & SENet  & 0.91 &  0.91 & \textbf{0.96} & 0.96 \\
+\hline
+ cifar-100 & ResNet  & 0.61 &  0.58 & \textbf{0.71} & 0.71 \\
+ cifar-100 & DIA-ResNet  & \textbf{0.96} &  0.85 & 0.90 & 0.89 \\
+ cifar-100 & SENet  & 0.97 &  0.96 & \textbf{0.98} & 0.98 \\
+ cifar-100 & WRN  & 0.32 &  0.04 & 0.66 & \textbf{0.69} \\
+\hline
+ svhn & ResNet  & 0.69 &  0.70 & \textbf{0.82} & 0.81 \\
+ svhn & DIA-ResNet  & 0.94 &  \textbf{0.95} & 0.78 & 0.77 \\
+ svhn & SENet  & \textbf{0.99} &  0.96 & 0.98 & 0.98 \\
+ svhn & WRN  & 0.13 &  0.10 & 0.20 & \textbf{0.21} \\
+ svhn & ResNeXt  & 0.87 &  \textbf{0.90} & 0.64 & 0.75 \\
+\hline
+ cub-200-2011 & ResNet  & 0.94 &  \textbf{0.95} & 0.08 & 0.08 \\
+ cub-200-2011 & SENet  & \textbf{0.66} &  0.59 & 0.41 & 0.38 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{$R^{2}$ Results for all CV model regressions. }
+\label{table:R2results}
+\end{table}

### 2020-02-12 submission/appendix.tex
appendix on two pages, more to do

@@
 
-We also remove the following outliers, as identified by visual inspection: \texttt{efficient\_b0,\_b2}.
-And we remove the entire \texttt{cifar100} \texttt{ResNeXT} series, which is the only example show no trends with the norm metrics.
-
+We remove the following outliers, as identified by visual inspection: \texttt{efficient\_b0,\_b2}.
+We also remove the entire \texttt{cifar100} \texttt{ResNeXT} series, which is the only example show no trends with the norm metrics.
+%
 The final datasets used are shown in Table~\ref{table:datasets}.
@@
     }
-    \qquad
+    %\qquad
     \subfigure[ CIFAR 10 ]{
@@
     }
-    \qquad
+    %\qquad
     \subfigure[ CIFAR 100 ]{
@@
     }
-    \qquad
+    %\qquad
     \subfigure[ SVHN ]{
@@
     }
-    \qquad
+    %\qquad
     \subfigure[ CUB 200 ]{
@@
     }
-    \caption{\charles{Preliminary charts:} PL exponent $\alpha$ vs. reported Top1 Test Accuracies for pretrained DNNs available\charles{ref} for 5 different data sets.}
+    \caption{% 
+             %\charles{Preliminary charts:} 
+             PL exponent $\alpha$ vs. reported Top1 Test Accuracies for pretrained DNNs available for five different data sets.
+            }
     \label{fig:DSalphas}
@@
 \begin{table}[t]
+\footnotesize  % \small
 \begin{center}
@@
 \end{center}
-\caption{MSE Results for all CV model regressions. }
+\caption{$MSE$ Results for all CV model regressions. }
 \label{table:MSEresults}

### 2020-02-13 submission/appendix.tex
tweaks to appendix, v1 submission to kdd

@@
 
-In this appendix, we provide more details on several issues that are important for reproducibility of our results.
+In this appendix, we provide more details on several issues that are important for the reproducibility of our results.
 
@@
 Each layer is a 4-index tensor of dimension $(w,h,in,out)$, with an $(w\times h)$ filter (or kernel) and $(in, out)$
-channels. When $w=h=k$,  giving $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\times out)$ each. 
+channels. When $w=h=k$, it gives $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\times out)$ each. 
 %
@@
 \begin{enumerate}
-\item run SVD on each pre-Activation Map $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values
-\item stack the maps into a single matrix of, say, dimension $((k\times k\times out)\times in)$, run SVD to get $in$ singular values
+\item run SVD on each pre-Activation Map $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values;
+\item stack the maps into a single matrix of, say, dimension $((k\times k\times out)\times in)$, run SVD to get $in$ singular values;
 \item compute the 2D Fourier Transform (FFT) for each of the $(in, out)$ pairs, and run SVD on the Fourier coeffients~\cite{CNNSVD}, leading to $\sim(k\times in\times out)$ non-zero singular values.
@@
 %\nred{Indeed, early theoretical work by Bartlett suggests that the test accuracy depends strongly on the ``total size'' of the weight matrics.}
-To apply RMT, we usually define $\mathbf{X}$ with $1/N$ normalization, assuming variance of $\sigma^{2}=1.0$.
+To apply RMT, we usually define $\mathbf{X}$ with a $1/N$ normalization, assuming variance of $\sigma^{2}=1.0$.
 %\footnote{For Heavy Tailed theorems, one typically needs a normalization such as \nred{$1/N^{\alpha-1}$. check this}}
@@
 
-NLP models, and other models with large initial embeddings require special care because the
-embedding layers frequently lack the implicit $1/\sqrt{N}$ normalization present in other layers.
-For example, in GPT, for most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$,
-but in the first embedding layer, the maximum is of order N (the number of words in the embedding), or
- $\lambda_{max}\sim\mathcal{O}(10^{5})$.  For GPT and GPT2, we treat all layers as-is (although one may to normalize
-the first 2 layers by  $\mathbf{X}$ by $1/N$, or to treat them as an outlier).
+NLP models, and other models with large initial embeddings, require special care because the embedding layers frequently lack the implicit $1/\sqrt{N}$ normalization present in other layers.
+For example, in GPT, for most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$, but in the first embedding layer, the maximum eigenvalue is of order $N$ (the number of words in the embedding), or $\lambda_{max}\sim\mathcal{O}(10^{5})$.
+For GPT and GPT2, we treat all layers as-is (although one may want to normalize the first 2 layers by  $\mathbf{X}$ by $1/N$, or to treat them as an outlier).
 
@@
 \end{center}
-\caption{Jupyter notebooks used to reproduce all results in Sections~\ref{sxn:cv} and~\ref{sxn:nlp}}
+\caption{Jupyter notebooks used to reproduce all results in Sections~\ref{sxn:cv} and~\ref{sxn:nlp}.}
 \label{table:notebooks}
@@
 There are over $450$ models in all, and we note that the \texttt{osmr/imgclsmob} repository is constantly being updated with new models.
-We omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets as there are fewer than 15 models for those datasets.  
-Also, we filter out regressions with fewer than five datapoints.
+We omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets, as there are fewer than 15 models for those datasets.  
+Also, we filter out regressions with fewer than 5 datapoints.
 
 We remove the following outliers, as identified by visual inspection: \texttt{efficient\_b0,\_b2}.
-We also remove the entire \texttt{cifar100} \texttt{ResNeXT} series, which is the only example show no trends with the norm metrics.
+We also remove the entire \texttt{cifar100} \texttt{ResNeXT} series, which is the only example to show no trends with the norm metrics.
 %
@@
 
-To generate the results in Table~\ref{table:results}, we run linear regressions for each architecture series in Table~\ref{table:architectures}, regressing each empirical Log Norm metric against the reported Top1 (and Top5) errors (as listed on the \texttt{osmr/imgclsmob} github repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}).
-We record the $R^{2}$ and $MSE$ for each metric, averaged over all regressions for all architectures and datasets.
-Plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.
-The final analysis includes 108 regressions total with 4 or more models, with a positive $R^2$.
-
 \begin{table}[t]
@@
 \hline
-Dataset & Model  & $\log\Vert\cdot\Vert_{F}$ & $\log\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
+Dataset & Model  & $\langle\log\Vert\cdot\Vert_{F}\rangle$ & $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ & $\hat{\alpha}$ & $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ \\
 \hline
@@
 \end{table}
+
+To generate the results in Table~\ref{table:results}, we run linear regressions for each architecture series in Table~\ref{table:architectures}, regressing each empirical Log Norm metric against the reported Top1 (and Top5) errors (as listed on the \texttt{osmr/imgclsmob} github repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}).
+We record the $R^{2}$ and $MSE$ for each metric, averaged over all regressions for all architectures and datasets.
+See Table~\ref{table:MSEresults} and Table~\ref{table:R2results}.
+In the repo, plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.
+The final analysis includes 108 regressions in all, those with 4 or more models, with a positive $R^2$.

### 2020-02-13 submission/introduction.tex
last par, whitespace, misc stuff
tweaks to appendix, v1 submission to kdd

@@
+\vspace{-1mm}
 \section{Introduction}
 \label{sxn:intro}
+%\vspace{-1mm}
 
@@
 
+\vspace{-1mm}
 \paragraph{The WeightWatcher Tool.}
@@
 All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
-To be fully reproducible, we only examine publically-available, pretrained models, and we also provide all Jupyter and Google Colab notebooks used in an accompanying github repository~\cite{kdd20_sub_repo}.
+To be fully reproducible, we only examine publically-available, pretrained models, and we also provide all Jupyter and Google Colab notebooks used in an accompanying github repository~\cite{kdd20_sub_repo_anonymized}.
 See Appendix~\ref{sxn:appendix} for details on how to reproduce all results.
 
+\vspace{-1mm}
 \paragraph{Organization of this paper.}

### 2020-02-15 submission/appendix.tex
pass to convert to TR format, e.g, whitespace, etc.

@@
     \subfigure[ImageNet 1K]{
-        \includegraphics[width=2.5cm]{img/imagenet1k_alpha.png}
+        %OLD% \includegraphics[width=2.5cm]{img/imagenet1k_alpha.png}
+        %KDD% \includegraphics[width=2.5cm]{img/all-imagenet-1k_alpha.png}
+        \includegraphics[width=4.9cm]{img/all-imagenet-1k_alpha.png}
         \label{fig:imagenet1k-alpha}
     }
-    %\qquad
+    %KDD% 
+    \qquad
     \subfigure[ CIFAR 10 ]{
-        \includegraphics[width=2.5cm]{img/cifar10_alpha.png}
+        %OLD% \includegraphics[width=2.5cm]{img/cifar10_alpha.png}
+        %KDD% \includegraphics[width=2.5cm]{img/all-cifar-10_alpha.png}
+        \includegraphics[width=4.9cm]{img/all-cifar-10_alpha.png}
         \label{fig:cifar10.alpha}
     }
-    %\qquad
+    %KDD% 
+    \qquad
     \subfigure[ CIFAR 100 ]{
-        \includegraphics[width=2.5cm]{img/cifar100_alpha.png}
+        %OLD% \includegraphics[width=2.5cm]{img/cifar100_alpha.png}
+        %KDD% \includegraphics[width=2.5cm]{img/all-cifar-100_alpha.png}
+        \includegraphics[width=4.9cm]{img/all-cifar-100_alpha.png}
         \label{fig:cifar100.alpha}
     }
-    %\qquad
+    %KDD% 
+    \qquad
     \subfigure[ SVHN ]{
-        \includegraphics[width=2.5cm]{img/svhn_alpha.png}
+        %OLD% \includegraphics[width=2.5cm]{img/svhn_alpha.png}
+        %KDD% \includegraphics[width=2.5cm]{img/all-svhn_alpha.png}
+        \includegraphics[width=4.9cm]{img/all-svhn_alpha.png}
         \label{fig:svhn.alpha}
     }
-    %\qquad
+    %KDD% 
+    \qquad
     \subfigure[ CUB 200 ]{
-        \includegraphics[width=2.5cm]{img/cub200_alpha.png}
+        %OLD% \includegraphics[width=2.5cm]{img/cub200_alpha.png}
+        %KDD% \includegraphics[width=2.5cm]{img/all-cub-200-2011_alpha.png}
+        \includegraphics[width=4.9cm]{img/all-cub-200-2011_alpha.png}
         \label{fig:cub200.alpha}

### 2020-02-15 submission/cv_models.tex
last par, whitespace, misc stuff
pass to convert to TR format, e.g, whitespace, etc.

@@
+%KDD% \vspace{-1mm}
 \section{Comparison of CV models}
 \label{sxn:cv}
+%\vspace{-1mm}
 
@@
 
+%KDD% \vspace{-1mm}
 \paragraph{Average Quality Metrics versus Reported Test Accuracies.}
@@
     \subfigure[Log Frobenius Norm, VGG ]{
-        %\includegraphics[width=5cm]{img/VGG_lognorm_accs.png}
-        \includegraphics[width=4.0cm]{img/VGG_lognorm_accs.png}
+        \includegraphics[width=4.9cm]{img/VGG_lognorm_accs.png}
+        %KDD% \includegraphics[width=4.0cm]{img/VGG_lognorm_accs.png}
         \label{fig:vgg-fnorm}
     }
-    %\qquad
+    %KDD% 
+    \qquad
     \subfigure[Log Spectral Norm, VGG ]{
-        %\includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
-        \includegraphics[width=4.0cm]{img/VGG_spectralnorm_accs.png}
+        \includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
+        %KDD% \includegraphics[width=4.0cm]{img/VGG_spectralnorm_accs.png}
         \label{fig:vgg-snorm}
     }
-    %\qquad
+    %KDD% 
+    \qquad
     \subfigure[ Weighted Alpha, VGG ]{
-        %\includegraphics[width=4.9cm]{img/VGG_alpha_weighted_accs.png}
-        \includegraphics[width=4.0cm]{img/VGG_alpha_weighted_accs.png}
+        \includegraphics[width=4.9cm]{img/VGG_alpha_weighted_accs.png}
+        %KDD% \includegraphics[width=4.0cm]{img/VGG_alpha_weighted_accs.png}
         \label{fig:vgg-walpha}
     }
-    %\qquad
+    %KDD% 
+    \qquad
     \subfigure[Log $\alpha$-Norm, VGG ]{
-        %\includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
-        \includegraphics[width=4.0cm]{img/VGG_logpnorm_accs.png}
+        \includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
+        %KDD% \includegraphics[width=4.0cm]{img/VGG_logpnorm_accs.png}
         \label{fig:vgg-pnorm}
@@
     \subfigure[ ResNet, Log $\alpha$-Norm ]{
-        %\includegraphics[width=4.2cm]{img/ResNet_logpnorm_accs.png}
-        \includegraphics[width=4.0cm]{img/ResNet_logpnorm_accs.png}
+        \includegraphics[width=4.9cm]{img/ResNet_logpnorm_accs.png}
+        %KDD% \includegraphics[width=4.0cm]{img/ResNet_logpnorm_accs.png}
         \label{fig:resnet-accuracy}
     }
-    %\qquad
+    %KDD% 
+    \qquad
     \subfigure[ ResNet-1K, Log $\alpha$-Norm ]{
-        %\includegraphics[width=4.5cm]{img/ResNet-1K_logpnorm_accs.png}
-        \includegraphics[width=4.0cm]{img/ResNet-1K_logpnorm_accs.png}
+        \includegraphics[width=4.9cm]{img/ResNet-1K_logpnorm_accs.png}
+        %KDD% \includegraphics[width=4.0cm]{img/ResNet-1K_logpnorm_accs.png}
         \label{fig:resnet1k-accuracy}
@@
 \begin{center}
-%\begin{tabular}{|p{1in}|c|c|c|c|c|}
-\begin{tabular}{|p{0.65in}|c|c|c|c|c|}
+\begin{tabular}{|p{1in}|c|c|c|c|c|}
+%KDD% \begin{tabular}{|p{0.65in}|c|c|c|c|c|}
 %\begin{tabular}{|l|c|c|c|c|c|}
@@
 
+%KDD% \vspace{-1mm}
 \paragraph{Variation in Data Set Size.}
@@
 
+%KDD% \vspace{-1mm}
 \paragraph{Layer Analysis: Metrics as a Function of Depth.}
@@
     \subfigure[ VGG ]{
-        %\includegraphics[width=4.5cm]{img/VGG_fnl_alpha_depth.png} 
-        \includegraphics[width=4.0cm]{img/VGG_fnl_alpha_depth.png} 
+        \includegraphics[width=4.9cm]{img/VGG_fnl_alpha_depth.png} 
+        %KDD% \includegraphics[width=4.0cm]{img/VGG_fnl_alpha_depth.png} 
                 \label{fig:vgg-alpha-layers}
     }
-    %\qquad
+    %KDD% 
+    \qquad
     \subfigure[ ResNet-1K ]{
-        %\includegraphics[width=4.5cm]{img/ResNet_fnl_alpha_depth.png} 
-        \includegraphics[width=4.0cm]{img/ResNet_fnl_alpha_depth.png} 
+        \includegraphics[width=4.9cm]{img/ResNet_fnl_alpha_depth.png} 
+        %KDD% \includegraphics[width=4.0cm]{img/ResNet_fnl_alpha_depth.png} 
         \label{fig:resnet-alpha-layer}
     }
-    %\qquad
+    %KDD% 
+    \qquad
     \subfigure[ DenseNet ]{
-        %\includegraphics[width=4.5cm]{img/DenseNet_fnl_alpha_depth.png} 
-        \includegraphics[width=4.0cm]{img/DenseNet_fnl_alpha_depth.png} 
+        \includegraphics[width=4.9cm]{img/DenseNet_fnl_alpha_depth.png} 
+        %KDD% \includegraphics[width=4.0cm]{img/DenseNet_fnl_alpha_depth.png} 
         \label{fig:densenet-alpha-layer}
-
-}    \subfigure[ ResNet-1K (overlaid) ]{
-        \includegraphics[width=4.0cm]{img/resnet_alpha_overlaid_depth.png} 
+    }    
+    %KDD% 
+    \qquad
+    \subfigure[ ResNet-1K (overlaid) ]{
+        \includegraphics[width=4.9cm]{img/resnet_alpha_overlaid_depth.png} 
+        %KDD% \includegraphics[width=4.0cm]{img/resnet_alpha_overlaid_depth.png} 
         \label{fig:resnet_alpha_overlaid_depth}
@@
 %The empirical manifestations of this (on weight matrices) are that fitted $\alpha$ values get much worse for deeper layers.
-%%\ncyan{
+
+%KDD% \vspace{-1mm}
 \paragraph{Comparison of VGG, ResNet, and DenseNet Architectures.}
@@
 ResNet models of up to 1000 layers have been trained~\cite{resnet1000}.
-%%}
+
 We conjuecture that the effeciency and effectivness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and strongly optimized.
@@
 
+%KDD% \vspace{-1mm}
 \paragraph{Correlation Flow.}
@@
    \subfigure[$\lambda_{max}$ for ResNet20 layers]{
-      %\includegraphics[scale=0.14]{img/resnet4d_maxev.png}
-      \includegraphics[width=4.0cm]{img/resnet4d_maxev.png}
-      %\includegraphics[width=4.5cm]{img/resnet4d_maxev.png}
+      %KDD% \includegraphics[width=4.0cm]{img/resnet4d_maxev.png}
+      \includegraphics[width=4.9cm]{img/resnet4d_maxev.png}
       \label{fig:resnet204Dmaxev}
    }
-   %\qquad
+   %KDD% 
+   \qquad
    \subfigure[$\alpha$ for ResNet20 layers]{
-      %\includegraphics[scale=0.14]{img/resnet4d_alphas.png}
-      \includegraphics[width=4.0cm]{img/resnet4d_alphas.png}
-      %\includegraphics[width=4.5cm]{img/resnet4d_alphas.png}
+      %KDD% \includegraphics[width=4.0cm]{img/resnet4d_alphas.png}
+      \includegraphics[width=4.9cm]{img/resnet4d_alphas.png}
       \label{fig:resnet204Dalpha}
@@
 
-%% %\paragraph{Layer Analysis: Behavior on Less Thoroughy Studied CV Models.}
-%% \paragraph{Correlation Flow: PL vs. Norm Metrics.}
+%KDD% \vspace{-1mm}
 \paragraph{Scale Collapse; or How Distillation May Break Models.}

### 2020-02-15 submission/nlp_models.tex
last par, whitespace, misc stuff
pass to convert to TR format, e.g, whitespace, etc.

@@
+%KDD% \vspace{-1mm}
 \section{Comparison of NLP Models}
 \label{sxn:nlp}
+%\vspace{-1mm}
 
@@
 
-%\paragraph{What does large $\alpha$ mean?}
+%KDD% \vspace{-1mm}
 \paragraph{What do large values of $\alpha$ mean?}
@@
 
+%KDD% \vspace{-1mm}
 \paragraph{OpenAI GPT Models.}
@@
 \begin{center}
-%\begin{tabular}{|p{1in}|c|c|c|c|c|}
-\begin{tabular}{|p{0.67in}|c|c|c|c|c|}
+\begin{tabular}{|p{1in}|c|c|c|c|c|}
+%KDD% \begin{tabular}{|p{0.67in}|c|c|c|c|c|}
 \hline
@@
 
+%KDD% \vspace{-1mm}
 \paragraph{Average Quality Metrics for GPT and GPT2.}
@@
 
+%KDD% \vspace{-1mm}
 \paragraph{Scale Collapse in Poorly Trained Models.}
@@
     \subfigure[PL exponent ($\alpha$)]{
-        \includegraphics[width=4.0cm]{img/GPT-alpha-hist.png}
+        %KDD% \includegraphics[width=4.0cm]{img/GPT-alpha-hist.png}
+        \includegraphics[width=5.5cm]{img/GPT-alpha-hist.png}
         \label{fig:GPT-alpha-hist}
        }
-    %\qquad
+    %KDD% 
+    \qquad
     \subfigure[Log Spectral Norm ($\log\Vert\mathbf{W}\Vert_{\infty}$)]{
-        \includegraphics[width=4.0cm]{img/GPT-snorm-hist.png}
+        %KDD% \includegraphics[width=4.0cm]{img/GPT-snorm-hist.png}
+        \includegraphics[width=5.5cm]{img/GPT-snorm-hist.png}
         \label{fig:GPT-snorm-hist}
@@
 
+%KDD% \vspace{-1mm}
 \paragraph{Layer Analysis: Correlation Flow and Scale Collapse in GPT and GPT2.}
@@
     \subfigure[PL exponent ($\alpha$)]{
-        \includegraphics[width=4.0cm]{img/GPT-alpha-depth.png}
+        %KDD% \includegraphics[width=4.0cm]{img/GPT-alpha-depth.png}
+        \includegraphics[width=5.5cm]{img/GPT-alpha-depth.png}
         \label{fig:gpt-alpha-layer}
     }
+    %KDD% 
+    \quad
     \subfigure[Log Spectral Norm ($\log\Vert\mathbf{W}\Vert_{\infty}$)]{
-        \includegraphics[width=4.0cm]{img/GPT-snorm-depth.png}
+        %KDD% \includegraphics[width=4.0cm]{img/GPT-snorm-depth.png}
+        \includegraphics[width=5.5cm]{img/GPT-snorm-depth.png}
         \label{fig:gpt-snorm-layer}
@@
 
+%KDD% \vspace{-1mm}
 \paragraph{GPT2: medium, large, xl.}
@@
     \subfigure[PL exponent ($\alpha$)]{
-        \includegraphics[width=4.0cm]{img/GPT2*_fnl_alpha_hist.png}
+        %KDD% \includegraphics[width=4.0cm]{img/GPT2*_fnl_alpha_hist.png}
+        \includegraphics[width=5.5cm]{img/GPT2*_fnl_alpha_hist.png}
         \label{fig:gpt2-alpha-hist}
     }
+    %KDD% 
+    \quad
     \subfigure[Log Alpha Norm]{
-        \includegraphics[width=4.0cm]{img/GPT2*_fnl_logpnorm_hist.png}
+        %KDD% \includegraphics[width=4.0cm]{img/GPT2*_fnl_logpnorm_hist.png}
+        \includegraphics[width=5.5cm]{img/GPT2*_fnl_logpnorm_hist.png}
         \label{fig:gpt2-pnorm-hist}

### 2020-02-16 submission/all_models.tex
pass to convert to TR format, e.g, whitespace, etc.
minor edits

@@
-\vspace{-1mm}
+%KDD% \vspace{-1mm}
 \section{Comparing Hundreds of CV Models}
@@
 \begin{center}
-%\begin{tabular}{|p{1in}|c|c|c|c|}
-\begin{tabular}{|p{0.75in}|c|c|c|c|}
+\begin{tabular}{|p{1in}|c|c|c|c|}
+%KDD% \begin{tabular}{|p{0.75in}|c|c|c|c|}
 \hline
@@
 
-We choose simple linear regression to analyze the relationship between quality metrics (computed with the \emph{WeightWatcher} tool) 
-and traditional accuracy metrics (publicly-reported from a test set).
-% (we avoid polynomial regressions as they are more prone to overfitting and does not make economic sense).                            
+We choose ordinary linear regression (OLS) to quantity the relationship between quality metrics (computed with the \emph{WeightWatcher} tool) 
+and the reported test error and/or accuracy metrics.
 We regress the metrics on the Top1 (and Top5) reported errors (as dependent variables).
@@
 (Pix.Acc.) and Intersection-Over-Union (IOU) for other models.
-We regress them individually on each of the norm-based and PL-based metrics, as described above.
+We regress them individually on each of the norm-based and PL-based metrics, as described in Section~\ref{sxn:cv}.
 

### 2020-02-16 submission/cv_models.tex
minor edits

@@
 Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best.
-These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among ``good-better-best'' models, with PL-based metrics performing somewhat (i.e., quantitatively) better.
+These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among
+a series of ``good-better-best'' models, with PL-based metrics performing somewhat (i.e., quantitatively) better.
 
@@
 Their ESDs can be well-fit by a (truncated) PL, with exponents $\alpha\in[2,4]$.
-Much larger values $(\alpha\gg 5\;\mbox{or}\;6)$ may reflect poorer PL fits, whereas smaller values $(\alpha\sim 2)$, are associated with models that generalize better.
+Much larger values $(\alpha\gg 6)$ may reflect poorer PL fits, whereas smaller values $(\alpha\sim 2)$, are associated with models that generalize better.
 %, but the presence of many adjacent layers with relatively good PL fits \emph{suggests} that those adjacent layers ``impedance match'' well and that the correlations that are captured by one layer propagate well to subsequent layers.
@@
 
-(That being said, Figure~\ref{fig:resnet204Dalpha} also depcits two very large $\alpha\gg 6$ values for the baseline, but not for the fine-tuned, model.
-This suggests the baseline model has at least two over-parameterized/under-trained layers, and that the distillation method does, in fact, improve the fine-tuned model by compressing these layers.)
+That being said, Figure~\ref{fig:resnet204Dalpha} also depcits two very large $\alpha\gg 6$ values for the baseline, but not for the fine-tuned, model.
+This suggests the baseline model has at least two over-parameterized/under-trained layers, and that the distillation method does, in fact, improve the fine-tuned model by compressing these layers.
 
@@
 %%has the unusual effect of increasing
-spuriously increases the norms of the $\mathbf{W}$ pre-activation maps for at least two of the Conv2D layers.
+causes the norms of the $\mathbf{W}$ pre-activation maps for two Conv2D layers to spuriously increase.
 This is difficult to diagnose by analyzing training/test curves, but it is easy to diagnose with our approach.

### 2020-02-16 submission/methods.tex
last par, whitespace, misc stuff
pass to convert to TR format, e.g, whitespace, etc.
minor edits

@@
+%KDD% \vspace{-1mm}
 \section{Methods}
 \label{sxn:methods}
+%\vspace{-1mm}
 
@@
 
+%KDD% \vspace{-1mm}
 \paragraph{DNN Empirical Quality Metrics.}
@@
 \begin{equation}
-\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}  ,
+\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}.
 \end{equation}
-where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
+%where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
 Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.
@@
 The PL exponent $\alpha$ measures of the amount of correlation in a DNN layer weight matrix $\mathbf{W}$. 
-It is valid for $\lambda<\lambda_{max}$, and it is scale-invariant, i.e., it does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.
+It is valid for $\lambda\le\lambda_{max}$, and it is scale-invariant, i.e., it does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.
 The $\lambda_{max}$ is a measure of the size, or scale, of $\mathbf{W}$.
@@
 
+%KDD% \vspace{-1mm}
 \paragraph{Convolutional Layers and Normalization issues.}

### 2020-02-16 submission/nlp_models.tex
minor edits

@@
 Phenomenologically, if we examine the ESD visually, we can usually describe these $\mathbf{W}$ as in the \emph{Bulk-Decay} or \emph{Bulk-plus-Spikes} phase~\cite{MM18_TR,MM19_HTSR_ICML}.
-Previous work~\cite{MM18_TR,MM19_HTSR_ICML} has conjectured that very well-trained DNNs would not have many \emph{outlier} $\alpha>6$; and improved versions of GPT (shown below) and BERT (not shown) confirm this.
+Previous work~\cite{MM18_TR,MM19_HTSR_ICML} has conjectured that very well-trained DNNs would not have many \emph{outlier} $\alpha\gg 6$; and improved versions of GPT (shown below) and BERT (not shown) confirm this.
 
@@
 %
-By comparing the poorly-trained (i.e., ``bad'') GPT to the well-trained (i.e., ``good'') GPT2, we can indentify empirical indidcators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
-By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously.
+By comparing the poorly-trained (i.e., ``bad'') GPT to the well-trained (i.e., ``good'') GPT2-small, we can indentify empirical indicators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
+By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously,
+or what we call a series of ``good-better-best'' models.
 
@@
 
-Note that Figure~\ref{fig:GPT-snorm-hist} also shows some unusually large Spectral Norms.
+(Note that Figure~\ref{fig:GPT-snorm-hist} also shows some unusually large Spectral Norms.
 Upon examination, e.g., from Figure \ref{fig:gpt-snorm-layer} (below), we see that these correspond to the first embedding layer(s).
@@
 We discuss this further in Appendix~\ref{sxn:appendix}.
-Here, we do not include them in our computed average metrics in Table~\ref{table:nlp}, and we do not include them in the histogram plot in Figure~\ref{fig:GPT-snorm-hist}.
+Here, we do not include them in our computed average metrics in Table~\ref{table:nlp}, and we do not include them in the histogram plot in Figure~\ref{fig:GPT-snorm-hist}.)
 

### 2020-02-16 submission/all_models.tex
some typos/tweaks, a few more to come

@@
 
-We choose ordinary linear regression (OLS) to quantity the relationship between quality metrics (computed with the \emph{WeightWatcher} tool) 
+We choose ordinary least squares (OLS) regression to quantify the relationship between quality metrics (computed with the \emph{WeightWatcher} tool) 
 and the reported test error and/or accuracy metrics.

### 2020-02-16 submission/background.tex
last par, whitespace, misc stuff
pass to convert to TR format, e.g, whitespace, etc.
some typos/tweaks, a few more to come

@@
+%KDD% \vspace{-1mm}
 \section{Background and Related Work}
 \label{sxn:background}
+%\vspace{-1mm}
 
@@
 
+%KDD% \vspace{-1mm}
 \paragraph{Statistical mechanics theory for DNNs.}
@@
 We \emph{use} statistical mechanics in a broader sense, drawing upon techniques from quantitative finance and random matrix theory.  % (RMT).
-Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statisical mechanics of heavy tailed and strongly correlated systems.
+Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statistical mechanics of heavy tailed and strongly correlated systems.
 % Probably best not to cite at this point %\charles{, and current work (in progress)\cite{blog}}
 
+%KDD% \vspace{-1mm}
 \paragraph{Norm-based capacity control theory.}

### 2020-02-16 submission/conclusion.tex
pass to convert to TR format, e.g, whitespace, etc.
some typos/tweaks, a few more to come

@@
-\vspace{-1mm}
+%KDD% \vspace{-1mm}
 \section{Conclusion}
@@
 Again, though, our results are the first to demonstrate this.
-It is also gratifying that this approach can be used to provide fine-scale insight (such as rationalizing the flow of correlations or the collapse of size scale) throughout a network.
+It is also gratifying that our approach can be used to provide fine-scale insight (such as rationalizing the flow of correlations or the collapse of size scale) throughout a network.
 
@@
 A drawback of the former approach is that it typically makes very strong assumptions on the data, and the strength of those assumptions can limit the practical applicability of the theory.
-Much of the work on the theory of DNNs focuses on the former type of theory.
+Nearly all of the work on the theory of DNNs focuses on the former type of theory.
 Our approach focuses on the latter type of theory.

### 2020-02-16 submission/cv_models.tex
some typos/tweaks, a few more to come

@@
 Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best.
-These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among
-a series of ``good-better-best'' models, with PL-based metrics performing somewhat (i.e., quantitatively) better.
+These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of ``good-better-best'' models, with PL-based metrics performing somewhat (i.e., quantitatively) better.
 
@@
 See Figure~\ref{fig:3models-alpha-layers}, which plots $\alpha$ for each layer (the first layer corresponds to data, the last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (no BN), ResNet, and DenseNet series.
-(Again, much more detailed set of plots is available at our repo; but note that the corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than the results we present here.)
+(Again, a much more detailed set of plots is available at our repo; but note that the corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than the results we present here.)
 
@@
 The reported baseline test accuracies (Top1$=91.45$ and Top5$=99.75$) are better than the reported fine-tuned test accuracies (Top1$=91.02$ and Top5$=99.67$).
-Because the baseline accuracy is greater,  the previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy}) suggests that the baseline Spectral Norms should be \emph{smaller} on average than the fine-tuned ones.
+Because the baseline accuracy is greater,  the previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy}) suggest that the baseline Spectral Norms should be \emph{smaller} on average than the fine-tuned ones.
 \emph{The oppopsite is observed.}
@@
 %%has the unusual effect of increasing
-causes the norms of the $\mathbf{W}$ pre-activation maps for two Conv2D layers to spuriously increase.
-This is difficult to diagnose by analyzing training/test curves, but it is easy to diagnose with our approach.
+causes the norms of the $\mathbf{W}$ pre-activation maps for two Conv2D layers to increase spuriously.
+This is difficult to diagnose by analyzing training/test curves, but it is easy to diagnose with our~approach.

### 2020-02-16 submission/introduction.tex
pass to convert to TR format, e.g, whitespace, etc.
included new repo ref
some typos/tweaks, a few more to come

--- 2020-02-12 submission/introduction.tex
@@
+%KDD% \vspace{-1mm}
 \section{Introduction}
 \label{sxn:intro}
+%\vspace{-1mm}
 
@@
 
+%KDD% \vspace{-1mm}
 \paragraph{The WeightWatcher Tool.}
@@
 All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
-To be fully reproducible, we only examine publically-available, pretrained models, and we also provide all Jupyter and Google Colab notebooks used in an accompanying github repository~\cite{kdd20_sub_repo}.
+To be fully reproducible, we only examine publicly-available, pretrained models, and we also provide all Jupyter and Google Colab notebooks used in an accompanying github repository~\cite{kdd20_sub_repo}.
 See Appendix~\ref{sxn:appendix} for details on how to reproduce all results.
 
+%KDD% \vspace{-1mm}
 \paragraph{Organization of this paper.}

### 2020-02-16 submission/methods.tex
some typos/tweaks, a few more to come

@@
 nor have we re-evaluated the test accuracies.
-We expect that most well-trained, production-quality models will employ one or more forms of on regularization, such as Batch Normalization (BN), Dropout, etc., and many will also contain additional structure such as Skip Connections, etc.
+We expect that most well-trained, production-quality models will employ one or more forms of regularization, such as Batch Normalization (BN), Dropout, etc., and many will also contain additional structure such as Skip Connections, etc.
 Here, we will ignore these details, and will focus only on the pretrained layer weight matrices $\mathbf{W}_{l}$.
@@
 The best performing empirical quality metrics depend on the norms and/or spectral properties of each weight matrix,
-$\mathbf{W}$ and/or, equivalently, it's \emph{Empirical Correlation Matrix}: $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$ .%
+$\mathbf{W}$, and/or, equivalently, it's \emph{Empirical Correlation Matrix}: $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.%
 
@@
 Here, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the $\mathbf{X}$, and $\lambda_{max}$ is the maximum eigenvalue.
-Recall that the eigenvalues are square of the singular values $\sigma_{i}$ of $\mathbf{W}$: $\lambda_{i}=\sigma^{2}_{i}$.
+Recall that the eigenvalues are squares of the singular values $\sigma_{i}$ of $\mathbf{W}$: $\lambda_{i}=\sigma^{2}_{i}$.
 Also, note that we do \emph{not} normalize $\mathbf{X}$ by $1/N$; see Appendix~\ref{sxn:appendix} for a discussion of this issue.
@@
 For norm-based metrics, we use the average of the log norm, and to the appropriate power.
-Informallly, this amounts to assuming that the layer weight matrices are statistically independent, in which case we can estimate the model complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm (which resembles a data dependent VC complexity),
+Informally, this amounts to assuming that the layer weight matrices are statistically independent, in which case we can estimate the model complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm (which resembles a data dependent VC complexity),
 \begin{equation}

### 2020-02-16 submission/nlp_models.tex
some typos/tweaks, a few more to come

@@
 By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously,
-or what we call a series of ``good-better-best'' models.
+an example of what we call a series of ``good-better-best'' models.
 
@@
 %
-The OpenAI GPT2 (English) models are: GPT2-small, GPT2-medium, GPT2-large, and GPT2-xl, having $12, 24, 36, \text{and } 48$ layers, respectively, with increasingly larger weight matrices.
+The OpenAI GPT2 (English) models are: GPT2-small, GPT2-medium, GPT2-large, and GPT2-xl, having $12$, $24$, $36$, and $48$ layers, respectively, with increasingly larger weight matrices.
 %The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
@@
 Indeed, we expect that a poorly-trained model will lack good (i.e., small $\alpha$) PL behavior in many/most layers.
-On the other hand, as expected, the newer improved GPT2-small model has, on average, smaller $\alpha$ values than the older GPT, with all $\alpha\le6$ and with smaller mean/median $\alpha$, and it also has far fewer unusually-large outlying $\alpha$ values than GPT.
+On the other hand, as expected, the newer improved GPT2-small model has, on average, smaller $\alpha$ values than the older GPT, with all $\alpha\le6$ and with smaller mean/median $\alpha$.
+It also has far fewer unusually-large outlying $\alpha$ values than GPT.
 From this (and other results not shown), we see that $\alpha$ provides a good quality metric for comparing these two models, the ``bad'' GPT versus the ``good'' GPT2-small.
@@
 Indeed, because there are so many anonymously small Spectral norms, it appears that the GPT model may be exhibiting a kind of \emph{Scale Collapse}, like that observed in the distilled CV models (in Figure~\ref{fig:resnet204D5L}).
-This is important because it demonstrates that, while the Spectral norm may correlate well with predicted test error, it is \emph{not} a good indicator of the overall model quality.
+This is important because it demonstrates that, while the Spectral (or Frobenius) norm may correlate well with predicted test error, it is \emph{not} a good indicator of the overall model quality.
 It can mispredict good-versus-bad questions in ways not seen with PL-based metrics.
@@
 In contrast, the well-trained GPT2-small model shows consistant and stable patterns, again with one stable $\alpha\sim 3.5$ (and below the GPT trend), and the other only slightly trending up, with $\alpha\le 6$.
-The scale-invariant $\alpha$ metric lets us identify potentially poorly-trained models; and these results show that the Correlation Flow differs significantly between GPT and GPT2-small (with the better GPT2-small looking more like the better ResNet-1K from Figure~\ref{fig:resnet-alpha-layer}).
+The scale-invariant $\alpha$ metric lets us identify potentially poorly-trained models.
+These results show that the Correlation Flow differs significantly between GPT and GPT2-small (with the better GPT2-small looking more like the better ResNet-1K from Figure~\ref{fig:resnet-alpha-layer}).
 
@@
     \caption{PL exponents ($\alpha$) (in (a)) and Log Spectral Norms ($\log\Vert\mathbf{W}\Vert_{\infty}$) (in (b)) for weight matrices from the OpenAI GPT and GPT2-small pretrained models.
-(Note that the Y axes on each plot are different.)
+(Note that the quantities being shown on each Y axis are different.)
              In the text, this is interpreted in terms of \emph{Correlation Flow} and \emph{Scale~Collapse}.
@@
 In general, as we move from GPT2-medium to GPT2-xl, histograms for both $\alpha$ exponents and the Log Norm metrics downshift from larger to smaller values. 
-For example, see Figure \ref{fig:gpt2-histograms}, which shows the histograms over the layer weight matrics for fitted PL exponent ($\alpha$) and the Log Alpha Norm ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$)~metric.
+For example, see Figure \ref{fig:gpt2-histograms}, which shows the histograms over the layer weight matrices for fitted PL exponent ($\alpha$) and the Log Alpha Norm ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$)~metric.
 

### 2020-02-16 submission/cv_models.tex
v1.4

@@
     }
-    \caption{Comparison of Average Log Norm and Weighted Alpha quality metrics versus reported test accuracy for pretrained VGG models (with and without BN), trained on ImageNet, available in pyTorch (v1.x).
+    \caption{Comparison of Average Log Norm and Weighted Alpha quality metrics versus reported test accuracy for pretrained VGG models (with and without BN), trained on ImageNet, available in pyTorch (v1.4).
 Metrics fit by linear regression, RMSE reported.

### 2020-02-16 submission/abstract_file.tex
full pass, fixed lots of typos, etc.; semifinal, nearly done

@@
 We also find that PL-based metrics do much better---quantitatively better at discriminating among a series of ``good-better-best'' models, and qualitatively better at discriminating ``good-versus-bad'' models.
-PL-based metrics can also be used to characterize fine-scale properties of these models, and we introduce the layer-wise \emph{Correlation Flow} as new quality assesment.
-We show how poorly-trained (and/or poorly fine-tuned) models may exhbit both \emph{Scale Collapse} and unsually large PL exponents, $\alpha\gg 6$, in particular for recent NLP models.
+PL-based metrics can also be used to characterize fine-scale properties of these models, and we introduce the layer-wise \emph{Correlation Flow} as new quality assessment.
+We show how poorly-trained (and/or poorly fine-tuned) models may exhibit both \emph{Scale Collapse} and unusually large PL exponents, $\alpha\gg 6$, in particular for recent NLP models.
 Our techniques, as implemented in the \emph{WeightWatcher} tool, can be used to identify when a pretrained DNN has problems that can not be detected simply by examining training/test accuracies.

### 2020-02-16 submission/appendix.tex
full pass, fixed lots of typos, etc.; semifinal, nearly done

@@
 In this appendix, we provide more details on several issues that are important for the reproducibility of our results.
+%
+All of our computations were performed with the \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
+More details and more results are available in an accompanying github repository~\cite{kdd20_sub_repo}.
 
@@
 \item run SVD on each pre-Activation Map $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values;
-\item stack the maps into a single matrix of, say, dimension $((k\times k\times out)\times in)$, run SVD to get $in$ singular values;
-\item compute the 2D Fourier Transform (FFT) for each of the $(in, out)$ pairs, and run SVD on the Fourier coeffients~\cite{CNNSVD}, leading to $\sim(k\times in\times out)$ non-zero singular values.
+\item stack the maps into a single matrix of, say, dimension $((k\times k\times out)\times in)$, and run SVD to get $in$ singular values;
+\item compute the 2D Fourier Transform (FFT) for each of the $(in, out)$ pairs, and run SVD on the Fourier coefficients~\cite{CNNSVD}, leading to $\sim(k\times in\times out)$ non-zero singular values.
 \end{enumerate}
@@
 For example, in GPT, for most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$, but in the first embedding layer, the maximum eigenvalue is of order $N$ (the number of words in the embedding), or $\lambda_{max}\sim\mathcal{O}(10^{5})$.
-For GPT and GPT2, we treat all layers as-is (although one may want to normalize the first 2 layers by  $\mathbf{X}$ by $1/N$, or to treat them as an outlier).
+For GPT and GPT2, we treat all layers as-is (although one may want to normalize the first 2 layers $\mathbf{X}$ by $1/N$, or to treat them as outliers).
 
@@
 
-We provide a github repository for this paper that includes Jupyter notebooks that fully reproduce all results (as well as many other results).
-All results have been produced using the \emph{WeightWatcher} tool (v0.2.7).
+We provide a github repository for this paper that includes Jupyter notebooks that fully reproduce all results (as well as many other results)~\cite{kdd20_sub_repo}.
+All results have been produced using the \emph{WeightWatcher} tool (v0.2.7)~\cite{weightwatcher_package}.
 The ImageNet and OpenAI GPT pretrained models are provided in the current 
-pyTorch~\cite{pytorch} and Huggingface~\cite{huggingface} distributions, as specified in the \texttt{requirements.txt} file.
+pyTorch~\cite{pytorch} and Huggingface~\cite{huggingface} distributions, as specified in the \texttt{requirements.txt}~file.
 
@@
 In the \texttt{distiller} folder of our github repo, 
-we provide the orginal Jupyter Notebooks, which use the Intel \texttt{distiller} framework.%
-\footnote{\url{https://nervanasystems.github.io/distiller}} 
+we provide the original Jupyter Notebooks, which use the Intel \texttt{distiller} framework~\cite{distiller}.   % \footnote{\url{https://nervanasystems.github.io/distiller}} 
 Figure~\ref{fig:resnet204D5L} is from the  \texttt{``...-Distiller-ResNet20.ipynb''} notebook (see Table~\ref{table:notebooks}).  
@@
 \begin{center}
-\begin{tabular}{|p{2in}|c|}
+%KDD% \begin{tabular}{|p{2in}|c|}
+\begin{tabular}{|p{3in}|c|}
 \hline
@@
              %\charles{Preliminary charts:} 
-             PL exponent $\alpha$ vs. reported Top1 Test Accuracies for pretrained DNNs available for five different data sets.
+             PL exponent $\alpha$ versus reported Top1 Test Accuracies for pretrained DNNs available for five different data sets.
             }
@@
 \begin{table}[t]
-\footnotesize  % \small
+%KDD% \footnotesize  % \small
+\scriptsize
 \begin{center}
@@
 \end{center}
+%KDD% 
+\vspace{-5mm}
 \caption{$MSE$ Results for all CV model regressions. }
@@
 \end{center}
+%KDD% 
+\vspace{-5mm}
 \caption{$R^{2}$ Results for all CV model regressions. }

### 2020-02-16 submission/background.tex
full pass, fixed lots of typos, etc.; semifinal, nearly done

@@
 We \emph{use} statistical mechanics in a broader sense, drawing upon techniques from quantitative finance and random matrix theory.  % (RMT).
-Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statistical mechanics of heavy tailed and strongly correlated systems.
+Thus, much more relevant for our methodological approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statistical mechanics of heavy tailed and strongly correlated systems.
 % Probably best not to cite at this point %\charles{, and current work (in progress)\cite{blog}}
@@
 %% \paragraph{Practical problems need a practical theory.}
-%% \nred{Exiting ML theory does not address many practical AI problems that our techiques directly target.}
+%% \nred{Exiting ML theory does not address many practical AI problems that our techniques directly target.}
 %% %There are many very practical problems in ML that are poorly addressed by existing theory and that either motivated our work or should be addressable by our techniques.
@@
 %% \item
-%% \textbf{Cost of acquring labeled data.}
+%% \textbf{Cost of acquiring labeled data.}
 %% \michael{XXX.  Put in two or at most three sentences.}
@@
 %% \charles{see comments in tex file}\michael{Let's touch base, not sure what to do.}
-%% % we adress this above
-%% %Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text or realiztic chatbots, developing self driving cars, or just clusterting user profiles.
+%% % we address this above
+%% %Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text or realistic chatbots, developing self driving cars, or just clustering user profiles.
 %% %\footnote{For example, current chatbots use perplexity as a proxy for passing a Turing test.}

### 2020-02-16 submission/conclusion.tex
full pass, fixed lots of typos, etc.; semifinal, nearly done

@@
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \section{Conclusion}
@@
 Our results are the first to demonstrate that they can be used for this important practical problem.
-That PL-based metrics perform better (than norm-based metrics) should not be surprising---at least to those familiar with the statisical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} (since our use of PL exponents is designed to capture the idea that well-trained models capture correlations over many size scales in the data).
+That PL-based metrics perform better (than norm-based metrics) should not be surprising---at least to those familiar with the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} (since our use of PL exponents is designed to capture the idea that well-trained models capture correlations over many size scales in the data).
 Again, though, our results are the first to demonstrate this.
@@
 Our approach focuses on the latter type of theory.
-Our results, which are based on our \emph{use} of sophisticated statistical mechanics theory to solve important practical DNN problems, suggests that our approach should be of interest more generally for those interested in developing a practical DNN theory.
+Our results, which are based on \emph{using} sophisticated statistical mechanics theory and \emph{solving} important practical DNN problems, suggests that the latter approach should be of interest more generally for those interested in developing a \emph{practical DNN~theory}.

### 2020-02-16 submission/cv_models.tex
full pass, fixed lots of typos, etc.; semifinal, nearly done

@@
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \section{Comparison of CV models}
@@
 
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \paragraph{Average Quality Metrics versus Reported Test Accuracies.}
@@
 To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BN), and it plots the four quality metrics versus the reported test accuracies~\cite{pytorch},%
-\footnote{That is, these test accuracies have been previously reported and made publicly-avaiable by others.
+\footnote{That is, these test accuracies have been previously reported and made publicly-available by others.
 We take them as given, and we do not attempt to reproduce/verify them, since we do not permit ourselves any access to training/test data.}
 as well as a basic linear regression line. 
-%with the Root Mean Squared Error (RMSE) shown, \michael{XXX.  WHERE IS RMSE.  IN THAT TABLE?}
 All four metrics correlate quite well with the reported Top1 accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalization (i.e., greater accuracy, lower error). 
@@
     %}
-    \caption{Comparison of Avergage $\alpha$-Norm quality metric ($\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle$) versus reported Top1 test accuracy for the ResNet and ResNet-1K pretrained (pyTorch) models.
+    \caption{Comparison of Average $\alpha$-Norm quality metric ($\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle$) versus reported Top1 test accuracy for the ResNet and ResNet-1K pretrained (pyTorch) models.
 }
@@
 \end{figure}
-
-%MM% \begin{figure}
-%MM% \centering
-%MM% \begin{minipage}[b]{.22\textwidth}
-%MM% %<Code for the first figure>
-%MM% XXX
-%MM% \caption{Caption}\label{label-a}
-%MM% \end{minipage}\qquad
-%MM% \begin{minipage}[b]{.22\textwidth}
-%MM% %<Code for the second figure>
-%MM% XXX
-%MM% \caption{Caption}\label{label-b}
-%MM% \end{minipage}
-%MM% \end{figure}
 
@@
 In our larger analysis, in Section~\ref{sxn:all_cv_models}, we will only include series with 5 or more models.
-%For completemnss and reproducibility here, the DenseNet Log $\alpha$-Norm plot is given in Appendix~\ref{sxn:appendix}.
+%For completeness and reproducibility here, the DenseNet Log $\alpha$-Norm plot is given in Appendix~\ref{sxn:appendix}.
 (Note that these and many other such plots can be seen on our publicly-available~repo.)
 
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \paragraph{Variation in Data Set Size.}
@@
 
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \paragraph{Layer Analysis: Metrics as a Function of Depth.}
@@
 
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \paragraph{Comparison of VGG, ResNet, and DenseNet Architectures.}
@@
 ResNet greatly improved on VGG by replacing the large FC layers, shrinking the Conv2D blocks, and introducing \emph{residual connections}.
-This optimized approach allows for greater accurary with far fewer parameters (and GPU memory requirements), and
+This optimized approach allows for greater accuracy with far fewer parameters (and GPU memory requirements), and
 ResNet models of up to 1000 layers have been trained~\cite{resnet1000}.
 
-We conjuecture that the effeciency and effectivness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and strongly optimized.
+We conjecture that the efficiency and effectiveness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and strongly optimized.
 Contrast this with the DenseNet models, which contains many connections between every layer.
@@
 
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \paragraph{Correlation Flow.}
@@
 More generally, we can understand the results presented in Figure~\ref{fig:3models-alpha-layers} in terms of what we will call the \emph{Correlation Flow} of the model. 
-Recall that the average Log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on ideas from the statisical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
+Recall that the average Log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on ideas from the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
 There, one expects the weight matrices of well-trained DNNs will exhibit correlations over many size scales. 
@@
 %, but the presence of many adjacent layers with relatively good PL fits \emph{suggests} that those adjacent layers ``impedance match'' well and that the correlations that are captured by one layer propagate well to subsequent layers.
-Informally, one would expect a DNN model to perform well when it fascilitates the propagation of information/features across layers.
+Informally, one would expect a DNN model to perform well when it facilitates the propagation of information/features across layers.
 Previous work argues this by computing the gradients over the input data.
@@
 
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \paragraph{Scale Collapse; or How Distillation May Break Models.}
@@
 More generally, do fitted $\alpha$ values contain information not captured by norms? 
-In examining hundreds of pretrained models, we have found several anamolies that demonsrate the power of our approach.
+In examining hundreds of pretrained models, we have found several anomalies that demonstrate the power of our approach.
 In particular, to show that $\alpha$ does capture something different, consider the following example, which looks at a compressed/distilled DNN model~\cite{CWZZ17_TR}.
@@
 Because the baseline accuracy is greater,  the previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy}) suggest that the baseline Spectral Norms should be \emph{smaller} on average than the fine-tuned ones.
-\emph{The oppopsite is observed.}
+\emph{The opposite is observed.}
 %\michael{Do we have something backwards here: where we observed that the average norm to increase with decreasing test error (as it ``should'' not), whereas the average PL exponent $\alpha$ decreases (as ``expected'' from HT-SR Theory).
@@
 
-That being said, Figure~\ref{fig:resnet204Dalpha} also depcits two very large $\alpha\gg 6$ values for the baseline, but not for the fine-tuned, model.
+That being said, Figure~\ref{fig:resnet204Dalpha} also depicts two very large $\alpha\gg 6$ values for the baseline, but not for the fine-tuned, model.
 This suggests the baseline model has at least two over-parameterized/under-trained layers, and that the distillation method does, in fact, improve the fine-tuned model by compressing these layers.

### 2020-02-16 submission/introduction.tex
full pass, fixed lots of typos, etc.; semifinal, nearly done

@@
 Here, one may not have access to either the training data or the testing data.
-Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it ``as is,'' or to fine-tune and/or compress it and then use it.
+Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it as-is, or to fine-tune and/or compress it and then use it.
 
@@
 Indeed, high quality, labeled data can be very expensive to acquire, and this cost can make or break a project.
-Methods that are developed and evaluated on any well-defined publicly-available coprus of data, no matter how large or diverse or interesting, are clearly not going to be well-suited to address problems such as this.
+Methods that are developed and evaluated on any well-defined publicly-available corpus of data, no matter how large or diverse or interesting, are clearly not going to be well-suited to address problems such as this.
 It is of great practical interest to have metrics to evaluate the quality of a trained model---in the absence of training/testing data and without any detailed knowledge of the training/testing process.
@@
 
-The \emph{quality metrics} we consider are based on the spectral properties of the layer weight matrics.
+The \emph{quality metrics} we consider are based on the spectral properties of the layer weight matrices.
 They are based on norms of weight matrices (such norms have been used in traditional statistical learning theory to bound capacity and construct regularizers) and/or parameters of power law (PL) fits of the eigenvalues of weight matrices (such PL fits are based on statistical mechanics approaches to DNNs).
 Note that, while we use traditional norm-based and PL-based metrics, our goals are not the traditional goals.
-Unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., as with hyperparmeter optimization).%
+Unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., as with hyperparameter optimization).%
 \footnote{One could of course use these techniques to improve training, and we have been asked about that, but we are not interested in that here.
@@
 Then, in Section~\ref{sxn:nlp}, we look at several variations of a popular NLP DNN architecture (the OpenAI GPT and GPT2 models); and we show how model quality and properties vary between several variants of GPT and GPT2, including how metrics behave similarly and differently.
-Then, in Section \ref{sxn:all_cv_models}, we present results based on an analysis of hundreds of pretraind DNN models, showing how well each metric predicts the reported test accuracies, and how the PL-based metrics perform remarkably well.
+Then, in Section \ref{sxn:all_cv_models}, we present results based on an analysis of hundreds of pretrained DNN models, showing how well each metric predicts the reported test accuracies, and how the PL-based metrics perform remarkably well.
 Finally, in Section~\ref{sxn:conc}, we provide a brief discussion and conclusion.

### 2020-02-16 submission/methods.tex
full pass, fixed lots of typos, etc.; semifinal, nearly done

@@
 The models have been trained and evaluated on labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using standard techniques.  
-The pretained pytroch model files are publicly-available, and the test accuracies have been reported online.  
+The pretrained pytroch model files are publicly-available, and the test accuracies have been reported online.  
 In this study, we do not have access to this data, and we have not trained any of the models ourselves,
@@
 \end{equation}
-where $L$ is the total nuumber of layer weight matrices.
+where $L$ is the total number of layer weight matrices.
 The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a limited set of models.

### 2020-02-16 submission/nlp_models.tex
full pass, fixed lots of typos, etc.; semifinal, nearly done

@@
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \section{Comparison of NLP Models}
@@
 %
-The Transformer architectures consist of blocks of so-called Attention layers, containing two large, Feed Forward (Linear) weight matrics~\cite{Attn2017}. 
+The Transformer architectures consist of blocks of so-called Attention layers, containing two large, Feed Forward (Linear) weight matrices~\cite{Attn2017}. 
 In contrast to smaller pre-Activation maps arising in Cond2D layers, Attention matrices are significantly larger.
 In general, we have found that they have larger PL exponents $\alpha$.
-Based on HT-SR Theory, 
-%(in particular, the interpretation of values of $\alpha \sim 2$ as modeling systems with good correations over many size scales~\cite{BouchaudPotters03, SornetteBook}), 
+Based on HT-SR Theory
+(in particular, the interpretation of values of $\alpha \sim 2$ as modeling systems with good correlations over many size scales~\cite{BouchaudPotters03, SornetteBook}), 
 this suggests that these models fail to capture successfully many of the correlations in the data (relative to their size) and thus are substantially \emph{under}-trained.
@@
 modern NLP models have larger weight matrices and display different spectral properties.
-%, both in terms of their architecture (e.g., wheter there are convolutions) as well as how they are used (e.g., whether one uses prediction error or perplexity).
+%, both in terms of their architecture (e.g., whether there are convolutions) as well as how they are used (e.g., whether one uses prediction error or perplexity).
 Thus, they provide a very different test for our empirical quality metrics.
@@
 While norm-based metrics perform reasonably well on well-trained NLP models, they often behave anomalously on poorly-trained models.
-Indeed, for such ``bad'' models, weight matrics may display rank collapse, decreased Frobeinus mass, or unsually small Spectral norms.
+Indeed, for such ``bad'' models, weight matrices may display rank collapse, decreased Frobenius mass, or unusually small Spectral norms.
 (This may be misinterpreted as ``smaller is better.'')
-In contrast, PL-based metrics, including the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $) display consistant behavior, even on poorly trained models.
-Indeed, we can use these metrics to help indentify when archiectures need repair and when more and/or better data are needed.
+In contrast, PL-based metrics, including the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $) display consistent behavior, even on poorly trained models.
+Indeed, we can use these metrics to help identify when architectures need repair and when more and/or better data are needed.
 
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \paragraph{What do large values of $\alpha$ mean?}
@@
 
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \paragraph{OpenAI GPT Models.}
@@
 %
-By comparing the poorly-trained (i.e., ``bad'') GPT to the well-trained (i.e., ``good'') GPT2-small, we can indentify empirical indicators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
+By comparing the poorly-trained (i.e., ``bad'') GPT to the well-trained (i.e., ``good'') GPT2-small, we can identify empirical indicators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
 By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously,
@@
 GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ layer Weight Matrices, $\mathbf{W}$.
-Each Block has 2 components, the Self Attention (attn) and the Projection (proj) matrics.  
+Each Block has 2 components, the Self Attention (attn) and the Projection (proj) matrices.  
 The self-attention  matrices are larger, of dimension ($2304\times 768$) or ($3072\times 768$).
@@
 \end{center}
-\caption{Average value for the average Log Norm and Weighted Alpha metrics for pretrainnd OpenAI GPT and GPT2 models.
+\caption{Average value for the average Log Norm and Weighted Alpha metrics for pretrained OpenAI GPT and GPT2 models.
 Column \# refers to number of layers treated.
@@
 
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \paragraph{Average Quality Metrics for GPT and GPT2.}
@@
 These two histograms are very different.
-The older deficient GPT has numerous unsually large $\alpha$ exponents---meaning they are not really well-described by a PL fit.
+The older deficient GPT has numerous unusually large $\alpha$ exponents---meaning they are not really well-described by a PL fit.
 Indeed, we expect that a poorly-trained model will lack good (i.e., small $\alpha$) PL behavior in many/most layers.
@@
 
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \paragraph{Scale Collapse in Poorly Trained Models.}
@@
 
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \paragraph{Layer Analysis: Correlation Flow and Scale Collapse in GPT and GPT2.}
@@
 The deficient GPT model displays two trends in $\alpha$, one stable with $\alpha\sim 4$, and one increasing with layer id, with $\alpha$ reaching as high as $12$.
-In contrast, the well-trained GPT2-small model shows consistant and stable patterns, again with one stable $\alpha\sim 3.5$ (and below the GPT trend), and the other only slightly trending up, with $\alpha\le 6$.
+In contrast, the well-trained GPT2-small model shows consistent and stable patterns, again with one stable $\alpha\sim 3.5$ (and below the GPT trend), and the other only slightly trending up, with $\alpha\le 6$.
 The scale-invariant $\alpha$ metric lets us identify potentially poorly-trained models.
@@
 
-%KDD% \vspace{-1mm}
+%KDD% 
+\vspace{-1mm}
 \paragraph{GPT2: medium, large, xl.}
@@
 
-We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticible between the differing good-better-best GTP2 models than between the good-versus-bad GPT and GPT2-small models.
+We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticeable between the differing good-better-best GTP2 models than between the good-versus-bad GPT and GPT2-small models.
 Unlike GPT, however, the layer Log Alpha Norms behave more as expected for GPT2 layers, with the larger models consistently having smaller norms. 
@@
               and Log Alpha Norm ($\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$) for weight matrices from models of different sizes in the GPT2 architecture series.
-(Plots omit the first 2 (embedding) layers, because they are normalized differently giving anamolously large values.)
+(Plots omit the first 2 (embedding) layers, because they are normalized differently giving anomalously large values.)
              }

### 2020-02-16 submission/empirical2_TRfmt.tex
cleaning up
full pass, fixed lots of typos, etc.; semifinal, nearly done
minor changes

@@
 
+\vspace{-2mm}
 \noindent
 \paragraph{Acknowledgements.}
-MWM would like to acknowledge ARO, DARPA, NSF, and ONR for providing partial support of this work.
+MWM would like to acknowledge ARO, DARPA, NSF, and ONR as well as the UC Berkeley BDD project and a gift from Intel for providing partial support of this work.
 We would also like to thank Amir Khosrowshahi and colleagues at Intel for helpful discussion regarding the Group Regularization distillation technique.
@@
 %\bibliography{gen_gap}
-%\bibliography{dnns}
-\bibliography{dnns,gen_gap}
+\bibliography{dnns}
+%\bibliography{dnns,gen_gap}
 }

### 2020-02-17 submission/FILES_TO_ARXIV_16feb20/T1/appendix.tex
files I uploaded to arXiv

--- 2020-02-16 submission/appendix.tex
@@
 In this appendix, we provide more details on several issues that are important for the reproducibility of our results.
-%
 All of our computations were performed with the \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
@@
 channels. When $w=h=k$, it gives $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\times out)$ each. 
-%
 We identify 3 different approaches for running SVD on a Conv2D layer:
@@
 Norm-based metrics, however, do depend strongly on the scale of the weight matrix---that is the point.
-%\nred{Indeed, early theoretical work by Bartlett suggests that the test accuracy depends strongly on the ``total size'' of the weight matrics.}
 To apply RMT, we usually define $\mathbf{X}$ with a $1/N$ normalization, assuming variance of $\sigma^{2}=1.0$.
-%\footnote{For Heavy Tailed theorems, one typically needs a normalization such as \nred{$1/N^{\alpha-1}$. check this}}
 Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with
@@
 
-\subsection{Reproducing Sections \ref{sxn:cv} and \ref{sxn:nlp} }   %% \subsection{Reproducing Sections 4 and 5}
+\subsection{Reproducing Sections \ref{sxn:cv} and \ref{sxn:nlp} }
 
@@
 In the \texttt{distiller} folder of our github repo, 
-we provide the original Jupyter Notebooks, which use the Intel \texttt{distiller} framework~\cite{distiller}.   % \footnote{\url{https://nervanasystems.github.io/distiller}} 
+we provide the original Jupyter Notebooks, which use the Intel \texttt{distiller} framework~\cite{distiller}.   
 Figure~\ref{fig:resnet204D5L} is from the  \texttt{``...-Distiller-ResNet20.ipynb''} notebook (see Table~\ref{table:notebooks}).  
@@
 We also remove the entire \texttt{cifar100} \texttt{ResNeXT} series, which is the only example to show no trends with the norm metrics.
-%
 The final datasets used are shown in Table~\ref{table:datasets}.
@@
 \begin{center}
-%KDD% \begin{tabular}{|p{2in}|c|}
 \begin{tabular}{|p{3in}|c|}
@@
     \subfigure[ImageNet 1K]{
-        %OLD% \includegraphics[width=2.5cm]{img/imagenet1k_alpha.png}
-        %KDD% \includegraphics[width=2.5cm]{img/all-imagenet-1k_alpha.png}
         \includegraphics[width=4.9cm]{img/all-imagenet-1k_alpha.png}
@@
     }
-    %KDD% 
     \qquad
     \subfigure[ CIFAR 10 ]{
-        %OLD% \includegraphics[width=2.5cm]{img/cifar10_alpha.png}
-        %KDD% \includegraphics[width=2.5cm]{img/all-cifar-10_alpha.png}
         \includegraphics[width=4.9cm]{img/all-cifar-10_alpha.png}
@@
     }
-    %KDD% 
     \qquad
     \subfigure[ CIFAR 100 ]{
-        %OLD% \includegraphics[width=2.5cm]{img/cifar100_alpha.png}
-        %KDD% \includegraphics[width=2.5cm]{img/all-cifar-100_alpha.png}
         \includegraphics[width=4.9cm]{img/all-cifar-100_alpha.png}
@@
     }
-    %KDD% 
     \qquad
     \subfigure[ SVHN ]{
-        %OLD% \includegraphics[width=2.5cm]{img/svhn_alpha.png}
-        %KDD% \includegraphics[width=2.5cm]{img/all-svhn_alpha.png}
         \includegraphics[width=4.9cm]{img/all-svhn_alpha.png}
@@
     }
-    %KDD% 
     \qquad
     \subfigure[ CUB 200 ]{
-        %OLD% \includegraphics[width=2.5cm]{img/cub200_alpha.png}
-        %KDD% \includegraphics[width=2.5cm]{img/all-cub-200-2011_alpha.png}
         \includegraphics[width=4.9cm]{img/all-cub-200-2011_alpha.png}
@@
     \caption{% 
-             %\charles{Preliminary charts:} 
              PL exponent $\alpha$ versus reported Top1 Test Accuracies for pretrained DNNs available for five different data sets.
@@
 \begin{table}[t]
-%KDD% \footnotesize  % \small
 \scriptsize
@@
 \end{center}
-%KDD% 
 \vspace{-5mm}

### 2020-02-17 submission/FILES_TO_ARXIV_16feb20/T1/cv_models.tex
files I uploaded to arXiv

--- 2020-02-16 submission/cv_models.tex
@@
-%KDD% 
 \vspace{-1mm}
@@
 \label{sxn:cv}
-%\vspace{-1mm}
 
@@
 
-%KDD% 
 \vspace{-1mm}
@@
         \includegraphics[width=4.9cm]{img/VGG_lognorm_accs.png}
-        %KDD% \includegraphics[width=4.0cm]{img/VGG_lognorm_accs.png}
         \label{fig:vgg-fnorm}
     }
-    %KDD% 
     \qquad
@@
         \includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
-        %KDD% \includegraphics[width=4.0cm]{img/VGG_spectralnorm_accs.png}
         \label{fig:vgg-snorm}
     }
-    %KDD% 
     \qquad
@@
         \includegraphics[width=4.9cm]{img/VGG_alpha_weighted_accs.png}
-        %KDD% \includegraphics[width=4.0cm]{img/VGG_alpha_weighted_accs.png}
         \label{fig:vgg-walpha}
     }
-    %KDD% 
     \qquad
@@
         \includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
-        %KDD% \includegraphics[width=4.0cm]{img/VGG_logpnorm_accs.png}
         \label{fig:vgg-pnorm}
@@
         \includegraphics[width=4.9cm]{img/ResNet_logpnorm_accs.png}
-        %KDD% \includegraphics[width=4.0cm]{img/ResNet_logpnorm_accs.png}
         \label{fig:resnet-accuracy}
     }
-    %KDD% 
     \qquad
@@
         \includegraphics[width=4.9cm]{img/ResNet-1K_logpnorm_accs.png}
-        %KDD% \includegraphics[width=4.0cm]{img/ResNet-1K_logpnorm_accs.png}
         \label{fig:resnet1k-accuracy}
     }
-    %\qquad
-    %\subfigure[ DenseNet, Log $\alpha$-Norm ]{
-    %    %\includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
-    %    \includegraphics[width=4.0cm]{img/DenseNet_logpnorm_accs.png}
-    %    \label{fig:densenet-accuracy}
-    %}
     \caption{Comparison of Average $\alpha$-Norm quality metric ($\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle$) versus reported Top1 test accuracy for the ResNet and ResNet-1K pretrained (pyTorch) models.
 }
-%, and DenseNet models. }The corresponding results for VGG are shown in Figure~\ref{fig:vgg-pnorm}.  }
     \label{fig:cv2-accuracy}
@@
 \begin{tabular}{|p{1in}|c|c|c|c|c|}
-%KDD% \begin{tabular}{|p{0.65in}|c|c|c|c|c|}
-%\begin{tabular}{|l|c|c|c|c|c|}
 \hline
@@
 \caption{RMSE (smaller is better) for linear fits of 
-         %Average Log Norm and Weighted Alpha 
          quality metrics to reported Top1 test error for pretrained models in each architecture series.
@@
 In our larger analysis, in Section~\ref{sxn:all_cv_models}, we will only include series with 5 or more models.
-%For completeness and reproducibility here, the DenseNet Log $\alpha$-Norm plot is given in Appendix~\ref{sxn:appendix}.
 (Note that these and many other such plots can be seen on our publicly-available~repo.)
 
-%KDD% 
 \vspace{-1mm}
@@
 See Figure~\ref{fig:cv2-accuracy}, which plots and compares the Log $\alpha$-Norm metric 
-%% MM: From Table 1 it is slightly worse than alpha-hat and we say that below %% --the best performing metric--
 for the full ResNet model, trained on the full ImageNet dataset, against the ResNet-1K model, which has been trained on a much smaller ImageNet-1K data set.
@@
 We can learn much more about a pretrained model by going beyond average values of quality metrics to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).
-% in the network.
-%The most interesting results are seen when we 
 For example, we can 
-plot (just) the PL exponent, $\alpha$, for each layer, %
-%\footnote{That is, here we consider just $\alpha$ for each layer, i.e., not $\hat{\alpha}$ or $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$.} 
+plot (just) the PL exponent, $\alpha$, for each layer, 
 as a function of depth.
-%
 See Figure~\ref{fig:3models-alpha-layers}, which plots $\alpha$ for each layer (the first layer corresponds to data, the last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (no BN), ResNet, and DenseNet series.
@@
         \includegraphics[width=4.9cm]{img/VGG_fnl_alpha_depth.png} 
-        %KDD% \includegraphics[width=4.0cm]{img/VGG_fnl_alpha_depth.png} 
                 \label{fig:vgg-alpha-layers}
     }
-    %KDD% 
     \qquad
-    \subfigure[ ResNet-1K ]{
+    \subfigure[ ResNet ]{
         \includegraphics[width=4.9cm]{img/ResNet_fnl_alpha_depth.png} 
-        %KDD% \includegraphics[width=4.0cm]{img/ResNet_fnl_alpha_depth.png} 
         \label{fig:resnet-alpha-layer}
     }
-    %KDD% 
     \qquad
@@
         \includegraphics[width=4.9cm]{img/DenseNet_fnl_alpha_depth.png} 
-        %KDD% \includegraphics[width=4.0cm]{img/DenseNet_fnl_alpha_depth.png} 
         \label{fig:densenet-alpha-layer}
     }    
-    %KDD% 
     \qquad
-    \subfigure[ ResNet-1K (overlaid) ]{
+    \subfigure[ ResNet (overlaid) ]{
         \includegraphics[width=4.9cm]{img/resnet_alpha_overlaid_depth.png} 
-        %KDD% \includegraphics[width=4.0cm]{img/resnet_alpha_overlaid_depth.png} 
         \label{fig:resnet_alpha_overlaid_depth}
@@
 However, as the ResNet-1K models get deeper, there is a wide range over which $\alpha$ values tend to remain quite small.
-%\michael{Can we say something about exceptions which are larger than for VGG or DenseNet.}
-%\charles{You have the same plots I do..what do you want to say >}
-%\michael{Are they small or different types of layers or something?}
-%\charles{Not that I am aware of.  Ill think on it}
 This is seen for other models in the ResNet-1K series, but it is most pronounced for the larger ResNet-1K (152) model, where $\alpha$ remains relatively stable at $\alpha\sim 2.0$, from the earliest layers all the way until we reach close to the final layers.
@@
 
-%Recall the differences---in particular, the number of residual connections---between the VGG, ResNet, and DenseNet architectures.
-%VGG, while a fine model in many ways, was limited by needing massive FC layers at the end of the model, requiring significant memory and computational resources, and it's poor conditioning led to problems with vanishing gradients.
-%The empirical manifestations of this (on weight matrices) are that fitted $\alpha$ values get much worse for deeper layers.
-
-%KDD% 
 \vspace{-1mm}
@@
 Much larger values $(\alpha\gg 6)$ may reflect poorer PL fits, whereas smaller values $(\alpha\sim 2)$, are associated with models that generalize better.
-%, but the presence of many adjacent layers with relatively good PL fits \emph{suggests} that those adjacent layers ``impedance match'' well and that the correlations that are captured by one layer propagate well to subsequent layers.
 Informally, one would expect a DNN model to perform well when it facilitates the propagation of information/features across layers.
@@
    \subfigure[$\lambda_{max}$ for ResNet20 layers]{
-      %KDD% \includegraphics[width=4.0cm]{img/resnet4d_maxev.png}
       \includegraphics[width=4.9cm]{img/resnet4d_maxev.png}
@@
    }
-   %KDD% 
    \qquad
    \subfigure[$\alpha$ for ResNet20 layers]{
-      %KDD% \includegraphics[width=4.0cm]{img/resnet4d_alphas.png}
       \includegraphics[width=4.9cm]{img/resnet4d_alphas.png}
@@
    }
-   \caption{%Correlation Flow.
+   \caption{%
             ResNet20, distilled with Group Regularization, as implemented in the \texttt{distiller} (4D\_regularized\_5Lremoved) pretrained models.  
@@
 The similarity between norm-based metrics and PL-based metrics suggests a question: is the Weighted Alpha metric just a variation of the more familiar norm-based metrics?  
-%%, or does it capture something different?
 More generally, do fitted $\alpha$ values contain information not captured by norms? 
@@
 \emph{The opposite is observed.}
-%\michael{Do we have something backwards here: where we observed that the average norm to increase with decreasing test error (as it ``should'' not), whereas the average PL exponent $\alpha$ decreases (as ``expected'' from HT-SR Theory).
-}
-%In both cases (Frobenius norm results not shown), we observe the opposite.
-%%\charles{WHy is this confusing? Call me...}
 Figure~\ref{fig:resnet204D5L} presents the Spectral Norm (here denoted $\log\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}$.%
@@
 On the other hand, the $\alpha$ values (in Figure~\ref{fig:resnet204Dalpha}) do not differ systematically between the baseline and fine-tuned models.
-%\michael{Some comment about big jump in norm; but fix previous backwards question first.}--SEE BELOW
 Also (not shown), the average (unweighted) baseline $\alpha$ is \emph{smaller} than the fine-tuned average (as predicted by HT-SR Theory, the basis of $\hat{\alpha}$).
@@
 The pretrained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the 
-VGG, ResNet, or DenseNet series.  %% we considered above.
+VGG, ResNet, or DenseNet series.  
 While norms make good regularizers for a single model, there is no reason \emph{a priori} to expect them correlate so well with test accuracies across different models.
@@
 is that the \texttt{distiller} Group Regularization technique 
-%%has the unusual effect of increasing
 causes the norms of the $\mathbf{W}$ pre-activation maps for two Conv2D layers to increase spuriously.

### 2020-02-17 submission/FILES_TO_ARXIV_16feb20/T1/nlp_models.tex
files I uploaded to arXiv

--- 2020-02-16 submission/nlp_models.tex
@@
 \label{sxn:nlp}
-%\vspace{-1mm}
 
 In this section, we examine empirical quality metrics described in Section~\ref{sxn:methods} for several NLP model architectures.
-%
-%In particular, 
 Within the past two years, nearly 100 open source, pretrained NLP DNNs based on the revolutionary Transformer architecture have emerged.
 These include variants of BERT, Transformer-XML, GPT, etc.
-%
 The Transformer architectures consist of blocks of so-called Attention layers, containing two large, Feed Forward (Linear) weight matrices~\cite{Attn2017}. 
@@
 this suggests that these models fail to capture successfully many of the correlations in the data (relative to their size) and thus are substantially \emph{under}-trained.
-%
 More generally, compared to the CV models of Section~\ref{sxn:cv},
 modern NLP models have larger weight matrices and display different spectral properties.
-%, both in terms of their architecture (e.g., whether there are convolutions) as well as how they are used (e.g., whether one uses prediction error or perplexity).
 Thus, they provide a very different test for our empirical quality metrics.
@@
 Later, OpenAI released a much improved model, GPT2-small, which has the same architecture and number of layers as GPT, but which has been trained on a larger and better data set (and with other changes), making it remarkably good at generating (near) human-quality fake text.
-%
 By comparing the poorly-trained (i.e., ``bad'') GPT to the well-trained (i.e., ``good'') GPT2-small, we can identify empirical indicators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
@@
 This gives $50$ large matrices.
-%
 Because GPT and GPT2 are trained on different data sets, the initial Embedding matrices differ in shape.
 GPT has an initial Token and Positional Embedding layers, of dimension $(40478\times 768)$ and $(512\times 768)$, respectively, whereas GPT2 has input Embeddings of shape $(50257\times 768)$ and $(1024\times 768)$, respectively.
-%
 The OpenAI GPT2 (English) models are: GPT2-small, GPT2-medium, GPT2-large, and GPT2-xl, having $12$, $24$, $36$, and $48$ layers, respectively, with increasingly larger weight matrices.
-%The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
 
@@
 \begin{tabular}{|p{1in}|c|c|c|c|c|}
-%KDD% \begin{tabular}{|p{0.67in}|c|c|c|c|c|}
 \hline
@@
 From this (and other results not shown), we see that $\alpha$ provides a good quality metric for comparing these two models, the ``bad'' GPT versus the ``good'' GPT2-small.
-%
 This should be contrasted with the behavior displayed by the Frobenius norm (not shown) and the Spectral~norm.
 
-%KDD% 
 \vspace{-1mm}
@@
     \subfigure[PL exponent ($\alpha$)]{
-        %KDD% \includegraphics[width=4.0cm]{img/GPT-alpha-hist.png}
         \includegraphics[width=5.5cm]{img/GPT-alpha-hist.png}
@@
        }
-    %KDD% 
     \qquad
     \subfigure[Log Spectral Norm ($\log\Vert\mathbf{W}\Vert_{\infty}$)]{
-        %KDD% \includegraphics[width=4.0cm]{img/GPT-snorm-hist.png}
         \includegraphics[width=5.5cm]{img/GPT-snorm-hist.png}
@@
 Upon examination, e.g., from Figure \ref{fig:gpt-snorm-layer} (below), we see that these correspond to the first embedding layer(s).
-%That is, they correspond to the change in normalization of the word embedding layer(s), and thus they can be ignored.
 These layers have a different effective normalization, and therefore a different scale.
-% \charles{We discuss this further in Appendix~\ref{sxn:appendix}.}
-% For example, in GPT, for most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$, but in the first embedding layer, the maximum is of order $N$ (the number of words in the embedding), or $\lambda_{max}\sim\mathcal{O}(10^{5})$.
-% For GPT and GPT2, we treat all layers as-is (although one may want to normalize the first 2 layers by $\mathbf{X}$ by $1/N$ or treat them as an outlier).
 We discuss this further in Appendix~\ref{sxn:appendix}.
@@
     \subfigure[PL exponent ($\alpha$)]{
-        %KDD% \includegraphics[width=4.0cm]{img/GPT-alpha-depth.png}
         \includegraphics[width=5.5cm]{img/GPT-alpha-depth.png}
@@
     }
-    %KDD% 
     \quad
     \subfigure[Log Spectral Norm ($\log\Vert\mathbf{W}\Vert_{\infty}$)]{
-        %KDD% \includegraphics[width=4.0cm]{img/GPT-snorm-depth.png}
         \includegraphics[width=5.5cm]{img/GPT-snorm-depth.png}
@@
     \subfigure[PL exponent ($\alpha$)]{
-        %KDD% \includegraphics[width=4.0cm]{img/GPT2*_fnl_alpha_hist.png}
-        \includegraphics[width=5.5cm]{img/GPT2*_fnl_alpha_hist.png}
+        \includegraphics[width=5.5cm]{img/GPT2xxx_fnl_alpha_hist.png}
         \label{fig:gpt2-alpha-hist}
     }
-    %KDD% 
     \quad
     \subfigure[Log Alpha Norm]{
-        %KDD% \includegraphics[width=4.0cm]{img/GPT2*_fnl_logpnorm_hist.png}
-        \includegraphics[width=5.5cm]{img/GPT2*_fnl_logpnorm_hist.png}
+        \includegraphics[width=5.5cm]{img/GPT2xxx_fnl_logpnorm_hist.png}
         \label{fig:gpt2-pnorm-hist}
@@
     \caption{Histogram of PL exponents ($\alpha$)
-             %Log Spectral norm ($\log\Vert\mathbf{W}\Vert_{\infty}$), 
               and Log Alpha Norm ($\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$) for weight matrices from models of different sizes in the GPT2 architecture series.

### 2020-02-17 submission/FILES_TO_ARXIV_16feb20/empirical2_TRfmt.tex
files I uploaded to arXiv

\documentclass[11pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\evensidemargin}{-0.5in}
\addtolength{\topmargin}{-1.0in}
\addtolength{\textheight}{1.7in}
\newlength{\defbaselineskip}
\setlength{\defbaselineskip}{\baselineskip}

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{framed}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{color}
\usepackage{xcolor}
\usepackage{paralist}

\usepackage{subfigure}

\usepackage{longtable} 
\usepackage{makecell}

\usepackage[multiple]{footmisc}
\usepackage[section]{placeins}

\usepackage{hyperref}
\hypersetup{
     colorlinks   = true,
     linkcolor    = blue,
     citecolor    = green
}

\newcommand{\fix}[1]{\textcolor{red}{#1}}
\newcommand{\comment}[1]{\textcolor{blue}{#1}}
\newcommand{\awk}[1]{\textcolor{green}{#1}}

\newcommand{\argmin}{\text{argmin}}
\newcommand{\Probab}[1]{\mbox{}{\bf{Pr}}\left[#1\right]}
\newcommand{\Expect}[1]{\mbox{}{\bf{E}}\left[#1\right]}
\newcommand{\ExpectBracket}[1]{\mbox{}\langle#1\rangle}

\newcommand {\nred}[1]{{\color{red}\sf{[#1]}}}
\newcommand {\ngreen}[1]{{\color{green}\sf{[#1]}}}
\newcommand {\ncyan}[1]{{\color{cyan}\sf{[#1]}}}
\newcommand {\michael}[1]{{\color{red}\sf{[michael: #1]}}}
\newcommand {\charles}[1]{{\color{blue}\sf{[charles: #1]}}}
\newcommand {\serena}[1]{{\color{orange}\sf{[charles: #1]}}}

\usepackage[normalem]{ulem}

\begin{document}

\title{%
Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data
}

\author{%
Charles H. Martin\thanks{Calculation Consulting, 8 Locksley Ave, 6B, San Francisco, CA 94122, \texttt{charles@CalculationConsulting.com}.} 
\and 
Tongsu (Serena) Peng\thanks{Calculation Consulting, 8 Locksley Ave, 6B, San Francisco, CA 94122, \texttt{serenapeng7@gmail.com}.}
\and
Michael W. Mahoney\thanks{ICSI and Department of Statistics, University of California at Berkeley, Berkeley, CA 94720, \texttt{mmahoney@stat.berkeley.edu}.}
}

\date{}
\maketitle

\begin{abstract}
\input{abstract_file}
\end{abstract}

\input{introduction}
\input{background}
\input{methods}
\input{cv_models}
\input{nlp_models}
\input{all_models}
\input{conclusion}

\vspace{-2mm}
\noindent
\paragraph{Acknowledgements.}
MWM would like to acknowledge ARO, DARPA, NSF, and ONR as well as the UC Berkeley BDD project and a gift from Intel for providing partial support of this work.
We would also like to thank Amir Khosrowshahi and colleagues at Intel for helpful discussion regarding the Group Regularization distillation technique.

   %% \bibliographystyle{unsrt}
   %% {\small
   %% \bibliography{dnns}
   %% }

{\small

\begin{thebibliography}{10}

\bibitem{weightwatcher_package}
{WeightWatcher}, 2018.
\newblock \url{https://pypi.org/project/WeightWatcher/}.

\bibitem{kdd20_sub_repo}
\url{https://github.com/CalculatedContent/ww-trends-2020}.

\bibitem{EB01_BOOK}
A.~Engel and C.~P. L.~Van den Broeck.
\newblock {\em Statistical mechanics of learning}.
\newblock Cambridge University Press, New York, NY, USA, 2001.

\bibitem{MM17_TR}
C.~H. Martin and M.~W. Mahoney.
\newblock Rethinking generalization requires revisiting old ideas: statistical
  mechanics approaches and complex learning behavior.
\newblock Technical Report Preprint: arXiv:1710.09553, 2017.

\bibitem{BKPx20}
Y.~Bahri, J.~Kadmon, J.~Pennington, S.~Schoenholz, J.~Sohl-Dickstein, and
  S.~Ganguli.
\newblock Statistical mechanics of deep learning.
\newblock {\em Annual Review of Condensed Matter Physics}, pages 000--000,
  2020.

\bibitem{MM18_TR}
C.~H. Martin and M.~W. Mahoney.
\newblock Implicit self-regularization in deep neural networks: Evidence from
  random matrix theory and implications for learning.
\newblock Technical Report Preprint: arXiv:1810.01075, 2018.

\bibitem{MM19_HTSR_ICML}
C.~H. Martin and M.~W. Mahoney.
\newblock Traditional and heavy-tailed self regularization in neural network
  models.
\newblock In {\em Proceedings of the 36th International Conference on Machine
  Learning}, pages 4284--4293, 2019.

\bibitem{MM19_KDD}
C.~H. Martin and M.~W. Mahoney.
\newblock Statistical mechanics methods for discovering knowledge from modern
  production quality neural networks.
\newblock In {\em Proceedings of the 25th Annual ACM SIGKDD Conference}, pages
  3239--3240, 2019.

\bibitem{MM20_SDM}
C.~H. Martin and M.~W. Mahoney.
\newblock Heavy-tailed {U}niversality predicts trends in test accuracies for
  very large pre-trained deep neural networks.
\newblock In {\em Proceedings of the 20th SIAM International Conference on Data
  Mining}, 2020.

\bibitem{BouchaudPotters03}
J.~P. Bouchaud and M.~Potters.
\newblock {\em Theory of Financial Risk and Derivative Pricing: From
  Statistical Physics to Risk Management}.
\newblock Cambridge University Press, 2003.

\bibitem{SornetteBook}
D.~Sornette.
\newblock {\em Critical phenomena in natural sciences: chaos, fractals,
  selforganization and disorder: concepts and tools}.
\newblock Springer-Verlag, Berlin, 2006.

\bibitem{BP11}
J.~P. Bouchaud and M.~Potters.
\newblock Financial applications of random matrix theory: a short review.
\newblock In G.~Akemann, J.~Baik, and P.~Di Francesco, editors, {\em The Oxford
  Handbook of Random Matrix Theory}. Oxford University Press, 2011.

\bibitem{bun2017}
J.~Bun, J.-P. Bouchaud, and M.~Potters.
\newblock Cleaning large correlation matrices: tools from random matrix theory.
\newblock {\em Physics Reports}, 666:1--109, 2017.

\bibitem{NTS15}
B.~Neyshabur, R.~Tomioka, and N.~Srebro.
\newblock Norm-based capacity control in neural networks.
\newblock In {\em Proceedings of the 28th Annual Conference on Learning
  Theory}, pages 1376--1401, 2015.

\bibitem{BFT17_TR}
P.~Bartlett, D.~J. Foster, and M.~Telgarsky.
\newblock Spectrally-normalized margin bounds for neural networks.
\newblock Technical Report Preprint: arXiv:1706.08498, 2017.

\bibitem{LMBx18_TR}
Q.~Liao, B.~Miranda, A.~Banburski, J.~Hidary, and T.~Poggio.
\newblock A surprising linear relationship predicts test performance in deep
  networks.
\newblock Technical Report Preprint: arXiv:1807.09659, 2018.

\bibitem{MM20_unpub_work}
C.~H. Martin and M.~W. Mahoney.
\newblock Unpublished results, 2020.

\bibitem{imagenet}
O.~Russakovsky et~al.
\newblock Imagenet large scale visual recognition challenge.
\newblock {\em International Journal of Computer Vision}, 115(3):211--252,
  2015.

\bibitem{pytorch}
A.~Paszke et~al.
\newblock Pytorch: An imperative style, high-performance deep learning library.
\newblock In {\em Annual Advances in Neural Information Processing Systems 32:
  Proceedings of the 2019 Conference}, pages 8024--8035, 2019.

\bibitem{osmr}
{Sandbox for training convolutional networks for computer vision}.
\newblock \url{https://github.com/osmr/imgclsmob}.

\bibitem{resnet1000}
K.~He, X.~Zhang, S.~Ren, and J.~Sun.
\newblock Identity mappings in deep residual networks.
\newblock Technical Report Preprint: arXiv:1603.05027, 2016.

\bibitem{CWZZ17_TR}
Y.~Cheng, D.~Wang, P.~Zhou, and T.~Zhang.
\newblock A survey of model compression and acceleration for deep neural
  networks.
\newblock Technical Report Preprint: arXiv:1710.09282, 2017.

\bibitem{distiller}
{Intel Distiller package}.
\newblock \url{https://nervanasystems.github.io/distiller}.

\bibitem{Attn2017}
A.~Vaswani et~al.
\newblock Attention is all you need.
\newblock Technical Report Preprint: arXiv:1706.03762, 2017.

\bibitem{huggingface}
T.~Wolf et~al.
\newblock Huggingface's transformers: State-of-the-art natural language
  processing.
\newblock Technical Report Preprint: arXiv:1910.03771, 2019.

\bibitem{gpt2-xl}
{OpenAI GPT-2: 1.5B Release}.
\newblock \url{https://openai.com/blog/gpt-2-1-5b-release/}.

\bibitem{CNNSVD}
H.~Sedghi, V.~Gupta, and P.~M. Long.
\newblock The singular values of convolutional layers.
\newblock Technical Report Preprint: arXiv:1805.10408, 2018.

\bibitem{GloBen10}
X.~Glorot and Y.~Bengio.
\newblock Understanding the difficulty of training deep feedforward neural
  networks.
\newblock In {\em Proceedings of the 13th International Workshop on Artificial
  Intelligence and Statistics}, pages 249--256, 2010.

\end{thebibliography}

}

\appendix
\input{appendix}

\end{document}

### 2020-02-17 submission/FILES_TO_ARXIV_16feb20/introduction.tex
files I uploaded to arXiv

--- 2020-02-16 submission/introduction.tex
@@
-%KDD% \vspace{-1mm}
 \section{Introduction}
 \label{sxn:intro}
-%\vspace{-1mm}
 
 A common problem in machine learning (ML) 
-%and artificial intelligence (AI) 
 is to evaluate the quality of a given model.
 A popular way to accomplish this
-%, in particular in academic environments, 
 is to train a model and then evaluate its training/testing error.
@@
 We consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
-%
 By now, there are many such state-of-the-art models that are publicly-available, e.g., 
@@
 \end{quote}
-
-%\begin{itemize}[leftmargin=*]
-%\item 
-%First, 
-%Motivated by practical problems, we formulate the question ``Can one predict trends in the quality of state-of-the-art neural networks without access to training or testing data''
-%\item
-%Second, 
-%To answer this question, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in CV and NLP.
-%\end{itemize}
 
@@
 We have examined hundreds of models, and we identify general trends, but we also highlight interesting exceptions.
-%%Several of the most interesting are described below.
 
-%KDD% \vspace{-1mm}
 \paragraph{The WeightWatcher Tool.}
@@
 
-%KDD% \vspace{-1mm}
 \paragraph{Organization of this paper.}

### 2020-02-17 submission/FILES_TO_ARXIV_16feb20/methods.tex
files I uploaded to arXiv

--- 2020-02-16 submission/methods.tex
@@
-%KDD% \vspace{-1mm}
 \section{Methods}
 \label{sxn:methods}
-%\vspace{-1mm}
 
@@
 
-%KDD% \vspace{-1mm}
 \paragraph{DNN Empirical Quality Metrics.}
@@
 The best performing empirical quality metrics depend on the norms and/or spectral properties of each weight matrix,
-$\mathbf{W}$, and/or, equivalently, it's \emph{Empirical Correlation Matrix}: $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.%
+$\mathbf{W}$, and/or, equivalently, it's \emph{Empirical Correlation Matrix}: $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
 
@@
 \end{equation}
-%where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
 Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.
@@
 The \emph{Weighted Alpha metric} is an average of $\alpha_l$ over all layers $l \in \{1,\ldots,l\}$, weighted by the size, or scale, or each matrix,
-%(and it approximates the average log $\alpha$-Shatten Norm metric),
 \begin{equation}
@@
 The $\lambda_{max}$ is a measure of the size, or scale, of $\mathbf{W}$.
-%
 Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric for the entire~DNN.
 
-%KDD% \vspace{-1mm}
 \paragraph{Convolutional Layers and Normalization issues.}

### 2020-07-02 submission/abstract_file.tex
full pass in light of kdd reviewer comments

--- 2020-02-12 submission/abstract_file.tex
@@
 In many practical applications, one works with deep neural network (DNN) models trained by someone else.
-For such \emph{pretrained models}, one typically does not have access to training data or test data.
-Moreover, one does not know many details about the model, such as the specifics of the training data, the loss function, the hyperparameter values, etc.
+For such \emph{pretrained models}, one often does not have access to training data or test data.
+Moreover, one may not know many details about the model, such as the specifics of the training data, the loss function, the hyperparameter values, etc.
 Given one or many pretrained models, can one say anything about the expected performance or quality of the models?
-Here, we present and evaluate empirical quality metrics for pretrained DNN models at scale.
+Here, we provide a detailed empirical meta-analysis to address this question, evaluating quality metrics for pretrained DNN models at scale.
 %%Our metrics are drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as Heavy-Tailed Random Matrix Theory (HT-RMT), as used in the recently-developed Theory of Heavy-Tailed Self Regularization (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models).
@@
 %
-Using the open-source \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision (CV) and natural language processing (NLP).
+Using the open-source \emph{WeightWatcher} tool to analyze DNN weight matrices, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision (CV) and natural language processing (NLP).
 We examine both familiar norm-based capacity control metrics (Frobenius and Spectral norms) as well as newer Power Law (PL) based metrics (including fitted PL exponents, $\alpha$, and the Weighted Alpha metric, $\hat{\alpha}$), from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
@@
 We find that norm-based metrics correlate well with reported test accuracies for well-trained models across nearly all CV architecture series.
-On the other hand, we find that norm-based metrics can not distinguish ``good-versus-bad'' models---which, arguably is the point of needing quality metrics.  
-Indeed, they may give spurious results.
-We also find that PL-based metrics do much better---quantitatively better at discriminating among a series of ``good-better-best'' models, and qualitatively better at discriminating ``good-versus-bad'' models.
-PL-based metrics can also be used to characterize fine-scale properties of these models, and we introduce the layer-wise \emph{Correlation Flow} as new quality assesment.
-We show how poorly-trained (and/or poorly fine-tuned) models may exhbit both \emph{Scale Collapse} and unsually large PL exponents, $\alpha\gg 6$, in particular for recent NLP models.
+On the other hand, we find that norm-based metrics can not distinguish well-trained versus poorly-trained models---which, arguably is the point of needing quality metrics.  
+%%%Indeed, they may give spurious results.
+We also find that PL-based metrics do much better---qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among a series of well-trained versus more-well-trained models.
+PL-based metrics can also be used to characterize fine-scale properties of these models.
+%%%, and we introduce the layer-wise \emph{Correlation Flow} as new quality assessment.
+%%%We show how poorly-trained (and/or poorly fine-tuned) models may exhibit both \emph{Scale Collapse} and unusually large PL exponents, $\alpha\gg 6$, in particular for recent NLP models.
 Our techniques, as implemented in the \emph{WeightWatcher} tool, can be used to identify when a pretrained DNN has problems that can not be detected simply by examining training/test accuracies.

### 2020-07-02 submission/all_models.tex
full pass, fixed lots of typos, etc.; semifinal, nearly done
full pass in light of kdd reviewer comments

@@
-%KDD% \vspace{-1mm}
-\section{Comparing Hundreds of CV Models}
+%KDD% 
+\vspace{-1mm}
+\section{Comparing Hundreds of Models}
 \label{sxn:all_cv_models}
@@
 
-In this section, we summarize results from a large-scale analysis of hundreds of CV models, including models developed for image 
+In this section, we summarize results from a large-scale analysis of hundreds of publicly-available  models, including models developed for image 
 classification, segmentation, and a range of related tasks.  
-Our aim is to complement the detailed results from Sections~\ref{sxn:cv} and~\ref{sxn:nlp} by providing broader conclusions. % about norm-based and PL-based metrics.
+Our aim is to provide a broad analysis that complements the detailed analysis in Sections~\ref{sxn:cv} and~\ref{sxn:nlp}.
+There, we considered CV and NLP models with similar structures. 
+Here, we provide much broader conclusions based on a much larger set of models with a much more diverse set of structures. % about norm-based and PL-based metrics.
 The models we consider have been pretrained on nine datasets.   %, shown in Table~\ref{table:datasets}.
@@
 \caption{Comparison of linear regression fits for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering 108 (out of over 400) different pretrained DNNs.
-         We include regressions only for architectures with five or more data points, and which are postively correlated with test error.
+         We include regressions only for architectures with five or more data points, and which are positively correlated with test error.
          These results can be readily reproduced using the Google Colab notebooks (see Appendix~\ref{sxn:appendix}). 

### 2020-07-02 submission/background.tex
full pass in light of kdd reviewer comments

@@
 There is very little work asking how to predict, in a theoretically-principled manner, the quality of large-scale state-of-the-art DNNs, and how to do so without access to training data or testing data or details of the training protocol, etc.
-Our 
-approach  %%  work 
-is, however, 
-%% loosely 
-related to 
-two  %%  several 
-other lines of work.
+%Our 
+%approach  %%  work 
+%is, however, 
+%%% loosely 
+%related to 
+%two  %%  several 
+%other lines of work.
+Similarly, there is very little work on performing weight matrix analysis.
+Here are several other lines of work most related to our approach.
 
@@
 our best-performing metrics (those using fitted PL exponents) are based on statistical mechanics~\cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM19_KDD, MM20_SDM}, in particular the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+See Appendix~\ref{sxn:theory-review-appendix} for a brief overview of this theory.
 We emphasize that the way in which we (and HT-SR Theory) \emph{use} statistical mechanics theory is quite different than the way it is more commonly formulated.
@@
 In this area, theoretical work aims to prove generalization bounds, and applied work uses these norms to construct regularizers to improve training.
-While we do find that norms provide relatively good quality metrics, at least for distinguishing good-better-best among well-trained models, we are not interested in proving generalization bounds or developing new regularizers.
+While we do find that norms provide relatively good quality metrics, at least for distinguishing among well-trained versus more-well-trained models, we are not interested in proving generalization bounds or developing new regularizers.
+
+\paragraph{Weight matrix analysis.}
+To our knowledge, our work, starting in 2018 with the \emph{WeightWatcher} tool~\cite{weightwatcher_package}, is the first to perform a detailed analysis of the weight matrices of DNN models~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+Subsequent to the release of the initial version of this paper, we became aware of two other works that were posted to the arXiv in 2020 within weeks of the initial arXiv version of this paper~\cite{EJRUY20_TR,UKGBT20_TR}.
+Both of these papers validate our basic result that one can gain substantial insight into model quality by examining weight matrices without access to any training or testing data.
+However, both of these papers consider smaller models drawn from a much narrower range of applications than we consider, and previous results in HT-SR Theory suggest that many of the insights from these smaller models may not extend to the state-of-the-art CV and NLP models we consider.
 

### 2020-07-02 submission/conclusion.tex
full pass in light of kdd reviewer comments

@@
 We have developed (based on strong theory) and evaluated (on a large corpus of publicly-available pretrained models from CV and NLP) methods to predict trends in the quality of state-of-the-art neural networks---without access to training or testing data.
-Prior to our work, it was not obvious that norm-based metrics would perform well to predict trends in quality \emph{across} models (as they are usually used \emph{within} a given model or parameterized model class, e.g., to bound generalization error or to construct regularizers).
-Our results are the first to demonstrate that they can be used for this important practical problem.
-That PL-based metrics perform better (than norm-based metrics) should not be surprising---at least to those familiar with the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} (since our use of PL exponents is designed to capture the idea that well-trained models capture correlations over many size scales in the data).
+Our main methodology involved performing a weight matrix meta-analysis, using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package}, and informed by the recently-developed HT-SR~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+Prior to our work, it was not even obvious that norm-based metrics would perform well to predict trends in quality \emph{across} models (as they are usually used \emph{within} a given model or parameterized model class, e.g., to bound generalization error or to construct regularizers).
+Importantly, our results are the first to demonstrate that they can be used for this important practical problem.
+
+Our results also demonstrate that PL-based metrics perform better (than norm-based metrics). 
+This should \emph{not} be surprising---at least to those familiar with the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}---since our use of PL exponents is designed to capture the idea that well-trained models capture information correlations over many size scales in the data.
 Again, though, our results are the first to demonstrate this.
+
 It is also gratifying that our approach can be used to provide fine-scale insight (such as rationalizing the flow of correlations or the collapse of size scale) throughout a network. 
+Clearly, in light of their role in our results on both CV and NLP models, both \emph{Correlation Flow} and \emph{Scale Collapse} are notions that future work should aim to understand better, both for improved training as well as for improved diagnostics on pretrained~models.
 

### 2020-07-02 submission/cv_models.tex
fixed fig 3
full pass in light of kdd reviewer comments

@@
 
-We perform \emph{coarse model analysis}, comparing and contrasting the four model series, and predicting trends in model quality. 
-We also perform \emph{fine layer analysis}, as a function of depth for these models, illustrating that PL-based metrics can provide novel insights among the VGG, ResNet/ResNet-1K, and DenseNet architectures.
+We first perform \emph{coarse model analysis}, comparing and contrasting the four model series, and predicting trends in model quality. 
+We then perform \emph{fine layer analysis}, as a function of depth for these models.
+This goes beyond simply predicting trends in model quality, instead illustrating that PL-based metrics can provide novel insights among the VGG, ResNet/ResNet-1K, and DenseNet architectures.
 
@@
 Similar results (not shown) are obtained for the Top5 accuracies.
-Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best.
-These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of ``good-better-best'' models, with PL-based metrics performing somewhat (i.e., quantitatively) better.
+Overall, for the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best.
+These model series are all very well-trodden, and thus our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of well-trained versus more-well-trained models, with PL-based metrics performing somewhat (i.e., quantitatively) better.
 
@@
 %For completeness and reproducibility here, the DenseNet Log $\alpha$-Norm plot is given in Appendix~\ref{sxn:appendix}.
-(Note that these and many other such plots can be seen on our publicly-available~repo.)
+(These and many other such plots can be seen on our publicly-available~repo.)
 
@@
 See Figure~\ref{fig:3models-alpha-layers}, which plots $\alpha$ for each layer (the first layer corresponds to data, the last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (no BN), ResNet, and DenseNet series.
-(Again, a much more detailed set of plots is available at our repo; but note that the corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than the results we present here.)
+(Again, many more such plots are available at our repo; but note that the corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than the results we present here.)
 
@@
     \qquad
-    \subfigure[ ResNet-1K ]{
+    \subfigure[ ResNet ]{
         \includegraphics[width=4.9cm]{img/ResNet_fnl_alpha_depth.png} 
@@
     \qquad
-    \subfigure[ ResNet-1K (overlaid) ]{
+    \subfigure[ ResNet (overlaid) ]{
         \includegraphics[width=4.9cm]{img/resnet_alpha_overlaid_depth.png} 
@@
 \paragraph{Comparison of VGG, ResNet, and DenseNet Architectures.}
-We can interpret these observations by recalling the architectural differences between the VGG, ResNet, and DenseNet architectures,
+We can interpret these layer-wise observations by recalling the architectural differences between the VGG, ResNet, and DenseNet architectures,
 and, in particular, the number of of residual connections. VGG resembles the  traditional convolutional architectures,
@@
 
-We conjecture that the efficiency and effectiveness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and strongly optimized.
+We conjecture that the efficiency and effectiveness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and more strongly optimized.
 Contrast this with the DenseNet models, which contains many connections between every layer.
-Our results (large $\alpha$, meaning they even a PL model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly optimized.
+Our results (large $\alpha$, meaning that even a PL model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly~optimized.
+
+Fine-scale measurements such as these enable us to form hypotheses as to the inner workings of DNN models, opening the door to an improved understanding of \emph{why} DNNs work, as well as \emph{how} to design better DNN models.
+We next turn to two such~examples.
 
@@
 
-More generally, we can understand the results presented in Figure~\ref{fig:3models-alpha-layers} in terms of what we will call the \emph{Correlation Flow} of the model. 
+We can understand the results presented in Figure~\ref{fig:3models-alpha-layers} in terms of what we will call the \emph{Correlation Flow} of the model. 
 Recall that the average Log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on ideas from the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
-There, one expects the weight matrices of well-trained DNNs will exhibit correlations over many size scales. 
-Their ESDs can be well-fit by a (truncated) PL, with exponents $\alpha\in[2,4]$.
+There, one expects that the weight matrices of well-trained DNNs will exhibit correlations over many size scales, as is well-known in other strongly-correlated systems~\cite{BouchaudPotters03, SornetteBook}.
+In particular, this would imply that their ESDs can be well-fit by a truncated PL, with exponents $\alpha\in[2,4]$.
 Much larger values $(\alpha\gg 6)$ may reflect poorer PL fits, whereas smaller values $(\alpha\sim 2)$, are associated with models that generalize better.
 %, but the presence of many adjacent layers with relatively good PL fits \emph{suggests} that those adjacent layers ``impedance match'' well and that the correlations that are captured by one layer propagate well to subsequent layers.
+
 Informally, one would expect a DNN model to perform well when it facilitates the propagation of information/features across layers.
 Previous work argues this by computing the gradients over the input data.
-In the absence of training/test data, one might hope that this leaves empirical signatures on weight matrices, and thus we can to try to quantify this by measuring the PL properties of weight matrices.
-In this case, smaller $\alpha$ values correspond to layers in which correlations across multiple scales are better captured~\cite{MM18_TR,SornetteBook}, and we expect that small $\alpha$ values that are stable across multiple layers enable better \emph{correlation flow} through the network.
+In the absence of training/test data, one might hypothesize that this flow of information leaves empirical signatures on weight matrices, and thus that we can quantify this by measuring the PL properties of weight matrices.
+In this case, given the statistical mechanical basis of HT-SR Theory, smaller $\alpha$ values would correspond to layers in which information correlations between data across multiple scales are better captured~\cite{MM18_TR,SornetteBook}, and thus we hypothesize that small $\alpha$ values that are stable across multiple layers enable better \emph{correlation flow} through the network.
 We have seen this in many models, including those shown in Figure~\ref{fig:3models-alpha-layers}.
+
+This is superficially similar to recent work on the information bottleneck~\cite{ST17_TR}, except that here we work in an entirely unsupervised setting.
+Developing theory to evaluate this hypothesis and exploit this insight is an intriguing direction for future work.
 
@@
 More generally, do fitted $\alpha$ values contain information not captured by norms? 
-In examining hundreds of pretrained models, we have found several anomalies that demonstrate the power of our approach.
+In examining hundreds of pretrained models, we have found several anomalies that are ``exceptions that prove the rule'' and that demonstrate the power of our approach.
 In particular, to show that $\alpha$ does capture something different, consider the following example, which looks at a compressed/distilled DNN model~\cite{CWZZ17_TR}.
@@
 
-The pretrained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the 
+The obvious interpretation is the following. 
+Pretrained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the 
 VGG, ResNet, or DenseNet series.  %% we considered above.
-While norms make good regularizers for a single model, there is no reason \emph{a priori} to expect them correlate so well with test accuracies across different models.
-We do expect, however, the PL $\alpha$ to do so because it effectively measures the amount of correlation in the model~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+While norms make good regularizers for a single model, there is no reason \emph{a priori} to expect them correlate so well with test accuracies across different models, and they may not differentiate well-trained versus poorly-trained models.
+We do expect, however, the PL $\alpha$ to do so, because it effectively measures the amount of information correlation in the model~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+Among other things, this suggests that the $\alpha$ values will improve, i.e., decrease, over time, as distillation techniques continue to improve.
 The reason for the anomalous behavior shown in 

### 2020-07-02 submission/introduction.tex
full pass in light of kdd reviewer comments

@@
 It is of great practical interest to have metrics to evaluate the quality of a trained model---in the absence of training/testing data and without any detailed knowledge of the training/testing process.
-We seek a practical theory for pretrained models which can predict how, when, and why such models can be expected to perform well or~poorly.
+There is a need for a \emph{practical theory} for pretrained models which can predict how, when, and why such models can be expected to perform well or~poorly.
 
-In this paper, we present and evaluate quality metrics for pretrained deep neural network (DNN) models, and we do so at scale.
-We consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
+In this paper, we provide a detailed empirical analysis to address these questions, evaluating quality metrics for pretrained deep neural network (DNN) models, and we do so at scale.
+Our approach may be viewed as a statistical meta-analysis of previously published work, where we consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
 %
@@
 \footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
-These provide a large corpus of models that by some community standard are state-of-the-art.%
-\footnote{Clearly, there is a selection bias or survivorship bias here---people tend not to make publicly-available their poorly-performing models---but these models are things in the world that (like social networks or the internet) can be analyzed for their properties.}
-Importantly, all of these models have been trained by someone else and have been viewed to be of sufficient interest/quality to be made publicly-available; and, for all of these models, we have no access to training data or testing data, and we have no knowledge of the training/testing protocols.
+These publicly-available models provide a large corpus of models that by some community standard are state-of-the-art. %
+%%%\footnote{Clearly, there is a selection bias or survivorship bias here---people tend not to make publicly-available their poorly-performing models---but these models are things in the world that (like social networks or the internet) can be analyzed for their properties.}
+Importantly, all of these models have been trained by someone else and have been viewed to be of sufficient interest/quality to be made publicly-available; and, for all of these models, we have no access to training data or testing data, and we have no specific knowledge of the training/testing protocols.
 
-The \emph{quality metrics} we consider are based on the spectral properties of the layer weight matrices.
-They are based on norms of weight matrices (such norms have been used in traditional statistical learning theory to bound capacity and construct regularizers) and/or parameters of power law (PL) fits of the eigenvalues of weight matrices (such PL fits are based on statistical mechanics approaches to DNNs).
-Note that, while we use traditional norm-based and PL-based metrics, our goals are not the traditional goals.
+In the absence of training and testing data, obvious quantities to examine are the weight matrices of pretrained models.
+Thus,
+the \emph{quality metrics} we consider are based on the spectral properties of the layer weight matrices.
+In particular, they are based on norms of weight matrices 
+and/or parameters of power law (PL) fits of the eigenvalues of weight matrices.
+Such norm-based metrics have been used in traditional statistical learning theory to bound capacity and construct regularizers; and such PL fits are based on statistical mechanics approaches to DNNs.
+While we use traditional norm-based and PL-based metrics, our goals are not the traditional goals.
 Unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., as with hyperparameter optimization).%
@@
 
-To answer this question, we analyze hundreds of publicly-available pretrained state-of-the-art CV and NLP models. 
+To answer this question, we perform a statistical meta-analysis, where we 
+perform weight matrix analysis on %%%analyze 
+hundreds of publicly-available pretrained state-of-the-art CV and NLP models. 
 Here is a summary of our main results.
@@
 \textbf{Norm-based metrics and poorly-trained models.}
-Norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.), exhibiting \emph{Scale Collapse} for these models. 
+Norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.).
+For example, they may exhibit what we call \emph{Scale Collapse} for these models. 
 \item 
@@
 PL-based metrics do much better at predicting quality trends in pretrained CV/NLP models.  
-They are quantitatively better at discriminating ``good-better-best'' trends, and qualitatively better at distinguishing ``good-versus-bad''~models.
+They are qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among a series of well-trained versus more-well-trained models.
 \item 
 \textbf{PL-based metrics and model diagnostics.}
-PL-based metrics can also be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow}) in well-trained and poorly-trained models, and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
+PL-based metrics can also be used to characterize fine-scale model properties (including what we call layer-wise \emph{Correlation Flow}) in well-trained and poorly-trained models, and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
 \end{itemize}
@@
 \noindent
-We emphasize that our goal is a practical theory to predict trends in the quality of state-of-the-art DNN models, i.e., not to make a statement about every publicly-available model.
+In light of our main results, several points are worth emphasizing.
+First, our goal is a practical theory to predict trends in the quality of state-of-the-art DNN models, i.e., not to make a statement about every publicly-available model.
 We have examined hundreds of models, and we identify general trends, but we also highlight interesting exceptions.
 %%Several of the most interesting are described below.
+Second, our work provides a theoretically-principled empirical evaluation---by far the largest, most detailed, and most comprehensive to date---and the theory we apply was developed previously~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+Performing such a meta-analysis of previously-published work is common in certain areas, but it is quite rare in machine learning, where the emphasis is on developing better training protocols.
+Third, in addition to addressing our main question, our weight matrix analysis can be used to obtain fine-scale insight into DNN models.
+This can be used to develop metrics for \emph{how} to train better models as well as to form hypotheses for \emph{why} DNNs perform as they do.
+We provide two such examples---\emph{Scale Collapse} and \emph{Correlation Flow}---that we have observed in both CV and NLP models and that deserve attention in future work.
 

### 2020-07-02 submission/methods.tex
full pass in light of kdd reviewer comments

@@
 
-Let us write the Energy Landscape (or optimization function, parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$,~as:
+Let us write the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$,~as:
 \begin{equation}
@@
 Here, we will ignore these details, and will focus only on the pretrained layer weight matrices $\mathbf{W}_{l}$.
+Our methodological approach is thus similar to a statistical meta-analysis, common in biomedical research.
 
@@
 
+Since we analyze pretrained models, most of our analysis involves analyzing DNN weight matrices.
 The best performing empirical quality metrics depend on the norms and/or spectral properties of each weight matrix,
@@
 Recall that the eigenvalues are squares of the singular values $\sigma_{i}$ of $\mathbf{W}$: $\lambda_{i}=\sigma^{2}_{i}$.
-Also, note that we do \emph{not} normalize $\mathbf{X}$ by $1/N$; see Appendix~\ref{sxn:appendix} for a discussion of this issue.
+Also, note that we do \emph{not} normalize $\mathbf{X}$ by $1/N$.
+See Appendix~\ref{sxn:appendix} for a discussion of this issue.
+Observe that all four metrics can be computed easily from DNN weight matrices.
 
-The first two norms are well-known in ML; the last two deserve special mention.
+The first two metrics are well-known in ML; the last two deserve special mention.
 The empirical parameter $\alpha$ is the Power Law (PL) exponent that arises in the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
-Operationally, $\alpha$ is determined by using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated PL,
+The bottom line from HT-SR Theory is that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models; see Appendix~\ref{sxn:theory-review-appendix} for more details.
+Operationally, $\alpha$ is determined by analyzing DNN weight matrices using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated PL,
 \begin{equation}
@@
 where $L$ is the total number of layer weight matrices.
-The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a limited set of models.
+The Weighted Alpha metric was introduced and justified previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a much smaller and more limited set of models.
 
@@
 \noindent
-For small $\alpha$, the Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}; and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm, and may track this metric in some~cases.
+For small $\alpha$, the Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}; and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm.
 
 To avoid confusion, let us clarify the relationship between $\alpha$ and $\hat{\alpha}$.  
-We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated PL, parameterized by 2 values: the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.  
-(Technically, we also need the minimum eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.)
-The PL exponent $\alpha$ measures of the amount of correlation in a DNN layer weight matrix $\mathbf{W}$. 
+We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated PL, parameterized by 2 values: the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.% 
+\footnote{Technically, we also need the minimum eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.}
+The PL exponent $\alpha$ measures the amount of correlation in a DNN layer weight matrix $\mathbf{W}$. 
 It is valid for $\lambda\le\lambda_{max}$, and it is scale-invariant, i.e., it does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.
@@
 \paragraph{Convolutional Layers and Normalization issues.}
-There are several technical issues (regarding spectral analysis of convolutional layers and normalization of empirical matrices) that are important for reproducibility of our results.
+There are several technical issues, e.g., regarding spectral analysis of convolutional layers and normalization of empirical matrices, that are important for reproducibility of our results.
 See Appendix~\ref{sxn:appendix} for a discussion.

### 2020-07-02 submission/nlp_models.tex
full pass in light of kdd reviewer comments

@@
 (in particular, the interpretation of values of $\alpha \sim 2$ as modeling systems with good correlations over many size scales~\cite{BouchaudPotters03, SornetteBook}), 
-this suggests that these models fail to capture successfully many of the correlations in the data (relative to their size) and thus are substantially \emph{under}-trained.
+this suggests that these models fail to capture successfully many of the information correlations in the data (relative to their size) and thus are substantially \emph{under}-trained.
 %
@@
 Indeed, for such ``bad'' models, weight matrices may display rank collapse, decreased Frobenius mass, or unusually small Spectral norms.
-(This may be misinterpreted as ``smaller is better.'')
-In contrast, PL-based metrics, including the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $) display consistent behavior, even on poorly trained models.
+This may be misinterpreted as ``smaller is better.'' 
+Instead, it should probably be interpreted as being due to a similar mechanism to how distillation can ``damage'' otherwise good models.
+In contrast to norm-based metrics, PL-based metrics, including the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $) display consistent behavior, even on poorly trained models.
 Indeed, we can use these metrics to help identify when architectures need repair and when more and/or better data are needed.
@@
 
-Many NLP models, such as GPT and BERT, have some weight matrices with unusually large PL exponents (e.g., $\alpha\gg 6$).
+We have observed that many NLP models, such as GPT and BERT, have some weight matrices with unusually large PL exponents (e.g., $\alpha\gg 6$).
 This indicates these matrices may be \emph{under}-correlated (i.e., over-parameterized, relative to the amount of data).
-In this regime, the truncated PL fit itself may not be very reliable because the MLE estimator it uses is unreliable in this range (i.e., the specific $\alpha$ values returned by the truncated PL fits are less reliable, but having large versus small values of $\alpha$ is reliable).
-Phenomenologically, if we examine the ESD visually, we can usually describe these $\mathbf{W}$ as in the \emph{Bulk-Decay} or \emph{Bulk-plus-Spikes} phase~\cite{MM18_TR,MM19_HTSR_ICML}.
-Previous work~\cite{MM18_TR,MM19_HTSR_ICML} has conjectured that very well-trained DNNs would not have many \emph{outlier} $\alpha\gg 6$; and improved versions of GPT (shown below) and BERT (not shown) confirm this.
+In this regime, the truncated PL fit itself may not be very reliable because the MLE estimator it uses is unreliable in this range.
+That is, the specific $\alpha$ values returned by the truncated PL fits are less reliable, but having large versus small values of $\alpha$ is reliable.
+Phenomenologically, if we examine the ESD visually, we can usually describe these $\mathbf{W}$ as in the \emph{Bulk-Decay} or \emph{Bulk-plus-Spikes} phase from HT-ST Theory~\cite{MM18_TR,MM19_HTSR_ICML}.
+Previous work~\cite{MM18_TR,MM19_HTSR_ICML} has conjectured that very well-trained DNNs would not have many \emph{outlier} $\alpha\gg 6$; and, consistent with this, improved versions of GPT (shown below) and BERT (not shown) confirm this.
 
@@
 By comparing the poorly-trained (i.e., ``bad'') GPT to the well-trained (i.e., ``good'') GPT2-small, we can identify empirical indicators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
-By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously,
-an example of what we call a series of ``good-better-best'' models.
+By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously, as well as analyze well-trained versus more-well-trained models.
 
@@
 Observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
-Notice that in the GPT model, being poorly trained, the norm metrics all exhibit \emph{Scale Collapse}, compared to GPT2-small.
+Notice that in the poorly-trained GPT model, the norm metrics all exhibit \emph{Scale Collapse}, compared to well-trained GPT2-small model.
 
@@
 Observe that (with one minor exception involving the log Frobenius norm metric) all four metrics decrease as one goes from medium to large to xl, indicating that the larger models indeed look better than the smaller models.
-Notice that, for these well-trained models, the norm metrics now behave as expected, decreasing with increasing accuracy.
+Notice that, for these well-trained models, the norm metrics now behave as expected, i.e., there is no \emph{Scale Collapse} and norms are decreasing with increasing~accuracy.
 
@@
 It also has far fewer unusually-large outlying $\alpha$ values than GPT.
-From this (and other results not shown), we see that $\alpha$ provides a good quality metric for comparing these two models, the ``bad'' GPT versus the ``good'' GPT2-small.
+From this (and other results not shown), we see that $\alpha$ provides a good quality metric for comparing these two models, the poorly-trained GPT versus the well-trained GPT2-small.
 %
@@
 We next describe the behavior of the Spectral norm in GPT versus GPT2-small.
-In Figure~\ref{fig:GPT-snorm-hist}, the ``bad'' GPT model has a smaller mean/median Spectral norm as well as, spuriously, many much smaller Spectral norms, compared to the ``good'' GPT2-small, violating the conventional wisdom that smaller Spectral norms are better.
+In Figure~\ref{fig:GPT-snorm-hist}, the poorly-trained GPT model has a smaller mean/median Spectral norm as well as, spuriously, many much smaller Spectral norms, compared to the well-trained GPT2-small, violating the conventional wisdom that smaller Spectral norms are better.
 Indeed, because there are so many anonymously small Spectral norms, it appears that the GPT model may be exhibiting a kind of \emph{Scale Collapse}, like that observed in the distilled CV models (in Figure~\ref{fig:resnet204D5L}).
-This is important because it demonstrates that, while the Spectral (or Frobenius) norm may correlate well with predicted test error, it is \emph{not} a good indicator of the overall model quality.
-It can mispredict good-versus-bad questions in ways not seen with PL-based metrics.
+
+This is important because it demonstrates that, while the Spectral (or Frobenius) norm may correlate well with predicted test error, at least among reasonably well-trained models, it is \emph{not} a good indicator of the overall model quality in general.
+In particular, it can mispredict well-trained versus poorly-trained questions in ways not seen with PL-based metrics.
 Using it as an empirical quality metric may give spurious results when applied to poorly-trained or otherwise deficient models.
@@
 
-(Note that Figure~\ref{fig:GPT-snorm-hist} also shows some unusually large Spectral Norms.
+We should note that Figure~\ref{fig:GPT-snorm-hist} also shows some unusually large Spectral Norms.
 Upon examination, e.g., from Figure \ref{fig:gpt-snorm-layer} (below), we see that these correspond to the first embedding layer(s).
@@
 We discuss this further in Appendix~\ref{sxn:appendix}.
-Here, we do not include them in our computed average metrics in Table~\ref{table:nlp}, and we do not include them in the histogram plot in Figure~\ref{fig:GPT-snorm-hist}.)
+Here, we do not include them in our computed average metrics in Table~\ref{table:nlp}, and we do not include them in the histogram plot in Figure~\ref{fig:GPT-snorm-hist}.
 
@@
 The scale-invariant $\alpha$ metric lets us identify potentially poorly-trained models.
-These results show that the Correlation Flow differs significantly between GPT and GPT2-small (with the better GPT2-small looking more like the better ResNet-1K from Figure~\ref{fig:resnet-alpha-layer}).
+These results suggest that the Correlation Flow is also a useful concept for NLP models, and that it differs significantly between GPT and GPT2-small, with the better GPT2-small looking more like the better ResNet-1K from Figure~\ref{fig:resnet-alpha-layer}.
 
@@
 This Scale Collapse for the poorly-trained GPT is similar to what we observed for the distilled ResNet20 model in Figure~\ref{fig:resnet204Dalpha}.
-Because of the anomalous scale collapse that is frequently observed in poorly-trained models, these results suggest that scale-dependent norm metrics should not be directly applied to distinguish good-versus-bad models.
+Because of the anomalous scale collapse that is frequently observed in poorly-trained models, these results suggest that scale-dependent norm metrics should not be directly applied to distinguish well-trained versus poorly-trained models.
 
@@
 
-We now look across series of increasingly improving GPT2 models (i.e., we consider good-better-best questions), by examining both the PL exponent $\alpha$ as well as the Log Norm metrics.  
+We now look across series of increasingly improving GPT2 models (i.e., we consider well-trained versus more-well-trained questions), by examining both the PL exponent $\alpha$ as well as the Log Norm metrics.
 In general, as we move from GPT2-medium to GPT2-xl, histograms for both $\alpha$ exponents and the Log Norm metrics downshift from larger to smaller values. 
@@
 
-We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticeable between the differing good-better-best GTP2 models than between the good-versus-bad GPT and GPT2-small models.
+We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticeable between the differing well-trained versus more-well-trained GTP2 models than between the poorly-trained versus well-trained GPT and GPT2-small models.
 Unlike GPT, however, the layer Log Alpha Norms behave more as expected for GPT2 layers, with the larger models consistently having smaller norms. 
 Similarly, the Log Spectral Norm also decreases on average with the larger models (not shown).  
-As expected, the norm metrics can indeed distinguish among good-better-best models among a series well-trained models.
+As expected, the norm metrics can indeed distinguish among well-trained versus more-well-trained models among a series well-trained models.
 

### 2020-07-02 submission/appendixB.tex
add file I added to the revision in light of kdd comments

\section{Overview of Heavy-Tailed Self-Regularization}
\label{sxn:theory-review-appendix}

For completeness, we provide here a brief overview of Martin and Mahoney's Theory of Heavy-Tailed Self-Regularization (HT-SR)~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.

Let us write the objective/optimization function for a typical DNN with $L$ layers, with activation functions $h_{l}(\cdot)$, and with $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$, as:
\begin{equation*}
E_{DNN} \hspace{-1mm} = \hspace{-1mm} h_{L}(\mathbf{W}_{L}\cdot h_{L-1}(\mathbf{W}_{L-1}\cdot h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})  .
\end{equation*}
Typically, this model would be trained on some labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using Backprop, by minimizing the loss $\mathcal{L}$.
For simplicity, we do not indicate the structural details of the layers (e.g., Dense or not, Convolutions or not, Residual/Skip Connections, etc.). 
Each layer is defined by one or more layer 2D weight matrices $\mathbf{W}_{l}$, and/or the 2D feature maps $\mathbf{W}_{l,i}$ extracted from 2D Convolutional (Conv2D) layers.
A typical modern DNN may have anywhere between 5 and 5000 2D layer matrices.%
\footnote{%
For each Linear Layer, we get a  single $(N\times M)$ (real-valued) 2D weight matrix, denoted $\mathbf{W}_{l}$, for layer $l$.  
This includes Dense or Fully-Connected (FC) layers, as well as 1D Convolutional (Conv1D) layers, Attention matrices, etc.
We ignore the bias terms $\mathbf{b}_{l}$ in this analysis. 
Let the aspect ratio be $Q=\frac{N}{M}$, with $Q\ge 1$.
For the Conv2D layers, we have a 4-index Tensor, of the form $(N\times M \times c\times d)$, consisting
of $c\times d$ 2D feature maps of shape $(N\times M)$.    
We  extract $n_{l}=c\times d$ 2D weight matrices $\mathbf{W}_{l,i}$, one for each feature map $i=[1,\dots,n_{l}]$ for layer $l$.
}

In the HT-SR Theory, we analyze the eigenvalue spectrum (the ESD) of the associated correlation matrices~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
From this, we can characterize the amount and form of correlation, and therefore implicit self-regularizartion, present in the DNN's weight matrices.
For each layer weight matrix $\mathbf{W}$, of size $N \times M$, construct the associated $M\times M$ (uncentered) correlation matrix $\mathbf{X}$. 
Dropping the $L$ and $l,i$ indices, we have
$$
\mathbf{X} = \frac{1}{N}\mathbf{W}^{T}\mathbf{W}.
$$
If we compute the eigenvalue spectrum of $\mathbf{X}$, i.e., $\lambda_i$ such that
$  % $$
\mathbf{X}\mathbf{v}_{i}=\lambda_{i}\mathbf{v}_{i} , 
$  % $$
then the ESD of eigenvalues, $\rho(\lambda)$, is just a histogram of the eigenvalues, formally written as
\begin{equation}
\rho(\lambda)=\sum\limits_{i=1}^{M}\delta(\lambda-\lambda_{i})  .
\label{eqn:eigenval_hist}
\end{equation}
Using HT-SR Theory, we can characterize the \emph{correlations} in a weight matrix by examining its ESD, $\rho(\lambda)$.
It can be well-fit to a truncated power law (PL) distribution, given~as
\begin{equation}
\rho(\lambda)\sim\lambda^{-\alpha}  ,
\label{eqn:eigenval_pl}
\end{equation}
which is (at least) valid within a bounded range of eigenvalues $\lambda\in[\lambda^{min},\lambda^{max}]$.  
We can determine $\alpha$ by fitting the   ESD to a truncated PL, using the commonly accepted Maximum Likelihood (MLE) method of Clauset et al.~\cite{CSN09_powerlaw,ABP14}.
This method works very well for exponents between $\alpha\in(2,4)$; and it is adequate, although imprecise, for smaller and especially larger $\alpha$~\cite{newman2005_zipf}.

The original work on HT-SR Theory~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM} considered NNs including AlexNet and InceptionV3 (as well as DenseNet, ResNet, and VGG), and it showed that for nearly every $\mathbf{W}$, the (bulk and tail) of the ESDs can be fit to a truncated PL and the PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$.
Moreover, smaller exponents $\alpha$ are correlated with more implicit self-regularization and, correspondingly, better generalization~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
Our meta-analysis, the main results of which are summarized in this paper, has shown that these results are ubiquitous.
For example, 
upon examining nearly 10,000 layer weight matrices $\mathbf{W}_{l,i}$ across hundreds of different modern pre-trained DNN architectures, the ESD of nearly every $\mathbf{W}$ layer matrix can be fit to a truncated PL:
$70-80\%$ of the time, the fitted PL exponent $\alpha$ lies in the range $\alpha\in(2,4)$; and  
$10-20\%$ of the time, the fitted PL exponent $\alpha$ lies in the range $\alpha< 2$.  
Of course, there are exceptions: in any real DNN, the fitted $\alpha$ may range anywhere from $\sim 1.5$ to $10$ or higher (and, of course, larger values of $\alpha$ may indicate that the PL is not a good model for the data).  
Still, overall, in nearly all large, pre-trained DNNs, the correlations in the  weight matrices exhibit a remarkable Universality, being both Heavy Tailed, and having small---but not too small---PL exponents.

As for the mechanism responsible for these properties, statistical physics offers several possibilities~\cite{SornetteBook}, although more work needs to be done.
Indeed, it is not even clear that one mechanism is responsible for every case.
One particularly intriguing possibility has to do with the presence of multiplicative noise in the stochastic optimization algorithms used to train these models~\cite{HodMah20A_TR,SorCon97}, and our results motivate the need to understand this multiplicative dynamics in greater detail.

### 2020-07-02 submission/abstract_file.tex
minor edits

@@
 %%%Indeed, they may give spurious results.
-We also find that PL-based metrics do much better---qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among a series of well-trained versus more-well-trained models.
+We also find that PL-based metrics do much better---qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among a series of well-trained models \nred{of increasing depth and accuracy}.
+%versus more-well-trained models.
 PL-based metrics can also be used to characterize fine-scale properties of these models.

### 2020-07-02 submission/introduction.tex
minor edits

@@
 This may be true in worst-case theory, but models are used in practice, and there is a need for a \emph{practical theory} to guide that practice.
-Moreover, if ML is to become an industrial process, then that process will become siloed: some groups will gather data, other groups will develop models, and other groups will use those models.
+Moreover, if ML is to become an industrial process, then that process will become \nred{comparmentalized in ortder to scale}: some groups will gather data, other groups will develop models, and other groups will use those models.
 Users of models can not be expected to know the precise details of how models were built, the specifics of data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized,~etc.

### 2020-07-02 submission/methods.tex
minor edits

@@
 
-Let us write the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$,~as:
+Let us write the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for aE DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$,~as \nred{the minimization of a general Loss Function $\mathcal{L}$ over  N (labeled) training data instances $(\mathbf{x}_{i},y_{i})$.
+For a typical supervised classification problem, one has
 \begin{equation}
-E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})  .
+\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N}\left(\mathcal{L}(E_{DNN}(\mathbf{x}_{i})-y_{i})\right)
+\end{equation}
+where }
+\begin{equation}
+E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L}),
 \label{eqn:dnn_energy}
 \end{equation}
+\nred{the loss function $\mathcal{L}()$ can take on a myriad of forms\cite{Janocha2017}, and $E_{DNN}$ is what we call the Energy Landscape.}
+
 Each DNN layer contains one or more layer 2D  $N\times M$ weight matrices, $\mathbf{W}_{l}$, or pre-activation maps, $\mathbf{W}_{i,l}$, extracted from 2D Convolutional layers, and where $N > M$.% 

### 2020-07-02 submission/methods.tex
minor edits

@@
 
-Let us write the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for aE DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$,~as \nred{the minimization of a general Loss Function $\mathcal{L}$ over  N (labeled) training data instances $(\mathbf{x}_{i},y_{i})$.
+Let us write the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for aE DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$,~as \nred{the minimization of a general Loss Function $\mathcal{L}$ over  N training data instances and labels  $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$
 For a typical supervised classification problem, one has
@@
 \end{equation}
-where }
+where the loss function $\mathcal{L}()$ can take on a myriad of forms\cite{Janocha2017}, and 
 \begin{equation}
-E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L}),
+E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})
 \label{eqn:dnn_energy}
 \end{equation}
-\nred{the loss function $\mathcal{L}()$ can take on a myriad of forms\cite{Janocha2017}, and $E_{DNN}$ is what we call the Energy Landscape.}
+is what we call the Energy Landscape function.  Notice the $E_{DNN}$ does not explicitly depend on the data, but, rather,
+maps data instances vectors $(\mathbf{x}_i)$ to predictions (i.e. labels).
+Therefore, we can analyze $E_{DNN}$ absent of any training or test data. }
 
@@
 Assume we are given several pretrained DNNs, e.g., as part of an architecture series.
-The models have been trained and evaluated on labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using standard techniques.  
+The models have been trained and evaluated on the labeled data $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$, using standard techniques.  
 The pretrained pytroch model files are publicly-available, and the test accuracies have been reported online.  
-In this study, we do not have access to this data, and we have not trained any of the models ourselves,
-nor have we re-evaluated the test accuracies.
+In this study, we do not have access to this data, and we have not trained any of the models ourselves. 
+%nor have we re-evaluated the test accuracies.
 We expect that most well-trained, production-quality models will employ one or more forms of regularization, such as Batch Normalization (BN), Dropout, etc., and many will also contain additional structure such as Skip Connections, etc.
@@
 \begin{equation}
-\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}.
+\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\in[\lambda_{min},\lambda_{max}].
 \end{equation}
-%where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
-Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.
+Each of these quantities is defined for a given layer $\mathbf{W}$ matrix and the PL fitting method. \cite{Clauset}
+\nred{Here, where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$, 
+and $\lambda_{min}$ is selected automaticlly to yield the best PL fit (which minimizes the K-S distance). }
 

### 2020-07-02 submission/background.tex
minor edits

@@
 In this area, theoretical work aims to prove generalization bounds, and applied work uses these norms to construct regularizers to improve training.
-While we do find that norms provide relatively good quality metrics, at least for distinguishing among well-trained versus more-well-trained models, we are not interested in proving generalization bounds or developing new regularizers.
+While we do find that norms provide relatively good quality metrics, at least for distinguishing among \nred{(moderatly)} well-trained versus 
+\nred{very-well-trained} models, \nred{of increasing deopth an accuracy}, 
+we are not interested in proving generalization bounds or developing new regularizers.
 

### 2020-07-02 submission/cv_models.tex
minor edits

@@
 We have examined the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
-To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BN), and it plots the four quality metrics versus the reported test accuracies~\cite{pytorch},%
+To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without \nred{BatchNormalization, denoted} BN), and it plots the four quality metrics versus the reported test accuracies~\cite{pytorch},%
 \footnote{That is, these test accuracies have been previously reported and made publicly-available by others.
@@
 Overall, for the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best.
-These model series are all very well-trodden, and thus our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of well-trained versus more-well-trained models, with PL-based metrics performing somewhat (i.e., quantitatively) better.
+These model series are all very well-trodden, and thus our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of well-trained versus very-well-trained models, with PL-based metrics performing somewhat (i.e., quantitatively) better.
 

### 2020-07-02 submission/nlp_models.tex
minor edits

@@
 By comparing the poorly-trained (i.e., ``bad'') GPT to the well-trained (i.e., ``good'') GPT2-small, we can identify empirical indicators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
-By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously, as well as analyze well-trained versus more-well-trained models.
+By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously, as well as analyze well-trained versus \nred{very-well-trained} models.
 
@@
 
-We now look across series of increasingly improving GPT2 models (i.e., we consider well-trained versus more-well-trained questions), by examining both the PL exponent $\alpha$ as well as the Log Norm metrics.
+We now look across series of increasingly improving GPT2 models (i.e., we consider well-trained versus \nred{very--well-trained} questions), by examining both the PL exponent $\alpha$ as well as the Log Norm metrics.
 In general, as we move from GPT2-medium to GPT2-xl, histograms for both $\alpha$ exponents and the Log Norm metrics downshift from larger to smaller values. 
@@
 
-We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticeable between the differing well-trained versus more-well-trained GTP2 models than between the poorly-trained versus well-trained GPT and GPT2-small models.
+We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticeable between the differing well-trained 
+versus \nred{very-well-trained} GTP2 models than between the poorly-trained versus well-trained GPT and GPT2-small models.
 Unlike GPT, however, the layer Log Alpha Norms behave more as expected for GPT2 layers, with the larger models consistently having smaller norms. 
 Similarly, the Log Spectral Norm also decreases on average with the larger models (not shown).  
-As expected, the norm metrics can indeed distinguish among well-trained versus more-well-trained models among a series well-trained models.
+As expected, the norm metrics can indeed distinguish among well-trained versus \nred{very-well-trained} models among a series well-trained models.
 

### 2020-07-06 submission/abstract_file.tex
removed red and fixed some typos

--- 2020-07-02 submission/abstract_file.tex
@@
 %%%Indeed, they may give spurious results.
-We also find that PL-based metrics do much better---qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among a series of well-trained versus more-well-trained models.
+We also find that PL-based metrics do much better---qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among a series of well-trained models of increasing depth and accuracy.
+%versus very-well-trained models.
 PL-based metrics can also be used to characterize fine-scale properties of these models.

### 2020-07-06 submission/background.tex
removed red and fixed some typos

--- 2020-07-02 submission/background.tex
@@
 In this area, theoretical work aims to prove generalization bounds, and applied work uses these norms to construct regularizers to improve training.
-While we do find that norms provide relatively good quality metrics, at least for distinguishing among well-trained versus more-well-trained models, we are not interested in proving generalization bounds or developing new regularizers.
+While we do find that norms provide relatively good quality metrics, at least for distinguishing among (moderately) well-trained versus 
+very-well-trained models, of increasing depth and accuracy, 
+we are not interested in proving generalization bounds or developing new regularizers.
 

### 2020-07-06 submission/cv_models.tex
removed red and fixed some typos

@@
 We have examined the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
-To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without \nred{BatchNormalization, denoted} BN), and it plots the four quality metrics versus the reported test accuracies~\cite{pytorch},%
+To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without Batch Normalization, denoted BN), and it plots the four quality metrics versus the reported test accuracies~\cite{pytorch},%
 \footnote{That is, these test accuracies have been previously reported and made publicly-available by others.

### 2020-07-06 submission/introduction.tex
removed red and fixed some typos

--- 2020-07-02 submission/introduction.tex
@@
 This may be true in worst-case theory, but models are used in practice, and there is a need for a \emph{practical theory} to guide that practice.
-Moreover, if ML is to become an industrial process, then that process will become siloed: some groups will gather data, other groups will develop models, and other groups will use those models.
+Moreover, if ML is to become an industrial process, then that process will become compartmentalized in order to scale: some groups will gather data, other groups will develop models, and other groups will use those models.
 Users of models can not be expected to know the precise details of how models were built, the specifics of data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized,~etc.
@@
 PL-based metrics do much better at predicting quality trends in pretrained CV/NLP models.  
-They are qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among a series of well-trained versus more-well-trained models.
+They are qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among a series of well-trained versus very-well-trained models.
 \item 
@@
 To be fully reproducible, we only examine publicly-available, pretrained models, and we also provide all Jupyter and Google Colab notebooks used in an accompanying github repository~\cite{kdd20_sub_repo}.
-See Appendix~\ref{sxn:appendix} for details on how to reproduce all results.
+See Appendix~\ref{sxn:appendix} for details on how to reproduce all~results.
 

### 2020-07-06 submission/methods.tex
removed red and fixed some typos

@@
 
-Let us write the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for aE DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$,~as \nred{the minimization of a general Loss Function $\mathcal{L}$ over  N training data instances and labels  $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$
+Let us write the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$, as
+the minimization of a general Loss Function $\mathcal{L}$ over $N$ training data instances and labels, $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$.
 For a typical supervised classification problem, one has
 \begin{equation}
-\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N}\left(\mathcal{L}(E_{DNN}(\mathbf{x}_{i})-y_{i})\right)
+\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N}\left(\mathcal{L}(E_{DNN}(\mathbf{x}_{i})-y_{i})\right)  ,
 \end{equation}
-where the loss function $\mathcal{L}()$ can take on a myriad of forms\cite{Janocha2017}, and 
+where the loss function $\mathcal{L}(\cdot)$ can take on a myriad of forms~\cite{JC17_TR}, and where
 \begin{equation}
@@
 \end{equation}
-is what we call the Energy Landscape function.  Notice the $E_{DNN}$ does not explicitly depend on the data, but, rather,
-maps data instances vectors $(\mathbf{x}_i)$ to predictions (i.e. labels).
-Therefore, we can analyze $E_{DNN}$ absent of any training or test data. }
+is what we call the Energy (or optimization) Landscape function.  
+Notice that $E_{DNN}$ does not \emph{explicitly} depend on the data, but rather it maps data instance vectors (i.e., $\mathbf{x}_i$ values) to predictions (i.e., labels).
+Therefore, we can analyze $E_{DNN}$ absent of any training or test data.
 
@@
 \begin{equation}
-\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\in[\lambda_{min},\lambda_{max}].
+\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\in[\lambda_{min},\lambda_{max}] ,
 \end{equation}
-Each of these quantities is defined for a given layer $\mathbf{W}$ matrix and the PL fitting method. \cite{Clauset}
-\nred{Here, where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$, 
-and $\lambda_{min}$ is selected automaticlly to yield the best PL fit (which minimizes the K-S distance). }
+where $\lambda_{max}$ is the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$, and 
+where $\lambda_{min}$ is selected automatically to yield the best PL fit (which minimizes the K-S distance). 
+Each of these quantities is defined for a given layer $\mathbf{W}$ matrix and the PL fitting method~\cite{CSN09_powerlaw,ABP14}.
 

### 2020-07-06 submission/nlp_models.tex
removed red and fixed some typos

--- 2020-07-02 submission/nlp_models.tex
@@
 By comparing the poorly-trained (i.e., ``bad'') GPT to the well-trained (i.e., ``good'') GPT2-small, we can identify empirical indicators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
-By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously, as well as analyze well-trained versus more-well-trained models.
+By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously, as well as analyze well-trained versus very-well-trained models.
 
@@
 In particular, it can mispredict well-trained versus poorly-trained questions in ways not seen with PL-based metrics.
-Using it as an empirical quality metric may give spurious results when applied to poorly-trained or otherwise deficient models.
+Using it as an empirical quality metric may give spurious results when applied to poorly-trained or otherwise deficient~models.
 
@@
 
-We now look across series of increasingly improving GPT2 models (i.e., we consider well-trained versus more-well-trained questions), by examining both the PL exponent $\alpha$ as well as the Log Norm metrics.
+We now look across series of increasingly improving GPT2 models (i.e., we consider well-trained versus very-well-trained questions), by examining both the PL exponent $\alpha$ as well as the Log Norm metrics.
 In general, as we move from GPT2-medium to GPT2-xl, histograms for both $\alpha$ exponents and the Log Norm metrics downshift from larger to smaller values. 
@@
 
-We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticeable between the differing well-trained versus more-well-trained GTP2 models than between the poorly-trained versus well-trained GPT and GPT2-small models.
+We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticeable between the differing well-trained 
+versus very-well-trained GTP2 models than between the poorly-trained versus well-trained GPT and GPT2-small models.
 Unlike GPT, however, the layer Log Alpha Norms behave more as expected for GPT2 layers, with the larger models consistently having smaller norms. 
 Similarly, the Log Spectral Norm also decreases on average with the larger models (not shown).  
-As expected, the norm metrics can indeed distinguish among well-trained versus more-well-trained models among a series well-trained models.
+As expected, the norm metrics can indeed distinguish among well-trained versus very-well-trained models among a series well-trained models.
 
-We do notice, however, that while the peaks of the $\alpha$ are getting smaller, towards $2.0$, the tails of the distribution shifts right, with larger GPT2 models having more usually large $\alpha$ (also not shown).
+We do notice, however, that while the peaks of the $\alpha$ are getting smaller, towards $2.0$, the tails of the distribution shifts right, with larger GPT2 models having more unusually large $\alpha$ (also not shown).
 We suspect this indicates that these larger GPT2 models are still under-optimized/over-parameterized (relative to the data on which they were trained) and that they have capacity to support datasets even larger than the recent XL $1.5B$ release~\cite{gpt2-xl}.

### 2020-09-04 submission/NatCom_abstract_file.tex
initial pass to modify for nat comm

In many applications, one works with neural network models trained by someone else.
For such \emph{pretrained models}, one often does not have access to training data or test data.
Moreover, one may not know details about the model, e.g., the specifics of the training data, the loss function, the hyperparameter values, etc.
Given one or many pretrained models, can one say anything about the expected performance or quality of the models?
Here, we address this question by providing a detailed meta-analysis of hundreds of publicly-available pretrained models.
We examine norm based capacity control metrics as well as power law based metrics from the recently-developed Theory of Heavy-Tailed Self Regularization.
We find that norm based metrics correlate well with reported test accuracies for well-trained models, but that they often can not distinguish well-trained versus poorly-trained models.
We also find that power law based metrics do much better---qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among series of well-trained models.
These methods can be used to identify when a pretrained neural network has problems that can not be detected simply by examining training/test accuracies.

### 2020-09-04 submission/NatCom_all_models.tex
initial pass to modify for nat comm

--- 2020-07-02 submission/all_models.tex
@@
-%KDD% 
-\vspace{-1mm}
-\section{Comparing Hundreds of Models}
+\subsection{Comparing Hundreds of Models}
 \label{sxn:all_cv_models}
-%\vspace{-1mm}
 
-In this section, we summarize results from a large-scale analysis of hundreds of publicly-available  models, including models developed for image 
-classification, segmentation, and a range of related tasks.  
-Our aim is to provide a broad analysis that complements the detailed analysis in Sections~\ref{sxn:cv} and~\ref{sxn:nlp}.
-There, we considered CV and NLP models with similar structures. 
-Here, we provide much broader conclusions based on a much larger set of models with a much more diverse set of structures. % about norm-based and PL-based metrics.
-The models we consider have been pretrained on nine datasets.   %, shown in Table~\ref{table:datasets}.
-%including ImageNet-1K, and CIFAR-10, CIFAR-100, Street View House Numbers (SVHN), Caltech-UCSD Birds-200-2011 (CUB-200-2011), Pascal VOC2012, ADE20K, Cityscapes, and Common Objects in Context (COCO).
-%%The pretrained models and their accuracy metrics are summarized in the osmr github~\cite{charles-add-link}.
-We provide full details about how to reproduce these results in Appendix~\ref{sxn:appendix}.
+We summarize results from a large-scale analysis of hundreds of publicly-available  models.
+This broader analysis on a much larger set of CV and NLP models, with a more diverse set of architectures, and developed for a wider range of tasks, complements the previous more detailed analysis on CV and NLP models.
+See Appendix~\ref{sxn:appendix} for details.
+To quantify the relationship between quality metrics and the reported test error and/or accuracy metrics, we use ordinary least squares to regress the metrics on the Top1 (and Top5) reported errors (as dependent variables).
+These include Top5 errors for the ImageNet-1K model, percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy (Pix.Acc.) and Intersection-Over-Union (IOU) for other models.
+We regress them individually on each of the norm-based and PL-based metrics.
 
@@
          We include regressions only for architectures with five or more data points, and which are positively correlated with test error.
-         These results can be readily reproduced using the Google Colab notebooks (see Appendix~\ref{sxn:appendix}). 
+         These results can be readily reproduced using the Google Colab notebooks. (See Appendix~\ref{sxn:appendix} for details.) 
         }
@@
 
-We choose ordinary least squares (OLS) regression to quantify the relationship between quality metrics (computed with the \emph{WeightWatcher} tool) 
-and the reported test error and/or accuracy metrics.
-We regress the metrics on the Top1 (and Top5) reported errors (as dependent variables).
-These include Top5 errors for the ImageNet-1K model, percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy
-(Pix.Acc.) and Intersection-Over-Union (IOU) for other models.
-We regress them individually on each of the norm-based and PL-based metrics, as described in Section~\ref{sxn:cv}.
-
-Our results are summarized in Table~\ref{table:results}.
+Results are summarized in Table~\ref{table:results}.
 For the mean, larger $R^{2}$ and smaller $MSE$ are desirable; and for the standard deviation, smaller values are desirable.
@@
 and PL-based metrics are much better for $MSE$ mean and standard deviation.
-These (and other) results suggest our conclusions from Sections~\ref{sxn:cv} and~\ref{sxn:nlp} hold much more generally, and they suggest
-obvious questions for future work.
+These and other results suggest our conclusions hold much more generally.

### 2020-09-04 submission/NatCom_appendix.tex
initial pass to modify for nat comm

--- 2020-02-16 submission/appendix.tex
@@
-\section{Appendix}
+\section{Reproducibility Appendix}
 \label{sxn:appendix}
@@
 More details and more results are available in an accompanying github repository~\cite{kdd20_sub_repo}.
+
+\subsection{Additional Details on Layer Weight Matrices}
+
+Recall that we can express the objective/optimization function for a typical DNN with $L$ layers, with activation functions $h_{l}(\cdot)$, and with $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$, as:
+\begin{equation*}
+E_{DNN} \hspace{-1mm} = \hspace{-1mm} h_{L}(\mathbf{W}_{L}\cdot h_{L-1}(\mathbf{W}_{L-1}\cdot h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})  .
+\end{equation*}
+We expect that most well-trained, production-quality models will employ one or more forms of regularization, such as Batch Normalization (BN), Dropout, etc., and many will also contain additional structure such as Skip Connections, etc.
+Here, we will ignore these details, and will focus only on the pretrained layer weight matrices $\mathbf{W}_{l}$.
+Typically, this model would be trained on some labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using Backprop, by minimizing the loss $\mathcal{L}$.
+For simplicity, we do not indicate the structural details of the layers (e.g., Dense or not, Convolutions or not, Residual/Skip Connections, etc.). 
+Each layer is defined by one or more layer 2D weight matrices $\mathbf{W}_{l}$, and/or the 2D feature maps $\mathbf{W}_{l,i}$ extracted from 2D Convolutional (Conv2D) layers.
+A typical modern DNN may have anywhere between 5 and 5000 2D layer matrices.
+
+For each Linear Layer, we get a  single $(N\times M)$ (real-valued) 2D weight matrix, denoted $\mathbf{W}_{l}$, for layer $l$.  
+This includes Dense or Fully-Connected (FC) layers, as well as 1D Convolutional (Conv1D) layers, Attention matrices, etc.
+We ignore the bias terms $\mathbf{b}_{l}$ in this analysis. 
+Let the aspect ratio be $Q=\frac{N}{M}$, with $Q\ge 1$.
+For the Conv2D layers, we have a 4-index Tensor, of the form $(N\times M \times c\times d)$, consisting
+of $c\times d$ 2D feature maps of shape $(N\times M)$.    
+We  extract $n_{l}=c\times d$ 2D weight matrices $\mathbf{W}_{l,i}$, one for each feature map $i=[1,\dots,n_{l}]$ for layer $l$.
+
+\subsection{Additional Details on HT-SR Theory}
+
+The original work on HT-SR Theory~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM} considered NNs including AlexNet and InceptionV3 (as well as DenseNet, ResNet, and VGG), and it showed that for nearly every $\mathbf{W}$, the (bulk and tail) of the ESDs can be fit to a truncated PL and the PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$.
+Our meta-analysis, the main results of which are summarized in this paper, has shown that these results are ubiquitous.
+For example, 
+upon examining nearly 10,000 layer weight matrices $\mathbf{W}_{l,i}$ across hundreds of different modern pre-trained DNN architectures, the ESD of nearly every $\mathbf{W}$ layer matrix can be fit to a truncated PL:
+$70-80\%$ of the time, the fitted PL exponent $\alpha$ lies in the range $\alpha\in(2,4)$; and  
+$10-20\%$ of the time, the fitted PL exponent $\alpha$ lies in the range $\alpha< 2$.  
+Of course, there are exceptions: in any real DNN, the fitted $\alpha$ may range anywhere from $\sim 1.5$ to $10$ or higher (and, of course, larger values of $\alpha$ may indicate that the PL is not a good model for the data).
+Still, overall, in nearly all large, pre-trained DNNs, the correlations in the  weight matrices exhibit a remarkable Universality, being both Heavy Tailed, and having small---but not too small---PL exponents.
 

### 2020-09-04 submission/NatCom_cv_models.tex
initial pass to modify for nat comm

--- 2020-07-02 submission/cv_models.tex
@@
-%KDD% 
-\vspace{-1mm}
-\section{Comparison of CV models}
+\subsection{Comparison of CV models}
 \label{sxn:cv}
-%\vspace{-1mm}
 
-In this section, we examine empirical quality metrics described in Section~\ref{sxn:methods} for several CV model architecture series.
-This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current open source pyTorch framework (version 1.4)~\cite{pytorch}.
-This also includes a larger set of ResNet models, trained on the ImageNet-1K dataset~\cite{imagenet}, provided on the OSMR ``Sandbox for training convolutional networks for computer vision''~\cite{osmr}, which we call the ResNet-1K series.
+Each of the VGG, ResNet, and DenseNet series of models consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each is distributed with the current open source pyTorch framework (version 1.4)~\cite{pytorch}.
+In addition, we examine a larger set of ResNet models, which we call the ResNet-1K series, trained on the ImageNet-1K dataset~\cite{imagenet} and provided on the OSMR Sandbox~\cite{osmr}.
+For these models, we first perform coarse model analysis, comparing and contrasting the four model series, and predicting trends in model quality.
+We then perform fine layer analysis, as a function of depth.
+This layer analysis goes beyond predicting trends in model quality, instead illustrating that PL-based metrics can provide novel insights among the VGG, ResNet/ResNet-1K, and DenseNet architectures.
 
-We first perform \emph{coarse model analysis}, comparing and contrasting the four model series, and predicting trends in model quality. 
-We then perform \emph{fine layer analysis}, as a function of depth for these models.
-This goes beyond simply predicting trends in model quality, instead illustrating that PL-based metrics can provide novel insights among the VGG, ResNet/ResNet-1K, and DenseNet architectures.
-
-%KDD% 
-\vspace{-1mm}
 \paragraph{Average Quality Metrics versus Reported Test Accuracies.}
 
-We have examined the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
-To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BN), and it plots the four quality metrics versus the reported test accuracies~\cite{pytorch},%
-\footnote{That is, these test accuracies have been previously reported and made publicly-available by others.
-We take them as given, and we do not attempt to reproduce/verify them, since we do not permit ourselves any access to training/test data.}
-as well as a basic linear regression line. 
-All four metrics correlate quite well with the reported Top1 accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalization (i.e., greater accuracy, lower error). 
-While all four metrics perform well, notice that the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) performs best (with an RMSE of $0.42$, see Table~\ref{table:cv-models}); and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $), which is an approximation to the Log $\alpha$-Norm metric~\cite{MM20_unpub_work}, performs second best (with an RMSE of $0.48$, see Table~\ref{table:cv-models}).
+We examine the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
+Figure~\ref{fig:vgg-metrics} 
+plots the four quality metrics versus reported test accuracies~\cite{pytorch},% 
+\footnote{These test accuracies have been previously reported and made publicly-available by others.
+We take them as given.
+We do not attempt to reproduce/verify them, since we do not permit ourselves access to training/test data.}
+as well as a basic linear regression line, for the VGG series.
+All four metrics correlate quite well with reported Top1 accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalization (i.e., greater accuracy, lower error). 
+The Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) performs best (RMSE of $0.42$, see Table~\ref{table:cv-models}); and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $) performs second best (RMSE of $0.48$, see Table~\ref{table:cv-models}).
 
@@
         \includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
-        %KDD% \includegraphics[width=4.0cm]{img/VGG_logpnorm_accs.png}
         \label{fig:vgg-pnorm}
     }
-    \caption{Comparison of Average Log Norm and Weighted Alpha quality metrics versus reported test accuracy for pretrained VGG models (with and without BN), trained on ImageNet, available in pyTorch (v1.4).
-Metrics fit by linear regression, RMSE reported.
-}
+    \caption{Comparison of Average Log Norm and Weighted Alpha quality metrics versus reported test accuracy for pretrained VGG models: 
+             VGG11, VGG13, VGG16, and VGG19, with and without Batch Normalization (BN),
+              trained on ImageNet, available in pyTorch (v1.4).  Metrics fit by linear regression, RMSE reported. }
     \label{fig:vgg-metrics}
@@
 
-See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics for the VGG, ResNet, and DenseNet series.
+See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics for the VGG, ResNet, ResNet-1K, and DenseNet series.
 Similar results (not shown) are obtained for the Top5 accuracies.
-Overall, for the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best.
-These model series are all very well-trodden, and thus our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of well-trained versus more-well-trained models, with PL-based metrics performing somewhat (i.e., quantitatively) better.
-
-The DenseNet series has similar behavior to what we see in Figures~\ref{fig:vgg-metrics} and~\ref{fig:cv2-accuracy} for the other models.
-However, as noted in Table~\ref{table:cv-models}, it has only 4 data points.  
-In our larger analysis, in Section~\ref{sxn:all_cv_models}, we will only include series with 5 or more models.
-%For completeness and reproducibility here, the DenseNet Log $\alpha$-Norm plot is given in Appendix~\ref{sxn:appendix}.
+Overall, all metrics perform relatively well; the Weighted Alpha metric performs best; and the Log $\alpha$-Norm metric performs second best.
+These model series are all very well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of well-trained versus very-well-trained models, with PL-based metrics performing somewhat (i.e., quantitatively) better.
+The DenseNet series has similar behavior.
 (These and many other such plots can be seen on our publicly-available~repo.)
 
-%KDD% 
-\vspace{-1mm}
 \paragraph{Variation in Data Set Size.}
 
-We are interested in how our four quality metrics depend on data set size.
-To examine this, we look at results on ResNet versus ResNet-1K.
-See Figure~\ref{fig:cv2-accuracy}, which plots and compares the Log $\alpha$-Norm metric 
-%% MM: From Table 1 it is slightly worse than alpha-hat and we say that below %% --the best performing metric--
-for the full ResNet model, trained on the full ImageNet dataset, against the ResNet-1K model, which has been trained on a much smaller ImageNet-1K data set.
-The Log $\alpha$-Norm is much better than the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is actually slightly worse than the Weighted Alpha metric).
+To examine how the four quality metrics depend on data set size, we look at results on ResNet versus ResNet-1K.
+Figure~\ref{fig:cv2-accuracy} compares the Log $\alpha$-Norm metric for the full ResNet model, trained on the full ImageNet dataset, against the ResNet-1K model, trained on a much smaller ImageNet-1K data set.
+The Log $\alpha$-Norm is much better than the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is slightly worse than the Weighted Alpha metric).
 The ResNet series has strong correlation, with an RMSE of $0.66$, whereas the ResNet-1K series also shows good correlation, but has a much larger RMSE of $1.9$.
 (Other metrics exhibit similar behavior.)
-As expected, the higher quality data set shows a better fit, even with fewer data~points.
+The higher quality data set shows a better fit, even with fewer data~points.
 
-%KDD% 
-\vspace{-1mm}
 \paragraph{Layer Analysis: Metrics as a Function of Depth.}
@@
 We can learn much more about a pretrained model by going beyond average values of quality metrics to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).
-% in the network.
-%The most interesting results are seen when we 
-For example, we can 
-plot (just) the PL exponent, $\alpha$, for each layer, %
-%\footnote{That is, here we consider just $\alpha$ for each layer, i.e., not $\hat{\alpha}$ or $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$.} 
-as a function of depth.
-%
-See Figure~\ref{fig:3models-alpha-layers}, which plots $\alpha$ for each layer (the first layer corresponds to data, the last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (no BN), ResNet, and DenseNet series.
-(Again, many more such plots are available at our repo; but note that the corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than the results we present here.)
+Figure~\ref{fig:3models-alpha-layers} plots the PL exponent $\alpha$, as a function of depth, for each layer (first layer corresponds to data, last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (no BN), ResNet, and DenseNet series.
+(Many more such plots are available at our repo.)
 
@@
         \includegraphics[width=4.9cm]{img/VGG_fnl_alpha_depth.png} 
-        %KDD% \includegraphics[width=4.0cm]{img/VGG_fnl_alpha_depth.png} 
                 \label{fig:vgg-alpha-layers}
     }
-    %KDD% 
     \qquad
@@
         \includegraphics[width=4.9cm]{img/ResNet_fnl_alpha_depth.png} 
-        %KDD% \includegraphics[width=4.0cm]{img/ResNet_fnl_alpha_depth.png} 
         \label{fig:resnet-alpha-layer}
     }
-    %KDD% 
     \qquad
@@
         \includegraphics[width=4.9cm]{img/DenseNet_fnl_alpha_depth.png} 
-        %KDD% \includegraphics[width=4.0cm]{img/DenseNet_fnl_alpha_depth.png} 
         \label{fig:densenet-alpha-layer}
     }    
-    %KDD% 
     \qquad
@@
         \includegraphics[width=4.9cm]{img/resnet_alpha_overlaid_depth.png} 
-        %KDD% \includegraphics[width=4.0cm]{img/resnet_alpha_overlaid_depth.png} 
         \label{fig:resnet_alpha_overlaid_depth}
@@
              This contrasts with how both VGG models gradually worsen in deeper layers and how the DenseNet models are much more erratic.  
-             In the text, this is interpreted in terms of \emph{Correlation Flow}.
+             In the text, this is interpreted in terms of Correlation Flow.
             }
@@
 In the VGG models, Figure~\ref{fig:vgg-alpha-layers} shows that the PL exponent $\alpha$ systematically increases as we move down the network, from data to labels, in the Conv2D layers, starting with $\alpha\lesssim 2.0$ and reaching all the way to $\alpha\sim 5.0$; and then, in the last three, large, fully-connected (FC) layers, $\alpha$ stabilizes back down to $\alpha\in[2,2.5]$.
-This is seen for all the VGG models (again, only the shallowest and deepest are shown in this figure), indicating that the main effect of increasing depth is to increase the range over which $\alpha$ increases, thus leading to larger $\alpha$ values in later Conv2D layers of the VGG models.
+This is seen for all the VGG models (again, only the shallowest and deepest are shown), indicating that the main effect of increasing depth is to increase the range over which $\alpha$ increases, thus leading to larger $\alpha$ values in later Conv2D layers of the VGG models.
 This is quite different than the behavior of either the ResNet-1K models or the DenseNet models.
 
-For the ResNet-1K models, Figure~\ref{fig:resnet-alpha-layer} shows that $\alpha$ also increases in the last few layers (more dramatically, in fact, than for VGG, observe the differing scales on the Y axes).
-However, as the ResNet-1K models get deeper, there is a wide range over which $\alpha$ values tend to remain quite small.
-%\michael{Can we say something about exceptions which are larger than for VGG or DenseNet.}
-%\charles{You have the same plots I do..what do you want to say >}
-%\michael{Are they small or different types of layers or something?}
-%\charles{Not that I am aware of.  Ill think on it}
+For the ResNet-1K models, Figure~\ref{fig:resnet-alpha-layer} shows that $\alpha$ also increases in the last few layers (more dramatically than for VGG, observe the differing scales on the Y axes).
+However, as the ResNet-1K models get deeper, there is a wide range over which $\alpha$ values tend to remain small.
 This is seen for other models in the ResNet-1K series, but it is most pronounced for the larger ResNet-1K (152) model, where $\alpha$ remains relatively stable at $\alpha\sim 2.0$, from the earliest layers all the way until we reach close to the final layers.
@@
 For the DenseNet models, Figure~\ref{fig:densenet-alpha-layer} shows that $\alpha$ tends to increase as the layer id increases, in particular for layers toward the end.
-While this is similar to what is seen in the VGG models, with the DenseNet models, $\alpha$ values increase almost immediately after the first few layers, and the variance is much larger (in particular for the earlier and middle layers, where it can range all the way to $\alpha\sim 8.0$) and much less systematic throughout the network.
+While this is similar to the VGG models, with the DenseNet models, $\alpha$ values increase almost immediately after the first few layers, and the variance is much larger (in particular for the earlier and middle layers, where it can range all the way to $\alpha\sim 8.0$) and much less systematic throughout the network.
 
-%Recall the differences---in particular, the number of residual connections---between the VGG, ResNet, and DenseNet architectures.
-%VGG, while a fine model in many ways, was limited by needing massive FC layers at the end of the model, requiring significant memory and computational resources, and it's poor conditioning led to problems with vanishing gradients.
-%The empirical manifestations of this (on weight matrices) are that fitted $\alpha$ values get much worse for deeper layers.
+\paragraph{Comparison of VGG, ResNet, and DenseNet Architectures.}
 
-%KDD% 
-\vspace{-1mm}
-\paragraph{Comparison of VGG, ResNet, and DenseNet Architectures.}
-We can interpret these layer-wise observations by recalling the architectural differences between the VGG, ResNet, and DenseNet architectures,
-and, in particular, the number of of residual connections. VGG resembles the  traditional convolutional architectures,
-such as LeNet5, and consists of several [Conv2D-Maxpool-ReLu] blocks, followed by 3 large Fully Connected (FC) layers.
-ResNet greatly improved on VGG by replacing the large FC layers, shrinking the Conv2D blocks, and introducing \emph{residual connections}.
-This optimized approach allows for greater accuracy with far fewer parameters (and GPU memory requirements), and
-ResNet models of up to 1000 layers have been trained~\cite{resnet1000}.
+These layer-wise observations can be understood in terms of architectural differences between VGG, ResNet, and DenseNet.
+VGG resembles the traditional convolutional architectures, such as LeNet5, and consists of several [Conv2D-Maxpool-ReLu] blocks, followed by 3 large Fully Connected (FC) layers.
+ResNet greatly improved on VGG by replacing the large FC layers, shrinking the Conv2D blocks, and introducing residual connections.
+This optimized approach allows for greater accuracy with far fewer parameters, and ResNet models of up to 1000 layers have been trained~\cite{resnet1000}.
 
-We conjecture that the efficiency and effectiveness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and more strongly optimized.
-Contrast this with the DenseNet models, which contains many connections between every layer.
-Our results (large $\alpha$, meaning that even a PL model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly~optimized.
+The efficiency and effectiveness of ResNet seems to be reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and more strongly optimized.
+This contrasts with the DenseNet models, which contains many connections between every layer.
+These results (large $\alpha$, meaning that even a PL model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly~optimized.
 
@@
 
-%KDD% 
-\vspace{-1mm}
 \paragraph{Correlation Flow.}
 
-We can understand the results presented in Figure~\ref{fig:3models-alpha-layers} in terms of what we will call the \emph{Correlation Flow} of the model. 
-Recall that the average Log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on ideas from the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
+Figure~\ref{fig:3models-alpha-layers} can be understood in terms of what we will call \emph{Correlation Flow}.
+The average Log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
 There, one expects that the weight matrices of well-trained DNNs will exhibit correlations over many size scales, as is well-known in other strongly-correlated systems~\cite{BouchaudPotters03, SornetteBook}.
-In particular, this would imply that their ESDs can be well-fit by a truncated PL, with exponents $\alpha\in[2,4]$.
+This would imply that their ESDs can be well-fit by a truncated PL, with exponents $\alpha\in[2,4]$.
 Much larger values $(\alpha\gg 6)$ may reflect poorer PL fits, whereas smaller values $(\alpha\sim 2)$, are associated with models that generalize better.
-%, but the presence of many adjacent layers with relatively good PL fits \emph{suggests} that those adjacent layers ``impedance match'' well and that the correlations that are captured by one layer propagate well to subsequent layers.
 
 Informally, one would expect a DNN model to perform well when it facilitates the propagation of information/features across layers.
-Previous work argues this by computing the gradients over the input data.
-In the absence of training/test data, one might hypothesize that this flow of information leaves empirical signatures on weight matrices, and thus that we can quantify this by measuring the PL properties of weight matrices.
-In this case, given the statistical mechanical basis of HT-SR Theory, smaller $\alpha$ values would correspond to layers in which information correlations between data across multiple scales are better captured~\cite{MM18_TR,SornetteBook}, and thus we hypothesize that small $\alpha$ values that are stable across multiple layers enable better \emph{correlation flow} through the network.
-We have seen this in many models, including those shown in Figure~\ref{fig:3models-alpha-layers}.
-
-This is superficially similar to recent work on the information bottleneck~\cite{ST17_TR}, except that here we work in an entirely unsupervised setting.
-Developing theory to evaluate this hypothesis and exploit this insight is an intriguing direction for future work.
+In the absence of training/test data, one might hypothesize that this flow of information leaves empirical signatures on weight matrices, and that we can quantify this by measuring the PL properties of weight matrices.
+In this case, smaller $\alpha$ values correspond to layers in which information correlations between data across multiple scales are better captured~\cite{MM18_TR,SornetteBook}.
+This leads to the hypothesis that small $\alpha$ values that are stable across multiple layers enable better correlation flow through the network.
+This is similar to recent work on the information bottleneck~\cite{ST17_TR}, except that here we work in an entirely unsupervised setting.
 
@@
 
-%KDD% 
-\vspace{-1mm}
 \paragraph{Scale Collapse; or How Distillation May Break Models.}
 
-The similarity between norm-based metrics and PL-based metrics suggests a question: is the Weighted Alpha metric just a variation of the more familiar norm-based metrics?
-%%, or does it capture something different?
+The similarity between norm-based metrics and PL-based metrics suggests a question: Is the Weighted Alpha metric just a variation of more familiar norm-based metrics?  
 More generally, do fitted $\alpha$ values contain information not captured by norms? 
-In examining hundreds of pretrained models, we have found several anomalies that are ``exceptions that prove the rule'' and that demonstrate the power of our approach.
-In particular, to show that $\alpha$ does capture something different, consider the following example, which looks at a compressed/distilled DNN model~\cite{CWZZ17_TR}.
-In this example, we show that some distillation methods may actually \emph{break} models unexpectedly by introducing what we call \emph{Scale Collapse}, where several distilled layers have unexpectedly small Spectral Norms.
+Among hundreds of pretrained models, there are ``exceptions that prove the rule.'' 
+To show that $\alpha$ does capture something different, we show that some compression/distillation methods \cite{CWZZ17_TR} may actually damage models unexpectedly, by introducing what we call \emph{Scale Collapse}, where several distilled layers have unexpectedly small Spectral Norms.
 
-We consider ResNet20, trained on CIFAR10, before and after applying the Group Regularization distillation technique, as implemented in the \texttt{distiller} package~\cite{distiller}.
+Consider ResNet20, trained on CIFAR10, before and after applying the Group Regularization distillation technique, as implemented in the \texttt{distiller} package~\cite{distiller}.
 We analyze the pretrained 4D\_regularized\_5Lremoved baseline and fine-tuned models. 
 The reported baseline test accuracies (Top1$=91.45$ and Top5$=99.75$) are better than the reported fine-tuned test accuracies (Top1$=91.02$ and Top5$=99.67$).  
-Because the baseline accuracy is greater,  the previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy}) suggest that the baseline Spectral Norms should be \emph{smaller} on average than the fine-tuned ones.
-\emph{The opposite is observed.}
-%\michael{Do we have something backwards here: where we observed that the average norm to increase with decreasing test error (as it ``should'' not), whereas the average PL exponent $\alpha$ decreases (as ``expected'' from HT-SR Theory).
-}
-%In both cases (Frobenius norm results not shown), we observe the opposite.
-%%\charles{WHy is this confusing? Call me...}
-Figure~\ref{fig:resnet204D5L} presents the Spectral Norm (here denoted $\log\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}$.%
-\footnote{Here, we only include layer matrices or feature maps with $M\ge50$.}
+Because the baseline accuracy is greater,  the previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy}) suggest that the baseline Spectral Norms should be smaller on average than the fine-tuned ones.
+The opposite is observed.
+Figure~\ref{fig:resnet204D5L} presents the Spectral Norm (here denoted $\log\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}$.
 On the other hand, the $\alpha$ values (in Figure~\ref{fig:resnet204Dalpha}) do not differ systematically between the baseline and fine-tuned models.
-%\michael{Some comment about big jump in norm; but fix previous backwards question first.}--SEE BELOW
-Also (not shown), the average (unweighted) baseline $\alpha$ is \emph{smaller} than the fine-tuned average (as predicted by HT-SR Theory, the basis of $\hat{\alpha}$).
-
-That being said, Figure~\ref{fig:resnet204Dalpha} also depicts two very large $\alpha\gg 6$ values for the baseline, but not for the fine-tuned, model.
+Also (not shown, but as predicted by HT-SR Theory, the basis of $\hat{\alpha}$), the average (unweighted) baseline $\alpha$ is smaller than the fine-tuned average.
+%
+In spite of this, Figure~\ref{fig:resnet204Dalpha} also depicts two very large $\alpha\gg 6$ values for the baseline, but not for the fine-tuned, model.
 This suggests the baseline model has at least two over-parameterized/under-trained layers, and that the distillation method does, in fact, improve the fine-tuned model by compressing these layers.
 
-The obvious interpretation is the following. 
-Pretrained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the 
-VGG, ResNet, or DenseNet series.  %% we considered above.
-While norms make good regularizers for a single model, there is no reason \emph{a priori} to expect them correlate so well with test accuracies across different models, and they may not differentiate well-trained versus poorly-trained models.
+Pretrained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the VGG, ResNet, or DenseNet series.  
+The obvious interpretation is that, while norms make good regularizers for a single model, there is no reason \emph{a priori} to expect them correlate well with test accuracies across different models, and they may not differentiate well-trained versus poorly-trained models.
 We do expect, however, the PL $\alpha$ to do so, because it effectively measures the amount of information correlation in the model~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
-Among other things, this suggests that the $\alpha$ values will improve, i.e., decrease, over time, as distillation techniques continue to improve.
+This suggests that the $\alpha$ values will improve, i.e., decrease, over time, as distillation techniques continue to improve.
 The reason for the anomalous behavior shown in 

### 2020-09-04 submission/NatCom_discussion.tex
initial pass to modify for nat comm

\section{Discussion}

\paragraph{Conclusions.}

We have developed and evaluated methods to predict trends in the quality of state-of-the-art neural networks---without access to training or testing data.
Our main methodology involved weight matrix meta-analysis, using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package}, and informed by the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
Prior to our work, it was not even obvious that norm-based metrics would perform well to predict trends in quality \emph{across} models (as they are usually used \emph{within} a given model or parameterized model class, e.g., to bound generalization error or to construct regularizers).
Our results are the first to demonstrate that they can be used for this important practical problem.
%
Our results also demonstrate that PL-based metrics perform better than norm-based metrics. 
This should \emph{not} be surprising---at least to those familiar with the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}---since our use of PL exponents is designed to capture the idea that well-trained models capture information correlations over many size scales in the data.
Again, though, our results are the first to demonstrate this.
%
Our approach can also be used to provide fine-scale insight (rationalizing the flow of correlations or the collapse of size scale) throughout a network. 
Both \emph{Correlation Flow} and \emph{Scale Collapse} are important for improved diagnostics on pretrained models as well as for improved training methodologies.

\paragraph{Related work.}

Statistical mechanics has long had influence on DNN theory and practice~\cite{EB01_BOOK, MM17_TR, BKPx20}.
Our best-performing PL-based metrics are based on statistical mechanics via HT-SR Theory~\cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM19_KDD, MM20_SDM}.
The way in which we (and HT-SR Theory) \emph{use} statistical mechanics theory is quite different than the way it is more commonly formulated~\cite{EB01_BOOK, BKPx20}.
Going beyond idealized models, we \emph{use} statistical mechanics in a broader sense, drawing upon techniques from quantitative finance, random matrix theory, and the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
%
There is also a large body of work in ML on using norm-based metrics to bound generalization error~\cite{NTS15, BFT17_TR, LMBx18_TR}.
This theoretical work aims to prove generalization bounds, and this applied work then uses these norms to construct regularizers to improve training.
Proving generalization bounds and developing new regularizers is very different than our focus of validating pretrained models.

Our work, starting in 2018 with the \emph{WeightWatcher} tool~\cite{weightwatcher_package}, is the first to perform a detailed analysis of the weight matrices of DNNs~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
Subsequent to the initial version of this paper, we became aware of two other works that were posted to the in 2020 within weeks of the initial version of this paper~\cite{EJRUY20_TR,UKGBT20_TR}.
Both of these papers validate our basic result that one can gain substantial insight into model quality by examining weight matrices without access to any training or testing data.
However, both consider smaller models drawn from a much narrower range of applications than we consider.
Previous results in HT-SR Theory suggest that insights from these smaller models may not extend to the state-of-the-art CV and NLP models we consider.

\paragraph{Looking forward.}

We conclude with a few comments on what a \emph{practical theory} of DNNs should look like.
To do so, we distinguish between two types of theories:
\emph{non-empirical or analogical theories}, in which one creates, often from general principles, a very simple toy model that can be analyzed rigorously, and one then argues that the model is relevant to the system of interest; and 
\emph{semi-empirical theories}, in which there exists a rigorous asymptotic theory, which comes with parameters, for the system of interest, and one then adjusts or fits those parameters to the finite non-asymptotic data.
A drawback of the former approach is that it typically makes very strong assumptions, and the strength of those assumptions can limit the practical applicability of the theory.
Nearly all of the work on DNN theory focuses on the former type of theory.
Our approach focuses on the latter type of theory.
Our results, which are based on \emph{using} sophisticated statistical mechanics theory and \emph{solving} important practical DNN problems, suggests that the latter approach should be of interest more generally for those interested in developing a \emph{practical DNN~theory}.

### 2020-09-04 submission/NatCom_introduction.tex
initial pass to modify for nat comm

--- 2020-07-06 submission/introduction.tex
@@
 and so on.
-A less well-known problem, but one that is increasingly important, in particular in industrial-scale artificial intelligence (AI), arises when the model \emph{user} is not the model \emph{developer}.
-Here, one may not have access to either the training data or the testing data.
-Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it as-is, or to fine-tune and/or compress it and then use it.
+A related problem, in particular in industrial-scale artificial intelligence (AI), arises when the model user is not the model developer.
+Then, one may not have access to the training data or the testing data.
+Instead, one may simply be given a model that has already been trained---a pretrained model---and need to use it as-is, or to fine-tune and/or compress it and then use it.
 
 Na\"{\i}vely---but in our experience commonly, among ML practitioners and ML theorists---if one does not have access to training or testing data, then one can say absolutely nothing about the quality of a ML model.
-This may be true in worst-case theory, but models are used in practice, and there is a need for a \emph{practical theory} to guide that practice.
+This may be true in worst-case theory, but models are used in practice, and there is a need for a practical theory to guide that practice.
 Moreover, if ML is to become an industrial process, then that process will become compartmentalized in order to scale: some groups will gather data, other groups will develop models, and other groups will use those models.
@@
 It is of great practical interest to have metrics to evaluate the quality of a trained model---in the absence of training/testing data and without any detailed knowledge of the training/testing process.
-There is a need for a \emph{practical theory} for pretrained models which can predict how, when, and why such models can be expected to perform well or~poorly.
+There is a need for a practical theory for pretrained models which can predict how, when, and why such models can be expected to perform well or~poorly.
 
-In this paper, we provide a detailed empirical analysis to address these questions, evaluating quality metrics for pretrained deep neural network (DNN) models, and we do so at scale.
+In the absence of training and testing data, obvious quantities to examine are the weight matrices of pretrained models, e.g., 
+properties such as norms of weight matrices and/or parameters of Power Law (PL) fits of the eigenvalues of weight matrices.
+Norm-based metrics have been used in traditional statistical learning theory to bound capacity and construct regularizers; and PL fits are based on statistical mechanics approaches to deep neural networks (DNNs).
+While we use traditional norm-based and PL-based metrics, our goals are not the traditional goals.
+Unlike more common ML approaches, we do not seek a bound on the generalization (e.g., by evaluating training/test errors), we do not seek a new regularizer, and we do not aim to evaluate a single model (e.g., as with hyperparameter optimization).
+Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves.
+In both cases, we ask:
+Can we predict trends in the quality of pretrained DNN models without access to training or testing data?
+
+To answer this question, we provide a detailed empirical analysis, evaluating quality metrics for pretrained DNN models, and we do so at scale.
 Our approach may be viewed as a statistical meta-analysis of previously published work, where we consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
+By now, there are many such state-of-the-art models that are publicly-available, e.g., 
+hundreds of pretrained models in CV ($\ge 500$) and NLP ($\approx 100$).%
+\footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
+For all these models, we have no access to training data or testing data, and we have no specific knowledge of the training/testing protocols. 
 %
-By now, there are many such state-of-the-art models that are publicly-available, e.g., 
-there are now hundreds of pretrained models in CV ($\ge 500$) and NLP ($\approx 100$).%
-\footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
-These publicly-available models provide a large corpus of models that by some community standard are state-of-the-art. %
-%%%\footnote{Clearly, there is a selection bias or survivorship bias here---people tend not to make publicly-available their poorly-performing models---but these models are things in the world that (like social networks or the internet) can be analyzed for their properties.}
-Importantly, all of these models have been trained by someone else and have been viewed to be of sufficient interest/quality to be made publicly-available; and, for all of these models, we have no access to training data or testing data, and we have no specific knowledge of the training/testing protocols.
-
-In the absence of training and testing data, obvious quantities to examine are the weight matrices of pretrained models.
-Thus,
-the \emph{quality metrics} we consider are based on the spectral properties of the layer weight matrices.
-In particular, they are based on norms of weight matrices 
-and/or parameters of power law (PL) fits of the eigenvalues of weight matrices.
-Such norm-based metrics have been used in traditional statistical learning theory to bound capacity and construct regularizers; and such PL fits are based on statistical mechanics approaches to DNNs.
-While we use traditional norm-based and PL-based metrics, our goals are not the traditional goals.
-Unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., as with hyperparameter optimization).%
-\footnote{One could of course use these techniques to improve training, and we have been asked about that, but we are not interested in that here.
-Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained DNN models.}
-Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves, and in both cases, we ask:
-\begin{quote}
-\emph{Can we predict trends in the quality of pretrained DNN models without access to training or testing data?}  
-\end{quote}
-
-%\begin{itemize}[leftmargin=*]
-%\item 
-%First, 
-%Motivated by practical problems, we formulate the question ``Can one predict trends in the quality of state-of-the-art neural networks without access to training or testing data''
-%\item
-%Second, 
-%To answer this question, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in CV and NLP.
-%\end{itemize}
-
-To answer this question, we perform a statistical meta-analysis, where we 
-perform weight matrix analysis on %%%analyze 
-hundreds of publicly-available pretrained state-of-the-art CV and NLP models. 
 Here is a summary of our main results.
-\begin{itemize}
-\item
-\textbf{Norm-based metrics and well-trained models.}
-Norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models.
-\item
-\textbf{Norm-based metrics and poorly-trained models.}
-Norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.).
+First, norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models.
+Second, norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.).
 For example, they may exhibit what we call \emph{Scale Collapse} for these models. 
-\item 
-\textbf{PL-based metrics and model quality.}
-PL-based metrics do much better at predicting quality trends in pretrained CV/NLP models.  
+Third, PL-based metrics do much better at predicting quality trends in pretrained CV/NLP models.  
 They are qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among a series of well-trained versus very-well-trained models.
-\item 
-\textbf{PL-based metrics and model diagnostics.}
-PL-based metrics can also be used to characterize fine-scale model properties (including what we call layer-wise \emph{Correlation Flow}) in well-trained and poorly-trained models, and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
-\end{itemize}
-
-\noindent
-In light of our main results, several points are worth emphasizing.
-First, our goal is a practical theory to predict trends in the quality of state-of-the-art DNN models, i.e., not to make a statement about every publicly-available model.
-We have examined hundreds of models, and we identify general trends, but we also highlight interesting exceptions.
-%%Several of the most interesting are described below.
-Second, our work provides a theoretically-principled empirical evaluation---by far the largest, most detailed, and most comprehensive to date---and the theory we apply was developed previously~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
-Performing such a meta-analysis of previously-published work is common in certain areas, but it is quite rare in machine learning, where the emphasis is on developing better training protocols.
-Third, in addition to addressing our main question, our weight matrix analysis can be used to obtain fine-scale insight into DNN models.
-This can be used to develop metrics for \emph{how} to train better models as well as to form hypotheses for \emph{why} DNNs perform as they do.
-We provide two such examples---\emph{Scale Collapse} and \emph{Correlation Flow}---that we have observed in both CV and NLP models and that deserve attention in future work.
-
-%KDD% \vspace{-1mm}
-\paragraph{The WeightWatcher Tool.}
-
-All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
-To be fully reproducible, we only examine publicly-available, pretrained models, and we also provide all Jupyter and Google Colab notebooks used in an accompanying github repository~\cite{kdd20_sub_repo}.
-See Appendix~\ref{sxn:appendix} for details on how to reproduce all~results.
-
-%KDD% \vspace{-1mm}
-\paragraph{Organization of this paper.}
-
-We start in Section~\ref{sxn:background} and Section~\ref{sxn:methods} with background and an overview of our general approach.
-In Section \ref{sxn:cv}, we study three well-known widely-available DNN CV architectures (the VGG, ResNet, and DenseNet series of models); and we provide an illustration of our basic methodology, both to evaluate the different metrics against reported test accuracies and to use quality metrics to understand model properties.
-Then, in Section~\ref{sxn:nlp}, we look at several variations of a popular NLP DNN architecture (the OpenAI GPT and GPT2 models); and we show how model quality and properties vary between several variants of GPT and GPT2, including how metrics behave similarly and differently.
-Then, in Section \ref{sxn:all_cv_models}, we present results based on an analysis of hundreds of pretrained DNN models, showing how well each metric predicts the reported test accuracies, and how the PL-based metrics perform remarkably well.
-Finally, in Section~\ref{sxn:conc}, we provide a brief discussion and conclusion.
+Fourth, PL-based metrics can also be used to characterize fine-scale model properties, including what we call layer-wise \emph{Correlation Flow}, in well-trained and poorly-trained models; and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
+%
+Our work provides a theoretically-principled empirical evaluation---by far the largest, most detailed, and most comprehensive to date---and the theory we apply was developed previously~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+Performing such a meta-analysis of previously-published work is common in certain areas, but it is quite rare in ML, where the emphasis is on developing better training protocols.

### 2020-09-04 submission/NatCom_methods.tex
initial pass to modify for nat comm

\section{Methods}
\label{sxn:methods}

Consider the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$, as 
the minimization of a general Loss Function $\mathcal{L}$ over $N$ training data instances and labels, $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$.
For a typical supervised classification problem, one has
\begin{equation}
\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N}\left(\mathcal{L}(E_{DNN}(\mathbf{x}_{i})-y_{i})\right)  ,
\end{equation}
where the loss function $\mathcal{L}(\cdot)$ can take on a myriad of forms~\cite{JC17_TR}, and where
\begin{equation}
E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})
\label{eqn:dnn_energy}
\end{equation}
is the Energy (or optimization) Landscape function.  
$E_{DNN}$ does not explicitly depend on the data, but rather it maps data instance vectors ($\mathbf{x}_i$ values) to predictions (labels).
Therefore, one can analyze $E_{DNN}$ in the absence of any training or test data.

Test accuracies have been reported online for publicly-available pretrained pytroch models.
These models have been trained and evaluated on labeled data $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$, using standard techniques.  
We do not have access to this data, and we have not trained any of the models ourselves. 
Our methodological approach is thus similar to a statistical meta-analysis, common in biomedical research, but uncommon in ML.
%
Computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
To be fully reproducible, we only examine publicly-available, pretrained models, and we provide all Jupyter and Google Colab notebooks used in an accompanying github repository~\cite{kdd20_sub_repo}.
See Appendix~\ref{sxn:appendix} for details.

\paragraph{Metrics for DNN Weight Matrices.}

Our approach involves analyzing DNN weight matrices.
Each DNN layer contains one or more layer 2D  $N\times M$ weight matrices, $\mathbf{W}_{l}$, or pre-activation maps, $\mathbf{W}_{i,l}$, e.g., extracted from 2D Convolutional layers, where $N > M$.
See Appendix~\ref{sxn:appendix} for details.
(We may drop the $i$ and/or $i,l$ subscripts below.)
The best performing quality metrics depend on the norms and/or spectral properties of each weight matrix,
$\mathbf{W}$, and/or, equivalently, it's Empirical Correlation Matrix, $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
We consider the following metrics:
\begin{eqnarray}
& & \text{Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{X}\Vert_{F}=\sum\nolimits_{i=1}^{M} \lambda_{i}$ } \\
& & \text{Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}^{2}=\Vert\mathbf{X}\Vert_{\infty}=\lambda_{max}$ } \\
& & \text{Weighted Alpha: $\hat{\alpha}=\alpha\log\lambda_{max}$ } \\
& & \text{$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{W}\Vert^{2\alpha}_{2\alpha}=\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum\nolimits_{i=1}^{M}\lambda_{i}^{\alpha}$. }
\end{eqnarray}
Here, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the $\mathbf{X}$, and $\lambda_{max}$ is the maximum eigenvalue.
These eigenvalues are squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, $\lambda_{i}=\sigma^{2}_{i}$.
%See Appendix~\ref{sxn:appendix} for details. 
All four metrics can be computed easily from DNN weight matrices.
The first two metrics are well-known in ML.
The last two metrics deserve special mention, as they depend on an empirical parameter $\alpha$ that is the PL exponent that arises in the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.

\paragraph{Overview of Heavy-Tailed Self-Regularization.}

In the HT-SR Theory, one analyzes the eigenvalue spectrum, i.e., the Empirical Spectral Density (ESD), of the associated correlation matrices~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
From this, one characterizes the amount and form of correlation, and therefore implicit self-regularizartion, present in the DNN's weight matrices.
For each layer weight matrix $\mathbf{W}$, of size $N \times M$, construct the associated $M\times M$ (uncentered) correlation matrix $\mathbf{X}$. 
Dropping the $L$ and $l,i$ indices, one~has
$$
\mathbf{X} = \frac{1}{N}\mathbf{W}^{T}\mathbf{W}.
$$
If we compute the eigenvalue spectrum of $\mathbf{X}$, i.e., $\lambda_i$ such that
$  % $$
\mathbf{X}\mathbf{v}_{i}=\lambda_{i}\mathbf{v}_{i} , 
$  % $$
then the ESD of eigenvalues, $\rho(\lambda)$, is just a histogram of the eigenvalues, formally written as
%\begin{equation}
$\rho(\lambda)=\sum\nolimits_{i=1}^{M}\delta(\lambda-\lambda_{i})  .$
%\label{eqn:eigenval_hist}
%\end{equation}
Using HT-SR Theory, one characterizes the correlations in a weight matrix by examining its ESD, $\rho(\lambda)$.
It can be well-fit to a truncated PL distribution, given~as
\begin{equation}
\rho(\lambda)\sim\lambda^{-\alpha}  ,
\label{eqn:eigenval_pl}
\end{equation}
which is (at least) valid within a bounded range of eigenvalues $\lambda\in[\lambda^{min},\lambda^{max}]$.

The original work on HT-SR Theory considered a small number of NNs, including AlexNet and InceptionV3. 
It showed that for nearly every $\mathbf{W}$, the (bulk and tail) of the ESDs can be fit to a truncated PL, and that PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
%
As for the mechanism responsible for these properties, statistical physics offers several possibilities~\cite{SornetteBook,nishimori01}, e.g., self organized criticality~\cite{SOC87,SOCat25yrs} or multiplicative noise in the stochastic optimization algorithms used to train these models~\cite{HodMah20A_TR,SorCon97}.
Our meta-analysis does not require knowledge of mechanisms; and it is not even clear that one mechanism is responsible for every case.
%
Crucially, HT-SR Theory predicts that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models.
Relatedly, previous work observed that smaller exponents $\alpha$ correspond to more implicit self-regularization and better generalization~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.

\paragraph{DNN Empirical Quality Metrics.}

For norm-based metrics, we use the average of the log norm, to the appropriate power.
Informally, this amounts to assuming that the layer weight matrices are statistically independent, in which case we can estimate the model complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm (which resembles a data dependent VC complexity),
\begin{equation}
\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert \times \cdots \times \Vert\mathbf{W}_{L}\Vert ,
\end{equation}
where $\Vert\cdot\Vert$ is a matrix norm.   
The log complexity,
\begin{equation}
\label{eqn:eqn:sum_log_norm}
\log\mathcal{C} \sim \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert + \cdots + \log\Vert\mathbf{W}_{L}\Vert = \sum\nolimits_l \log\Vert\mathbf{W}_{l}\Vert ,
\end{equation}
 takes the form of an average Log Norm.
For the \emph{Frobenius Norm metric} and \emph{Spectral Norm metric}, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly (since, when taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor).

The \emph{Weighted Alpha metric} is an average of $\alpha_l$ over all layers $l \in \{1,\ldots,l\}$, weighted by the size, or scale, or each matrix,
\begin{equation}
\hat{\alpha} = \dfrac{1}{L}\sum_l \alpha_l\log\lambda_{max,l}\approx\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle    ,
\end{equation}
where $L$ is the total number of layer weight matrices.
The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a much smaller and more limited set of models than we consider here.

Based on this, in this paper, we introduce and evaluate the \emph{$\alpha$-Shatten Norm metric},
\begin{equation}
\label{eqn:sum_log_alpha_norm_alpha}
\sum\nolimits_l \log \Vert\mathbf{X}_l\Vert_{\alpha_l}^{\alpha_l} 
=
\sum\nolimits_l \alpha_l \log \Vert\mathbf{X}_l\Vert_{\alpha_l} .
\end{equation}
For the $\alpha$-Shatten Norm metric, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it can not be taken out of the sum.
For small $\alpha$, the Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}; and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm.

One determines $\alpha$ for a given layer by fitting the ESD of that layer's weight matrix to a truncated PL, using the commonly accepted Maximum Likelihood method~\cite{CSN09_powerlaw,ABP14}.
This method works very well for exponents between $\alpha\in(2,4)$; and it is adequate, although imprecise, for smaller and especially larger $\alpha$~\cite{newman2005_zipf}. 
%
Operationally, $\alpha$ is determined by using the \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the histogram of eigenvalues, $\rho(\lambda)$, to a truncated PL, 
\begin{equation}
\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\in[\lambda_{min},\lambda_{max}] ,
\end{equation}
where $\lambda_{max}$ is the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$, and 
where $\lambda_{min}$ is selected automatically to yield the best (in the sense of minimizing the K-S distance) PL fit.
Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.
See Figure~\ref{fig:ww} for an illustration.

\begin{figure}[t]
    \centering
    \includegraphics[width=15.0cm]{img/WeightWatcher_v2}
    \caption{Schematic of analyzing DNN layer weight matrices $\mathbf{W}$.  Norm-based metrics (not shown) and PL-based metrics (depending on fitting the histogram of eigenvalues to a truncated PL) can be used to compare models (or for other tasks such as improved model training).}
    \label{fig:ww}
\end{figure}

To avoid confusion, let us clarify the relationship between $\alpha$ and $\hat{\alpha}$.  
We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated PL, parameterized by 2 values: the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.
%%\footnote{Technically, we also need the minimum eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.}
The PL exponent $\alpha$ measures the amount of correlation in a DNN layer weight matrix $\mathbf{W}$. 
It is valid for $\lambda\le\lambda_{max}$, and it is scale-invariant, i.e., it does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.
The $\lambda_{max}$ is a measure of the size, or scale, of $\mathbf{W}$.
Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric $\hat{\alpha}$ for the entire~DNN.

### 2020-09-04 submission/NatCom_nlp_models.tex
initial pass to modify for nat comm

--- 2020-07-06 submission/nlp_models.tex
@@
-%KDD% 
-\vspace{-1mm}
-\section{Comparison of NLP Models}
+\subsection{Comparison of NLP Models}
 \label{sxn:nlp}
-%\vspace{-1mm}
 
-In this section, we examine empirical quality metrics described in Section~\ref{sxn:methods} for several NLP model architectures.
-%
-%In particular, 
-Within the past two years, nearly 100 open source, pretrained NLP DNNs based on the revolutionary Transformer architecture have emerged.
+Within the past few years, nearly 100 open source, pretrained NLP DNNs based on the revolutionary Transformer architecture have emerged.
 These include variants of BERT, Transformer-XML, GPT, etc.
-%
 The Transformer architectures consist of blocks of so-called Attention layers, containing two large, Feed Forward (Linear) weight matrices~\cite{Attn2017}. 
 In contrast to smaller pre-Activation maps arising in Cond2D layers, Attention matrices are significantly larger.
-In general, we have found that they have larger PL exponents $\alpha$.
+In general, they have larger PL exponents $\alpha$.
 Based on HT-SR Theory
@@
 this suggests that these models fail to capture successfully many of the information correlations in the data (relative to their size) and thus are substantially \emph{under}-trained.
-%
-More generally, compared to the CV models of Section~\ref{sxn:cv},
+More generally, compared to CV models,
 modern NLP models have larger weight matrices and display different spectral properties.
-%, both in terms of their architecture (e.g., whether there are convolutions) as well as how they are used (e.g., whether one uses prediction error or perplexity).
-Thus, they provide a very different test for our empirical quality metrics.
 
 While norm-based metrics perform reasonably well on well-trained NLP models, they often behave anomalously on poorly-trained models.
-Indeed, for such ``bad'' models, weight matrices may display rank collapse, decreased Frobenius mass, or unusually small Spectral norms.
+For such models, weight matrices may display rank collapse, decreased Frobenius mass, or unusually small Spectral norms.
 This may be misinterpreted as ``smaller is better.'' 
 Instead, it should probably be interpreted as being due to a similar mechanism to how distillation can ``damage'' otherwise good models.
-In contrast to norm-based metrics, PL-based metrics, including the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $) display consistent behavior, even on poorly trained models.
-Indeed, we can use these metrics to help identify when architectures need repair and when more and/or better data are needed.
+In contrast to norm-based metrics, PL-based metrics, including the Log $\alpha$-Norm metric and the Weighted Alpha metric display consistent behavior, even on poorly trained models.
+One can use these metrics to help identify when architectures need repair and when more and/or better data are needed.
 
-%KDD% 
-\vspace{-1mm}
 \paragraph{What do large values of $\alpha$ mean?}
 
-We have observed that many NLP models, such as GPT and BERT, have some weight matrices with unusually large PL exponents (e.g., $\alpha\gg 6$).
+Many NLP models, such as early variants of GPT and BERT, have weight matrices with unusually large PL exponents (e.g., $\alpha\gg 6$).
 This indicates these matrices may be \emph{under}-correlated (i.e., over-parameterized, relative to the amount of data).
-In this regime, the truncated PL fit itself may not be very reliable because the MLE estimator it uses is unreliable in this range.
-That is, the specific $\alpha$ values returned by the truncated PL fits are less reliable, but having large versus small values of $\alpha$ is reliable.
-Phenomenologically, if we examine the ESD visually, we can usually describe these $\mathbf{W}$ as in the \emph{Bulk-Decay} or \emph{Bulk-plus-Spikes} phase from HT-ST Theory~\cite{MM18_TR,MM19_HTSR_ICML}.
-Previous work~\cite{MM18_TR,MM19_HTSR_ICML} has conjectured that very well-trained DNNs would not have many \emph{outlier} $\alpha\gg 6$; and, consistent with this, improved versions of GPT (shown below) and BERT (not shown) confirm this.
+In this regime, the truncated PL fit itself may not be very reliable because the Maximum Likelihood estimator it uses is unreliable in this range.
+The specific $\alpha$ values returned by the truncated PL fits are less reliable, but having large versus small $\alpha$ is reliable.
+If the ESD is visually examined, one can usually describe these $\mathbf{W}$ as in the \emph{Bulk-Decay} or \emph{Bulk-plus-Spikes} phase from HT-ST Theory~\cite{MM18_TR,MM19_HTSR_ICML}.
+Previous work~\cite{MM18_TR,MM19_HTSR_ICML} has conjectured that very well-trained DNNs would not have many \emph{outlier} $\alpha\gg 6$.
+Consistent with this, more recent improved versions of GPT (shown below) and BERT (not shown) confirm this.
 
-%KDD% 
-\vspace{-1mm}
 \paragraph{OpenAI GPT Models.}
 
-The OpenAI GPT and GPT2 models provide us with the opportunity to analyze two effects: training the same model with different data set sizes; and increasing the sizes of both the data set and the architectures simultaneously.
-These models have the remarkable ability to generate fake text that appears to the human to be real, and they have generated significant media attention because of the potential for their misuse.
+The OpenAI GPT and GPT2 models provide the opportunity to analyze two effects: training the same model with different data set sizes; and increasing the sizes of both the data set and the architectures simultaneously.
+These models have the ability to generate fake text that appears to the human to be real, and they have generated media attention because of the potential for their misuse.
 For this reason, the original GPT model released by OpenAI was trained on a deficient data set, rendering the model interesting but not fully functional.  
-Later, OpenAI released a much improved model, GPT2-small, which has the same architecture and number of layers as GPT, but which has been trained on a larger and better data set (and with other changes), making it remarkably good at generating (near) human-quality fake text.
-%
-By comparing the poorly-trained (i.e., ``bad'') GPT to the well-trained (i.e., ``good'') GPT2-small, we can identify empirical indicators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
+Later, OpenAI released a much improved model, GPT2-small, which has the same architecture and number of layers as GPT, but which has been trained on a larger and better data set, making it remarkably good at generating (near) human-quality fake text.
+By comparing the poorly-trained GPT to the well-trained GPT2-small, we can identify empirical indicators for when a model has been poorly-trained and thus may perform poorly when deployed.
 By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously, as well as analyze well-trained versus very-well-trained models.
-
 The GPT models we analyze are deployed with the popular HuggingFace PyTorch library~\cite{huggingface}.
-GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ layer Weight Matrices, $\mathbf{W}$.
-Each Block has 2 components, the Self Attention (attn) and the Projection (proj) matrices.  
-The self-attention  matrices are larger, of dimension ($2304\times 768$) or ($3072\times 768$).
-The projection layer concatenates the self-attention results into a vector (of dimension $768$).
-This gives $50$ large matrices.
-%
-Because GPT and GPT2 are trained on different data sets, the initial Embedding matrices differ in shape.
-GPT has an initial Token and Positional Embedding layers, of dimension $(40478\times 768)$ and $(512\times 768)$, respectively, whereas GPT2 has input Embeddings of shape $(50257\times 768)$ and $(1024\times 768)$, respectively.
-%
-The OpenAI GPT2 (English) models are: GPT2-small, GPT2-medium, GPT2-large, and GPT2-xl, having $12$, $24$, $36$, and $48$ layers, respectively, with increasingly larger weight matrices.
-%The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
 
@@
 Column \# refers to number of layers treated.  
-Note that the averages do not include the first embedding layer(s) because they are not (implicitly) normalized.
+Averages do not include the first embedding layer(s) because they are not (implicitly) normalized.  
+GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ layer Weight Matrices, $\mathbf{W}$.
+Each Block has 2 components, the Self Attention (attn) and the Projection (proj) matrices.  
+Self-attention  matrices are larger, of dimension ($2304\times 768$) or ($3072\times 768$).
+The projection layer concatenates the self-attention results into a vector (of dimension $768$).
+This gives $50$ large matrices.
+Because GPT and GPT2 are trained on different data sets, the initial Embedding matrices differ in shape.
+GPT has an initial Token and Positional Embedding layers, of dimension $(40478\times 768)$ and $(512\times 768)$, respectively, whereas GPT2 has input Embeddings of shape $(50257\times 768)$ and $(1024\times 768)$, respectively.
+The OpenAI GPT2 (English) models are: GPT2-small, GPT2-medium, GPT2-large, and GPT2-xl, having $12$, $24$, $36$, and $48$ layers, respectively, with increasingly larger weight matrices.
 }
@@
 
-%KDD% 
-\vspace{-1mm}
 \paragraph{Average Quality Metrics for GPT and GPT2.}
 
-We have analyzed the four quality metrics described in Section~\ref{sxn:methods} for the OpenAI GPT and GPT2 pretrained models.
-See Table \ref{table:nlp} for a summary of results.
-We start by examining trends between GPT and GPT2-small.
-Observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
-Notice that in the poorly-trained GPT model, the norm metrics all exhibit \emph{Scale Collapse}, compared to well-trained GPT2-small model.
+See Table \ref{table:nlp} for a summary of results for the four quality metrics for the OpenAI GPT and GPT2 pretrained models.
+Comparing trends between GPT and GPT2-small, 
+observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
+In the poorly-trained GPT model, the norm metrics all exhibit \emph{Scale Collapse}, compared to well-trained GPT2-small model.
 
-We next examine trends between GPT2-medium to GPT2-large to GPT2-xl.
-Observe that (with one minor exception involving the log Frobenius norm metric) all four metrics decrease as one goes from medium to large to xl, indicating that the larger models indeed look better than the smaller models.
-Notice that, for these well-trained models, the norm metrics now behave as expected, i.e., there is no \emph{Scale Collapse} and norms are decreasing with increasing~accuracy.
+Comparing trends between GPT2-medium to GPT2-large to GPT2-xl,
+observe that (with one minor exception involving the log Frobenius norm metric) all four metrics decrease as one goes from medium to large to xl, indicating that the larger models indeed look better than the smaller models.
+For these well-trained models, the norm metrics now behave as expected, i.e., there is no \emph{Scale Collapse} and norms are decreasing with increasing~accuracy.
 
 Going beyond average values, Figure~\ref{fig:GPT-alpha-hist} shows the histogram (empirical density), for all layers, of $\alpha$ for GPT and GPT2-small.  
-These two histograms are very different.
-The older deficient GPT has numerous unusually large $\alpha$ exponents---meaning they are not really well-described by a PL fit.
+The older deficient GPT has numerous unusually large $\alpha$ exponents---meaning they are not well-described by a PL fit.
 Indeed, we expect that a poorly-trained model will lack good (i.e., small $\alpha$) PL behavior in many/most layers.
-On the other hand, as expected, the newer improved GPT2-small model has, on average, smaller $\alpha$ values than the older GPT, with all $\alpha\le6$ and with smaller mean/median $\alpha$.
+On the other hand, the newer improved GPT2-small model has, on average, smaller $\alpha$ values than the older GPT, with all $\alpha\le6$ and with smaller mean/median $\alpha$.
 It also has far fewer unusually-large outlying $\alpha$ values than GPT.
-From this (and other results not shown), we see that $\alpha$ provides a good quality metric for comparing these two models, the poorly-trained GPT versus the well-trained GPT2-small.
-%
+From this (and other results not shown), we see that $\alpha$ provides a good quality metric for comparing the poorly-trained GPT versus the well-trained GPT2-small.
 This should be contrasted with the behavior displayed by the Frobenius norm (not shown) and the Spectral~norm.
 
-%KDD% 
-\vspace{-1mm}
 \paragraph{Scale Collapse in Poorly Trained Models.}
 
-We next describe the behavior of the Spectral norm in GPT versus GPT2-small.
-In Figure~\ref{fig:GPT-snorm-hist}, the poorly-trained GPT model has a smaller mean/median Spectral norm as well as, spuriously, many much smaller Spectral norms, compared to the well-trained GPT2-small, violating the conventional wisdom that smaller Spectral norms are better.
-Indeed, because there are so many anonymously small Spectral norms, it appears that the GPT model may be exhibiting a kind of \emph{Scale Collapse}, like that observed in the distilled CV models (in Figure~\ref{fig:resnet204D5L}).
+We next describe the Spectral norm in GPT versus GPT2-small.
+In Figure~\ref{fig:GPT-snorm-hist}, the poorly-trained GPT model has a smaller mean/median Spectral norm as well as, spuriously, many much smaller Spectral norms, compared to the well-trained GPT2-small.
+This violates the conventional wisdom that smaller Spectral norms are better.
+Because there are so many anonymously small Spectral norms, the GPT model may be exhibiting a kind of \emph{Scale Collapse}, like that observed in the distilled CV models (Figure~\ref{fig:resnet204D5L}).
 
-This is important because it demonstrates that, while the Spectral (or Frobenius) norm may correlate well with predicted test error, at least among reasonably well-trained models, it is \emph{not} a good indicator of the overall model quality in general.
-In particular, it can mispredict well-trained versus poorly-trained questions in ways not seen with PL-based metrics.
+This demonstrates that, while the Spectral (or Frobenius) norm may correlate well with predicted test error, at least among reasonably well-trained models, it is \emph{not} a good indicator of the overall model quality in general.
+It can mispredict well-trained versus poorly-trained questions in ways not seen with PL-based metrics.
 Using it as an empirical quality metric may give spurious results when applied to poorly-trained or otherwise deficient~models.
@@
 
-We should note that Figure~\ref{fig:GPT-snorm-hist} also shows some unusually large Spectral Norms.
-Upon examination, e.g., from Figure \ref{fig:gpt-snorm-layer} (below), we see that these correspond to the first embedding layer(s).
-%That is, they correspond to the change in normalization of the word embedding layer(s), and thus they can be ignored.
+Figure~\ref{fig:GPT-snorm-hist} also shows some unusually large Spectral Norms.
+Upon examination, e.g., from Figure \ref{fig:gpt-snorm-layer} (below), these correspond to the first embedding layer(s).
 These layers have a different effective normalization, and therefore a different scale.
-% \charles{We discuss this further in Appendix~\ref{sxn:appendix}.}
-% For example, in GPT, for most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$, but in the first embedding layer, the maximum is of order $N$ (the number of words in the embedding), or $\lambda_{max}\sim\mathcal{O}(10^{5})$.
-% For GPT and GPT2, we treat all layers as-is (although one may want to normalize the first 2 layers by $\mathbf{X}$ by $1/N$ or treat them as an outlier).
-We discuss this further in Appendix~\ref{sxn:appendix}.
+See Appendix~\ref{sxn:appendix} for details.
 Here, we do not include them in our computed average metrics in Table~\ref{table:nlp}, and we do not include them in the histogram plot in Figure~\ref{fig:GPT-snorm-hist}.
 
-%KDD% 
-\vspace{-1mm}
 \paragraph{Layer Analysis: Correlation Flow and Scale Collapse in GPT and GPT2.}
 
-We also examine in Figure~\ref{fig:gpt-alpha-layers} the PL exponent $\alpha$ and Log Spectral Norm versus layer id, for GPT and GPT2-small.
-Let's start with Figure~\ref{fig:gpt-alpha-layer}, which plots $\alpha$ versus the depth (i.e., layer id) for each model.
+See Figure~\ref{fig:gpt-alpha-layers} for the PL exponent $\alpha$ and Log Spectral Norm versus layer id, for GPT and GPT2-small.
+Figure~\ref{fig:gpt-alpha-layer} plots $\alpha$ versus the depth (layer id) for each model.
 The deficient GPT model displays two trends in $\alpha$, one stable with $\alpha\sim 4$, and one increasing with layer id, with $\alpha$ reaching as high as $12$.
@@
 
-%KDD% 
-\vspace{-1mm}
 \paragraph{GPT2: medium, large, xl.}
 
-We now look across series of increasingly improving GPT2 models (i.e., we consider well-trained versus very-well-trained questions), by examining both the PL exponent $\alpha$ as well as the Log Norm metrics.
+We now look across series of increasingly improving GPT2 models (well-trained versus very-well-trained models), by examining both the PL exponent $\alpha$ as well as the Log Norm metrics.  
 In general, as we move from GPT2-medium to GPT2-xl, histograms for both $\alpha$ exponents and the Log Norm metrics downshift from larger to smaller values. 
-For example, see Figure \ref{fig:gpt2-histograms}, which shows the histograms over the layer weight matrices for fitted PL exponent ($\alpha$) and the Log Alpha Norm ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$)~metric.
+Figure \ref{fig:gpt2-histograms} shows the histograms over the layer weight matrices for fitted PL exponent $\alpha$ and the Log Alpha Norm metric.
 
-We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticeable between the differing well-trained 
+The average $\alpha$ decreases with increasing model size, although the differences are less noticeable between the differing well-trained 
 versus very-well-trained GTP2 models than between the poorly-trained versus well-trained GPT and GPT2-small models.
-Unlike GPT, however, the layer Log Alpha Norms behave more as expected for GPT2 layers, with the larger models consistently having smaller norms. 
+Unlike GPT, the layer Log Alpha Norms behave more as expected for GPT2 layers, with the larger models consistently having smaller norms. 
 Similarly, the Log Spectral Norm also decreases on average with the larger models (not shown).  
-As expected, the norm metrics can indeed distinguish among well-trained versus very-well-trained models among a series well-trained models.
+As expected, the norm metrics can indeed distinguish among well-trained versus very-well-trained models.
 
-We do notice, however, that while the peaks of the $\alpha$ are getting smaller, towards $2.0$, the tails of the distribution shifts right, with larger GPT2 models having more unusually large $\alpha$ (also not shown).
-We suspect this indicates that these larger GPT2 models are still under-optimized/over-parameterized (relative to the data on which they were trained) and that they have capacity to support datasets even larger than the recent XL $1.5B$ release~\cite{gpt2-xl}.
+However, while the peaks of the $\alpha$ are getting smaller, towards $2.0$, the tails of the distribution shifts right, with larger GPT2 models having more unusually large $\alpha$ (also not shown). 
+This suggests that these larger GPT2 models are still under-optimized/over-parameterized (relative to the data on which they were trained) and that they have capacity to support datasets even larger than the recent XL $1.5B$ release~\cite{gpt2-xl}.
 
@@
     \subfigure[PL exponent ($\alpha$)]{
-        %KDD% \includegraphics[width=4.0cm]{img/GPT2*_fnl_alpha_hist.png}
         \includegraphics[width=5.5cm]{img/GPT2*_fnl_alpha_hist.png}
@@
     }
-    %KDD% 
     \quad
     \subfigure[Log Alpha Norm]{
-        %KDD% \includegraphics[width=4.0cm]{img/GPT2*_fnl_logpnorm_hist.png}
         \includegraphics[width=5.5cm]{img/GPT2*_fnl_logpnorm_hist.png}

### 2020-09-04 submission/empirical2_NatComfmt.tex
initial pass to modify for nat comm

--- 2020-02-16 submission/empirical2_TRfmt.tex
@@
 \begin{abstract}
-\input{abstract_file}
+\input{NatCom_abstract_file}
 \end{abstract}
@@
 
-\input{introduction}
-\input{background}
-\input{methods}
-\input{cv_models}
-\input{nlp_models}
-\input{all_models}
-\input{conclusion}
+\input{NatCom_introduction}
+\input{NatCom_methods}
 
-\vspace{-2mm}
+\section{Results}
+We study in detail three well-known CV architectures (the VGG, ResNet, and DenseNet series of models).
+Then, we look in detail at several variations of a popular NLP architecture (the OpenAI GPT and GPT2 models).
+Finally, we present results from a broader analysis of hundreds of pretrained DNN~models.
+
+\input{NatCom_cv_models}
+\input{NatCom_nlp_models}
+\input{NatCom_all_models}
+
+\input{NatCom_discussion}
+
 \noindent
@@
 MWM would like to acknowledge ARO, DARPA, NSF, and ONR as well as the UC Berkeley BDD project and a gift from Intel for providing partial support of this work.
+Our conclusions do not necessarily reflect the position or the policy of our sponsors, and no official endorsement should be inferred.
 We would also like to thank Amir Khosrowshahi and colleagues at Intel for helpful discussion regarding the Group Regularization distillation technique.
@@
 \appendix
-\input{appendix}
+\input{NatCom_appendix}
 

### 2020-10-01 submission/NatCom_abstract_file.tex
pass to address most of the presubmission comments for nat comm

@@
 In many applications, one works with neural network models trained by someone else.
-For such \emph{pretrained models}, one often does not have access to training data or test data.
+For such pretrained models, one may not have access to training data or test data.
 Moreover, one may not know details about the model, e.g., the specifics of the training data, the loss function, the hyperparameter values, etc.
-Given one or many pretrained models, can one say anything about the expected performance or quality of the models?
-Here, we address this question by providing a detailed meta-analysis of hundreds of publicly-available pretrained models.
+Given one or many pretrained models, it is a challenge to say anything about the expected performance or quality of the models.
+Here, we address this challenge by providing a detailed meta-analysis of hundreds of publicly-available pretrained models.
 We examine norm based capacity control metrics as well as power law based metrics from the recently-developed Theory of Heavy-Tailed Self Regularization.

### 2020-10-01 submission/NatCom_appendix.tex
pass to address most of the presubmission comments for nat comm

--- 2020-02-16 submission/appendix.tex
@@
-\section{Appendix}
+%\section{Reproducibility Appendix}
+\section{Supplementary Information}
 \label{sxn:appendix}
 
-In this appendix, we provide more details on several issues that are important for the reproducibility of our results.
-%
-All of our computations were performed with the \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
-More details and more results are available in an accompanying github repository~\cite{kdd20_sub_repo}.
+\subsection{Supplementary Tables}
 
-\subsection{Reproducibility Considerations}
-
-\paragraph{SVD of Convolutional 2D Layers.}
-
-There is some ambiguity in performing spectral analysis on Conv2D layers.  
-Each layer is a 4-index tensor of dimension $(w,h,in,out)$, with an $(w\times h)$ filter (or kernel) and $(in, out)$
-channels. When $w=h=k$, it gives $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\times out)$ each. 
-%
-We identify 3 different approaches for running SVD on a Conv2D layer:
-\begin{enumerate}
-\item run SVD on each pre-Activation Map $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values;
-\item stack the maps into a single matrix of, say, dimension $((k\times k\times out)\times in)$, and run SVD to get $in$ singular values;
-\item compute the 2D Fourier Transform (FFT) for each of the $(in, out)$ pairs, and run SVD on the Fourier coefficients~\cite{CNNSVD}, leading to $\sim(k\times in\times out)$ non-zero singular values.
-\end{enumerate}
-Each method has tradeoffs.  
-Method (3) is mathematically sound, but computationally expensive. Method (2) is ambiguous.
-For our analysis, because we need thousands of runs, we select method (1), which is the fastest (and is easiest to reproduce).
-
-\paragraph{Normalization of Empirical Matrices.}
-
-Normalization is an important, if underappreciated, practical issue.
-Importantly, the normalization of weight matrices does \emph{not} affect the PL fits because $\alpha$ is scale-invariant.
-Norm-based metrics, however, do depend strongly on the scale of the weight matrix---that is the point.
-%\nred{Indeed, early theoretical work by Bartlett suggests that the test accuracy depends strongly on the ``total size'' of the weight matrics.}
-To apply RMT, we usually define $\mathbf{X}$ with a $1/N$ normalization, assuming variance of $\sigma^{2}=1.0$.
-%\footnote{For Heavy Tailed theorems, one typically needs a normalization such as \nred{$1/N^{\alpha-1}$. check this}}
-Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with
- $\sigma^{2}\sim 1/\sqrt{N}$, or some variant, e.g., the Glorot/Xavier normalization~\cite{GloBen10}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers. With this implicit scale, 
-we do \emph{not} ``renormalize'' the empirical weight matrices, i.e., we use them as-is.
-The only exception is that \emph{we do rescale} the Conv2D pre-activation maps $\mathbf{W}_{i,L}$ 
-by $k/\sqrt{2}$ so that they are on the same scale as the Linear / Fully Connected (FC) layers.
-
-\paragraph{Special consideration for NLP models.}
-
-NLP models, and other models with large initial embeddings, require special care because the embedding layers frequently lack the implicit $1/\sqrt{N}$ normalization present in other layers.
-For example, in GPT, for most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$, but in the first embedding layer, the maximum eigenvalue is of order $N$ (the number of words in the embedding), or $\lambda_{max}\sim\mathcal{O}(10^{5})$.
-For GPT and GPT2, we treat all layers as-is (although one may want to normalize the first 2 layers $\mathbf{X}$ by $1/N$, or to treat them as outliers).
-
-\subsection{Reproducing Sections \ref{sxn:cv} and \ref{sxn:nlp} }   %% \subsection{Reproducing Sections 4 and 5}
+\paragraph{Reproducing Sections \ref{sxn:cv} and \ref{sxn:nlp}. }
 
 We provide a github repository for this paper that includes Jupyter notebooks that fully reproduce all results (as well as many other results)~\cite{kdd20_sub_repo}.
-All results have been produced using the \emph{WeightWatcher} tool (v0.2.7)~\cite{weightwatcher_package}.
+All results have been produced using the \texttt{WeightWatcher} tool (v0.2.7)~\cite{weightwatcher_package}.
 The ImageNet and OpenAI GPT pretrained models are provided in the current 
@@
 
-\subsection{Reproducing Figure~\ref{fig:resnet204D5L}, for the Distiller Model}
+\paragraph{Reproducing Figure~\ref{fig:resnet204D5L}, for the Distiller Model.}
 
@@
 Figure~\ref{fig:resnet204D5L} is from the  \texttt{``...-Distiller-ResNet20.ipynb''} notebook (see Table~\ref{table:notebooks}).  
-For completeness, we provide both the results described here, as well as additional results on other pretrained and distilled models using the \emph{WeightWatcher} tool.
+For completeness, we provide both the results described here, as well as additional results on other pretrained and distilled models using the \texttt{WeightWatcher} tool.
 
-\subsection{Reproducing Table~\ref{table:results} in Section~\ref{sxn:all_cv_models} }
+\paragraph{Reproducing Table~\ref{table:results} in Section~\ref{sxn:all_cv_models}. }
 
 In the \texttt{ww-colab} folder of our github repo, we provide several Google Colab notebooks which can be used to reproduce the results of Section~\ref{sxn:all_cv_models}.
-The ImageNet-1K and other pretrained models are taken from the pytorch models in the \texttt{omsr/imgclsmob} 
+The ImageNet-1K and other pretrained models are taken from the pyTorch models in the \texttt{omsr/imgclsmob} 
 ``Sandbox for training convolutional networks for computer vision'' github repository~\cite{osmr}.
@@
 
-To explain further how to reproduce our analysis, we run three batches of linear regressions. 
-First, at the global level, we divide models by datasets and run regressions separately on all models of a certain dataset, regardless of the architecture. 
-At this level, the plots are quite noisy and clustered, as each architecture has its own accuracy trend; but one can
- still see that most plots show positive relationship with positive coefficients. 
-Example regressions are shown in Figure~\ref{fig:DSalphas}, as available in the results notebook.
-
 \begin{table}[t]
-%KDD% \footnotesize  % \small
 \scriptsize
@@
 \end{table}
+
+\subsection{Supplementary Figures}
+
+To explain further how to reproduce our analysis, we run three batches of linear regressions. 
+First, at the global level, we divide models by datasets and run regressions separately on all models of a certain dataset, regardless of the architecture. 
+At this level, the plots are quite noisy and clustered, as each architecture has its own accuracy trend; but one can
+ still see that most plots show positive relationship with positive coefficients. 
+Example regressions are shown in Figure~\ref{fig:DSalphas}, as available in the results notebook.
 
@@
 The final analysis includes 108 regressions in all, those with 4 or more models, with a positive $R^2$.
+
+\subsection{Supplementary Discussion: Additional Details on HT-SR Theory}
+
+The original work on HT-SR Theory~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM} considered NNs including AlexNet and InceptionV3 (as well as DenseNet, ResNet, and VGG), and it showed that for nearly every $\mathbf{W}$, the (bulk and tail) of the ESDs can be fit to a truncated PL and the PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$.
+Our meta-analysis, the main results of which are summarized in this paper, has shown that these results are ubiquitous.
+For example, 
+upon examining nearly 10,000 layer weight matrices $\mathbf{W}_{l,i}$ across hundreds of different modern pre-trained DNN architectures, the ESD of nearly every $\mathbf{W}$ layer matrix can be fit to a truncated PL:
+$70-80\%$ of the time, the fitted PL exponent $\alpha$ lies in the range $\alpha\in(2,4)$; and  
+$10-20\%$ of the time, the fitted PL exponent $\alpha$ lies in the range $\alpha< 2$.  
+Of course, there are exceptions: in any real DNN, the fitted $\alpha$ may range anywhere from $\sim 1.5$ to $10$ or higher (and, of course, larger values of $\alpha$ may indicate that the PL is not a good model for the data).
+Still, overall, in nearly all large, pre-trained DNNs, the correlations in the  weight matrices exhibit a remarkable Universality, being both Heavy Tailed, and having small---but not too small---PL exponents.

### 2020-10-01 submission/NatCom_cv_models.tex
pass to address most of the presubmission comments for nat comm

@@
              VGG11, VGG13, VGG16, and VGG19, with and without Batch Normalization (BN),
-              trained on ImageNet, available in pyTorch (v1.4).  Metrics fit by linear regression, RMSE reported. }
+             trained on ImageNet, available in pyTorch (v1.4).  
+             Metrics fit by linear regression, RMSE reported.  }
     \label{fig:vgg-metrics}
@@
     }
-    \caption{Comparison of Average $\alpha$-Norm quality metric ($\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle$) versus reported Top1 test accuracy for the ResNet and ResNet-1K pretrained (pyTorch) models.
+    \caption{Comparison of Average $\alpha$-Norm quality metric versus reported Top1 test accuracy for the ResNet and ResNet-1K pretrained (pyTorch) models. 
+             \michael{Symbols and text are way to small on these and other figures; axes/ticks way too small; need different symbols, not just red versus green; etc.}
             }
@@
          Column \# refers to number of models.  
-VGG, ResNet, and DenseNet were pretrained on ImageNet, and ResNet-1K was pretrained on ImageNet-1K.
+         VGG, ResNet, and DenseNet were pretrained on ImageNet.  
+         ResNet-1K was pretrained on ImageNet-1K. 
 }
@@
 (These and many other such plots can be seen on our publicly-available~repo.)
-
-\paragraph{Variation in Data Set Size.}
 
@@
 
-\paragraph{Comparison of VGG, ResNet, and DenseNet Architectures.}
+\paragraph{Correlation Flow; or How $\alpha$ Varies Across Layers.}
 
-These layer-wise observations can be understood in terms of architectural differences between VGG, ResNet, and DenseNet.
-VGG resembles the traditional convolutional architectures, such as LeNet5, and consists of several [Conv2D-Maxpool-ReLu] blocks, followed by 3 large Fully Connected (FC) layers.
-ResNet greatly improved on VGG by replacing the large FC layers, shrinking the Conv2D blocks, and introducing residual connections.
-This optimized approach allows for greater accuracy with far fewer parameters, and ResNet models of up to 1000 layers have been trained~\cite{resnet1000}.
-
-The efficiency and effectiveness of ResNet seems to be reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and more strongly optimized.
-This contrasts with the DenseNet models, which contains many connections between every layer.
-These results (large $\alpha$, meaning that even a PL model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly~optimized.
-
-Fine-scale measurements such as these enable us to form hypotheses as to the inner workings of DNN models, opening the door to an improved understanding of \emph{why} DNNs work, as well as \emph{how} to design better DNN models.
-We next turn to two such~examples.
-
-\paragraph{Correlation Flow.}
-
-Figure~\ref{fig:3models-alpha-layers} can be understood in terms of what we will call \emph{Correlation Flow}.
+Figure~\ref{fig:3models-alpha-layers} can be understood in terms of what we will call Correlation Flow.
 The average Log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
@@
 
-The similarity between norm-based metrics and PL-based metrics suggests a question: Is the Weighted Alpha metric just a variation of more familiar norm-based metrics?  
-More generally, do fitted $\alpha$ values contain information not captured by norms? 
-Among hundreds of pretrained models, there are ``exceptions that prove the rule.'' 
-To show that $\alpha$ does capture something different, we show that some compression/distillation methods \cite{CWZZ17_TR} may actually damage models unexpectedly, by introducing what we call \emph{Scale Collapse}, where several distilled layers have unexpectedly small Spectral Norms.
+The similarity between norm-based metrics and PL-based metrics may lead one to wonder whether the Weighted Alpha metric is just a variation of more familiar norm-based metrics.  
+Among hundreds of pretrained models, there are ``exceptions that prove the rule,'' and these can be used to show that fitted $\alpha$ values do contain information not captured by norms. 
+To illustrate this, we show that some compression/distillation methods \cite{CWZZ17_TR} may actually damage models unexpectedly, by introducing what we call Scale Collapse, where several distilled layers have unexpectedly small Spectral Norms.
 
@@
 Pretrained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the VGG, ResNet, or DenseNet series.  
-The obvious interpretation is that, while norms make good regularizers for a single model, there is no reason \emph{a priori} to expect them correlate well with test accuracies across different models, and they may not differentiate well-trained versus poorly-trained models.
+The obvious interpretation is that, while norms make good regularizers for a single model, there is no reason a priori to expect them correlate well with test accuracies across different models, and they may not differentiate well-trained versus poorly-trained models.
 We do expect, however, the PL $\alpha$ to do so, because it effectively measures the amount of information correlation in the model~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.

### 2020-10-01 submission/NatCom_discussion.tex
pass to address most of the presubmission comments for nat comm

@@
 
-\paragraph{Conclusions.}
+\paragraph{Comparison of VGG, ResNet, and DenseNet Architectures.}
 
-We have developed and evaluated methods to predict trends in the quality of state-of-the-art neural networks---without access to training or testing data.
-Our main methodology involved weight matrix meta-analysis, using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package}, and informed by the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
-Prior to our work, it was not even obvious that norm-based metrics would perform well to predict trends in quality \emph{across} models (as they are usually used \emph{within} a given model or parameterized model class, e.g., to bound generalization error or to construct regularizers).
-Our results are the first to demonstrate that they can be used for this important practical problem.
+Going beyond the goal of predicting trends in the quality of state-of-the-art neural networks without access to training or testing data, observations such as
+the layer-wise observations we described in Figure~\ref{fig:3models-alpha-layers} can be understood in terms of architectural differences between VGG, ResNet, and DenseNet.
+VGG resembles the traditional convolutional architectures, such as LeNet5, and consists of several [Conv2D-Maxpool-ReLu] blocks, followed by 3 large Fully Connected (FC) layers.
+ResNet greatly improved on VGG by replacing the large FC layers, shrinking the Conv2D blocks, and introducing residual connections.
+This optimized approach allows for greater accuracy with far fewer parameters, and ResNet models of up to 1000 layers have been trained~\cite{resnet1000}.
+
+The efficiency and effectiveness of ResNet seems to be reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and more strongly optimized.
+This contrasts with the DenseNet models, which contains many connections between every layer.
+These results (large $\alpha$, meaning that even a PL model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly~optimized.
 %
-Our results also demonstrate that PL-based metrics perform better than norm-based metrics. 
-This should \emph{not} be surprising---at least to those familiar with the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}---since our use of PL exponents is designed to capture the idea that well-trained models capture information correlations over many size scales in the data.
-Again, though, our results are the first to demonstrate this.
-%
-Our approach can also be used to provide fine-scale insight (rationalizing the flow of correlations or the collapse of size scale) throughout a network. 
-Both \emph{Correlation Flow} and \emph{Scale Collapse} are important for improved diagnostics on pretrained models as well as for improved training methodologies.
+Fine-scale measurements such as these enable us to form hypotheses as to the inner workings of DNN models, opening the door to an improved understanding of why DNNs work, as well as how to design better DNN models.
+Correlation Flow and Scale Collapse are two such~examples.
 
@@
 Our best-performing PL-based metrics are based on statistical mechanics via HT-SR Theory~\cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM19_KDD, MM20_SDM}.
-The way in which we (and HT-SR Theory) \emph{use} statistical mechanics theory is quite different than the way it is more commonly formulated~\cite{EB01_BOOK, BKPx20}.
-Going beyond idealized models, we \emph{use} statistical mechanics in a broader sense, drawing upon techniques from quantitative finance, random matrix theory, and the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
-%
+The way in which we (and HT-SR Theory) use statistical mechanics theory is quite different than the way it is more commonly formulated~\cite{EB01_BOOK, BKPx20}.
+Going beyond idealized models, we use statistical mechanics in a broader sense, drawing upon techniques from quantitative finance, random matrix theory, and the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
 There is also a large body of work in ML on using norm-based metrics to bound generalization error~\cite{NTS15, BFT17_TR, LMBx18_TR}.
@@
 
-Our work, starting in 2018 with the \emph{WeightWatcher} tool~\cite{weightwatcher_package}, is the first to perform a detailed analysis of the weight matrices of DNNs~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+Our work, starting in 2018 with the \texttt{WeightWatcher} tool~\cite{weightwatcher_package}, is the first to perform a detailed analysis of the weight matrices of DNNs~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
 Subsequent to the initial version of this paper, we became aware of two other works that were posted to the in 2020 within weeks of the initial version of this paper~\cite{EJRUY20_TR,UKGBT20_TR}.
@@
 
+\paragraph{Conclusions.}
+
+We have developed and evaluated methods to predict trends in the quality of state-of-the-art neural networks---without access to training or testing data.
+Our main methodology involved weight matrix meta-analysis, using the publicly-available \texttt{WeightWatcher} tool~\cite{weightwatcher_package}, and informed by the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+Prior to our work, it was not even obvious that norm-based metrics would perform well to predict trends in quality across models (as they are usually used within a given model or parameterized model class, e.g., to bound generalization error or to construct regularizers).
+Our results are the first to demonstrate that they can be used for this important practical problem.
+Our results also demonstrate that PL-based metrics perform better than norm-based metrics. 
+This should not be surprising---at least to those familiar with the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}---since our use of PL exponents is designed to capture the idea that well-trained models capture information correlations over many size scales in the data.
+Again, though, our results are the first to demonstrate this.
+Our approach can also be used to provide fine-scale insight (rationalizing the flow of correlations or the collapse of size scale) throughout a network. 
+Both Correlation Flow and Scale Collapse are important for improved diagnostics on pretrained models as well as for improved training methodologies.
+
 \paragraph{Looking forward.}
 
-We conclude with a few comments on what a \emph{practical theory} of DNNs should look like.
-To do so, we distinguish between two types of theories:
-\emph{non-empirical or analogical theories}, in which one creates, often from general principles, a very simple toy model that can be analyzed rigorously, and one then argues that the model is relevant to the system of interest; and
-\emph{semi-empirical theories}, in which there exists a rigorous asymptotic theory, which comes with parameters, for the system of interest, and one then adjusts or fits those parameters to the finite non-asymptotic data.
+More generally, our results suggest what a practical theory of DNNs should look like.
+To see this, let's distinguish between two types of theories:
+non-empirical or analogical theories, in which one creates, often from general principles, a very simple toy model that can be analyzed rigorously, and one then claims that the model is relevant to the system of interest; and
+semi-empirical theories, in which there exists a rigorous asymptotic theory, which comes with parameters, for the system of interest, and one then adjusts or fits those parameters to the finite non-asymptotic data, to make predictions about practical problems.
 A drawback of the former approach is that it typically makes very strong assumptions, and the strength of those assumptions can limit the practical applicability of the theory.
@@
 Our approach focuses on the latter type of theory.
-Our results, which are based on \emph{using} sophisticated statistical mechanics theory and \emph{solving} important practical DNN problems, suggests that the latter approach should be of interest more generally for those interested in developing a \emph{practical DNN~theory}.
+Our results, which are based on using sophisticated statistical mechanics theory and solving important practical DNN problems, suggests that the latter approach should be of interest more generally for those interested in developing a practical DNN~theory.

### 2020-10-01 submission/NatCom_introduction.tex
pass to address most of the presubmission comments for nat comm

@@
 
-Most importantly, in industry, one faces unique practical problems such as: do we have enough data for this model? 
+Most importantly, in industry, one faces unique practical problems such as determining whether one has enough data for a given model. 
 Indeed, high quality, labeled data can be very expensive to acquire, and this cost can make or break a project.
@@
 Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves.
-In both cases, we ask:
-Can we predict trends in the quality of pretrained DNN models without access to training or testing data?
+In both cases, one can ask whether it is possible to predict trends in the quality of pretrained DNN models without access to training or testing data.
 
@@
 Second, norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.).
-For example, they may exhibit what we call \emph{Scale Collapse} for these models. 
+For example, they may exhibit what we call Scale Collapse for these models. 
 Third, PL-based metrics do much better at predicting quality trends in pretrained CV/NLP models.  
 They are qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among a series of well-trained versus very-well-trained models.
-Fourth, PL-based metrics can also be used to characterize fine-scale model properties, including what we call layer-wise \emph{Correlation Flow}, in well-trained and poorly-trained models; and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
+Fourth, PL-based metrics can also be used to characterize fine-scale model properties, including what we call layer-wise Correlation Flow, in well-trained and poorly-trained models; and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
 %

### 2020-10-01 submission/NatCom_methods.tex
pass to address most of the presubmission comments for nat comm

@@
 
-Consider the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$, as
-the minimization of a general Loss Function $\mathcal{L}$ over $N$ training data instances and labels, $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$.
-For a typical supervised classification problem, one has
-\begin{equation}
-\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N}\left(\mathcal{L}(E_{DNN}(\mathbf{x}_{i})-y_{i})\right)  ,
-\end{equation}
-where the loss function $\mathcal{L}(\cdot)$ can take on a myriad of forms~\cite{JC17_TR}, and where
-\begin{equation}
-E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})
-\label{eqn:dnn_energy}
-\end{equation}
-is the Energy (or optimization) Landscape function.  
-$E_{DNN}$ does not explicitly depend on the data, but rather it maps data instance vectors ($\mathbf{x}_i$ values) to predictions (labels).
-Therefore, one can analyze $E_{DNN}$ in the absence of any training or test data.
+To be fully reproducible, we only examine publicly-available, pretrained models.
+All of our computations were performed with the \texttt{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}, and we provide all Jupyter and Google Colab notebooks used in an accompanying github repository~\cite{kdd20_sub_repo}, which includes more details and more results.
 
-Test accuracies have been reported online for publicly-available pretrained pytroch models.
-These models have been trained and evaluated on labeled data $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$, using standard techniques.  
-We do not have access to this data, and we have not trained any of the models ourselves. 
-Our methodological approach is thus similar to a statistical meta-analysis, common in biomedical research, but uncommon in ML.
+\paragraph{Additional Details on Layer Weight Matrices}
+
+Recall that we can express the objective/optimization function for a typical DNN with $L$ layers, with activation functions $h_{l}(\cdot)$, and with $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$ as Equation~(\ref{eqn:dnn_energy}).
+We expect that most well-trained, production-quality models will employ one or more forms of regularization, such as Batch Normalization (BN), Dropout, etc., and many will also contain additional structure such as Skip Connections, etc.
+Here, we will ignore these details, and will focus only on the pretrained layer weight matrices $\mathbf{W}_{l}$.
+Typically, this model would be trained on some labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using Backprop, by minimizing the loss $\mathcal{L}$.
+For simplicity, we do not indicate the structural details of the layers (e.g., Dense or not, Convolutions or not, Residual/Skip Connections, etc.). 
+Each layer is defined by one or more layer 2D weight matrices $\mathbf{W}_{l}$, and/or the 2D feature maps $\mathbf{W}_{l,i}$ extracted from 2D Convolutional (Conv2D) layers.
+A typical modern DNN may have anywhere between 5 and 5000 2D layer matrices.
+
+For each Linear Layer, we get a  single $(N\times M)$ (real-valued) 2D weight matrix, denoted $\mathbf{W}_{l}$, for layer $l$.  
+This includes Dense or Fully-Connected (FC) layers, as well as 1D Convolutional (Conv1D) layers, Attention matrices, etc.
+We ignore the bias terms $\mathbf{b}_{l}$ in this analysis. 
+Let the aspect ratio be $Q=\frac{N}{M}$, with $Q\ge 1$.
+For the Conv2D layers, we have a 4-index Tensor, of the form $(N\times M \times c\times d)$, consisting
+of $c\times d$ 2D feature maps of shape $(N\times M)$.    
+We  extract $n_{l}=c\times d$ 2D weight matrices $\mathbf{W}_{l,i}$, one for each feature map $i=[1,\dots,n_{l}]$ for layer $l$.
+
+\paragraph{SVD of Convolutional 2D Layers.}
+
+There is some ambiguity in performing spectral analysis on Conv2D layers.  
+Each layer is a 4-index tensor of dimension $(w,h,in,out)$, with an $(w\times h)$ filter (or kernel) and $(in, out)$
+channels. When $w=h=k$, it gives $(k\times k)$ tensor slices, or pre-Activation Maps, $\mathbf{W}_{i,L}$ of dimension $(in\times out)$ each. 
 %
-Computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
-To be fully reproducible, we only examine publicly-available, pretrained models, and we provide all Jupyter and Google Colab notebooks used in an accompanying github repository~\cite{kdd20_sub_repo}.
-See Appendix~\ref{sxn:appendix} for details.
+We identify 3 different approaches for running SVD on a Conv2D layer:
+\begin{enumerate}
+\item run SVD on each pre-Activation Map $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values;
+\item stack the maps into a single matrix of, say, dimension $((k\times k\times out)\times in)$, and run SVD to get $in$ singular values;
+\item compute the 2D Fourier Transform (FFT) for each of the $(in, out)$ pairs, and run SVD on the Fourier coefficients~\cite{CNNSVD}, leading to $\sim(k\times in\times out)$ non-zero singular values.
+\end{enumerate}
+Each method has tradeoffs.  
+Method (3) is mathematically sound, but computationally expensive. Method (2) is ambiguous.
+For our analysis, because we need thousands of runs, we select method (1), which is the fastest (and is easiest to reproduce).
 
-\paragraph{Metrics for DNN Weight Matrices.}
+\paragraph{Normalization of Empirical Matrices.}
 
-Our approach involves analyzing DNN weight matrices.
-Each DNN layer contains one or more layer 2D  $N\times M$ weight matrices, $\mathbf{W}_{l}$, or pre-activation maps, $\mathbf{W}_{i,l}$, e.g., extracted from 2D Convolutional layers, where $N > M$.
-See Appendix~\ref{sxn:appendix} for details.
-(We may drop the $i$ and/or $i,l$ subscripts below.)
-The best performing quality metrics depend on the norms and/or spectral properties of each weight matrix,
-$\mathbf{W}$, and/or, equivalently, it's Empirical Correlation Matrix, $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
-We consider the following metrics:
-\begin{eqnarray}
-& & \text{Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{X}\Vert_{F}=\sum\nolimits_{i=1}^{M} \lambda_{i}$ } \\
-& & \text{Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}^{2}=\Vert\mathbf{X}\Vert_{\infty}=\lambda_{max}$ } \\
-& & \text{Weighted Alpha: $\hat{\alpha}=\alpha\log\lambda_{max}$ } \\
-& & \text{$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{W}\Vert^{2\alpha}_{2\alpha}=\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum\nolimits_{i=1}^{M}\lambda_{i}^{\alpha}$. }
-\end{eqnarray}
-Here, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the $\mathbf{X}$, and $\lambda_{max}$ is the maximum eigenvalue.
-These eigenvalues are squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, $\lambda_{i}=\sigma^{2}_{i}$.
-%See Appendix~\ref{sxn:appendix} for details. 
-All four metrics can be computed easily from DNN weight matrices.
-The first two metrics are well-known in ML.
-The last two metrics deserve special mention, as they depend on an empirical parameter $\alpha$ that is the PL exponent that arises in the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+Normalization is an important, if underappreciated, practical issue.
+Importantly, the normalization of weight matrices does \emph{not} affect the PL fits because $\alpha$ is scale-invariant.
+Norm-based metrics, however, do depend strongly on the scale of the weight matrix---that is the point.
+To apply RMT, we usually define $\mathbf{X}$ with a $1/N$ normalization, assuming variance of $\sigma^{2}=1.0$.
+Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with $\sigma^{2}\sim 1/\sqrt{N}$, or some variant, e.g., the Glorot/Xavier normalization~\cite{GloBen10}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.
+With this implicit scale, we do \emph{not} ``renormalize'' the empirical weight matrices, i.e., we use them as-is.
+The only exception is that \emph{we do rescale} the Conv2D pre-activation maps $\mathbf{W}_{i,L}$ by $k/\sqrt{2}$ so that they are on the same scale as the Linear / Fully Connected (FC) layers.
 
-\paragraph{Overview of Heavy-Tailed Self-Regularization.}
+\paragraph{Special consideration for NLP models.}
 
-In the HT-SR Theory, one analyzes the eigenvalue spectrum, i.e., the Empirical Spectral Density (ESD), of the associated correlation matrices~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
-From this, one characterizes the amount and form of correlation, and therefore implicit self-regularizartion, present in the DNN's weight matrices.
-For each layer weight matrix $\mathbf{W}$, of size $N \times M$, construct the associated $M\times M$ (uncentered) correlation matrix $\mathbf{X}$. 
-Dropping the $L$ and $l,i$ indices, one~has
-$$
-\mathbf{X} = \frac{1}{N}\mathbf{W}^{T}\mathbf{W}.
-$$
-If we compute the eigenvalue spectrum of $\mathbf{X}$, i.e., $\lambda_i$ such that
-$  % $$
-\mathbf{X}\mathbf{v}_{i}=\lambda_{i}\mathbf{v}_{i} , 
-$  % $$
-then the ESD of eigenvalues, $\rho(\lambda)$, is just a histogram of the eigenvalues, formally written as
-%\begin{equation}
-$\rho(\lambda)=\sum\nolimits_{i=1}^{M}\delta(\lambda-\lambda_{i})  .$
-%\label{eqn:eigenval_hist}
-%\end{equation}
-Using HT-SR Theory, one characterizes the correlations in a weight matrix by examining its ESD, $\rho(\lambda)$.
-It can be well-fit to a truncated PL distribution, given~as
-\begin{equation}
-\rho(\lambda)\sim\lambda^{-\alpha}  ,
-\label{eqn:eigenval_pl}
-\end{equation}
-which is (at least) valid within a bounded range of eigenvalues $\lambda\in[\lambda^{min},\lambda^{max}]$.
-
-The original work on HT-SR Theory considered a small number of NNs, including AlexNet and InceptionV3. 
-It showed that for nearly every $\mathbf{W}$, the (bulk and tail) of the ESDs can be fit to a truncated PL, and that PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
-%
-As for the mechanism responsible for these properties, statistical physics offers several possibilities~\cite{SornetteBook,nishimori01}, e.g., self organized criticality~\cite{SOC87,SOCat25yrs} or multiplicative noise in the stochastic optimization algorithms used to train these models~\cite{HodMah20A_TR,SorCon97}.
-Our meta-analysis does not require knowledge of mechanisms; and it is not even clear that one mechanism is responsible for every case.
-%
-Crucially, HT-SR Theory predicts that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models.
-Relatedly, previous work observed that smaller exponents $\alpha$ correspond to more implicit self-regularization and better generalization~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
-
-\paragraph{DNN Empirical Quality Metrics.}
-
-For norm-based metrics, we use the average of the log norm, to the appropriate power.
-Informally, this amounts to assuming that the layer weight matrices are statistically independent, in which case we can estimate the model complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm (which resembles a data dependent VC complexity),
-\begin{equation}
-\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert \times \cdots \times \Vert\mathbf{W}_{L}\Vert ,
-\end{equation}
-where $\Vert\cdot\Vert$ is a matrix norm.   
-The log complexity,
-\begin{equation}
-\label{eqn:eqn:sum_log_norm}
-\log\mathcal{C} \sim \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert + \cdots + \log\Vert\mathbf{W}_{L}\Vert = \sum\nolimits_l \log\Vert\mathbf{W}_{l}\Vert ,
-\end{equation}
- takes the form of an average Log Norm.
-For the \emph{Frobenius Norm metric} and \emph{Spectral Norm metric}, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly (since, when taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor).
-
-The \emph{Weighted Alpha metric} is an average of $\alpha_l$ over all layers $l \in \{1,\ldots,l\}$, weighted by the size, or scale, or each matrix,
-\begin{equation}
-\hat{\alpha} = \dfrac{1}{L}\sum_l \alpha_l\log\lambda_{max,l}\approx\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle    ,
-\end{equation}
-where $L$ is the total number of layer weight matrices.
-The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a much smaller and more limited set of models than we consider here.
-
-Based on this, in this paper, we introduce and evaluate the \emph{$\alpha$-Shatten Norm metric},
-\begin{equation}
-\label{eqn:sum_log_alpha_norm_alpha}
-\sum\nolimits_l \log \Vert\mathbf{X}_l\Vert_{\alpha_l}^{\alpha_l} 
-=
-\sum\nolimits_l \alpha_l \log \Vert\mathbf{X}_l\Vert_{\alpha_l} .
-\end{equation}
-For the $\alpha$-Shatten Norm metric, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it can not be taken out of the sum.
-For small $\alpha$, the Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}; and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm.
-
-One determines $\alpha$ for a given layer by fitting the ESD of that layer's weight matrix to a truncated PL, using the commonly accepted Maximum Likelihood method~\cite{CSN09_powerlaw,ABP14}.
-This method works very well for exponents between $\alpha\in(2,4)$; and it is adequate, although imprecise, for smaller and especially larger $\alpha$~\cite{newman2005_zipf}. 
-%
-Operationally, $\alpha$ is determined by using the \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the histogram of eigenvalues, $\rho(\lambda)$, to a truncated PL, 
-\begin{equation}
-\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\in[\lambda_{min},\lambda_{max}] ,
-\end{equation}
-where $\lambda_{max}$ is the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$, and 
-where $\lambda_{min}$ is selected automatically to yield the best (in the sense of minimizing the K-S distance) PL fit.
-Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.
-See Figure~\ref{fig:ww} for an illustration.
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=15.0cm]{img/WeightWatcher_v2}
-    \caption{Schematic of analyzing DNN layer weight matrices $\mathbf{W}$.
-Norm-based metrics (not shown) and PL-based metrics (depending on fitting the histogram of eigenvalues to a truncated PL) can be used to compare models (or for other tasks such as improved model training).}
-    \label{fig:ww}
-\end{figure}
-
-To avoid confusion, let us clarify the relationship between $\alpha$ and $\hat{\alpha}$.  
-We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated PL, parameterized by 2 values: the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.
-%%\footnote{Technically, we also need the minimum eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.}
-The PL exponent $\alpha$ measures the amount of correlation in a DNN layer weight matrix $\mathbf{W}$. 
-It is valid for $\lambda\le\lambda_{max}$, and it is scale-invariant, i.e., it does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.
-The $\lambda_{max}$ is a measure of the size, or scale, of $\mathbf{W}$.
-Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric $\hat{\alpha}$ for the entire~DNN.
+NLP models, and other models with large initial embeddings, require special care because the embedding layers frequently lack the implicit $1/\sqrt{N}$ normalization present in other layers.
+For example, in GPT, for most layers, the maximum eigenvalue $\lambda_{max}\sim\mathcal{O}(10-100)$, but in the first embedding layer, the maximum eigenvalue is of order $N$ (the number of words in the embedding), or $\lambda_{max}\sim\mathcal{O}(10^{5})$.
+For GPT and GPT2, we treat all layers as-is (although one may want to normalize the first 2 layers $\mathbf{X}$ by $1/N$, or to treat them as outliers).

### 2020-10-01 submission/NatCom_nlp_models.tex
pass to address most of the presubmission comments for nat comm

@@
 (in particular, the interpretation of values of $\alpha \sim 2$ as modeling systems with good correlations over many size scales~\cite{BouchaudPotters03, SornetteBook}), 
-this suggests that these models fail to capture successfully many of the information correlations in the data (relative to their size) and thus are substantially \emph{under}-trained.
+this suggests that these models fail to capture successfully many of the information correlations in the data (relative to their size) and thus are substantially under-trained.
 More generally, compared to CV models,
@@
 
-\paragraph{What do large values of $\alpha$ mean?}
-
 Many NLP models, such as early variants of GPT and BERT, have weight matrices with unusually large PL exponents (e.g., $\alpha\gg 6$).
-This indicates these matrices may be \emph{under}-correlated (i.e., over-parameterized, relative to the amount of data).
+This indicates these matrices may be under-correlated (i.e., over-parameterized, relative to the amount of data).
 In this regime, the truncated PL fit itself may not be very reliable because the Maximum Likelihood estimator it uses is unreliable in this range.
 The specific $\alpha$ values returned by the truncated PL fits are less reliable, but having large versus small $\alpha$ is reliable.
-If the ESD is visually examined, one can usually describe these $\mathbf{W}$ as in the \emph{Bulk-Decay} or \emph{Bulk-plus-Spikes} phase from HT-ST Theory~\cite{MM18_TR,MM19_HTSR_ICML}.
-Previous work~\cite{MM18_TR,MM19_HTSR_ICML} has conjectured that very well-trained DNNs would not have many \emph{outlier} $\alpha\gg 6$.
-Consistent with this, more recent improved versions of GPT (shown below) and BERT (not shown) confirm this.
+If the ESD is visually examined, one can usually describe these $\mathbf{W}$ as in the Bulk-Decay or Bulk-plus-Spikes phase from HT-ST Theory~\cite{MM18_TR,MM19_HTSR_ICML}.
+Previous work~\cite{MM18_TR,MM19_HTSR_ICML} has conjectured that very well-trained DNNs would not have many outlier $\alpha\gg 6$.
+Consistent with this, more recent improved versions of GPT (shown below) and BERT (not shown) confirm~this.
 
@@
 GPT has an initial Token and Positional Embedding layers, of dimension $(40478\times 768)$ and $(512\times 768)$, respectively, whereas GPT2 has input Embeddings of shape $(50257\times 768)$ and $(1024\times 768)$, respectively.
-The OpenAI GPT2 (English) models are: GPT2-small, GPT2-medium, GPT2-large, and GPT2-xl, having $12$, $24$, $36$, and $48$ layers, respectively, with increasingly larger weight matrices.
+The OpenAI GPT2 (English) models are: GPT2-small, GPT2-medium, GPT2-large, and GPT2-xl, having $12$, $24$, $36$, and $48$ layers, respectively, with increasingly larger weight~matrices.
 }
@@
 observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
-In the poorly-trained GPT model, the norm metrics all exhibit \emph{Scale Collapse}, compared to well-trained GPT2-small model.
+In the poorly-trained GPT model, the norm metrics all exhibit Scale Collapse, compared to well-trained GPT2-small model.
 
@@
 observe that (with one minor exception involving the log Frobenius norm metric) all four metrics decrease as one goes from medium to large to xl, indicating that the larger models indeed look better than the smaller models.
-For these well-trained models, the norm metrics now behave as expected, i.e., there is no \emph{Scale Collapse} and norms are decreasing with increasing~accuracy.
+For these well-trained models, the norm metrics now behave as expected, i.e., there is no Scale Collapse and norms are decreasing with increasing~accuracy.
 
@@
 This violates the conventional wisdom that smaller Spectral norms are better.
-Because there are so many anonymously small Spectral norms, the GPT model may be exhibiting a kind of \emph{Scale Collapse}, like that observed in the distilled CV models (Figure~\ref{fig:resnet204D5L}).
+Because there are so many anonymously small Spectral norms, the GPT model may be exhibiting a kind of Scale Collapse, like that observed in the distilled CV models (Figure~\ref{fig:resnet204D5L}).
 
-This demonstrates that, while the Spectral (or Frobenius) norm may correlate well with predicted test error, at least among reasonably well-trained models, it is \emph{not} a good indicator of the overall model quality in general.
-It can mispredict well-trained versus poorly-trained questions in ways not seen with PL-based metrics.
+This demonstrates that, while the Spectral (or Frobenius) norm may correlate well with predicted test error, at least among reasonably well-trained models, it is not a good indicator of the overall model quality in general.
 Using it as an empirical quality metric may give spurious results when applied to poorly-trained or otherwise deficient~models.
@@
        }
-   \caption{Histogram of PL exponents ($\alpha$) and Log Spectral Norms ($\log\Vert\mathbf{W}\Vert_{\infty}$) for weight matrices from the OpenAI GPT and GPT2-small pretrained models.}
+   \caption{Histogram of PL exponents 
+           %($\alpha$) 
+           and Log Spectral Norms 
+           %($\log\Vert\mathbf{W}\Vert_{\infty}$) 
+           for weight matrices from the OpenAI GPT and GPT2-small pretrained models.}
 
@@
 These layers have a different effective normalization, and therefore a different scale.
-See Appendix~\ref{sxn:appendix} for details.
+See the Supplementary Information
+%~\ref{sxn:appendix} 
+for details.
 Here, we do not include them in our computed average metrics in Table~\ref{table:nlp}, and we do not include them in the histogram plot in Figure~\ref{fig:GPT-snorm-hist}.
@@
     }
-    \caption{PL exponents ($\alpha$) (in (a)) and Log Spectral Norms ($\log\Vert\mathbf{W}\Vert_{\infty}$) (in (b)) for weight matrices from the OpenAI GPT and GPT2-small pretrained models.
-(Note that the quantities being shown on each Y axis are different.)
-             In the text, this is interpreted in terms of \emph{Correlation Flow} and \emph{Scale~Collapse}.
+    \caption{PL exponents 
+             %($\alpha$) 
+             (in (a)) and Log Spectral Norms 
+             %($\log\Vert\mathbf{W}\Vert_{\infty}$) 
+             (in (b)) for weight matrices from the OpenAI GPT and GPT2-small pretrained models.  (Note that the quantities being shown on each Y axis are different.)
+             In the text, this is interpreted in terms of Correlation Flow and Scale~Collapse.
             }
@@
     }
-    \caption{Histogram of PL exponents ($\alpha$)
-              and Log Alpha Norm ($\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$) for weight matrices from models of different sizes in the GPT2 architecture series.
+    \caption{Histogram of PL exponents 
+             %($\alpha$)
+             and Log Alpha Norm 
+             %($\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}$) 
+             for weight matrices from models of different sizes in the GPT2 architecture series.
 (Plots omit the first 2 (embedding) layers, because they are normalized differently giving anomalously large values.)

### 2020-10-01 submission/NatCom_overall_approach.tex
pass to address most of the presubmission comments for nat comm

\subsection{Overall approach}
\label{sxn:approach}

Consider the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$, as 
the minimization of a general Loss Function $\mathcal{L}$ over $N$ training data instances and labels, $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$.
For a typical supervised classification problem, one has
\begin{equation}
\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N}\left(\mathcal{L}(E_{DNN}(\mathbf{x}_{i})-y_{i})\right)  ,
\end{equation}
where the loss function $\mathcal{L}(\cdot)$ can take on a myriad of forms~\cite{JC17_TR}, and where
\begin{equation}
E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})
\label{eqn:dnn_energy}
\end{equation}
is the Energy (or optimization) Landscape function.  
\michael{See comments for math equations, explain $h_{L}$, scalar valued function, $b_L$, times symbol, etc., confusing about role or $y$, and whether we are dealing with products of matrices.  E.g., $L$ layers, with activation functions $h_{l}(\cdot)$, and with $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$; etc.  }
$E_{DNN}$ does not explicitly depend on the data, but rather it maps data instance vectors ($\mathbf{x}_i$ values) to predictions (labels).
\michael{Mark found that sentence confusing.}
Therefore, one can analyze $E_{DNN}$ in the absence of any training or test~data.

Test accuracies have been reported online for publicly-available pretrained pyTorch models~\cite{osmr}.
These models have been trained and evaluated on labeled data $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$, using standard techniques.  
We do not have access to this data, and we have not trained any of the models ourselves. 
Our methodological approach is thus similar to a statistical meta-analysis, common in biomedical research, but uncommon in ML.
%
Computations were performed with the publicly-available \texttt{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
To be fully reproducible, we only examine publicly-available, pretrained models, and we provide all Jupyter and Google Colab notebooks used in an accompanying github repository~\cite{kdd20_sub_repo}.
See the Supplementary Information
%~\ref{sxn:appendix} 
for details.

\paragraph{Metrics for DNN Weight Matrices.}

Our approach involves analyzing individual DNN weight matrices, for (depending on the architecture) fully-connected and/or convolutional layers.
Each DNN layer contains one or more layer 2D  $N\times M$ weight matrices, $\mathbf{W}_{l}$, or pre-activation maps, $\mathbf{W}_{i,l}$, e.g., extracted from 2D Convolutional layers, where $N > M$.
See the Supplementary Information
%~\ref{sxn:appendix} 
for details.
(We may drop the $i$ and/or $i,l$ subscripts below.)
The best performing quality metrics depend on the norms and/or spectral properties of each weight matrix,
$\mathbf{W}$, and/or, equivalently, it's Empirical Correlation Matrix, $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
We consider the following metrics:
\begin{eqnarray}
& & \text{Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{X}\Vert_{F}=\sum\nolimits_{i=1}^{M} \lambda_{i}$ } \\
& & \text{Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}^{2}=\Vert\mathbf{X}\Vert_{\infty}=\lambda_{max}$ } \\
& & \text{Weighted Alpha: $\hat{\alpha}=\alpha\log\lambda_{max}$ } \\
& & \text{$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{W}\Vert^{2\alpha}_{2\alpha}=\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum\nolimits_{i=1}^{M}\lambda_{i}^{\alpha}$. }
\end{eqnarray}
Here, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the $\mathbf{X}$, and $\lambda_{max}$ is the maximum eigenvalue.
These eigenvalues are squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, $\lambda_{i}=\sigma^{2}_{i}$.
All four metrics can be computed easily from DNN weight matrices.
The first two metrics are well-known in ML.
The last two metrics deserve special mention, as they depend on an empirical parameter $\alpha$ that is the PL exponent that arises in the recently-developed Heavy Tailed Self Regularization (HT-SR) Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.

\paragraph{Overview of Heavy-Tailed Self-Regularization.}

In the HT-SR Theory, one analyzes the eigenvalue spectrum, i.e., the Empirical Spectral Density (ESD), of the associated correlation matrices~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
From this, one characterizes the amount and form of correlation, and therefore implicit self-regularizartion, present in the DNN's weight matrices.
For each layer weight matrix $\mathbf{W}$, of size $N \times M$, construct the associated $M\times M$ (uncentered) correlation matrix $\mathbf{X}$. 
Dropping the $L$ and $l,i$ indices, one~has
$$
\mathbf{X} = \frac{1}{N}\mathbf{W}^{T}\mathbf{W}.
$$
If we compute the eigenvalue spectrum of $\mathbf{X}$, i.e., $\lambda_i$ such that
$  % $$
\mathbf{X}\mathbf{v}_{i}=\lambda_{i}\mathbf{v}_{i} , 
$  % $$
then the ESD of eigenvalues, $\rho(\lambda)$, is just a histogram of the eigenvalues, formally written as
%\begin{equation}
$\rho(\lambda)=\sum\nolimits_{i=1}^{M}\delta(\lambda-\lambda_{i})  .$
%\label{eqn:eigenval_hist}
%\end{equation}
Using HT-SR Theory, one characterizes the correlations in a weight matrix by examining its ESD, $\rho(\lambda)$.
It can be well-fit to a truncated PL distribution, given~as
\begin{equation}
\rho(\lambda)\sim\lambda^{-\alpha}  ,
\label{eqn:eigenval_pl}
\end{equation}
which is (at least) valid within a bounded range of eigenvalues $\lambda\in[\lambda^{min},\lambda^{max}]$.

The original work on HT-SR Theory considered a small number of NNs, including AlexNet and InceptionV3. 
It showed that for nearly every $\mathbf{W}$, the (bulk and tail) of the ESDs can be fit to a truncated PL, and that PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
%
As for the mechanism responsible for these properties, statistical physics offers several possibilities~\cite{SornetteBook,nishimori01}, e.g., self organized criticality~\cite{SOC87,SOCat25yrs} or multiplicative noise in the stochastic optimization algorithms used to train these models~\cite{HodMah20A_TR,SorCon97}.
Our meta-analysis does not require knowledge of mechanisms; and it is not even clear that one mechanism is responsible for every case.
%
Crucially, HT-SR Theory predicts that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models.
Relatedly, previous work observed that smaller exponents $\alpha$ correspond to more implicit self-regularization and better generalization~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.

\paragraph{DNN Empirical Quality Metrics.}

For norm-based metrics, we use the average of the log norm, to the appropriate power.
Informally, this amounts to assuming that the layer weight matrices are statistically independent, in which case we can estimate the model complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm (which resembles a data dependent VC complexity),
\begin{equation}
\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert \times \cdots \times \Vert\mathbf{W}_{L}\Vert ,
\end{equation}
where $\Vert\cdot\Vert$ is a matrix norm.   
The log complexity,
\begin{equation}
\label{eqn:eqn:sum_log_norm}
\log\mathcal{C} \sim \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert + \cdots + \log\Vert\mathbf{W}_{L}\Vert = \sum\nolimits_l \log\Vert\mathbf{W}_{l}\Vert ,
\end{equation}
 takes the form of an average Log Norm.
For the Frobenius Norm metric and Spectral Norm metric, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly (since, when taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor).

The Weighted Alpha metric is an average of $\alpha_l$ over all layers $l \in \{1,\ldots,l\}$, weighted by the size, or scale, or each matrix,
\begin{equation}
\hat{\alpha} = \dfrac{1}{L}\sum_l \alpha_l\log\lambda_{max,l}\approx\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle    ,
\end{equation}
where $L$ is the total number of layer weight matrices.
The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a much smaller and more limited set of models than we consider here.

Based on this, in this paper, we introduce and evaluate the $\alpha$-Shatten Norm metric,
\begin{equation}
\label{eqn:sum_log_alpha_norm_alpha}
\sum\nolimits_l \log \Vert\mathbf{X}_l\Vert_{\alpha_l}^{\alpha_l} 
=
\sum\nolimits_l \alpha_l \log \Vert\mathbf{X}_l\Vert_{\alpha_l} .
\end{equation}
For the $\alpha$-Shatten Norm metric, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it can not be taken out of the sum.
For small $\alpha$, the Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}; and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm.

One determines $\alpha$ for a given layer by fitting the ESD of that layer's weight matrix to a truncated PL, using the commonly accepted Maximum Likelihood method~\cite{CSN09_powerlaw,ABP14}.
This method works very well for exponents between $\alpha\in(2,4)$; and it is adequate, although imprecise, for smaller and especially larger $\alpha$~\cite{newman2005_zipf}. 
%
Operationally, $\alpha$ is determined by using the \texttt{WeightWatcher} tool~\cite{weightwatcher_package} to fit the histogram of eigenvalues, $\rho(\lambda)$, to a truncated PL, 
\begin{equation}
\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\in[\lambda_{min},\lambda_{max}] ,
\end{equation}
where $\lambda_{max}$ is the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$, and 
where $\lambda_{min}$ is selected automatically to yield the best (in the sense of minimizing the K-S distance) PL fit.
Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.
See Figure~\ref{fig:ww} for an illustration.

\begin{figure}[t]
    \centering
    %\includegraphics[width=15.0cm]{img/WeightWatcher_v2}
    \includegraphics[width=15.0cm]{../img2/WeightWatcher_v2}
    \caption{Schematic of analyzing DNN layer weight matrices $\mathbf{W}$.  
             Given an individual layer weight matrix $\mathbf{W}$, from either a fully-connected layer or a convolutional layer, perform a Singular Value Decomposition (SVD) to obtain $\mathbf{W} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^{T}$, and examine the histogram of eigenvalues of $\mathbf{W}^{T}\mathbf{W}$.
             Norm-based metrics (not shown) and PL-based metrics (that depend on fitting the histogram of eigenvalues to a truncated PL) can be used to compare models.
             For example, one can analyze one layer of a pre-trained model, compare multiple layers of a pre-trained model, make comparisons across model architectures, monitor neural network properties during training, etc. 
             \michael{Put the Analyze, Compare, Monitor comments in the caption, not the figure.}
}
    \label{fig:ww}
\end{figure}

To avoid confusion, let us clarify the relationship between $\alpha$ and $\hat{\alpha}$.  
We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated PL, parameterized by 2 values: the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.
The PL exponent $\alpha$ measures the amount of correlation in a DNN layer weight matrix $\mathbf{W}$. 
It is valid for $\lambda\le\lambda_{max}$, and it is scale-invariant, i.e., it does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.
The $\lambda_{max}$ is a measure of the size, or scale, of $\mathbf{W}$.
Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric $\hat{\alpha}$ for the entire~DNN.

### 2020-10-01 submission/empirical2_NatComfmt.tex
pass to address most of the presubmission comments for nat comm

@@
 \newcommand {\ncyan}[1]{{\color{cyan}\sf{[#1]}}}
+
+%% MWM: Do the following to remove the comments from the pdf.
+%\newcommand {\michael}[1]{{ }}
 \newcommand {\michael}[1]{{\color{red}\sf{[michael: #1]}}}
+
 \newcommand {\charles}[1]{{\color{blue}\sf{[charles: #1]}}}
@@
 Michael W. Mahoney\thanks{ICSI and Department of Statistics, University of California at Berkeley, Berkeley, CA 94720, \texttt{mmahoney@stat.berkeley.edu}.}
+\thanks{Corresponding author.}
 }
@@
 \input{NatCom_introduction}
-\input{NatCom_methods}
 
 \section{Results}
-We study in detail three well-known CV architectures (the VGG, ResNet, and DenseNet series of models).
-Then, we look in detail at several variations of a popular NLP architecture (the OpenAI GPT and GPT2 models).
-Finally, we present results from a broader analysis of hundreds of pretrained DNN~models.
+After describing our overall approach, we study in detail three well-known CV architectures (the VGG, ResNet, and DenseNet series of models).
+Then, we look in detail at several variations of a popular NLP architecture (the OpenAI GPT and GPT2 models), and we present results from a broader analysis of hundreds of pretrained DNN~models.
 
+\input{NatCom_overall_approach}
 \input{NatCom_cv_models}
@@
 \input{NatCom_discussion}
+
+\input{NatCom_methods}
 

### 2020-10-02 submission/NatCom_methods.tex
more on address presubmission comments

@@
 
-Recall that we can express the objective/optimization function for a typical DNN with $L$ layers, with activation functions $h_{l}(\cdot)$, and with $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$ as Equation~(\ref{eqn:dnn_energy}).
+Recall that we can express the objective/optimization function for a typical DNN with $L$ layers and with $N\times M$ weight matrices $\mathbf{W}_{l}$ and bias vectors $\mathbf{b}_{l}$ as Equation~(\ref{eqn:dnn_energy}).
 We expect that most well-trained, production-quality models will employ one or more forms of regularization, such as Batch Normalization (BN), Dropout, etc., and many will also contain additional structure such as Skip Connections, etc.
@@
 Each layer is defined by one or more layer 2D weight matrices $\mathbf{W}_{l}$, and/or the 2D feature maps $\mathbf{W}_{l,i}$ extracted from 2D Convolutional (Conv2D) layers.
-A typical modern DNN may have anywhere between 5 and 5000 2D layer matrices.
+A typical modern DNN may have anywhere between 5 and 5000 2D layer~matrices.
 

### 2020-10-02 submission/NatCom_overall_approach.tex
adjust img2 to img
more on address presubmission comments

@@
 
-Consider the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$, as
-the minimization of a general Loss Function $\mathcal{L}$ over $N$ training data instances and labels, $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$.
-For a typical supervised classification problem, one has
+Consider the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, and weight matrices $\mathbf{W}_{l}$ and bias vectors $\mathbf{b}_{l}$, as
+the minimization of a general Loss Function $\mathcal{L}$ over the training data instances and labels, $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$.
+For a typical supervised classification problem, the goal of training is to construct $\mathbf{W}_{l}$ and $\mathbf{b}_{l}$ that capture correlations in the data, in the sense of solving
 \begin{equation}
@@
 \end{equation}
-where the loss function $\mathcal{L}(\cdot)$ can take on a myriad of forms~\cite{JC17_TR}, and where
+where the loss function $\mathcal{L}(\cdot)$ can take on a myriad of forms~\cite{JC17_TR}, and where the Energy (or optimization) Landscape function
 \begin{equation}
-E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})
+%E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})
+E_{DNN} = f(\mathbf{W}_{1},\ldots,\mathbf{W}_{L},\mathbf{b}_{1},\ldots,\mathbf{b}_{L})
 \label{eqn:dnn_energy}
 \end{equation}
-is the Energy (or optimization) Landscape function.  
-\michael{See comments for math equations, explain $h_{L}$, scalar valued function, $b_L$, times symbol, etc., confusing about role or $y$, and whether we are dealing with products of matrices.
-E.g., $L$ layers, with activation functions $h_{l}(\cdot)$, and with $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$; etc.
-}
-$E_{DNN}$ does not explicitly depend on the data, but rather it maps data instance vectors ($\mathbf{x}_i$ values) to predictions (labels).
+depends on the weights and biases.
+\michael{See comments for math equations.}
+For a trained model, $E_{DNN}$ does not explicitly depend on the data, but rather it maps data instance vectors ($\mathbf{x}_i$ values) to predictions ($y_{i}$ labels).
 \michael{Mark found that sentence confusing.}
@@
 Our approach involves analyzing individual DNN weight matrices, for (depending on the architecture) fully-connected and/or convolutional layers.
-Each DNN layer contains one or more layer 2D  $N\times M$ weight matrices, $\mathbf{W}_{l}$, or pre-activation maps, $\mathbf{W}_{i,l}$, e.g., extracted from 2D Convolutional layers, where $N > M$.
+Each DNN layer contains one or more layer 2D  $N_{l}\times M_{l}$ weight matrices, $\mathbf{W}_{l}$, or pre-activation maps, $\mathbf{W}_{i,l}$, e.g., extracted from 2D Convolutional layers, where $N > M$.
 See the Supplementary Information
@@
 As for the mechanism responsible for these properties, statistical physics offers several possibilities~\cite{SornetteBook,nishimori01}, e.g., self organized criticality~\cite{SOC87,SOCat25yrs} or multiplicative noise in the stochastic optimization algorithms used to train these models~\cite{HodMah20A_TR,SorCon97}.
+Alternatively, related techniques have been used to analyze correlations and information propogation in actual spiking neurons~\cite{SYYRP11,YKYP14}.
 Our meta-analysis does not require knowledge of mechanisms; and it is not even clear that one mechanism is responsible for every case.
@@
     %\includegraphics[width=15.0cm]{img/WeightWatcher_v2}
-    \includegraphics[width=15.0cm]{../img2/WeightWatcher_v2}
+    \includegraphics[width=15.0cm]{img/WeightWatcher_v3}
     \caption{Schematic of analyzing DNN layer weight matrices $\mathbf{W}$.  
@@
              For example, one can analyze one layer of a pre-trained model, compare multiple layers of a pre-trained model, make comparisons across model architectures, monitor neural network properties during training, etc.
-             \michael{Put the Analyze, Compare, Monitor comments in the caption, not the figure.}
 }

### 2020-10-03 submission/NatCom_all_models.tex
pass to address most of the presubmission comments for nat comm
minor changes before submit to Nat Comm

@@
 This broader analysis on a much larger set of CV and NLP models, with a more diverse set of architectures, and developed for a wider range of tasks, complements the previous more detailed analysis on CV and NLP models.
-See Appendix~\ref{sxn:appendix} for details.
-To quantify the relationship between quality metrics and the reported test error and/or accuracy metrics, we use ordinary least squares to regress the metrics on the Top1 (and Top5) reported errors (as dependent variables).
+See the Supplementary Information
+%~\ref{sxn:appendix} 
+for details.
+To quantify the relationship between quality metrics and the reported test error and/or accuracy metrics, we use ordinary least squares (similar results are seen with rank correlation metrics) to regress the metrics on the Top1 (and Top5) reported errors (as dependent variables).
 These include Top5 errors for the ImageNet-1K model, percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy (Pix.Acc.) and Intersection-Over-Union (IOU) for other models.
@@
 \begin{tabular}{|p{1in}|c|c|c|c|}
-%KDD% \begin{tabular}{|p{0.75in}|c|c|c|c|}
 \hline
@@
          We include regressions only for architectures with five or more data points, and which are positively correlated with test error.
-         These results can be readily reproduced using the Google Colab notebooks. (See Appendix~\ref{sxn:appendix} for details.) 
+         These results can be readily reproduced using the Google Colab notebooks. 
+         (See the Supplementary Information
+         %~\ref{sxn:appendix} 
+         for details.) 
         }

### 2020-10-03 submission/NatCom_overall_approach.tex
minor changes before submit to Nat Comm

@@
 Consider the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, and weight matrices $\mathbf{W}_{l}$ and bias vectors $\mathbf{b}_{l}$, as
-the minimization of a general Loss Function $\mathcal{L}$ over the training data instances and labels, $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$.
-For a typical supervised classification problem, the goal of training is to construct $\mathbf{W}_{l}$ and $\mathbf{b}_{l}$ that capture correlations in the data, in the sense of solving
+the minimization of a general loss function $\mathcal{L}$ over the training data instances and labels, $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$.
+For a typical supervised classification problem, the goal of training is to construct (or learn) $\mathbf{W}_{l}$ and $\mathbf{b}_{l}$ that capture correlations in the data, in the sense of solving
 \begin{equation}
@@
 \end{equation}
-where the loss function $\mathcal{L}(\cdot)$ can take on a myriad of forms~\cite{JC17_TR}, and where the Energy (or optimization) Landscape function
+where the loss function $\mathcal{L}(\cdot)$ can take on a myriad of forms~\cite{JC17_TR}, and where the energy (or optimization) landscape function
 \begin{equation}
 %E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})
-E_{DNN} = f(\mathbf{W}_{1},\ldots,\mathbf{W}_{L},\mathbf{b}_{1},\ldots,\mathbf{b}_{L})
+E_{DNN} = f(\mathbf{x}_{i} ; \mathbf{W}_{1},\ldots,\mathbf{W}_{L},\mathbf{b}_{1},\ldots,\mathbf{b}_{L})
 \label{eqn:dnn_energy}
 \end{equation}
-depends on the weights and biases.
+depends parametrically on the weights and biases.
 \michael{See comments for math equations.}
@@
 The best performing quality metrics depend on the norms and/or spectral properties of each weight matrix,
-$\mathbf{W}$, and/or, equivalently, it's Empirical Correlation Matrix, $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
+$\mathbf{W}$, and/or, equivalently, it's empirical correlation matrix, $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
 We consider the following metrics:

### 2020-12-28 submission/resp_to_revs_NatCom.tex
response to reviewers of nat comm

\documentclass[11pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\evensidemargin}{-0.5in}
\addtolength{\topmargin}{-1.0in}
\addtolength{\textheight}{1.7in}
\newlength{\defbaselineskip}
\setlength{\defbaselineskip}{\baselineskip}

%-----------------------------------------------------------------------

\begin{document}

\section*{Response to Reviewers
``Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data''
}

Thanks to the two reviewers for providing feedback.
One reviewer had questions about additional results to justify our main claims.
In some cases, we had these results previously, but they were not included due to space constraints.
In other cases, these results were easy to generate, and we are happy to include them.
This reviewer also had several misunderstandings, which we will clarify.
The other reviewer had several general questions, which we also happy to include and clarify.
We go through each of the points in turn below, with the reviewer comments in italics and our response immediately after.

\subsection*{Reviewer 1}

\emph{%
The paper presents a set of metrics to evaluate the quality of trained neural networks without having access to the training/testing data used to train/validate the models. 
This is achieved by taking advantage of regularization techniques to train neural networks. 
The paper evaluates different approaches such as norms of the learned weight matrices or parameters from power law fits of the eigenvalues of weight matrices. 
These proposed metrics are therefore evaluated on a large number of pre-trained convolutional neural network architectures and natural language processing models.
}

\begin{quote}
To clarify a potential misunderstanding, it is not correct that the metrics we consider are
``achieved by taking advantage of regularization techniques to train neural networks.'' 
Norms may be used to train neural networks, but power law fits are not.
Instead, for the power law fits, we use ideas from statistical mechanics and strongly correlated system theory to try to capture the idea that trained neural networks encode correlations from the data.
It is fair to say that the power law exponents characterize is the implicit regularization, whether or not the model trainer knew that that was what he or she was doing, but not any regularization explicitly used to train.
We clarify this point at Line XXX of the revised manuscript.
\end{quote}

\noindent
\emph{%
I like the idea of proposing a metric to evaluate the quality of already trained models. 
However, I believe the paper arrives to conclusions without using proper statistical analysis and it is too speculative:
}

\noindent
\emph{%
- The average quality vs Top 1 accuracy is tested using linear regression. However, the p-values are not reported and only the RMSE and $R^2$ is provided. Moreover, the assumption of normality of residuals is not tested. Therefore, we cannot conclude with RMSE and $R^2$ that there is a correlation between the metrics and the Top 1 accuracy.
}

\begin{quote}
XXX.  CM: 
The best thing is to run R2 and Ktau, not p-values, and then also look at the plots.
I'm open to other suggestions here.
I'll need to set this up this week, and also create a repo with ALL the plots.
\end{quote}

\noindent
\emph{%
- In the layer analysis, the paper analyzes how the PL exponent varies in the different layers. Figure 4 shows how the distribution of alpha among layers is architecture-dependent, having different distributions. Computing the mean of this value among all layers is in my opinion not a good metric to compare different models. Despite the fact, that the results of the linear regression used in Figure 9 are not reported, it seems to be the case as seen in the figure, that there is a poor correlation between trained models with different architectures.
}

\begin{quote}
(((XXX.  CM: 
The whole point is that a crude average works pretty well. 
We never claim that there is a correlation between different architectures.
)))\\
The reviewer is correct that the distribution of $\alpha$ is different for different architectures.
That is the point.
Different architectures can be more or less appropriate for a given task, and within an architecture series the metrics we present can be used to predict trends in model quality.
We never claim that there is a correlation between different architectures, and we know that there is not.
There are many reasons for this, e.g., different normalizations, etc.
As for For Figure 9, XXX.
%
We clarify these points at Line XXX, Line XXX, and Line XXX of the revised manuscript.
\end{quote}

\noindent
\emph{%
- In the first paragraph of the NLP models (line 325) the paper indicates that lower quality metrics represent better model, but Table 2 indicates the opposite. Also, when comparing the metrics with the depth of the network the paper compares GPT2-medium, large, and xl. However, it does not include in the comparison the small model, which will not present the nice trend of decreasing values of the metrics when increasing the depth of the model.
}

\begin{quote}
XXX.  CM: 
IDK what this is talking about.  We must have explained things improperly.  We should review.
\end{quote}

\noindent
\emph{%
- The paper also concludes that the alpha metric is a good metric to explain the quality based on Figure 6 and 8. However, these histograms look very similar with the exception of a few outliers. This should be analyzed in more detail.
}

\begin{quote}
(((XXX.  CM:
We can analyze in more detail sure.  Space provided.
)))\\
The reviewer is correct that the plots are similar.
The important point is how they are different.
For the GPT $\alpha$ plots, the overall distribution shifts slightly to smaller values of $\alpha$ and there are fewer large outliers, both indicating that the larger model is
XXX MM TO REFINE.
\end{quote}

\noindent
\emph{%
- The paper refers to results to support the conclusion, but these results are not reported in the paper or supplementary material -- ``not shown here''. If these results support the conclusion, they should be presented in the paper or at least the supplementary material.
}

\begin{quote}
XXX.  CM:
We can add a supplementary github repo.
\end{quote}

\noindent
\emph{%
So, while I like the idea and believe that the authors tackle an important topic, I believe that the conclusions they arrive at, are not necessarily supported by the experiments and the shown results. Hence, I believe the manuscript needs a major revision and additional experiments in order to support the claims. For these reasons my recommendation is to reject the paper, and encourage the authors to proceed with their work.
}

\begin{quote}
XXX.  CM:
We disagree.  We are happy to provide a more detailed experimental analysis and in gross detail.
Space is limited here, however. 
\end{quote}

\subsection*{Reviewer 3}

\noindent
\emph{%
Nowadays, much research effort is being put into developing a theory for deep learning that could predict a network's generalization error. Yet, for many of these theories, the rubber never meets the road, as no concrete guidelines are provided for practitioners that would allow them to predict the generalization error of state-of-the-art networks, trained on real data, across various hyperparameters, at scale. The authors of this paper undertake this important challenge in the specific, yet prevalent setting of transfer learning (and also model distillation), proposing novel statistics for predicting generalization performance, based on the spectrum of layer-wise weight matrices. Of particular interest is the methodology that the authors chose, relying heavily on data scientific experiments. This work exemplifies the potency of data science in delivering knowledge in the modern scientific era. I, therefore, recommend the acceptance of this publication to Nature Communications,
once the authors address the minor issues raised below.
}

\begin{quote}
Thanks for articulating at least as well as we did the broad goal of the paper.
\end{quote}

\noindent
\emph{%
- ``DNNs will exhibit correlations over many size scales'' -- could the authors elaborate on this statement and explain it for readers unaware of theories pertaining to strongly-correlated systems?
}

\begin{quote}
(((XXX.  CM:
Easy.
)))\\
We have elaborated on this at Line XXX of the revised manuscript.
\end{quote}

\noindent
\emph{%
- Would the authors consider the following logic to be correct: \\
- Smaller power-law exponents result in a more condensed spectrum; \\
- More condensed spectrum results in a smaller condition number; \\
- A smaller condition number facilitates the propagation of information/features across layers.
}

\begin{quote}
(((XXX.  CM:
We can discuss in detail.  The trick is in how the tail is estimated ...
)))\\
More important than the smaller condition number, which has to do with the ratio of the largest to smallest eigenvalue, is the ``shape'' of the spectral distribution, and in particular the details of the tail and how the tail is estimated.
We clarify this point at Line XXX of the revised manuscript.
\end{quote}

\noindent
\emph{%
- Could the authors elucidate the relation between the Information Bottleneck and the Theory of Heavy-Tailed Self Regularization?
}

\begin{quote}
(((XXX.  CM:
Pftt IDK.  We can think something up.
)))\\
Certainly.
We clarify this point at Line XXX of the revised manuscript.
\end{quote}

\noindent
\emph{%
- Would the authors consider rank collapse a phenomenon detrimental or beneficial to the network's performance? Why?
}

\begin{quote}
(((XXX.  CM:
Detrimental.  This is known in GANs, where regularizers are used to prevent rank collapse.
I'll have to think of a bigger reason.
)))\\
In general, it would be detrimental.
This is known in GANs, where regularizers are used to prevent rank collapse, but it appears to hold more generally.
We clarify this point at Line XXX of the revised manuscript.
\end{quote}

\noindent
\emph{%
- Could the authors define mathematically what is ``Scale Collapse''?
}

\begin{quote}
(((XXX.  CM:
This just means that the scale of the W matrix changes suddenly and unexpectedly.
)))\\
XXX MM QUES: ``suddenly and unexpectedly'' MEANS AS A FUNCTION OF SOMETHING THAT IS CHANGING.
\end{quote}

\noindent
\emph{%
- If one were to multiply all weights in the network by a constant, the spectral norm would change, but the classification error would not. Isn't the spectral norm of the weights therefore irrelevant for diagnosing deep networks?
}

\begin{quote}
(((XXX.  CM:
Yes and No.  Recall we are comparing networks with similar architectures.
One then has to choose a canonical way to normalize the network so that similar architectures can be compared.
While different networks do have different normalizations, we have not found any discrepancies across 
models in the same architecture series.
)))\\
Yes and No.  
Recall we are comparing networks with similar architectures.
One then has to choose a canonical way to normalize the network so that similar architectures can be compared.
While different networks do have different normalizations, we have not found any discrepancies across 
models in the same architecture series.
We clarify this point at Line XXX of the revised manuscript.
\end{quote}

\noindent
\emph{%
- The authors mention in the paper that weight matrices can be ``over-parameterized, relative to the amount of data.'' Is there any empirical evidence in the literature showing that networks can have worse generalization performance due to overparameterization? Isn't the double descent phenomenon showing that generalization error keeps decreasing after the second descent despite increasing overparameterization?
}

\begin{quote}
(((XXX.  CM:
I'll let you handle this one.
)))\\
XXX.  MM TO DO.
\end{quote}

\noindent
\emph{%
- Not all losses used in deep learning are a function of the difference between the network's prediction and the label (for example, cross-entropy). I would suggest changing the loss in Equation 1 to include two inputs: the prediction and the label.
}

\begin{quote}
(((XXX.  CM:
That's fair.
)))\\
That is a good point.
We clarify this point at Line XXX of the revised manuscript.
\end{quote}

\bibliographystyle{unsrt}
%\bibliographystyle{plain}
{\small
%\bibliography{gen_gap}
\bibliography{dnns}
%\bibliography{dnns,gen_gap}
}

\end{document}

### 2020-12-30 submission/resp_to_revs_NatCom.tex
v1 of response, without line nubmers

@@
 \begin{quote}
-XXX.  CM: 
-The best thing is to run R2 and Ktau, not p-values, and then also look at the plots.
-I'm open to other suggestions here.
-I'll need to set this up this week, and also create a repo with ALL the plots.
+%(((XXX.  CM: 
+%The best thing is to run R2 and Ktau, not p-values, and then also look at the plots.
+%I'm open to other suggestions here.
+%I'll need to set this up this week, and also create a repo with ALL the plots.
+%)))\\
+The reviewer is correct.  
+We intended for the reader to examine plots to see that the trends are qualitatively in the correct (or incorrect, in some cases for norm-based metrics) direction, and then to look at the numbers to get quantative insight.
+We know that errors are not normal; so, for this and other reasons, in our experience, reporting $p$-values won't clarify the situation.
+We have expanded all of the tables that report RMSE and/or $R^2$ to report the Kendal-$\tau$ rank correlation measure.
+Also, a point we didn't emphasize enough is that from HT-SR theory we expect that PL-based metrics to have strong linear and not just rank correlation.
+We have clarified these points at Line XXX, Line XXX, and Line XXX of the revised manuscript, and in the tables and plots throughout the manuscript.
+(Also, additional results beyond what we have in the main text and Supplementary Material are available in our fully-reproducible and publicly-available repo, and it is our hope that that level of transparency will aid the reader's understanding.)
 \end{quote}
@@
 \begin{quote}
-(((XXX.  CM: 
-The whole point is that a crude average works pretty well. 
-We never claim that there is a correlation between different architectures.
-)))\\
+%(((XXX.  CM: 
+%The whole point is that a crude average works pretty well. 
+%We never claim that there is a correlation between different architectures.
+%)))\\
 The reviewer is correct that the distribution of $\alpha$ is different for different architectures.
@@
 There are many reasons for this, e.g., different normalizations, etc.
-As for For Figure 9, XXX.
+We clarify this point at Line XXX of the revised manuscript.
+
+As for For Figure 9, this should be read in light of Tables 7, 8, and 9, as well as the discussion in Sec 2.4.  
+Our metrics for predicting trends in model quality perform well within a given architectural series, and we have discussed this in detail in the main text.
+We also want to show that our general results hold for the hundreds of publicly-available CV and NLP models.
+All of this is available in our publicly-available repo, and so we decided in the text and Supplementary Material to go with just a summary.
+In the main text, we have R2, MSE, and Kendall-Tau; and in the appendix, we have more granular results for several model series.
+Since Figure 9 partitions based on data and not model series, we think that it will likely confuse the reader.
+Thus, we removed it, and we will simply point to the publicly-available repo.
 %
@@
 \begin{quote}
-XXX.  CM: 
-IDK what this is talking about.  We must have explained things improperly.  We should review.
+%(((XXX.  CM: 
+%IDK what this is talking about.  We must have explained things improperly.  We should review.
+%)))\\
+Sorry for the confusion.
+We have adjusted that table, and we have substantially reworded the text.
+
+The first point is there are two things we are illustrating: one is when we increase both data size and model size (GPT2-small to GPT2-medium to GPT2-large to GPT2-xl); and the second is when a give model is trained to high-quality verus low-quality data (GPT2-small versus GPT).
+For the former, we see a general trend with all the metrics decreasing, with the largest difference between GPT2-medium and GPT2-large, and with the difference between GPT2-small and GPT2-medium very small (near the noise level, and slightly in the opposite direction).
+For the latter, we see GPT seems to violate what we expect. 
+Thus, we look in more detail at the contributions to alpha-hat, i.e., to the PL exponent alpha and the spectral norm.
+Doing this makes it clear what while the alpha part behaves as HT-SR theory would predict, i.e., with better models the PL exponents get smaller, the norms behave anomalously and exhibit scale collapes.
+The bottom line is that the metrics we discuss can be sued to distinguish between well-trained versus very well trained models, but that one has to look at the PL exponents to diagnose gross model problems, which was one of our original points.
+We clarify these points in revised text between Line XXX and Line XXX of the revised manuscript.
 \end{quote}
@@
 \begin{quote}
-(((XXX.  CM:
-We can analyze in more detail sure.  Space provided.
-)))\\
-The reviewer is correct that the plots are similar.
-The important point is how they are different.
-For the GPT $\alpha$ plots, the overall distribution shifts slightly to smaller values of $\alpha$ and there are fewer large outliers, both indicating that the larger model is
-XXX MM TO REFINE.
+%(((XXX.  CM:
+%We can analyze in more detail sure.  Space provided.
+%)))\\
+The details matter a lot in both of these figures.
+The reviewer is correct that the plots are similar---the important point is how they are different.
+In Fig 6a, the alphas behave as they should and as as would be predicted by HT-SR theory, i.e., the better GPT2 has the bulk of the alphas shifted to the left and all of the large outlying alphas removed.
+In Fig 6b, the spectral norm fails to perform well; while it is true that the higher-quality GPT2 does not exhibit Scale Collapes, the distribution of norms is shifted slightly to the right.
+In Fig 8a, there is a slight shift of alphas to the left as the models get larger (with average values now given in the text at Line XXX).
+Interesting the larger models still have some larger alpha values, and we conjecture at Line XXX that the reason for this is that the models are not fully optimized.  
+Subsequent developments in NLP have shown this to be the case.
+In Fig 8b, which shows the trand more stronlgy, the Shatten norm shifts to the left with higher quality models, as we would expect.
+We have put this discussion and clarified these points in the text.
 \end{quote}
@@
 \begin{quote}
-XXX.  CM:
-We can add a supplementary github repo.
+%(((XXX.  CM:
+%We can add a supplementary github repo.
+%)))\\
+We have removed those statements and only refer to the main text or the Supplementary Material.
+We do point out, however, that we intend out results to be fully-reproducible by the community, and thus our publicly-available repo is designed to contain the enormous quantity of data underlying out analysis.
 \end{quote}
@@
 \begin{quote}
-XXX.  CM:
-We disagree.  We are happy to provide a more detailed experimental analysis and in gross detail.
-Space is limited here, however. 
+%(((XXX.  CM:
+%We disagree.  We are happy to provide a more detailed experimental analysis and in gross detail.
+%Space is limited here, however. 
+%)))\\
+Thanks for the comments.
+Space is limited in the main text, but hopefully we have addressed your main concerns.
 \end{quote}
@@
 \begin{quote}
-(((XXX.  CM:
-Easy.
-)))\\
+%(((XXX.  CM:
+%Easy.
+%)))\\
 We have elaborated on this at Line XXX of the revised manuscript.
@@
 \begin{quote}
-(((XXX.  CM:
-We can discuss in detail.  The trick is in how the tail is estimated ...
-)))\\
+%(((XXX.  CM:
+%We can discuss in detail.  The trick is in how the tail is estimated ...
+%)))\\
 More important than the smaller condition number, which has to do with the ratio of the largest to smallest eigenvalue, is the ``shape'' of the spectral distribution, and in particular the details of the tail and how the tail is estimated.
@@
 \begin{quote}
-(((XXX.  CM:
-Pftt IDK.  We can think something up.
-)))\\
+%(((XXX.  CM:
+%Pftt IDK.  We can think something up.
+%)))\\
 Certainly.
@@
 \begin{quote}
-(((XXX.  CM:
-Detrimental.  This is known in GANs, where regularizers are used to prevent rank collapse.
-I'll have to think of a bigger reason.
-)))\\
+%(((XXX.  CM:
+%Detrimental.  This is known in GANs, where regularizers are used to prevent rank collapse.
+%I'll have to think of a bigger reason.
+%)))\\
 In general, it would be detrimental.
@@
 \begin{quote}
-(((XXX.  CM:
-This just means that the scale of the W matrix changes suddenly and unexpectedly.
-)))\\
-XXX MM QUES: ``suddenly and unexpectedly'' MEANS AS A FUNCTION OF SOMETHING THAT IS CHANGING.
+%(((XXX.  CM:
+%This just means that the scale of the W matrix changes suddenly and unexpectedly.
+%)))\\
+%XXX MM QUES: ``suddenly and unexpectedly'' MEANS AS A FUNCTION OF SOMETHING THAT IS CHANGING.
+By Scale Collapes, we mean that the size scale or norm of one or more layers changes a lot, while the size scale or norm of the remaining layers changes very little, as a function of some change to or perturbation of a model.
+Examples of such changes could include model compression or distillation, data augmentation, additional training, model fine-tuning, etc.
+A more precise definition would have to do with the stability of the model under those sorts of perturbations, but we think that including such a precise definition in the main text would only confuse the reader.
+We clarify this point at Line XXX of the revised manuscript.
 \end{quote}
@@
 \begin{quote}
-(((XXX.  CM:
-Yes and No.  Recall we are comparing networks with similar architectures.
-One then has to choose a canonical way to normalize the network so that similar architectures can be compared.
-While different networks do have different normalizations, we have not found any discrepancies across 
-models in the same architecture series.
-)))\\
+%(((XXX.  CM:
+%Yes and No.  Recall we are comparing networks with similar architectures.
+%One then has to choose a canonical way to normalize the network so that similar architectures can be compared.
+%While different networks do have different normalizations, we have not found any discrepancies across 
+%models in the same architecture series.
+%)))\\
 Yes and No.  
+In theory, one could do that; but in practice, for state-of-the-art models, one does not do that.
 Recall we are comparing networks with similar architectures.
-One then has to choose a canonical way to normalize the network so that similar architectures can be compared.
+One then has to choose (explicitly, or more commonly implicitly, by design choices in the training of the model) a canonical way to normalize the network so that similar architectures can be compared.
 While different networks do have different normalizations, we have not found any discrepancies across 
@@
 \begin{quote}
-(((XXX.  CM:
-I'll let you handle this one.
-)))\\
-XXX.  MM TO DO.
+%(((XXX.  CM:
+%I'll let you handle this one.
+%)))\\
+Superficially, work on double descent may be interpreted as ``more parameters are better,'' but in practice the situation is not that simple, e.g., since it can be difficult to define an appropriate effective number of parameters in realistic architectures and since realistic models are not fully optimized.
+We comment on and clarify this point at Line XXX of the revised manuscript.
 \end{quote}
@@
 \begin{quote}
-(((XXX.  CM:
-That's fair.
-)))\\
+%(((XXX.  CM:
+%That's fair.
+%)))\\
 That is a good point.

### 2020-12-30 submission/NatCom_abstract_file.tex
update roughly pages one to five

@@
 We examine norm based capacity control metrics as well as power law based metrics from the recently-developed Theory of Heavy-Tailed Self Regularization.
-We find that norm based metrics correlate well with reported test accuracies for well-trained models, but that they often can not distinguish well-trained versus poorly-trained models.
-We also find that power law based metrics do much better---qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among series of well-trained models.
-These methods can be used to identify when a pretrained neural network has problems that can not be detected simply by examining training/test accuracies.
+We find that norm based metrics correlate well with reported test accuracies for well-trained models, but that they often cannot distinguish well-trained versus poorly-trained models.
+We also find that power law based metrics can do much better---quantitatively better at discriminating among series of well-trained models with a given architecture; and qualitatively better at discriminating well-trained versus poorly-trained models.
+These methods can be used to identify when a pretrained neural network has problems that cannot be detected simply by examining training/test accuracies.

### 2020-12-30 submission/NatCom_cv_models.tex
update roughly pages one to five

@@
 
-Each of the VGG, ResNet, and DenseNet series of models consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each is distributed with the current open source pyTorch framework (version 1.4)~\cite{pytorch}.
+Each of the VGG, ResNet, and DenseNet series of models consists of several pretrained DNN models, with a given base architecture, trained on the full ImageNet~\cite{imagenet} dataset, and each is distributed with the current open source pyTorch framework (version 1.4)~\cite{pytorch}.
 In addition, we examine a larger set of ResNet models, which we call the ResNet-1K series, trained on the ImageNet-1K dataset~\cite{imagenet} and provided on the OSMR Sandbox~\cite{osmr}.
@@
     \caption{Comparison of Average $\alpha$-Norm quality metric versus reported Top1 test accuracy for the ResNet and ResNet-1K pretrained (pyTorch) models. 
-             \michael{Symbols and text are way to small on these and other figures; axes/ticks way too small; need different symbols, not just red versus green; etc.}
+             \michael{FROM V1: Symbols and text are way to small on these and other figures; axes/ticks way too small; need different symbols, not just red versus green; etc.}
             }
@@
 See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics for the VGG, ResNet, ResNet-1K, and DenseNet series.
-Similar results (not shown) are obtained for the Top5 accuracies.
+Similar results are obtained for the Top5 accuracies.
 Overall, all metrics perform relatively well; the Weighted Alpha metric performs best; and the Log $\alpha$-Norm metric performs second best.

### 2020-12-30 submission/NatCom_introduction.tex
update roughly pages one to five

@@
 Moreover, if ML is to become an industrial process, then that process will become compartmentalized in order to scale: some groups will gather data, other groups will develop models, and other groups will use those models.
-Users of models can not be expected to know the precise details of how models were built, the specifics of data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized,~etc.
+Users of models cannot be expected to know the precise details of how models were built, the specifics of data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized,~etc.
 
@@
 For example, they may exhibit what we call Scale Collapse for these models. 
-Third, PL-based metrics do much better at predicting quality trends in pretrained CV/NLP models.  
-They are qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among a series of well-trained versus very-well-trained models.
+Third, PL-based metrics can do much better at predicting quality trends in pretrained CV/NLP models.  
+In particular, 
+a weighted PL exponent (weighted by the log of the spectral norm of the corresponding layer) is 
+quantitatively better at discriminating among a series of well-trained versus very-well-trained models
+within a given architecture series; and
+the (unweighted) average PL exponent is 
+qualitatively better at discriminating well-trained versus poorly-trained models.
 Fourth, PL-based metrics can also be used to characterize fine-scale model properties, including what we call layer-wise Correlation Flow, in well-trained and poorly-trained models; and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).

### 2020-12-30 submission/NatCom_overall_approach.tex
update roughly pages one to five

@@
 \end{equation}
+\michael{how precisely to change eqn for R3 comment}
 where the loss function $\mathcal{L}(\cdot)$ can take on a myriad of forms~\cite{JC17_TR}, and where the energy (or optimization) landscape function
@@
 \end{equation}
+\michael{sentences around here probably need to be changed in light of change to that eqn R3 wanted}
 depends parametrically on the weights and biases.
-\michael{See comments for math equations.}
+\michael{FROM V1: See comments for math equations.}
 For a trained model, $E_{DNN}$ does not explicitly depend on the data, but rather it maps data instance vectors ($\mathbf{x}_i$ values) to predictions ($y_{i}$ labels).
-\michael{Mark found that sentence confusing.}
+\michael{FROM V1: Mark found that sentence confusing.}
 Therefore, one can analyze $E_{DNN}$ in the absence of any training or test~data.
@@
 $\mathbf{W}$, and/or, equivalently, it's empirical correlation matrix, $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
-We consider the following metrics:
+To evaluate the quality of state-of-the-art DNNs, 
+we consider the following metrics:
 \begin{eqnarray}
@@
 \end{eqnarray}
-Here, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the $\mathbf{X}$, and $\lambda_{max}$ is the maximum eigenvalue.
+%To diagnose problems, 
+To perform diagnostics on potentially-problematic DNNs,
+we will decompose $\hat{\alpha}$ into its two components, $\alpha$ and $\lambda_{max}$.
+Here, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the $\mathbf{X}$, $\lambda_{max}$ is the maximum eigenvalue, and $\alpha$ is the fitted PL exponent. 
 These eigenvalues are squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, $\lambda_{i}=\sigma^{2}_{i}$.
@@
 Crucially, HT-SR Theory predicts that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models.
-Relatedly, previous work observed that smaller exponents $\alpha$ correspond to more implicit self-regularization and better generalization~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
+The notion of ``size scale'' is well-defined in physical systems, to which this style of analysis is usually applied, but it is less well-defined in CV and NLP applications.
+Informally, it would correspond to pixel groups that are at a greater distance in some metric, or between sentence parts that are at a greater distance in text.
+Relatedly, previous work observed that smaller exponents $\alpha$ correspond to more implicit self-regularization and better generalization, and that we expect a linear correlation between $\hat{\alpha}$ and model quality~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
 
@@
 \end{equation}
-For the $\alpha$-Shatten Norm metric, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it can not be taken out of the sum.
+For the $\alpha$-Shatten Norm metric, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it cannot be taken out of the sum.
 For small $\alpha$, the Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}; and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm.
+
+Finally, although it does less well for predicting trends in state-of-the-art model series, e.g., as depth changes, the average value of $\alpha$, i.e., 
+\begin{equation}
+\bar{\alpha} = \dfrac{1}{L}\sum_l \alpha_l = \langle\alpha\rangle    ,
+\end{equation}
+can be used to perform model diagnostics, to identify problems that cannot be detected by examining training/test accuracies, and to discriminate poorly-trained models from well-trained~models.
 
@@
              Given an individual layer weight matrix $\mathbf{W}$, from either a fully-connected layer or a convolutional layer, perform a Singular Value Decomposition (SVD) to obtain $\mathbf{W} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^{T}$, and examine the histogram of eigenvalues of $\mathbf{W}^{T}\mathbf{W}$.
-             Norm-based metrics (not shown) and PL-based metrics (that depend on fitting the histogram of eigenvalues to a truncated PL) can be used to compare models.
+             Norm-based metrics and PL-based metrics (that depend on fitting the histogram of eigenvalues to a truncated PL) can be used to compare models.
              For example, one can analyze one layer of a pre-trained model, compare multiple layers of a pre-trained model, make comparisons across model architectures, monitor neural network properties during training, etc.

### 2020-12-30 submission/empirical2_NatComfmt.tex
more on address presubmission comments
minor changes before submit to Nat Comm
update roughly pages one to five

@@
 
+\usepackage{lineno}
+\linenumbers
+
 %-----------------------------------------------------------------------
@@
 \section{Results}
-After describing our overall approach, we study in detail three well-known CV architectures (the VGG, ResNet, and DenseNet series of models).
-Then, we look in detail at several variations of a popular NLP architecture (the OpenAI GPT and GPT2 models), and we present results from a broader analysis of hundreds of pretrained DNN~models.
+After describing our overall approach, we study in detail three well-known CV architecture series (the VGG, ResNet, and DenseNet series of models).
+Then, we look in detail at several variations of a popular NLP architecture series (the OpenAI GPT and GPT2 series of models), and we present results from a broader analysis of hundreds of pretrained DNN~models.
 

### 2020-12-30 submission/NatCom_cv_models.tex
first part of cv section, including placeholder for extended table one

@@
 All four metrics correlate quite well with reported Top1 accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalization (i.e., greater accuracy, lower error). 
-The Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) performs best (RMSE of $0.42$, see Table~\ref{table:cv-models}); and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $) performs second best (RMSE of $0.48$, see Table~\ref{table:cv-models}).
+\michael{The following, text and numbers, will need to be updated once we finalize the figures/tables.}
+The Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) performs best (RMSE of $0.42$, $R^2$ of $XXX$, and Kendall-$\tau$ of $XXX$, see Table~\ref{table:cv-models}); and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $) performs second best (RMSE of $0.48$, $R^2$ of $XXX$, and Kendall-$\tau$ of $XXX$, see Table~\ref{table:cv-models}).
+\michael{Put comment on why these three metrics, that HT-SR says that $\hat{\alpha}$ should exhibit good linear correlation, and a few comments on non-normality to address R1 concerns.}
 
@@
 
+%V1VERSION% \begin{table}[t]
+%V1VERSION% \small
+%V1VERSION% \begin{center}
+%V1VERSION% \begin{tabular}{|p{1in}|c|c|c|c|c|}
+%V1VERSION% \hline
+%V1VERSION%  Series    &\#   & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\hat{\alpha}$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
+%V1VERSION% \hline
+%V1VERSION%  VGG       &  6 & 0.56 & 0.53 & 0.48          & \textbf{0.42}  \\
+%V1VERSION%  ResNet    &  5 & 0.9  & 1.4  & \textbf{0.61} & 0.66           \\
+%V1VERSION%  ResNet-1K & 19 & 2.4  & 3.6  & \textbf{1.8}  & 1.9            \\
+%V1VERSION%  DenseNet  &  4 & 0.3  & 0.26 & \textbf{0.16} & 0.21           \\
+%V1VERSION% \hline
+%V1VERSION% \end{tabular}
+%V1VERSION% \end{center}
+%V1VERSION% \caption{RMSE (smaller is better) for linear fits of 
+%V1VERSION%          quality metrics to reported Top1 test error for pretrained models in each architecture series.  
+%V1VERSION%          Column \# refers to number of models.  
+%V1VERSION%          VGG, ResNet, and DenseNet were pretrained on ImageNet.  
+%V1VERSION%          ResNet-1K was pretrained on ImageNet-1K. 
+%V1VERSION% }
+%V1VERSION% \label{table:cv-models}
+%V1VERSION% \end{table}
+
 \begin{table}[t]
@@
 \begin{center}
-\begin{tabular}{|p{1in}|c|c|c|c|c|}
+\begin{tabular}{|p{1in}|c|c|c|c|c|c|}
 \hline
- Series    &\#   & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\hat{\alpha}$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
+ Series                      & \#                    & Metric         & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\hat{\alpha}$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
 \hline
- VGG       &  6 & 0.56 & 0.53 & 0.48          & \textbf{0.42}  \\
- ResNet    &  5 & 0.9  & 1.4  & \textbf{0.61} & 0.66           \\
- ResNet-1K & 19 & 2.4  & 3.6  & \textbf{1.8}  & 1.9            \\
- DenseNet  &  4 & 0.3  & 0.26 & \textbf{0.16} & 0.21           \\
+\multirow{3}{4em}{VGG}       & \multirow{3}{1em}{6}  & RMSE           & 0.56 & 0.53 & 0.48          & \textbf{0.42}  \\
+                             &                       & $R^2$          &  XXX &  XXX &  XXX          &  XXX           \\
+                             &                       & Kendall-$\tau$ &  XXX &  XXX &  XXX          &  XXX           \\
+\hline
+\multirow{3}{4em}{ResNet}    & \multirow{3}{1em}{5}  & 0.9            & 1.4  & \textbf{0.61} & 0.66           \\
+                             &                       & $R^2$          &  XXX &  XXX &  XXX          &  XXX           \\
+                             &                       & Kendall-$\tau$ &  XXX &  XXX &  XXX          &  XXX           \\
+\hline
+\multirow{3}{4em}{ResNet-1K} & \multirow{3}{1em}{19} & 2.4            & 3.6  & \textbf{1.8}  & 1.9            \\
+                             &                       & $R^2$          &  XXX &  XXX &  XXX          &  XXX           \\
+                             &                       & Kendall-$\tau$ &  XXX &  XXX &  XXX          &  XXX           \\
+\hline
+\multirow{3}{4em}{DenseNet}  & \multirow{3}{1em}{4}  & 0.3            & 0.26 & \textbf{0.16} & 0.21           \\
+                             &                       & $R^2$          &  XXX &  XXX &  XXX          &  XXX           \\
+                             &                       & Kendall-$\tau$ &  XXX &  XXX &  XXX          &  XXX           \\
 \hline
@@
 \end{center}
-\caption{RMSE (smaller is better) for linear fits of 
-         quality metrics to reported Top1 test error for pretrained models in each architecture series.  
+\caption{Quality metrics 
+         (RMSE, smaller is better; $R^2$, XXX is better; and Kendall-$\tau$ rank correlation, XXX is better)
+         for reported Top1 test error for pretrained models in each architecture series.  
          Column \# refers to number of models.  
@@
 
+\michael{The following, text and numbers, will need to be updated once we finalize the figures/tables.}
+Overall, all metrics perform relatively well; the Weighted Alpha metric performs best; and the Log $\alpha$-Norm metric performs second best.
 See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics for the VGG, ResNet, ResNet-1K, and DenseNet series.
 Similar results are obtained for the Top5 accuracies.
-Overall, all metrics perform relatively well; the Weighted Alpha metric performs best; and the Log $\alpha$-Norm metric performs second best.
 These model series are all very well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of well-trained versus very-well-trained models, with PL-based metrics performing somewhat (i.e., quantitatively) better.
@@
 Figure~\ref{fig:cv2-accuracy} compares the Log $\alpha$-Norm metric for the full ResNet model, trained on the full ImageNet dataset, against the ResNet-1K model, trained on a much smaller ImageNet-1K data set.
+\michael{The following, text and numbers, will need to be updated once we finalize the figures/tables.}
 The Log $\alpha$-Norm is much better than the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is slightly worse than the Weighted Alpha metric).
-The ResNet series has strong correlation, with an RMSE of $0.66$, whereas the ResNet-1K series also shows good correlation, but has a much larger RMSE of $1.9$.
+The ResNet series has strong correlation (RMSE of $0.66$, $R^2$ of $XXX$, and Kendall-$\tau$ of $XXX$), whereas the ResNet-1K series also shows good but weaker correlation (much larger RMSE of $1.9$, $R^2$ of $XXX$, and Kendall-$\tau$ of $XXX$).
 (Other metrics exhibit similar behavior.)
@@
 
-We can learn much more about a pretrained model by going beyond average values of quality metrics to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).
+We can learn much more about a pretrained model by using our metrics to perform model diagnostics.
+This involves going beyond average values of quality metrics to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).  
+As an example,
 Figure~\ref{fig:3models-alpha-layers} plots the PL exponent $\alpha$, as a function of depth, for each layer (first layer corresponds to data, last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (no BN), ResNet, and DenseNet series.
@@
 
+Overall, Figure~\ref{fig:3models-alpha-layers} demonstrates that the distribution of $\alpha$ values among layers is architecture dependent, 
+and that it can vary in a systematic way within an architecture series.
+This is to be expected, since some architectures enable better extraction of signal from the data.
+This also suggests that, while performing very well at predicting trends within an architecture series, PL-based metrics (as well as norm-based metrics) should be used with caution when comparing models with very different architectures.
+
 \paragraph{Correlation Flow; or How $\alpha$ Varies Across Layers.}
@@
 Figure~\ref{fig:3models-alpha-layers} can be understood in terms of what we will call Correlation Flow.
-The average Log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
+Recall that the average Log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}.
 There, one expects that the weight matrices of well-trained DNNs will exhibit correlations over many size scales, as is well-known in other strongly-correlated systems~\cite{BouchaudPotters03, SornetteBook}.

### 2020-12-30 submission/NatCom_cv_models.tex
fix typo in table

@@
 \hline
-\multirow{3}{4em}{ResNet}    & \multirow{3}{1em}{5}  & 0.9            & 1.4  & \textbf{0.61} & 0.66           \\
+\multirow{3}{4em}{ResNet}    & \multirow{3}{1em}{5}  & RMSE           &  0.9 & 1.4  & \textbf{0.61} & 0.66           \\
                              &                       & $R^2$          &  XXX &  XXX &  XXX          &  XXX           \\
@@
 \hline
-\multirow{3}{4em}{ResNet-1K} & \multirow{3}{1em}{19} & 2.4            & 3.6  & \textbf{1.8}  & 1.9            \\
+\multirow{3}{4em}{ResNet-1K} & \multirow{3}{1em}{19} & RMSE           &  2.4 & 3.6  & \textbf{1.8}  & 1.9            \\
                              &                       & $R^2$          &  XXX &  XXX &  XXX          &  XXX           \\
@@
 \hline
-\multirow{3}{4em}{DenseNet}  & \multirow{3}{1em}{4}  & 0.3            & 0.26 & \textbf{0.16} & 0.21           \\
+\multirow{3}{4em}{DenseNet}  & \multirow{3}{1em}{4}  & RMSE           &  0.3 & 0.26 & \textbf{0.16} & 0.21           \\
                              &                       & $R^2$          &  XXX &  XXX &  XXX          &  XXX           \\

### 2020-12-30 submission/NatCom_cv_models.tex
finished cv section

@@
 
+Going beyond coarse averages to examining quality metrics for each layer weight matrix as a function of depth (or layer id), our metrics can be used to perform model diagnostics and to identify fine-scale properties in a pretrained model.
+We provide three examples of this.
+
 \paragraph{Layer Analysis: Metrics as a Function of Depth.}
 
-We can learn much more about a pretrained model by using our metrics to perform model diagnostics.
-This involves going beyond average values of quality metrics to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).  
-As an example,
+%%We can learn much more about a pretrained model by using our metrics to perform model diagnostics.
+%%This involves going beyond average values of quality metrics to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).  
+%%As an example,
 Figure~\ref{fig:3models-alpha-layers} plots the PL exponent $\alpha$, as a function of depth, for each layer (first layer corresponds to data, last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (no BN), ResNet, and DenseNet series.
@@
 To illustrate this, we show that some compression/distillation methods \cite{CWZZ17_TR} may actually damage models unexpectedly, by introducing what we call Scale Collapse, where several distilled layers have unexpectedly small Spectral Norms.
+%
+By Scale Collapse, we mean that the size scale, e.g., as measured by the Spectral or Frobenius Norm, of one or more layers changes dramatically, while the size scale of other layers changes very little, as a function of some change to or perturbation of a model.
+The size scales of different parts of a DNN model are typically defined implicitly by the model training process, and they typically vary in a gradual way for high-quality models.
+Examples of changes of interest include model compression or distillation (discussed here for a CV model), data augmentation (discussed below for an NLP model), additional training, model fine-tuning, etc.
 
@@
 On the other hand, the $\alpha$ values (in Figure~\ref{fig:resnet204Dalpha}) do not differ systematically between the baseline and fine-tuned models.
-Also (not shown, but as predicted by HT-SR Theory, the basis of $\hat{\alpha}$), the average (unweighted) baseline $\alpha$ is smaller than the fine-tuned average.
+Also, $\bar{\alpha}$, the average unweighted baseline $\alpha$, is smaller for the original model than for the fine-tuned model 
+(as predicted by HT-SR Theory, the basis of $\hat{\alpha}$).
 %

### 2020-12-30 submission/NatCom_cv_models.tex
images updated

@@
     \subfigure[Log Frobenius Norm, VGG ]{
-        \includegraphics[width=4.9cm]{img/VGG_lognorm_accs.png}
+        \includegraphics[width=4.9cm]{img/VGG_log_norm_accs.png}
         \label{fig:vgg-fnorm}
@@
     \subfigure[Log Spectral Norm, VGG ]{
-        \includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
+        \includegraphics[width=4.9cm]{img/VGG_log_spectral_norm_accs.png}
         \label{fig:vgg-snorm}
@@
     \subfigure[Log $\alpha$-Norm, VGG ]{
-        \includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
+        \includegraphics[width=4.9cm]{img/VGG_log_alpha_norm_accs.png}
         \label{fig:vgg-pnorm}

### 2020-12-30 submission/NatCom_all_models.tex
putting placehodlers in appendix

@@
 
-We summarize results from a large-scale analysis of hundreds of publicly-available  models.
-This broader analysis on a much larger set of CV and NLP models, with a more diverse set of architectures, and developed for a wider range of tasks, complements the previous more detailed analysis on CV and NLP models.
+We have performed a large-scale analysis of hundreds of publicly-available models.
+This broader analysis is on a much larger set of CV and NLP models, with a more diverse set of architectures, that have been developed for a wider range of tasks; and it complements the previous more detailed analysis on CV and NLP models, where we have analyzed only a single architecture series at a time.
 See the Supplementary Information
 %~\ref{sxn:appendix} 
+(and our publicly-available repo)
 for details.

### 2020-12-30 submission/NatCom_appendix.tex
putting placehodlers in appendix

@@
 ESPNetv2                                   & 6 \\
-HRNet                      z                & 6 \\
+HRNet                                      & 6 \\
 SqueezeNet/SqueezeResNet                   & 6 \\
@@
     }
-    \caption{% 
-             %\charles{Preliminary charts:} 
-             PL exponent $\alpha$ versus reported Top1 Test Accuracies for pretrained DNNs available for five different data sets.
+    \caption{PL exponent $\alpha$ versus reported Top1 Test Accuracies for pretrained DNNs available for different datasets (as segmented in Table~\ref{table:datasets}).
             }
     \label{fig:DSalphas}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \subfigure[ ResNet ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_01}
+    }
+    \subfigure[ SENet/SE-ResNet/SE-PreResNet/SE-ResNeXt ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_02}
+    }
+    \subfigure[ DIA-ResNet/DIA-PreResNet ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_03}
+    }
+    \subfigure[ ResNeXt ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_04}
+    }
+    \subfigure[ WRN ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_05}
+    }
+    \subfigure[ DLA ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_06}
+    }
+    \subfigure[ PreResNet ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_07}
+    }
+    \subfigure[ ProxylessNAS ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_08}
+    }
+    \subfigure[ VGG/BN-VGG ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_09}
+    }
+    \subfigure[ IGCV3 ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_10}
+    }
+    \subfigure[ EfficientNet ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_11}
+    }
+    \subfigure[ SqueezeNext/SqNxt ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_12}
+    }
+    \subfigure[ ShuffleNet ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_13}
+    }
+    \subfigure[ DRN-C/DRN-D ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_14}
+    }
+    \subfigure[ ESPNetv2 ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_15}
+    }
+    \subfigure[ HRNet ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_16}
+    }
+    \subfigure[ SqueezeNet/SqueezeResNet ]{
+        \includegraphics[width=2.7cm]{img/hex.png}
+        \label{fig:ARCHalphas_17}
+    }
+    \caption{PL exponent $\alpha$ versus reported Top1 Test Accuracies for pretrained DNNs available for different architecture series (as segmented in Table~\ref{table:architectures}).
+            }
+    \label{fig:ARCHalphas}
 \end{figure}
@@
 \vspace{-5mm}
-\caption{$MSE$ Results for all CV model regressions. }
+\caption{$MSE$ results for our analysis of all CV models in Table~\ref{table:datasets}. }
 \label{table:MSEresults}
@@
 \vspace{-5mm}
-\caption{$R^{2}$ Results for all CV model regressions. }
+\caption{$R^{2}$ results for our analysis of all CV models in Table~\ref{table:datasets}. }
 \label{table:R2results}
@@
 
+\begin{table}[t]
+\scriptsize
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+Dataset & Model  & $\langle\log\Vert\cdot\Vert_{F}\rangle$ & $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ & $\hat{\alpha}$ & $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ \\
+\hline
+ imagenet-1k  & ResNet        &  XXX &   XXX &  XXX &  XXX \\
+ imagenet-1k  & EfficientNet  &  XXX &   XXX &  XXX &  XXX \\
+ imagenet-1k  & PreResNet     &  XXX &   XXX &  XXX &  XXX \\
+ imagenet-1k  & ShuffleNet    &  XXX &   XXX &  XXX &  XXX \\
+ imagenet-1k  & VGG           &  XXX &   XXX &  XXX &  XXX \\
+ imagenet-1k  & DLA           &  XXX &   XXX &  XXX &  XXX \\
+ imagenet-1k  & HRNet         &  XXX &   XXX &  XXX &  XXX \\
+ imagenet-1k  & DRN-C         &  XXX &   XXX &  XXX &  XXX \\
+ imagenet-1k  & SqueezeNext   &  XXX &   XXX &  XXX &  XXX \\
+ imagenet-1k  & ESPNetv2      &  XXX &   XXX &  XXX &  XXX \\
+ imagenet-1k  & IGCV3         &  XXX &   XXX &  XXX &  XXX \\
+ imagenet-1k  & SqueezeNet    &  XXX &   XXX &  XXX &  XXX \\
+ imagenet-1k  & ProxylessNAS  &  XXX &   XXX &  XXX &  XXX \\
+\hline
+ cifar-10     & ResNet        &  XXX &   XXX &  XXX &  XXX \\
+ cifar-10     & DIA-ResNet    &  XXX &   XXX &  XXX &  XXX \\
+ cifar-10     & SENet         &  XXX &   XXX &  XXX &  XXX \\
+\hline
+ cifar-100    & ResNet        &  XXX &   XXX &  XXX &  XXX \\
+ cifar-100    & DIA-ResNet    &  XXX &   XXX &  XXX &  XXX \\
+ cifar-100    & SENet         &  XXX &   XXX &  XXX &  XXX \\
+ cifar-100    & WRN           &  XXX &   XXX &  XXX &  XXX \\
+\hline
+ svhn         & ResNet        &  XXX &   XXX &  XXX &  XXX \\
+ svhn         & DIA-ResNet    &  XXX &   XXX &  XXX &  XXX \\
+ svhn         & SENet         &  XXX &   XXX &  XXX &  XXX \\
+ svhn         & WRN           &  XXX &   XXX &  XXX &  XXX \\
+ svhn         & ResNeXt       &  XXX &   XXX &  XXX &  XXX \\
+\hline
+ cub-200-2011 & ResNet        &  XXX &   XXX &  XXX &  XXX \\
+ cub-200-2011 & SENet         &  XXX &   XXX &  XXX &  XXX \\
+\hline
+\end{tabular}
+\end{center}
+\vspace{-5mm}
+\caption{Kendall-$\tau$ results for our analysis of all CV models in Table~\ref{table:datasets}. }
+\label{table:KTresults}
+\end{table}
+
 \subsection{Supplementary Figures}
+
+\michael{The text in this section needs to be changed---probably not removed, but instead discuss things, not point to the figure we remove, and point to the repo. }
+
+\michael{Do we get more visually compelling results if we present figures, but segmented by architectures, as listed in Table~\ref{table:architectures}; if so, perhaps we should put those, so we don't tell R1 that we just removed what looked bad; or maybe actually both, since we segment by dataset in Table~\ref{table:datasets}.
+}
 
@@
  still see that most plots show positive relationship with positive coefficients. 
-Example regressions are shown in Figure~\ref{fig:DSalphas}, as available in the results notebook.
+% V1 REMOVED FOR V2 % Example regressions are shown in Figure~\ref{fig:DSalphas}, as available in the results notebook.
+Example regressions are shown in 
+Figure~\ref{fig:DSalphas}, for models segmented by dataset, as listed in Table~\ref{table:datasets}, 
+and in 
+Figure~\ref{fig:ARCHalphas}, for models segmented by architecture, as listed in Table~\ref{table:architectures}.
+\michael{Comment here on differences.}
 
@@
 We record the $R^{2}$ and $MSE$ for each metric, averaged over all regressions for all architectures and datasets.
-See Table~\ref{table:MSEresults} and Table~\ref{table:R2results}.
+See Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
 In the repo, plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.

### 2020-12-30 submission/NatCom_cv_models.tex
images updated
getting appendix in place

@@
     \subfigure[ ResNet-1K, Log $\alpha$-Norm ]{
-        \includegraphics[width=4.9cm]{img/ResNet-1K_logpnorm_accs.png}
+        \includegraphics[width=4.9cm]{img/ResNet-1K_log_alpha_norm_accs.png}
         \label{fig:resnet1k-accuracy}
@@
 This leads to the hypothesis that small $\alpha$ values that are stable across multiple layers enable better correlation flow through the network.
-This is similar to recent work on the information bottleneck~\cite{ST17_TR}, except that here we work in an entirely unsupervised setting.
+This is similar to recent work on the information bottleneck~\cite{TZ15,ST17_TR}, except that here we work in an entirely unsupervised setting.
 

### 2020-12-30 submission/NatCom_overall_approach.tex
getting appendix in place

@@
 For the $\alpha$-Shatten Norm metric, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it cannot be taken out of the sum.
-For small $\alpha$, the Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}; and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm.
+For small $\alpha$, the Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation;
+%% \cite{MM20_unpub_work}; 
+and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm.
 

### 2020-12-31 submission/NatCom_all_models.tex
getting appendix in place
pass on everything except nlp subsection

@@
 \label{sxn:all_cv_models}
+
+\michael{Refine text in this subsection after we get the final version of all the tables and figures in the appendix.
+Things to emphasize: we have a huge amount of data that we are summarizing very succicently; the table here is the tip of the iceberg; the results show our metrics are good, as a sanity check; they show that you can get some results in general, but looking at the details in the appendix show that results are more reliable within a given architecture series than across different architectures; point to out publicly-availab e repo; mention that we do Kendall-$\tau$ rank correlation also.}
 
@@
 \begin{center}
-\begin{tabular}{|p{1in}|c|c|c|c|}
+\begin{tabular}{|p{1.25in}|c|c|c|c|}
 \hline
@@
 %    & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\langle\hat{\alpha}=\alpha\log\lambda_{max}\rangle$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
-Series        & $\log\Vert\cdot\Vert_{F}$ & $\log\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
+                        & $\log\Vert\cdot\Vert_{F}$ & $\log\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
 \hline
@@
 $MSE$ (std)             & 8.69 & 23.06 &          5.14 & \textbf{5.00} \\
+\hline
+Kendall-$\tau$ (mean)   &  XXX &   XXX &           XXX &           XXX \\
+Kendall-$\tau$ (std)    &  XXX &   XXX &           XXX &           XXX \\
 \hline

### 2020-12-31 submission/NatCom_discussion.tex
getting appendix in place
pass on everything except nlp subsection

@@
 
-Our work, starting in 2018 with the \texttt{WeightWatcher} tool~\cite{weightwatcher_package}, is the first to perform a detailed analysis of the weight matrices of DNNs~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
+Our work also has intriguing similarities and differences with work on understanding DNNs with the information bottleneck principle~\cite{TZ15,ST17_TR}, which posits that DNNs can be quantified by the mutual information between their layers and the input and output variables.
+Most importantly, our approach does not require access to any data, while information measures used in the information bottleneck approach do require this.
+Nevertheless, several results from HT-SR Theory, on which our metrics are based, have parallels in the information bottleneck approach.
+Perhaps most notably, the quick transition from a \textsc{Random-like} phase to \textsc{Bulk+Spikes} phase, followed by slow transition to a \textsc{Heavy-Tailed} phase, as noted previously~\cite{MM18_TR}, is reminiscent of the dynamics on the Information Plane~\cite{ST17_TR}.
+
+Finally, 
+our work, starting in 2018 with the \texttt{WeightWatcher} tool~\cite{weightwatcher_package}, is the first to perform a detailed analysis of the weight matrices of DNNs~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
 Subsequent to the initial version of this paper, we became aware of two other works that were posted to the in 2020 within weeks of the initial version of this paper~\cite{EJRUY20_TR,UKGBT20_TR}.

### 2020-12-31 submission/NatCom_appendix.tex
getting appendix in place
tables updated

@@
 
-\subsection{Supplementary Tables}
+%%\subsection{Supplementary Tables} 
+\subsection{Supplementary Details}
 
@@
 The final architecture series used are shown in  Table~\ref{table:architectures}, with the number of models in each.
+
+Tables and figures summarizing this analysis (in a more fine-grained way than provided by Table~\ref{table:results}) are presented next.
 
@@
 \hline
- imagenet-1k  & ResNet        &  XXX &   XXX &  XXX &  XXX \\
- imagenet-1k  & EfficientNet  &  XXX &   XXX &  XXX &  XXX \\
- imagenet-1k  & PreResNet     &  XXX &   XXX &  XXX &  XXX \\
- imagenet-1k  & ShuffleNet    &  XXX &   XXX &  XXX &  XXX \\
- imagenet-1k  & VGG           &  XXX &   XXX &  XXX &  XXX \\
- imagenet-1k  & DLA           &  XXX &   XXX &  XXX &  XXX \\
- imagenet-1k  & HRNet         &  XXX &   XXX &  XXX &  XXX \\
- imagenet-1k  & DRN-C         &  XXX &   XXX &  XXX &  XXX \\
- imagenet-1k  & SqueezeNext   &  XXX &   XXX &  XXX &  XXX \\
- imagenet-1k  & ESPNetv2      &  XXX &   XXX &  XXX &  XXX \\
- imagenet-1k  & IGCV3         &  XXX &   XXX &  XXX &  XXX \\
- imagenet-1k  & SqueezeNet    &  XXX &   XXX &  XXX &  XXX \\
- imagenet-1k  & ProxylessNAS  &  XXX &   XXX &  XXX &  XXX \\
+imagenet-1k & ResNet  & 0.77 &  0.74 & 0.88 & 0.88 \\
+ imagenet-1k & EfficientNet  & 0.67 &  0.79 & 0.66 & 0.66 \\
+ imagenet-1k & PreResNet  & 0.65 &  0.54 & 0.87 & 0.87 \\
+ imagenet-1k & ShuffleNet  & 0.39 &  0.09 & 0.85 & 0.82 \\
+ imagenet-1k & VGG  & 0.82 &  0.76 & 0.52 & 0.61 \\
+ imagenet-1k & DLA  & 0.51 &  0.82 & 0.78 & 0.69 \\
+ imagenet-1k & HRNet  & 0.56 &  0.44 & 0.61 & 0.61 \\
+ imagenet-1k & DRN-C  & 0.90 &  0.81 & 0.90 & 0.90 \\
+ imagenet-1k & SqueezeNext  & 0.47 &  -0.33 & 0.73 & 0.73 \\
+ imagenet-1k & ESPNetv2  & 0.00 &  0.80 & 1.00 & 1.00 \\
+ imagenet-1k & SqueezeNet  & 0.33 &  0.00 & 0.67 & 0.67 \\
+ imagenet-1k & ProxylessNAS  & 0.33 &  0.33 & 0.33 & 0.33 \\
+ imagenet-1k & IGCV3  & 1.00 &  0.00 & 1.00 & 1.00 \\
 \hline
- cifar-10     & ResNet        &  XXX &   XXX &  XXX &  XXX \\
- cifar-10     & DIA-ResNet    &  XXX &   XXX &  XXX &  XXX \\
- cifar-10     & SENet         &  XXX &   XXX &  XXX &  XXX \\
+ cifar-10 & ResNet  & 0.64 &  0.64 & 0.71 & 0.71 \\
+ cifar-10 & DIA-ResNet  & 0.79 &  0.50 & 0.86 & 0.79 \\
+ cifar-10 & SENet  & 0.74 &  0.95 & 0.95 & 0.95 \\
 \hline
- cifar-100    & ResNet        &  XXX &   XXX &  XXX &  XXX \\
- cifar-100    & DIA-ResNet    &  XXX &   XXX &  XXX &  XXX \\
- cifar-100    & SENet         &  XXX &   XXX &  XXX &  XXX \\
- cifar-100    & WRN           &  XXX &   XXX &  XXX &  XXX \\
+ cifar-100 & ResNet  & 0.64 &  0.57 & 0.64 & 0.64 \\
+ cifar-100 & DIA-ResNet  & 0.93 &  0.43 & 0.93 & 0.93 \\
+ cifar-100 & SENet  & 1.00 &  1.00 & 1.00 & 1.00 \\
+ cifar-100 & WRN  & 0.67 &  -0.67 & 0.00 & 0.00 \\
 \hline
- svhn         & ResNet        &  XXX &   XXX &  XXX &  XXX \\
- svhn         & DIA-ResNet    &  XXX &   XXX &  XXX &  XXX \\
- svhn         & SENet         &  XXX &   XXX &  XXX &  XXX \\
- svhn         & WRN           &  XXX &   XXX &  XXX &  XXX \\
- svhn         & ResNeXt       &  XXX &   XXX &  XXX &  XXX \\
+ svhn & ResNet  & 0.81 &  0.71 & 0.81 & 0.81 \\
+ svhn & DIA-ResNet  & 0.86 &  0.86 & 0.57 & 0.57 \\
+ svhn & SENet  & 1.00 &  1.00 & 0.67 & 0.67 \\
+ svhn & WRN  & -0.33 &  -0.33 & 0.67 & 0.67 \\
+ svhn & ResNeXt  & 0.67 &  0.67 & 0.33 & 0.33 \\
 \hline
- cub-200-2011 & ResNet        &  XXX &   XXX &  XXX &  XXX \\
- cub-200-2011 & SENet         &  XXX &   XXX &  XXX &  XXX \\
+ cub-200-2011 & ResNet  & 1.00 &  1.00 & -0.33 & -0.33 \\
+ cub-200-2011 & SENet  & 0.87 &  0.87 & -0.20 & -0.20 \\
 \hline
@@
 
-\subsection{Supplementary Figures}
+%%\subsection{Supplementary Figures}
+\subsection{Supplementary Tables and Figures}
 
@@
 
+\michael{Probably some discussion of why these metrics are the right ones, e.g., as opposed to $p$ values, to address R1 comments.}
+
 \subsection{Supplementary Discussion: Additional Details on HT-SR Theory}

### 2020-12-31 submission/NatCom_cv_models.tex
figs updated
minor text added to CV

@@
 \michael{Put comment on why these three metrics, that HT-SR says that $\hat{\alpha}$ should exhibit good linear correlation, and a few comments on non-normality to address R1 concerns.}
+\charles{The Log Spectral Norm metric correlates very well with small, simpler data sets like the VGG and DenseNet architectures, however, when moving to the larger and more complex ResNet series, the PL-based metrics Log $\alpha$-Norm ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$)
+and  Weighted Alpha ($\hat\alpha =\alpha\log\lambda_{max} $) metrics perform better across all measures.}
 
@@
              trained on ImageNet, available in pyTorch (v1.4).  
-             Metrics fit by linear regression, RMSE reported.  }
+             Metrics fit by linear regression, RMSE, R2, and the Kendal-tau rank correlation metric reported.  }
     \label{fig:vgg-metrics}
@@
     \subfigure[ ResNet, Log $\alpha$-Norm ]{
-        \includegraphics[width=4.9cm]{img/ResNet_logpnorm_accs.png}
+        \includegraphics[width=4.9cm]{img/ResNet_log_alpha_norm_accs.png}
         \label{fig:resnet-accuracy}
@@
              \michael{FROM V1: Symbols and text are way to small on these and other figures; axes/ticks way too small; need different symbols, not just red versus green; etc.}
+\charles{See changes}
             }
@@
 \hline
- Series                      & \#                    & Metric         & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\hat{\alpha}$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
+ Series                      & \#                    & Metric         & $\langle\log\Vert\mathbf{W}\Vert^{2}_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$ & $\hat{\alpha}$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
 \hline
-\multirow{3}{4em}{VGG}       & \multirow{3}{1em}{6}  & RMSE           & 0.56 & 0.53 & 0.48          & \textbf{0.42}  \\
-                             &                       & $R^2$          &  XXX &  XXX &  XXX          &  XXX           \\
-                             &                       & Kendall-$\tau$ &  XXX &  XXX &  XXX          &  XXX           \\
+\multirow{3}{4em}{VGG}       & \multirow{3}{1em}{6}  & RMSE           & 0.56 & \textbf{0.23} & 0.48          & 0.34  \\
+                             &                       & $R^2$          & 0.88 & \textbf{0.98} &  0.92         &  0.96           \\
+                             &                       & Kendall-$\tau$ & -0.79 &  \textbf{-0.93} & \textbf{-0.93} & \textbf{-0.93}        \\
 \hline
-\multirow{3}{4em}{ResNet}    & \multirow{3}{1em}{5}  & RMSE           &  0.9 & 1.4  & \textbf{0.61} & 0.66           \\
-                             &                       & $R^2$          &  XXX &  XXX &  XXX          &  XXX           \\
-                             &                       & Kendall-$\tau$ &  XXX &  XXX &  XXX          &  XXX           \\
+\multirow{3}{4em}{ResNet}    & \multirow{3}{1em}{5}  & RMSE           &  0.9 & 0.97 & \textbf{0.61} & 0.66        \\
+                             &                       & $R^2$          &  0.92 &  0.9 &  \textbf{0.96}   &  0.9         \\
+                             &                       & Kendall-$\tau$ &  -1.0 &  -1.0 &  -1.0          &  -1.0        \\
 \hline
-\multirow{3}{4em}{ResNet-1K} & \multirow{3}{1em}{19} & RMSE           &  2.4 & 3.6  & \textbf{1.8}  & 1.9            \\
-                             &                       & $R^2$          &  XXX &  XXX &  XXX          &  XXX           \\
-                             &                       & Kendall-$\tau$ &  XXX &  XXX &  XXX          &  XXX           \\
+\multirow{3}{4em}{ResNet-1K} & \multirow{3}{1em}{19} & RMSE           &  2.4 & 2.8  & \textbf{1.8}  & 1.9          \\
+                             &                       & $R^2$          &  0.81 &  0.74 &  \textbf{0.89}    &  0.88           \\
+                             &                       & Kendall-$\tau$ &  -0.79 &  -0.79 &  \textbf{-0.89}  &  -0.88           \\
 \hline
-\multirow{3}{4em}{DenseNet}  & \multirow{3}{1em}{4}  & RMSE           &  0.3 & 0.26 & \textbf{0.16} & 0.21           \\
-                             &                       & $R^2$          &  XXX &  XXX &  XXX          &  XXX           \\
-                             &                       & Kendall-$\tau$ &  XXX &  XXX &  XXX          &  XXX           \\
+\multirow{3}{4em}{DenseNet}  & \multirow{3}{1em}{4}  & RMSE           &  0.3 & \textbf{0.11}  & 0.16 & 0.21       \\
+                             &                       & $R^2$          &  0.93 &  \textbf{0.99} &  0.98      &  0.97    \\
+                             &                       & Kendall-$\tau$ &  -1.0 &  -1.0 &  -1.0         &  -1.0     \\
 \hline

### 2020-12-31 submission/NatCom_nlp_models.tex
part way through nlp

@@
 Instead, it should probably be interpreted as being due to a similar mechanism to how distillation can ``damage'' otherwise good models.
-In contrast to norm-based metrics, PL-based metrics, including the Log $\alpha$-Norm metric and the Weighted Alpha metric display consistent behavior, even on poorly trained models.
-One can use these metrics to help identify when architectures need repair and when more and/or better data are needed.
+In contrast to norm-based metrics, PL-based metrics, including the Log $\alpha$-Norm metric and the Weighted Alpha metric, display more consistent behavior, even on less well-trained models.
+To help identify when architectures need repair and when more and/or better data are needed,
+one can use these metrics, 
+as well as the decomposition of the Weighted Alpha metric ($\alpha\log\lambda_{max}$) into its PL component ($\alpha$) and its norm component ($\log\lambda_{max}$), for each layer.
 
@@
 
-The OpenAI GPT and GPT2 models provide the opportunity to analyze two effects: training the same model with different data set sizes; and increasing the sizes of both the data set and the architectures simultaneously.
+The OpenAI GPT and GPT2 series of models provide the opportunity to analyze two effects: increasing the sizes of both the data set and the architectures simultaneously; and training the same model with low-quality data versus high-quality data.
 These models have the ability to generate fake text that appears to the human to be real, and they have generated media attention because of the potential for their misuse.
@@
 Later, OpenAI released a much improved model, GPT2-small, which has the same architecture and number of layers as GPT, but which has been trained on a larger and better data set, making it remarkably good at generating (near) human-quality fake text.
+Subsequent models in the GPT2 were larger and trained to more data.
+By comparing GPT2-small to GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously, as well as analyze well-trained versus very-well-trained models.
 By comparing the poorly-trained GPT to the well-trained GPT2-small, we can identify empirical indicators for when a model has been poorly-trained and thus may perform poorly when deployed.
-By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously, as well as analyze well-trained versus very-well-trained models.
 The GPT models we analyze are deployed with the popular HuggingFace PyTorch library~\cite{huggingface}.
@@
 GPT2-small & 49 & 2.04  & 2.54& 9.62 & 9.87 \\
-\hline
+%\hline
 GPT2-medium & 98 & 2.08 & 2.58& 9.74 & 10.01 \\
@@
 
-See Table \ref{table:nlp} for a summary of results for the four quality metrics for the OpenAI GPT and GPT2 pretrained models.
+We examine the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) for the OpenAI GPT and GPT2 pretrained models.
+See Table \ref{table:nlp} for a summary of results.
+%
+Comparing trends between GPT2-medium to GPT2-large to GPT2-xl,
+observe that (with one minor exception involving the log Frobenius norm metric) all four metrics decrease as one goes from medium to large to xl.
+This indicates that the larger models indeed look better than the smaller models, as expected.
+(((
+For these well-trained models, the norm metrics now behave as expected, i.e., there is no Scale Collapse and norms are decreasing with increasing~accuracy.
+XXX.  MOVE.
+)))
+GPT2-small violates this general trend, but only very slightly.
+This could be due to under-optimization of the GPT2-small model, or since it is the smallest, and these metrics are most relevant for models at scale.
+%
 Comparing trends between GPT and GPT2-small, 
-observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are smaller for the higher-quality model (higher quality since GPT was trained to better data), when the number of layers is held fixed.
+observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are larger for the higher-quality model (higher quality since GPT2-small was trained to better data), when the number of layers is held fixed.
+XXX.  POINT OUT AND DRILL DOWN.
+This is unexpected XXX AND DRILL DOWN.
+
 In the poorly-trained GPT model, the norm metrics all exhibit Scale Collapse, compared to well-trained GPT2-small model.
-
-Comparing trends between GPT2-medium to GPT2-large to GPT2-xl,
-observe that (with one minor exception involving the log Frobenius norm metric) all four metrics decrease as one goes from medium to large to xl, indicating that the larger models indeed look better than the smaller models.
-For these well-trained models, the norm metrics now behave as expected, i.e., there is no Scale Collapse and norms are decreasing with increasing~accuracy.
+XXX.  WHERE.
 
@@
 We now look across series of increasingly improving GPT2 models (well-trained versus very-well-trained models), by examining both the PL exponent $\alpha$ as well as the Log Norm metrics.  
-In general, as we move from GPT2-medium to GPT2-xl, histograms for both $\alpha$ exponents and the Log Norm metrics downshift from larger to smaller values. 
 Figure \ref{fig:gpt2-histograms} shows the histograms over the layer weight matrices for fitted PL exponent $\alpha$ and the Log Alpha Norm metric. 
+In general, and as expected, as we move from GPT2-medium to GPT2-xl, histograms for both $\alpha$ exponents and the Log Norm metrics downshift from larger to smaller values.
 
-The average $\alpha$ decreases with increasing model size, although the differences are less noticeable between the differing well-trained 
-versus very-well-trained GTP2 models than between the poorly-trained versus well-trained GPT and GPT2-small models.
-Unlike GPT, the layer Log Alpha Norms behave more as expected for GPT2 layers, with the larger models consistently having smaller norms. 
-Similarly, the Log Spectral Norm also decreases on average with the larger models (not shown).  
+From Figure~\ref{fig:gpt2-alpha-hist}, we see that
+$\bar{\alpha}$, the average $\alpha$ value, decreases with increasing model size (XXX for GPT2-medium, XXX for GPT2-large, and XXX for GPT2-xl), although the differences are less noticeable between the differing well-trained versus very-well-trained GTP2 models than between the poorly-trained versus well-trained GPT and GPT2-small models.
+Also, from Figure~\ref{fig:gpt2-pnorm-hist}, we see that, 
+unlike GPT, the layer Log Alpha Norms behave more as expected for GPT2 layers, with the larger models consistently having smaller norms (XXX for GPT2-medium, XXX for GPT2-large, and XXX for GPT2-xl). 
+Similarly, the Log Spectral Norm also decreases on average with the larger models (XXX for GPT2-medium, XXX for GPT2-large, and XXX for GPT2-xl).
+% (not shown).  
 As expected, the norm metrics can indeed distinguish among well-trained versus very-well-trained models.
 
-However, while the peaks of the $\alpha$ are getting smaller, towards $2.0$, the tails of the distribution shifts right, with larger GPT2 models having more unusually large $\alpha$ (also not shown). 
-This suggests that these larger GPT2 models are still under-optimized/over-parameterized (relative to the data on which they were trained) and that they have capacity to support datasets even larger than the recent XL $1.5B$ release~\cite{gpt2-xl}.
+However, while the means and peaks of the $\alpha$ distributions are getting smaller, towards $2.0$, Figure~\ref{fig:gpt2-alpha-hist} also shows that the tails of the distribution shifts right, with larger GPT2 models having more unusually large $\alpha$ values.
+This is unexpected, and it suggests that these larger GPT2 models are still under-optimized/over-parameterized (relative to the data on which they were trained) and that they have capacity to support datasets even larger than the recent XL $1.5B$ release~\cite{gpt2-xl}.
+This does not contradict recent theoretical work on the benefits of over-parameterization~\cite{BHMM19}, e.g., since in practice these extremely large models are not fully optimized.
+Subsequent refinements to these models, and other models such as BERT, indicate that this is likely the~case.
 

### 2020-12-31 submission/NatCom_appendix.tex
draft for appendix figures

@@
 
-\begin{figure}[t]
-    \centering
-    \subfigure[ImageNet 1K]{
-        \includegraphics[width=4.9cm]{img/all-imagenet-1k_alpha.png}
-        \label{fig:imagenet1k-alpha}
-    }
-    \qquad
-    \subfigure[ CIFAR 10 ]{
-        \includegraphics[width=4.9cm]{img/all-cifar-10_alpha.png}
-        \label{fig:cifar10.alpha}
-    }
-    \qquad
-    \subfigure[ CIFAR 100 ]{
-        \includegraphics[width=4.9cm]{img/all-cifar-100_alpha.png}
-        \label{fig:cifar100.alpha}
-    }
-    \qquad
-    \subfigure[ SVHN ]{
-        \includegraphics[width=4.9cm]{img/all-svhn_alpha.png}
-        \label{fig:svhn.alpha}
-    }
-    \qquad
-    \subfigure[ CUB 200 ]{
-        \includegraphics[width=4.9cm]{img/all-cub-200-2011_alpha.png}
-        \label{fig:cub200.alpha}
-    }
-    \caption{PL exponent $\alpha$ versus reported Top1 Test Accuracies for pretrained DNNs available for different datasets (as segmented in Table~\ref{table:datasets}).
-            }
-    \label{fig:DSalphas}
-\end{figure}
-
-\begin{figure}[t]
-    \centering
-    \subfigure[ ResNet ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_01}
-    }
-    \subfigure[ SENet/SE-ResNet/SE-PreResNet/SE-ResNeXt ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_02}
-    }
-    \subfigure[ DIA-ResNet/DIA-PreResNet ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_03}
-    }
-    \subfigure[ ResNeXt ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_04}
-    }
-    \subfigure[ WRN ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_05}
-    }
-    \subfigure[ DLA ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_06}
-    }
-    \subfigure[ PreResNet ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_07}
-    }
-    \subfigure[ ProxylessNAS ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_08}
-    }
-    \subfigure[ VGG/BN-VGG ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_09}
-    }
-    \subfigure[ IGCV3 ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_10}
-    }
-    \subfigure[ EfficientNet ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_11}
-    }
-    \subfigure[ SqueezeNext/SqNxt ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_12}
-    }
-    \subfigure[ ShuffleNet ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_13}
-    }
-    \subfigure[ DRN-C/DRN-D ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_14}
-    }
-    \subfigure[ ESPNetv2 ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_15}
-    }
-    \subfigure[ HRNet ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_16}
-    }
-    \subfigure[ SqueezeNet/SqueezeResNet ]{
-        \includegraphics[width=2.7cm]{img/hex.png}
-        \label{fig:ARCHalphas_17}
-    }
-    \caption{PL exponent $\alpha$ versus reported Top1 Test Accuracies for pretrained DNNs available for different architecture series (as segmented in Table~\ref{table:architectures}).
-            }
-    \label{fig:ARCHalphas}
-\end{figure}
+%%\begin{figure}[t]
+%%    \centering
+%%    \subfigure[ImageNet 1K]{
+%%        \includegraphics[width=4.9cm]{img/all-imagenet-1k_alpha.png}
+%%        \label{fig:imagenet1k-alpha}
+%%    }
+%%    \qquad
+%%    \subfigure[ CIFAR 10 ]{
+%%        \includegraphics[width=4.9cm]{img/all-cifar-10_alpha.png}
+%%        \label{fig:cifar10.alpha}
+%%    }
+%%    \qquad
+%%    \subfigure[ CIFAR 100 ]{
+%%        \includegraphics[width=4.9cm]{img/all-cifar-100_alpha.png}
+%%        \label{fig:cifar100.alpha}
+%%    }
+%%    \qquad
+%%    \subfigure[ SVHN ]{
+%%        \includegraphics[width=4.9cm]{img/all-svhn_alpha.png}
+%%        \label{fig:svhn.alpha}
+%%    }
+%%    \qquad
+%%    \subfigure[ CUB 200 ]{
+%%        \includegraphics[width=4.9cm]{img/all-cub-200-2011_alpha.png}
+%%        \label{fig:cub200.alpha}
+%%    }
+%%    \caption{PL exponent $\alpha$ versus reported Top1 Test Accuracies for pretrained DNNs available for different datasets (as segmented in Table~\ref{table:datasets}).
+%%            }
+%%    \label{fig:DSalphas}
+%%\end{figure}
+%%
+%% 
+%%\begin{figure}[t]
+%%    \centering
+%%    \subfigure[ ResNet ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_01}
+%%    }
+%%    \subfigure[ SENet/SE-ResNet/SE-PreResNet/SE-ResNeXt ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_02}
+%%    }
+%%    \subfigure[ DIA-ResNet/DIA-PreResNet ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_03}
+%%    }
+%%    \subfigure[ ResNeXt ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_04}
+%%    }
+%%    \subfigure[ WRN ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_05}
+%%    }
+%%    \subfigure[ DLA ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_06}
+%%    }
+%%    \subfigure[ PreResNet ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_07}
+%%    }
+%%    \subfigure[ ProxylessNAS ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_08}
+%%    }
+%%    \subfigure[ VGG/BN-VGG ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_09}
+%%    }
+%%    \subfigure[ IGCV3 ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_10}
+%%    }
+%%    \subfigure[ EfficientNet ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_11}
+%%    }
+%%    \subfigure[ SqueezeNext/SqNxt ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_12}
+%%    }
+%%    \subfigure[ ShuffleNet ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_13}
+%%    }
+%%    \subfigure[ DRN-C/DRN-D ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_14}
+%%    }
+%%    \subfigure[ ESPNetv2 ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_15}
+%%    }
+%%    \subfigure[ HRNet ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_16}
+%%    }
+%%    \subfigure[ SqueezeNet/SqueezeResNet ]{
+%%        \includegraphics[width=2.7cm]{img/hex.png}
+%%        \label{fig:ARCHalphas_17}
+%%    }
+%%    \caption{PL exponent $\alpha$ versus reported Top1 Test Accuracies for pretrained DNNs available for different architecture series (as segmented in Table~\ref{table:architectures}).
+%%            }
+%%    \label{fig:ARCHalphas}
+%%\end{figure}
 
@@
 Still, overall, in nearly all large, pre-trained DNNs, the correlations in the  weight matrices exhibit a remarkable Universality, being both Heavy Tailed, and having small---but not too small---PL exponents.
+
+\begin{figure}[t]
+    \centering
+    \subfigure[ XXX, XXX ]{
+        \includegraphics[width=3.7cm]{img/hex.png}
+        \label{fig:summary_regressions_A_01}
+    }
+    \subfigure[ XXX, XXX ]{
+        \includegraphics[width=3.7cm]{img/hex.png}
+        \label{fig:summary_regressions_A_02}
+    }
+    \subfigure[ XXX, XXX ]{
+        \includegraphics[width=3.7cm]{img/hex.png}
+        \label{fig:summary_regressions_A_03}
+    }
+    \subfigure[ XXX, XXX ]{
+        \includegraphics[width=3.7cm]{img/hex.png}
+        \label{fig:summary_regressions_A_04}
+    }
+    \subfigure[ XXX, XXX ]{
+        \includegraphics[width=3.7cm]{img/hex.png}
+        \label{fig:summary_regressions_A_05}
+    }
+    \subfigure[ XXX, XXX ]{
+        \includegraphics[width=3.7cm]{img/hex.png}
+        \label{fig:summary_regressions_A_06}
+    }
+    \subfigure[ XXX, XXX ]{
+        \includegraphics[width=3.7cm]{img/hex.png}
+        \label{fig:summary_regressions_A_07}
+    }
+    \subfigure[ XXX, XXX ]{
+        \includegraphics[width=3.7cm]{img/hex.png}
+        \label{fig:summary_regressions_A_08}
+    }
+    \subfigure[ XXX, XXX ]{
+        \includegraphics[width=3.7cm]{img/hex.png}
+        \label{fig:summary_regressions_A_09}
+    }
+    \subfigure[ XXX, XXX ]{
+        \includegraphics[width=3.7cm]{img/hex.png}
+        \label{fig:summary_regressions_A_10}
+    }
+    \subfigure[ XXX, XXX ]{
+        \includegraphics[width=3.7cm]{img/hex.png}
+        \label{fig:summary_regressions_A_11}
+    }
+    \subfigure[ XXX, XXX ]{
+        \includegraphics[width=3.7cm]{img/hex.png}
+        \label{fig:summary_regressions_A_12}
+    }
+    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
+             Each row corresponds to a different dataset-model pair:
+             imagenet-1k and ResNet; 
+             imagenet-1k and EfficientNet; and 
+             imagenet-1k and PreResNet, repsectively.
+             Each column column corresponds to a different metric:
+             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
+             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
+             $\hat{\alpha}$; and
+             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$, repsectively.
+            }
+    \label{fig:summary_regressions_A}
+\end{figure}

### 2021-01-01 submission/NatCom_appendix.tex
put all those figs in appendix

@@
 % V1 REMOVED FOR V2 % Example regressions are shown in Figure~\ref{fig:DSalphas}, as available in the results notebook.
-Example regressions are shown in 
-Figure~\ref{fig:DSalphas}, for models segmented by dataset, as listed in Table~\ref{table:datasets}, 
-and in 
-Figure~\ref{fig:ARCHalphas}, for models segmented by architecture, as listed in Table~\ref{table:architectures}.
+%%Example regressions are shown in 
+%%Figure~\ref{fig:DSalphas}, for models segmented by dataset, as listed in Table~\ref{table:datasets}, 
+%%and in 
+%%Figure~\ref{fig:ARCHalphas}, for models segmented by architecture, as listed in Table~\ref{table:architectures}.
 \michael{Comment here on differences.}
@@
 We record the $R^{2}$ and $MSE$ for each metric, averaged over all regressions for all architectures and datasets.
+\michael{Mention KTau.}
 See Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
@@
 
+See
+Figure~\ref{fig:summary_regressions_A},
+Figure~\ref{fig:summary_regressions_B},
+Figure~\ref{fig:summary_regressions_C},
+Figure~\ref{fig:summary_regressions_D},
+Figure~\ref{fig:summary_regressions_E},
+Figure~\ref{fig:summary_regressions_F},
+Figure~\ref{fig:summary_regressions_G},
+Figure~\ref{fig:summary_regressions_H}, and
+Figure~\ref{fig:summary_regressions_A} for more details.
+\michael{Put a bit of discussion, e.g., on best and worst.}
+
 \michael{Probably some discussion of why these metrics are the right ones, e.g., as opposed to $p$ values, to address R1 comments.}
+
+\input{figs_largeScale_A}
+\input{figs_largeScale_B}
+\input{figs_largeScale_C}
+\input{figs_largeScale_D}
+\input{figs_largeScale_E}
+\input{figs_largeScale_F}
+\input{figs_largeScale_G}
+\input{figs_largeScale_H}
+\input{figs_largeScale_I}
 
@@
 Still, overall, in nearly all large, pre-trained DNNs, the correlations in the  weight matrices exhibit a remarkable Universality, being both Heavy Tailed, and having small---but not too small---PL exponents.
-
-\begin{figure}[t]
-    \centering
-    \subfigure[ XXX, XXX ]{
-        \includegraphics[width=3.7cm]{img/hex.png}
-        \label{fig:summary_regressions_A_01}
-    }
-    \subfigure[ XXX, XXX ]{
-        \includegraphics[width=3.7cm]{img/hex.png}
-        \label{fig:summary_regressions_A_02}
-    }
-    \subfigure[ XXX, XXX ]{
-        \includegraphics[width=3.7cm]{img/hex.png}
-        \label{fig:summary_regressions_A_03}
-    }
-    \subfigure[ XXX, XXX ]{
-        \includegraphics[width=3.7cm]{img/hex.png}
-        \label{fig:summary_regressions_A_04}
-    }
-    \subfigure[ XXX, XXX ]{
-        \includegraphics[width=3.7cm]{img/hex.png}
-        \label{fig:summary_regressions_A_05}
-    }
-    \subfigure[ XXX, XXX ]{
-        \includegraphics[width=3.7cm]{img/hex.png}
-        \label{fig:summary_regressions_A_06}
-    }
-    \subfigure[ XXX, XXX ]{
-        \includegraphics[width=3.7cm]{img/hex.png}
-        \label{fig:summary_regressions_A_07}
-    }
-    \subfigure[ XXX, XXX ]{
-        \includegraphics[width=3.7cm]{img/hex.png}
-        \label{fig:summary_regressions_A_08}
-    }
-    \subfigure[ XXX, XXX ]{
-        \includegraphics[width=3.7cm]{img/hex.png}
-        \label{fig:summary_regressions_A_09}
-    }
-    \subfigure[ XXX, XXX ]{
-        \includegraphics[width=3.7cm]{img/hex.png}
-        \label{fig:summary_regressions_A_10}
-    }
-    \subfigure[ XXX, XXX ]{
-        \includegraphics[width=3.7cm]{img/hex.png}
-        \label{fig:summary_regressions_A_11}
-    }
-    \subfigure[ XXX, XXX ]{
-        \includegraphics[width=3.7cm]{img/hex.png}
-        \label{fig:summary_regressions_A_12}
-    }
-    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
-             Each row corresponds to a different dataset-model pair:
-             imagenet-1k and ResNet; 
-             imagenet-1k and EfficientNet; and 
-             imagenet-1k and PreResNet, repsectively.
-             Each column column corresponds to a different metric:
-             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
-             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
-             $\hat{\alpha}$; and
-             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$, repsectively.
-            }
-    \label{fig:summary_regressions_A}
-\end{figure}

### 2021-01-01 submission/figs_largeScale_A.tex
put all those figs in appendix

\begin{figure}[t]
    \centering
    \subfigure[ imagenet-1k--ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ResNet_lognorm}
        \label{fig:summary_regressions_A_01}
    }
    \subfigure[ imagenet-1k--ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ResNet_spectralnormlog}
        \label{fig:summary_regressions_A_02}
    }
    \subfigure[ imagenet-1k--ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ResNet_alpha_weighted}
        \label{fig:summary_regressions_A_03}
    }
    \subfigure[ imagenet-1k--ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ResNet_logpnorm}
        \label{fig:summary_regressions_A_04}
    }
    \subfigure[ imagenet-1k--EfficientNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_EfficientNet_lognorm}
        \label{fig:summary_regressions_A_05}
    }
    \subfigure[ imagenet-1k--EfficientNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_EfficientNet_spectralnormlog}
        \label{fig:summary_regressions_A_06}
    }
    \subfigure[ imagenet-1k--EfficientNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_EfficientNet_alpha_weighted}
        \label{fig:summary_regressions_A_07}
    }
    \subfigure[ imagenet-1k--EfficientNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_EfficientNet_logpnorm}
        \label{fig:summary_regressions_A_08}
    }
    \subfigure[ imagenet-1k--PreResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_PreResNet_lognorm}
        \label{fig:summary_regressions_A_09}
    }
    \subfigure[ imagenet-1k--PreResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_PreResNet_spectralnormlog}
        \label{fig:summary_regressions_A_10}
    }
    \subfigure[ imagenet-1k--PreResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_PreResNet_alpha_weighted}
        \label{fig:summary_regressions_A_11}
    }
    \subfigure[ imagenet-1k--PreResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_PreResNet_logpnorm}
        \label{fig:summary_regressions_A_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             imagenet-1k--ResNet; 
             imagenet-1k--EfficientNet; 
             and 
             imagenet-1k--PreResNet; 
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_A}
\end{figure}

### 2021-01-01 submission/figs_largeScale_B.tex
put all those figs in appendix

 
\begin{figure}[t]
    \centering
    \subfigure[ imagenet-1k--ShuffleNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ShuffleNet_lognorm}
        \label{fig:summary_regressions_B_01}
    }
    \subfigure[ imagenet-1k--ShuffleNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ShuffleNet_spectralnormlog}
        \label{fig:summary_regressions_B_02}
    }
    \subfigure[ imagenet-1k--ShuffleNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ShuffleNet_alpha_weighted}
        \label{fig:summary_regressions_B_03}
    }
    \subfigure[ imagenet-1k--ShuffleNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ShuffleNet_logpnorm}
        \label{fig:summary_regressions_B_04}
    }
    \subfigure[ imagenet-1k--VGG, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_VGG_lognorm}
        \label{fig:summary_regressions_B_05}
    }
    \subfigure[ imagenet-1k--VGG, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_VGG_spectralnormlog}
        \label{fig:summary_regressions_B_06}
    }
    \subfigure[ imagenet-1k--VGG, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_VGG_alpha_weighted}
        \label{fig:summary_regressions_B_07}
    }
    \subfigure[ imagenet-1k--VGG, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_VGG_logpnorm}
        \label{fig:summary_regressions_B_08}
    }
    \subfigure[ imagenet-1k--DLA, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DLA_lognorm}
        \label{fig:summary_regressions_B_09}
    }
    \subfigure[ imagenet-1k--DLA, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DLA_spectralnormlog}
        \label{fig:summary_regressions_B_10}
    }
    \subfigure[ imagenet-1k--DLA, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DLA_alpha_weighted}
        \label{fig:summary_regressions_B_11}
    }
    \subfigure[ imagenet-1k--DLA, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DLA_logpnorm}
        \label{fig:summary_regressions_B_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             imagenet-1k--ShuffleNet;
             imagenet-1k--VGG;
             and
             imagenet-1k--DLA;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_B}
\end{figure}

### 2021-01-01 submission/figs_largeScale_C.tex
put all those figs in appendix

 
\begin{figure}[t]
    \centering
    \subfigure[ imagenet-1k--HRNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_HRNet_lognorm}
        \label{fig:summary_regressions_C_01}
    }
    \subfigure[ imagenet-1k--HRNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_HRNet_spectralnormlog}
        \label{fig:summary_regressions_C_02}
    }
    \subfigure[ imagenet-1k--HRNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_HRNet_alpha_weighted}
        \label{fig:summary_regressions_C_03}
    }
    \subfigure[ imagenet-1k--HRNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_HRNet_logpnorm}
        \label{fig:summary_regressions_C_04}
    }
    \subfigure[ imagenet-1k--DRN-C, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DRN_C_lognorm}
        \label{fig:summary_regressions_C_05}
    }
    \subfigure[ imagenet-1k--DRN-C, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DRN_C_spectralnormlog}
        \label{fig:summary_regressions_C_06}
    }
    \subfigure[ imagenet-1k--DRN-C, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DRN_C_alpha_weighted}
        \label{fig:summary_regressions_C_07}
    }
    \subfigure[ imagenet-1k--DRN-C, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DRN_C_logpnorm}
        \label{fig:summary_regressions_C_08}
    }
    \subfigure[ imagenet-1k--SqueezeNext, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNext_lognorm}
        \label{fig:summary_regressions_C_09}
    }
    \subfigure[ imagenet-1k--SqueezeNext, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNext_spectralnormlog}
        \label{fig:summary_regressions_C_10}
    }
    \subfigure[ imagenet-1k--SqueezeNext, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNext_alpha_weighted}
        \label{fig:summary_regressions_C_11}
    }
    \subfigure[ imagenet-1k--SqueezeNext, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNext_logpnorm}
        \label{fig:summary_regressions_C_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             imagenet-1k--HRNet;
             imagenet-1k--DRN-C;
             and
             imagenet-1k--SqueezeNext;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_C}
\end{figure}

### 2021-01-01 submission/figs_largeScale_D.tex
put all those figs in appendix

 
\begin{figure}[t]
    \centering
    \subfigure[ imagenet-1k--ESPNetv2, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ESPNetv2_lognorm}
        \label{fig:summary_regressions_D_01}
    }
    \subfigure[ imagenet-1k--ESPNetv2, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ESPNetv2_spectralnormlog}
        \label{fig:summary_regressions_D_02}
    }
    \subfigure[ imagenet-1k--ESPNetv2, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ESPNetv2_alpha_weighted}
        \label{fig:summary_regressions_D_03}
    }
    \subfigure[ imagenet-1k--ESPNetv2, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ESPNetv2_logpnorm}
        \label{fig:summary_regressions_D_04}
    }
    \subfigure[ imagenet-1k--SqueezeNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNet_lognorm}
        \label{fig:summary_regressions_D_05}
    }
    \subfigure[ imagenet-1k--SqueezeNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNet_spectralnormlog}
        \label{fig:summary_regressions_D_06}
    }
    \subfigure[ imagenet-1k--SqueezeNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNet_alpha_weighted}
        \label{fig:summary_regressions_D_07}
    }
    \subfigure[ imagenet-1k--SqueezeNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNet_logpnorm}
        \label{fig:summary_regressions_D_08}
    }
    \subfigure[ imagenet-1k--ProxylessNAS, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ProxylessNAS_lognorm}
        \label{fig:summary_regressions_D_09}
    }
    \subfigure[ imagenet-1k--ProxylessNAS, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ProxylessNAS_spectralnormlog}
        \label{fig:summary_regressions_D_10}
    }
    \subfigure[ imagenet-1k--ProxylessNAS, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ProxylessNAS_alpha_weighted}
        \label{fig:summary_regressions_D_11}
    }
    \subfigure[ imagenet-1k--ProxylessNAS, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ProxylessNAS_logpnorm}
        \label{fig:summary_regressions_D_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             imagenet-1k--ESPNetv2;
             imagenet-1k--SqueezeNet;
             and
             imagenet-1k--ProxylessNAS;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_D}
\end{figure}

### 2021-01-01 submission/figs_largeScale_E.tex
put all those figs in appendix

 
\begin{figure}[t]
    \centering
    \subfigure[ imagenet-1k--IGCV3, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_IGCV3_lognorm}
        \label{fig:summary_regressions_E_01}
    }
    \subfigure[ imagenet-1k--IGCV3, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_IGCV3_spectralnormlog}
        \label{fig:summary_regressions_E_02}
    }
    \subfigure[ imagenet-1k--IGCV3, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_IGCV3_alpha_weighted}
        \label{fig:summary_regressions_E_03}
    }
    \subfigure[ imagenet-1k--IGCV3, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_IGCV3_logpnorm}
        \label{fig:summary_regressions_E_04}
    }
    \subfigure[ cifar-10--ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_ResNet_lognorm}
        \label{fig:summary_regressions_E_05}
    }
    \subfigure[ cifar-10--ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_ResNet_spectralnormlog}
        \label{fig:summary_regressions_E_06}
    }
    \subfigure[ cifar-10--ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_ResNet_alpha_weighted}
        \label{fig:summary_regressions_E_07}
    }
    \subfigure[ cifar-10--ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_ResNet_logpnorm}
        \label{fig:summary_regressions_E_08}
    }
    \subfigure[ cifar-10--DIA-ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_DIA_ResNet_lognorm}
        \label{fig:summary_regressions_E_09}
    }
    \subfigure[ cifar-10--DIA-ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_DIA_ResNet_spectralnormlog}
        \label{fig:summary_regressions_E_10}
    }
    \subfigure[ cifar-10--DIA-ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_DIA_ResNet_alpha_weighted}
        \label{fig:summary_regressions_E_11}
    }
    \subfigure[ cifar-10--DIA-ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_DIA_ResNet_logpnorm}
        \label{fig:summary_regressions_E_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             imagenet-1k--IGCV3;
             cifar-10--ResNet;
             and
             cifar-10--DIA-ResNet;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_E}
\end{figure}

### 2021-01-01 submission/figs_largeScale_F.tex
put all those figs in appendix

 
\begin{figure}[t]
    \centering
    \subfigure[ cifar-10--SENet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_SENet_lognorm}
        \label{fig:summary_regressions_F_01}
    }
    \subfigure[ cifar-10--SENet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_SENet_spectralnormlog}
        \label{fig:summary_regressions_F_02}
    }
    \subfigure[ cifar-10--SENet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_SENet_alpha_weighted}
        \label{fig:summary_regressions_F_03}
    }
    \subfigure[ cifar-10--SENet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_SENet_logpnorm}
        \label{fig:summary_regressions_F_04}
    }
    \subfigure[ cifar-100--ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_ResNet_lognorm}
        \label{fig:summary_regressions_F_05}
    }
    \subfigure[ cifar-100--ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_ResNet_spectralnormlog}
        \label{fig:summary_regressions_F_06}
    }
    \subfigure[ cifar-100--ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_ResNet_alpha_weighted}
        \label{fig:summary_regressions_F_07}
    }
    \subfigure[ cifar-100--ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_ResNet_logpnorm}
        \label{fig:summary_regressions_F_08}
    }
    \subfigure[ cifar-100--DIA-ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_DIA_ResNet_lognorm}
        \label{fig:summary_regressions_F_09}
    }
    \subfigure[ cifar-100--DIA-ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_DIA_ResNet_spectralnormlog}
        \label{fig:summary_regressions_F_10}
    }
    \subfigure[ cifar-100--DIA-ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_DIA_ResNet_alpha_weighted}
        \label{fig:summary_regressions_F_11}
    }
    \subfigure[ cifar-100--DIA-ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_DIA_ResNet_logpnorm}
        \label{fig:summary_regressions_F_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             cifar-10--SENet;
             cifar-100--ResNet;
             and
             cifar-100--DIA-ResNet;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_F}
\end{figure}

### 2021-01-01 submission/figs_largeScale_G.tex
put all those figs in appendix

 
\begin{figure}[t]
    \centering
    \subfigure[ cifar-100--SENet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_SENet_lognorm}
        \label{fig:summary_regressions_G_01}
    }
    \subfigure[ cifar-100--SENet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_SENet_spectralnormlog}
        \label{fig:summary_regressions_G_02}
    }
    \subfigure[ cifar-100--SENet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_SENet_alpha_weighted}
        \label{fig:summary_regressions_G_03}
    }
    \subfigure[ cifar-100--SENet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_SENet_logpnorm}
        \label{fig:summary_regressions_G_04}
    }
    \subfigure[ cifar-100--WRN, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_WRN_lognorm}
        \label{fig:summary_regressions_G_05}
    }
    \subfigure[ cifar-100--WRN, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_WRN_spectralnormlog}
        \label{fig:summary_regressions_G_06}
    }
    \subfigure[ cifar-100--WRN, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_WRN_alpha_weighted}
        \label{fig:summary_regressions_G_07}
    }
    \subfigure[ cifar-100--WRN, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_WRN_logpnorm}
        \label{fig:summary_regressions_G_08}
    }
    \subfigure[ svhn--ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_ResNet_lognorm}
        \label{fig:summary_regressions_G_09}
    }
    \subfigure[ svhn--ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_ResNet_spectralnormlog}
        \label{fig:summary_regressions_G_10}
    }
    \subfigure[ svhn--ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_ResNet_alpha_weighted}
        \label{fig:summary_regressions_G_11}
    }
    \subfigure[ svhn--ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_ResNet_logpnorm}
        \label{fig:summary_regressions_G_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             cifar-100--SENet;
             cifar-100--WRN;
             and
             svhn--ResNet;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_G}
\end{figure}

### 2021-01-01 submission/figs_largeScale_H.tex
put all those figs in appendix

 
\begin{figure}[t]
    \centering
    \subfigure[ svhn--DIA-ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_DIA_ResNet_lognorm}
        \label{fig:summary_regressions_H_01}
    }
    \subfigure[ svhn--DIA-ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_DIA_ResNet_spectralnormlog}
        \label{fig:summary_regressions_H_02}
    }
    \subfigure[ svhn--DIA-ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_DIA_ResNet_alpha_weighted}
        \label{fig:summary_regressions_H_03}
    }
    \subfigure[ svhn--DIA-ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_DIA_ResNet_logpnorm}
        \label{fig:summary_regressions_H_04}
    }
    \subfigure[ svhn--SENet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_SENet_lognorm}
        \label{fig:summary_regressions_H_05}
    }
    \subfigure[ svhn--SENet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_SENet_spectralnormlog}
        \label{fig:summary_regressions_H_06}
    }
    \subfigure[ svhn--SENet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_SENet_alpha_weighted}
        \label{fig:summary_regressions_H_07}
    }
    \subfigure[ svhn--SENet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_SENet_logpnorm}
        \label{fig:summary_regressions_H_08}
    }
    \subfigure[ svhn--WRN, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_WRN_lognorm}
        \label{fig:summary_regressions_H_09}
    }
    \subfigure[ svhn--WRN, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_WRN_spectralnormlog}
        \label{fig:summary_regressions_H_10}
    }
    \subfigure[ svhn--WRN, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_WRN_alpha_weighted}
        \label{fig:summary_regressions_H_11}
    }
    \subfigure[ svhn--WRN, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_WRN_logpnorm}
        \label{fig:summary_regressions_H_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             svhn--DIA-ResNet;
             svhn--SENet;
             and
             svhn--WRN;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_H}
\end{figure}

### 2021-01-01 submission/figs_largeScale_I.tex
put all those figs in appendix

 
\begin{figure}[t]
    \centering
    \subfigure[ svhn--ResNeXt, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_ResNeXt_lognorm}
        \label{fig:summary_regressions_I_01}
    }
    \subfigure[ svhn--ResNeXt, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_ResNeXt_spectralnormlog}
        \label{fig:summary_regressions_I_02}
    }
    \subfigure[ svhn--ResNeXt, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_ResNeXt_alpha_weighted}
        \label{fig:summary_regressions_I_03}
    }
    \subfigure[ svhn--ResNeXt, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_ResNeXt_logpnorm}
        \label{fig:summary_regressions_I_04}
    }
    \subfigure[ cub-200-2011--ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cub_200_2011_ResNet_lognorm}
        \label{fig:summary_regressions_I_05}
    }
    \subfigure[ cub-200-2011--ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cub_200_2011_ResNet_spectralnormlog}
        \label{fig:summary_regressions_I_06}
    }
    \subfigure[ cub-200-2011--ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cub_200_2011_ResNet_alpha_weighted}
        \label{fig:summary_regressions_I_07}
    }
    \subfigure[ cub-200-2011--ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cub_200_2011_ResNet_logpnorm}
        \label{fig:summary_regressions_I_08}
    }
    \subfigure[ cub-200-2011--SENet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cub_200_2011_SENet_lognorm}
        \label{fig:summary_regressions_I_09}
    }
    \subfigure[ cub-200-2011--SENet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cub_200_2011_SENet_spectralnormlog}
        \label{fig:summary_regressions_I_10}
    }
    \subfigure[ cub-200-2011--SENet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cub_200_2011_SENet_alpha_weighted}
        \label{fig:summary_regressions_I_11}
    }
    \subfigure[ cub-200-2011--SENet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cub_200_2011_SENet_logpnorm}
        \label{fig:summary_regressions_I_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             svhn--ResNeXt;
             cub-200-2011--ResNet;
             and
             cub-200-2011--SENet;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_I}
\end{figure}

### 2021-01-01 submission/NatCom_appendix.tex
updating tables

@@
 \end{center}
-\caption{Datasets used}
+\caption{Datasets used \charles{This table needs updated}}
 \label{table:datasets}
@@
 \begin{center}
-\begin{tabular}{|p{3in}|c|}
+\begin{tabular}{|c|c|c|c|c|c|c|}
 \hline
-Architecture & $\#$ of Models \\
+Architecture & $\#$ of Models & Datasets & & & & \\
 \hline
-ResNet                                     & 30 \\
-SENet/SE-ResNet/SE-PreResNet/SE-ResNeXt    & 24 \\
-DIA-ResNet/DIA-PreResNet                   & 18 \\
-ResNeXt                                    & 12 \\
-WRN                                        & 12 \\
-DLA                                        & 6 \\
-PreResNet                                  & 6 \\
-ProxylessNAS                               & 6 \\
-VGG/BN-VGG                                 & 6 \\
-IGCV3                                      & 6 \\
-EfficientNet                               & 6 \\
-SqueezeNext/SqNxt                          & 6 \\
-ShuffleNet                                 & 6 \\
-DRN-C/DRN-D                                & 6 \\
-ESPNetv2                                   & 6 \\
-HRNet                                      & 6 \\
-SqueezeNet/SqueezeResNet                   & 6 \\
+  & total &imagenet-1k & cifar-10 & cifar-100 & svhn & cub-200-2011 \\
+\hline
+ResNet & 51 &22 & 8 & 8 & 7 & 6 \\
+EfficientNet & 20 &20 & 0 & 0 & 0 & 0 \\
+PreResNet & 14 &14 & 0 & 0 & 0 & 0 \\
+VGG/BN-VGG & 12 &12 & 0 & 0 & 0 & 0 \\
+ShuffleNet & 12 &12 & 0 & 0 & 0 & 0 \\
+DLA & 10 &10 & 0 & 0 & 0 & 0 \\
+HRNet & 9 &9 & 0 & 0 & 0 & 0 \\
+DRN-C/DRN-D & 7 &7 & 0 & 0 & 0 & 0 \\
+SqueezeNext/SqNxt & 6 &6 & 0 & 0 & 0 & 0 \\
+ESPNetv2 & 5 &5 & 0 & 0 & 0 & 0 \\
+SqueezeNet/SqueezeResNet & 4 &4 & 0 & 0 & 0 & 0 \\
+IGCV3 & 4 &4 & 0 & 0 & 0 & 0 \\
+ProxylessNAS & 4 &4 & 0 & 0 & 0 & 0 \\
+DIA-ResNet/DIA-PreResNet & 24 &0 & 8 & 8 & 8 & 0 \\
+SENet/SE-ResNet/... & 20 &0 & 5 & 5 & 4 & 6 \\
+WRN & 8 &0 & 0 & 4 & 4 & 0 \\
+ResNeXt & 4 &0 & 0 & 0 & 4 & 0 \\
 \hline
@@
 \end{center}
-\caption{Architectures used}
+\caption{Architectures used in total and per dataset \charles{I need to double check this}}
 \label{table:architectures}

### 2021-01-01 submission/NatCom_appendix.tex
OMSR Analysis almost done

@@
 \begin{center}
-\begin{tabular}{|p{1in}|c|}
+\begin{tabular}{|c|c|c|}
 \hline
-Figure & Jupyter Notebook \\
+Table & Figure & Jupyter Notebook \\
 \hline
-\ref{fig:vgg-metrics}                                 & WeightWatcher-VGG.ipynb \\
-\ref{fig:resnet-accuracy}                             & WeightWatcher-ResNet.ipynb \\
-\ref{fig:resnet1k-accuracy}                           & WeightWatcher-ResNet-1K.ipynb \\
-\ref{fig:vgg-alpha-layers}                            & WeightWatcher-VGG.ipynb \\
-\ref{fig:resnet-alpha-layer}                          & WeightWatcher-ResNet.ipynb \\
-\ref{fig:densenet-alpha-layer}                        & WeightWatcher-DenseNet.ipynb \\
+& \ref{fig:vgg-metrics}                                 & WeightWatcher-VGG.ipynb \\
+& \ref{fig:resnet-accuracy}                             & WeightWatcher-ResNet.ipynb \\
+& \ref{fig:resnet1k-accuracy}                           & WeightWatcher-ResNet-1K.ipynb \\
+& \ref{fig:vgg-alpha-layers}                            & WeightWatcher-VGG.ipynb \\
+& \ref{fig:resnet-alpha-layer}                          & WeightWatcher-ResNet.ipynb \\
+& \ref{fig:densenet-alpha-layer}                        & WeightWatcher-DenseNet.ipynb \\
 \hline
-\ref{fig:resnet204D5L}                                & WeightWatcher-Intel-Distiller-ResNet20.ipynb \\
+& \ref{fig:resnet204D5L}                                & WeightWatcher-Intel-Distiller-ResNet20.ipynb \\
 \hline
-\ref{fig:GPT-hist}                                    & WeightWatcher-OpenAI-GPT.ipynb \\
-\ref{fig:gpt-alpha-layers}, \ref{fig:gpt2-histograms} & WeightWatcher-OpenAI-GPT2.ipynb \\
+& \ref{fig:GPT-hist}                                    & WeightWatcher-OpenAI-GPT.ipynb \\
+& \ref{fig:gpt-alpha-layers}, \ref{fig:gpt2-histograms} & WeightWatcher-OpenAI-GPT2.ipynb \\
+\hline
+3,7,8,9 & Appendix & OSMR-Analysis.ipynb \\
 \hline
@@
 \end{center}
-\caption{Jupyter notebooks used to reproduce all results in Sections~\ref{sxn:cv} and~\ref{sxn:nlp}.}
+\caption{Jupyter notebooks used to reproduce all results in Sections~\ref{sxn:cv} and~\ref{sxn:nlp}.\charles{Need to update referendes}}
 \label{table:notebooks}
@@
 \hline
- imagenet-1k & ResNet  & 5.96 &  11.03 & \textbf{3.51} & 4.01 \\
- imagenet-1k & EfficientNet  & 2.67 &  \textbf{1.23} & 2.56 & 2.50 \\
- imagenet-1k & PreResNet  & 6.59 &  15.44 & \textbf{3.59} & 3.71 \\
- imagenet-1k & ShuffleNet  & 35.38 &  89.58 & 19.54 & \textbf{18.48} \\
- imagenet-1k & VGG  & 0.84 &  \textbf{0.68} & 1.89 & 1.59 \\
- imagenet-1k & DLA  & 22.41 &  \textbf{8.49} & 14.69 & 15.68 \\
- imagenet-1k & HRNet  & 0.47 &  0.51 & \textbf{0.16} & 0.16 \\
- imagenet-1k & DRN-C  & 0.60 &  0.66 & \textbf{0.40} & 0.48 \\
- imagenet-1k & SqueezeNext  & 21.94 &  21.39 & 13.31 & \textbf{13.23} \\
- imagenet-1k & ESPNetv2  & 13.77 &  14.74 & \textbf{1.87} & 2.53 \\
- imagenet-1k & IGCV3  & 1.94 &  87.76 & 8.48 & \textbf{1.09} \\
- imagenet-1k & ProxylessNAS  & \textbf{0.19} &  0.26 & 0.28 & 0.26 \\
- imagenet-1k & SqueezeNet  & 0.11 &  0.11 & \textbf{0.07} & 0.08 \\
+imagenet-1k & ResNet  & 0.79 &  0.63 & \textbf{0.88} & 0.86 \\
+ imagenet-1k & EfficientNet  & 0.65 &  \textbf{0.84} & 0.67 & 0.67 \\
+ imagenet-1k & PreResNet  & 0.73 &  0.36 & \textbf{0.85} & 0.85 \\
+ imagenet-1k & ShuffleNet  & 0.63 &  0.06 & 0.80 & \textbf{0.81} \\
+ imagenet-1k & VGG  & 0.71 &  \textbf{0.76} & 0.35 & 0.46 \\
+ imagenet-1k & DLA  & 0.11 &  \textbf{0.65} & 0.40 & 0.36 \\
+ imagenet-1k & HRNet  & 0.91 &  0.90 & \textbf{0.97} & 0.97 \\
+ imagenet-1k & DRN-C  & 0.81 &  0.79 & \textbf{0.87} & 0.85 \\
+ imagenet-1k & SqueezeNext  & 0.05 &  0.07 & 0.42 & \textbf{0.43} \\
+ imagenet-1k & ESPNetv2  & 0.42 &  0.38 & \textbf{0.92} & 0.89 \\
+ imagenet-1k & ProxylessNAS  & \textbf{0.68} &  0.56 & 0.53 & 0.58 \\
+ imagenet-1k & SqueezeNet  & 0.01 &  0.00 & \textbf{0.38} & 0.26 \\
+ imagenet-1k & IGCV3  & 0.98 &  0.12 & 0.92 & \textbf{0.99} \\
 \hline
- cifar-10 & ResNet  & 0.31 &  0.30 & \textbf{0.28} & 0.28 \\
- cifar-10 & DIA-ResNet  & \textbf{0.05} &  0.08 & 0.28 & 0.32 \\
- cifar-10 & SENet  & 0.09 &  0.09 & \textbf{0.04} & 0.04 \\
+ cifar-10 & ResNet  & 0.58 &  0.59 & \textbf{0.62} & 0.61 \\
+ cifar-10 & DIA-ResNet  & \textbf{0.96} &  0.93 & 0.74 & 0.71 \\
+ cifar-10 & SENet  & 0.91 &  0.91 & \textbf{0.96} & 0.96 \\
 \hline
- cifar-100 & ResNet  & 4.13 &  4.50 & \textbf{3.06} & 3.06 \\
- cifar-100 & DIA-ResNet  & \textbf{0.36} &  1.38 & 0.93 & 1.02 \\
- cifar-100 & SENet  & 0.36 &  0.43 & \textbf{0.26} & 0.26 \\
- cifar-100 & WRN  & 0.14 &  0.20 & 0.07 & \textbf{0.06} \\
+ cifar-100 & ResNet  & 0.61 &  0.58 & \textbf{0.71} & 0.71 \\
+ cifar-100 & DIA-ResNet  & \textbf{0.96} &  0.85 & 0.90 & 0.89 \\
+ cifar-100 & SENet  & 0.97 &  0.96 & \textbf{0.98} & 0.98 \\
+ cifar-100 & WRN  & 0.32 &  0.04 & 0.66 & \textbf{0.69} \\
 \hline
- svhn & ResNet  & 0.04 &  0.04 & \textbf{0.02} & 0.02 \\
- svhn & DIA-ResNet  & 0.00 &  \textbf{0.00} & 0.02 & 0.02 \\
- svhn & SENet  & \textbf{0.00} &  0.00 & 0.00 & 0.00 \\
- svhn & WRN  & 0.01 &  0.01 & 0.01 & \textbf{0.01} \\
- svhn & ResNeXt  & 0.00 &  \textbf{0.00} & 0.01 & 0.01 \\
+ svhn & ResNet  & 0.69 &  0.70 & \textbf{0.82} & 0.81 \\
+ svhn & DIA-ResNet  & 0.94 &  \textbf{0.95} & 0.78 & 0.77 \\
+ svhn & SENet  & \textbf{0.99} &  0.96 & 0.98 & 0.98 \\
+ svhn & WRN  & 0.13 &  0.10 & 0.20 & \textbf{0.21} \\
+ svhn & ResNeXt  & 0.87 &  \textbf{0.90} & 0.64 & 0.75 \\
 \hline
- cub-200-2011 & ResNet  & 0.20 &  \textbf{0.18} & 3.19 & 3.21 \\
- cub-200-2011 & SENet  & \textbf{1.07} &  1.29 & 1.85 & 1.95 \\
+ cub-200-2011 & ResNet  & 0.94 &  \textbf{0.95} & 0.08 & 0.08 \\
+ cub-200-2011 & SENet  & \textbf{0.66} &  0.59 & 0.41 & 0.38 \\
 \hline
@@
 \vspace{-5mm}
-\caption{$MSE$ results for our analysis of all CV models in Table~\ref{table:datasets}. }
+\caption{$RMSE$ results for our analysis of all CV models in Table~\ref{table:datasets}. }
 \label{table:MSEresults}
@@
 \hline
- imagenet-1k & ResNet  & 0.82 &  0.67 & \textbf{0.90} & 0.88 \\
+imagenet-1k & ResNet  & 0.79 &  0.63 & \textbf{0.88} & 0.86 \\
  imagenet-1k & EfficientNet  & 0.65 &  \textbf{0.84} & 0.67 & 0.67 \\
@@
  imagenet-1k & ShuffleNet  & 0.63 &  0.06 & 0.80 & \textbf{0.81} \\
- imagenet-1k & VGG  & 0.71 &  \textbf{0.76} & 0.35 & 0.45 \\
- imagenet-1k & DLA  & 0.13 &  \textbf{0.67} & 0.43 & 0.39 \\
+ imagenet-1k & VGG  & 0.71 &  \textbf{0.76} & 0.35 & 0.46 \\
+ imagenet-1k & DLA  & 0.11 &  \textbf{0.65} & 0.40 & 0.36 \\
  imagenet-1k & HRNet  & 0.91 &  0.90 & \textbf{0.97} & 0.97 \\
@@
  imagenet-1k & ESPNetv2  & 0.42 &  0.38 & \textbf{0.92} & 0.89 \\
+ imagenet-1k & ProxylessNAS  & \textbf{0.68} &  0.56 & 0.53 & 0.58 \\
+ imagenet-1k & SqueezeNet  & 0.01 &  0.00 & \textbf{0.38} & 0.26 \\
  imagenet-1k & IGCV3  & 0.98 &  0.12 & 0.92 & \textbf{0.99} \\
- imagenet-1k & SqueezeNet  & 0.01 &  0.00 & \textbf{0.38} & 0.26 \\
- imagenet-1k & ProxylessNAS  & \textbf{0.68} &  0.56 & 0.53 & 0.58 \\
 \hline
@@
 \hline
- imagenet-1k & ResNet  & 0.77 &  0.74 & 0.88 & 0.88 \\
- imagenet-1k & EfficientNet  & 0.67 &  0.79 & 0.66 & 0.66 \\
- imagenet-1k & PreResNet  & 0.65 &  0.54 & 0.87 & 0.87 \\
- imagenet-1k & ShuffleNet  & 0.39 &  0.09 & 0.85 & 0.82 \\
- imagenet-1k & VGG  & 0.82 &  0.76 & 0.52 & 0.61 \\
- imagenet-1k & DLA  & 0.51 &  0.82 & 0.78 & 0.69 \\
- imagenet-1k & HRNet  & 0.56 &  0.44 & 0.61 & 0.61 \\
- imagenet-1k & DRN-C  & 0.90 &  0.81 & 0.90 & 0.90 \\
- imagenet-1k & SqueezeNext  & 0.47 &  -0.33 & 0.73 & 0.73 \\
- imagenet-1k & ESPNetv2  & 0.00 &  0.80 & 1.00 & 1.00 \\
- imagenet-1k & SqueezeNet  & 0.33 &  0.00 & 0.67 & 0.67 \\
- imagenet-1k & ProxylessNAS  & 0.33 &  0.33 & 0.33 & 0.33 \\
- imagenet-1k & IGCV3  & 1.00 &  0.00 & 1.00 & 1.00 \\
+imagenet-1k & ResNet  & 0.77 &  0.74 & \textbf{0.88} & 0.88 \\
+ imagenet-1k & EfficientNet  & 0.67 &  \textbf{0.79} & 0.66 & 0.66 \\
+ imagenet-1k & PreResNet  & 0.65 &  0.54 & \textbf{0.87} & 0.87 \\
+ imagenet-1k & ShuffleNet  & 0.39 &  0.09 & \textbf{0.85} & 0.82 \\
+ imagenet-1k & VGG  & \textbf{0.82} &  0.76 & 0.52 & 0.61 \\
+ imagenet-1k & DLA  & 0.51 &  \textbf{0.82} & 0.78 & 0.69 \\
+ imagenet-1k & HRNet  & 0.56 &  0.44 & \textbf{0.61} & 0.61 \\
+ imagenet-1k & DRN-C  & \textbf{0.90} &  0.81 & 0.90 & 0.90 \\
+ imagenet-1k & SqueezeNext  & 0.47 &  -0.33 & \textbf{0.73} & 0.73 \\
+ imagenet-1k & ESPNetv2  & 0.00 &  0.80 & \textbf{1.00} & 1.00 \\
+ imagenet-1k & ProxylessNAS  & \textbf{0.33} &  0.33 & 0.33 & 0.33 \\
+ imagenet-1k & SqueezeNet  & 0.33 &  0.00 & \textbf{0.67} & 0.67 \\
+ imagenet-1k & IGCV3  & \textbf{1.00} &  0.00 & 1.00 & 1.00 \\
 \hline
- cifar-10 & ResNet  & 0.64 &  0.64 & 0.71 & 0.71 \\
- cifar-10 & DIA-ResNet  & 0.79 &  0.50 & 0.86 & 0.79 \\
- cifar-10 & SENet  & 0.74 &  0.95 & 0.95 & 0.95 \\
+ cifar-10 & ResNet  & 0.64 &  0.64 & \textbf{0.71} & 0.71 \\
+ cifar-10 & DIA-ResNet  & 0.79 &  0.50 & \textbf{0.86} & 0.79 \\
+ cifar-10 & SENet  & 0.74 &  \textbf{0.95} & 0.95 & 0.95 \\
 \hline
- cifar-100 & ResNet  & 0.64 &  0.57 & 0.64 & 0.64 \\
- cifar-100 & DIA-ResNet  & 0.93 &  0.43 & 0.93 & 0.93 \\
- cifar-100 & SENet  & 1.00 &  1.00 & 1.00 & 1.00 \\
- cifar-100 & WRN  & 0.67 &  -0.67 & 0.00 & 0.00 \\
+ cifar-100 & ResNet  & \textbf{0.64} &  0.57 & 0.64 & 0.64 \\
+ cifar-100 & DIA-ResNet  & \textbf{0.93} &  0.43 & 0.93 & 0.93 \\
+ cifar-100 & SENet  & \textbf{1.00} &  1.00 & 1.00 & 1.00 \\
+ cifar-100 & WRN  & \textbf{0.67} &  -0.67 & 0.00 & 0.00 \\
 \hline
- svhn & ResNet  & 0.81 &  0.71 & 0.81 & 0.81 \\
- svhn & DIA-ResNet  & 0.86 &  0.86 & 0.57 & 0.57 \\
- svhn & SENet  & 1.00 &  1.00 & 0.67 & 0.67 \\
- svhn & WRN  & -0.33 &  -0.33 & 0.67 & 0.67 \\
- svhn & ResNeXt  & 0.67 &  0.67 & 0.33 & 0.33 \\
+ svhn & ResNet  & \textbf{0.81} &  0.71 & 0.81 & 0.81 \\
+ svhn & DIA-ResNet  & \textbf{0.86} &  0.86 & 0.57 & 0.57 \\
+ svhn & SENet  & \textbf{1.00} &  1.00 & 0.67 & 0.67 \\
+ svhn & WRN  & -0.33 &  -0.33 & \textbf{0.67} & 0.67 \\
+ svhn & ResNeXt  & \textbf{0.67} &  0.67 & 0.33 & 0.33 \\
 \hline
- cub-200-2011 & ResNet  & 1.00 &  1.00 & -0.33 & -0.33 \\
- cub-200-2011 & SENet  & 0.87 &  0.87 & -0.20 & -0.20 \\
+ cub-200-2011 & ResNet  & \textbf{1.00} &  1.00 & -0.33 & -0.33 \\
+ cub-200-2011 & SENet  & \textbf{0.87} &  0.87 & -0.20 & -0.20 \\
 \hline

### 2021-01-01 submission/NatCom_cv_models.tex
got nlp section to local opt

@@
 Going beyond coarse averages to examining quality metrics for each layer weight matrix as a function of depth (or layer id), our metrics can be used to perform model diagnostics and to identify fine-scale properties in a pretrained model.
-We provide three examples of this.
+Doing so involves separating $\hat{\alpha}$ into its two components, $\alpha$ and $\lambda_{max}$, and examining the distributions of each.
+We provide examples of~this.
 
@@
 This leads to the hypothesis that small $\alpha$ values that are stable across multiple layers enable better correlation flow through the network.
-This is similar to recent work on the information bottleneck~\cite{TZ15,ST17_TR}, except that here we work in an entirely unsupervised setting.
+This is similar to recent work on the information bottleneck~\cite{TZ15,ST17_TR}, except that here we work in an entirely unsupervised~setting.
 
@@
 On the other hand, the $\alpha$ values (in Figure~\ref{fig:resnet204Dalpha}) do not differ systematically between the baseline and fine-tuned models.
-Also, $\bar{\alpha}$, the average unweighted baseline $\alpha$, is smaller for the original model than for the fine-tuned model 
+Also, $\bar{\alpha}$, the average unweighted baseline $\alpha$, 
+from Eqn.~(\ref{eqn:alpha_bar}),
+is smaller for the original model than for the fine-tuned model 
 (as predicted by HT-SR Theory, the basis of $\hat{\alpha}$).

### 2021-01-01 submission/NatCom_nlp_models.tex
got nlp section to local opt

@@
 In this regime, the truncated PL fit itself may not be very reliable because the Maximum Likelihood estimator it uses is unreliable in this range.
-The specific $\alpha$ values returned by the truncated PL fits are less reliable, but having large versus small $\alpha$ is reliable.
-If the ESD is visually examined, one can usually describe these $\mathbf{W}$ as in the Bulk-Decay or Bulk-plus-Spikes phase from HT-ST Theory~\cite{MM18_TR,MM19_HTSR_ICML}.
+In this case, the specific $\alpha$ values returned by the truncated PL fits are less reliable, but having large versus small $\alpha$ is reliable.
+If the ESD is visually examined, one can usually describe these $\mathbf{W}$ as in the \textsc{Bulk-Decay} or \textsc{Bulk-plus-Spikes} phase from HT-ST Theory~\cite{MM18_TR,MM19_HTSR_ICML}.
 Previous work~\cite{MM18_TR,MM19_HTSR_ICML} has conjectured that very well-trained DNNs would not have many outlier $\alpha\gg 6$.
@@
 This indicates that the larger models indeed look better than the smaller models, as expected.
-(((
-For these well-trained models, the norm metrics now behave as expected, i.e., there is no Scale Collapse and norms are decreasing with increasing~accuracy.
-XXX.  MOVE.
-)))
 GPT2-small violates this general trend, but only very slightly.
-This could be due to under-optimization of the GPT2-small model, or since it is the smallest, and these metrics are most relevant for models at scale.
-%
-Comparing trends between GPT and GPT2-small, 
-observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are larger for the higher-quality model (higher quality since GPT2-small was trained to better data), when the number of layers is held fixed.
-XXX.  POINT OUT AND DRILL DOWN.
-This is unexpected XXX AND DRILL DOWN.
+This could be due to under-optimization of the GPT2-small model, or since it is the smallest of the GPT2 series, and the metrics we present are most relevant for models at scale.
+Aside from this minor discrepancy, overall for these well-trained models, all these metrics now behave as expected, i.e., there is no Scale Collapse and norms are decreasing with increasing~accuracy.
 
-In the poorly-trained GPT model, the norm metrics all exhibit Scale Collapse, compared to well-trained GPT2-small model.
-XXX.  WHERE.
+Comparing trends between GPT and GPT2-small reveals a different story. 
+Observe that all four metrics increase when going from GPT to GPT2-small, i.e., they are larger for the higher-quality model (higher quality since GPT2-small was trained to better data) and smaller for the lower-quality model, when the number of layers is held fixed.
+This is unexpected.
+Here, too, we can perform model diagnostics, by separating $\hat{\alpha}$ into its two components, $\alpha$ and $\lambda_{max}$, and examining the distributions of each.
+In doing so, we see additional examples of Scale Collapse and additional evidence for Correlation Flow.
 
-Going beyond average values, Figure~\ref{fig:GPT-alpha-hist} shows the histogram (empirical density), for all layers, of $\alpha$ for GPT and GPT2-small.  
-The older deficient GPT has numerous unusually large $\alpha$ exponents---meaning they are not well-described by a PL fit.
-Indeed, we expect that a poorly-trained model will lack good (i.e., small $\alpha$) PL behavior in many/most layers.
-On the other hand, the newer improved GPT2-small model has, on average, smaller $\alpha$ values than the older GPT, with all $\alpha\le6$ and with smaller mean/median $\alpha$.
-It also has far fewer unusually-large outlying $\alpha$ values than GPT.
-From this (and other results not shown), we see that $\alpha$ provides a good quality metric for comparing the poorly-trained GPT versus the well-trained GPT2-small.
-This should be contrasted with the behavior displayed by the Frobenius norm (not shown) and the Spectral~norm.
+\paragraph{Layer Analysis: Scale Collapse in GPT and GPT2.}
 
-\paragraph{Scale Collapse in Poorly Trained Models.}
-
-We next describe the Spectral norm in GPT versus GPT2-small.
+We next examine the Spectral norm in GPT versus GPT2-small.
 In Figure~\ref{fig:GPT-snorm-hist}, the poorly-trained GPT model has a smaller mean/median Spectral norm as well as, spuriously, many much smaller Spectral norms, compared to the well-trained GPT2-small.
 This violates the conventional wisdom that smaller Spectral norms are better.
-Because there are so many anonymously small Spectral norms, the GPT model may be exhibiting a kind of Scale Collapse, like that observed in the distilled CV models (Figure~\ref{fig:resnet204D5L}).
+Because there are so many anomalously small Spectral norms, the GPT model appears to be exhibiting a kind of Scale Collapse, like that observed (in Figure~\ref{fig:resnet204D5L}) for the distilled CV models.
+This demonstrates that, while the Spectral (or Frobenius) norm may correlate well with predicted test error, at least among reasonably well-trained models, it is not a good indicator of the overall model quality in general.
+Na\"ively using it as an empirical quality metric may give spurious results when applied to poorly-trained or otherwise deficient~models.
 
-This demonstrates that, while the Spectral (or Frobenius) norm may correlate well with predicted test error, at least among reasonably well-trained models, it is not a good indicator of the overall model quality in general.
-Using it as an empirical quality metric may give spurious results when applied to poorly-trained or otherwise deficient~models.
-
-\begin{figure}[h]
+\begin{figure}[hbt!] %[t]
     \centering
+    \subfigure[Log Spectral Norm ($\log\Vert\mathbf{W}\Vert_{\infty}$)]{
+        \includegraphics[width=5.5cm]{img/GPT-snorm-hist.png}
+        \label{fig:GPT-snorm-hist}
+       }
+    \qquad
     \subfigure[PL exponent ($\alpha$)]{
@@
         \label{fig:GPT-alpha-hist}
-       }
-    \qquad
-    \subfigure[Log Spectral Norm ($\log\Vert\mathbf{W}\Vert_{\infty}$)]{
-        \includegraphics[width=5.5cm]{img/GPT-snorm-hist.png}
-        \label{fig:GPT-snorm-hist}
        }
@@
 
-Figure~\ref{fig:GPT-snorm-hist} also shows some unusually large Spectral Norms.
-Upon examination, e.g., from Figure \ref{fig:gpt-snorm-layer} (below), these correspond to the first embedding layer(s).
+\begin{figure}[hbt!] %[t]
+    \centering
+    \subfigure[Log Spectral Norm ($\log\Vert\mathbf{W}\Vert_{\infty}$)]{
+        \includegraphics[width=5.5cm]{img/GPT-snorm-depth.png}
+        \label{fig:gpt-snorm-layer}
+    }
+    \quad
+    \subfigure[PL exponent ($\alpha$)]{
+        \includegraphics[width=5.5cm]{img/GPT-alpha-depth.png}
+        \label{fig:gpt-alpha-layer}
+    }
+    \caption{%
+             Log Spectral Norms 
+             %($\log\Vert\mathbf{W}\Vert_{\infty}$) 
+             (in (a)) 
+             and 
+             PL exponents 
+             %($\alpha$) 
+             (in (b)) 
+             for weight matrices from the OpenAI GPT and GPT2-small pretrained models.  
+             (Note that the quantities shown on each Y axis are different.)
+             In the text, this is interpreted in terms of 
+             Scale Collapse
+             and 
+             Correlation~Flow.
+            }
+    \label{fig:gpt-alpha-layers}
+\end{figure}
+
+Figure~\ref{fig:gpt-snorm-layer} shows the Spectral norm as a function of depth (layer id).
+This illustrates two phenomenon.
+First, the large value of Spectral norm (in Figure~\ref{fig:GPT-snorm-hist}) corresponds to the first embedding layer(s).
 These layers have a different effective normalization, and therefore a different scale.
@@
 for details.
-Here, we do not include them in our computed average metrics in Table~\ref{table:nlp}, and we do not include them in the histogram plot in Figure~\ref{fig:GPT-snorm-hist}.
+We do not include them in our computed average metrics in Table~\ref{table:nlp}.
+%% and we do not include them in the histogram plot in Figure~\ref{fig:GPT-snorm-hist}.
+Second, for GPT, there seems to be two types of layers with very different Spectral norms (an effect which is seen, but to a much weaker extent, for GPT2-small).
+Recall that attention models have two types of layers, one small and large; and the Spectral norm (in particular, other norms do too) displays unusually small values for some of these layers for GPT.
+This Scale Collapse for the poorly-trained GPT is similar to what we observed for the distilled ResNet20 model in Figure~\ref{fig:resnet204Dalpha}.
+Because of the anomalous Scale Collapse that is frequently observed in poorly-trained models, these results suggest that scale-dependent norm metrics should not be directly applied to distinguish well-trained versus poorly-trained models.
 
-\paragraph{Layer Analysis: Correlation Flow and Scale Collapse in GPT and GPT2.}
+\paragraph{Layer Analysis: Correlation Flow in GPT and GPT2.}
 
-See Figure~\ref{fig:gpt-alpha-layers} for the PL exponent $\alpha$ and Log Spectral Norm versus layer id, for GPT and GPT2-small.
+We next examine the distribution of $\alpha$ values in GPT versus GPT2-small.
+Figure~\ref{fig:GPT-alpha-hist} shows the histogram (empirical density), for all layers, of $\alpha$ for GPT and GPT2-small.  
+The older deficient GPT has numerous unusually large $\alpha$ exponents---meaning they are not well-described by a PL fit.
+Indeed, we expect that a poorly-trained model will lack good (i.e., small $\alpha$) PL behavior in many/most layers.
+On the other hand, the newer improved GPT2-small model has, on average, smaller $\alpha$ values than the older GPT, with all $\alpha\le6$ and with smaller mean/median $\alpha$.
+It also has far fewer unusually-large outlying $\alpha$ values than GPT.
+From this (and other results not shown), we see that $\bar{\alpha}$ 
+from Eqn.~(\ref{eqn:alpha_bar}),
+provides a good quality metric for comparing the poorly-trained GPT versus the well-trained GPT2-small.
+This should be contrasted with the behavior displayed by scale-dependent metrics such as the Frobenius norm (not shown) and the Spectral~norm.
+This also reveals why $\hat{\alpha}$ performs unusually in Table~\ref{table:nlp}.
+The PL exponent $\alpha$ behaves as expected, and thus the scale-invariant $\bar{\alpha}$ metric lets us identify potentially poorly-trained models.
+It is the Scale Collapse that causes problems for $\hat{\alpha}$ (recall that the scale enters into $\hat{\alpha}$ via the weights $\log\lambda_{max}$).
+
 Figure~\ref{fig:gpt-alpha-layer} plots $\alpha$ versus the depth (layer id) for each model.
@@
 In contrast, the well-trained GPT2-small model shows consistent and stable patterns, again with one stable $\alpha\sim 3.5$ (and below the GPT trend), and the other only slightly trending up, with $\alpha\le 6$.
-The scale-invariant $\alpha$ metric lets us identify potentially poorly-trained models.
-These results suggest that the Correlation Flow is also a useful concept for NLP models, and that it differs significantly between GPT and GPT2-small, with the better GPT2-small looking more like the better ResNet-1K from Figure~\ref{fig:resnet-alpha-layer}.
-
-These results should be contrasted with the corresponding results for Spectral Norms, shown in Figure~\ref{fig:gpt-snorm-layer}.
-Attention models have two types of layers, one small and large; and the Spectral Norm, in particular, displays unusually small values for some of these layers for GPT.
-This Scale Collapse for the poorly-trained GPT is similar to what we observed for the distilled ResNet20 model in Figure~\ref{fig:resnet204Dalpha}.
-Because of the anomalous scale collapse that is frequently observed in poorly-trained models, these results suggest that scale-dependent norm metrics should not be directly applied to distinguish well-trained versus poorly-trained models.
-
-\begin{figure}[htb]
-    \centering
-    \subfigure[PL exponent ($\alpha$)]{
-        \includegraphics[width=5.5cm]{img/GPT-alpha-depth.png}
-        \label{fig:gpt-alpha-layer}
-    }
-    \quad
-    \subfigure[Log Spectral Norm ($\log\Vert\mathbf{W}\Vert_{\infty}$)]{
-        \includegraphics[width=5.5cm]{img/GPT-snorm-depth.png}
-        \label{fig:gpt-snorm-layer}
-    }
-    \caption{PL exponents 
-             %($\alpha$) 
-             (in (a)) and Log Spectral Norms 
-             %($\log\Vert\mathbf{W}\Vert_{\infty}$) 
-             (in (b)) for weight matrices from the OpenAI GPT and GPT2-small pretrained models.  (Note that the quantities being shown on each Y axis are different.)
-             In the text, this is interpreted in terms of Correlation Flow and Scale~Collapse.
-            }
-    \label{fig:gpt-alpha-layers}
-\end{figure}
+These results show that the behavior of $\alpha$ across layers differs significantly between GPT and GPT2-small, with the better GPT2-small looking more like the better ResNet-1K from Figure~\ref{fig:resnet-alpha-layer}.
+These results also suggest that smaller more stable values of $\alpha$ across depth is beneficial, i.e., that the Correlation Flow is also a useful concept for NLP~models.
 
@@
 
-However, while the means and peaks of the $\alpha$ distributions are getting smaller, towards $2.0$, Figure~\ref{fig:gpt2-alpha-hist} also shows that the tails of the distribution shifts right, with larger GPT2 models having more unusually large $\alpha$ values.
-This is unexpected, and it suggests that these larger GPT2 models are still under-optimized/over-parameterized (relative to the data on which they were trained) and that they have capacity to support datasets even larger than the recent XL $1.5B$ release~\cite{gpt2-xl}.
+While the means and peaks of the $\alpha$ distributions are getting smaller, towards $2.0$, as expected, Figure~\ref{fig:gpt2-alpha-hist} also shows that the tails of the $\alpha$ distributions shift right, with larger GPT2 models having more unusually large $\alpha$ values.
+This is unexpected.
+It suggests that these larger GPT2 models are still under-optimized/over-parameterized (relative to the data on which they were trained) and that they have capacity to support datasets even larger than the recent XL $1.5B$ release~\cite{gpt2-xl}.
 This does not contradict recent theoretical work on the benefits of over-parameterization~\cite{BHMM19}, e.g., since in practice these extremely large models are not fully optimized.

### 2021-01-01 submission/NatCom_overall_approach.tex
got nlp section to local opt

@@
 \bar{\alpha} = \dfrac{1}{L}\sum_l \alpha_l = \langle\alpha\rangle    ,
+\label{eqn:alpha_bar}
 \end{equation}
@@
 Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric $\hat{\alpha}$ for the entire~DNN.
+We will see that for well-trained CV and NLP models, $\hat{\alpha}$ performs quite well and as expected, but for CV and NLP models that are potentially-problematic or less well-trained, metrics that depend on the scale of the problem can perform anomalously.
+In these cases, separating $\hat{\alpha}$ into its two components, $\alpha$ and $\lambda_{max}$, and examining the distributions of each, can be helpful.

### 2021-01-01 submission/NatCom_appendix.tex
tables 7,8,9 autogenerated

@@
 %%\end{figure}
-
-\begin{table}[t]
-\scriptsize
-\begin{center}
-\begin{tabular}{|c|c|c|c|c|c|}
-\hline
-Dataset & Model  & $\langle\log\Vert\cdot\Vert_{F}\rangle$ & $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ & $\hat{\alpha}$ & $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ \\
-\hline
-imagenet-1k & ResNet  & 0.79 &  0.63 & \textbf{0.88} & 0.86 \\
- imagenet-1k & EfficientNet  & 0.65 &  \textbf{0.84} & 0.67 & 0.67 \\
- imagenet-1k & PreResNet  & 0.73 &  0.36 & \textbf{0.85} & 0.85 \\
- imagenet-1k & ShuffleNet  & 0.63 &  0.06 & 0.80 & \textbf{0.81} \\
- imagenet-1k & VGG  & 0.71 &  \textbf{0.76} & 0.35 & 0.46 \\
- imagenet-1k & DLA  & 0.11 &  \textbf{0.65} & 0.40 & 0.36 \\
- imagenet-1k & HRNet  & 0.91 &  0.90 & \textbf{0.97} & 0.97 \\
- imagenet-1k & DRN-C  & 0.81 &  0.79 & \textbf{0.87} & 0.85 \\
- imagenet-1k & SqueezeNext  & 0.05 &  0.07 & 0.42 & \textbf{0.43} \\
- imagenet-1k & ESPNetv2  & 0.42 &  0.38 & \textbf{0.92} & 0.89 \\
- imagenet-1k & ProxylessNAS  & \textbf{0.68} &  0.56 & 0.53 & 0.58 \\
- imagenet-1k & SqueezeNet  & 0.01 &  0.00 & \textbf{0.38} & 0.26 \\
- imagenet-1k & IGCV3  & 0.98 &  0.12 & 0.92 & \textbf{0.99} \\
-\hline
- cifar-10 & ResNet  & 0.58 &  0.59 & \textbf{0.62} & 0.61 \\
- cifar-10 & DIA-ResNet  & \textbf{0.96} &  0.93 & 0.74 & 0.71 \\
- cifar-10 & SENet  & 0.91 &  0.91 & \textbf{0.96} & 0.96 \\
-\hline
- cifar-100 & ResNet  & 0.61 &  0.58 & \textbf{0.71} & 0.71 \\
- cifar-100 & DIA-ResNet  & \textbf{0.96} &  0.85 & 0.90 & 0.89 \\
- cifar-100 & SENet  & 0.97 &  0.96 & \textbf{0.98} & 0.98 \\
- cifar-100 & WRN  & 0.32 &  0.04 & 0.66 & \textbf{0.69} \\
-\hline
- svhn & ResNet  & 0.69 &  0.70 & \textbf{0.82} & 0.81 \\
- svhn & DIA-ResNet  & 0.94 &  \textbf{0.95} & 0.78 & 0.77 \\
- svhn & SENet  & \textbf{0.99} &  0.96 & 0.98 & 0.98 \\
- svhn & WRN  & 0.13 &  0.10 & 0.20 & \textbf{0.21} \\
- svhn & ResNeXt  & 0.87 &  \textbf{0.90} & 0.64 & 0.75 \\
-\hline
- cub-200-2011 & ResNet  & 0.94 &  \textbf{0.95} & 0.08 & 0.08 \\
- cub-200-2011 & SENet  & \textbf{0.66} &  0.59 & 0.41 & 0.38 \\
-\hline
-\end{tabular}
-\end{center}
-\vspace{-5mm}
-\caption{$RMSE$ results for our analysis of all CV models in Table~\ref{table:datasets}. }
-\label{table:MSEresults}
-\end{table}
-
-\begin{table}[t]
-\scriptsize
-\begin{center}
-\begin{tabular}{|c|c|c|c|c|c|}
-\hline
-Dataset & Model  & $\langle\log\Vert\cdot\Vert_{F}\rangle$ & $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ & $\hat{\alpha}$ & $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ \\
-\hline
-imagenet-1k & ResNet  & 0.79 &  0.63 & \textbf{0.88} & 0.86 \\
- imagenet-1k & EfficientNet  & 0.65 &  \textbf{0.84} & 0.67 & 0.67 \\
- imagenet-1k & PreResNet  & 0.73 &  0.36 & \textbf{0.85} & 0.85 \\
- imagenet-1k & ShuffleNet  & 0.63 &  0.06 & 0.80 & \textbf{0.81} \\
- imagenet-1k & VGG  & 0.71 &  \textbf{0.76} & 0.35 & 0.46 \\
- imagenet-1k & DLA  & 0.11 &  \textbf{0.65} & 0.40 & 0.36 \\
- imagenet-1k & HRNet  & 0.91 &  0.90 & \textbf{0.97} & 0.97 \\
- imagenet-1k & DRN-C  & 0.81 &  0.79 & \textbf{0.87} & 0.85 \\
- imagenet-1k & SqueezeNext  & 0.05 &  0.07 & 0.42 & \textbf{0.43} \\
- imagenet-1k & ESPNetv2  & 0.42 &  0.38 & \textbf{0.92} & 0.89 \\
- imagenet-1k & ProxylessNAS  & \textbf{0.68} &  0.56 & 0.53 & 0.58 \\
- imagenet-1k & SqueezeNet  & 0.01 &  0.00 & \textbf{0.38} & 0.26 \\
- imagenet-1k & IGCV3  & 0.98 &  0.12 & 0.92 & \textbf{0.99} \\
-\hline
- cifar-10 & ResNet  & 0.58 &  0.59 & \textbf{0.62} & 0.61 \\
- cifar-10 & DIA-ResNet  & \textbf{0.96} &  0.93 & 0.74 & 0.71 \\
- cifar-10 & SENet  & 0.91 &  0.91 & \textbf{0.96} & 0.96 \\
-\hline
- cifar-100 & ResNet  & 0.61 &  0.58 & \textbf{0.71} & 0.71 \\
- cifar-100 & DIA-ResNet  & \textbf{0.96} &  0.85 & 0.90 & 0.89 \\
- cifar-100 & SENet  & 0.97 &  0.96 & \textbf{0.98} & 0.98 \\
- cifar-100 & WRN  & 0.32 &  0.04 & 0.66 & \textbf{0.69} \\
-\hline
- svhn & ResNet  & 0.69 &  0.70 & \textbf{0.82} & 0.81 \\
- svhn & DIA-ResNet  & 0.94 &  \textbf{0.95} & 0.78 & 0.77 \\
- svhn & SENet  & \textbf{0.99} &  0.96 & 0.98 & 0.98 \\
- svhn & WRN  & 0.13 &  0.10 & 0.20 & \textbf{0.21} \\
- svhn & ResNeXt  & 0.87 &  \textbf{0.90} & 0.64 & 0.75 \\
-\hline
- cub-200-2011 & ResNet  & 0.94 &  \textbf{0.95} & 0.08 & 0.08 \\
- cub-200-2011 & SENet  & \textbf{0.66} &  0.59 & 0.41 & 0.38 \\
-\hline
-\end{tabular}
-\end{center}
-\vspace{-5mm}
-\caption{$R^{2}$ results for our analysis of all CV models in Table~\ref{table:datasets}. }
-\label{table:R2results}
-\end{table}
-
-\begin{table}[t]
-\scriptsize
-\begin{center}
-\begin{tabular}{|c|c|c|c|c|c|}
-\hline
-Dataset & Model  & $\langle\log\Vert\cdot\Vert_{F}\rangle$ & $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ & $\hat{\alpha}$ & $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ \\
-\hline
-imagenet-1k & ResNet  & 0.77 &  0.74 & \textbf{0.88} & 0.88 \\
- imagenet-1k & EfficientNet  & 0.67 &  \textbf{0.79} & 0.66 & 0.66 \\
- imagenet-1k & PreResNet  & 0.65 &  0.54 & \textbf{0.87} & 0.87 \\
- imagenet-1k & ShuffleNet  & 0.39 &  0.09 & \textbf{0.85} & 0.82 \\
- imagenet-1k & VGG  & \textbf{0.82} &  0.76 & 0.52 & 0.61 \\
- imagenet-1k & DLA  & 0.51 &  \textbf{0.82} & 0.78 & 0.69 \\
- imagenet-1k & HRNet  & 0.56 &  0.44 & \textbf{0.61} & 0.61 \\
- imagenet-1k & DRN-C  & \textbf{0.90} &  0.81 & 0.90 & 0.90 \\
- imagenet-1k & SqueezeNext  & 0.47 &  -0.33 & \textbf{0.73} & 0.73 \\
- imagenet-1k & ESPNetv2  & 0.00 &  0.80 & \textbf{1.00} & 1.00 \\
- imagenet-1k & ProxylessNAS  & \textbf{0.33} &  0.33 & 0.33 & 0.33 \\
- imagenet-1k & SqueezeNet  & 0.33 &  0.00 & \textbf{0.67} & 0.67 \\
- imagenet-1k & IGCV3  & \textbf{1.00} &  0.00 & 1.00 & 1.00 \\
-\hline
- cifar-10 & ResNet  & 0.64 &  0.64 & \textbf{0.71} & 0.71 \\
- cifar-10 & DIA-ResNet  & 0.79 &  0.50 & \textbf{0.86} & 0.79 \\
- cifar-10 & SENet  & 0.74 &  \textbf{0.95} & 0.95 & 0.95 \\
-\hline
- cifar-100 & ResNet  & \textbf{0.64} &  0.57 & 0.64 & 0.64 \\
- cifar-100 & DIA-ResNet  & \textbf{0.93} &  0.43 & 0.93 & 0.93 \\
- cifar-100 & SENet  & \textbf{1.00} &  1.00 & 1.00 & 1.00 \\
- cifar-100 & WRN  & \textbf{0.67} &  -0.67 & 0.00 & 0.00 \\
-\hline
- svhn & ResNet  & \textbf{0.81} &  0.71 & 0.81 & 0.81 \\
- svhn & DIA-ResNet  & \textbf{0.86} &  0.86 & 0.57 & 0.57 \\
- svhn & SENet  & \textbf{1.00} &  1.00 & 0.67 & 0.67 \\
- svhn & WRN  & -0.33 &  -0.33 & \textbf{0.67} & 0.67 \\
- svhn & ResNeXt  & \textbf{0.67} &  0.67 & 0.33 & 0.33 \\
-\hline
- cub-200-2011 & ResNet  & \textbf{1.00} &  1.00 & -0.33 & -0.33 \\
- cub-200-2011 & SENet  & \textbf{0.87} &  0.87 & -0.20 & -0.20 \\
-\hline
-\end{tabular}
-\end{center}
-\vspace{-5mm}
-\caption{Kendall-$\tau$ results for our analysis of all CV models in Table~\ref{table:datasets}. }
-\label{table:KTresults}
-\end{table}
+\input{table_RMSE.tex}
+\input{table_R2.tex}
+\input{table_Ktau.tex}
 

### 2021-01-01 submission/NatCom_appendix.tex
autogen table 6

@@
 
-\begin{table}[t]
-\small
-\begin{center}
-\begin{tabular}{|c|c|c|c|c|c|c|}
-\hline
-Architecture & $\#$ of Models & Datasets & & & & \\
-\hline
-  & total &imagenet-1k & cifar-10 & cifar-100 & svhn & cub-200-2011 \\
-\hline
-ResNet & 51 &22 & 8 & 8 & 7 & 6 \\
-EfficientNet & 20 &20 & 0 & 0 & 0 & 0 \\
-PreResNet & 14 &14 & 0 & 0 & 0 & 0 \\
-VGG/BN-VGG & 12 &12 & 0 & 0 & 0 & 0 \\
-ShuffleNet & 12 &12 & 0 & 0 & 0 & 0 \\
-DLA & 10 &10 & 0 & 0 & 0 & 0 \\
-HRNet & 9 &9 & 0 & 0 & 0 & 0 \\
-DRN-C/DRN-D & 7 &7 & 0 & 0 & 0 & 0 \\
-SqueezeNext/SqNxt & 6 &6 & 0 & 0 & 0 & 0 \\
-ESPNetv2 & 5 &5 & 0 & 0 & 0 & 0 \\
-SqueezeNet/SqueezeResNet & 4 &4 & 0 & 0 & 0 & 0 \\
-IGCV3 & 4 &4 & 0 & 0 & 0 & 0 \\
-ProxylessNAS & 4 &4 & 0 & 0 & 0 & 0 \\
-DIA-ResNet/DIA-PreResNet & 24 &0 & 8 & 8 & 8 & 0 \\
-SENet/SE-ResNet/... & 20 &0 & 5 & 5 & 4 & 6 \\
-WRN & 8 &0 & 0 & 4 & 4 & 0 \\
-ResNeXt & 4 &0 & 0 & 0 & 4 & 0 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{Architectures used in total and per dataset \charles{I need to double check this}}
-\label{table:architectures}
-\end{table}
+\input{table_6.tex}
 

### 2021-01-02 submission/figs_largeScale_A.tex
fixed tables refs in figs...

@@
     }
-    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
+    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}.
              Each row corresponds to a different dataset-model pair:

### 2021-01-02 submission/NatCom_all_models.tex
trying to include table 3
table 3 updated

@@
 
-\begin{table}[t]
-\small
-\begin{center}
-\begin{tabular}{|p{1.25in}|c|c|c|c|}
-\hline
-%    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
-%    & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\langle\hat{\alpha}=\alpha\log\lambda_{max}\rangle$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
-                        & $\log\Vert\cdot\Vert_{F}$ & $\log\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
-\hline
-$R^{2}$ (mean)          & 0.63 &  0.55 & \textbf{0.64} & \textbf{0.64} \\
-$R^{2}$ (std)           & 0.34 &  0.36 & \textbf{0.29} &          0.30 \\
-\hline
-$MSE$ (mean)            & 4.54 &  9.62 &          3.14 & \textbf{2.92} \\
-$MSE$ (std)             & 8.69 & 23.06 &          5.14 & \textbf{5.00} \\
-\hline
-Kendall-$\tau$ (mean)   &  XXX &   XXX &           XXX &           XXX \\
-Kendall-$\tau$ (std)    &  XXX &   XXX &           XXX &           XXX \\
-\hline
-\end{tabular}
-\end{center}
-\caption{Comparison of linear regression fits for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering 108 (out of over 400) different pretrained DNNs.
-         We include regressions only for architectures with five or more data points, and which are positively correlated with test error.
-         These results can be readily reproduced using the Google Colab notebooks. 
-         (See the Supplementary Information
-         %~\ref{sxn:appendix} 
-         for details.) 
-        }
-\label{table:results}
-\end{table}
+\input{table_3}
+\charles{Tabel 3 is autogenerated now: I think R2 adjusted may be a better statistic for us}
 

### 2021-01-02 submission/table_3.tex
table 3 updated

\begin{table}[t]
\small
\begin{center}
\begin{tabular}{|p{1.25in}|c|c|c|c|}
\hline
  & $\log\Vert\cdot\Vert^{2}_{F}$ & $\log\Vert\cdot\Vert^{2}_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
\hline

RMSE (mean) & 4.84 & 5.56 & 4.58& 4.54 \\
RMSE (std) & 9.14 & 9.17 & 9.16& 9.17 \\
\hline
R2 (mean) & 3.9 & 3.85 & 3.89& 3.89 \\
R2 (std) & 9.34 & 9.36 & 9.34& 9.34 \\
\hline
R2 adjusted (mean) & 3.82 & 3.77 & 3.82& 3.83 \\
R2 adjusted (std) & 9.38 & 9.4 & 9.37& 9.37 \\
\hline
Kendal-tau (mean) & 3.84 & 3.77 & 3.85& 3.85 \\
Kendal-tau (std) & 9.37 & 9.4 & 9.36& 9.36 \\
\hline

\hline
\end{tabular}
\end{center}
\caption{Comparison of linear regression fits for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering \
108 (out of over 400) different pretrained DNNs.
         We include regressions only for architectures with five or more data points, and which are positively correlated with test error.
         These results can be readily reproduced using the Google Colab notebooks.
         (See the Supplementary Information
         %~\ref{sxn:appendix}                                                                                                                             
         for details.)}
\label{table:results}
\end{table}

### 2021-01-04 submission/NatCom_appendix.tex
updated appendix
uodated figs

@@
 \hline
-& \ref{fig:vgg-metrics}                                 & WeightWatcher-VGG.ipynb \\
-& \ref{fig:resnet-accuracy}                             & WeightWatcher-ResNet.ipynb \\
-& \ref{fig:resnet1k-accuracy}                           & WeightWatcher-ResNet-1K.ipynb \\
-& \ref{fig:vgg-alpha-layers}                            & WeightWatcher-VGG.ipynb \\
-& \ref{fig:resnet-alpha-layer}                          & WeightWatcher-ResNet.ipynb \\
-& \ref{fig:densenet-alpha-layer}                        & WeightWatcher-DenseNet.ipynb \\
+1 & \ref{fig:vgg-metrics}                                 & WeightWatcher-VGG.ipynb \\
+1 & \ref{fig:resnet-accuracy}                             & WeightWatcher-ResNet.ipynb \\
+1 & \ref{fig:resnet1k-accuracy}                           & WeightWatcher-ResNet-1K.ipynb \\
+1 & \ref{fig:vgg-alpha-layers}                            & WeightWatcher-VGG.ipynb \\
+1 & \ref{fig:resnet-alpha-layer}                          & WeightWatcher-ResNet.ipynb \\
+1 & \ref{fig:densenet-alpha-layer}                        & WeightWatcher-DenseNet.ipynb \\
 \hline
@@
 \hline
-& \ref{fig:GPT-hist}                                    & WeightWatcher-OpenAI-GPT.ipynb \\
-& \ref{fig:gpt-alpha-layers}, \ref{fig:gpt2-histograms} & WeightWatcher-OpenAI-GPT2.ipynb \\
+2 & \ref{fig:GPT-hist}                                    & WeightWatcher-OpenAI-GPT.ipynb \\
+2 & \ref{fig:gpt-alpha-layers}, \ref{fig:gpt2-histograms} & WeightWatcher-OpenAI-GPT2.ipynb \\
 \hline
@@
 %
-The final datasets used are shown in Table~\ref{table:datasets}.
+%The final datasets used are shown in Table~\ref{table:datasets}.
 The final architecture series used are shown in  Table~\ref{table:architectures}, with the number of models in each.
@@
 Tables and figures summarizing this analysis (in a more fine-grained way than provided by Table~\ref{table:results}) are presented next.
-
-\begin{table}[t]
-\small
-\begin{center}
-\begin{tabular}{|p{1in}|c|}
-\hline
-Dataset & $\#$ of Models \\
-\hline
-imagenet-1k   &  76 \\
-svhn          &  30 \\
-cifar-100     &  26 \\
-cifar-10      &  18 \\
-cub-200-2011  &  12 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{Datasets used \charles{This table needs updated}}
-\label{table:datasets}
-\end{table}
-
-\input{table_6.tex}
 
@@
 %%\end{figure}
-\input{table_RMSE.tex}
-\input{table_R2.tex}
-\input{table_Ktau.tex}
+\input{table_RMSE}
+\input{table_R2}
+\input{table_Ktau}
 
@@
 \michael{Mention KTau.}
-See Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
+See Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}.
 In the repo, plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.

### 2021-01-04 submission/NatCom_appendix.tex
made change to table in appendix
full pass to revise cv discussion, revise comments, revise response

@@
 %%\end{figure}
+
+\input{table_6}
+
 \input{table_RMSE}
@@
 
-\michael{The text in this section needs to be changed---probably not removed, but instead discuss things, not point to the figure we remove, and point to the repo. }
+\michael{Charles, would you put in a couple of paragraphs about how you generated the tables/figures in the appendix, and a few things we learn from them, i.e., basically enough to convince that reviewer that we paid attention to their concerns.}
 
-\michael{Do we get more visually compelling results if we present figures, but segmented by architectures, as listed in Table~\ref{table:architectures}; if so, perhaps we should put those, so we don't tell R1 that we just removed what looked bad; or maybe actually both, since we segment by dataset in Table~\ref{table:datasets}.
-}
+\michael{Table~\ref{table:architectures} still needs to be fixed, since it does not include the overall results partitioned by dataset that we had in the old Table~\ref{table:datasets}. }
 
@@
 %%Figure~\ref{fig:ARCHalphas}, for models segmented by architecture, as listed in Table~\ref{table:architectures}.
-\michael{Comment here on differences.}
+\michael{Charles, this paragraph is stale since it refers to the old figures that the reviewer did not like, so would you update it.}
 
 To generate the results in Table~\ref{table:results}, we run linear regressions for each architecture series in Table~\ref{table:architectures}, regressing each empirical Log Norm metric against the reported Top1 (and Top5) errors (as listed on the \texttt{osmr/imgclsmob} github repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}).
-We record the $R^{2}$ and $MSE$ for each metric, averaged over all regressions for all architectures and datasets.
-\michael{Mention KTau.}
-See Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}.
+We record the $RMSE$, $R^{2}$, and Kendall-$\tau$ for each metric, averaged over all regressions for all architectures and datasets;
+see Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}, respectively.
+\michael{Charles, the label is broken in the caption of Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}.}
 In the repo, plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.
@@
 Figure~\ref{fig:summary_regressions_H}, and
-Figure~\ref{fig:summary_regressions_A} for more details.
-\michael{Put a bit of discussion, e.g., on best and worst.}
-
-\michael{Probably some discussion of why these metrics are the right ones, e.g., as opposed to $p$ values, to address R1 comments.}
+Figure~\ref{fig:summary_regressions_I} for more details.
+\michael{Charles, would you put a bit of discussion, e.g., on best and worst, again just a teaser to satisfy that reviewer.}
 

### 2021-01-04 submission/NatCom_cv_models.tex
full pass to revise cv discussion, revise comments, revise response

@@
 We examine the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
-Figure~\ref{fig:vgg-metrics} 
-plots the four quality metrics versus reported test accuracies~\cite{pytorch},% 
+%
+Figure~\ref{fig:vgg-metrics} plots the four quality metrics versus reported test accuracies~\cite{pytorch},% 
 \footnote{These test accuracies have been previously reported and made publicly-available by others.
@@
 as well as a basic linear regression line, for the VGG series.
-All four metrics correlate quite well with reported Top1 accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalization (i.e., greater accuracy, lower error). 
-\michael{The following, text and numbers, will need to be updated once we finalize the figures/tables.}
-The Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) performs best (RMSE of $0.42$, $R^2$ of $XXX$, and Kendall-$\tau$ of $XXX$, see Table~\ref{table:cv-models}); and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $) performs second best (RMSE of $0.48$, $R^2$ of $XXX$, and Kendall-$\tau$ of $XXX$, see Table~\ref{table:cv-models}).
-\michael{Put comment on why these three metrics, that HT-SR says that $\hat{\alpha}$ should exhibit good linear correlation, and a few comments on non-normality to address R1 concerns.}
-\charles{The Log Spectral Norm metric correlates very well with small, simpler data sets like the VGG and DenseNet architectures, however, when moving to the larger and more complex ResNet series, the PL-based metrics Log $\alpha$-Norm ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$)
-and  Weighted Alpha ($\hat\alpha =\alpha\log\lambda_{max} $) metrics perform better across all measures.}
+Here, smaller norms and smaller values of $\hat{\alpha}$ imply better generalization (i.e., greater accuracy, lower error). 
+Quantitatively, Log Spectral norm is the best; but, visually, all four metrics correlate quite well with reported Top1 accuracies.
+The DenseNet series has similar behavior.
+(These and many other such plots can be seen on our publicly-available~repo.)
+
+To examine visually how the four quality metrics depend on data set size on a larger, more complex model series, we next look at results on ResNet versus ResNet-1K.
+Figure~\ref{fig:cv2-accuracy} compares the Log $\alpha$-Norm metric for the full ResNet model, trained on the full ImageNet dataset, against the ResNet-1K model, trained on a much smaller ImageNet-1K data set.
+Here, the Log $\alpha$-Norm is much better than the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is slightly worse than the Weighted Alpha metric).
+The ResNet series has strong correlation (RMSE of $0.66$, $R^2$ of $0.9$, and Kendall-$\tau$ of $-1.0$), whereas the ResNet-1K series also shows good but weaker correlation (much larger RMSE of $1.9$, $R^2$ of $0.88$, and Kendall-$\tau$ of $-0.88$).
+%%(Other metrics exhibit similar behavior.)
+%%The higher quality data set shows a better fit, even with fewer data~points.
 
@@
              trained on ImageNet, available in pyTorch (v1.4).  
-             Metrics fit by linear regression, RMSE, R2, and the Kendal-tau rank correlation metric reported.  }
+             Metrics fit by linear regression, RMSE, R2, and the Kendal-tau rank correlation metric reported.  
+            }
     \label{fig:vgg-metrics}
@@
     \caption{Comparison of Average $\alpha$-Norm quality metric versus reported Top1 test accuracy for the ResNet and ResNet-1K pretrained (pyTorch) models. 
-             \michael{FROM V1: Symbols and text are way to small on these and other figures; axes/ticks way too small; need different symbols, not just red versus green; etc.}
-\charles{See changes}
+             Metrics fit by linear regression, RMSE, R2, and the Kendal-tau rank correlation metric reported.  
             }
@@
 \end{figure}
+
+See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics for the VGG, ResNet, ResNet-1K, and DenseNet series.
+Similar results are obtained for the Top5 accuracies.
+The Log Frobenius norm performs well but not extremely well;
+the Log Spectral norm performs very well on smaller, simpler models like the VGG and DenseNet architectures; and, 
+when moving to the larger, more complex ResNet series, the PL-based metrics, Weighted Alpha and the Log $\alpha$-Norm, perform the best.
+Overall, though, these model series are all very well-trodden; and our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of well-trained versus very-well-trained models, with PL-based metrics performing somewhat (i.e., quantitatively) better on the larger, more complex ResNet series.
+
+%%\michael{Put comment on why these three metrics, that HT-SR says that $\hat{\alpha}$ should exhibit good linear correlation, and a few comments on non-normality to address R1 concerns.}
 
@@
 \caption{Quality metrics 
-         (RMSE, smaller is better; $R^2$, XXX is better; and Kendall-$\tau$ rank correlation, XXX is better)
+         (for RMSE, smaller is better; for $R^2$, larger is better; and for Kendall-$\tau$ rank correlation, larger magnitude is better)
          for reported Top1 test error for pretrained models in each architecture series.  
@@
 
-\michael{The following, text and numbers, will need to be updated once we finalize the figures/tables.}
-Overall, all metrics perform relatively well; the Weighted Alpha metric performs best; and the Log $\alpha$-Norm metric performs second best.
-See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics for the VGG, ResNet, ResNet-1K, and DenseNet series.
-Similar results are obtained for the Top5 accuracies.
-These model series are all very well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of well-trained versus very-well-trained models, with PL-based metrics performing somewhat (i.e., quantitatively) better.
-The DenseNet series has similar behavior.
-(These and many other such plots can be seen on our publicly-available~repo.)
-
-To examine how the four quality metrics depend on data set size, we look at results on ResNet versus ResNet-1K.
-Figure~\ref{fig:cv2-accuracy} compares the Log $\alpha$-Norm metric for the full ResNet model, trained on the full ImageNet dataset, against the ResNet-1K model, trained on a much smaller ImageNet-1K data set.
-\michael{The following, text and numbers, will need to be updated once we finalize the figures/tables.}
-The Log $\alpha$-Norm is much better than the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is slightly worse than the Weighted Alpha metric).
-The ResNet series has strong correlation (RMSE of $0.66$, $R^2$ of $XXX$, and Kendall-$\tau$ of $XXX$), whereas the ResNet-1K series also shows good but weaker correlation (much larger RMSE of $1.9$, $R^2$ of $XXX$, and Kendall-$\tau$ of $XXX$).
-(Other metrics exhibit similar behavior.)
-The higher quality data set shows a better fit, even with fewer data~points.
+\michael{Charles, any more sentences to add to the last three paragraphs to clarify anything?  If not, then just comment out this comment.}
 

### 2021-01-04 submission/NatCom_overall_approach.tex
full pass to revise cv discussion, revise comments, revise response

@@
 \begin{equation}
-\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N}\left(\mathcal{L}(E_{DNN}(\mathbf{x}_{i})-y_{i})\right)  ,
+%%%\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N}\left(\mathcal{L}(E_{DNN}(\mathbf{x}_{i})-y_{i})\right)  ,
+\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N} \mathcal{L}(E_{DNN}(\mathbf{x}_{i}),y_{i})   ,
 \end{equation}
-\michael{how precisely to change eqn for R3 comment}
-where the loss function $\mathcal{L}(\cdot)$ can take on a myriad of forms~\cite{JC17_TR}, and where the energy (or optimization) landscape function
+where the loss function $\mathcal{L}(\cdot,\cdot)$ can take on a myriad of forms~\cite{JC17_TR}, and where the energy (or optimization) landscape function
 \begin{equation}
@@
 \end{equation}
-\michael{sentences around here probably need to be changed in light of change to that eqn R3 wanted}
 depends parametrically on the weights and biases.
-\michael{FROM V1: See comments for math equations.}
-For a trained model, $E_{DNN}$ does not explicitly depend on the data, but rather it maps data instance vectors ($\mathbf{x}_i$ values) to predictions ($y_{i}$ labels).
-\michael{FROM V1: Mark found that sentence confusing.}
-Therefore, one can analyze $E_{DNN}$ in the absence of any training or test~data.
+For a trained model, the form of the function $E_{DNN}$ does not explicitly depend on the data (but it does explicitly depend on the weights and biases).
+The function $E_{DNN}$ maps data instance vectors ($\mathbf{x}_i$ values) to predictions ($y_{i}$ labels), and thus the output of this function does depend on the data.
+Therefore, one can analyze the form of $E_{DNN}$ in the absence of any training or test~data.
 

### 2021-01-04 submission/resp_to_revs_NatCom.tex
full pass to revise cv discussion, revise comments, revise response

@@
 One reviewer had questions about additional results to justify our main claims.
-In some cases, we had these results previously, but they were not included due to space constraints.
-In other cases, these results were easy to generate, and we are happy to include them.
+In some cases, we had these results previously, but they were not included due to space constraints; and, in other cases, these results were easy to generate.
+To do so, we have completely redone our analysis, including the requested analysis as well as additional analysis, and we are happy to include these additional results.
 This reviewer also had several misunderstandings, which we will clarify.
-The other reviewer had several general questions, which we also happy to include and clarify.
+The other reviewer had several general questions about how our conclusions fit into the bigger picture that is emerging in the area, and we are also happy to address and clarify these issues.
 We go through each of the points in turn below, with the reviewer comments in italics and our response immediately after.
@@
 It is fair to say that the power law exponents characterize is the implicit regularization, whether or not the model trainer knew that that was what he or she was doing, but not any regularization explicitly used to train.
-We clarify this point at Line XXX of the revised manuscript.
+We have clarified this point at Line XXX in the revised manuscript.
 \end{quote}
@@
 The reviewer is correct.  
-We intended for the reader to examine plots to see that the trends are qualitatively in the correct (or incorrect, in some cases for norm-based metrics) direction, and then to look at the numbers to get quantative insight.
+We intended for the reader to examine plots to see that the trends are qualitatively in the correct (or incorrect, in some cases for norm-based metrics) direction, and then to look at the numbers to get quantitative insight.
 We know that errors are not normal; so, for this and other reasons, in our experience, reporting $p$-values won't clarify the situation.
@@
 (Also, additional results beyond what we have in the main text and Supplementary Material are available in our fully-reproducible and publicly-available repo, and it is our hope that that level of transparency will aid the reader's understanding.)
+XXX.  MM STILL TO REFINE THIS ANSWER.
 \end{quote}
@@
 We clarify these points at Line XXX, Line XXX, and Line XXX of the revised manuscript.
+
+XXX.  MM STILL TO REFINE THIS ANSWER.
 \end{quote}
@@
 We clarify these points in revised text between Line XXX and Line XXX of the revised manuscript.
+
+XXX.  MM STILL TO REFINE THIS ANSWER.
 \end{quote}
@@
 We have put this discussion and clarified these points in the text.
+
+XXX.  MM STILL TO REFINE THIS ANSWER.
 \end{quote}
@@
 We clarify this point at Line XXX of the revised manuscript.
+
+XXX.  MM STILL TO DO THIS IN TEXT MAYBE.
 \end{quote}
@@
 That is a good point.
-We clarify this point at Line XXX of the revised manuscript.
+We have modified this equation and adjusted the surrounding and subsequent text.
 \end{quote}

### 2021-01-09 submission/NatCom_nlp_models.tex
full pass to revise cv discussion, revise comments, revise response
figure 8 updated
GPT2 images ren amed and fixed
XXX filled in

@@
         \includegraphics[width=5.5cm]{img/GPT-snorm-hist.png}
+
         \label{fig:GPT-snorm-hist}
@@
 From Figure~\ref{fig:gpt2-alpha-hist}, we see that
-$\bar{\alpha}$, the average $\alpha$ value, decreases with increasing model size (XXX for GPT2-medium, XXX for GPT2-large, and XXX for GPT2-xl), although the differences are less noticeable between the differing well-trained versus very-well-trained GTP2 models than between the poorly-trained versus well-trained GPT and GPT2-small models.
+$\bar{\alpha}$, the average $\alpha$ value, decreases with increasing model size ($3.82$ for GPT2-medium, $3.97$ for GPT2-large, and $3.81$ for GPT2-xl), although the differences are less noticeable between the differing well-trained versus very-well-trained GTP2 models than between the poorly-trained versus well-trained GPT and GPT2-small models.
 Also, from Figure~\ref{fig:gpt2-pnorm-hist}, we see that, 
-unlike GPT, the layer Log Alpha Norms behave more as expected for GPT2 layers, with the larger models consistently having smaller norms (XXX for GPT2-medium, XXX for GPT2-large, and XXX for GPT2-xl). 
-Similarly, the Log Spectral Norm also decreases on average with the larger models (XXX for GPT2-medium, XXX for GPT2-large, and XXX for GPT2-xl).
-% (not shown).  
+unlike GPT, the layer Log Alpha Norms behave more as expected for GPT2 layers, with the larger models consistently having smaller norms ($9.96$ for GPT2-medium, $7.982$ for GPT2-large, and $7.49$ for GPT2-xl).
+Similarly, the Log Spectral Norm also decreases on average with the larger models ($2.58$ for GPT2-medium, $1.99$ for GPT2-large, and $1.92$ for GPT2-xl).
 As expected, the norm metrics can indeed distinguish among well-trained versus very-well-trained models.
@@
     \centering
-
     \subfigure[PL exponent ($\alpha$)]{
-        \includegraphics[width=5.5cm]{img/GPT2*_fnl_alpha_hist.png}
+        \includegraphics[width=5.5cm]{img/GPT2_fnl_alpha_hist.png}
         \label{fig:gpt2-alpha-hist}
@@
     \subfigure[Log Alpha Norm]{
-        \includegraphics[width=5.5cm]{img/GPT2*_fnl_logpnorm_hist.png}
+        \includegraphics[width=5.5cm]{img/GPT2_fnl_log_alpha_norm_hist.png}
         \label{fig:gpt2-pnorm-hist}

### 2021-01-09 submission/NatCom_appendix.tex
appendix plots updates

@@
-%\section{Reproducibility Appendix}
+>%\section{Reproducibility Appendix}
 \section{Supplementary Information}
@@
 \michael{Charles, would you put a bit of discussion, e.g., on best and worst, again just a teaser to satisfy that reviewer.}
+From these Figures, we recognize fits of varying quality, ranging from remarkably good to completely uncorrelated.
+Starting with the best, consider the Imagenet-1K PreResNet results.  Figure~\ref{fig:summary_regressions_A_10} shows 
+the Avg. Log Spectral Norm $\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$, 
+which has a rather large $RMSE=3.93$, and rather small $R2=0.36$, and Kendal-$\tau=0.54$, and
+has 6 out of 13 points outside the $95\%$ confidence bands.  In contrast, 
+the Avg. Log Alpha Norm $\langle\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}\rangle$, 
+in  Figure~\ref{fig:summary_regressions_A_12} , 
+has a much smaller $RMSE=1.93$, and much larger  $R2=0.85$, and Kendal-$\tau=0.87$, 
+and only 2 points outside the $95\%$ confidence bands.
+For examples of lower quality fits, consider the SqueezeNext results, as shown Figures~\ref{fig:summary_regressions_C_10} and 
+~\ref{fig:summary_regressions_C_12}.  
+The Avg. Log Spectral Norm $\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$,
+is visually anti-correlated  with the test accuracies
+(as it is with ShuffleNet, in Figure ~\ref{fig:summary_regressions_B_02}).
+with a very large $95\%$ confidence band, with only 2 points close to the regression line,
+a large RMSE,  $R2=0.07$ (i.e zero), and small Kendal-$\tau=0.33$.
+The Avg. Log Alpha Norm $\langle\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}\rangle$ is always positively-correlated
+with test accuracies, but with $R2=43$, which is not great--but not zero, and a reasonable Kendal-$\tau=0.73$.
 

### 2021-01-09 submission/appendixB.tex
udpated appendix

@@
 
-The original work on HT-SR Theory~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM} considered NNs including AlexNet and InceptionV3 (as well as DenseNet, ResNet, and VGG), and it showed that for nearly every $\mathbf{W}$, the (bulk and tail) of the ESDs can be fit to a truncated PL and the PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$.
+The original work on HT-SR Theory~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM} considered the Linear weight matrics of AlexNet and InceptionV3 (as well as DenseNet, ResNet, and VGG), and it showed that for nearly every $\mathbf{W}$ examined, the (bulk and tail) of the ESDs can be fit to a truncated PL and the PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$.
 Moreover, smaller exponents $\alpha$ are correlated with more implicit self-regularization and, correspondingly, better generalization~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.

### 2021-01-09 submission/NatCom_appendix.tex
appendix

@@
 
-The original work on HT-SR Theory~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM} considered NNs including AlexNet and InceptionV3 (as well as DenseNet, ResNet, and VGG), and it showed that for nearly every $\mathbf{W}$, the (bulk and tail) of the ESDs can be fit to a truncated PL and the PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$.
+The original work on HT-SR Theory~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM} considered DNNs including AlexNet and InceptionV3 (as well as DenseNet, ResNet, and VGG), and it showed that for nearly every $\mathbf{W}$, the (bulk and tail) of the ESDs can be fit to a truncated PL and the PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$.
 Our meta-analysis, the main results of which are summarized in this paper, has shown that these results are ubiquitous.

### 2021-01-09 submission/NatCom_appendix.tex
minor update table 3

@@
 \michael{Charles, would you put in a couple of paragraphs about how you generated the tables/figures in the appendix, and a few things we learn from them, i.e., basically enough to convince that reviewer that we paid attention to their concerns.}
-
+\charles{fixed below}
 \michael{Table~\ref{table:architectures} still needs to be fixed, since it does not include the overall results partitioned by dataset that we had in the old Table~\ref{table:datasets}. }
 
-To explain further how to reproduce our analysis, we run three batches of linear regressions. 
-First, at the global level, we divide models by datasets and run regressions separately on all models of a certain dataset, regardless of the architecture. 
-At this level, the plots are quite noisy and clustered, as each architecture has its own accuracy trend; but one can
- still see that most plots show positive relationship with positive coefficients. 
+%To explain further how to reproduce our analysis, we run three batches of linear regressions. 
+%First, at the global level, we divide models by datasets and run regressions separately on all models of a certain dataset, regardless of the architecture. 
+%At this level, the plots are quite noisy and clustered, as each architecture has its own accuracy trend; but one can
+% still see that most plots show positive relationship with positive coefficients. 
 % V1 REMOVED FOR V2 % Example regressions are shown in Figure~\ref{fig:DSalphas}, as available in the results notebook.
@@
 %%Figure~\ref{fig:ARCHalphas}, for models segmented by architecture, as listed in Table~\ref{table:architectures}.
-\michael{Charles, this paragraph is stale since it refers to the old figures that the reviewer did not like, so would you update it.}
+%\michael{Charles, this paragraph is stale since it refers to the old figures that the reviewer did not like, so would you update it.}
 
-To generate the results in Table~\ref{table:results}, we run linear regressions for each architecture series in Table~\ref{table:architectures}, regressing each empirical Log Norm metric against the reported Top1 (and Top5) errors (as listed on the \texttt{osmr/imgclsmob} github repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}).
-We record the $RMSE$, $R^{2}$, and Kendall-$\tau$ for each metric, averaged over all regressions for all architectures and datasets;
-see Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}, respectively.
-\michael{Charles, the label is broken in the caption of Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}.}
-In the repo, plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.
-The final analysis includes 108 regressions in all, those with 4 or more models, with a positive $R^2$.
+Table~\ref{table:results} and Figures~\ref{fig:summary_regressions_A}-\ref{fig:summary_regressions_I} below present results for OLS 
+(Ordinary Least Squares) regressions for every dataset and architecture series listed in Table~\ref{table:architectures},
+ created by running the \texttt{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}
+on numerous pretrained models taken from the \texttt{OSMR/imgclsmob Sandbox} github repository of pretrained CV DNN model
+ggithub repository. \michael{citation here ?}
+In each Figure,  each row of subfigures, for a given pretrained model and dataset, depicts the average Norm-based and Power Law metrics
+ ($\langle\log\Vert\cdot\Vert_{F}\rangle$,    $\langle\log\Vert\cdot\Vert_{\infty}\rangle$,
+$\hat{\alpha}$,  $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$)
+ against the Top1 Test Accuracy, as reported in the github repository README file~\cite{osmr} ).
+, along with a shaded area
+representing the $95\%$ confidence bound.
+For each regression, we report the RMSE, R2 regresssion metrics, and Kendal-$\tau$ rank correlation metric,
+both in the title of each subfiguire, and in Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}, respectively.
+To reproduce these Figures and Tables, see the \texttt{OSMR-Analysis.ipynb} python Jupyter Notebook, as listed in Table~\ref{},
+and provided github repo accompanying this paper.
+The reader may regenerate these Figures and Tables,
+along with more fine grained results, by rerunning
+ the \texttt{OSMR-Analysis.ipynb} python Jupyter Notebook (see Table~\ref{table:notebooks}), which 
+analyzes the data in the \texttt{df\_all.xlsx} file. 
+This repository also contains the original \texttt{Google Colab} notebooks, run in January 2020, 
+which downloads the pretrained models and  and runs the \texttt{WeightWatcher} tool on them.
+%To generate the results in Table~\ref{table:results}, we run linear regressions for each architecture series in Table~\ref{table:architectures}, regressing each empirical Log Norm metric against the reported Top1 rrors (as listed on the \texttt{osmr/imgclsmob} github repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}).
+%We record the $RMSE$, $R^{2}$, and Kendall-$\tau$ for each metric, averaged over all regressions for all architectures and datasets;
+%see Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}, respectively.
+%captions fixed in table Ktau.  
+%In the repo, plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.
+The reader may also run the most current version of  \texttt{WeightWatcher} (ww0.4.1) on individual 
+pretrained models (which may also have changed slightly since January 2020), 
+such as the ResNet models, trained on the ImageNet-1K dataset, using the \texttt{WeightWatcher-ResNet-1K.ipynb} notebook.
+The final analysis includes 108 regressions in all.%, those with 4 or more models.%, with a positive $R^2$.
 
@@
 \michael{Charles, would you put a bit of discussion, e.g., on best and worst, again just a teaser to satisfy that reviewer.}
+\charles{below}
 From these Figures, we recognize fits of varying quality, ranging from remarkably good to completely uncorrelated.

### 2021-01-09 submission/table_3.tex
fixed ResNet-1K data
minor update table 3

@@
 
-RMSE (mean) & 4.84 & 5.56 & 4.58& 4.54 \\
-RMSE (std) & 9.14 & 9.17 & 9.16& 9.17 \\
+RMSE (mean) & 4.84 & 5.57 & 4.58& 4.55 \\
+RMSE (std) & 9.14 & 9.16 & 9.16& 9.17 \\
 \hline
@@
 \hline
-R2 adjusted (mean) & 3.82 & 3.77 & 3.82& 3.83 \\
-R2 adjusted (std) & 9.38 & 9.4 & 9.37& 9.37 \\
-\hline
-Kendal-tau (mean) & 3.84 & 3.77 & 3.85& 3.85 \\
+Kendal-tau (mean) & 3.84 & 3.77 & 3.86& 3.85 \\
 Kendal-tau (std) & 9.37 & 9.4 & 9.36& 9.36 \\
@@
          We include regressions only for architectures with five or more data points, and which are positively correlated with test error.
-         These results can be readily reproduced using the Google Colab notebooks.
+         These results can be readily reproduced using the Jupyter notebooks and data provided 
          (See the Supplementary Information

### 2021-01-09 submission/NatCom_appendix.tex
appendix updated

@@
 The ImageNet and OpenAI GPT pretrained models are provided in the current 
-pyTorch~\cite{pytorch} and Huggingface~\cite{huggingface} distributions, as specified in the \texttt{requirements.txt}~file.
+pyTorch~\cite{pytorch} and Huggingface~\cite{huggingface} distributions.
 
@@
 
-In the \texttt{ww-colab} folder of our github repo, we provide several Google Colab notebooks which can be used to reproduce the results of Section~\ref{sxn:all_cv_models}.
+In our github repository, we provide several python Juoyter Google Colab anotebooks,
+as well as previously generated data,  which can be used to reproduce the results of Section~\ref{sxn:all_cv_models}, 
+and using the \texttt{WeightWatcher} tool (version 0.2.7), as described in Table~\ref{table:notebooks}.
 The ImageNet-1K and other pretrained models are taken from the pyTorch models in the \texttt{omsr/imgclsmob} 
 ``Sandbox for training convolutional networks for computer vision'' github repository~\cite{osmr}.
-The data for each regression can be generated in parallel by running each Google Colab notebook (i.e., \texttt{ww\_colab\_0\_100.ipynb}) simultaneously on the same account.
-The data generated are analyzed with \texttt{ww\_colab\_results.ipynb}, which runs all regressions and which tabulates the results presented in Table~\ref{table:results}.
+The full  \texttt{WeightWatcher} results are provided in the datasets: \texttt{[data/osmr/data...xlsx]},
+last generated in January 2020, using the Google Colab notebooks: \texttt{[ww\_colab/ww\_colab...ipynb]}.
+These data files are analyzed with the \texttt{OSMR-Analysis.ipynb} python Juyter Notebook,
+which runs all regressions and which tabulates the results presented in Table~\ref{table:results}
+and generates the figures in Figures~\ref{fig:summary_regressions_A}-\ref{fig:summary_regressions_I}.
 

### 2021-01-09 submission/NatCom_appendix.tex
appendix updated

@@
 
-In our github repository, we provide several python Juoyter Google Colab anotebooks,
-as well as previously generated data,  which can be used to reproduce the results of Section~\ref{sxn:all_cv_models}, 
-and using the \texttt{WeightWatcher} tool (version 0.2.7), as described in Table~\ref{table:notebooks}.
+The results of Section~\ref{sxn:all_cv_models} can be regenerated using
+using the \texttt{WeightWatcher} tool (version 0.2.7), as described in Table~\ref{table:notebooks};
+the reader may regenerate all the \texttt{WeightWatcher} results using the Google Colab notebooks (in \texttt{ww-colab}),
+and/or simply reproduce the Tables and Figures using the Jupyter notebooks (in Table~ref{table:notebooks},
+and the pre-computed \texttt{WeightWatcher} datasets (in the \texttt{data/osmr} folder).
 The ImageNet-1K and other pretrained models are taken from the pyTorch models in the \texttt{omsr/imgclsmob} 

### 2021-01-09 submission/NatCom_appendix.tex
appendix updated

--- 2021-01-09 submission/NatCom_appendix.tex
@@
 
-In our github repository, we provide several python Juoyter Google Colab anotebooks,
-as well as previously generated data,  which can be used to reproduce the results of Section~\ref{sxn:all_cv_models}, 
-and using the \texttt{WeightWatcher} tool (version 0.2.7), as described in Table~\ref{table:notebooks}.
+The reader may regenerat all of the  results of Section~\ref{sxn:all_cv_models} 
+\texttt{WeightWatcher} results using the Google Colab Jupyter notebooks (in the \texttt{ww-colab} folder)
+and  the \texttt{WeightWatcher} tool, 
+and/or simply reproduce the Tables and Figures using the Jupyter notebooks (shown in Table~\ref{table:notebooks}),
+and the pre-computed \texttt{WeightWatcher} datasets (in the \texttt{data/osmr} folder).
 The ImageNet-1K and other pretrained models are taken from the pyTorch models in the \texttt{omsr/imgclsmob} 
@@
 We attempt to run linear regressions for all pyTorch models for each architecture series for all datasets provided.  
-There are over $450$ models in all, and we note that the \texttt{osmr/imgclsmob} repository is constantly being updated with new models.
+There are over $450$ models in all to consider, and we note that the \texttt{osmr/imgclsmob} repository is constantly being updated with new models.
 We omit the results for CUB-200-2011, Pascal-VOC2012, ADE20K, and COCO datasets, as there are fewer than 15 models for those datasets.  
@@
 See
-Figure~\ref{fig:summary_regressions_A},
-Figure~\ref{fig:summary_regressions_B},
-Figure~\ref{fig:summary_regressions_C},
-Figure~\ref{fig:summary_regressions_D},
-Figure~\ref{fig:summary_regressions_E},
-Figure~\ref{fig:summary_regressions_F},
-Figure~\ref{fig:summary_regressions_G},
-Figure~\ref{fig:summary_regressions_H}, and
-Figure~\ref{fig:summary_regressions_I} for more details.
+Figure~\ref{fig:summary_regressions_A}-\ref{fig:summary_regressions_I} for more details.
+%5Figure~\ref{fig:summary_regressions_B},
+%Figure~\ref{fig:summary_regressions_C},
+%Figure~\ref{fig:summary_regressions_D},
+%Figure~\ref{fig:summary_regressions_E},
+%Figure~\ref{fig:summary_regressions_F},
+%Figure~\ref{fig:summary_regressions_G},
+%Figure~\ref{fig:summary_regressions_H}, and
+
 \michael{Charles, would you put a bit of discussion, e.g., on best and worst, again just a teaser to satisfy that reviewer.}
@@
 a large RMSE,  $R2=0.07$ (i.e zero), and small Kendal-$\tau=0.33$.
-The Avg. Log Alpha Norm $\langle\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}\rangle$ is always positively-correlated
+The Avg. Log Alpha Norm $\langle\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}\rangle$ is (as always) positively-correlated
 with test accuracies, but with $R2=43$, which is not great--but not zero, and a reasonable Kendal-$\tau=0.73$.

### 2021-01-09 submission/NatCom_appendix.tex
appendix updated

@@
 We provide a github repository for this paper that includes Jupyter notebooks that fully reproduce all results (as well as many other results)~\cite{kdd20_sub_repo}.
-All results have been produced using the \texttt{WeightWatcher} tool (v0.2.7)~\cite{weightwatcher_package}.
+All results have been produced using the \texttt{WeightWatcher} tool)~\cite{weightwatcher_package}.
 The ImageNet and OpenAI GPT pretrained models are provided in the current 
@@
 
-The reader may regenerat all of the  results of Section~\ref{sxn:all_cv_models} 
+The reader may regenerate all of the  results of Section~\ref{sxn:all_cv_models} 
 \texttt{WeightWatcher} results using the Google Colab Jupyter notebooks (in the \texttt{ww-colab} folder)
 and  the \texttt{WeightWatcher} tool, 
-and/or simply reproduce the Tables and Figures using the Jupyter notebooks (shown in Table~\ref{table:notebooks}),
+and/or simply reproduce 
+Table~\ref{table:results}, as well as Tables~\ref{table:RMSEresults}-\ref{table:Ktauresults}
+and  Figures~\ref{fig:summary_regressions_A}-\ref{fig:summary_regressions_I},
+using the Jupyter notebooks (shown in Table~\ref{table:notebooks}),
 and the pre-computed \texttt{WeightWatcher} datasets (in the \texttt{data/osmr} folder).
-The ImageNet-1K and other pretrained models are taken from the pyTorch models in the \texttt{omsr/imgclsmob} 
+The pretrained models, trained on ImageNet-1K and the other datasets, 
+are taken from the pyTorch models in the \texttt{omsr/imgclsmob} 
 ``Sandbox for training convolutional networks for computer vision'' github repository~\cite{osmr}.
 The full  \texttt{WeightWatcher} results are provided in the datasets: \texttt{[data/osmr/data...xlsx]},
-last generated in January 2020, using the Google Colab notebooks: \texttt{[ww\_colab/ww\_colab...ipynb]}.
-These data files are analyzed with the \texttt{OSMR-Analysis.ipynb} python Juyter Notebook,
-which runs all regressions and which tabulates the results presented in Table~\ref{table:results}
-and generates the figures in Figures~\ref{fig:summary_regressions_A}-\ref{fig:summary_regressions_I}.
+last generated in January 2020, using the Google Colab notebooks: \texttt{[ww\_colab/ww\_colab...ipynb]},
+and \texttt{WeightWatcher} version ww0.2.7.  Results can be recomputed using the current version (ww0.4.1)
+using the \texttt{ww2x=True} back compatability option, although note the pretrained models must
+be downloaded and  may have changed slightly.
+The data files currently provided are analyzed with the \texttt{OSMR-Analysis.ipynb} python Juyter Notebook,
+which runs all regressions and which tabulates the results presented in 
+Table~\ref{table:results}
+and generates the figures in Figures~\ref{fig:summary_regressions_A}-\ref{fig:summary_regressions_I}
+and Tables
 
@@
 $\hat{\alpha}$,  $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$)
- against the Top1 Test Accuracy, as reported in the github repository README file~\cite{osmr} ).
-, along with a shaded area
-representing the $95\%$ confidence bound.
+ against the Top1 Test Accuracy, as reported in the github repository README file~\cite{osmr} ),
+along with a shaded area representing the $95\%$ confidence bound.
 For each regression, we report the RMSE, R2 regresssion metrics, and Kendal-$\tau$ rank correlation metric,
@@
 The reader may regenerate these Figures and Tables,
-along with more fine grained results, by rerunning
+and even more fine grained results, by rerunning
  the \texttt{OSMR-Analysis.ipynb} python Jupyter Notebook (see Table~\ref{table:notebooks}), which 
-analyzes the data in the \texttt{df\_all.xlsx} file. 
+analyzes the precomputed data in the \texttt{df\_all.xlsx} file. 
 This repository also contains the original \texttt{Google Colab} notebooks, run in January 2020, 
@@
 %In the repo, plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.
-The reader may also run the most current version of  \texttt{WeightWatcher} (ww0.4.1) on individual 
-pretrained models (which may also have changed slightly since January 2020), 
+The reader may also run the  \texttt{WeightWatcher} locally on each of the 
+pretrained models,
 such as the ResNet models, trained on the ImageNet-1K dataset, using the \texttt{WeightWatcher-ResNet-1K.ipynb} notebook.
+(However, note that the models may have changed, giving slightly different results).
 The final analysis includes 108 regressions in all.%, those with 4 or more models.%, with a positive $R^2$.
-
-See
-Figure~\ref{fig:summary_regressions_A}-\ref{fig:summary_regressions_I} for more details.
+See Figure~\ref{fig:summary_regressions_A}-\ref{fig:summary_regressions_I} for more details.
 %5Figure~\ref{fig:summary_regressions_B},
@@
 ~\ref{fig:summary_regressions_C_12}.  
-The Avg. Log Spectral Norm $\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$,
-is visually anti-correlated  with the test accuracies
+The Avg. Log Spectral Norm $\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$ appears
+visually anti-correlated with the test accuracies
 (as it is with ShuffleNet, in Figure ~\ref{fig:summary_regressions_B_02}).
-with a very large $95\%$ confidence band, with only 2 points close to the regression line,
+It has a very large $95\%$ confidence band, with only 2 points close to the regression line, 
 a large RMSE,  $R2=0.07$ (i.e zero), and small Kendal-$\tau=0.33$.  
 The Avg. Log Alpha Norm $\langle\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}\rangle$ is (as always) positively-correlated
-with test accuracies, but with $R2=43$, which is not great--but not zero, and a reasonable Kendal-$\tau=0.73$.
+with test accuracies, but with $R2=0.43$, show some linear correaltion, and a reasonable Kendal-$\tau=0.73$,
+showing moderately strong rank correlation.
 

### 2021-01-10 submission/NatCom_cv_models.tex
gpt images added
CV updated

@@
 
-%%\michael{Put comment on why these three metrics, that HT-SR says that $\hat{\alpha}$ should exhibit good linear correlation, and a few comments on non-normality to address R1 concerns.}
-
 %V1VERSION% \begin{table}[t]
@@
 \michael{Charles, any more sentences to add to the last three paragraphs to clarify anything?  If not, then just comment out this comment.}
+\charles{In partcular, the PL-based AlphaHat $\hat{\alpha}$ and Shatten-Norm $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ metrics 
+tend to perform better when there is a wider variation in the hyper-parameters going beyond just increasing the depth.  In fact sometimes the
+purely norm-based metrics such as the Spectral Norm can be uncorrelated or even
+anti-correlated with the test accuracy, as seen in Section~\ref{} with
+ShuffleNet in Figure~\ref{fig:summary_regressions_B_02},
+SqueezeNext in Figure~\ref{fig:summary_regressions_C_10},
+and WRN in Figure~\ref{fig:summary_regressions_G_06}).
+}
 

### 2021-01-10 submission/NatCom_cv_models.tex
CV updated

@@
 purely norm-based metrics such as the Spectral Norm can be uncorrelated or even
-anti-correlated with the test accuracy, as seen in Section~\ref{} with
+anti-correlated with the test accuracy, as seen in Section~\ref{sxn:appendix} with
 ShuffleNet in Figure~\ref{fig:summary_regressions_B_02},

### 2021-01-11 submission/NatCom_cv_models.tex
updated tablesd

--- 2021-01-10 submission/NatCom_cv_models.tex
@@
 purely norm-based metrics such as the Spectral Norm can be uncorrelated or even
-anti-correlated with the test accuracy, as seen in Section~\ref{} with
+anti-correlated with the test accuracy, where the PL-metrrics are positively-correlated.
+This is seen in Section~\ref{sxn:appendix} with
 ShuffleNet in Figure~\ref{fig:summary_regressions_B_02},
 SqueezeNext in Figure~\ref{fig:summary_regressions_C_10},
-and WRN in Figure~\ref{fig:summary_regressions_G_06}).
-}
+and WRN in Figure~\ref{fig:summary_regressions_G_06}).}
 

### 2021-01-11 submission/NatCom_all_models.tex
full pass to revise cv discussion, revise comments, revise response
smooth out a bit

@@
 
-\michael{Refine text in this subsection after we get the final version of all the tables and figures in the appendix.
+%%\michael{Refine text in this subsection after the appendix is finished.
 Things to emphasize: we have a huge amount of data that we are summarizing very succicently; the table here is the tip of the iceberg; the results show our metrics are good, as a sanity check; they show that you can get some results in general, but looking at the details in the appendix show that results are more reliable within a given architecture series than across different architectures; point to out publicly-availab e repo; mention that we do Kendall-$\tau$ rank correlation also.}
@@
 for details.
-To quantify the relationship between quality metrics and the reported test error and/or accuracy metrics, we use ordinary least squares (similar results are seen with rank correlation metrics) to regress the metrics on the Top1 (and Top5) reported errors (as dependent variables).
+To quantify the relationship between quality metrics and the reported test error and/or accuracy metrics, we use ordinary least squares 
+%(similar results are seen with rank correlation metrics) 
+to regress the metrics on the Top1 (and Top5) reported errors (as dependent variables), and we report the RMSE, the $R^2$ (R2) regresssion metric, and the Kendal-$\tau$ rank correlation metric.
 These include Top5 errors for the ImageNet-1K model, percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy (Pix.Acc.) and Intersection-Over-Union (IOU) for other models.
@@
 \input{table_3}
-\charles{Tabel 3 is autogenerated now: I think R2 adjusted may be a better statistic for us}
 
-Results are summarized in Table~\ref{table:results}.
-For the mean, larger $R^{2}$ and smaller $MSE$ are desirable; and for the standard deviation, smaller values are desirable.
+Results are summarized in Table~\ref{table:results} (and Figures~\ref{fig:summary_regressions_A}--\ref{fig:summary_regressions_I} of the Supplementary Material).
+For the mean, 
+smaller RMSE, 
+larger $R^2$, and 
+larger Kendal-$\tau$ 
+are desirable; 
+and, for the standard deviation, smaller values are desirable.
 Taken as a whole, over the entire corpus of data, PL-based metrics are somewhat better for both the $R^{2}$ mean and standard deviation;
-and PL-based metrics are much better for $MSE$ mean and standard deviation.
-These and other results suggest our conclusions hold much more generally.
+and PL-based metrics are much better for RMSE mean and standard deviation.
+Model diagnostics (Supplementary Material) indicate many outliers and imperfect fits.
+Overall, though, these and other results suggest our conclusions hold much more generally.

### 2021-01-11 submission/NatCom_appendix.tex
table 5 updated with totals per dataset
smooth out a bit

@@
->%\section{Reproducibility Appendix}
+%\section{Reproducibility Appendix}
 \section{Supplementary Information}
@@
 
-\input{table_6}
+\input{table_arch}
 
@@
 \subsection{Supplementary Tables and Figures}
-
-\michael{Charles, would you put in a couple of paragraphs about how you generated the tables/figures in the appendix, and a few things we learn from them, i.e., basically enough to convince that reviewer that we paid attention to their concerns.}
-\charles{fixed below}
-\michael{Table~\ref{table:architectures} still needs to be fixed, since it does not include the overall results partitioned by dataset that we had in the old Table~\ref{table:datasets}. }
 
@@
 
-Table~\ref{table:results} and Figures~\ref{fig:summary_regressions_A}-\ref{fig:summary_regressions_I} below present results for OLS 
-(Ordinary Least Squares) regressions for every dataset and architecture series listed in Table~\ref{table:architectures},
- created by running the \texttt{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}
-on numerous pretrained models taken from the \texttt{OSMR/imgclsmob Sandbox} github repository of pretrained CV DNN model
-ggithub repository. \michael{citation here ?}
-In each Figure,  each row of subfigures, for a given pretrained model and dataset, depicts the average Norm-based and Power Law metrics
- ($\langle\log\Vert\cdot\Vert_{F}\rangle$,    $\langle\log\Vert\cdot\Vert_{\infty}\rangle$,
-$\hat{\alpha}$,  $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$)
- against the Top1 Test Accuracy, as reported in the github repository README file~\cite{osmr} ),
-along with a shaded area representing the $95\%$ confidence bound.
-For each regression, we report the RMSE, R2 regresssion metrics, and Kendal-$\tau$ rank correlation metric,
-both in the title of each subfiguire, and in Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}, respectively.
-To reproduce these Figures and Tables, see the \texttt{OSMR-Analysis.ipynb} python Jupyter Notebook, as listed in Table~\ref{},
-and provided github repo accompanying this paper.
-The reader may regenerate these Figures and Tables,
-and even more fine grained results, by rerunning
- the \texttt{OSMR-Analysis.ipynb} python Jupyter Notebook (see Table~\ref{table:notebooks}), which 
-analyzes the precomputed data in the \texttt{df\_all.xlsx} file. 
-This repository also contains the original \texttt{Google Colab} notebooks, run in January 2020, 
-which downloads the pretrained models and  and runs the \texttt{WeightWatcher} tool on them.
+Here, we present a more detailed discussion of our large-scale analysis of hundreds of models, which were summarized in Table~\ref{table:results} of Section~\ref{sxn:all_cv_models}. 
+We ran the \texttt{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}
+on numerous pretrained models taken from the \texttt{OSMR/imgclsmob Sandbox} github repository of pretrained CV DNN models~\cite{osmr},   
+performing OLS (Ordinary Least Squares) regressions for every dataset and architecture series listed in Table~\ref{table:architectures}.
+Table~\ref{table:results} summarized the overall results, and Figures~\ref{fig:summary_regressions_A}--\ref{fig:summary_regressions_I} below present a more detailed visual summary, which enables model diagnostics.
+
+For each Figure, each row of subfigures considers a given pretrained model and dataset, depicting the average Norm-based and Power Law 
+metrics---Log Frobenius norm
+          ($\langle\log\Vert\mathbf{W}\Vert^{2}_{F}\rangle$),
+          Log Spectral norm
+          ($\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$),
+          Weighted Alpha
+          ($\hat{\alpha}$),
+          and Log $\alpha$-Norm
+          ($\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$)---against 
+the Top1 Test Accuracy, as reported in the github repository README file~\cite{osmr}, along with a shaded area representing the $95\%$ confidence bound.
+For each regression, we report the RMSE, the R2 regresssion metric, and the Kendal-$\tau$ rank correlation metric in the title of each subfigure.
+We also present these same numerical values in Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}, respectively.
+To reproduce these Figures and Tables, see the \texttt{OSMR-Analysis.ipynb} python Jupyter Notebook, as listed in Table~\ref{table:notebooks}.
+(These are provided in the github repo accompanying this paper.)
+The reader may regenerate these Figures and Tables, as well as more fine grained results, by rerunning the \texttt{OSMR-Analysis.ipynb} python Jupyter Notebook (see Table~\ref{table:notebooks}), which analyzes the precomputed data in the \texttt{df\_all.xlsx} file.
+This repository also contains the original \texttt{Google Colab} notebooks, run in January 2020, which download the pretrained models and run the \texttt{WeightWatcher} tool on them.
 %To generate the results in Table~\ref{table:results}, we run linear regressions for each architecture series in Table~\ref{table:architectures}, regressing each empirical Log Norm metric against the reported Top1 rrors (as listed on the \texttt{osmr/imgclsmob} github repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}).
@@
 such as the ResNet models, trained on the ImageNet-1K dataset, using the \texttt{WeightWatcher-ResNet-1K.ipynb} notebook.
-(However, note that the models may have changed, giving slightly different results).
-The final analysis includes 108 regressions in all.%, those with 4 or more models.%, with a positive $R^2$.
-See Figure~\ref{fig:summary_regressions_A}-\ref{fig:summary_regressions_I} for more details.
+(We should note, however, that the publicly-available versions of these models may have changed slightly, giving slightly different results).
+Our final analysis includes 108 regressions in all. %, those with 4 or more models.%, with a positive $R^2$.
+See Figure~\ref{fig:summary_regressions_A}--\ref{fig:summary_regressions_I} for more details.
 %5Figure~\ref{fig:summary_regressions_B},
@@
 
-\michael{Charles, would you put a bit of discussion, e.g., on best and worst, again just a teaser to satisfy that reviewer.}
-\charles{below}
+% Log Frobenius norm
+% Log Spectral norm
+% Weighted Alpha
+% Log $\alpha$-Norm
+
 From these Figures, we recognize fits of varying quality, ranging from remarkably good to completely uncorrelated.
-Starting with the best, consider the Imagenet-1K PreResNet results.  Figure~\ref{fig:summary_regressions_A_10} shows 
-the Avg. Log Spectral Norm $\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$, 
+Starting with some of the best, consider the Imagenet-1K PreResNet results.  
+For example,
+Figure~\ref{fig:summary_regressions_A_10} shows the 
+Log Spectral norm, %Avg. Log Spectral Norm $\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$, 
 which has a rather large $RMSE=3.93$, and rather small $R2=0.36$, and Kendal-$\tau=0.54$, and 
-has 6 out of 13 points outside the $95\%$ confidence bands.  In contrast, 
-the Avg. Log Alpha Norm $\langle\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}\rangle$, 
-in  Figure~\ref{fig:summary_regressions_A_12} , 
-has a much smaller $RMSE=1.93$, and much larger  $R2=0.85$, and Kendal-$\tau=0.87$, 
-and only 2 points outside the $95\%$ confidence bands.
-For examples of lower quality fits, consider the SqueezeNext results, as shown Figures~\ref{fig:summary_regressions_C_10} and 
-~\ref{fig:summary_regressions_C_12}.  
-The Avg. Log Spectral Norm $\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$ appears
-visually anti-correlated with the test accuracies
-(as it is with ShuffleNet, in Figure ~\ref{fig:summary_regressions_B_02}).
-It has a very large $95\%$ confidence band, with only 2 points close to the regression line, 
-a large RMSE,  $R2=0.07$ (i.e zero), and small Kendal-$\tau=0.33$.  
-The Avg. Log Alpha Norm $\langle\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}\rangle$ is (as always) positively-correlated
-with test accuracies, but with $R2=0.43$, show some linear correaltion, and a reasonable Kendal-$\tau=0.73$,
-showing moderately strong rank correlation.
+which has 6 out of 13 points outside the $95\%$ confidence bands.  
+In contrast, the 
+Log Frobenius norm %Avg. Log Alpha Norm $\langle\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}\rangle$, 
+in Figure~\ref{fig:summary_regressions_A_12}
+has a much smaller $RMSE=1.93$, a much larger $R2=0.85$, and Kendal-$\tau=0.87$, and 
+has only 2 points outside the $95\%$ confidence bands.
+For examples of lower quality fits, consider the SqueezeNext results, as shown Figures~\ref{fig:summary_regressions_C_10} and~\ref{fig:summary_regressions_C_12}.  
+The 
+Log Spectral norm  %Avg. Log Spectral Norm $\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$ 
+appears visually anti-correlated with the test accuracies (as it is with ShuffleNet, in Figure~\ref{fig:summary_regressions_B_02}).
+It has a very large $95\%$ confidence band, with only 2 points close to the regression line, a large RMSE, $R2=0.07$ (i.e., near zero), and small Kendal-$\tau=0.33$.  
+The 
+Log Frobenius norm %Avg. Log Alpha Norm $\langle\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}\rangle$ 
+is (as always) positively-correlated with test accuracies, but with $R2=0.43$, it show some linear correlation, and a reasonable Kendal-$\tau=0.73$, showing moderately strong rank correlation.
+
+Many more such conclusions can be drawn by examining these Tables and Figures and reproducing the results from our publicly-available repo.
 

### 2021-01-11 submission/NatCom_cv_models.tex
smooth out a bit

--- 2021-01-10 submission/NatCom_cv_models.tex
@@
 
-We examine the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
+We examine the performance of the four quality metrics---Log Frobenius norm
+                                                         ($\langle\log\Vert\mathbf{W}\Vert^{2}_{F}\rangle$), 
+                                                         Log Spectral norm
+                                                         ($\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$),
+                                                         Weighted Alpha
+                                                         ($\hat{\alpha}$),
+                                                         and Log $\alpha$-Norm
+                                                         ($\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$)---applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
 %
@@
 
-\michael{Charles, any more sentences to add to the last three paragraphs to clarify anything?  If not, then just comment out this comment.}
-\charles{In partcular, the PL-based AlphaHat $\hat{\alpha}$ and Shatten-Norm $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ metrics 
-tend to perform better when there is a wider variation in the hyper-parameters going beyond just increasing the depth.  In fact sometimes the
-purely norm-based metrics such as the Spectral Norm can be uncorrelated or even
-anti-correlated with the test accuracy, as seen in Section~\ref{} with
+In particular, the PL-based 
+Weighted Alpha and Log $\alpha$-Norm
+metrics tend to perform better when there is a wider variation in the hyperparameters, going beyond just increasing the depth.  
+In addition, sometimes the purely norm-based metrics such as the Log Spectral norm can be uncorrelated or even anti-correlated with the test accuracy, while the PL-metrrics are positively-correlated.
+This is seen in the Supplementary Material, 
 ShuffleNet in Figure~\ref{fig:summary_regressions_B_02},
-SqueezeNext in Figure~\ref{fig:summary_regressions_C_10},
-and WRN in Figure~\ref{fig:summary_regressions_G_06}).
-}
+SqueezeNext in Figure~\ref{fig:summary_regressions_C_10}, and 
+WRN in Figure~\ref{fig:summary_regressions_G_06}).
 

### 2021-01-11 submission/resp_to_revs_NatCom.tex
revised response letter, semifinal version

@@
 It is fair to say that the power law exponents characterize is the implicit regularization, whether or not the model trainer knew that that was what he or she was doing, but not any regularization explicitly used to train.
-We have clarified this point at Line XXX in the revised manuscript.
+We have clarified this point
+in the revised manuscript.
 \end{quote}
@@
 We know that errors are not normal; so, for this and other reasons, in our experience, reporting $p$-values won't clarify the situation.
-We have expanded all of the tables that report RMSE and/or $R^2$ to report the Kendal-$\tau$ rank correlation measure.
+Instead, to clarify understanding, we have expanded all of the tables that report RMSE and/or $R^2$ to report the Kendal-$\tau$ rank correlation measure; and we have provided figures in the Supplementary Material for all model-dataset pairs, which indicates trends more clearly than the aggregated data.
 Also, a point we didn't emphasize enough is that from HT-SR theory we expect that PL-based metrics to have strong linear and not just rank correlation.
-We have clarified these points at Line XXX, Line XXX, and Line XXX of the revised manuscript, and in the tables and plots throughout the manuscript.
+We have clarified these points 
+in the text 
+of the revised manuscript, and in the tables and plots throughout the manuscript.
 (Also, additional results beyond what we have in the main text and Supplementary Material are available in our fully-reproducible and publicly-available repo, and it is our hope that that level of transparency will aid the reader's understanding.)
-XXX.  MM STILL TO REFINE THIS ANSWER.
+%XXX.  MM STILL TO REFINE THIS ANSWER.
 \end{quote}
@@
 There are many reasons for this, e.g., different normalizations, etc.
-We clarify this point at Line XXX of the revised manuscript.
+We have clarified this point
+in the revised manuscript.
 
@@
 In the main text, we have R2, MSE, and Kendall-Tau; and in the appendix, we have more granular results for several model series.
-Since Figure 9 partitions based on data and not model series, we think that it will likely confuse the reader.
+Since Figure 9 partitions based on data and not model series, we think that it will likely confuse the reader, as opposed to presenting results for all model-dataset pairs, as we now do.
 Thus, we removed it, and we will simply point to the publicly-available repo.
 %
-We clarify these points at Line XXX, Line XXX, and Line XXX of the revised manuscript.
+We have clarified these points
+in the revised manuscript.
 
-XXX.  MM STILL TO REFINE THIS ANSWER.
+%XXX.  MM STILL TO REFINE THIS ANSWER.
 \end{quote}
@@
 The bottom line is that the metrics we discuss can be sued to distinguish between well-trained versus very well trained models, but that one has to look at the PL exponents to diagnose gross model problems, which was one of our original points.
-We clarify these points in revised text between Line XXX and Line XXX of the revised manuscript.
+We have clarified these points
+in the revised manuscript.
 
-XXX.  MM STILL TO REFINE THIS ANSWER.
+%XXX.  MM STILL TO REFINE THIS ANSWER.
 \end{quote}
@@
 In Fig 6b, the spectral norm fails to perform well; while it is true that the higher-quality GPT2 does not exhibit Scale Collapes, the distribution of norms is shifted slightly to the right.
-In Fig 8a, there is a slight shift of alphas to the left as the models get larger (with average values now given in the text at Line XXX).
-Interesting the larger models still have some larger alpha values, and we conjecture at Line XXX that the reason for this is that the models are not fully optimized.  
+In Fig 8a, there is a slight shift of alphas to the left as the models get larger (with average values now given 
+in the text).
+Interesting the larger models still have some larger alpha values, and we conjecture 
+that the reason for this is that the models are not fully optimized.  
 Subsequent developments in NLP have shown this to be the case.
@@
 
-XXX.  MM STILL TO REFINE THIS ANSWER.
+%XXX.  MM STILL TO REFINE THIS ANSWER.
 \end{quote}
@@
 %)))\\
-We have elaborated on this at Line XXX of the revised manuscript.
+We have elaborated on this 
+in the revised manuscript.
 \end{quote}
@@
 More important than the smaller condition number, which has to do with the ratio of the largest to smallest eigenvalue, is the ``shape'' of the spectral distribution, and in particular the details of the tail and how the tail is estimated.
-We clarify this point at Line XXX of the revised manuscript.
+We have clarified this point
+in the revised manuscript.
 
-XXX.  MM STILL TO DO THIS IN TEXT MAYBE.
+%XXX.  MM STILL TO DO THIS IN TEXT MAYBE.
 \end{quote}
@@
 Certainly.
-We clarify this point at Line XXX of the revised manuscript.
+We have clarified this point
+in the revised manuscript.
 \end{quote}
@@
 This is known in GANs, where regularizers are used to prevent rank collapse, but it appears to hold more generally.
-We clarify this point at Line XXX of the revised manuscript.
+We have clarified this point
+in the revised manuscript.
 \end{quote}
@@
 A more precise definition would have to do with the stability of the model under those sorts of perturbations, but we think that including such a precise definition in the main text would only confuse the reader.
-We clarify this point at Line XXX of the revised manuscript.
+We have clarified this point
+in the revised manuscript.
 \end{quote}
@@
 models in the same architecture series.
-We clarify this point at Line XXX of the revised manuscript.
+We have clarified this point
+in the revised manuscript.
 \end{quote}
@@
 Superficially, work on double descent may be interpreted as ``more parameters are better,'' but in practice the situation is not that simple, e.g., since it can be difficult to define an appropriate effective number of parameters in realistic architectures and since realistic models are not fully optimized.
-We comment on and clarify this point at Line XXX of the revised manuscript.
+We comment on and clarify this point 
+in the revised manuscript.
 \end{quote}

### 2021-01-11 submission/resp_to_revs_NatCom.tex
some edits to resp to revs

@@
 
-The first point is there are two things we are illustrating: one is when we increase both data size and model size (GPT2-small to GPT2-medium to GPT2-large to GPT2-xl); and the second is when a give model is trained to high-quality verus low-quality data (GPT2-small versus GPT).
-For the former, we see a general trend with all the metrics decreasing, with the largest difference between GPT2-medium and GPT2-large, and with the difference between GPT2-small and GPT2-medium very small (near the noise level, and slightly in the opposite direction).
-For the latter, we see GPT seems to violate what we expect. 
-Thus, we look in more detail at the contributions to alpha-hat, i.e., to the PL exponent alpha and the spectral norm.
-Doing this makes it clear what while the alpha part behaves as HT-SR theory would predict, i.e., with better models the PL exponents get smaller, the norms behave anomalously and exhibit scale collapes.
-The bottom line is that the metrics we discuss can be sued to distinguish between well-trained versus very well trained models, but that one has to look at the PL exponents to diagnose gross model problems, which was one of our original points.
-We have clarified these points
-in the revised manuscript.
+The first point is there are two things we are illustrating: one is when we increase both data size and model size (GPT2-small to GPT2-medium to GPT2-large to GPT2-xl); and the second is when a given model is trained to high-quality verus low-quality data (GPT2-small versus GPT).
+For the former, as the model size increases, we see a general trend with all the metrics decreasing, with the largest difference between GPT2-medium and GPT2-large, and with the difference between GPT2-small and GPT2-medium being very small (near the noise level, and slightly in the opposite direction).
+For the latter, when comparing the well trained to poorly trained model, we see that the smaller and poorly trained GPT model seems to violate what we expect.
+Thus, we look in more detail at the two contributions to alpha-hat, the PL exponent alpha and the spectral norm.
+Doing this makes it clear what while the alpha part behaves as HT-SR theory would predict, and that with better models the PL exponents get smaller,
+but the norms behave anomalously and exhibit scale collape.
+The bottom line is that the metrics we discuss can be used to distinguish between well-trained versus very well trained models, but that one has to look at the PL exponents and Spectral Norm to diagnose the model problems; this was one of our original points.
+We have clarified these points in the revised manuscript.
 

### 2021-01-11 submission/resp_to_revs_NatCom.tex
fixed typos, etc.

@@
 For the latter, when comparing the well trained to poorly trained model, we see that the smaller and poorly trained GPT model seems to violate what we expect.
-Thus, we look in more detail at the two contributions to alpha-hat, the PL exponent alpha and the spectral norm.
-Doing this makes it clear what while the alpha part behaves as HT-SR theory would predict, and that with better models the PL exponents get smaller,
-but the norms behave anomalously and exhibit scale collape.
-The bottom line is that the metrics we discuss can be used to distinguish between well-trained versus very well trained models, but that one has to look at the PL exponents and Spectral Norm to diagnose the model problems; this was one of our original points.
+Thus, we look in more detail at the two contributions to 
+$\hat{\alpha}$, 
+the PL exponent, 
+$\alpha$, 
+and the spectral norm,
+$\|\cdot\|_{\infty}$.
+Doing this makes it clear what while the 
+$\alpha$ 
+part behaves as HT-SR theory would predict, 
+i.e., 
+that with better models the PL exponents get smaller, 
+the norms behave anomalously and exhibit 
+Scale Collapse.
+The bottom line is that the metrics we discuss can be used to distinguish between well-trained versus very well trained models, but that one has to look at both the PL exponents and Spectral Norm to diagnose the model problems; this was one of our original points.
 We have clarified these points in the revised manuscript.

### 2021-01-11 submission/resp_to_revs_NatCom.tex
add other comment about confidence bands

@@
 We know that errors are not normal; so, for this and other reasons, in our experience, reporting $p$-values won't clarify the situation.
-Instead, to clarify understanding, we have expanded all of the tables that report RMSE and/or $R^2$ to report the Kendal-$\tau$ rank correlation measure; and we have provided figures in the Supplementary Material for all model-dataset pairs, which indicates trends more clearly than the aggregated data.
+Instead, to clarify understanding, we have expanded all of the tables that report RMSE and/or $R^2$ to report the Kendal-$\tau$ rank correlation measure; and we have provided figures in the Supplementary Material for all model-dataset pairs, which indicates trends more clearly than the aggregated data, and which (among other things) show that the plots display the $95\%$ confidence band for every regression along with tabulated results for all metrics.
 Also, a point we didn't emphasize enough is that from HT-SR theory we expect that PL-based metrics to have strong linear and not just rank correlation.

### 2021-01-12 submission/NatCom_all_models.tex
last minute changes before resubmission

@@
 
-Results are summarized in Table~\ref{table:results} (and Figures~\ref{fig:summary_regressions_A}--\ref{fig:summary_regressions_I} of the Supplementary Material).
+Results are summarized in Table~\ref{table:results} (and Figures~\ref{fig:summary_regressions_A}--\ref{fig:summary_regressions_I} of the Supplementary Information).
 For the mean, 
@@
 and PL-based metrics are much better for RMSE mean and standard deviation.
-Model diagnostics (Supplementary Material) indicate many outliers and imperfect fits.
+Model diagnostics (Supplementary Information) indicate many outliers and imperfect fits.
 Overall, though, these and other results suggest our conclusions hold much more generally.

### 2021-01-12 submission/NatCom_cv_models.tex
last minute changes before resubmission

@@
 In addition, sometimes the purely norm-based metrics such as the Log Spectral norm can be uncorrelated or even anti-correlated with the test accuracy, while the PL-metrrics are positively-correlated.
-This is seen in the Supplementary Material, 
+This is seen in the Supplementary Information, 
 ShuffleNet in Figure~\ref{fig:summary_regressions_B_02},

### 2021-01-12 submission/resp_to_revs_NatCom.tex
last minute changes before resubmission

--- 2021-01-11 submission/resp_to_revs_NatCom.tex
@@
 We know that errors are not normal; so, for this and other reasons, in our experience, reporting $p$-values won't clarify the situation.
-Instead, to clarify understanding, we have expanded all of the tables that report RMSE and/or $R^2$ to report the Kendal-$\tau$ rank correlation measure; and we have provided figures in the Supplementary Material for all model-dataset pairs, which indicates trends more clearly than the aggregated data.
+Instead, to clarify understanding, we have expanded all of the tables that report RMSE and/or $R^2$ to report the Kendal-$\tau$ rank correlation measure; and we have provided figures in the Supplementary Information for all model-dataset pairs, which indicate trends more clearly than the aggregated data, and which (among other things) show that the plots display the $95\%$ confidence band for every regression along with tabulated results for all metrics.
 Also, a point we didn't emphasize enough is that from HT-SR theory we expect that PL-based metrics to have strong linear and not just rank correlation.
@@
 of the revised manuscript, and in the tables and plots throughout the manuscript.
-(Also, additional results beyond what we have in the main text and Supplementary Material are available in our fully-reproducible and publicly-available repo, and it is our hope that that level of transparency will aid the reader's understanding.)
+(Also, additional results beyond what we have in the main text and Supplementary Information are available in our fully-reproducible and publicly-available repo, and it is our hope that that level of transparency will aid the reader's understanding.)
 %XXX.  MM STILL TO REFINE THIS ANSWER.
@@
 
-As for For Figure 9, this should be read in light of Tables 7, 8, and 9, as well as the discussion in Sec 2.4.  
+As for (previous) Figure 9, this should be read in light of Tables 7, 8, and 9, as well as the discussion in Sec 2.4.  
 Our metrics for predicting trends in model quality perform well within a given architectural series, and we have discussed this in detail in the main text.
 We also want to show that our general results hold for the hundreds of publicly-available CV and NLP models.
-All of this is available in our publicly-available repo, and so we decided in the text and Supplementary Material to go with just a summary.
+All of this is available in our publicly-available repo, and so we decided in the text and Supplementary Information to go with just a summary.
 In the main text, we have R2, MSE, and Kendall-Tau; and in the appendix, we have more granular results for several model series.
-Since Figure 9 partitions based on data and not model series, we think that it will likely confuse the reader, as opposed to presenting results for all model-dataset pairs, as we now do.
+Since (previous) Figure 9 partitions based on data and not model series, we think that it will likely confuse the reader, as opposed to presenting results for all model-dataset pairs, as we now do.
 Thus, we removed it, and we will simply point to the publicly-available repo.
@@
 %)))\\
-We have removed those statements and only refer to the main text or the Supplementary Material.
+We have removed those statements and only refer to the main text or the Supplementary Information.
 We do point out, however, that we intend out results to be fully-reproducible by the community, and thus our publicly-available repo is designed to contain the enormous quantity of data underlying out analysis.

### 2021-02-18 submission/TOPML_Workshop/paper.tex
initial version of files for TOPML submission

%%%%%%%% ICML 2021 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2021} with \usepackage[nohyperref]{icml2021} above.
\usepackage{hyperref}

\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{overpic}
\usepackage{wrapfig}

\usepackage{siunitx}
\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}

% \DeclareMathOperator*{\cgrad}{grad}
\newcommand{\cgrad}{\operatorname{grad}}
\DeclareMathOperator*{\cdiv}{div}
\DeclareMathOperator*{\dd}{d}
\DeclareMathOperator*{\diag}{diag}
\DeclareMathOperator*{\vect}{vec}
\DeclareMathOperator*{\real}{Re}

\newcommand{\code}[1]{\texttt{#1}}
\newtheorem{prop}{Proposition}
\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}

\newcommand{\michael}[1]{{\color{red}\sf{[Michael: #1]}}}

\newcommand{\fakesection}[1]{%
  \par\refstepcounter{section}% Increase section counter
  \sectionmark{#1}% Add section mark (header)
  \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}% Add section to ToC
  % Add more content here, if needed.
}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2021}

% If accepted, instead use the following line for the camera-ready submission:
% \usepackage[accepted]{icml2021}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{A Differential Geometry Perspective on Orthogonal Recurrent Models}

\begin{document}

\twocolumn[
\icmltitle{XXX TITLE}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2021
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Charles H. Martin}{CC}
\icmlauthor{Michael W. Mahoney}{UCB}

\end{icmlauthorlist}

\icmlaffiliation{CC}{Department of Computer Science at Ben-Gurion University of the Negev, Israel.}
\icmlaffiliation{UCB}{ICSI and Department of Statistics at UC Berkeley, CA, USA.}

\icmlcorrespondingauthor{Omri Azencot}{azencot@cs.bgu.ac.il}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Differential Geometry, Sequence Models}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
% \printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}
XXX.  ABSTRACT.  THE NOMINAL CONTRIBUTION IS TO APPLY HT-SR THEORY AND WEIGHTWATCHER TO DO AN ANALYSIS OF DATA ANS IDENTIFY SIMPSONS LIKE PARADOXES.
\end{abstract}

\section{Introduction}

XXX.  INTRO AND OVERVIEW, BASICALLY THE CORE DUMP CHARLES DID ON THE PHONE.

\michael{Charles, put in a core dump of the intro that you said.  Maybe half a column to one column, in this two column format.}

\section{Statistical Mechanics and Heavy-Tailed Self Regularization}

XXX.  SUMMARY OF STAT MECH AND HTSR THEORY FOR DNNS.

\cite{MM17_TR} is Charles: Rethinking generalization requires revisiting old ideas: statistical mechanics approaches and complex learning behavior''

\cite{MM18_TR} is Charles: ``Implicit Self-Regularization in Deep Neural Networks: Evidence from Random Matrix Theory and Implications for Learning''

\cite{MM19_HTSR_ICML} is Charles: ``Traditional and Heavy-Tailed Self Regularization in Neural Network Models''

\cite{MM20_SDM} is Charles: ``Heavy-Tailed {U}niversality Predicts Trends in Test Accuracies for Very Large Pre-Trained Deep Neural Networks''

\cite{MM19_KDD} is Charles: ``Statistical Mechanics Methods for Discovering Knowledge from Modern Production Quality Neural Networks''

XXX.  ABSTRACT FROM \cite{MM18_TR}:
Random Matrix Theory (RMT) is applied to analyze the weight matrices of Deep Neural Networks (DNNs), including both production quality, pre-trained models such as AlexNet and Inception, and smaller models trained from scratch, such as LeNet5 and a miniature-AlexNet.
Empirical and theoretical results clearly indicate that the DNN training process itself implicitly implements a form of \emph{Self-Regularization}, implicitly sculpting a more regularized energy or penalty landscape.
In particular, the empirical spectral density (ESD) of DNN layer matrices displays signatures of traditionally-regularized statistical models, even in the absence of exogenously specifying traditional forms of explicit regularization, such as Dropout or Weight Norm constraints.
Building on relatively recent results in RMT, most notably its extension to Universality classes of Heavy-Tailed matrices, and applying them to these empirical results,
we develop a theory to identify \emph{5+1 Phases of Training}, corresponding to increasing amounts of \emph{Implicit Self-Regularization}.
These phases can be observed during the training process as well as in the final learned DNNs.
For smaller and/or older DNNs, this Implicit Self-Regularization is like traditional Tikhonov regularization, in that there is a ``size scale'' separating signal from noise.
For state-of-the-art DNNs, however, we identify a novel form of \emph{Heavy-Tailed Self-Regularization}, similar to the self-organization seen in the statistical physics of disordered systems (such as classical models of actual neural activity).
This results from correlations arising at all size scales, which for DNNs arises implicitly due to the training process itself.
This implicit Self-Regularization can depend strongly on the many knobs of the training process.
In particular, by exploiting the generalization gap phenomena,
we demonstrate that we can cause a small model to exhibit all 5+1 phases of training simply by changing the batch size.
This demonstrates that---all else being equal---DNN optimization with larger batch sizes leads to less-well implicitly-regularized models, and it provides an explanation for the generalization gap phenomena.
Our results suggest that large, well-trained DNN architectures should exhibit Heavy-Tailed Self-Regularization, and we discuss the theoretical and practical implications of~this.

\subsection{Overall approach}
\label{sxn:approach}

XXX.  THE FOLLOWING IS TROM THE PREDICTING TRENDS PAPER.

Consider the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, and weight matrices $\mathbf{W}_{l}$ and bias vectors $\mathbf{b}_{l}$, as 
the minimization of a general loss function $\mathcal{L}$ over the training data instances and labels, $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$.
For a typical supervised classification problem, the goal of training is to construct (or learn) $\mathbf{W}_{l}$ and $\mathbf{b}_{l}$ that capture correlations in the data, in the sense of solving
\begin{equation}
%%%\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N}\left(\mathcal{L}(E_{DNN}(\mathbf{x}_{i})-y_{i})\right)  ,
\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N} \mathcal{L}(E_{DNN}(\mathbf{x}_{i}),y_{i})   ,
\end{equation}
where the loss function $\mathcal{L}(\cdot,\cdot)$ can take on a myriad of forms, and where the energy (or optimization) landscape function
\begin{equation}
%E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})
E_{DNN} = f(\mathbf{x}_{i} ; \mathbf{W}_{1},\ldots,\mathbf{W}_{L},\mathbf{b}_{1},\ldots,\mathbf{b}_{L})
\label{eqn:dnn_energy}
\end{equation}
depends parametrically on the weights and biases.
For a trained model, the form of the function $E_{DNN}$ does not explicitly depend on the data (but it does explicitly depend on the weights and biases).
The function $E_{DNN}$ maps data instance vectors ($\mathbf{x}_i$ values) to predictions ($y_{i}$ labels), and thus the output of this function does depend on the data.
Therefore, one can analyze the form of $E_{DNN}$ in the absence of any training or test~data.

Test accuracies have been reported online for publicly-available pretrained pyTorch models~\cite{osmr}.
These models have been trained and evaluated on labeled data $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$, using standard techniques.  
We do not have access to this data, and we have not trained any of the models ourselves. 
Our methodological approach is thus similar to a statistical meta-analysis, common in biomedical research, but uncommon in ML.
%
Computations were performed with the publicly-available \texttt{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.

\paragraph{Metrics for DNN Weight Matrices.}

Our approach involves analyzing individual DNN weight matrices, for (depending on the architecture) fully-connected and/or convolutional layers.
Each DNN layer contains one or more layer 2D  $N_{l}\times M_{l}$ weight matrices, $\mathbf{W}_{l}$, or pre-activation maps, $\mathbf{W}_{i,l}$, e.g., extracted from 2D Convolutional layers, where $N > M$.
See the Supplementary Information
%~\ref{sxn:appendix} 
for details.
(We may drop the $i$ and/or $i,l$ subscripts below.)
The best performing quality metrics depend on the norms and/or spectral properties of each weight matrix,
$\mathbf{W}$, and/or, equivalently, it's empirical correlation matrix, $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
To evaluate the quality of state-of-the-art DNNs, 
we consider the following metrics:
\begin{eqnarray}
& & \text{Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{X}\Vert_{F}=\sum\nolimits_{i=1}^{M} \lambda_{i}$ } \\
& & \text{Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}^{2}=\Vert\mathbf{X}\Vert_{\infty}=\lambda_{max}$ } \\
& & \text{Weighted Alpha: $\hat{\alpha}=\alpha\log\lambda_{max}$ } \\
& & \text{$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{W}\Vert^{2\alpha}_{2\alpha}=\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum\nolimits_{i=1}^{M}\lambda_{i}^{\alpha}$. }
\end{eqnarray}
%To diagnose problems, 
To perform diagnostics on potentially-problematic DNNs,
we will decompose $\hat{\alpha}$ into its two components, $\alpha$ and $\lambda_{max}$.
Here, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the $\mathbf{X}$, $\lambda_{max}$ is the maximum eigenvalue, and $\alpha$ is the fitted PL exponent. 
These eigenvalues are squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, $\lambda_{i}=\sigma^{2}_{i}$.
All four metrics can be computed easily from DNN weight matrices.
The first two metrics are well-known in ML.
The last two metrics deserve special mention, as they depend on an empirical parameter $\alpha$ that is the PL exponent that arises in the recently-developed Heavy Tailed Self Regularization (HT-SR) Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.

\paragraph{Overview of Heavy-Tailed Self-Regularization.}

In the HT-SR Theory, one analyzes the eigenvalue spectrum, i.e., the Empirical Spectral Density (ESD), of the associated correlation matrices~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
From this, one characterizes the amount and form of correlation, and therefore implicit self-regularizartion, present in the DNN's weight matrices.
For each layer weight matrix $\mathbf{W}$, of size $N \times M$, construct the associated $M\times M$ (uncentered) correlation matrix $\mathbf{X}$. 
Dropping the $L$ and $l,i$ indices, one~has
$$
\mathbf{X} = \frac{1}{N}\mathbf{W}^{T}\mathbf{W}.
$$
If we compute the eigenvalue spectrum of $\mathbf{X}$, i.e., $\lambda_i$ such that
$  % $$
\mathbf{X}\mathbf{v}_{i}=\lambda_{i}\mathbf{v}_{i} , 
$  % $$
then the ESD of eigenvalues, $\rho(\lambda)$, is just a histogram of the eigenvalues, formally written as
%\begin{equation}
$\rho(\lambda)=\sum\nolimits_{i=1}^{M}\delta(\lambda-\lambda_{i})  .$
%\label{eqn:eigenval_hist}
%\end{equation}
Using HT-SR Theory, one characterizes the correlations in a weight matrix by examining its ESD, $\rho(\lambda)$.
It can be well-fit to a truncated PL distribution, given~as
\begin{equation}
\rho(\lambda)\sim\lambda^{-\alpha}  ,
\label{eqn:eigenval_pl}
\end{equation}
which is (at least) valid within a bounded range of eigenvalues $\lambda\in[\lambda^{min},\lambda^{max}]$.

The original work on HT-SR Theory considered a small number of NNs, including AlexNet and InceptionV3. 
It showed that for nearly every $\mathbf{W}$, the (bulk and tail) of the ESDs can be fit to a truncated PL, and that PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
%
%%As for the mechanism responsible for these properties, statistical physics offers several possibilities~\cite{SornetteBook,nishimori01}, e.g., self organized criticality~\cite{SOC87,SOCat25yrs} or multiplicative noise in the stochastic optimization algorithms used to train these models~\cite{HodMah20A_TR,SorCon97}.
%%Alternatively, related techniques have been used to analyze correlations and information propogation in actual spiking neurons~\cite{SYYRP11,YKYP14}.
%%Our meta-analysis does not require knowledge of mechanisms; and it is not even clear that one mechanism is responsible for every case.
%
Crucially, HT-SR Theory predicts that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models.
The notion of ``size scale'' is well-defined in physical systems, to which this style of analysis is usually applied, but it is less well-defined in CV and NLP applications.
Informally, it would correspond to pixel groups that are at a greater distance in some metric, or between sentence parts that are at a greater distance in text.
Relatedly, previous work observed that smaller exponents $\alpha$ correspond to more implicit self-regularization and better generalization, and that we expect a linear correlation between $\hat{\alpha}$ and model quality~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.

\paragraph{DNN Empirical Quality Metrics.}

For norm-based metrics, we use the average of the log norm, to the appropriate power.
Informally, this amounts to assuming that the layer weight matrices are statistically independent, in which case we can estimate the model complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm (which resembles a data dependent VC complexity),
\begin{equation}
\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert \times \cdots \times \Vert\mathbf{W}_{L}\Vert ,
\end{equation}
where $\Vert\cdot\Vert$ is a matrix norm.   
The log complexity,
\begin{equation}
\label{eqn:eqn:sum_log_norm}
\log\mathcal{C} \sim \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert + \cdots + \log\Vert\mathbf{W}_{L}\Vert = \sum\nolimits_l \log\Vert\mathbf{W}_{l}\Vert ,
\end{equation}
 takes the form of an average Log Norm.
For the Frobenius Norm metric and Spectral Norm metric, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly (since, when taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor).

The Weighted Alpha metric is an average of $\alpha_l$ over all layers $l \in \{1,\ldots,l\}$, weighted by the size, or scale, or each matrix,
\begin{equation}
\hat{\alpha} = \dfrac{1}{L}\sum_l \alpha_l\log\lambda_{max,l}\approx\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle    ,
\end{equation}
where $L$ is the total number of layer weight matrices.
The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a much smaller and more limited set of models than we consider here.

Based on this, in this paper, we introduce and evaluate the $\alpha$-Shatten Norm metric,
\begin{equation}
\label{eqn:sum_log_alpha_norm_alpha}
\sum\nolimits_l \log \Vert\mathbf{X}_l\Vert_{\alpha_l}^{\alpha_l} 
=
\sum\nolimits_l \alpha_l \log \Vert\mathbf{X}_l\Vert_{\alpha_l} .
\end{equation}
For the $\alpha$-Shatten Norm metric, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it cannot be taken out of the sum.
For small $\alpha$, the Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation;
%% \cite{MM20_unpub_work}; 
and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm.

Finally, although it does less well for predicting trends in state-of-the-art model series, e.g., as depth changes, the average value of $\alpha$, i.e., 
\begin{equation}
\bar{\alpha} = \dfrac{1}{L}\sum_l \alpha_l = \langle\alpha\rangle    ,
\label{eqn:alpha_bar}
\end{equation}
can be used to perform model diagnostics, to identify problems that cannot be detected by examining training/test accuracies, and to discriminate poorly-trained models from well-trained~models.

One determines $\alpha$ for a given layer by fitting the ESD of that layer's weight matrix to a truncated PL, using the commonly accepted Maximum Likelihood method~\cite{CSN09_powerlaw,ABP14}.
%%This method works very well for exponents between $\alpha\in(2,4)$; and it is adequate, although imprecise, for smaller and especially larger $\alpha$~\cite{newman2005_zipf}. 
%
Operationally, $\alpha$ is determined by using the \texttt{WeightWatcher} tool~\cite{weightwatcher_package} to fit the histogram of eigenvalues, $\rho(\lambda)$, to a truncated PL, 
%%\begin{equation}
%%\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\in[\lambda_{min},\lambda_{max}] ,
%%\end{equation}
%%where $\lambda_{max}$ is the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$, and 
%%where $\lambda_{min}$ is selected automatically to yield the best (in the sense of minimizing the K-S distance) PL fit.
%%Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.
%%See Figure~\ref{fig:ww} for an illustration.

%\begin{figure}[t]
%    \centering
%    %\includegraphics[width=15.0cm]{img/WeightWatcher_v2}
%    \includegraphics[width=15.0cm]{img/WeightWatcher_v3}
%    \caption{Schematic of analyzing DNN layer weight matrices $\mathbf{W}$.  
%             Given an individual layer weight matrix $\mathbf{W}$, from either a fully-connected layer or a convolutional layer, perform a Singular Value Decomposition (SVD) to obtain $\mathbf{W} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^{T}$, and examine the histogram of eigenvalues of $\mathbf{W}^{T}\mathbf{W}$.
%             Norm-based metrics and PL-based metrics (that depend on fitting the histogram of eigenvalues to a truncated PL) can be used to compare models.
%             For example, one can analyze one layer of a pre-trained model, compare multiple layers of a pre-trained model, make comparisons across model architectures, monitor neural network properties during training, etc. 
%}
%    \label{fig:ww}
%\end{figure}

To avoid confusion, let us clarify the relationship between $\alpha$ and $\hat{\alpha}$.  
We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated PL, parameterized by 2 values: the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.
The PL exponent $\alpha$ measures the amount of correlation in a DNN layer weight matrix $\mathbf{W}$. 
It is valid for $\lambda\le\lambda_{max}$, and it is scale-invariant, i.e., it does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.
The $\lambda_{max}$ is a measure of the size, or scale, of $\mathbf{W}$.
Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric $\hat{\alpha}$ for the entire~DNN.
We will see that for well-trained CV and NLP models, $\hat{\alpha}$ performs quite well and as expected, but for CV and NLP models that are potentially-problematic or less well-trained, metrics that depend on the scale of the problem can perform anomalously.  
In these cases, separating $\hat{\alpha}$ into its two components, $\alpha$ and $\lambda_{max}$, and examining the distributions of each, can be helpful.

\section{Predicting Trends in the Quality of State-of-the-art Models}

XXX.  THIS IS A VERY PRACTICAL PROBLEM THAT OUR THEORETICAL APPROACH, STRONGLY COUPLED AND MOTIVATED BY EMPIRICAL PRACTICE, CAN SOLVE.

\cite{MM20a_trends_TR} is Charles: ``Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data''

XXX.  ABSTRACT FROM \cite{MM20a_trends_TR}:
In many applications, one works with neural network models trained by someone else.
For such pretrained models, one may not have access to training data or test data.
Moreover, one may not know details about the model, e.g., the specifics of the training data, the loss function, the hyperparameter values, etc.
Given one or many pretrained models, it is a challenge to say anything about the expected performance or quality of the models.
Here, we address this challenge by providing a detailed meta-analysis of hundreds of publicly-available pretrained models.
We examine norm based capacity control metrics as well as power law based metrics from the recently-developed Theory of Heavy-Tailed Self Regularization.
We find that norm based metrics correlate well with reported test accuracies for well-trained models, but that they often cannot distinguish well-trained versus poorly-trained models.
We also find that power law based metrics can do much better---quantitatively better at discriminating among series of well-trained models with a given architecture; and qualitatively better at discriminating well-trained versus poorly-trained models.
These methods can be used to identify when a pretrained neural network has problems that cannot be detected simply by examining training/test accuracies.

XXX.  CITE ALSO BARTLETT AND POGGIO AS DOING SOMETHING LOOSELY SIMILAR, AND THE GOOGLE CONTEST AS DOING SOMETHING LOOSELY SIMILAR BUT DIFFERENT.

\cite{BFT17_TR} is Bartlett: ``Spectrally-normalized margin bounds for neural networks''

\cite{LMBx18_TR} is Poggio: ``A surprising linear relationship predicts test performance in deep networks''

\subsection{Introduction OF PREDICTING TRENDS PAPER}
\label{sxn:intro}

XXX.  THIS IS THE INTRO FROM THE PREDICTING TRENDS PAPER.

A common problem in machine learning (ML) 
is to evaluate the quality of a given model.
A popular way to accomplish this
is to train a model and then evaluate its training/testing error.
There are many problems with this approach.
The training/testing curves give very limited insight into the overall properties of the model; 
they do not take into account the (often large human and CPU/GPU) time for hyperparameter fiddling;
they typically do not correlate with other properties of interest such as robustness or fairness or interpretability; 
and so on.
A related problem, in particular in industrial-scale artificial intelligence (AI), arises when the model user is not the model developer.
Then, one may not have access to the training data or the testing data.
Instead, one may simply be given a model that has already been trained---a pretrained model---and need to use it as-is, or to fine-tune and/or compress it and then use it.

Na\"{\i}vely---but in our experience commonly, among ML practitioners and ML theorists---if one does not have access to training or testing data, then one can say absolutely nothing about the quality of a ML model.
This may be true in worst-case theory, but models are used in practice, and there is a need for a practical theory to guide that practice.
Moreover, if ML is to become an industrial process, then that process will become compartmentalized in order to scale: some groups will gather data, other groups will develop models, and other groups will use those models.
Users of models cannot be expected to know the precise details of how models were built, the specifics of data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized,~etc.

Moreover, for many large scale, practical applications, there is no obvious way to define an ideal test metric. 
For example, models that generate fake text or conversational chatbots may use a proxy, like perplexity, as a test metric.
In the end, however, they really require human evaluation. 
Alternatively, models that cluster user profiles, which are widely used in areas such as marketing and advertising, are unsupervised and have no obvious labels for comparison and/or evaluation.
In these and other areas, ML objectives can be poor proxies for downstream goals.

Most importantly, in industry, one faces unique practical problems such as determining whether one has enough data for a given model. 
Indeed, high quality, labeled data can be very expensive to acquire, and this cost can make or break a project.
Methods that are developed and evaluated on any well-defined publicly-available corpus of data, no matter how large or diverse or interesting, are clearly not going to be well-suited to address problems such as this.
It is of great practical interest to have metrics to evaluate the quality of a trained model---in the absence of training/testing data and without any detailed knowledge of the training/testing process.  
There is a need for a practical theory for pretrained models which can predict how, when, and why such models can be expected to perform well or~poorly.

In the absence of training and testing data, obvious quantities to examine are the weight matrices of pretrained models, e.g., 
properties such as norms of weight matrices and/or parameters of Power Law (PL) fits of the eigenvalues of weight matrices.
Norm-based metrics have been used in traditional statistical learning theory to bound capacity and construct regularizers; and PL fits are based on statistical mechanics approaches to deep neural networks (DNNs).
While we use traditional norm-based and PL-based metrics, our goals are not the traditional goals.
Unlike more common ML approaches, we do not seek a bound on the generalization (e.g., by evaluating training/test errors), we do not seek a new regularizer, and we do not aim to evaluate a single model (e.g., as with hyperparameter optimization).
Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves.
In both cases, one can ask whether it is possible to predict trends in the quality of pretrained DNN models without access to training or testing data.

To answer this question, we provide a detailed empirical analysis, evaluating quality metrics for pretrained DNN models, and we do so at scale.
Our approach may be viewed as a statistical meta-analysis of previously published work, where we consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
By now, there are many such state-of-the-art models that are publicly-available, e.g., 
hundreds of pretrained models in CV ($\ge 500$) and NLP ($\approx 100$).%
\footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
For all these models, we have no access to training data or testing data, and we have no specific knowledge of the training/testing protocols. 
%
Here is a summary of our main results.
First, norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models.
Second, norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.).
For example, they may exhibit what we call Scale Collapse for these models. 
Third, PL-based metrics can do much better at predicting quality trends in pretrained CV/NLP models.  
In particular, 
a weighted PL exponent (weighted by the log of the spectral norm of the corresponding layer) is 
quantitatively better at discriminating among a series of well-trained versus very-well-trained models
within a given architecture series; and
the (unweighted) average PL exponent is 
qualitatively better at discriminating well-trained versus poorly-trained models.
Fourth, PL-based metrics can also be used to characterize fine-scale model properties, including what we call layer-wise Correlation Flow, in well-trained and poorly-trained models; and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
%
Our work provides a theoretically-principled empirical evaluation---by far the largest, most detailed, and most comprehensive to date---and the theory we apply was developed previously~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
Performing such a meta-analysis of previously-published work is common in certain areas, but it is quite rare in ML, where the emphasis is on developing better training protocols.

\section{A Simpson's Paradox---and the Need for Finer-scale Analysis}

XXX.  IF WANT A ONE SIZE FITS ALL METRIC, THEN CAN EXPECT A SIMPSONS PARADOX.  OUR TECHNIQUES CAN DIAGNOST THIS.  AND IN FACT WE SEE THAT.

\michael{Charles, put in a paragraph and the figure for the Simpson's paradox here.}

\cite{JNBx19_fantastic_TR} is the terrible fantastic paper: ``Fantastic Generalization Measures and Where to Find Them''

\cite{JFYx20_contest_v10} is the contest: ``{NeurIPS} 2020 Competition: Predicting Generalization in Deep Learning (Version 1.0)''

\cite{BHO75} is Bickel: ``Sex Bias in Graduate Admissions: Data from Berkeley''

\section{Conclusion}

We have shown that weight matrix analysis derived from HT-SR Theory can be used to predict trends in the quality of state-of-the-art neural networks without access to training or testing data as well as identify Simpson's-like paradoxes in publicly-available NN models.
The HT-SR Theory underlying our analysis is quite different than theoretical approaches popular in ML, and indeed it is quite different than typical statistical mechanics approaches to NNs \cite{BKPx20}.
Recently, however, work motivated by HT-SR Theory has 
demonstrated that multiplicative noise in stochastic optimization leads to heavy tails~\cite{HodMah20A_TR,GSZ20_TR}; 
provided a random matrix analysis of random {F}ourier features, going beyond the {G}aussian kernel, to provide a precise phase transition and illustrate the corresponding double descent~\cite{ZCM20_TR}; 
provided exact expressions for double descent and implicit regularization via a notion of surrogate random design \cite{DLM19_Exact_TR}; 
demonstrated that good classifiers are abundant in the interpolating regime \cite{TKM20_TR};
shown how injecting noise into hidden states of recurrent NNs leads to implicit regularization~\cite{LEHM21_TR,Mah12}; and 
shown that analogous phase transitions hold for the {C}olumn {S}ubset {S}election and the {N}ystrom method~\cite{DKM20_TR}. 
We expect that the empirical results identified by tools from HT-SR Theory will provide guidance and constraints for future development of overparameterized machine learning models.

\bibliographystyle{icml2021}
\bibliography{dnns}

\end{document}

### 2021-02-18 submission/TOPML_Workshop/paper.tex
TOPML updated

@@
-%%%%%%%% ICML 2021 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-
 \documentclass{article}
@@
 \newcommand{\michael}[1]{{\color{red}\sf{[Michael: #1]}}}
+\newcommand{\charles}[1]{{\color{blue}\sf{[Charles: #1]}}}
 
@@
 
+\charles{We apply our theory of Heavy Tailed Self-Regularization (HTSR) to predict the generalization accuracies of the pre-trained models provided in this contest, as a function of both model depth (L) and the solver hyperparameters (batch size, dropout, weight decay), and compare quality metrics that describe both the heavy-tailed shape and norm-based scale of the model layer weight matrices.
+In doing so, we identify what amounts to a Simpson’s paradox, in which
+shape and scale metrics show opposite behavior, and a combine metric resolves the paradox}
+
+\begin{figure}[t!] 
+    \centering
+    \subfigure[Spectral Norm $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$]{
+      \includegraphics[width=4.0cm]{img/simpsons-task2-logspectralnorm.png}
+    }
+    \subfigure[Alpha Hat $\hat{\alpha}$]{
+     \includegraphics[width=4cm]{img/simpsons-task2-alpha_weighted.png}
+    }
+    \caption{Test accuracy as a function of the
+Log Spectral Norm $\log\Vert\mathbf{W}\Vert_{\infty}$ 
+and the HTSR Alpha Hat $\hat{\alpha}$ metrics
+as model hyperparameters and depth are varied. 
+Models from the NeurIPS Predicting Generalization contest.}
+    \label{fig:simpsons}
+\end{figure}
+
 \cite{JNBx19_fantastic_TR} is the terrible fantastic paper: ``Fantastic Generalization Measures and Where to Find Them''

### 2021-02-18 submission/TOPML_Workshop/intro.tex
added into tex

We present our Theory of Heavy Tailed Self Regularization (HTSR) and
its application to the (overparameterized)  Deep Neural Networks (DNNs) 
to  predict the generalization accuracy and improve fine-tuning
of pretrained models by capturing the dominant correlation structure.
We analyze the correlations using techniques from 
Heavy Tailed Random Matrix Theory (HT-RMT)
and from  strongly correlated systems from theoretical
chemistry and  physics.  
We show how to predict trends in the reported test accuracies
across hundreds of pretrained DNNs without needing access
to training or test data by charactering the heavy tailed
structure of the empirical spectral densities (ESD) of 
the layer weight matrices.
In doing so, we show that different metrics, which
either depend on the shape or scale of the ESDs,
exhbit a Simpson's paradox when varying DNN model depth
vs regularization hyperparameters.
In particular,  the familar scale-based spectral norm
metric behave opposite theoretical predictions,
whereas our HT alpha-metric acts as predicted.
Moreover, the paradoxical behavior  can be resolved
by combining these shape and scale metrics with
our HT alpha-hat metric.
Finally,  And with access to training data, we show how to
smooth and capture the low rank structure to both accurately
predict the generalization errors and potentially improve
finetuning of pretrained DNNs.

### 2021-02-18 submission/TOPML_Workshop/paper.tex
paper on four pages, lots of fat removed

@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{A Differential Geometry Perspective on Orthogonal Recurrent Models}
+\icmltitlerunning{Using Heavy-Tailed Self-Regularization Theory}
 
@@
 \twocolumn[
-\icmltitle{XXX TITLE}
+\icmltitle{Using Heavy-Tailed Self-Regularization Theory to Characterize Correlations in and Perform Diagnostics on State-of-the-art Overparameterized Models}
 
@@
 \begin{abstract}
-XXX.  ABSTRACT.  THE NOMINAL CONTRIBUTION IS TO APPLY HT-SR THEORY AND WEIGHTWATCHER TO DO AN ANALYSIS OF DATA ANS IDENTIFY SIMPSONS LIKE PARADOXES.
+Understanding state-of-the-art overparameterized neural networks requires detailed empirical studies as well as the development of predictive theory in light of those studies. 
+Building on the theory of Heavy-Tailed Self-Regularization, we show that correlations in the weight matrices of state-of-the-art overparameterized neural networks are well-described by heavy-tailed random matrix theory.
+Using the publicly-available WeightWatcher tool, we perform spectral analylsis on these weight matrices, and we show that fitted truncated power law exponents are strongly predictive of model quality, for publicly-available state-of-the-art pre-trained models, even in the absence of training or testing data.
+We also show that these metrics can also be used to identify a Simpson's paradox in the data from a recent contest aimed at developing metrics to predice generalization in deep learning.
 \end{abstract}
@@
 
-XXX.  INTRO AND OVERVIEW, BASICALLY THE CORE DUMP CHARLES DID ON THE PHONE.
-
-\michael{Charles, put in a core dump of the intro that you said.  Maybe half a column to one column, in this two column format.}
+\michael{Here is Charles core dump of intro; still to refine and incorporate with rest of text.}
+We present our Theory of Heavy Tailed Self Regularization (HTSR) and
+its application to the (overparameterized)  Deep Neural Networks (DNNs) 
+to  predict the generalization accuracy and improve fine-tuning
+of pretrained models by capturing the dominant correlation structure.
+We analyze the correlations using techniques from 
+Heavy Tailed Random Matrix Theory (HT-RMT)
+and from  strongly correlated systems from theoretical
+chemistry and  physics.  
+We show how to predict trends in the reported test accuracies
+across hundreds of pretrained DNNs without needing access
+to training or test data by charactering the heavy tailed
+structure of the empirical spectral densities (ESD) of 
+the layer weight matrices.
+In doing so, we show that different metrics, which
+either depend on the shape or scale of the ESDs,
+exhbit a Simpson's paradox when varying DNN model depth
+vs regularization hyperparameters.
+In particular,  the familar scale-based spectral norm
+metric behave opposite theoretical predictions,
+whereas our HT alpha-metric acts as predicted.
+Moreover, the paradoxical behavior  can be resolved
+by combining these shape and scale metrics with
+our HT alpha-hat metric.
+Finally,  And with access to training data, we show how to
+smooth and capture the low rank structure to both accurately
+predict the generalization errors and potentially improve
+finetuning of pretrained DNNs.
 
@@
 
-XXX.  SUMMARY OF STAT MECH AND HTSR THEORY FOR DNNS.
+%\citep{MM17_TR} is Charles: Rethinking generalization requires revisiting old ideas: statistical mechanics approaches and complex learning behavior''
+%
+%\citep{MM18_TR} is Charles: ``Implicit Self-Regularization in Deep Neural Networks: Evidence from Random Matrix Theory and Implications for Learning''
+%
+%\citep{MM19_HTSR_ICML} is Charles: ``Traditional and Heavy-Tailed Self Regularization in Neural Network Models''
+%
+%\citep{MM20_SDM} is Charles: ``Heavy-Tailed {U}niversality Predicts Trends in Test Accuracies for Very Large Pre-Trained Deep Neural Networks''
+%
+%\citep{MM19_KDD} is Charles: ``Statistical Mechanics Methods for Discovering Knowledge from Modern Production Quality Neural Networks''
 
-\cite{MM17_TR} is Charles: Rethinking generalization requires revisiting old ideas: statistical mechanics approaches and complex learning behavior''
+Recently, Random Matrix Theory (RMT) was applied to analyze empirically the weight matrices of Deep Neural Networks (DNNs), including both production quality, pre-trained models such as AlexNet and Inception, and smaller models trained from scratch, such as LeNet5 and a miniature-AlexNet~\citep{MM18_TR} (see also \citep{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM20_SDM, MM19_KDD}).
+%%Empirical results clearly indicate that the empirical spectral density (ESD) of DNN layer matrices displays signatures of traditionally-regularized statistical models, even in the absence of exogenously specifying traditional forms of explicit regularization, such as dropout or weight norm~constraints.
+%%
+Building on these empirical results, 
+%%and using relatively recent results in RMT, most notably its extension to Universality classes of heavy-tailed matrices, 
+\citet{MM18_TR} then developed a phenomenological (statistical mechanics based, as opposed to statistical learning theory based) theory to identify ``5+1 Phases of Training,'' corresponding to increasing amounts of implicit self-regularization.
+These phases can be observed during the training process as well as in the final learned DNNs.
+For smaller and/or older DNNs, this implicit self-regularization is like traditional Tikhonov regularization, in that there is a ``size scale'' separating signal from noise.
+For state-of-the-art (SOTA) DNNs, however, they identified a novel form of \emph{Heavy-Tailed Self-Regularization} (HT-ST).
+This is similar to the self-organization seen in the statistical physics of disordered systems (such as classical models of actual neural activity); and 
+it results from correlations arising at all size scales.
+For DNNs, this arises due to the training process itself, when applied to very overparameterized models.
 
-\cite{MM18_TR} is Charles: ``Implicit Self-Regularization in Deep Neural Networks: Evidence from Random Matrix Theory and Implications for Learning''
-
-\cite{MM19_HTSR_ICML} is Charles: ``Traditional and Heavy-Tailed Self Regularization in Neural Network Models''
-
-\cite{MM20_SDM} is Charles: ``Heavy-Tailed {U}niversality Predicts Trends in Test Accuracies for Very Large Pre-Trained Deep Neural Networks''
-
-\cite{MM19_KDD} is Charles: ``Statistical Mechanics Methods for Discovering Knowledge from Modern Production Quality Neural Networks''
-
-XXX.  ABSTRACT FROM \cite{MM18_TR}:
-Random Matrix Theory (RMT) is applied to analyze the weight matrices of Deep Neural Networks (DNNs), including both production quality, pre-trained models such as AlexNet and Inception, and smaller models trained from scratch, such as LeNet5 and a miniature-AlexNet.
-Empirical and theoretical results clearly indicate that the DNN training process itself implicitly implements a form of \emph{Self-Regularization}, implicitly sculpting a more regularized energy or penalty landscape.
-In particular, the empirical spectral density (ESD) of DNN layer matrices displays signatures of traditionally-regularized statistical models, even in the absence of exogenously specifying traditional forms of explicit regularization, such as Dropout or Weight Norm constraints.
-Building on relatively recent results in RMT, most notably its extension to Universality classes of Heavy-Tailed matrices, and applying them to these empirical results,
-we develop a theory to identify \emph{5+1 Phases of Training}, corresponding to increasing amounts of \emph{Implicit Self-Regularization}.
-These phases can be observed during the training process as well as in the final learned DNNs.
-For smaller and/or older DNNs, this Implicit Self-Regularization is like traditional Tikhonov regularization, in that there is a ``size scale'' separating signal from noise.
-For state-of-the-art DNNs, however, we identify a novel form of \emph{Heavy-Tailed Self-Regularization}, similar to the self-organization seen in the statistical physics of disordered systems (such as classical models of actual neural activity).
-This results from correlations arising at all size scales, which for DNNs arises implicitly due to the training process itself.
-This implicit Self-Regularization can depend strongly on the many knobs of the training process.
-In particular, by exploiting the generalization gap phenomena,
-we demonstrate that we can cause a small model to exhibit all 5+1 phases of training simply by changing the batch size.
-This demonstrates that---all else being equal---DNN optimization with larger batch sizes leads to less-well implicitly-regularized models, and it provides an explanation for the generalization gap phenomena.
-Our results suggest that large, well-trained DNN architectures should exhibit Heavy-Tailed Self-Regularization, and we discuss the theoretical and practical implications of~this.
-
-\subsection{Overall approach}
-\label{sxn:approach}
-
-XXX.  THE FOLLOWING IS TROM THE PREDICTING TRENDS PAPER.
-
-Consider the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a DNN with $L$ layers, and weight matrices $\mathbf{W}_{l}$ and bias vectors $\mathbf{b}_{l}$, as
-the minimization of a general loss function $\mathcal{L}$ over the training data instances and labels, $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$.
-For a typical supervised classification problem, the goal of training is to construct (or learn) $\mathbf{W}_{l}$ and $\mathbf{b}_{l}$ that capture correlations in the data, in the sense of solving
-\begin{equation}
-%%%\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N}\left(\mathcal{L}(E_{DNN}(\mathbf{x}_{i})-y_{i})\right)  ,
-\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N} \mathcal{L}(E_{DNN}(\mathbf{x}_{i}),y_{i})   ,
-\end{equation}
-where the loss function $\mathcal{L}(\cdot,\cdot)$ can take on a myriad of forms, and where the energy (or optimization) landscape function
-\begin{equation}
-%E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})
-E_{DNN} = f(\mathbf{x}_{i} ; \mathbf{W}_{1},\ldots,\mathbf{W}_{L},\mathbf{b}_{1},\ldots,\mathbf{b}_{L})
-\label{eqn:dnn_energy}
-\end{equation}
-depends parametrically on the weights and biases.
-For a trained model, the form of the function $E_{DNN}$ does not explicitly depend on the data (but it does explicitly depend on the weights and biases).
-The function $E_{DNN}$ maps data instance vectors ($\mathbf{x}_i$ values) to predictions ($y_{i}$ labels), and thus the output of this function does depend on the data.
-Therefore, one can analyze the form of $E_{DNN}$ in the absence of any training or test~data.
-
-Test accuracies have been reported online for publicly-available pretrained pyTorch models~\cite{osmr}.
-These models have been trained and evaluated on labeled data $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$, using standard techniques.  
-We do not have access to this data, and we have not trained any of the models ourselves. 
-Our methodological approach is thus similar to a statistical meta-analysis, common in biomedical research, but uncommon in ML.
-%
-Computations were performed with the publicly-available \texttt{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
-
-\paragraph{Metrics for DNN Weight Matrices.}
-
-Our approach involves analyzing individual DNN weight matrices, for (depending on the architecture) fully-connected and/or convolutional layers.
-Each DNN layer contains one or more layer 2D  $N_{l}\times M_{l}$ weight matrices, $\mathbf{W}_{l}$, or pre-activation maps, $\mathbf{W}_{i,l}$, e.g., extracted from 2D Convolutional layers, where $N > M$.
-See the Supplementary Information
-%~\ref{sxn:appendix} 
-for details.
-(We may drop the $i$ and/or $i,l$ subscripts below.)
-The best performing quality metrics depend on the norms and/or spectral properties of each weight matrix,
-$\mathbf{W}$, and/or, equivalently, it's empirical correlation matrix, $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
-To evaluate the quality of state-of-the-art DNNs, 
-we consider the following metrics:
-\begin{eqnarray}
-& & \text{Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{X}\Vert_{F}=\sum\nolimits_{i=1}^{M} \lambda_{i}$ } \\
-& & \text{Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}^{2}=\Vert\mathbf{X}\Vert_{\infty}=\lambda_{max}$ } \\
-& & \text{Weighted Alpha: $\hat{\alpha}=\alpha\log\lambda_{max}$ } \\
-& & \text{$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{W}\Vert^{2\alpha}_{2\alpha}=\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum\nolimits_{i=1}^{M}\lambda_{i}^{\alpha}$. }
-\end{eqnarray}
-%To diagnose problems, 
-To perform diagnostics on potentially-problematic DNNs,
-we will decompose $\hat{\alpha}$ into its two components, $\alpha$ and $\lambda_{max}$.
-Here, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the $\mathbf{X}$, $\lambda_{max}$ is the maximum eigenvalue, and $\alpha$ is the fitted PL exponent. 
-These eigenvalues are squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, $\lambda_{i}=\sigma^{2}_{i}$.
-All four metrics can be computed easily from DNN weight matrices.
-The first two metrics are well-known in ML.
-The last two metrics deserve special mention, as they depend on an empirical parameter $\alpha$ that is the PL exponent that arises in the recently-developed Heavy Tailed Self Regularization (HT-SR) Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
-
-\paragraph{Overview of Heavy-Tailed Self-Regularization.}
-
-In the HT-SR Theory, one analyzes the eigenvalue spectrum, i.e., the Empirical Spectral Density (ESD), of the associated correlation matrices~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
-From this, one characterizes the amount and form of correlation, and therefore implicit self-regularizartion, present in the DNN's weight matrices.
-For each layer weight matrix $\mathbf{W}$, of size $N \times M$, construct the associated $M\times M$ (uncentered) correlation matrix $\mathbf{X}$. 
-Dropping the $L$ and $l,i$ indices, one~has
-$$
-\mathbf{X} = \frac{1}{N}\mathbf{W}^{T}\mathbf{W}.
-$$
-If we compute the eigenvalue spectrum of $\mathbf{X}$, i.e., $\lambda_i$ such that
-$  % $$
-\mathbf{X}\mathbf{v}_{i}=\lambda_{i}\mathbf{v}_{i} , 
-$  % $$
-then the ESD of eigenvalues, $\rho(\lambda)$, is just a histogram of the eigenvalues, formally written as
-%\begin{equation}
-$\rho(\lambda)=\sum\nolimits_{i=1}^{M}\delta(\lambda-\lambda_{i})  .$
-%\label{eqn:eigenval_hist}
-%\end{equation}
-Using HT-SR Theory, one characterizes the correlations in a weight matrix by examining its ESD, $\rho(\lambda)$.
-It can be well-fit to a truncated PL distribution, given~as
-\begin{equation}
-\rho(\lambda)\sim\lambda^{-\alpha}  ,
-\label{eqn:eigenval_pl}
-\end{equation}
+In the HT-SR Theory, one analyzes the eigenvalue spectrum, i.e., the Empirical Spectral Density (ESD), of the associated correlation matrices~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
+From this, one characterizes the amount and form of correlation present in the DNN's weight matrices.
+For each layer weight matrix $\mathbf{W}$, if we compute the eigenvalue spectrum of 
+$ \mathbf{X} = \frac{1}{N}\mathbf{W}^{T}\mathbf{W} $
+then HT-SR theory characterizes the correlations in a weight matrix by examining its ESD, $\rho(\lambda)$.
+Empirically, the ESD can be well-fit to a truncated power law (PL) distribution, given as
+$\rho(\lambda)\sim\lambda^{-\alpha} $,
 which is (at least) valid within a bounded range of eigenvalues $\lambda\in[\lambda^{min},\lambda^{max}]$.
 
-The original work on HT-SR Theory considered a small number of NNs, including AlexNet and InceptionV3. 
-It showed that for nearly every $\mathbf{W}$, the (bulk and tail) of the ESDs can be fit to a truncated PL, and that PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
+The original work on HT-SR Theory showed that for nearly every $\mathbf{W}$, i.e., among publicly-available SOTA models, the (bulk and tail) of the ESDs can be fit to a truncated PL, and that PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
 %
-%%As for the mechanism responsible for these properties, statistical physics offers several possibilities~\cite{SornetteBook,nishimori01}, e.g., self organized criticality~\cite{SOC87,SOCat25yrs} or multiplicative noise in the stochastic optimization algorithms used to train these models~\cite{HodMah20A_TR,SorCon97}.
-%%Alternatively, related techniques have been used to analyze correlations and information propogation in actual spiking neurons~\cite{SYYRP11,YKYP14}.
+%%As for the mechanism responsible for these properties, statistical physics offers several possibilities~\citep{SornetteBook,nishimori01}, e.g., self organized criticality~\citep{SOC87,SOCat25yrs} or multiplicative noise in the stochastic optimization algorithms used to train these models~\citep{HodMah20A_TR,SorCon97}.
+%%Alternatively, related techniques have been used to analyze correlations and information propogation in actual spiking neurons~\citep{SYYRP11,YKYP14}.
 %%Our meta-analysis does not require knowledge of mechanisms; and it is not even clear that one mechanism is responsible for every case.
@@
 Crucially, HT-SR Theory predicts that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models.
-The notion of ``size scale'' is well-defined in physical systems, to which this style of analysis is usually applied, but it is less well-defined in CV and NLP applications.
-Informally, it would correspond to pixel groups that are at a greater distance in some metric, or between sentence parts that are at a greater distance in text.
-Relatedly, previous work observed that smaller exponents $\alpha$ correspond to more implicit self-regularization and better generalization, and that we expect a linear correlation between $\hat{\alpha}$ and model quality~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
+%%The notion of ``size scale'' is well-defined in physical systems, to which this style of analysis is usually applied, but it is less well-defined in CV and NLP applications.
+%%Informally, it would correspond to pixel groups that are at a greater distance in some metric, or between sentence parts that are at a greater distance in text.
+%%Relatedly, previous work observed that smaller exponents $\alpha$ correspond to more implicit self-regularization and better generalization, and that we expect a linear correlation between $\hat{\alpha}$ and model quality~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
 
-\paragraph{DNN Empirical Quality Metrics.}
+\section{Overall approach}
 
-For norm-based metrics, we use the average of the log norm, to the appropriate power.
-Informally, this amounts to assuming that the layer weight matrices are statistically independent, in which case we can estimate the model complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm (which resembles a data dependent VC complexity),
-\begin{equation}
-\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert \times \cdots \times \Vert\mathbf{W}_{L}\Vert ,
-\end{equation}
-where $\Vert\cdot\Vert$ is a matrix norm.   
-The log complexity,
-\begin{equation}
-\label{eqn:eqn:sum_log_norm}
-\log\mathcal{C} \sim \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert + \cdots + \log\Vert\mathbf{W}_{L}\Vert = \sum\nolimits_l \log\Vert\mathbf{W}_{l}\Vert ,
-\end{equation}
- takes the form of an average Log Norm.
-For the Frobenius Norm metric and Spectral Norm metric, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly (since, when taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor).
-
-The Weighted Alpha metric is an average of $\alpha_l$ over all layers $l \in \{1,\ldots,l\}$, weighted by the size, or scale, or each matrix,
-\begin{equation}
-\hat{\alpha} = \dfrac{1}{L}\sum_l \alpha_l\log\lambda_{max,l}\approx\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle    ,
-\end{equation}
-where $L$ is the total number of layer weight matrices.
-The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a much smaller and more limited set of models than we consider here.
-
-Based on this, in this paper, we introduce and evaluate the $\alpha$-Shatten Norm metric,
-\begin{equation}
-\label{eqn:sum_log_alpha_norm_alpha}
-\sum\nolimits_l \log \Vert\mathbf{X}_l\Vert_{\alpha_l}^{\alpha_l} 
-=
-\sum\nolimits_l \alpha_l \log \Vert\mathbf{X}_l\Vert_{\alpha_l} .
-\end{equation}
-For the $\alpha$-Shatten Norm metric, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it cannot be taken out of the sum.
-For small $\alpha$, the Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation;
-%% \cite{MM20_unpub_work}; 
-and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm.
-
-Finally, although it does less well for predicting trends in state-of-the-art model series, e.g., as depth changes, the average value of $\alpha$, i.e., 
-\begin{equation}
-\bar{\alpha} = \dfrac{1}{L}\sum_l \alpha_l = \langle\alpha\rangle    ,
-\label{eqn:alpha_bar}
-\end{equation}
-can be used to perform model diagnostics, to identify problems that cannot be detected by examining training/test accuracies, and to discriminate poorly-trained models from well-trained~models.
-
-One determines $\alpha$ for a given layer by fitting the ESD of that layer's weight matrix to a truncated PL, using the commonly accepted Maximum Likelihood method~\cite{CSN09_powerlaw,ABP14}.
-%%This method works very well for exponents between $\alpha\in(2,4)$; and it is adequate, although imprecise, for smaller and especially larger $\alpha$~\cite{newman2005_zipf}. 
+Test accuracies have been reported online for publicly-available pretrained pyTorch models~\citep{osmr}.
+These models have been trained and evaluated on labeled data, using standard techniques.  
+We do not have access to this data, and we have not trained any of the models ourselves. 
+Our methodological approach is thus similar to a statistical meta-analysis, common in biomedical research, but uncommon in machine learning.
 %
-Operationally, $\alpha$ is determined by using the \texttt{WeightWatcher} tool~\cite{weightwatcher_package} to fit the histogram of eigenvalues, $\rho(\lambda)$, to a truncated PL, 
-%%\begin{equation}
-%%\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\in[\lambda_{min},\lambda_{max}] ,
-%%\end{equation}
-%%where $\lambda_{max}$ is the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$, and 
-%%where $\lambda_{min}$ is selected automatically to yield the best (in the sense of minimizing the K-S distance) PL fit.
-%%Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.
-%%See Figure~\ref{fig:ww} for an illustration.
-
-%\begin{figure}[t]
-%    \centering
-%    %\includegraphics[width=15.0cm]{img/WeightWatcher_v2}
-%    \includegraphics[width=15.0cm]{img/WeightWatcher_v3}
-%    \caption{Schematic of analyzing DNN layer weight matrices $\mathbf{W}$.  
-%             Given an individual layer weight matrix $\mathbf{W}$, from either a fully-connected layer or a convolutional layer, perform a Singular Value Decomposition (SVD) to obtain $\mathbf{W} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^{T}$, and examine the histogram of eigenvalues of $\mathbf{W}^{T}\mathbf{W}$.
-%             Norm-based metrics and PL-based metrics (that depend on fitting the histogram of eigenvalues to a truncated PL) can be used to compare models.
-%             For example, one can analyze one layer of a pre-trained model, compare multiple layers of a pre-trained model, make comparisons across model architectures, monitor neural network properties during training, etc.
-%}
-%    \label{fig:ww}
-%\end{figure}
-
-To avoid confusion, let us clarify the relationship between $\alpha$ and $\hat{\alpha}$.  
-We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated PL, parameterized by 2 values: the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.
-The PL exponent $\alpha$ measures the amount of correlation in a DNN layer weight matrix $\mathbf{W}$. 
-It is valid for $\lambda\le\lambda_{max}$, and it is scale-invariant, i.e., it does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.
-The $\lambda_{max}$ is a measure of the size, or scale, of $\mathbf{W}$.
-Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric $\hat{\alpha}$ for the entire~DNN.
-We will see that for well-trained CV and NLP models, $\hat{\alpha}$ performs quite well and as expected, but for CV and NLP models that are potentially-problematic or less well-trained, metrics that depend on the scale of the problem can perform anomalously.
-In these cases, separating $\hat{\alpha}$ into its two components, $\alpha$ and $\lambda_{max}$, and examining the distributions of each, can be helpful.
+To evaluate the quality of SOTA DNNs, we consider the following metrics:
+Frobenius Norm, $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{X}\Vert_{F}=\sum\nolimits_{i=1}^{M} \lambda_{i}$; 
+Spectral Norm, $\Vert\mathbf{W}\Vert_{\infty}^{2}=\Vert\mathbf{X}\Vert_{\infty}=\lambda_{max}$;
+Weighted Alpha, $\hat{\alpha}=\alpha\log\lambda_{max}$; and
+$\alpha$-Shatten Norm, $\Vert\mathbf{W}\Vert^{2\alpha}_{2\alpha}=\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum\nolimits_{i=1}^{M}\lambda_{i}^{\alpha}$. 
+%\begin{eqnarray*}
+%& & \text{Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{X}\Vert_{F}=\sum\nolimits_{i=1}^{M} \lambda_{i}$ } \\
+%& & \text{Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}^{2}=\Vert\mathbf{X}\Vert_{\infty}=\lambda_{max}$ } \\
+%& & \text{Weighted Alpha: $\hat{\alpha}=\alpha\log\lambda_{max}$ } \\
+%& & \text{$\alpha$-Shatten Norm: $\Vert\mathbf{W}\Vert^{2\alpha}_{2\alpha}=\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum\nolimits_{i=1}^{M}\lambda_{i}^{\alpha}$. }
+%\end{eqnarray*}
+To perform diagnostics on potentially-problematic DNNs, we will decompose $\hat{\alpha}$ into its two components, $\alpha$ and $\lambda_{max}$.
+One determines $\alpha$ for a given layer by fitting the ESD of that layer's weight matrix to a truncated PL, using the commonly accepted Maximum Likelihood method~\citep{CSN09_powerlaw,ABP14}.
+Computations were performed with the publicly-available \texttt{WeightWatcher} tool (version 0.2.7)~\citep{weightwatcher_package}.
 
@@
 
-XXX.  THIS IS A VERY PRACTICAL PROBLEM THAT OUR THEORETICAL APPROACH, STRONGLY COUPLED AND MOTIVATED BY EMPIRICAL PRACTICE, CAN SOLVE.
+%XXX.  ABSTRACT FROM \citep{MM20a_trends_TR}: ``Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data''
 
-\cite{MM20a_trends_TR} is Charles: ``Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data''
-
-XXX.  ABSTRACT FROM \cite{MM20a_trends_TR}:
 In many applications, one works with neural network models trained by someone else.
@@
 Given one or many pretrained models, it is a challenge to say anything about the expected performance or quality of the models.
-Here, we address this challenge by providing a detailed meta-analysis of hundreds of publicly-available pretrained models.
-We examine norm based capacity control metrics as well as power law based metrics from the recently-developed Theory of Heavy-Tailed Self Regularization.
-We find that norm based metrics correlate well with reported test accuracies for well-trained models, but that they often cannot distinguish well-trained versus poorly-trained models.
-We also find that power law based metrics can do much better---quantitatively better at discriminating among series of well-trained models with a given architecture; and qualitatively better at discriminating well-trained versus poorly-trained models.
-These methods can be used to identify when a pretrained neural network has problems that cannot be detected simply by examining training/test accuracies.
-
-XXX.  CITE ALSO BARTLETT AND POGGIO AS DOING SOMETHING LOOSELY SIMILAR, AND THE GOOGLE CONTEST AS DOING SOMETHING LOOSELY SIMILAR BUT DIFFERENT.
-
-\cite{BFT17_TR} is Bartlett: ``Spectrally-normalized margin bounds for neural networks''
-
-\cite{LMBx18_TR} is Poggio: ``A surprising linear relationship predicts test performance in deep networks''
-
-\subsection{Introduction OF PREDICTING TRENDS PAPER}
-\label{sxn:intro}
-
-XXX.  THIS IS THE INTRO FROM THE PREDICTING TRENDS PAPER.
-
-A common problem in machine learning (ML) 
-is to evaluate the quality of a given model.
-A popular way to accomplish this
-is to train a model and then evaluate its training/testing error.
-There are many problems with this approach.
-The training/testing curves give very limited insight into the overall properties of the model; 
-they do not take into account the (often large human and CPU/GPU) time for hyperparameter fiddling;
-they typically do not correlate with other properties of interest such as robustness or fairness or interpretability; 
-and so on.
-A related problem, in particular in industrial-scale artificial intelligence (AI), arises when the model user is not the model developer.
-Then, one may not have access to the training data or the testing data.
-Instead, one may simply be given a model that has already been trained---a pretrained model---and need to use it as-is, or to fine-tune and/or compress it and then use it.
-
-Na\"{\i}vely---but in our experience commonly, among ML practitioners and ML theorists---if one does not have access to training or testing data, then one can say absolutely nothing about the quality of a ML model.
-This may be true in worst-case theory, but models are used in practice, and there is a need for a practical theory to guide that practice.
-Moreover, if ML is to become an industrial process, then that process will become compartmentalized in order to scale: some groups will gather data, other groups will develop models, and other groups will use those models.
-Users of models cannot be expected to know the precise details of how models were built, the specifics of data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized,~etc.
-
-Moreover, for many large scale, practical applications, there is no obvious way to define an ideal test metric. 
-For example, models that generate fake text or conversational chatbots may use a proxy, like perplexity, as a test metric.
-In the end, however, they really require human evaluation. 
-Alternatively, models that cluster user profiles, which are widely used in areas such as marketing and advertising, are unsupervised and have no obvious labels for comparison and/or evaluation.
-In these and other areas, ML objectives can be poor proxies for downstream goals.
-
-Most importantly, in industry, one faces unique practical problems such as determining whether one has enough data for a given model. 
-Indeed, high quality, labeled data can be very expensive to acquire, and this cost can make or break a project.
-Methods that are developed and evaluated on any well-defined publicly-available corpus of data, no matter how large or diverse or interesting, are clearly not going to be well-suited to address problems such as this.
-It is of great practical interest to have metrics to evaluate the quality of a trained model---in the absence of training/testing data and without any detailed knowledge of the training/testing process.
-There is a need for a practical theory for pretrained models which can predict how, when, and why such models can be expected to perform well or~poorly.
-
-In the absence of training and testing data, obvious quantities to examine are the weight matrices of pretrained models, e.g., 
-properties such as norms of weight matrices and/or parameters of Power Law (PL) fits of the eigenvalues of weight matrices.
-Norm-based metrics have been used in traditional statistical learning theory to bound capacity and construct regularizers; and PL fits are based on statistical mechanics approaches to deep neural networks (DNNs).
-While we use traditional norm-based and PL-based metrics, our goals are not the traditional goals.
-Unlike more common ML approaches, we do not seek a bound on the generalization (e.g., by evaluating training/test errors), we do not seek a new regularizer, and we do not aim to evaluate a single model (e.g., as with hyperparameter optimization).
-Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves.
-In both cases, one can ask whether it is possible to predict trends in the quality of pretrained DNN models without access to training or testing data.
+We have applied HT-ST Theory to provide a detailed meta-analysis of hundreds of publicly-available pretrained models, examining both 
+norm based capacity control metrics from statistical learning theory as well as PL based metrics from HT-ST theory.
 
@@
 For all these models, we have no access to training data or testing data, and we have no specific knowledge of the training/testing protocols.
+
+We find that norm based metrics correlate well with reported test accuracies for well-trained models, but that they often cannot distinguish well-trained versus poorly-trained models.
+We also find that power law based metrics can do much better---quantitatively better at discriminating among series of well-trained models with a given architecture; and qualitatively better at discriminating well-trained versus poorly-trained models.
+These methods can be used to identify when a pretrained neural network has problems that cannot be detected simply by examining training/test accuracies.
 %
-Here is a summary of our main results.
-First, norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models.
-Second, norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.).
-For example, they may exhibit what we call Scale Collapse for these models. 
-Third, PL-based metrics can do much better at predicting quality trends in pretrained CV/NLP models.  
-In particular, 
-a weighted PL exponent (weighted by the log of the spectral norm of the corresponding layer) is 
-quantitatively better at discriminating among a series of well-trained versus very-well-trained models
-within a given architecture series; and
-the (unweighted) average PL exponent is 
-qualitatively better at discriminating well-trained versus poorly-trained models.
-Fourth, PL-based metrics can also be used to characterize fine-scale model properties, including what we call layer-wise Correlation Flow, in well-trained and poorly-trained models; and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
+%Here is a summary of our main results.
+%First, norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models.
+%Second, norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.).
+%For example, they may exhibit what we call Scale Collapse for these models. 
+%Third, PL-based metrics can do much better at predicting quality trends in pretrained CV/NLP models.  
+%In particular, 
+%a weighted PL exponent (weighted by the log of the spectral norm of the corresponding layer) is 
+%quantitatively better at discriminating among a series of well-trained versus very-well-trained models
+%within a given architecture series; and
+%the (unweighted) average PL exponent is 
+%qualitatively better at discriminating well-trained versus poorly-trained models.
+%Fourth, PL-based metrics can also be used to characterize fine-scale model properties, including what we call layer-wise Correlation Flow, in well-trained and poorly-trained models; and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
 %
-Our work provides a theoretically-principled empirical evaluation---by far the largest, most detailed, and most comprehensive to date---and the theory we apply was developed previously~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
-Performing such a meta-analysis of previously-published work is common in certain areas, but it is quite rare in ML, where the emphasis is on developing better training protocols.
+Our work provides a theoretically-principled empirical evaluation---by far the largest, most detailed, and most comprehensive to date---and it uses HT-SR theory.%
+\footnote{Performing such a meta-analysis of previously-published work is common in certain areas, but it is quite rare in ML, where the emphasis is on developing better training protocols.}
+
+%\citep{BFT17_TR} is Bartlett: ``Spectrally-normalized margin bounds for neural networks''
+%\citep{LMBx18_TR} is Poggio: ``A surprising linear relationship predicts test performance in deep networks''
+There is related work \citep{BFT17_TR,LMBx18_TR}, but that work provides generalization bounds, modifies the models, requires access to data, etc., and so is quite different.
 
@@
 
-\cite{JNBx19_fantastic_TR} is the terrible fantastic paper: ``Fantastic Generalization Measures and Where to Find Them''
+\citep{JNBx19_fantastic_TR} is the terrible fantastic paper: ``Fantastic Generalization Measures and Where to Find Them''
 
-\cite{JFYx20_contest_v10} is the contest: ``{NeurIPS} 2020 Competition: Predicting Generalization in Deep Learning (Version 1.0)''
+\citep{JFYx20_contest_v10} is the contest: ``{NeurIPS} 2020 Competition: Predicting Generalization in Deep Learning (Version 1.0)''
 
-\cite{BHO75} is Bickel: ``Sex Bias in Graduate Admissions: Data from Berkeley''
+\citep{BHO75} is Bickel: ``Sex Bias in Graduate Admissions: Data from Berkeley''
 
@@
 
+\michael{I may cut this stuff.}
 We have shown that weight matrix analysis derived from HT-SR Theory can be used to predict trends in the quality of state-of-the-art neural networks without access to training or testing data as well as identify Simpson's-like paradoxes in publicly-available NN models.
-The HT-SR Theory underlying our analysis is quite different than theoretical approaches popular in ML, and indeed it is quite different than typical statistical mechanics approaches to NNs \cite{BKPx20}.
+The HT-SR Theory underlying our analysis is quite different than theoretical approaches popular in ML, and indeed it is quite different than typical statistical mechanics approaches to NNs \citep{BKPx20}.
 Recently, however, work motivated by HT-SR Theory has 
-demonstrated that multiplicative noise in stochastic optimization leads to heavy tails~\cite{HodMah20A_TR,GSZ20_TR}; 
-provided a random matrix analysis of random {F}ourier features, going beyond the {G}aussian kernel, to provide a precise phase transition and illustrate the corresponding double descent~\cite{ZCM20_TR};
-provided exact expressions for double descent and implicit regularization via a notion of surrogate random design \cite{DLM19_Exact_TR}; 
-demonstrated that good classifiers are abundant in the interpolating regime \cite{TKM20_TR};
-shown how injecting noise into hidden states of recurrent NNs leads to implicit regularization~\cite{LEHM21_TR,Mah12}; and 
-shown that analogous phase transitions hold for the {C}olumn {S}ubset {S}election and the {N}ystrom method~\cite{DKM20_TR}. 
+demonstrated that multiplicative noise in stochastic optimization leads to heavy tails~\citep{HodMah20A_TR,GSZ20_TR}; 
+provided a random matrix analysis of random {F}ourier features, going beyond the {G}aussian kernel, to provide a precise phase transition and illustrate the corresponding double descent~\citep{ZCM20_TR};
+provided exact expressions for double descent and implicit regularization via a notion of surrogate random design \citep{DLM19_Exact_TR}; 
+demonstrated that good classifiers are abundant in the interpolating regime \citep{TKM20_TR};
+shown how injecting noise into hidden states of recurrent NNs leads to implicit regularization~\citep{LEHM21_TR,Mah12}; and 
+shown that analogous phase transitions hold for the {C}olumn {S}ubset {S}election and the {N}ystrom method~\citep{DKM20_TR}. 
 We expect that the empirical results identified by tools from HT-SR Theory will provide guidance and constraints for future development of overparameterized machine learning models.

### 2021-02-18 submission/TOPML_Workshop/paper.tex
small edit
changes to papaer first two sections and add table

@@
 
-\begin{abstract}
-Understanding state-of-the-art overparameterized neural networks requires detailed empirical studies as well as the development of predictive theory in light of those studies. 
-Building on the theory of Heavy-Tailed Self-Regularization, we show that correlations in the weight matrices of state-of-the-art overparameterized neural networks are well-described by heavy-tailed random matrix theory.
-Using the publicly-available WeightWatcher tool, we perform spectral analylsis on these weight matrices, and we show that fitted truncated power law exponents are strongly predictive of model quality, for publicly-available state-of-the-art pre-trained models, even in the absence of training or testing data.
-We also show that these metrics can also be used to identify a Simpson's paradox in the data from a recent contest aimed at developing metrics to predice generalization in deep learning.
-\end{abstract}
+%\begin{abstract}
+%\end{abstract}
 
-\section{Introduction}
+\section*{Introduction}
 
-\michael{Here is Charles core dump of intro; still to refine and incorporate with rest of text.}
-We present our Theory of Heavy Tailed Self Regularization (HTSR) and
-its application to the (overparameterized)  Deep Neural Networks (DNNs) 
-to  predict the generalization accuracy and improve fine-tuning
-of pretrained models by capturing the dominant correlation structure.
-We analyze the correlations using techniques from 
-Heavy Tailed Random Matrix Theory (HT-RMT)
-and from  strongly correlated systems from theoretical
-chemistry and  physics.  
-We show how to predict trends in the reported test accuracies
-across hundreds of pretrained DNNs without needing access
-to training or test data by charactering the heavy tailed
-structure of the empirical spectral densities (ESD) of 
-the layer weight matrices.
-In doing so, we show that different metrics, which
-either depend on the shape or scale of the ESDs,
-exhbit a Simpson's paradox when varying DNN model depth
-vs regularization hyperparameters.
-In particular,  the familar scale-based spectral norm
-metric behave opposite theoretical predictions,
-whereas our HT alpha-metric acts as predicted.
-Moreover, the paradoxical behavior  can be resolved
-by combining these shape and scale metrics with
-our HT alpha-hat metric.
-Finally,  And with access to training data, we show how to
-smooth and capture the low rank structure to both accurately
+Understanding state-of-the-art overparameterized deep neural networks (DNNs) requires detailed empirical studies as well as the development of predictive theory in light of those studies. 
+We use the theory of Heavy-Tailed Self-Regularization (HT-SR) to predict the generalization accuracy and improve the fine-tuning of (overparameterized) pretrained DNN models by capturing the dominant correlation structure.
+HT-SR is based on Heavy-Tailed Random Matrix Theory (HT-RMT) and strongly-correlated systems theory from theoretical chemistry and theoretical physics, and methods based on it are implemented in the publicly-available WeightWatcher tool.
+Using WeightWatcher, we can characterize HT structure in the empirical spectral densities (ESDs) of the layer weight matrices, and we can use this to predict trends in the reported test accuracies across hundreds of pretrained DNNs---without needing access to training or test data.
+Going beyond popular norm-based metrics, we show that different metrics, which either depend on the scale (norms) or shape (fitted truncated power law (PL) exponents) of the ESDs, exhbit a Simpson's paradox when varying DNN model depth versus solver hyperparameters.
+The familar scale-based spectral norm metric behaves opposite to theoretical predictions, while the HT $\alpha$-metric (the exponent in a truncated PL fit) acts as predicted.
+The paradoxical behavior  can be resolved by combining these shape and scale metrics with the HT $\hat{\alpha}$ metric.
+Finally, and with access to training data, we show how to
+smooth out and capture the low rank structure to both accurately
 predict the generalization errors and potentially improve
-finetuning of pretrained DNNs.
+fine-tuning of pretrained DNNs.
 
-\section{Statistical Mechanics and Heavy-Tailed Self Regularization}
+%\michael{Here is Charles initial core dump of intro; still to refine and incorporate with rest of text.}
+%We present our Theory of Heavy Tailed Self Regularization (HTSR) and
+%its application to the (overparameterized)  Deep Neural Networks (DNNs) 
+%to  predict the generalization accuracy and improve fine-tuning
+%of pretrained models by capturing the dominant correlation structure.
+%We analyze the correlations using techniques from 
+%Heavy Tailed Random Matrix Theory (HT-RMT)
+%and from  strongly correlated systems from theoretical
+%chemistry and  physics.  
+%We show how to predict trends in the reported test accuracies
+%across hundreds of pretrained DNNs without needing access
+%to training or test data by charactering the heavy tailed
+%structure of the empirical spectral densities (ESD) of 
+%the layer weight matrices.
+%In doing so, we show that different metrics, which
+%either depend on the shape or scale of the ESDs,
+%exhbit a Simpson's paradox when varying DNN model depth
+%vs regularization hyperparameters.
+%In particular,  the familar scale-based spectral norm
+%metric behaves opposite to theoretical predictions,
+%whereas our HT alpha-metric acts as predicted.
+%Moreover, the paradoxical behavior  can be resolved
+%by combining these shape and scale metrics with
+%our HT alpha-hat metric.
+%Finally,  And with access to training data, we show how to
+%smooth and capture the low rank structure to both accurately
+%predict the generalization errors and potentially improve
+%finetuning of pretrained DNNs.
+
+\section*{Statistical Mechanics and Heavy-Tailed Self-Regularization}
 
@@
 
-Recently, Random Matrix Theory (RMT) was applied to analyze empirically the weight matrices of Deep Neural Networks (DNNs), including both production quality, pre-trained models such as AlexNet and Inception, and smaller models trained from scratch, such as LeNet5 and a miniature-AlexNet~\citep{MM18_TR} (see also \citep{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM20_SDM, MM19_KDD}).
-%%Empirical results clearly indicate that the empirical spectral density (ESD) of DNN layer matrices displays signatures of traditionally-regularized statistical models, even in the absence of exogenously specifying traditional forms of explicit regularization, such as dropout or weight norm~constraints.
-%%
+Recently, RMT was applied to analyze empirically the weight matrices of DNNs, including both production quality, pretrained models such as AlexNet and Inception, and smaller models trained from scratch, such as LeNet5 and a miniature-AlexNet~\citep{MM18_TR} (see also \citep{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM20_SDM, MM19_KDD}).
 Building on these empirical results, 
-%%and using relatively recent results in RMT, most notably its extension to Universality classes of heavy-tailed matrices, 
-\citet{MM18_TR} then developed a phenomenological (statistical mechanics based, as opposed to statistical learning theory based) theory to identify ``5+1 Phases of Training,'' corresponding to increasing amounts of implicit self-regularization.
-These phases can be observed during the training process as well as in the final learned DNNs.
+\citet{MM18_TR} developed a phenomenological (statistical mechanics based, as opposed to statistical learning theory based) theory to identify ``5+1 Phases of Training,'' corresponding to increasing amounts of implicit self-regularization.
+These phases can be observed during the training process as well as in the final learned (pretrained) DNNs.
 For smaller and/or older DNNs, this implicit self-regularization is like traditional Tikhonov regularization, in that there is a ``size scale'' separating signal from noise.
@@
 
-In the HT-SR Theory, one analyzes the eigenvalue spectrum, i.e., the Empirical Spectral Density (ESD), of the associated correlation matrices~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
+In the HT-SR theory, one analyzes the eigenvalue spectrum, i.e., the ESD, of the associated correlation matrices~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
 From this, one characterizes the amount and form of correlation present in the DNN's weight matrices.
@@
 then HT-SR theory characterizes the correlations in a weight matrix by examining its ESD, $\rho(\lambda)$.
-Empirically, the ESD can be well-fit to a truncated power law (PL) distribution, given as
-$\rho(\lambda)\sim\lambda^{-\alpha} $,
-which is (at least) valid within a bounded range of eigenvalues $\lambda\in[\lambda^{min},\lambda^{max}]$.
-
-The original work on HT-SR Theory showed that for nearly every $\mathbf{W}$, i.e., among publicly-available SOTA models, the (bulk and tail) of the ESDs can be fit to a truncated PL, and that PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
+Empirically, the ESD can be well-fit to a truncated PL distribution, 
+$\rho(\lambda)\sim\lambda^{-\alpha} $.
+%
+The original work on HT-SR theory showed that for nearly every $\mathbf{W}$, i.e., among publicly-available SOTA models, the (bulk and tail) of the ESDs can be fit to a truncated PL, and that PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
 %
@@
 %
-Crucially, HT-SR Theory predicts that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models.
+Crucially, HT-SR theory predicts that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models.
 %%The notion of ``size scale'' is well-defined in physical systems, to which this style of analysis is usually applied, but it is less well-defined in CV and NLP applications.
@@
 
-\section{Overall approach}
+\section*{Overall approach}
 
@@
 
-\section{Predicting Trends in the Quality of State-of-the-art Models}
+\section*{Predicting Trends in the Quality of State-of-the-art Models}
 
@@
 
-\section{A Simpson's Paradox---and the Need for Finer-scale Analysis}
+\begin{table}[t]
+\small
+\begin{center}
+\begin{tabular}{|p{0.83in}|c|c|c|c|}
+\hline
+  & $\log\Vert\cdot\Vert^{2}_{F}$ & $\log\Vert\cdot\Vert^{2}_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
+\hline
+
+RMSE (mean) & 4.84 & 5.57 & 4.58& 4.55 \\
+RMSE (std) & 9.14 & 9.16 & 9.16& 9.17 \\
+\hline
+R2 (mean) & 3.9 & 3.85 & 3.89& 3.89 \\
+R2 (std) & 9.34 & 9.36 & 9.34& 9.34 \\
+\hline
+Kendal-$\tau$ (mean) & 3.84 & 3.77 & 3.86& 3.85 \\
+Kendal-$\tau$ (std) & 9.37 & 9.4 & 9.36& 9.36 \\
+\hline
+
+\hline
+\end{tabular}
+\end{center}
+\caption{Comparison of linear regressions for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering \
+108 (out of $>400$) different pretrained~DNNs.
+}
+\label{table:results}
+\end{table}
+
+\section*{A Simpson's Paradox---and the Need for Finer-scale Analysis}
 
@@
 
-\charles{We apply our theory of Heavy Tailed Self-Regularization (HTSR) to predict the generalization accuracies of the pre-trained models provided in this contest, as a function of both model depth (L) and the solver hyperparameters (batch size, dropout, weight decay), and compare quality metrics that describe both the heavy-tailed shape and norm-based scale of the model layer weight matrices.
+\charles{We apply our theory of Heavy Tailed Self-Regularization (HTSR) to predict the generalization accuracies of the pretrained models provided in this contest, as a function of both model depth (L) and the solver hyperparameters (batch size, dropout, weight decay), and compare quality metrics that describe both the heavy-tailed shape and norm-based scale of the model layer weight matrices.
 In doing so, we identify what amounts to a Simpson’s paradox, in which
@@
 
-\section{Conclusion}
+\section*{Conclusion}
 

### 2021-02-18 submission/TOPML_Workshop/paper.tex
minor changes

@@
 We use the theory of Heavy-Tailed Self-Regularization (HT-SR) to predict the generalization accuracy and improve the fine-tuning of (overparameterized) pretrained DNN models by capturing the dominant correlation structure.
-HT-SR is based on Heavy-Tailed Random Matrix Theory (HT-RMT) and strongly-correlated systems theory from theoretical chemistry and theoretical physics, and methods based on it are implemented in the publicly-available WeightWatcher tool.
-Using WeightWatcher, we can characterize HT structure in the empirical spectral densities (ESDs) of the layer weight matrices, and we can use this to predict trends in the reported test accuracies across hundreds of pretrained DNNs---without needing access to training or test data.
+HT-SR is based on Heavy-Tailed Random Matrix Theory (HT-RMT) and strongly-correlated systems theory from theoretical chemistry and theoretical physics, and methods based on it are implemented in the publicly-available \texttt{WeightWatcher} tool.
+Using \texttt{WeightWatcher}, we can characterize HT structure in the empirical spectral densities (ESDs) of the layer weight matrices, and we can use this to predict trends in the reported test accuracies across hundreds of pretrained DNNs---without needing access to training or test data.
 Going beyond popular norm-based metrics, we show that different metrics, which either depend on the scale (norms) or shape (fitted truncated power law (PL) exponents) of the ESDs, exhbit a Simpson's paradox when varying DNN model depth versus solver hyperparameters.
@@
 For smaller and/or older DNNs, this implicit self-regularization is like traditional Tikhonov regularization, in that there is a ``size scale'' separating signal from noise.
-For state-of-the-art (SOTA) DNNs, however, they identified a novel form of \emph{Heavy-Tailed Self-Regularization} (HT-ST).
+For state-of-the-art (SOTA) DNNs, however, they identified a novel form of \emph{Heavy-Tailed Self-Regularization} (HT-SR).
 This is similar to the self-organization seen in the statistical physics of disordered systems (such as classical models of actual neural activity); and 
@@
 
-In many applications, one works with neural network models trained by someone else.
+In many applications, one works with DNN models trained by someone else.
 For such pretrained models, one may not have access to training data or test data.
 Moreover, one may not know details about the model, e.g., the specifics of the training data, the loss function, the hyperparameter values, etc.
+(So, one is \emph{not} training models, one is validating pretrained models.)
 Given one or many pretrained models, it is a challenge to say anything about the expected performance or quality of the models.
-We have applied HT-ST Theory to provide a detailed meta-analysis of hundreds of publicly-available pretrained models, examining both 
-norm based capacity control metrics from statistical learning theory as well as PL based metrics from HT-ST theory.
+We have applied HT-SR theory to provide a detailed meta-analysis of hundreds of publicly-available pretrained models, examining both norm based capacity control metrics (from statistical learning theory) as well as PL based metrics (from HT-SR theory).
 
-To answer this question, we provide a detailed empirical analysis, evaluating quality metrics for pretrained DNN models, and we do so at scale.
-Our approach may be viewed as a statistical meta-analysis of previously published work, where we consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
+We consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
 By now, there are many such state-of-the-art models that are publicly-available, e.g., 
-hundreds of pretrained models in CV ($\ge 500$) and NLP ($\approx 100$).%
-\footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
+hundreds of pretrained models in CV ($\ge 500$) and NLP ($\approx 100$).
 For all these models, we have no access to training data or testing data, and we have no specific knowledge of the training/testing protocols. 
+Thus, performing weight matrix analysis is a natural option.
 
@@
     \centering
-    \subfigure[Spectral Norm $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$]{
-      \includegraphics[width=4.0cm]{img/simpsons-task2-logspectralnorm.png}
+    \subfigure[Log Spectral Norm]{
+       \includegraphics[width=3.5cm]{img/simpsons-task2-logspectralnorm.png}
     }
-    \subfigure[Alpha Hat $\hat{\alpha}$]{
-     \includegraphics[width=4cm]{img/simpsons-task2-alpha_weighted.png}
+    \subfigure[Weighted Alpha $\hat{\alpha}$]{
+       \includegraphics[width=3.5cm]{img/simpsons-task2-alpha_weighted.png}
     }
-    \caption{Test accuracy as a function of the
-Log Spectral Norm $\log\Vert\mathbf{W}\Vert_{\infty}$ 
-and the HTSR Alpha Hat $\hat{\alpha}$ metrics
+    \caption{Test accuracy as a function of 
+             the Log Spectral Norm and 
+             the (HT-SR) Weighted Alpha ($\hat{\alpha}$) metrics,
              as model hyperparameters and depth are varied. 
-Models from the NeurIPS Predicting Generalization contest.}
+             Models are from the NeurIPS Predicting Generalization contest.
+             A Simpson's paradox is clearly visible.
+            }
     \label{fig:simpsons}

### 2021-02-18 submission/TOPML_Workshop/paper.tex
work on simpsons section, its pretty mature now

@@
 We find that norm based metrics correlate well with reported test accuracies for well-trained models, but that they often cannot distinguish well-trained versus poorly-trained models.
-We also find that power law based metrics can do much better---quantitatively better at discriminating among series of well-trained models with a given architecture; and qualitatively better at discriminating well-trained versus poorly-trained models.
-These methods can be used to identify when a pretrained neural network has problems that cannot be detected simply by examining training/test accuracies.
+We also find that PL based metrics can do much better---quantitatively better at discriminating among series of well-trained models with a given architecture; and qualitatively better at discriminating well-trained versus poorly-trained models.
+These methods can be used to identify when a pretrained DNN has problems that cannot be detected simply by examining training/test accuracies.
 %
@@
 %Fourth, PL-based metrics can also be used to characterize fine-scale model properties, including what we call layer-wise Correlation Flow, in well-trained and poorly-trained models; and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
+See Table~\ref{table:results} for an overview summary of our results.
+
+Our work provides a theoretically-principled empirical evaluation---by far the largest, most detailed, and most comprehensive to date---and it uses HT-SR theory.
 %
-Our work provides a theoretically-principled empirical evaluation---by far the largest, most detailed, and most comprehensive to date---and it uses HT-SR theory.%
-\footnote{Performing such a meta-analysis of previously-published work is common in certain areas, but it is quite rare in ML, where the emphasis is on developing better training protocols.}
-
 %\citep{BFT17_TR} is Bartlett: ``Spectrally-normalized margin bounds for neural networks''
@@
 
-XXX.  IF WANT A ONE SIZE FITS ALL METRIC, THEN CAN EXPECT A SIMPSONS PARADOX.  OUR TECHNIQUES CAN DIAGNOST THIS.  AND IN FACT WE SEE THAT.
-
-\michael{Charles, put in a paragraph and the figure for the Simpson's paradox here.}
-
-\charles{We apply our theory of Heavy Tailed Self-Regularization (HTSR) to predict the generalization accuracies of the pretrained models provided in this contest, as a function of both model depth (L) and the solver hyperparameters (batch size, dropout, weight decay), and compare quality metrics that describe both the heavy-tailed shape and norm-based scale of the model layer weight matrices.
-In doing so, we identify what amounts to a Simpson’s paradox, in which
-shape and scale metrics show opposite behavior, and a combine metric resolves the paradox}
+%\citep{JNBx19_fantastic_TR} is the terrible fantastic paper: ``Fantastic Generalization Measures and Where to Find Them''
+%
+%\citep{JFYx20_contest_v10} is the contest: ``{NeurIPS} 2020 Competition: Predicting Generalization in Deep Learning (Version 1.0)''
+%
+%\citep{BHO75} is Bickel: ``Sex Bias in Graduate Admissions: Data from Berkeley''
 
@@
              as model hyperparameters and depth are varied. 
-             Models are from the NeurIPS Predicting Generalization contest.
+             Models are from the NeurIPS 2020 ``Predicting Generalization'' contest.
              A Simpson's paradox is clearly visible.
@@
 
-\citep{JNBx19_fantastic_TR} is the terrible fantastic paper: ``Fantastic Generalization Measures and Where to Find Them''
+Subsequent to the development of HT-SR theory, and our initial results on analyzing pretrained models at scale, there was at NeurIPS 2020 a ``Predicting Generalization'' workshop and contest \citep{JNBx19_fantastic_TR, JFYx20_contest_v10}.
+This contest made publicly-available pretrained models.
+This corpus of models was much more narrow (in terms of domain) but more extensive (in terms of depth and hyperparameter variation) than the previous SOTA corpus we considered~\citep{osmr}.
 
-\citep{JFYx20_contest_v10} is the contest: ``{NeurIPS} 2020 Competition: Predicting Generalization in Deep Learning (Version 1.0)''
+In particular, we apply HT-SR theory to predict the generalization accuracies of the pretrained models provided in this contest, as a function of both model depth ($L$) and the solver hyperparameters (batch size, dropout, weight decay), and we compare quality metrics that describe the norm-based scale and the HT/PL shape of the model layer weight matrices.
+See Figure~\ref{fig:simpsons} for a summary.
+For norm-based metrics, we identify what amounts to a Simpson’s paradox \citep{BHO75}, in which scale and shape metrics show opposite behavior, depending on whether the data are partitioned by depth or solver hyperparameters values.
+Since $\hat{\alpha}$ combines both scale and shape information, it resolves the paradox.
 
-\citep{BHO75} is Bickel: ``Sex Bias in Graduate Admissions: Data from Berkeley''
+%\charles{We apply our theory of Heavy Tailed Self-Regularization (HTSR) to predict the generalization accuracies of the pretrained models provided in this contest, as a function of both model depth (L) and the solver hyperparameters (batch size, dropout, weight decay), and compare quality metrics that describe both the heavy-tailed shape and norm-based scale of the model layer weight matrices.
+In doing so, we identify what amounts to a Simpson’s paradox, in which shape and scale metrics show opposite behavior, and a combine metric resolves the paradox}
 
@@
 
-\michael{I may cut this stuff.}
-We have shown that weight matrix analysis derived from HT-SR Theory can be used to predict trends in the quality of state-of-the-art neural networks without access to training or testing data as well as identify Simpson's-like paradoxes in publicly-available NN models.
-The HT-SR Theory underlying our analysis is quite different than theoretical approaches popular in ML, and indeed it is quite different than typical statistical mechanics approaches to NNs \citep{BKPx20}.
-Recently, however, work motivated by HT-SR Theory has 
-demonstrated that multiplicative noise in stochastic optimization leads to heavy tails~\citep{HodMah20A_TR,GSZ20_TR}; 
-provided a random matrix analysis of random {F}ourier features, going beyond the {G}aussian kernel, to provide a precise phase transition and illustrate the corresponding double descent~\citep{ZCM20_TR};
-provided exact expressions for double descent and implicit regularization via a notion of surrogate random design \citep{DLM19_Exact_TR}; 
-demonstrated that good classifiers are abundant in the interpolating regime \citep{TKM20_TR};
-shown how injecting noise into hidden states of recurrent NNs leads to implicit regularization~\citep{LEHM21_TR,Mah12}; and 
-shown that analogous phase transitions hold for the {C}olumn {S}ubset {S}election and the {N}ystrom method~\citep{DKM20_TR}. 
-We expect that the empirical results identified by tools from HT-SR Theory will provide guidance and constraints for future development of overparameterized machine learning models.
+The HT-SR theory underlying our analysis is quite different than theoretical approaches popular in machine learning, and it is quite different than typical statistical mechanics approaches to DNNs \citep{BKPx20}.
+%
+%Recently, however, work motivated by HT-SR theory has 
+%demonstrated that multiplicative noise in stochastic optimization leads to heavy tails~\citep{HodMah20A_TR,GSZ20_TR}; 
+%provided a random matrix analysis of random {F}ourier features, going beyond the {G}aussian kernel, to provide a precise phase transition and illustrate the corresponding double descent~\citep{ZCM20_TR};
+%provided exact expressions for double descent and implicit regularization via a notion of surrogate random design \citep{DLM19_Exact_TR}; 
+%demonstrated that good classifiers are abundant in the interpolating regime \citep{TKM20_TR};
+%shown how injecting noise into hidden states of recurrent NNs leads to implicit regularization~\citep{LEHM21_TR,Mah12}; and 
+%shown that analogous phase transitions hold for the {C}olumn {S}ubset {S}election and the {N}ystrom method~\citep{DKM20_TR}. 
+%We expect that the empirical results identified by tools from HT-SR theory will provide guidance and constraints for future development of overparameterized machine learning models.
+%
+However, recent work, work motivated by HT-SR theory, has begun to appear
+\citep{HodMah20A_TR,GSZ20_TR, ZCM20_TR, DLM19_Exact_TR, TKM20_TR, LEHM21_TR,Mah12, DKM20_TR}. 
+We expect that the empirical results and theoretical insights HT-SR theory provides will provide guidance and constraints for future development of overparameterized machine learning models.
 

### 2021-02-18 submission/TOPML_Workshop/paper.tex
edits added

@@
 Using \texttt{WeightWatcher}, we can characterize HT structure in the empirical spectral densities (ESDs) of the layer weight matrices, and we can use this to predict trends in the reported test accuracies across hundreds of pretrained DNNs---without needing access to training or test data.
-Going beyond popular norm-based metrics, we show that different metrics, which either depend on the scale (norms) or shape (fitted truncated power law (PL) exponents) of the ESDs, exhbit a Simpson's paradox when varying DNN model depth versus solver hyperparameters.
-The familar scale-based spectral norm metric behaves opposite to theoretical predictions, while the HT $\alpha$-metric (the exponent in a truncated PL fit) acts as predicted.
+Going beyond popular norm-based metrics, we show that different metrics, which either depend on the scale or shap of the ESDs, exhibit a Simpson's paradox when varying DNN model depth versus regularization hyperparameters.
+The familiar scale-based spectral norm metric behaves opposite to theoretical predictions, while the HT $\alpha$-metric acts as predicted.
 The paradoxical behavior  can be resolved by combining these shape and scale metrics with the HT $\hat{\alpha}$ metric.
@@
 Building on these empirical results, 
-\citet{MM18_TR} developed a phenomenological (statistical mechanics based, as opposed to statistical learning theory based) theory to identify ``5+1 Phases of Training,'' corresponding to increasing amounts of implicit self-regularization.
+\citet{MM18_TR} developed a phenomenological theory, based in statistical mechanics (as opposed to statistical learning theory), to identify ``5+1 Phases of Training,'' 
+corresponding to increasing amounts of implicit self-regularization.
 These phases can be observed during the training process as well as in the final learned (pretrained) DNNs.
@@
 For state-of-the-art (SOTA) DNNs, however, they identified a novel form of \emph{Heavy-Tailed Self-Regularization} (HT-SR).
-This is similar to the self-organization seen in the statistical physics of disordered systems (such as classical models of actual neural activity); and 
+This is similar to the self-organization seen in the statistical physics of disordered systems, such classical models of actual spiking neural activity; and 
 it results from correlations arising at all size scales.
@@
 
-In the HT-SR theory, one analyzes the eigenvalue spectrum, i.e., the ESD, of the associated correlation matrices~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
-From this, one characterizes the amount and form of correlation present in the DNN's weight matrices.
+In the HT-SR theory, one analyzes the eigenvalue spectrum, or empirical spectral density (ESD) 
+of the associated layer correlation matrices~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
+and uses this to characterizes the amount and form of correlation present in the DNN's weight matrices.
 For each layer weight matrix $\mathbf{W}$, if we compute the eigenvalue spectrum of 
 $ \mathbf{X} = \frac{1}{N}\mathbf{W}^{T}\mathbf{W} $
-then HT-SR theory characterizes the correlations in a weight matrix by examining its ESD, $\rho(\lambda)$.
-Empirically, the ESD can be well-fit to a truncated PL distribution, 
-$\rho(\lambda)\sim\lambda^{-\alpha} $.
+then HT-SR theory characterizes the correlations in a weight matrix by fitting its ESD, $\rho(\lambda)$.
+to a truncated Power Law (TPL) distribution,  $\rho(\lambda)\sim\lambda^{-\alpha} $.
 %
-The original work on HT-SR theory showed that for nearly every $\mathbf{W}$, i.e., among publicly-available SOTA models, the (bulk and tail) of the ESDs can be fit to a truncated PL, and that PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
+The original work on HT-SR showed that among numerous publicly-available SOTA models, for nearly every $\mathbf{W}$,  the tail of the ESDs can be fit to a  TPL, and that Power Law (PL) exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
 %
@@
 %
-Crucially, HT-SR theory predicts that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models.
+Crucially, HT-SR theory predicts that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models,
+and therefore higher generalization accuracies.
 %%The notion of ``size scale'' is well-defined in physical systems, to which this style of analysis is usually applied, but it is less well-defined in CV and NLP applications.

### 2021-02-18 submission/TOPML_Workshop/paper.tex
edits added

@@
 %
-The original work on HT-SR showed that among numerous publicly-available SOTA models, for nearly every $\mathbf{W}$,  the tail of the ESDs can be fit to a  TPL, and that Power Law (PL) exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
+The original work on HT-SR showed that among numerous publicly-available SOTA models, for nearly every $\mathbf{W}$,  the tail of the ESDs can be fit to a  TPL, and that
+the Power Law (PL) exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
 %
@@
 
-We find that norm based metrics correlate well with reported test accuracies for well-trained models, but that they often cannot distinguish well-trained versus poorly-trained models.
+We find that norm based metrics (may) correlate well with reported test accuracies for well-trained models, 
+but that they often cannot distinguish well-trained versus poorly-trained models.
 We also find that PL based metrics can do much better---quantitatively better at discriminating among series of well-trained models with a given architecture; and qualitatively better at discriminating well-trained versus poorly-trained models.

### 2021-02-18 submission/TOPML_Workshop/paper.tex
added figure 2

@@
 
+\begin{figure}[t!] 
+    \centering
+    \subfigure[Sharpness]{
+       \includegraphics[width=3.5cm]{img/Sharpness_task2.png}
+    }
+    \subfigure[SVDSmoothing]{
+       \includegraphics[width=3.5cm]{img/SVDSmoothing_task2.png}
+    }
+    \caption{Applying the Sharpness (Shape) and SVDSMoothing (Scale) transforms to pretrained DNNs, we can correct for overparemeterization and predict test accuracies remarkably well by evaluating the training accuracy on the transformed models}
+    \label{fig:smoothing}
+\end{figure}
+
 \section*{Conclusion}

### 2021-02-19 submission/TOPML_Workshop/paper.tex
full pass to whitespace, typo check; fin_v1

--- 2021-02-18 submission/TOPML_Workshop/paper.tex
@@
 \twocolumn[
-\icmltitle{Using Heavy-Tailed Self-Regularization Theory to Characterize Correlations in and Perform Diagnostics on State-of-the-art Overparameterized Models}
+\icmltitle{Using Heavy-Tailed Self-Regularization Theory to Characterize Correlations in and Perform Diagnostics on State-of-the-Art Overparameterized Models}
 
@@
 
-\section*{Introduction}
+\paragraph{Introduction.}
+%\subsection*{Introduction}
 
 Understanding state-of-the-art overparameterized deep neural networks (DNNs) requires detailed empirical studies as well as the development of predictive theory in light of those studies. 
-We use the theory of Heavy-Tailed Self-Regularization (HT-SR) to predict the generalization accuracy and improve the fine-tuning of (overparameterized) pretrained DNN models by capturing the dominant correlation structure.
+We use the theory of Heavy-Tailed Self-Regularization (HT-SR) to predict the generalization accuracy and improve the fine-tuning of (overparameterized) pretrained DNN models by capturing the dominant correlational structure.
 HT-SR is based on Heavy-Tailed Random Matrix Theory (HT-RMT) and strongly-correlated systems theory from theoretical chemistry and theoretical physics, and methods based on it are implemented in the publicly-available \texttt{WeightWatcher} tool.
-Using \texttt{WeightWatcher}, we can characterize HT structure in the empirical spectral densities (ESDs) of the layer weight matrices, and we can use this to predict trends in the reported test accuracies across hundreds of pretrained DNNs---without needing access to training or test data.
-Going beyond popular norm-based metrics, we show that different metrics, which either depend on the scale or shap of the ESDs, exhibit a Simpson's paradox when varying DNN model depth versus regularization hyperparameters.
+Using \texttt{WeightWatcher}, we can characterize HT structure in the empirical spectral densities (ESDs) of layer weight matrices, and we can use this to predict trends in the reported test accuracies across hundreds of pretrained DNNs---without needing access to training or test data.
+Going beyond popular norm-based metrics, we show that different metrics, which either depend on the scale or shape of the ESDs, exhibit a Simpson's paradox when varying DNN model depth versus regularization hyperparameters.
 The familiar scale-based spectral norm metric behaves opposite to theoretical predictions, while the HT $\alpha$-metric acts as predicted.
-The paradoxical behavior  can be resolved by combining these shape and scale metrics with the HT $\hat{\alpha}$ metric.
-Finally, and with access to training data, we show how to
+The paradoxical behavior  can be resolved by combining shape and scale metrics with the HT $\hat{\alpha}$ metric.
+Finally, and with access to training data, we can 
 smooth out and capture the low rank structure to both accurately
@@
 
-\section*{Statistical Mechanics and Heavy-Tailed Self-Regularization}
+\vspace{-4mm}
+\paragraph{Statistical Mechanics and Heavy-Tailed Self-Regularization.}
+%\subsection*{Statistical Mechanics and Heavy-Tailed Self-Regularization}
 
-%\citep{MM17_TR} is Charles: Rethinking generalization requires revisiting old ideas: statistical mechanics approaches and complex learning behavior''
+%\cite{MM17_TR} is Charles: Rethinking generalization requires revisiting old ideas: statistical mechanics approaches and complex learning behavior''
 %
-%\citep{MM18_TR} is Charles: ``Implicit Self-Regularization in Deep Neural Networks: Evidence from Random Matrix Theory and Implications for Learning''
+%\cite{MM18_TR} is Charles: ``Implicit Self-Regularization in Deep Neural Networks: Evidence from Random Matrix Theory and Implications for Learning''
 %
-%\citep{MM19_HTSR_ICML} is Charles: ``Traditional and Heavy-Tailed Self Regularization in Neural Network Models''
+%\cite{MM19_HTSR_ICML} is Charles: ``Traditional and Heavy-Tailed Self Regularization in Neural Network Models''
 %
-%\citep{MM20_SDM} is Charles: ``Heavy-Tailed {U}niversality Predicts Trends in Test Accuracies for Very Large Pre-Trained Deep Neural Networks''
+%\cite{MM20_SDM} is Charles: ``Heavy-Tailed {U}niversality Predicts Trends in Test Accuracies for Very Large Pre-Trained Deep Neural Networks''
 %
-%\citep{MM19_KDD} is Charles: ``Statistical Mechanics Methods for Discovering Knowledge from Modern Production Quality Neural Networks''
+%\cite{MM19_KDD} is Charles: ``Statistical Mechanics Methods for Discovering Knowledge from Modern Production Quality Neural Networks''
 
-Recently, RMT was applied to analyze empirically the weight matrices of DNNs, including both production quality, pretrained models such as AlexNet and Inception, and smaller models trained from scratch, such as LeNet5 and a miniature-AlexNet~\citep{MM18_TR} (see also \citep{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM20_SDM, MM19_KDD}).
+RMT was applied to analyze empirically weight matrices of DNNs, including both production quality, pretrained models such as AlexNet and Inception, and smaller models trained from scratch, such as LeNet5 and a miniature-AlexNet~\cite{MM18_TR} (see also \cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM20_SDM, MM19_KDD}).
 Building on these empirical results, 
-\citet{MM18_TR} developed a phenomenological theory, based in statistical mechanics (as opposed to statistical learning theory), to identify ``5+1 Phases of Training,'' 
+\citet{MM18_TR} developed a phenomenological theory, based on statistical mechanics (as opposed to statistical learning theory), to identify ``5+1 Phases of Training,'' 
 corresponding to increasing amounts of implicit self-regularization.
 These phases can be observed during the training process as well as in the final learned (pretrained) DNNs.
-For smaller and/or older DNNs, this implicit self-regularization is like traditional Tikhonov regularization, in that there is a ``size scale'' separating signal from noise.
+For smaller/older DNNs, this implicit self-regularization is like Tikhonov regularization, in that there is a ``size scale'' separating signal from noise.
 For state-of-the-art (SOTA) DNNs, however, they identified a novel form of \emph{Heavy-Tailed Self-Regularization} (HT-SR).
-This is similar to the self-organization seen in the statistical physics of disordered systems, such classical models of actual spiking neural activity; and 
+This is similar to the self-organization seen in the statistical physics of disordered systems, such as classical models of actual spiking neurons; and 
 it results from correlations arising at all size scales.
-For DNNs, this arises due to the training process itself, when applied to very overparameterized models.
+For DNNs, this arises due to the training process itself, when applied to overparameterized models.
 
-In the HT-SR theory, one analyzes the eigenvalue spectrum, or empirical spectral density (ESD) 
-of the associated layer correlation matrices~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
+In HT-SR theory, one analyzes the eigenvalue spectrum, or empirical spectral density (ESD) 
+of the associated layer correlation matrices~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}
 and uses this to characterizes the amount and form of correlation present in the DNN's weight matrices.
 For each layer weight matrix $\mathbf{W}$, if we compute the eigenvalue spectrum of 
-$ \mathbf{X} = \frac{1}{N}\mathbf{W}^{T}\mathbf{W} $
-then HT-SR theory characterizes the correlations in a weight matrix by fitting its ESD, $\rho(\lambda)$.
+$ \mathbf{X} = \frac{1}{N}\mathbf{W}^{T}\mathbf{W} $,
+then HT-SR theory characterizes the correlations in a weight matrix by fitting its ESD, $\rho(\lambda)$,
 to a truncated Power Law (TPL) distribution,  $\rho(\lambda)\sim\lambda^{-\alpha} $.
 %
-The original work on HT-SR showed that among numerous publicly-available SOTA models, for nearly every $\mathbf{W}$,  the tail of the ESDs can be fit to a  TPL, and that
-the Power Law (PL) exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
+The original work on HT-SR showed that, among numerous publicly-available SOTA models, for nearly every $\mathbf{W}$, the tail of the ESDs can be fit to a  TPL, and that
+PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
 %
-%%As for the mechanism responsible for these properties, statistical physics offers several possibilities~\citep{SornetteBook,nishimori01}, e.g., self organized criticality~\citep{SOC87,SOCat25yrs} or multiplicative noise in the stochastic optimization algorithms used to train these models~\citep{HodMah20A_TR,SorCon97}.
-%%Alternatively, related techniques have been used to analyze correlations and information propogation in actual spiking neurons~\citep{SYYRP11,YKYP14}.
+%%As for the mechanism responsible for these properties, statistical physics offers several possibilities~\cite{SornetteBook,nishimori01}, e.g., self organized criticality~\cite{SOC87,SOCat25yrs} or multiplicative noise in the stochastic optimization algorithms used to train these models~\cite{HodMah20A_TR,SorCon97}.
+%%Alternatively, related techniques have been used to analyze correlations and information propogation in actual spiking neurons~\cite{SYYRP11,YKYP14}.
 %%Our meta-analysis does not require knowledge of mechanisms; and it is not even clear that one mechanism is responsible for every case.
 %
-Crucially, HT-SR theory predicts that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models,
-and therefore higher generalization accuracies.
+Crucially, HT-SR theory predicts that smaller values of $\alpha$ correspond to models with better correlation over multiple size scales and thus to better models,
+and therefore higher generalization~accuracies.
 %%The notion of ``size scale'' is well-defined in physical systems, to which this style of analysis is usually applied, but it is less well-defined in CV and NLP applications.
 %%Informally, it would correspond to pixel groups that are at a greater distance in some metric, or between sentence parts that are at a greater distance in text.
-%%Relatedly, previous work observed that smaller exponents $\alpha$ correspond to more implicit self-regularization and better generalization, and that we expect a linear correlation between $\hat{\alpha}$ and model quality~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
+%%Relatedly, previous work observed that smaller exponents $\alpha$ correspond to more implicit self-regularization and better generalization, and that we expect a linear correlation between $\hat{\alpha}$ and model quality~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
 
-\section*{Overall approach}
+\vspace{-5mm}
+\paragraph{Overall Approach.}
+%\subsection*{Overall Approach}
 
-Test accuracies have been reported online for publicly-available pretrained pyTorch models~\citep{osmr}.
+Test accuracies have been reported online for publicly-available pretrained pyTorch models~\cite{osmr}.
 These models have been trained and evaluated on labeled data, using standard techniques.  
 We do not have access to this data, and we have not trained any of the models ourselves. 
-Our methodological approach is thus similar to a statistical meta-analysis, common in biomedical research, but uncommon in machine learning.
+Our methodological approach is thus similar to a statistical meta-analysis, common in biomedical research, but uncommon in machine learning (ML).
 %
@@
 %\end{eqnarray*}
-To perform diagnostics on potentially-problematic DNNs, we will decompose $\hat{\alpha}$ into its two components, $\alpha$ and $\lambda_{max}$.
-One determines $\alpha$ for a given layer by fitting the ESD of that layer's weight matrix to a truncated PL, using the commonly accepted Maximum Likelihood method~\citep{CSN09_powerlaw,ABP14}.
-Computations were performed with the publicly-available \texttt{WeightWatcher} tool (version 0.2.7)~\citep{weightwatcher_package}.
+To perform diagnostics on potentially-problematic DNNs, we decompose $\hat{\alpha}$ into its two components, $\alpha$ and $\lambda_{max}$.
+One determines $\alpha$ for a given layer by fitting the ESD of that layer's weight matrix to a TPL, using the commonly accepted Maximum Likelihood method~\cite{CSN09_powerlaw,ABP14}.
+Computations were performed with the publicly-available \texttt{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
 
-\section*{Predicting Trends in the Quality of State-of-the-art Models}
+\vspace{-4mm}
+\paragraph{Predicting Trends in the Quality of State-of-the-art Models.}
+%\subsection*{Predicting Trends in the Quality of State-of-the-art Models}
 
-%XXX.  ABSTRACT FROM \citep{MM20a_trends_TR}: ``Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data''
+%XXX.  ABSTRACT FROM \cite{MM20a_trends_TR}: ``Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data''
 
@@
 Moreover, one may not know details about the model, e.g., the specifics of the training data, the loss function, the hyperparameter values, etc.
-(So, one is \emph{not} training models, one is validating pretrained models.)
+%(So, one is \emph{not} training models, one is validating pretrained models.)
 Given one or many pretrained models, it is a challenge to say anything about the expected performance or quality of the models.
@@
 
-We consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
-By now, there are many such state-of-the-art models that are publicly-available, e.g., 
-hundreds of pretrained models in CV ($\ge 500$) and NLP ($\approx 100$).
-For all these models, we have no access to training data or testing data, and we have no specific knowledge of the training/testing protocols. 
+We consider a corpus of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
+By now, there are 
+hundreds of such pretrained 
+state-of-the-art models that are publicly-available, 
+in both CV ($\ge 500$) and NLP ($\approx 100$).
+For these models, we have no access to training data or testing data, and we have no specific knowledge of training/testing protocols. 
 Thus, performing weight matrix analysis is a natural option.
@@
 %
-%\citep{BFT17_TR} is Bartlett: ``Spectrally-normalized margin bounds for neural networks''
-%\citep{LMBx18_TR} is Poggio: ``A surprising linear relationship predicts test performance in deep networks''
-There is related work \citep{BFT17_TR,LMBx18_TR}, but that work provides generalization bounds, modifies the models, requires access to data, etc., and so is quite different.
+%\cite{BFT17_TR} is Bartlett: ``Spectrally-normalized margin bounds for neural networks''
+%\cite{LMBx18_TR} is Poggio: ``A surprising linear relationship predicts test performance in deep networks''
+There is related work \cite{BFT17_TR,LMBx18_TR}, but that work provides generalization bounds, modifies the models, requires access to data, etc., and so is quite different.
 
@@
 \end{center}
+\vspace{-4mm}
 \caption{Comparison of linear regressions for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering \
@@
 
-\section*{A Simpson's Paradox---and the Need for Finer-scale Analysis}
+\vspace{-4mm}
+\paragraph{A Simpson's Paradox---and the Need for Finer-scale Analysis.}
+%\subsection*{A Simpson's Paradox---and the Need for Finer-scale Analysis}
 
-%\citep{JNBx19_fantastic_TR} is the terrible fantastic paper: ``Fantastic Generalization Measures and Where to Find Them''
+%\cite{JNBx19_fantastic_TR} is the terrible fantastic paper: ``Fantastic Generalization Measures and Where to Find Them''
 %
-%\citep{JFYx20_contest_v10} is the contest: ``{NeurIPS} 2020 Competition: Predicting Generalization in Deep Learning (Version 1.0)''
+%\cite{JFYx20_contest_v10} is the contest: ``{NeurIPS} 2020 Competition: Predicting Generalization in Deep Learning (Version 1.0)''
 %
-%\citep{BHO75} is Bickel: ``Sex Bias in Graduate Admissions: Data from Berkeley''
+%\cite{BHO75} is Bickel: ``Sex Bias in Graduate Admissions: Data from Berkeley''
 
 \begin{figure}[t!] 
+\vspace{-4mm}
     \centering
@@
     }
+\vspace{-4mm}
     \caption{Test accuracy as a function of 
-             the Log Spectral Norm and 
-             the (HT-SR) Weighted Alpha ($\hat{\alpha}$) metrics,
-             as model hyperparameters and depth are varied. 
-             Models are from the NeurIPS 2020 ``Predicting Generalization'' contest.
+             Log Spectral Norm and 
+             (HT-SR) Weighted Alpha ($\hat{\alpha}$) metrics,
+             as hyperparameters and depth are varied, for 
+             models from the NeurIPS 2020 ``Predicting Generalization'' contest.
              A Simpson's paradox is clearly visible.
@@
 
-Subsequent to the development of HT-SR theory, and our initial results on analyzing pretrained models at scale, there was at NeurIPS 2020 a ``Predicting Generalization'' workshop and contest \citep{JNBx19_fantastic_TR, JFYx20_contest_v10}.
-This contest made publicly-available pretrained models.
-This corpus of models was much more narrow (in terms of domain) but more extensive (in terms of depth and hyperparameter variation) than the previous SOTA corpus we considered~\citep{osmr}.
+Subsequent to the development of HT-SR theory, and our initial results on analyzing pretrained models at scale, there was a NeurIPS 2020 workshop/contest on ``Predicting Generalization'' 
+\cite{JNBx19_fantastic_TR, JFYx20_contest_v10} that made pretrained models publicly-available. 
+This corpus of models was much more narrow (in terms of domain) but more extensive (in terms of depth and hyperparameter variation) than the previous SOTA corpus we considered~\cite{osmr}.
 
-In particular, we apply HT-SR theory to predict the generalization accuracies of the pretrained models provided in this contest, as a function of both model depth ($L$) and the solver hyperparameters (batch size, dropout, weight decay), and we compare quality metrics that describe the norm-based scale and the HT/PL shape of the model layer weight matrices.
+We apply HT-SR theory to predict the generalization accuracies of the pretrained models provided in this contest, as a function of both model depth ($L$) and the solver hyperparameters (batch size, dropout, weight decay), and we compare quality metrics that describe the norm-based scale and the HT/PL shape of the model layer weight matrices.
 See Figure~\ref{fig:simpsons} for a summary.
-For norm-based metrics, we identify what amounts to a Simpson’s paradox \citep{BHO75}, in which scale and shape metrics show opposite behavior, depending on whether the data are partitioned by depth or solver hyperparameters values.
+For norm-based metrics, we identify what amounts to a Simpson’s paradox \cite{BHO75}, in which scale and shape metrics show opposite behavior, depending on whether the data are partitioned by depth or solver hyperparameters values.
 Since $\hat{\alpha}$ combines both scale and shape information, it resolves the paradox.
@@
 
-\section*{Conclusion}
+%\begin{figure}[t!] 
+%    \centering
+%    \subfigure[Sharpness]{
+%       \includegraphics[width=3.5cm]{img/Sharpness_task2.png}
+%    }
+%    \subfigure[SVDSmoothing]{
+%       \includegraphics[width=3.5cm]{img/SVDSmoothing_task2.png}
+%    }
+%    \caption{Applying the Sharpness (Shape) and SVDSMoothing (Scale) transforms to pretrained DNNs, we can correct for overparemeterization and predict test accuracies remarkably well by evaluating the training accuracy on the transformed models}
+%    \label{fig:smoothing}
+%\end{figure}
 
-The HT-SR theory underlying our analysis is quite different than theoretical approaches popular in machine learning, and it is quite different than typical statistical mechanics approaches to DNNs \citep{BKPx20}.
+\vspace{-5mm}
+\paragraph{Conclusion.}
+%\subsection*{Conclusion}
+
+The HT-SR theory underlying our analysis is quite different than theoretical approaches popular in ML, and it is quite different than typical statistical mechanics approaches to DNNs \cite{BKPx20}.
 %
 %Recently, however, work motivated by HT-SR theory has 
-%demonstrated that multiplicative noise in stochastic optimization leads to heavy tails~\citep{HodMah20A_TR,GSZ20_TR}; 
-%provided a random matrix analysis of random {F}ourier features, going beyond the {G}aussian kernel, to provide a precise phase transition and illustrate the corresponding double descent~\citep{ZCM20_TR};
-%provided exact expressions for double descent and implicit regularization via a notion of surrogate random design \citep{DLM19_Exact_TR}; 
-%demonstrated that good classifiers are abundant in the interpolating regime \citep{TKM20_TR};
-%shown how injecting noise into hidden states of recurrent NNs leads to implicit regularization~\citep{LEHM21_TR,Mah12}; and 
-%shown that analogous phase transitions hold for the {C}olumn {S}ubset {S}election and the {N}ystrom method~\citep{DKM20_TR}. 
-%We expect that the empirical results identified by tools from HT-SR theory will provide guidance and constraints for future development of overparameterized machine learning models.
+%demonstrated that multiplicative noise in stochastic optimization leads to heavy tails~\cite{HodMah20A_TR,GSZ20_TR}; 
+%provided a random matrix analysis of random {F}ourier features, going beyond the {G}aussian kernel, to provide a precise phase transition and illustrate the corresponding double descent~\cite{ZCM20_TR};
+%provided exact expressions for double descent and implicit regularization via a notion of surrogate random design \cite{DLM19_Exact_TR}; 
+%demonstrated that good classifiers are abundant in the interpolating regime \cite{TKM20_TR};
+%shown how injecting noise into hidden states of recurrent NNs leads to implicit regularization~\cite{LEHM21_TR,Mah12}; and 
+%shown that analogous phase transitions hold for the {C}olumn {S}ubset {S}election and the {N}ystrom method~\cite{DKM20_TR}. 
+%We expect that the empirical results identified by tools from HT-SR theory will provide guidance and constraints for future development of overparameterized ML models.
 %
-However, recent work, work motivated by HT-SR theory, has begun to appear
-\citep{HodMah20A_TR,GSZ20_TR, ZCM20_TR, DLM19_Exact_TR, TKM20_TR, LEHM21_TR,Mah12, DKM20_TR}. 
-We expect that the empirical results and theoretical insights HT-SR theory provides will provide guidance and constraints for future development of overparameterized machine learning models.
+However, recent ML work, motivated by HT-SR theory, has begun to appear
+\cite{HodMah20A_TR,GSZ20_TR, ZCM20_TR, DLM19_Exact_TR, TKM20_TR, LEHM21_TR,Mah12, DKM20_TR}. 
+We expect that the empirical results and theoretical insights HT-SR theory provides will provide guidance and constraints for future development of overparameterized ML~models.
 
